### 2008-09-10 developments/PDB/BIT_VECTOR_DEFINITION.tex

\documentstyle[11pt,supertabular]{article}
\setlength{\topmargin}{13mm}
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\textheight}{225mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm}
\input{psfig}

\begin{document}
\begin{sloppypar}

% DEFINITION DES CARACTERES MATHEMATIQUES B
%------------------------------------------
\def\@setmcodes#1#2#3{{\count0=#1 \count1=#3
	\loop \global\mathcode\count0=\count1 \ifnum \count0<#2
	\advance\count0 by1 \advance\count1 by1 \repeat}}

\@setmcodes{`A}{`Z}{"7441}
\@setmcodes{`a}{`z}{"7461}

\mathcode`\;="8000 % Makes ; active in math mode
{\catcode`\;=\active \gdef;{\semicolon\;}}
\mathchardef\semicolon="003B
%    Nominal distance from top of paper to top of page
\topmargin 0 pt
\textheight 53\baselineskip

%   Left margin on odd-numbered pages
\oddsidemargin  0.15 in
%   Left margin on even-numbered pages
\evensidemargin 0.35 in
%   Width of marginal notes.
\marginparwidth 1 in
%   Note that \oddsidemargin = \evensidemargin
\oddsidemargin 0.25 in
\evensidemargin 0.25 in
\marginparwidth 0.75 in
\textwidth 5.875 in % Width of text line.

\setlength{\parindent}{0pt}
\setlength{\parskip}{0ex}

% DEFINITION DES FONTS
%---------------------
% The AMS extra symbol fonts are loaded.
% Note: sometimes called euxm10
\font\msx=msam10
% Note: sometimes called euym10
\font\msy=msbm10

\newfam\msxfam \textfont\msxfam=\msx
\newfam\msyfam \textfont\msyfam=\msy

\def\famletter#1{\ifcase #1 0\or 1\or 2\or 3\or 4\or 5\or 6\or 7\or
	8\or 9\or A\or B\or C\or D\or E\or F\fi}

\edef\fx{\famletter\msxfam}
\edef\fy{\famletter\msyfam}

\def\bbold{\fam\msyfam \msy}

% SYMBOLES B
%-----------
% makes a quoted expression in mathematical text
\def\token#1{\hbox{`$#1$'}}
% used for error messages in Z specs
\def\report#1{\hbox{`{\tt #1}'}}

% \@myop makes an operator, with a strut to defeat TeX's vertical adjustment.
\def\@myop#1{\mathop{\mathstrut{#1}}\nolimits}

% This underscore doesn't have the little kern --- you get an italic
% correction anyway in math mode.
\def\_{\leavevmode \vbox{\hrule width0.5em}}

% Save \q as \xq for quantifiers q.
\let\xforall=\forall
\let\xexists=\exists
\let\xlambda=\lambda
\let\xmu=\mu

% \p and \f make arrows with 1 and 2 crossings resp.
\def\p#1{\mathrel{\ooalign{\hfil$\mapstochar\mkern 5mu$\hfil\cr$#1$}}}
\def\f#1{\mathrel{\ooalign{\hfil
	$\mapstochar\mkern 3mu\mapstochar\mkern 5mu$\hfil\cr$#1$}}}

\let\mc=\mathchardef

\def	\pow		{\mbox{${\cal P}$}}
\def	\po1		{\mbox{${\cal P}_1$}}
\let	\cross		\times
\def	\lambda		{\@myop{\xlambda}}
\def	\lnot		{\neg\;}
\def	\land		{\mathrel{\wedge}}
\def	\lor		{\mathrel{\vee}}
\let	\implies	\Rightarrow
\let	\iff		\Leftrightarrow
\def	\forall		{\@myop{\xforall}}
\def	\exists		{\@myop{\xexists}}
\def	\semi		{\mathrel{\comp}}
\def	\ssemi		{\mathbin{\rm ;}}
\let	\ensembleVide	\emptyset
\let	\rel		\leftrightarrow
\def	\dom		{\@myop{\sf dom}}
\def	\ran		{\@myop{\sf ran}}
\def	\id		{\@myop{\sf id}}
\def	\comp		{\mathbin{\raise
			0.6ex\hbox{\oalign{\hfil$\scriptscriptstyle
			\rm o$\hfil\cr\hfil$\scriptscriptstyle\rm 9$\hfil}}}}
\def	\para		{\mbox{$\mid\mid$}}
\mc	\dres		"2\fx43
\mc	\rres		"2\fx42
\def	\ndres		{\mathbin{{\dres} \llap{$-$}}}
\def	\nrres		{\mathbin{{\rres}\llap{$-$}}}
\def	\lover		{\mathbin{{\dres} \llap{$-\!\!\!\!-\!$}}}
\def	\rover		{\mathbin{{\rres}\llap{$\!-\!\!\!-$}}}
\let	\fun		\rightarrow
\def	\pfun		{\p\fun}
\def	\pinj		{\p\inj}
\mc	\inj		"3\fx1A
\def	\psurj		{\p\surj}
\mc	\surj		"3\fx10
\def	\bij		{\surj\!\!\!\!\!\!\!\inj}
\def	\nat		{\mbox{${\cal N}$}}
\def	\na1		{\mbox{${\cal N}_1$}}
\def	\num		{\mbox{${\cal Z}$}}
\def	\int		{\mbox{${\cal Z}$}}
\def	\rat		{\mbox{${\cal Q}$}}
\def	\div		{\mathbin{\rm /}}
\def	\mod		{\mathbin{\bf mod}}
\def	\upto		{\mathbin{\ldotp\ldotp}}
\def	\finset		{\mbox{${\cal F}$}}
\def	\finse1		{\mbox{${\cal F}_1$}}
\def	\ffun		{\f\fun}
\def	\finj		{\f\inj}
\def	\seq		{\@myop{\rm seq}}
\def	\cat		{\mathbin{\raise 0.8ex\hbox{$\mathchar"2\fx61$}}}
\def	\sep		{\hspace*{.05in}}

\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{0}

%-------------------%
% Debut du document %
%-------------------%

\bf MACHINE

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION

\vspace*{4mm}
\bf SEES

\hspace*{0.20in}\it BIT\_DEFINITION

\vspace*{4mm}
\bf CONSTANTS

\hspace*{0.20in}\it BIT\_VECTOR\rm ,

\hspace*{0.20in}\it bv\_size\rm , \it bv\_catenate\rm , \it bv\_sub\rm , \it bv\_zero\rm , \it bv\_one\rm ,

\hspace*{0.20in}\it bv\_not\rm ,\hspace*{0.20in}\it bv\_and\rm ,\hspace*{0.20in}\it bv\_or\rm , \it bv\_xor\rm ,

\hspace*{0.20in}\it bv\_at\rm , \it bv\_set\rm , \it bv\_clear

\vspace*{4mm}
\bf PROPERTIES

\hspace*{0.20in}\it BIT\_VECTOR \rm = \bf seq1 \rm (\it BIT\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_size  $\in$  \it BIT\_VECTOR  $\fun$   $ \na1 $   $\land$

\hspace*{0.20in}\it bv\_size \rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv  $\in$  \it BIT\_VECTOR  $\mid$  \bf size\rm (\it bv\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_catenate \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm ) \rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2 $\in$  \it BIT\_VECTOR  $\mid$  \it v2  $\cat$  \it v1\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_sub  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\times$   $\nat$   $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_sub \rm =  $\lambda$ \rm (\it bv\rm , \it low\rm , \it high\rm ) \rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it low $\in$   $\nat$   $\land$  \it high $\in$   $\nat$   $\land$  \it low  $\leq$  \it high  $\mid$  \rm 0 $\upto$ \rm (\it high\rm -\it low\rm )  $\dres$  \it bv\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_zero  $\in$   $ \na1 $   $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_zero \rm =  $\lambda$ \rm (\it sz\rm ) \rm . \rm (\it sz  $\in$   $ \na1 $   $\mid$  \rm (\rm 1 $\upto$ \it sz\rm )  $\times$  \rm \{\rm 0\rm \}\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_one  $\in$   $ \na1 $   $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_one \rm =  $\lambda$ \rm (\it sz\rm ) \rm . \rm (\it sz  $\in$   $ \na1 $   $\mid$  \rm (\rm 1 $\upto$ \it sz\rm )  $\times$  \rm \{\rm 1\rm \}\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_not  $\in$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_not \rm =  $\lambda$ \rm (\it v1\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\mid$   $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_not\rm (\it v1\rm (\it idx\rm )\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_and  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_and \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \bf size\rm (\it v1\rm ) \rm = \bf size\rm (\it v2\rm )  $\mid$

\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_and\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_or  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_or \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \bf size\rm (\it v1\rm ) \rm = \bf size\rm (\it v2\rm )  $\mid$

\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_or\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_xor  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_xor \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \bf size\rm (\it v1\rm ) \rm = \bf size\rm (\it v2\rm )  $\mid$

\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_xor\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_at  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it bv\_at \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$   $\nat$   $\land$  \it idx $<$ \bf size\rm (\it v1\rm )  $\mid$  \it v1\rm (\it idx\rm +\rm 1\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_set  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_set \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$   $\nat$   $\land$  \it idx $<$ \bf size\rm (\it v1\rm )  $\mid$  \it v1  $\lover$  \rm \{ \it idx\rm +\rm 1  $\mapsto$  \rm 1 \rm \}\rm )  $\land$

\hspace*{0.20in}\it bv\_clear  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_clear \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$   $\nat$   $\land$  \it idx $<$ \bf size\rm (\it v1\rm )  $\mid$  \it v1  $\lover$  \rm \{ \it idx\rm +\rm 1  $\mapsto$  \rm 0 \rm \}\rm )

\vspace*{4mm}
\bf ASSERTIONS

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \it bv\_size\rm (\it bv\_not\rm (\it bv\rm )\rm ) \rm = \it bv\_size\rm (\it bv\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_not\rm (\it bv\_not\rm (\it bv\rm )\rm ) \rm = \it bv\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_catenate\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm ) \rm + \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it low\rm ,\it high\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it low  $\in$   $\nat$   $\land$  \it high  $\in$   $\nat$   $\land$  \it low  $\leq$  \it high  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_sub\rm (\it bv\rm , \it low\rm , \it high\rm )\rm ) \rm = \rm 1\rm +\it high\rm -\it low\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR\hspace*{0.10in} $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm ) \rm = \it bv\_and\rm (\it v2\rm , \it v1\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it v3\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR $\land$  \it v3  $\in$  \it BIT\_VECTOR\hspace*{0.10in} $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v3\rm )  $\implies$

\hspace*{0.60in}\rm (\it bv\_and\rm (\it v1\rm , \it bv\_and\rm (\it v2\rm ,\it v3\rm )\rm ) \rm = \it bv\_and\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm , \it v3\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_and\rm (\it bv\rm , \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_and\rm (\it bv\rm , \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm ) \rm = \it bv\_or\rm (\it v2\rm , \it v1\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it v3\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR $\land$  \it v3  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v3\rm )  $\implies$

\hspace*{0.60in}\rm (\it bv\_or\rm (\it v1\rm , \it bv\_or\rm (\it v2\rm ,\it v3\rm )\rm ) \rm = \it bv\_or\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm , \it v3\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_or\rm (\it bv\rm , \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_or\rm (\it bv\rm , \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm ) \rm = \it bv\_xor\rm (\it v2\rm , \it v1\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it bv\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \it bv\_xor\rm (\it bv\rm ,\it bv\rm ) \rm = \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )

\bf END\newpage
\end{sloppypar}\end{document}

### 2008-09-10 developments/HARDWARE_LIBRARY/PDB/BIT_DEFINITION.tex

--- 2008-09-10 developments/PDB/BIT_VECTOR_DEFINITION.tex
@@
 
-\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION
-
-\vspace*{4mm}
-\bf SEES
-
 \hspace*{0.20in}\it BIT\_DEFINITION
@@
 
-\hspace*{0.20in}\it BIT\_VECTOR\rm ,
+\hspace*{0.20in}\it BIT\rm ,
 
-\hspace*{0.20in}\it bv\_size\rm , \it bv\_catenate\rm , \it bv\_sub\rm , \it bv\_zero\rm , \it bv\_one\rm ,
+\hspace*{0.20in}\it bit\_not\rm ,
 
-\hspace*{0.20in}\it bv\_not\rm ,\hspace*{0.20in}\it bv\_and\rm ,\hspace*{0.20in}\it bv\_or\rm , \it bv\_xor\rm ,
+\hspace*{0.20in}\it bit\_and\rm ,
 
-\hspace*{0.20in}\it bv\_at\rm , \it bv\_set\rm , \it bv\_clear
+\hspace*{0.20in}\it bit\_or\rm ,
+
+\hspace*{0.20in}\it bit\_xor\rm ,
+
+\hspace*{0.20in}\it bool\_to\_bit
 
@@
 
-\hspace*{0.20in}\it BIT\_VECTOR \rm = \bf seq1 \rm (\it BIT\rm )  $\land$
+\hspace*{0.20in}\it BIT \rm = \rm 0 $\upto$ \rm 1  $\land$
 
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_size  $\in$  \it BIT\_VECTOR  $\fun$   $ \na1 $   $\land$
+\hspace*{0.20in}\it bit\_not  $\in$  \it BIT  $\fun$  \it BIT  $\land$
 
-\hspace*{0.20in}\it bv\_size \rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv  $\in$  \it BIT\_VECTOR  $\mid$  \bf size\rm (\it bv\rm )\rm )  $\land$
+\hspace*{0.20in} $\forall$  \rm (\it bb\rm )\rm . \rm (\it bb  $\in$  \it BIT  $\implies$  \it bit\_not\rm (\it bb\rm ) \rm = \rm 1\rm -\it bb\rm )  $\land$
 
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$
+\hspace*{0.20in}\it bit\_and  $\in$  \it BIT  $\times$  \it BIT  $\fun$  \it BIT  $\land$
 
-\hspace*{0.20in}\it bv\_catenate \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm ) \rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2 $\in$  \it BIT\_VECTOR  $\mid$  \it v2  $\cat$  \it v1\rm )  $\land$
+\hspace*{0.20in} $\forall$  \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\rm (\it bit\_and\rm (\it b1\rm , \it b2\rm ) \rm = \rm 1\rm )  $\iff$  \rm (\it b1 \rm = \rm 1\rm )  $\land$  \rm (\it b2 \rm = \rm 1\rm )\rm )\rm )  $\land$
 
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_sub  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\times$   $\nat$   $\fun$  \it BIT\_VECTOR  $\land$
+\hspace*{0.20in}\it bit\_or  $\in$  \it BIT  $\times$  \it BIT  $\fun$  \it BIT  $\land$
 
-\hspace*{0.20in}\it bv\_sub \rm =  $\lambda$ \rm (\it bv\rm , \it low\rm , \it high\rm ) \rm .
-\rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it low $\in$   $\nat$   $\land$  \it high $\in$   $\nat$   $\land$  \it low  $\leq$  \it high  $\mid$  \rm 0 $\upto$ \rm (\it high\rm -\it low\rm )  $\dres$  \it bv\rm )  $\land$
+\hspace*{0.20in} $\forall$  \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\rm (\it bit\_or\rm (\it b1\rm , \it b2\rm ) \rm = \rm 1\rm )  $\iff$  \rm (\it b1 \rm = \rm 1\rm )  $\lor$  \rm (\it b2 \rm = \rm 1\rm )\rm )\rm )  $\land$
 
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_zero  $\in$   $ \na1 $   $\fun$  \it BIT\_VECTOR  $\land$
+\hspace*{0.20in}\it bit\_xor  $\in$  \it BIT  $\times$  \it BIT  $\fun$  \it BIT  $\land$
 
-\hspace*{0.20in}\it bv\_zero \rm =  $\lambda$ \rm (\it sz\rm ) \rm . \rm (\it sz  $\in$   $ \na1 $   $\mid$  \rm (\rm 1 $\upto$ \it sz\rm )  $\times$  \rm \{\rm 0\rm \}\rm )  $\land$
+\hspace*{0.20in} $\forall$  \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\rm (\it bit\_xor\rm (\it b1\rm , \it b2\rm ) \rm = \rm 1\rm )  $\iff$  \rm (\it b1  $\not =$  \it b2\rm )\rm )\rm )  $\land$
 
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_one  $\in$   $ \na1 $   $\fun$  \it BIT\_VECTOR  $\land$
+\hspace*{0.20in}\it bool\_to\_bit  $\in$  \bf BOOL  $\fun$  \it BIT  $\land$
 
-\hspace*{0.20in}\it bv\_one \rm =  $\lambda$ \rm (\it sz\rm ) \rm . \rm (\it sz  $\in$   $ \na1 $   $\mid$  \rm (\rm 1 $\upto$ \it sz\rm )  $\times$  \rm \{\rm 1\rm \}\rm )  $\land$
-
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_not  $\in$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$
-
-\hspace*{0.20in}\it bv\_not \rm =  $\lambda$ \rm (\it v1\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\mid$   $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_not\rm (\it v1\rm (\it idx\rm )\rm )\rm )\rm )  $\land$
-
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_and  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$
-
-\hspace*{0.20in}\it bv\_and \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \bf size\rm (\it v1\rm ) \rm = \bf size\rm (\it v2\rm )  $\mid$
-
-\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_and\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )  $\land$
-
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_or  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$
-
-\hspace*{0.20in}\it bv\_or \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \bf size\rm (\it v1\rm ) \rm = \bf size\rm (\it v2\rm )  $\mid$
-
-\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_or\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )  $\land$
-
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_xor  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$
-
-\hspace*{0.20in}\it bv\_xor \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \bf size\rm (\it v1\rm ) \rm = \bf size\rm (\it v2\rm )  $\mid$
-
-\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_xor\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )  $\land$
-
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_at  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT  $\land$
-
-\hspace*{0.20in}\it bv\_at \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$   $\nat$   $\land$  \it idx $<$ \bf size\rm (\it v1\rm )  $\mid$  \it v1\rm (\it idx\rm +\rm 1\rm )\rm )  $\land$
-
-\vspace*{4mm}
-\hspace*{0.20in}\it bv\_set  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT\_VECTOR  $\land$
-
-\hspace*{0.20in}\it bv\_set \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$   $\nat$   $\land$  \it idx $<$ \bf size\rm (\it v1\rm )  $\mid$  \it v1  $\lover$  \rm \{ \it idx\rm +\rm 1  $\mapsto$  \rm 1 \rm \}\rm )  $\land$
-
-\hspace*{0.20in}\it bv\_clear  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT\_VECTOR  $\land$
-
-\hspace*{0.20in}\it bv\_clear \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$   $\nat$   $\land$  \it idx $<$ \bf size\rm (\it v1\rm )  $\mid$  \it v1  $\lover$  \rm \{ \it idx\rm +\rm 1  $\mapsto$  \rm 0 \rm \}\rm )
+\hspace*{0.20in}\it bool\_to\_bit \rm = \rm \{ \bf TRUE  $\mapsto$  \rm 1\rm , \bf FALSE  $\mapsto$  \rm 0 \rm \}
 
@@
 
-\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \it bv\_size\rm (\it bv\_not\rm (\it bv\rm )\rm ) \rm = \it bv\_size\rm (\it bv\rm )\rm )\rm ;
+\hspace*{0.20in}\it bit\_not\rm (\rm 0\rm ) \rm = \rm 1\rm ;
 
-\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_not\rm (\it bv\_not\rm (\it bv\rm )\rm ) \rm = \it bv\rm )\rm )\rm ;
+\hspace*{0.20in}\it bit\_not\rm (\rm 1\rm ) \rm = \rm 0\rm ;
+
+\hspace*{0.20in} $\forall$ \rm (\it bb\rm )\rm .\rm (\it bb  $\in$  \it BIT  $\implies$  \it bit\_not\rm (\it bit\_not\rm (\it bb\rm )\rm ) \rm = \it bb\rm )\rm ;
 
 \vspace*{4mm}
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\implies$
+\hspace*{0.20in}\it bit\_and\rm (\rm 0\rm ,\rm 0\rm ) \rm = \rm 0\rm ;
 
-\hspace*{0.60in}\it bv\_size\rm (\it bv\_catenate\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm ) \rm + \it bv\_size\rm (\it v2\rm )\rm )\rm ;
+\hspace*{0.20in}\it bit\_and\rm (\rm 0\rm ,\rm 1\rm ) \rm = \rm 0\rm ;
+
+\hspace*{0.20in}\it bit\_and\rm (\rm 1\rm ,\rm 0\rm ) \rm = \rm 0\rm ;
+
+\hspace*{0.20in}\it bit\_and\rm (\rm 1\rm ,\rm 1\rm ) \rm = \rm 1\rm ;
+
+\hspace*{0.20in} $\forall$ \rm (\it b1\rm ,\it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\it bit\_and\rm (\it b1\rm , \it b2\rm ) \rm = \it bit\_and\rm (\it b2\rm ,\it b1\rm )\rm )\rm )\rm ;
+
+\hspace*{0.20in} $\forall$ \rm (\it b1\rm ,\it b2\rm ,\it b3\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it b3  $\in$  \it BIT  $\implies$  \rm (\it bit\_and\rm (\it b1\rm , \it bit\_and\rm (\it b2\rm ,\it b3\rm )\rm ) \rm = \it bit\_and\rm (\it bit\_and\rm (\it b1\rm ,\it b2\rm )\rm ,\it b3\rm )\rm )\rm )\rm ;
+
+\hspace*{0.20in} $\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\implies$  \rm (\it bit\_and\rm (\it b1\rm , \rm 1\rm ) \rm = \it b1\rm )\rm )\rm ;
+
+\hspace*{0.20in} $\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\implies$  \rm (\it bit\_and\rm (\it b1\rm , \rm 0\rm ) \rm = \rm 0\rm )\rm )\rm ;
 
 \vspace*{4mm}
-\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it low\rm ,\it high\rm )\rm .
-\rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it low  $\in$   $\nat$   $\land$  \it high  $\in$   $\nat$   $\land$  \it low  $\leq$  \it high  $\implies$
+\hspace*{0.20in}\it bit\_or\rm (\rm 0\rm ,\rm 0\rm ) \rm = \rm 0\rm ;
 
-\hspace*{0.60in}\it bv\_size\rm (\it bv\_sub\rm (\it bv\rm , \it low\rm , \it high\rm )\rm ) \rm = \rm 1\rm +\it high\rm -\it low\rm )\rm ;
+\hspace*{0.20in}\it bit\_or\rm (\rm 0\rm ,\rm 1\rm ) \rm = \rm 0\rm ;
+
+\hspace*{0.20in}\it bit\_or\rm (\rm 1\rm ,\rm 0\rm ) \rm = \rm 0\rm ;
+
+\hspace*{0.20in}\it bit\_or\rm (\rm 1\rm ,\rm 1\rm ) \rm = \rm 1\rm ;
+
+\hspace*{0.20in} $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\it bit\_or\rm (\it b1\rm , \it b2\rm ) \rm = \it bit\_or\rm (\it b2\rm ,\it b1\rm )\rm )\rm )\rm ;
+
+\hspace*{0.20in} $\forall$ \rm (\it b1\rm ,\it b2\rm ,\it b3\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it b3  $\in$  \it BIT  $\implies$  \rm (\it bit\_or\rm (\it b1\rm , \it bit\_or\rm (\it b2\rm ,\it b3\rm )\rm ) \rm = \it bit\_or\rm (\it bit\_or\rm (\it b1\rm ,\it b2\rm )\rm ,\it b3\rm )\rm )\rm )\rm ;
+
+\hspace*{0.20in} $\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\implies$  \rm (\it bit\_or\rm (\it b1\rm , \rm 1\rm ) \rm = \rm 1\rm )\rm )\rm ;
+
+\hspace*{0.20in} $\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\implies$  \rm (\it bit\_or\rm (\it b1\rm , \rm 0\rm ) \rm = \it b1\rm )\rm )\rm ;
 
 \vspace*{4mm}
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$
+\hspace*{0.20in}\it bit\_xor\rm (\rm 0\rm ,\rm 0\rm ) \rm = \rm 0\rm ;
 
-\hspace*{0.60in}\it bv\_size\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;
+\hspace*{0.20in}\it bit\_xor\rm (\rm 0\rm ,\rm 1\rm ) \rm = \rm 1\rm ;
 
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR\hspace*{0.10in} $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$
+\hspace*{0.20in}\it bit\_xor\rm (\rm 1\rm ,\rm 0\rm ) \rm = \rm 1\rm ;
 
-\hspace*{0.60in}\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm ) \rm = \it bv\_and\rm (\it v2\rm , \it v1\rm )\rm )\rm )\rm ;
+\hspace*{0.20in}\it bit\_xor\rm (\rm 1\rm ,\rm 1\rm ) \rm = \rm 0\rm ;
 
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it v3\rm )\rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR $\land$  \it v3  $\in$  \it BIT\_VECTOR\hspace*{0.10in} $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v3\rm )  $\implies$
+\hspace*{0.20in} $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\it bit\_xor\rm (\it b1\rm , \it b2\rm ) \rm = \it bit\_xor\rm (\it b2\rm ,\it b1\rm )\rm )\rm )\rm ;
 
-\hspace*{0.60in}\rm (\it bv\_and\rm (\it v1\rm , \it bv\_and\rm (\it v2\rm ,\it v3\rm )\rm ) \rm = \it bv\_and\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm , \it v3\rm )\rm )\rm )\rm ;
+\hspace*{0.20in} $\forall$ \rm (\it b1\rm ,\it b2\rm ,\it b3\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it b3  $\in$  \it BIT  $\implies$  \rm (\it bit\_xor\rm (\it b1\rm , \it bit\_xor\rm (\it b2\rm ,\it b3\rm )\rm ) \rm = \it bit\_xor\rm (\it bit\_xor\rm (\it b1\rm ,\it b2\rm )\rm ,\it b3\rm )\rm )\rm )\rm ;
 
-\hspace*{0.20in} $\forall$  \it bv\rm .
-\rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_and\rm (\it bv\rm , \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm )\rm ;
-
-\hspace*{0.20in} $\forall$  \it bv\rm .
-\rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_and\rm (\it bv\rm , \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\rm )\rm )\rm ;
+\hspace*{0.20in} $\forall$ \rm (\it bb\rm )\rm .\rm (\it bb  $\in$  \it BIT  $\implies$  \it bit\_xor\rm (\it bb\rm ,\it bb\rm ) \rm = \rm 0\rm )\rm ;
 
 \vspace*{4mm}
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$
+\hspace*{0.20in}\it bool\_to\_bit\rm (\bf TRUE\rm ) \rm = \rm 1\rm ;
 
-\hspace*{0.60in}\it bv\_size\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm )\rm )\rm ;
-
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$
-
-\hspace*{0.60in}\it bv\_size\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;
-
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$
-
-\hspace*{0.60in}\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm ) \rm = \it bv\_or\rm (\it v2\rm , \it v1\rm )\rm )\rm )\rm ;
-
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it v3\rm )\rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR $\land$  \it v3  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v3\rm )  $\implies$
-
-\hspace*{0.60in}\rm (\it bv\_or\rm (\it v1\rm , \it bv\_or\rm (\it v2\rm ,\it v3\rm )\rm ) \rm = \it bv\_or\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm , \it v3\rm )\rm )\rm )\rm ;
-
-\hspace*{0.20in} $\forall$  \it bv\rm .
-\rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_or\rm (\it bv\rm , \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm )\rm ;
-
-\hspace*{0.20in} $\forall$  \it bv\rm .
-\rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_or\rm (\it bv\rm , \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\rm )\rm )\rm ;
+\hspace*{0.20in}\it bool\_to\_bit\rm (\bf FALSE\rm ) \rm = \rm 0
 
 \vspace*{4mm}
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$
-
-\hspace*{0.60in}\it bv\_size\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm )\rm )\rm ;
-
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$
-
-\hspace*{0.60in}\it bv\_size\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;
-
-\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm .
-\rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$
-
-\hspace*{0.60in}\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm ) \rm = \it bv\_xor\rm (\it v2\rm , \it v1\rm )\rm )\rm )\rm ;
-
-\hspace*{0.20in} $\forall$  \rm (\it bv\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \it bv\_xor\rm (\it bv\rm ,\it bv\rm ) \rm = \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )
-
 \bf END\newpage

### 2008-09-10 developments/PIC16C432/PDB/ALU.tex

\documentstyle[11pt,supertabular]{article}
\setlength{\topmargin}{13mm}
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\textheight}{225mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm}
\input{psfig}

\begin{document}
\begin{sloppypar}

% DEFINITION DES CARACTERES MATHEMATIQUES B
%------------------------------------------
\def\@setmcodes#1#2#3{{\count0=#1 \count1=#3
	\loop \global\mathcode\count0=\count1 \ifnum \count0<#2
	\advance\count0 by1 \advance\count1 by1 \repeat}}

\@setmcodes{`A}{`Z}{"7441}
\@setmcodes{`a}{`z}{"7461}

\mathcode`\;="8000 % Makes ; active in math mode
{\catcode`\;=\active \gdef;{\semicolon\;}}
\mathchardef\semicolon="003B
%    Nominal distance from top of paper to top of page
\topmargin 0 pt
\textheight 53\baselineskip

%   Left margin on odd-numbered pages
\oddsidemargin  0.15 in
%   Left margin on even-numbered pages
\evensidemargin 0.35 in
%   Width of marginal notes.
\marginparwidth 1 in
%   Note that \oddsidemargin = \evensidemargin
\oddsidemargin 0.25 in
\evensidemargin 0.25 in
\marginparwidth 0.75 in
\textwidth 5.875 in % Width of text line.

\setlength{\parindent}{0pt}
\setlength{\parskip}{0ex}

% DEFINITION DES FONTS
%---------------------
% The AMS extra symbol fonts are loaded.
% Note: sometimes called euxm10
\font\msx=msam10
% Note: sometimes called euym10
\font\msy=msbm10

\newfam\msxfam \textfont\msxfam=\msx
\newfam\msyfam \textfont\msyfam=\msy

\def\famletter#1{\ifcase #1 0\or 1\or 2\or 3\or 4\or 5\or 6\or 7\or
	8\or 9\or A\or B\or C\or D\or E\or F\fi}

\edef\fx{\famletter\msxfam}
\edef\fy{\famletter\msyfam}

\def\bbold{\fam\msyfam \msy}

% SYMBOLES B
%-----------
% makes a quoted expression in mathematical text
\def\token#1{\hbox{`$#1$'}}
% used for error messages in Z specs
\def\report#1{\hbox{`{\tt #1}'}}

% \@myop makes an operator, with a strut to defeat TeX's vertical adjustment.
\def\@myop#1{\mathop{\mathstrut{#1}}\nolimits}

% This underscore doesn't have the little kern --- you get an italic
% correction anyway in math mode.
\def\_{\leavevmode \vbox{\hrule width0.5em}}

% Save \q as \xq for quantifiers q.
\let\xforall=\forall
\let\xexists=\exists
\let\xlambda=\lambda
\let\xmu=\mu

% \p and \f make arrows with 1 and 2 crossings resp.
\def\p#1{\mathrel{\ooalign{\hfil$\mapstochar\mkern 5mu$\hfil\cr$#1$}}}
\def\f#1{\mathrel{\ooalign{\hfil
	$\mapstochar\mkern 3mu\mapstochar\mkern 5mu$\hfil\cr$#1$}}}

\let\mc=\mathchardef

\def	\pow		{\mbox{${\cal P}$}}
\def	\po1		{\mbox{${\cal P}_1$}}
\let	\cross		\times
\def	\lambda		{\@myop{\xlambda}}
\def	\lnot		{\neg\;}
\def	\land		{\mathrel{\wedge}}
\def	\lor		{\mathrel{\vee}}
\let	\implies	\Rightarrow
\let	\iff		\Leftrightarrow
\def	\forall		{\@myop{\xforall}}
\def	\exists		{\@myop{\xexists}}
\def	\semi		{\mathrel{\comp}}
\def	\ssemi		{\mathbin{\rm ;}}
\let	\ensembleVide	\emptyset
\let	\rel		\leftrightarrow
\def	\dom		{\@myop{\sf dom}}
\def	\ran		{\@myop{\sf ran}}
\def	\id		{\@myop{\sf id}}
\def	\comp		{\mathbin{\raise
			0.6ex\hbox{\oalign{\hfil$\scriptscriptstyle
			\rm o$\hfil\cr\hfil$\scriptscriptstyle\rm 9$\hfil}}}}
\def	\para		{\mbox{$\mid\mid$}}
\mc	\dres		"2\fx43
\mc	\rres		"2\fx42
\def	\ndres		{\mathbin{{\dres} \llap{$-$}}}
\def	\nrres		{\mathbin{{\rres}\llap{$-$}}}
\def	\lover		{\mathbin{{\dres} \llap{$-\!\!\!\!-\!$}}}
\def	\rover		{\mathbin{{\rres}\llap{$\!-\!\!\!-$}}}
\let	\fun		\rightarrow
\def	\pfun		{\p\fun}
\def	\pinj		{\p\inj}
\mc	\inj		"3\fx1A
\def	\psurj		{\p\surj}
\mc	\surj		"3\fx10
\def	\bij		{\surj\!\!\!\!\!\!\!\inj}
\def	\nat		{\mbox{${\cal N}$}}
\def	\na1		{\mbox{${\cal N}_1$}}
\def	\num		{\mbox{${\cal Z}$}}
\def	\int		{\mbox{${\cal Z}$}}
\def	\rat		{\mbox{${\cal Q}$}}
\def	\div		{\mathbin{\rm /}}
\def	\mod		{\mathbin{\bf mod}}
\def	\upto		{\mathbin{\ldotp\ldotp}}
\def	\finset		{\mbox{${\cal F}$}}
\def	\finse1		{\mbox{${\cal F}_1$}}
\def	\ffun		{\f\fun}
\def	\finj		{\f\inj}
\def	\seq		{\@myop{\rm seq}}
\def	\cat		{\mathbin{\raise 0.8ex\hbox{$\mathchar"2\fx61$}}}
\def	\sep		{\hspace*{.05in}}

\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{0}

%-------------------%
% Debut du document %
%-------------------%

\bf MACHINE

\hspace*{0.20in}\it ALU

\vspace*{4mm}
\bf SEES

\hspace*{0.20in}\it TYPES\rm ,

\hspace*{0.20in}\it BIT\_DEFINITION\rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_ARITHMETICS\rm ,

\hspace*{0.20in}\it BYTE\_DEFINITION

\vspace*{4mm}
\bf CONSTANTS

\hspace*{0.20in}\it add\rm , \it substract\rm ,

\hspace*{0.20in}\it and\rm , \it ior\rm , \it xor\rm ,

\hspace*{0.20in}\it bitclear\rm , \it bitset\rm , \it bitget\rm ,

\hspace*{0.20in}\it complement\rm , \it swap\rm , \it rotateleft\rm , \it rotateright

\vspace*{4mm}
\bf PROPERTIES

\hspace*{0.20in}\it add \rm : \rm (\it UCHAR  $\times$  \it UCHAR\rm )  $\fun$  \rm (\it UCHAR  $\times$  \bf BOOL  $\times$  \bf BOOL  $\times$  \bf BOOL\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it substract \rm : \rm (\it UCHAR  $\times$  \it UCHAR\rm )  $\fun$  \rm (\it UCHAR  $\times$  \bf BOOL  $\times$  \bf BOOL\rm )  $\land$

\hspace*{0.20in} $\forall$  \rm (\it w1\rm , \it w2\rm , \it diff\rm )\rm .

\hspace*{0.30in}\rm (\it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it UCHAR  $\land$  \it diff  $\in$   $\int$   $\land$  \it diff \rm = \it w1\rm -\it w2  $\implies$

\hspace*{0.40in}\rm (\rm (\it diff $<$ \rm 0  $\implies$  \it substract\rm (\it w1\rm , \it w2\rm ) \rm = \rm (\it diff\rm +\rm 2\rm 5\rm 6\rm , \bf FALSE\rm , \bf TRUE\rm )\rm )  $\land$

\hspace*{0.45in}\rm (\it diff  $\geq$  \rm 0  $\implies$  \it substract\rm (\it w1\rm , \it w2\rm ) \rm = \rm (\it diff\rm , \bf bool\rm (\it diff\rm =\rm 0\rm )\rm , \bf FALSE\rm )\rm )\rm )\rm )  $\land$

\hspace*{0.20in}\it and \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \rm (\it BYTE  $\times$  \bf BOOL\rm )  $\land$

\hspace*{0.20in} $\forall$  \rm (\it w1\rm , \it w2\rm , \it ww\rm )\rm .

\hspace*{0.30in}\rm (\it w1  $\in$  \it BYTE  $\land$  \it w2  $\in$  \it BYTE  $\land$  \it ww  $\in$  \it BYTE  $\land$  \it ww \rm = \it bv\_and\rm (\it w1\rm , \it w2\rm )  $\implies$

\hspace*{0.60in}\it and\rm (\it w1\rm , \it w2\rm ) \rm = \rm (\it ww\rm , \bf bool\rm (\it bv\_to\_nat\rm (\it ww\rm ) \rm = \rm 0\rm )\rm )\rm )  $\land$

\hspace*{0.20in}\it ior \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \rm (\it BYTE  $\times$  \bf BOOL\rm )  $\land$

\hspace*{0.20in} $\forall$  \rm (\it w1\rm , \it w2\rm , \it ww\rm )\rm .

\hspace*{0.30in}\rm (\it w1  $\in$  \it BYTE  $\land$  \it w2  $\in$  \it BYTE  $\land$  \it ww  $\in$  \it BYTE  $\land$  \it ww \rm = \it bv\_or\rm (\it w1\rm , \it w2\rm )  $\implies$

\hspace*{0.60in}\it ior\rm (\it w1\rm , \it w2\rm ) \rm = \rm (\it ww\rm , \bf bool\rm (\it bv\_to\_nat\rm (\it ww\rm ) \rm = \rm 0\rm )\rm )\rm )  $\land$

\hspace*{0.20in}\it xor \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \rm (\it BYTE  $\times$  \bf BOOL\rm )  $\land$

\hspace*{0.20in} $\forall$  \rm (\it w1\rm , \it w2\rm , \it ww\rm )\rm .

\hspace*{0.30in}\rm (\it w1  $\in$  \it BYTE  $\land$  \it w2  $\in$  \it BYTE  $\land$  \it ww  $\in$  \it BYTE  $\implies$

\hspace*{0.40in}\rm (\it ww \rm = \it bv\_xor\rm (\it w1\rm , \it w2\rm )  $\implies$

\hspace*{0.60in}\it xor\rm (\it w1\rm , \it w2\rm ) \rm = \rm (\it ww\rm , \bf bool\rm (\it bv\_to\_nat\rm (\it ww\rm ) \rm = \rm 0\rm )\rm )\rm )\rm )  $\land$

\hspace*{0.20in}\it bitget \rm : \rm (\it BYTE  $\times$  \it BYTE\_INDEX\rm )  $\fun$  \it BIT  $\land$

\hspace*{0.20in} $\forall$  \rm (\it ww\rm , \it ii\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it BYTE\_INDEX  $\implies$  \it bitget\rm (\it ww\rm ,\it ii\rm ) \rm = \it ww\rm (\it ii\rm )\rm )  $\land$

\hspace*{0.20in}\it bitset \rm : \rm (\it BYTE  $\times$  \it BYTE\_INDEX\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in} $\forall$  \rm (\it ww\rm , \it ii\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it BYTE\_INDEX  $\implies$  \it bitset\rm (\it ww\rm ,\it ii\rm ) \rm = \it bv\_set\rm (\it ww\rm , \it ii\rm )\rm )  $\land$

\hspace*{0.20in}\it bitclear \rm : \rm (\it BYTE  $\times$  \it BYTE\_INDEX\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in} $\forall$  \rm (\it ww\rm , \it ii\rm , \it bb\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it BYTE\_INDEX  $\land$  \it bb  $\in$  \it BIT  $\implies$  \it bitclear\rm (\it ww\rm ,\it ii\rm ) \rm = \it bv\_clear\rm (\it ww\rm , \it ii\rm )\rm )  $\land$

\hspace*{0.20in}\it complement  $\in$  \it BYTE  $\fun$  \it BYTE  $\land$

\hspace*{0.20in} $\forall$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\implies$  \it complement\rm (\it ww\rm ) \rm = \it bv\_not\rm (\it ww\rm )\rm )  $\land$

\hspace*{0.20in}\it swap  $\in$  \it BYTE  $\fun$  \it BYTE  $\land$

\hspace*{0.20in} $\forall$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\implies$

\hspace*{0.30in}\rm (\it swap\rm (\it ww\rm ) \rm = \rm \{\rm 0 $\mapsto$ \it ww\rm (\rm 4\rm )\rm , \rm 1 $\mapsto$ \it ww\rm (\rm 5\rm )\rm , \rm 2 $\mapsto$ \it ww\rm (\rm 6\rm )\rm , \rm 3 $\mapsto$ \it ww\rm (\rm 7\rm )\rm , \rm 4 $\mapsto$ \it ww\rm (\rm 0\rm )\rm , \rm 5 $\mapsto$ \it ww\rm (\rm 1\rm )\rm , \rm 6 $\mapsto$ \it ww\rm (\rm 2\rm )\rm , \rm 7 $\mapsto$ \it ww\rm (\rm 3\rm )\rm \}\rm )\rm )  $\land$

\hspace*{0.20in}\it rotateleft  $\in$  \it BYTE  $\fun$  \it BYTE  $\times$  \bf BOOL  $\land$

\hspace*{0.20in} $\forall$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\implies$

\hspace*{0.30in}\rm (\it rotateleft\rm (\it ww\rm ) \rm = \rm (\rm \{\rm 0 $\mapsto$ \it ww\rm (\rm 7\rm )\rm , \rm 1 $\mapsto$ \it ww\rm (\rm 0\rm )\rm , \rm 2 $\mapsto$ \it ww\rm (\rm 1\rm )\rm , \rm 3 $\mapsto$ \it ww\rm (\rm 2\rm )\rm , \rm 4 $\mapsto$ \it ww\rm (\rm 3\rm )\rm , \rm 5 $\mapsto$ \it ww\rm (\rm 4\rm )\rm , \rm 6 $\mapsto$ \it ww\rm (\rm 5\rm )\rm , \rm 7 $\mapsto$ \it ww\rm (\rm 6\rm )\rm \}\rm ,

\hspace*{0.80in}\bf bool\rm (\it ww\rm (\rm 7\rm )\rm =\rm 1\rm )\rm )\rm )\rm )  $\land$

\hspace*{0.20in}\it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE  $\times$  \bf BOOL\hspace*{0.10in} $\land$

\hspace*{0.20in} $\forall$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it BYTE\hspace*{0.10in} $\implies$

\hspace*{0.30in}\rm (\it rotateright\rm (\it ww\rm ) \rm = \rm (\rm \{\rm 0 $\mapsto$ \it ww\rm (\rm 1\rm )\rm , \rm 1 $\mapsto$ \it ww\rm (\rm 2\rm )\rm , \rm 2 $\mapsto$ \it ww\rm (\rm 3\rm )\rm , \rm 3 $\mapsto$ \it ww\rm (\rm 4\rm )\rm , \rm 4 $\mapsto$ \it ww\rm (\rm 5\rm )\rm , \rm 5 $\mapsto$ \it ww\rm (\rm 6\rm )\rm , \rm 6 $\mapsto$ \it ww\rm (\rm 7\rm )\rm , \rm 7 $\mapsto$ \it ww\rm (\rm 0\rm )\rm \}\rm ,

\hspace*{0.80in}\bf bool\rm (\it ww\rm (\rm 0\rm )\rm =\rm 1\rm )\rm )\rm )\rm )

\vspace*{4mm}
\bf ASSERTIONS

\hspace*{0.20in}\bf dom\rm (\it add\rm ) \rm = \it UCHAR  $\times$  \it UCHAR\rm ;

\hspace*{0.20in}\bf ran\rm (\it add\rm )  $\subseteq$  \it UCHAR  $\times$  \bf BOOL  $\times$  \bf BOOL  $\times$  \bf BOOL\rm ;

\hspace*{0.20in}\bf dom\rm (\it substract\rm ) \rm = \it UCHAR  $\times$  \it UCHAR\rm ;

\hspace*{0.20in}\bf ran\rm (\it substract\rm )  $\subseteq$  \it UCHAR  $\times$  \bf BOOL  $\times$  \bf BOOL\rm ;

\hspace*{0.20in}\bf dom\rm (\it and\rm ) \rm = \it BYTE  $\times$  \it BYTE\rm ;

\hspace*{0.20in}\bf ran\rm (\it and\rm )  $\subseteq$  \it BYTE  $\times$  \bf BOOL\rm ;

\hspace*{0.20in}\bf dom\rm (\it ior\rm ) \rm = \it BYTE  $\times$  \it BYTE\rm ;

\hspace*{0.20in}\bf ran\rm (\it ior\rm )  $\subseteq$  \it BYTE  $\times$  \bf BOOL\rm ;

\hspace*{0.20in}\bf dom\rm (\it xor\rm ) \rm = \it BYTE  $\times$  \it BYTE\rm ;

\hspace*{0.20in}\bf ran\rm (\it xor\rm )  $\subseteq$  \it BYTE  $\times$  \bf BOOL\rm ;

\hspace*{0.20in}\bf dom\rm (\it bitclear\rm ) \rm = \it BYTE  $\times$  \it BYTE\_INDEX\rm ;

\hspace*{0.20in}\bf ran\rm (\it bitclear\rm )  $\subseteq$  \it BYTE\rm ;

\hspace*{0.20in}\bf dom\rm (\it bitset\rm ) \rm = \it BYTE  $\times$  \it BYTE\_INDEX\rm ;

\hspace*{0.20in}\bf ran\rm (\it bitset\rm )  $\subseteq$  \it BYTE\rm ;

\hspace*{0.20in}\bf dom\rm (\it bitget\rm ) \rm = \it BYTE  $\times$  \it BYTE\_INDEX\rm ;

\hspace*{0.20in}\bf ran\rm (\it bitget\rm )  $\subseteq$  \it BIT\rm ;

\hspace*{0.20in}\bf dom\rm (\it complement\rm ) \rm = \it BYTE\rm ;

\hspace*{0.20in}\bf ran\rm (\it complement\rm )  $\subseteq$  \it BYTE\rm ;

\hspace*{0.20in}\bf dom\rm (\it swap\rm ) \rm = \it BYTE\rm ;

\hspace*{0.20in}\bf ran\rm (\it swap\rm )  $\subseteq$  \it BYTE\rm ;

\hspace*{0.20in}\bf ran\rm (\it rotateleft\rm )  $\subseteq$  \it BYTE  $\times$  \bf BOOL\rm ;

\hspace*{0.20in}\bf dom\rm (\it rotateleft\rm ) \rm = \it BYTE\rm ;

\hspace*{0.20in}\bf dom\rm (\it rotateright\rm ) \rm = \it BYTE\rm ;

\hspace*{0.20in}\bf ran\rm (\it rotateright\rm )  $\subseteq$  \it BYTE  $\times$  \bf BOOL\rm ;

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it add\rm (\rm 0\rm ,\rm 0\rm ) \rm = \rm (\rm 0\rm , \bf TRUE\rm , \bf FALSE\rm , \bf FALSE\rm )\rm ;

\hspace*{0.20in}\it add\rm (\rm 2\rm 5\rm 5\rm ,\rm 1\rm ) \rm = \rm (\rm 0\rm , \bf TRUE\rm , \bf FALSE\rm , \bf TRUE\rm )

\bf END

\newpage
\end{sloppypar}\end{document}

### 2008-09-10 developments/PIC16C432/PDB/PIC.tex

--- 2008-09-10 developments/PIC16C432/PDB/ALU.tex
@@
 
+\vspace*{4mm}
 \bf MACHINE
+
+\hspace*{0.20in}\it PIC
+
+\vspace*{4mm}
+\bf SEES
+
+\hspace*{0.20in}\it BIT\_DEFINITION\rm ,
+
+\hspace*{0.20in}\it BYTE\_DEFINITION\rm ,
+
+\hspace*{0.20in}\it TYPES
+
+\vspace*{4mm}
+\bf USES
 
@@
 \vspace*{4mm}
-\bf SEES
+\bf INCLUDES
 
-\hspace*{0.20in}\it TYPES\rm ,
+\hspace*{0.20in}\it DATA\_MEMORY
 
-\hspace*{0.20in}\it BIT\_DEFINITION\rm ,
+\vspace*{4mm}
+\bf VARIABLES
 
-\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,
+\hspace*{0.20in}\it W\_REGISTER\rm ,
 
-\hspace*{0.20in}\it BIT\_VECTOR\_ARITHMETICS\rm ,
+\hspace*{0.20in}\it pc\rm ,
 
-\hspace*{0.20in}\it BYTE\_DEFINITION
+\hspace*{0.20in}\it stack\rm ,
+
+\hspace*{0.20in}\it sp
+
+\vspace*{4mm}
+\bf INVARIANT
+
+\hspace*{0.20in}\it W\_REGISTER  $\in$  \it BYTE  $\land$
+
+\hspace*{0.20in}\it pc  $\in$  \it INSTRUCTION  $\land$
+
+\hspace*{0.20in}\it sp  $\in$   $\nat$   $\land$
+
+\hspace*{0.20in}\it stack  $\in$   $\nat$   $\pfun$  \it INSTRUCTION  $\land$
+
+\hspace*{0.20in}\bf dom\rm (\it stack\rm ) \rm = \rm 0 $\upto$ \rm (\it sp\rm -\rm 1\rm )
 
@@
 
-\hspace*{0.20in}\it add\rm , \it substract\rm ,
+\hspace*{0.20in}
 
-\hspace*{0.20in}\it and\rm , \it ior\rm , \it xor\rm ,
+\hspace*{0.20in}\it TO\_POS\rm , \it PD\_POS\rm ,
 
-\hspace*{0.20in}\it bitclear\rm , \it bitset\rm , \it bitget\rm ,
+\hspace*{0.20in}\it ZERO\_POS\rm , \it DCARRY\_POS\rm , \it CARRY\_POS\rm ,
 
-\hspace*{0.20in}\it complement\rm , \it swap\rm , \it rotateleft\rm , \it rotateright
+\vspace*{4mm}
+\hspace*{0.20in}
+
+\hspace*{0.20in}\it get\_address\rm ,
+
+\hspace*{0.20in}\it get\_offset
 
@@
 
-\hspace*{0.20in}\it add \rm : \rm (\it UCHAR  $\times$  \it UCHAR\rm )  $\fun$  \rm (\it UCHAR  $\times$  \bf BOOL  $\times$  \bf BOOL  $\times$  \bf BOOL\rm )  $\land$
-
-\vspace*{4mm}
-\hspace*{0.20in}\it substract \rm : \rm (\it UCHAR  $\times$  \it UCHAR\rm )  $\fun$  \rm (\it UCHAR  $\times$  \bf BOOL  $\times$  \bf BOOL\rm )  $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it w1\rm , \it w2\rm , \it diff\rm )\rm .
-
-\hspace*{0.30in}\rm (\it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it UCHAR  $\land$  \it diff  $\in$   $\int$   $\land$  \it diff \rm = \it w1\rm -\it w2  $\implies$
-
-\hspace*{0.40in}\rm (\rm (\it diff $<$ \rm 0  $\implies$  \it substract\rm (\it w1\rm , \it w2\rm ) \rm = \rm (\it diff\rm +\rm 2\rm 5\rm 6\rm , \bf FALSE\rm , \bf TRUE\rm )\rm )  $\land$
-
-\hspace*{0.45in}\rm (\it diff  $\geq$  \rm 0  $\implies$  \it substract\rm (\it w1\rm , \it w2\rm ) \rm = \rm (\it diff\rm , \bf bool\rm (\it diff\rm =\rm 0\rm )\rm , \bf FALSE\rm )\rm )\rm )\rm )  $\land$
-
-\hspace*{0.20in}\it and \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \rm (\it BYTE  $\times$  \bf BOOL\rm )  $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it w1\rm , \it w2\rm , \it ww\rm )\rm .
-
-\hspace*{0.30in}\rm (\it w1  $\in$  \it BYTE  $\land$  \it w2  $\in$  \it BYTE  $\land$  \it ww  $\in$  \it BYTE  $\land$  \it ww \rm = \it bv\_and\rm (\it w1\rm , \it w2\rm )  $\implies$
-
-\hspace*{0.60in}\it and\rm (\it w1\rm , \it w2\rm ) \rm = \rm (\it ww\rm , \bf bool\rm (\it bv\_to\_nat\rm (\it ww\rm ) \rm = \rm 0\rm )\rm )\rm )  $\land$
-
-\hspace*{0.20in}\it ior \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \rm (\it BYTE  $\times$  \bf BOOL\rm )  $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it w1\rm , \it w2\rm , \it ww\rm )\rm .
-
-\hspace*{0.30in}\rm (\it w1  $\in$  \it BYTE  $\land$  \it w2  $\in$  \it BYTE  $\land$  \it ww  $\in$  \it BYTE  $\land$  \it ww \rm = \it bv\_or\rm (\it w1\rm , \it w2\rm )  $\implies$
-
-\hspace*{0.60in}\it ior\rm (\it w1\rm , \it w2\rm ) \rm = \rm (\it ww\rm , \bf bool\rm (\it bv\_to\_nat\rm (\it ww\rm ) \rm = \rm 0\rm )\rm )\rm )  $\land$
-
-\hspace*{0.20in}\it xor \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \rm (\it BYTE  $\times$  \bf BOOL\rm )  $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it w1\rm , \it w2\rm , \it ww\rm )\rm .
-
-\hspace*{0.30in}\rm (\it w1  $\in$  \it BYTE  $\land$  \it w2  $\in$  \it BYTE  $\land$  \it ww  $\in$  \it BYTE  $\implies$
-
-\hspace*{0.40in}\rm (\it ww \rm = \it bv\_xor\rm (\it w1\rm , \it w2\rm )  $\implies$
-
-\hspace*{0.60in}\it xor\rm (\it w1\rm , \it w2\rm ) \rm = \rm (\it ww\rm , \bf bool\rm (\it bv\_to\_nat\rm (\it ww\rm ) \rm = \rm 0\rm )\rm )\rm )\rm )  $\land$
-
-\hspace*{0.20in}\it bitget \rm : \rm (\it BYTE  $\times$  \it BYTE\_INDEX\rm )  $\fun$  \it BIT  $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it ww\rm , \it ii\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it BYTE\_INDEX  $\implies$  \it bitget\rm (\it ww\rm ,\it ii\rm ) \rm = \it ww\rm (\it ii\rm )\rm )  $\land$
-
-\hspace*{0.20in}\it bitset \rm : \rm (\it BYTE  $\times$  \it BYTE\_INDEX\rm )  $\fun$  \it BYTE  $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it ww\rm , \it ii\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it BYTE\_INDEX  $\implies$  \it bitset\rm (\it ww\rm ,\it ii\rm ) \rm = \it bv\_set\rm (\it ww\rm , \it ii\rm )\rm )  $\land$
-
-\hspace*{0.20in}\it bitclear \rm : \rm (\it BYTE  $\times$  \it BYTE\_INDEX\rm )  $\fun$  \it BYTE  $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it ww\rm , \it ii\rm , \it bb\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it BYTE\_INDEX  $\land$  \it bb  $\in$  \it BIT  $\implies$  \it bitclear\rm (\it ww\rm ,\it ii\rm ) \rm = \it bv\_clear\rm (\it ww\rm , \it ii\rm )\rm )  $\land$
-
-\hspace*{0.20in}\it complement  $\in$  \it BYTE  $\fun$  \it BYTE  $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\implies$  \it complement\rm (\it ww\rm ) \rm = \it bv\_not\rm (\it ww\rm )\rm )  $\land$
-
-\hspace*{0.20in}\it swap  $\in$  \it BYTE  $\fun$  \it BYTE  $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\implies$
-
-\hspace*{0.30in}\rm (\it swap\rm (\it ww\rm ) \rm = \rm \{\rm 0 $\mapsto$ \it ww\rm (\rm 4\rm )\rm , \rm 1 $\mapsto$ \it ww\rm (\rm 5\rm )\rm , \rm 2 $\mapsto$ \it ww\rm (\rm 6\rm )\rm , \rm 3 $\mapsto$ \it ww\rm (\rm 7\rm )\rm , \rm 4 $\mapsto$ \it ww\rm (\rm 0\rm )\rm , \rm 5 $\mapsto$ \it ww\rm (\rm 1\rm )\rm , \rm 6 $\mapsto$ \it ww\rm (\rm 2\rm )\rm , \rm 7 $\mapsto$ \it ww\rm (\rm 3\rm )\rm \}\rm )\rm )  $\land$
-
-\hspace*{0.20in}\it rotateleft  $\in$  \it BYTE  $\fun$  \it BYTE  $\times$  \bf BOOL  $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\implies$
-
-\hspace*{0.30in}\rm (\it rotateleft\rm (\it ww\rm ) \rm = \rm (\rm \{\rm 0 $\mapsto$ \it ww\rm (\rm 7\rm )\rm , \rm 1 $\mapsto$ \it ww\rm (\rm 0\rm )\rm , \rm 2 $\mapsto$ \it ww\rm (\rm 1\rm )\rm , \rm 3 $\mapsto$ \it ww\rm (\rm 2\rm )\rm , \rm 4 $\mapsto$ \it ww\rm (\rm 3\rm )\rm , \rm 5 $\mapsto$ \it ww\rm (\rm 4\rm )\rm , \rm 6 $\mapsto$ \it ww\rm (\rm 5\rm )\rm , \rm 7 $\mapsto$ \it ww\rm (\rm 6\rm )\rm \}\rm ,
-
-\hspace*{0.80in}\bf bool\rm (\it ww\rm (\rm 7\rm )\rm =\rm 1\rm )\rm )\rm )\rm )  $\land$
-
-\hspace*{0.20in}\it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE  $\times$  \bf BOOL\hspace*{0.10in} $\land$
-
-\hspace*{0.20in} $\forall$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it BYTE\hspace*{0.10in} $\implies$
-
-\hspace*{0.30in}\rm (\it rotateright\rm (\it ww\rm ) \rm = \rm (\rm \{\rm 0 $\mapsto$ \it ww\rm (\rm 1\rm )\rm , \rm 1 $\mapsto$ \it ww\rm (\rm 2\rm )\rm , \rm 2 $\mapsto$ \it ww\rm (\rm 3\rm )\rm , \rm 3 $\mapsto$ \it ww\rm (\rm 4\rm )\rm , \rm 4 $\mapsto$ \it ww\rm (\rm 5\rm )\rm , \rm 5 $\mapsto$ \it ww\rm (\rm 6\rm )\rm , \rm 6 $\mapsto$ \it ww\rm (\rm 7\rm )\rm , \rm 7 $\mapsto$ \it ww\rm (\rm 0\rm )\rm \}\rm ,
-
-\hspace*{0.80in}\bf bool\rm (\it ww\rm (\rm 0\rm )\rm =\rm 1\rm )\rm )\rm )\rm )
-
-\vspace*{4mm}
-\bf ASSERTIONS
-
-\hspace*{0.20in}\bf dom\rm (\it add\rm ) \rm = \it UCHAR  $\times$  \it UCHAR\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it add\rm )  $\subseteq$  \it UCHAR  $\times$  \bf BOOL  $\times$  \bf BOOL  $\times$  \bf BOOL\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it substract\rm ) \rm = \it UCHAR  $\times$  \it UCHAR\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it substract\rm )  $\subseteq$  \it UCHAR  $\times$  \bf BOOL  $\times$  \bf BOOL\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it and\rm ) \rm = \it BYTE  $\times$  \it BYTE\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it and\rm )  $\subseteq$  \it BYTE  $\times$  \bf BOOL\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it ior\rm ) \rm = \it BYTE  $\times$  \it BYTE\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it ior\rm )  $\subseteq$  \it BYTE  $\times$  \bf BOOL\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it xor\rm ) \rm = \it BYTE  $\times$  \it BYTE\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it xor\rm )  $\subseteq$  \it BYTE  $\times$  \bf BOOL\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it bitclear\rm ) \rm = \it BYTE  $\times$  \it BYTE\_INDEX\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it bitclear\rm )  $\subseteq$  \it BYTE\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it bitset\rm ) \rm = \it BYTE  $\times$  \it BYTE\_INDEX\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it bitset\rm )  $\subseteq$  \it BYTE\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it bitget\rm ) \rm = \it BYTE  $\times$  \it BYTE\_INDEX\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it bitget\rm )  $\subseteq$  \it BIT\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it complement\rm ) \rm = \it BYTE\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it complement\rm )  $\subseteq$  \it BYTE\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it swap\rm ) \rm = \it BYTE\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it swap\rm )  $\subseteq$  \it BYTE\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it rotateleft\rm )  $\subseteq$  \it BYTE  $\times$  \bf BOOL\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it rotateleft\rm ) \rm = \it BYTE\rm ;
-
-\hspace*{0.20in}\bf dom\rm (\it rotateright\rm ) \rm = \it BYTE\rm ;
-
-\hspace*{0.20in}\bf ran\rm (\it rotateright\rm )  $\subseteq$  \it BYTE  $\times$  \bf BOOL\rm ;
-
 \vspace*{4mm}
@@
 
-\hspace*{0.20in}\it add\rm (\rm 0\rm ,\rm 0\rm ) \rm = \rm (\rm 0\rm , \bf TRUE\rm , \bf FALSE\rm , \bf FALSE\rm )\rm ;
+\hspace*{0.20in}\it TO\_POS  $\in$  \it BYTE\_INDEX  $\land$ \hspace*{0.20in}\it TO\_POS \rm = \rm 4  $\land$
 
-\hspace*{0.20in}\it add\rm (\rm 2\rm 5\rm 5\rm ,\rm 1\rm ) \rm = \rm (\rm 0\rm , \bf TRUE\rm , \bf FALSE\rm , \bf TRUE\rm )
+\hspace*{0.20in}\it PD\_POS  $\in$  \it BYTE\_INDEX  $\land$ \hspace*{0.20in}\it PD\_POS \rm = \rm 3  $\land$
+
+\hspace*{0.20in}\it ZERO\_POS  $\in$  \it BYTE\_INDEX  $\land$ \hspace*{0.20in}\it ZERO\_POS \rm = \rm 2  $\land$
+
+\hspace*{0.20in}\it DCARRY\_POS  $\in$  \it BYTE\_INDEX  $\land$ \hspace*{0.20in}\it DCARRY\_POS \rm = \rm 1  $\land$
+
+\hspace*{0.20in}\it CARRY\_POS  $\in$  \it BYTE\_INDEX  $\land$ \hspace*{0.20in}\it CARRY\_POS \rm = \rm 0
+
+\vspace*{4mm}
+\bf INITIALISATION
+
+\hspace*{0.20in}\it W\_REGISTER  $:\in$  \it BYTE  $\para$
+
+\hspace*{0.20in}\it pc  $:\in$  \it INSTRUCTION  $\para$
+
+\hspace*{0.20in}\it stack \rm :=  $\ensembleVide$   $\para$
+
+\hspace*{0.20in}\it sp \rm := \rm 0
+
+\vspace*{4mm}
+\bf OPERATIONS
+
+\hspace*{0.20in}\bf CALL \rm (\it kk\rm ) \rm =
+
+\hspace*{0.40in}\bf PRE \it kk  $\in$  \it INSTRUCTION \bf THEN
+
+\hspace*{0.60in}\bf stack\rm (\it sp\rm ) \rm := \it instruction\_next\rm (\it pc\rm )  $\para$
+
+\hspace*{0.60in}\it sp \rm := \it sp \rm + \rm 1  $\para$
+
+\hspace*{0.60in}\it pc \rm := \it kk
+
+\hspace*{0.40in}\bf END
+
+\hspace*{0.20in}\rm ;
+
+\hspace*{0.20in}\bf RETURN \rm =
+
+\hspace*{0.40in}\bf PRE \it sp $>$ \rm 0 \bf THEN
+
+\hspace*{0.60in}\it stack \rm := \rm \{\it sp \rm - \rm 1\rm \}  $\ndres$  \it stack  $\para$
+
+\hspace*{0.60in}\it pc \rm := \it stack\rm (\it sp \rm - \rm 1\rm )  $\para$
+
+\hspace*{0.60in}\it sp \rm := \it sp \rm - \rm 1
+
+\hspace*{0.40in}\bf END
+
+\hspace*{0.20in}\rm ;
+
+\hspace*{0.20in}\bf RETLW \rm (\it kk\rm ) \rm =
+
+\hspace*{0.40in}\bf PRE \it kk  $\in$  \it UCHAR  $\land$  \it sp $>$ \rm 0 \bf THEN
+
+\hspace*{0.60in}\it pc \rm := \it stack\rm (\it sp \rm - \rm 1\rm )  $\para$
+
+\hspace*{0.60in}\it W\_REGISTER \rm := \it uchar\_to\_byte\rm (\it kk\rm )
+
+\hspace*{0.40in}\bf END
+
+\hspace*{0.20in}\rm ;
+
+\hspace*{0.20in}\bf ADDLW \rm (\it kk\rm ) \rm =
+
+\hspace*{0.40in}\bf PRE \it kk  $\in$  \it UCHAR \bf THEN
+
+\hspace*{0.60in}\bf ANY \it result\rm , \it carry\rm , \it digit\_carry\rm , \it zero \bf WHERE
+
+\hspace*{0.80in}\it result  $\in$  \it UCHAR  $\land$  \it digit\_carry  $\in$  \bf BOOL  $\land$  \it carry  $\in$  \bf BOOL  $\land$  \it zero  $\in$  \bf BOOL  $\land$
+
+\hspace*{0.80in}\it result\rm , \it digit\_carry\rm , \it carry\rm , \it zero \rm = \it add\rm (\it kk\rm , \it byte\_to\_uchar\rm (\it W\_REGISTER\rm )\rm )
+
+\hspace*{0.60in}\bf THEN
+
+\hspace*{0.80in}\it W\_REGISTER \rm := \it uchar\_to\_byte\rm (\it result\rm )  $\para$
+
+\hspace*{0.80in}\it mem \rm := \it update\_status\rm (\it zero\rm , \it digit\_carry\rm , \it carry\rm )
+
+\hspace*{0.60in}\bf END  $\para$
+
+\hspace*{0.60in}\it pc \rm := \it instruction\_next\rm (\it pc\rm )
+
+\hspace*{0.40in}\bf END
+
+\hspace*{0.20in}\rm ;
+
+\hspace*{0.20in}\bf ADDWF \rm (\it ff\rm , \it dd\rm ) \rm =
+
+\hspace*{0.40in}\bf PRE \it ff  $\in$  \it REGISTER0  $\land$  \it dd  $\in$  \it BIT \bf THEN
+
+\hspace*{0.60in}\bf ANY \it dest\rm , \it result\rm , \it carry\rm , \it digit\_carry\rm , \it zero \bf WHERE
+
+\hspace*{0.80in}\it dest  $\in$  \it REGISTER  $\land$
+
+\hspace*{0.80in}\it result  $\in$  \it UCHAR  $\land$  \it digit\_carry  $\in$  \bf BOOL  $\land$  \it carry  $\in$  \bf BOOL  $\land$  \it zero  $\in$  \bf BOOL  $\land$
+
+\hspace*{0.80in}\it dest \rm = \it get\_address\rm (\it ff\rm , \it RP0\_VAL\rm , \it byte\_to\_uchar\rm (\it FSR\_VAL\rm )\rm )  $\land$
+
+\hspace*{0.80in}\it result\rm , \it digit\_carry\rm , \it carry\rm , \it zero \rm = \it add\rm (\it byte\_to\_uchar\rm (\it mem\rm (\it dest\rm )\rm )\rm , \it byte\_to\_uchar\rm (\it W\_REGISTER\rm )\rm )
+
+\hspace*{0.60in}\bf THEN
+
+\hspace*{0.80in}\bf ANY \it mem2\rm , \it bv \bf WHERE
+
+\hspace*{1.00in}\it bv  $\in$  \it BYTE  $\land$  \it bv \rm = \it uchar\_to\_byte\rm (\it result\rm )  $\land$
+
+\hspace*{1.00in}\it mem2  $\in$  \it REGISTER  $\fun$  \it BYTE  $\land$
+
+\hspace*{1.00in}\rm (\it dd  $\not =$  \rm 0  $\land$  \it dest  $\not\in$  \it UNIMPLEMENTED\_LOCATIONS  $\implies$  \it mem2 \rm = \it mem  $\lover$  \rm \{ \it dest  $\mapsto$  \it bv \rm \}\rm )  $\land$
+
+\hspace*{1.00in}\rm (\it dd \rm = \rm 0  $\lor$  \it dest  $\in$  \it UNIMPLEMENTED\_LOCATIONS  $\implies$  \it mem2 \rm = \it mem\rm )
+
+\hspace*{0.80in}\bf THEN
+
+\hspace*{1.00in}\bf IF \it dd \rm = \rm 0 \bf THEN
+
+\hspace*{1.20in}\it W\_REGISTER \rm := \it bv
+
+\hspace*{1.00in}\bf END  $\para$
+
+\hspace*{1.00in}\bf IF \it ff  $\not =$  \it STATUS\_ADDR0 \bf THEN
+
+\hspace*{1.20in}\it mem \rm := \it update\_status\rm (\it zero\rm , \it digit\_carry\rm , \it carry\rm )
+
+\hspace*{1.00in}\bf END
+
+\hspace*{0.80in}\bf END
+
+\hspace*{0.60in}\bf END  $\para$
+
+\hspace*{0.60in}\it pc \rm := \it instruction\_next\rm (\it pc\rm )
+
+\hspace*{0.40in}\bf END
 

### 2008-10-02 developments/platforms/8051H/bdp/ALU.n.tex
8051H com memória separada - incompleto

\documentstyle[11pt,supertabular]{article}
\setlength{\topmargin}{13mm}
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\textheight}{225mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm}

\begin{document}
\begin{sloppypar}

% DEFINITION DES CARACTERES MATHEMATIQUES B
%------------------------------------------
\def\@setmcodes#1#2#3{{\count0=#1 \count1=#3
	\loop \global\mathcode\count0=\count1 \ifnum \count0<#2
	\advance\count0 by1 \advance\count1 by1 \repeat}}

\@setmcodes{`A}{`Z}{"7441}
\@setmcodes{`a}{`z}{"7461}

\mathcode`\;="8000 % Makes ; active in math mode
{\catcode`\;=\active \gdef;{\semicolon\;}}
\mathchardef\semicolon="003B
%    Nominal distance from top of paper to top of page
\topmargin 0 pt
\textheight 53\baselineskip

%   Left margin on odd-numbered pages
\oddsidemargin  0.15 in
%   Left margin on even-numbered pages
\evensidemargin 0.35 in
%   Width of marginal notes.
\marginparwidth 1 in
%   Note that \oddsidemargin = \evensidemargin
\oddsidemargin 0.25 in
\evensidemargin 0.25 in
\marginparwidth 0.75 in
\textwidth 5.875 in % Width of text line.

\setlength{\parindent}{0pt}
\setlength{\parskip}{0ex}

% DEFINITION DES FONTS
%---------------------
% The AMS extra symbol fonts are loaded.
% Note: sometimes called euxm10
\font\msx=msam10
% Note: sometimes called euym10
\font\msy=msbm10

\newfam\msxfam \textfont\msxfam=\msx
\newfam\msyfam \textfont\msyfam=\msy

\def\famletter#1{\ifcase #1 0\or 1\or 2\or 3\or 4\or 5\or 6\or 7\or
	8\or 9\or A\or B\or C\or D\or E\or F\fi}

\edef\fx{\famletter\msxfam}
\edef\fy{\famletter\msyfam}

\def\bbold{\fam\msyfam \msy}

% SYMBOLES B
%-----------
% makes a quoted expression in mathematical text
\def\token#1{\hbox{`$#1$'}}
% used for error messages in Z specs
\def\report#1{\hbox{`{\tt #1}'}}

% \@myop makes an operator, with a strut to defeat TeX's vertical adjustment.
\def\@myop#1{\mathop{\mathstrut{#1}}\nolimits}

% This underscore doesn't have the little kern --- you get an italic
% correction anyway in math mode.
\def\_{\leavevmode \vbox{\hrule width0.5em}}

% Save \q as \xq for quantifiers q.
\let\xforall=\forall
\let\xexists=\exists
\let\xlambda=\lambda
\let\xmu=\mu

% \p and \f make arrows with 1 and 2 crossings resp.
\def\p#1{\mathrel{\ooalign{\hfil$\mapstochar\mkern 5mu$\hfil\cr$#1$}}}
\def\f#1{\mathrel{\ooalign{\hfil
	$\mapstochar\mkern 3mu\mapstochar\mkern 5mu$\hfil\cr$#1$}}}

\let\mc=\mathchardef

\def	\pow		{\mbox{${\cal P}$}}
\def	\po1		{\mbox{${\cal P}_1$}}
\let	\cross		\times
\def	\lambda		{\@myop{\xlambda}}
\def	\lnot		{\neg\;}
\def	\land		{\mathrel{\wedge}}
\def	\lor		{\mathrel{\vee}}
\let	\implies	\Rightarrow
\let	\iff		\Leftrightarrow
\def	\forall		{\@myop{\xforall}}
\def	\exists		{\@myop{\xexists}}
\def	\semi		{\mathrel{\comp}}
\def	\ssemi		{\mathbin{\rm ;}}
\let	\ensembleVide	\emptyset
\let	\rel		\leftrightarrow
\def	\dom		{\@myop{\sf dom}}
\def	\ran		{\@myop{\sf ran}}
\def	\id		{\@myop{\sf id}}
\def	\comp		{\mathbin{\raise
			0.6ex\hbox{\oalign{\hfil$\scriptscriptstyle
			\rm o$\hfil\cr\hfil$\scriptscriptstyle\rm 9$\hfil}}}}
\def	\para		{\mbox{$\mid\mid$}}
\mc	\dres		"2\fx43
\mc	\rres		"2\fx42
\def	\ndres		{\mathbin{{\dres} \llap{$-$}}}
\def	\nrres		{\mathbin{{\rres}\llap{$-$}}}
\def	\lover		{\mathbin{{\dres} \llap{$-\!\!\!\!-\!$}}}
\def	\rover		{\mathbin{{\rres}\llap{$\!-\!\!\!-$}}}
\let	\fun		\rightarrow
\def	\pfun		{\p\fun}
\def	\pinj		{\p\inj}
\mc	\inj		"3\fx1A
\def	\psurj		{\p\surj}
\mc	\surj		"3\fx10
\def	\bij		{\surj\!\!\!\!\!\!\!\inj}
\def	\nat		{\mbox{${\cal N}$}}
\def	\na1		{\mbox{${\cal N}_1$}}
\def	\num		{\mbox{${\cal Z}$}}
\def	\int		{\mbox{${\cal Z}$}}
\def	\rat		{\mbox{${\cal Q}$}}
\def	\div		{\mathbin{\rm /}}
\def	\mod		{\mathbin{\bf mod}}
\def	\upto		{\mathbin{\ldotp\ldotp}}
\def	\finset		{\mbox{${\cal F}$}}
\def	\finse1		{\mbox{${\cal F}_1$}}
\def	\ffun		{\f\fun}
\def	\finj		{\f\inj}
\def	\seq		{\@myop{\rm seq}}
\def	\cat		{\mathbin{\raise 0.8ex\hbox{$\mathchar"2\fx61$}}}
\def	\sep		{\hspace*{.05in}}

\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{0}

%-------------------%
% Debut du document %
%-------------------%

\bf MACHINE

\hspace*{0.10in}\it ALU

\bf SEES

\hspace*{0.10in}\it TYPES

\bf CONSTANTS

\hspace*{0.10in}\it parity\rm ,\it soma\rm ,\it bitget\rm ,\it bitset\rm ,\it bitclear\rm ,\it bitcomplement\rm ,\it bitput\rm ,\it pswcalc\rm ,\it accparity\rm ,\it add\rm ,\it diff\rm ,\it substract\rm ,\it and\rm ,\it ior\rm ,\it xor\rm ,\it \\complement\rm ,\it swap\rm ,\it rotateleft\rm ,\it rotateright

\bf PROPERTIES

\hspace*{0.10in}\it parity $\in$  \it WORD  $\fun$  \it BIT  $\land$

\hspace*{0.10in} $\forall$ \it w1\rm .\rm (\it w1 $\in$  \it WORD  $\implies$  \it parity\rm (\it w1\rm ) \rm = \it BV8\_PAR\rm (\it WORD\_TO\_BV\rm (\it w1\rm )\rm )\rm )  $\land$

\hspace*{0.10in}\it soma $\in$  \it WORD $\times$ \it WORD  $\fun$  \it WORD $\times$ \it BIT  $\land$

\hspace*{0.10in} $\forall$ \rm (\it w1\rm ,\it w2\rm ,\it su\rm )\rm .\rm (\it w1 $\in$  \it WORD  $\land$  \it w2 $\in$  \it WORD  $\land$  \it su $\in$   $\nat$   $\land$  \it su \rm = \it w2\rm +\it w1  $\implies$  \rm (\it su $\leq$ \rm 2\rm 5\rm 5  $\implies$  \it soma\rm (\it w1\rm ,\it w2\rm ) \rm = \it su\rm ,\rm 0\rm )  $\land$  \rm (\it su $\geq$ \rm 2\rm 5\rm 6  $\implies$  \it soma\rm (\it w1\rm ,\it w2\rm ) \rm = \it su\rm -\rm 2\rm 5\rm 5\rm ,\rm 1\rm )\rm )  $\land$

\hspace*{0.10in}\it bitget $\in$  \it WORD $\times$ \it WORD\_POSITION  $\fun$  \it BIT  $\land$

\hspace*{0.10in} $\forall$ \rm (\it wd\rm ,\it ind\rm )\rm .\rm (\it wd $\in$  \it WORD  $\land$  \it ind $\in$  \it WORD\_POSITION  $\implies$  \it bitget\rm (\it wd\rm ,\it ind\rm ) \rm = \it WORD\_TO\_BV\rm (\it wd\rm )\rm (\it ind\rm )\rm )  $\land$

\hspace*{0.10in}\it bitset $\in$  \it WORD $\times$ \it WORD\_POSITION  $\fun$  \it WORD  $\land$

\hspace*{0.10in} $\forall$ \rm (\it wd\rm ,\it ind\rm )\rm .\rm (\it wd $\in$  \it WORD  $\land$  \it ind $\in$  \it WORD\_POSITION  $\implies$  \it bitset\rm (\it wd\rm ,\it ind\rm ) \rm = \it BV\_TO\_WORD\rm (\it BV8\_SET\_BIT\rm (\it WORD\_TO\_BV\rm (\it wd\rm )\rm ,\it ind\rm ,\rm 1\rm )\rm )\rm )  $\land$

\hspace*{0.10in}\it bitclear $\in$  \it WORD $\times$ \it WORD\_POSITION  $\fun$  \it WORD  $\land$

\hspace*{0.10in} $\forall$ \rm (\it wd\rm ,\it ind\rm ,\it bit\rm )\rm .\rm (\it wd $\in$  \it WORD  $\land$  \it ind $\in$  \it WORD\_POSITION  $\land$  \it bit $\in$  \it BIT  $\implies$  \it bitclear\rm (\it wd\rm ,\it ind\rm ) \rm = \it BV\_TO\_WORD\rm (\it BV8\_SET\_BIT\rm (\it WORD\_TO\_BV\rm (\it wd\rm )\rm ,\it ind\rm ,\rm 0\rm )\rm )\rm )  $\land$

\hspace*{0.10in}\it bitcomplement $\in$  \it WORD $\times$ \it WORD\_POSITION  $\fun$  \it WORD  $\land$

\hspace*{0.10in} $\forall$ \rm (\it wd\rm ,\it ind\rm ,\it bit\rm )\rm .\rm (\it wd $\in$  \it WORD  $\land$  \it ind $\in$  \it WORD\_POSITION  $\land$  \it bit $\in$  \it BIT  $\land$  \it bit \rm = \it bitget\rm (\it wd\rm ,\it ind\rm )  $\implies$  \rm (\it bit \rm = \rm 1  $\implies$  \it bitcomplement\rm (\it wd\rm ,\it ind\rm ) \rm = \it bitclear\rm (\it wd\rm ,\it ind\rm )  $\land$  \it bit $\not =$ \rm 1  $\implies$  \it bitcomplement\rm (\it wd\rm ,\it ind\rm ) \rm = \it bitset\rm (\it wd\rm ,\it ind\rm )\rm )\rm )  $\land$

\hspace*{0.10in}\it bitput $\in$  \it WORD $\times$ \it WORD\_POSITION $\times$ \it BIT  $\fun$  \it WORD  $\land$

\hspace*{0.10in} $\forall$ \rm (\it wd\rm ,\it ind\rm ,\it bit\rm )\rm .\rm (\it wd $\in$  \it WORD  $\land$  \it ind $\in$  \it WORD\_POSITION  $\land$  \it bit $\in$  \it BIT  $\implies$  \rm (\it bit \rm = \rm 0  $\implies$  \it bitput\rm (\it wd\rm ,\it ind\rm ,\it bit\rm ) \rm = \it bitclear\rm (\it wd\rm ,\it ind\rm )  $\land$  \it bit \rm = \rm 1  $\implies$  \it bitput\rm (\it wd\rm ,\it ind\rm ,\it bit\rm ) \rm = \it bitset\rm (\it wd\rm ,\it ind\rm )\rm )\rm )  $\land$

\hspace*{0.10in}\it pswcalc $\in$  \it WORD $\times$ \it WORD  $\fun$  \it BIT $\times$ \it BIT $\times$ \it BIT $\times$ \it BIT $\times$ \it BIT  $\land$

\hspace*{0.10in} $\forall$ \rm (\it psw\rm ,\it acc\rm ,\it psw\_new\rm )\rm .\rm (\it psw $\in$  \it WORD  $\land$  \it acc $\in$  \it WORD  $\land$  \it psw\_new $\in$  \it BV8  $\land$  \it psw\_new\rm (\rm 0\rm ) \rm = \it parity\rm (\it acc\rm )  $\land$  \it psw\_new\rm (\rm 1\rm ) \rm = \it bitget\rm (\it psw\rm ,\rm 1\rm )  $\land$  \it psw\_new\rm (\rm 3\rm ) \rm = \it bitget\rm (\it psw\rm ,\rm 3\rm )  $\land$  \it psw\_new\rm (\rm 4\rm ) \rm = \it bitget\rm (\it psw\rm ,\rm 4\rm )  $\land$  \it psw\_new\rm (\rm 5\rm ) \rm = \it bitget\rm (\it psw\rm ,\rm 5\rm )  $\implies$  \it pswcalc\rm (\it psw\rm ,\it acc\rm ) \rm = \it psw\_new\rm (\rm 0\rm )\rm ,\it psw\_new\rm (\rm 1\rm )\rm ,\it psw\_new\rm (\rm 3\rm )\rm ,\it psw\_new\rm (\rm 4\rm )\rm ,\it psw\_new\rm (\rm 5\rm )\rm )  $\land$

\hspace*{0.10in}\it accparity $\in$  \it WORD $\times$ \it WORD  $\fun$  \it WORD  $\land$

\hspace*{0.10in} $\forall$ \rm (\it psw\rm ,\it acc\rm ,\it psw\_new\rm )\rm .\rm (\it psw $\in$  \it WORD  $\land$  \it psw\_new $\in$  \it BV8  $\land$  \it acc $\in$  \it WORD  $\land$  \it psw\_new\rm (\rm 0\rm )\rm ,\it psw\_new\rm (\rm 1\rm )\rm ,\it psw\_new\rm (\rm 3\rm )\rm ,\it psw\_new\rm (\rm 4\rm )\rm ,\it psw\_new\rm (\rm 5\rm ) \rm = \it pswcalc\rm (\it psw\rm ,\it acc\rm )  $\land$  \it psw\_new\rm (\rm 6\rm ) \rm = \it bitget\rm (\it psw\rm ,\rm 6\rm )  $\land$  \it psw\_new\rm (\rm 7\rm ) \rm = \it bitget\rm (\it psw\rm ,\rm 7\rm )  $\implies$  \it accparity\rm (\it acc\rm ,\it psw\rm ) \rm = \it BV\_TO\_WORD\rm (\it psw\_new\rm )\rm )  $\land$

\hspace*{0.10in}\it add $\in$  \it WORD $\times$ \it WORD $\times$ \it WORD  $\fun$  \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in} $\forall$ \rm (\it ac\rm ,\it w2\rm ,\it psw\rm ,\it sun\rm ,\it psw\_new\rm )\rm .\rm (\it ac $\in$  \it WORD  $\land$  \it sun $\in$  \it WORD  $\land$  \it w2 $\in$  \it WORD  $\land$  \it psw $\in$  \it WORD  $\land$  \it psw\_new $\in$  \it BV8  $\land$  \it psw\_new\rm (\rm 6\rm ) \rm = \it BV8\_ADD\_AUX\_CARRY\rm (\it WORD\_TO\_BV\rm (\it ac\rm )\rm ,\it WORD\_TO\_BV\rm (\it w2\rm )\rm )  $\land$  \it sun\rm ,\it psw\_new\rm (\rm 7\rm ) \rm = \it soma\rm (\it ac\rm ,\it w2\rm )  $\land$  \it psw\_new\rm (\rm 0\rm )\rm ,\it psw\_new\rm (\rm 1\rm )\rm ,\it psw\_new\rm (\rm 3\rm )\rm ,\it psw\_new\rm (\rm 4\rm )\rm ,\it psw\_new\rm (\rm 5\rm ) \rm = \it pswcalc\rm (\it psw\rm ,\it sun\rm )  $\implies$  \it add\rm (\it ac\rm ,\it w2\rm ,\it psw\rm ) \rm = \it sun\rm ,\it BV\_TO\_WORD\rm (\it psw\_new\rm )\rm )  $\land$

\hspace*{0.10in}\it diff $\in$  \it WORD $\times$ \it WORD $\times$ \it BIT  $\fun$  \it WORD $\times$ \it BIT  $\land$

\hspace*{0.10in} $\forall$ \rm (\it ac\rm ,\it wd\rm ,\it ca\rm ,\it df\rm )\rm .\rm (\it ac $\in$  \it WORD  $\land$  \it wd $\in$  \it WORD  $\land$  \it ca $\in$  \it BIT  $\land$  \it df $\in$   $\int$   $\land$  \it df \rm = \it ac\rm -\rm (\it wd\rm +\it ca\rm )  $\implies$  \rm (\it df$<$\rm 0  $\implies$  \it diff\rm (\it ac\rm ,\it wd\rm ,\it ca\rm ) \rm = \it df\rm +\rm 2\rm 5\rm 5\rm ,\rm 1\rm )  $\land$  \rm (\it df $\geq$ \rm 0  $\implies$  \it diff\rm (\it ac\rm ,\it wd\rm ,\it ca\rm ) \rm = \it df\rm ,\rm 1\rm )\rm )  $\land$

\hspace*{0.10in}\it substract $\in$  \it WORD $\times$ \it WORD $\times$ \it WORD  $\fun$  \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in} $\forall$ \rm (\it ac\rm ,\it w2\rm ,\it psw\rm ,\it psw\_new\rm ,\it df\rm )\rm .\rm (\it ac $\in$  \it WORD  $\land$  \it w2 $\in$  \it WORD  $\land$  \it psw $\in$  \it WORD  $\land$  \it psw\_new $\in$  \it BV8  $\land$  \it df $\in$  \it WORD  $\land$  \it psw\_new\rm (\rm 2\rm ) \rm = \it OVERFLOW\_SUB\rm (\it WORD\_TO\_COMP2\rm (\it ac\rm )\rm ,\it WORD\_TO\_COMP2\rm (\it w2\rm )\rm ,\it bitget\rm (\it psw\rm ,\rm 7\rm )\rm )  $\land$  \it psw\_new\rm (\rm 6\rm ) \rm = \it BV8\_SUB\_AUX\_CARRY\rm (\it WORD\_TO\_BV\rm (\it ac\rm )\rm ,\it WORD\_TO\_BV\rm (\it w2\rm )\rm ,\it bitget\rm (\it psw\rm ,\rm 7\rm )\rm )  $\land$  \it psw\_new\rm (\rm 7\rm )\rm ,\it df \rm = \it diff\rm (\it ac\rm ,\it w2\rm ,\it bitget\rm (\it psw\rm ,\rm 7\rm )\rm )  $\land$  \it psw\_new\rm (\rm 0\rm )\rm ,\it psw\_new\rm (\rm 1\rm )\rm ,\it psw\_new\rm (\rm 3\rm )\rm ,\it psw\_new\rm (\rm 4\rm )\rm ,\it psw\_new\rm (\rm 5\rm ) \rm = \it pswcalc\rm (\it psw\rm ,\it df\rm )  $\implies$  \it substract\rm (\it ac\rm ,\it w2\rm ,\it psw\rm ) \rm = \it df\rm ,\it BV\_TO\_WORD\rm (\it psw\_new\rm )\rm )  $\land$

\hspace*{0.10in}\it and $\in$  \it WORD $\times$ \it WORD  $\fun$  \it WORD  $\land$

\hspace*{0.10in} $\forall$ \rm (\it w1\rm ,\it w2\rm ,\it wd\rm )\rm .\rm (\it w1 $\in$  \it WORD  $\land$  \it w2 $\in$  \it WORD  $\land$  \it wd $\in$  \it WORD  $\land$  \it wd \rm = \it BV\_TO\_WORD\rm (\it BV8\_AND\rm (\it WORD\_TO\_BV\rm (\it w1\rm )\rm ,\it WORD\_TO\_BV\rm (\it w2\rm )\rm )\rm )  $\implies$  \it and\rm (\it w1\rm ,\it w2\rm ) \rm = \it wd\rm )  $\land$

\hspace*{0.10in}\it ior $\in$  \it WORD $\times$ \it WORD  $\fun$  \it WORD $\times$ \bf BOOL  $\land$

\hspace*{0.10in} $\forall$ \rm (\it w1\rm ,\it w2\rm ,\it wd\rm )\rm .\rm (\it w1 $\in$  \it WORD  $\land$  \it w2 $\in$  \it WORD  $\land$  \it wd $\in$  \it WORD  $\land$  \it wd \rm = \it BV\_TO\_WORD\rm (\it BV8\_IOR\rm (\it WORD\_TO\_BV\rm (\it w1\rm )\rm ,\it WORD\_TO\_BV\rm (\it w2\rm )\rm )\rm )  $\implies$  \it ior\rm (\it w1\rm ,\it w2\rm ) \rm = \it wd\rm ,\bf bool\rm (\it wd \rm = \rm 0\rm )\rm )  $\land$

\hspace*{0.10in}\it xor $\in$  \it WORD $\times$ \it WORD  $\fun$  \it WORD $\times$ \bf BOOL  $\land$

\hspace*{0.10in} $\forall$ \rm (\it w1\rm ,\it w2\rm ,\it wd\rm )\rm .\rm (\it w1 $\in$  \it WORD  $\land$  \it w2 $\in$  \it WORD  $\land$  \it wd $\in$  \it WORD  $\land$  \it wd \rm = \it BV\_TO\_WORD\rm (\it BV8\_XOR\rm (\it WORD\_TO\_BV\rm (\it w1\rm )\rm ,\it WORD\_TO\_BV\rm (\it w2\rm )\rm )\rm )  $\implies$  \it xor\rm (\it w1\rm ,\it w2\rm ) \rm = \it wd\rm ,\bf bool\rm (\it wd \rm = \rm 0\rm )\rm )  $\land$

\hspace*{0.10in}\it complement $\in$  \it WORD  $\fun$  \it WORD  $\land$

\hspace*{0.10in} $\forall$ \it wd\rm .\rm (\it wd $\in$  \it WORD  $\implies$  \it complement\rm (\it wd\rm ) \rm = \it BV\_TO\_WORD\rm (\it BV8\_COMPLEMENT\rm (\it WORD\_TO\_BV\rm (\it wd\rm )\rm )\rm )\rm )  $\land$

\hspace*{0.10in}\it swap $\in$  \it WORD  $\fun$  \it WORD  $\land$

\hspace*{0.10in} $\forall$ \rm (\it wd\rm ,\it bv\rm )\rm .\rm (\it wd $\in$  \it WORD  $\land$  \it bv $\in$  \it BV8  $\implies$  \rm (\it bv \rm = \it WORD\_TO\_BV\rm (\it wd\rm )  $\implies$  \it swap\rm (\it wd\rm ) \rm = \it BV\_TO\_WORD\rm (\rm \{\rm 0 $\mapsto$ \it bv\rm (\rm 4\rm )\rm ,\rm 1 $\mapsto$ \it bv\rm (\rm 5\rm )\rm ,\rm 2 $\mapsto$ \it bv\rm (\rm 6\rm )\rm ,\rm 3 $\mapsto$ \it bv\rm (\rm 7\rm )\rm ,\rm 4 $\mapsto$ \it bv\rm (\rm 0\rm )\rm ,\rm 5 $\mapsto$ \it bv\rm (\rm 1\rm )\rm ,\rm 6 $\mapsto$ \it bv\rm (\rm 2\rm )\rm ,\rm 7 $\mapsto$ \it bv\rm (\rm 3\rm )\rm \}\rm )\rm )\rm )  $\land$

\hspace*{0.10in}\it rotateleft $\in$  \it WORD  $\fun$  \it WORD $\times$ \bf BOOL  $\land$

\hspace*{0.10in} $\forall$ \rm (\it wd\rm ,\it bv\rm )\rm .\rm (\it wd $\in$  \it WORD  $\land$  \it bv $\in$  \it BV8  $\implies$  \rm (\it bv \rm = \it WORD\_TO\_BV\rm (\it wd\rm )  $\implies$  \it rotateleft\rm (\it wd\rm ) \rm = \it BV\_TO\_WORD\rm (\rm \{\rm 0 $\mapsto$ \it bv\rm (\rm 7\rm )\rm ,\rm 1 $\mapsto$ \it bv\rm (\rm 0\rm )\rm ,\rm 2 $\mapsto$ \it bv\rm (\rm 1\rm )\rm ,\rm 3 $\mapsto$ \it bv\rm (\rm 2\rm )\rm ,\rm 4 $\mapsto$ \it bv\rm (\rm 3\rm )\rm ,\rm 5 $\mapsto$ \it bv\rm (\rm 4\rm )\rm ,\rm 6 $\mapsto$ \it bv\rm (\rm 5\rm )\rm ,\rm 7 $\mapsto$ \it bv\rm (\rm 6\rm )\rm \}\rm )\rm ,\bf bool\rm (\it bv\rm (\rm 7\rm ) \rm = \rm 1\rm )\rm )\rm )  $\land$

\hspace*{0.10in}\it rotateright $\in$  \it WORD  $\fun$  \it WORD $\times$ \bf BOOL  $\land$

\hspace*{0.10in} $\forall$ \rm (\it wd\rm ,\it bv\rm )\rm .\rm (\it wd $\in$  \it WORD  $\land$  \it bv $\in$  \it BV8  $\implies$  \rm (\it bv \rm = \it WORD\_TO\_BV\rm (\it wd\rm )  $\implies$  \it rotateright\rm (\it wd\rm ) \rm = \it BV\_TO\_WORD\rm (\rm \{\rm 0 $\mapsto$ \it bv\rm (\rm 1\rm )\rm ,\rm 1 $\mapsto$ \it bv\rm (\rm 2\rm )\rm ,\rm 2 $\mapsto$ \it bv\rm (\rm 3\rm )\rm ,\rm 3 $\mapsto$ \it bv\rm (\rm 4\rm )\rm ,\rm 4 $\mapsto$ \it bv\rm (\rm 5\rm )\rm ,\rm 5 $\mapsto$ \it bv\rm (\rm 6\rm )\rm ,\rm 6 $\mapsto$ \it bv\rm (\rm 7\rm )\rm ,\rm 7 $\mapsto$ \it bv\rm (\rm 0\rm )\rm \}\rm )\rm ,\bf bool\rm (\it bv\rm (\rm 0\rm ) \rm = \rm 1\rm )\rm )\rm )

\hspace*{0.10in}

\hspace*{0.10in}

\bf ASSERTIONS

\hspace*{0.10in}\bf dom\rm (\it parity\rm ) \rm = \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it parity\rm )  $\subseteq$  \it BIT  $\land$

\hspace*{0.10in}\bf dom\rm (\it soma\rm ) \rm = \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it soma\rm )  $\subseteq$  \it WORD $\times$ \it BIT  $\land$

\hspace*{0.10in}\bf dom\rm (\it bitget\rm ) \rm = \it WORD $\times$ \it WORD\_POSITION  $\land$

\hspace*{0.10in}\bf ran\rm (\it bitget\rm )  $\subseteq$  \it BIT  $\land$

\hspace*{0.10in}\bf dom\rm (\it bitset\rm ) \rm = \it WORD $\times$ \it WORD\_POSITION  $\land$

\hspace*{0.10in}\bf ran\rm (\it bitset\rm )  $\subseteq$  \it WORD  $\land$

\hspace*{0.10in}\bf dom\rm (\it bitclear\rm ) \rm = \it WORD $\times$ \it WORD\_POSITION  $\land$

\hspace*{0.10in}\bf ran\rm (\it bitclear\rm )  $\subseteq$  \it WORD  $\land$

\hspace*{0.10in}\bf dom\rm (\it bitcomplement\rm ) \rm = \it WORD $\times$ \it WORD\_POSITION  $\land$

\hspace*{0.10in}\bf ran\rm (\it bitcomplement\rm )  $\subseteq$  \it WORD  $\land$

\hspace*{0.10in}\bf dom\rm (\it bitput\rm ) \rm = \it WORD $\times$ \it WORD\_POSITION $\times$ \it BIT  $\land$

\hspace*{0.10in}\bf ran\rm (\it bitput\rm )  $\subseteq$  \it WORD  $\land$

\hspace*{0.10in}\bf dom\rm (\it pswcalc\rm ) \rm = \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it pswcalc\rm )  $\subseteq$  \it BIT $\times$ \it BIT $\times$ \it BIT $\times$ \it BIT $\times$ \it BIT  $\land$

\hspace*{0.10in}\bf dom\rm (\it accparity\rm ) \rm = \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it accparity\rm )  $\subseteq$  \it WORD  $\land$

\hspace*{0.10in}\bf dom\rm (\it add\rm ) \rm = \it WORD $\times$ \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it add\rm )  $\subseteq$  \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in}\bf dom\rm (\it diff\rm ) \rm = \it WORD $\times$ \it WORD $\times$ \it BIT  $\land$

\hspace*{0.10in}\bf ran\rm (\it diff\rm )  $\subseteq$  \it WORD $\times$ \it BIT  $\land$

\hspace*{0.10in}\bf dom\rm (\it substract\rm ) \rm = \it WORD $\times$ \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it substract\rm )  $\subseteq$  \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in}\bf dom\rm (\it and\rm ) \rm = \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it and\rm )  $\subseteq$  \it WORD  $\land$

\hspace*{0.10in}\bf dom\rm (\it ior\rm ) \rm = \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it ior\rm )  $\subseteq$  \it WORD $\times$ \bf BOOL  $\land$

\hspace*{0.10in}\bf dom\rm (\it xor\rm ) \rm = \it WORD $\times$ \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it xor\rm )  $\subseteq$  \it WORD $\times$ \bf BOOL  $\land$

\hspace*{0.10in}\bf dom\rm (\it complement\rm ) \rm = \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it complement\rm )  $\subseteq$  \it WORD  $\land$

\hspace*{0.10in}\bf dom\rm (\it swap\rm ) \rm = \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it swap\rm )  $\subseteq$  \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it rotateleft\rm )  $\subseteq$  \it WORD $\times$ \bf BOOL  $\land$

\hspace*{0.10in}\bf dom\rm (\it rotateleft\rm ) \rm = \it WORD  $\land$

\hspace*{0.10in}\bf dom\rm (\it rotateright\rm ) \rm = \it WORD  $\land$

\hspace*{0.10in}\bf ran\rm (\it rotateright\rm )  $\subseteq$  \it WORD $\times$ \bf BOOL

\bf END

\newpage
\end{sloppypar}\end{document}

### 2008-11-12 trunk/developments/platforms/Old/8051H/bdp/A8051.n.tex
Arrumando o repositorio

\documentstyle[11pt,supertabular]{article}
\setlength{\topmargin}{13mm}
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\textheight}{225mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm}
\input{psfig}

\begin{document}
\begin{sloppypar}

% DEFINITION DES CARACTERES MATHEMATIQUES B
%------------------------------------------
\def\@setmcodes#1#2#3{{\count0=#1 \count1=#3
	\loop \global\mathcode\count0=\count1 \ifnum \count0<#2
	\advance\count0 by1 \advance\count1 by1 \repeat}}

\@setmcodes{`A}{`Z}{"7441}
\@setmcodes{`a}{`z}{"7461}

\mathcode`\;="8000 % Makes ; active in math mode
{\catcode`\;=\active \gdef;{\semicolon\;}}
\mathchardef\semicolon="003B
%    Nominal distance from top of paper to top of page
\topmargin 0 pt
\textheight 53\baselineskip

%   Left margin on odd-numbered pages
\oddsidemargin  0.15 in
%   Left margin on even-numbered pages
\evensidemargin 0.35 in
%   Width of marginal notes.
\marginparwidth 1 in
%   Note that \oddsidemargin = \evensidemargin
\oddsidemargin 0.25 in
\evensidemargin 0.25 in
\marginparwidth 0.75 in
\textwidth 5.875 in % Width of text line.

\setlength{\parindent}{0pt}
\setlength{\parskip}{0ex}

% DEFINITION DES FONTS
%---------------------
% The AMS extra symbol fonts are loaded.
% Note: sometimes called euxm10
\font\msx=msam10
% Note: sometimes called euym10
\font\msy=msbm10

\newfam\msxfam \textfont\msxfam=\msx
\newfam\msyfam \textfont\msyfam=\msy

\def\famletter#1{\ifcase #1 0\or 1\or 2\or 3\or 4\or 5\or 6\or 7\or
	8\or 9\or A\or B\or C\or D\or E\or F\fi}

\edef\fx{\famletter\msxfam}
\edef\fy{\famletter\msyfam}

\def\bbold{\fam\msyfam \msy}

% SYMBOLES B
%-----------
% makes a quoted expression in mathematical text
\def\token#1{\hbox{`$#1$'}}
% used for error messages in Z specs
\def\report#1{\hbox{`{\tt #1}'}}

% \@myop makes an operator, with a strut to defeat TeX's vertical adjustment.
\def\@myop#1{\mathop{\mathstrut{#1}}\nolimits}

% This underscore doesn't have the little kern --- you get an italic
% correction anyway in math mode.
\def\_{\leavevmode \vbox{\hrule width0.5em}}

% Save \q as \xq for quantifiers q.
\let\xforall=\forall
\let\xexists=\exists
\let\xlambda=\lambda
\let\xmu=\mu

% \p and \f make arrows with 1 and 2 crossings resp.
\def\p#1{\mathrel{\ooalign{\hfil$\mapstochar\mkern 5mu$\hfil\cr$#1$}}}
\def\f#1{\mathrel{\ooalign{\hfil
	$\mapstochar\mkern 3mu\mapstochar\mkern 5mu$\hfil\cr$#1$}}}

\let\mc=\mathchardef

\def	\pow		{\mbox{${\cal P}$}}
\def	\po1		{\mbox{${\cal P}_1$}}
\let	\cross		\times
\def	\lambda		{\@myop{\xlambda}}
\def	\lnot		{\neg\;}
\def	\land		{\mathrel{\wedge}}
\def	\lor		{\mathrel{\vee}}
\let	\implies	\Rightarrow
\let	\iff		\Leftrightarrow
\def	\forall		{\@myop{\xforall}}
\def	\exists		{\@myop{\xexists}}
\def	\semi		{\mathrel{\comp}}
\def	\ssemi		{\mathbin{\rm ;}}
\let	\ensembleVide	\emptyset
\let	\rel		\leftrightarrow
\def	\dom		{\@myop{\sf dom}}
\def	\ran		{\@myop{\sf ran}}
\def	\id		{\@myop{\sf id}}
\def	\comp		{\mathbin{\raise
			0.6ex\hbox{\oalign{\hfil$\scriptscriptstyle
			\rm o$\hfil\cr\hfil$\scriptscriptstyle\rm 9$\hfil}}}}
\def	\para		{\mbox{$\mid\mid$}}
\mc	\dres		"2\fx43
\mc	\rres		"2\fx42
\def	\ndres		{\mathbin{{\dres} \llap{$-$}}}
\def	\nrres		{\mathbin{{\rres}\llap{$-$}}}
\def	\lover		{\mathbin{{\dres} \llap{$-\!\!\!\!-\!$}}}
\def	\rover		{\mathbin{{\rres}\llap{$\!-\!\!\!-$}}}
\let	\fun		\rightarrow
\def	\pfun		{\p\fun}
\def	\pinj		{\p\inj}
\mc	\inj		"3\fx1A
\def	\psurj		{\p\surj}
\mc	\surj		"3\fx10
\def	\bij		{\surj\!\!\!\!\!\!\!\inj}
\def	\nat		{\mbox{${\cal N}$}}
\def	\na1		{\mbox{${\cal N}_1$}}
\def	\num		{\mbox{${\cal Z}$}}
\def	\int		{\mbox{${\cal Z}$}}
\def	\rat		{\mbox{${\cal Q}$}}
\def	\div		{\mathbin{\rm /}}
\def	\mod		{\mathbin{\bf mod}}
\def	\upto		{\mathbin{\ldotp\ldotp}}
\def	\finset		{\mbox{${\cal F}$}}
\def	\finse1		{\mbox{${\cal F}_1$}}
\def	\ffun		{\f\fun}
\def	\finj		{\f\inj}
\def	\seq		{\@myop{\rm seq}}
\def	\cat		{\mathbin{\raise 0.8ex\hbox{$\mathchar"2\fx61$}}}
\def	\sep		{\hspace*{.05in}}

\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{0}

%-------------------%
% Debut du document %
%-------------------%

\bf MACHINE

\hspace*{0.10in}\it A8051

\bf SEES

\hspace*{0.10in}\it TYPES\rm ,\it ALU\rm ,\it MEMORY

\bf CONCRETE\_VARIABLES

\hspace*{0.10in}\it mem\rm ,\it pc

\bf INVARIANT

\hspace*{0.10in}\it mem $\in$  \it MEM\_ADDR  $\pfun$  \it WORD  $\land$

\hspace*{0.10in}\it pc $\in$  \it INSTRUCTION  $\land$

\hspace*{0.10in}\it bitget\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 0\rm ) \rm = \it parity\rm (\it mem\rm (\it ACC\rm )\rm )

\bf CONSTANTS

\hspace*{0.10in}\it bitjump

\bf PROPERTIES

\hspace*{0.10in}\it bitjump $\in$  \it BIT $\times$ \it BIT $\times$ \it WORD $\times$ \it INSTRUCTION  $\fun$  \it INSTRUCTION  $\land$

\hspace*{0.10in} $\forall$ \rm (\it bit\rm ,\it test\rm ,\it jump\rm ,\it pc\rm )\rm .\rm (\it bit $\in$  \it BIT  $\land$  \it test $\in$  \it BIT  $\land$  \it pc $\in$  \it INSTRUCTION  $\land$  \it jump $\in$  \it INSTRUCTION  $\implies$  \rm (\it bit \rm = \it test  $\implies$  \it bitjump\rm (\it bit\rm ,\it test\rm ,\it jump\rm ,\it pc\rm ) \rm = \it INSTRUCTION\_IJUMP\rm (\it pc\rm ,\it jump\rm )  $\land$  \it bit $\not =$ \it test  $\implies$  \it bitjump\rm (\it bit\rm ,\it test\rm ,\it jump\rm ,\it pc\rm ) \rm = \it INSTRUCTION\_NEXT\rm (\it pc\rm )\rm )\rm )

\bf ASSERTIONS

\hspace*{0.10in}\bf dom\rm (\it bitjump\rm ) \rm = \it BIT $\times$ \it BIT $\times$ \it WORD $\times$ \it INSTRUCTION  $\land$

\hspace*{0.10in}\bf ran\rm (\it bitjump\rm ) \rm = \it INSTRUCTION

\bf INITIALISATION

\hspace*{0.10in}\it mem $:\in$  \it MEM\_ADDR  $\surj$  \it WORD  $\para$

\hspace*{0.10in}\it pc\rm :=\rm 0

\bf OPERATIONS

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}\bf INIT \rm = \bf BEGIN

\hspace*{0.20in}\it pc\rm :=\rm 0

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}

\hspace*{0.10in}\bf CJNE\rm (\it addr\rm ,\it data\rm ,\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it data $\in$  \it WORD  $\land$

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf IF  $\neg$ \rm (\it mem\rm (\it addr\rm ) \rm = \it data\rm ) \bf THEN

\hspace*{0.30in}\bf BEGIN

\hspace*{0.40in}\bf IF \it mem\rm (\it addr\rm )$<$\it data \bf THEN

\hspace*{0.50in}\bf mem\rm (\it PSW\rm )\rm :=\it bitset\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )

\hspace*{0.40in}\bf ELSE

\hspace*{0.50in}\bf mem\rm (\it PSW\rm )\rm :=\it bitclear\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )

\hspace*{0.40in}\bf END

\hspace*{0.30in}\bf END  $\para$

\hspace*{0.30in}\it pc\rm :=\it INSTRUCTION\_IJUMP\rm (\it pc\rm ,\it jump\rm )

\hspace*{0.20in}\bf ELSE

\hspace*{0.30in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}

\hspace*{0.10in}\bf CJNEI\rm (\it Rn\rm ,\it data\rm ,\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it Rn $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.20in}\rm (\it Rn \rm = \it R1  $\lor$  \it Rn \rm = \it R2\rm )  $\land$

\hspace*{0.20in}\it data $\in$  \it WORD  $\land$

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION  $\land$

\hspace*{0.20in}\it mem\rm (\it Rn\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it addr \rm = \it mem\rm (\it Rn\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF  $\neg$ \rm (\it mem\rm (\it addr\rm ) \rm = \it data\rm ) \bf THEN

\hspace*{0.40in}\bf BEGIN

\hspace*{0.50in}\bf IF \it mem\rm (\it addr\rm )$<$\it data \bf THEN

\hspace*{0.60in}\bf mem\rm (\it PSW\rm )\rm :=\it bitset\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )

\hspace*{0.50in}\bf ELSE

\hspace*{0.60in}\bf mem\rm (\it PSW\rm )\rm :=\it bitclear\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )

\hspace*{0.50in}\bf END

\hspace*{0.40in}\bf END  $\para$

\hspace*{0.40in}\it pc\rm :=\it INSTRUCTION\_IJUMP\rm (\it pc\rm ,\it jump\rm )

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf CJNEM\rm (\it dest\rm ,\it src\rm ,\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it dest $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it src $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf IF  $\neg$ \rm (\it mem\rm (\it dest\rm ) \rm = \it mem\rm (\it src\rm )\rm ) \bf THEN

\hspace*{0.30in}\bf BEGIN

\hspace*{0.40in}\bf IF \it mem\rm (\it dest\rm )$<$\it mem\rm (\it src\rm ) \bf THEN

\hspace*{0.50in}\bf mem\rm (\it PSW\rm )\rm :=\it bitset\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )

\hspace*{0.40in}\bf ELSE

\hspace*{0.50in}\bf mem\rm (\it PSW\rm )\rm :=\it bitclear\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )

\hspace*{0.40in}\bf END

\hspace*{0.30in}\bf END  $\para$

\hspace*{0.30in}\it pc\rm :=\it INSTRUCTION\_IJUMP\rm (\it pc\rm ,\it jump\rm )

\hspace*{0.20in}\bf ELSE

\hspace*{0.30in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf SJMP\rm (\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_IJUMP\rm (\it pc\rm ,\it jump\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf AJMP\rm (\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\it pc\rm :=\it jump

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}\bf MOV\rm (\it dest\rm ,\it data\rm ) \rm = \bf PRE

\hspace*{0.20in}\it dest $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it data $\in$  \it WORD

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf IF \it dest \rm = \it ACC \bf THEN

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it dest $\mapsto$ \it data\rm ,\it PSW $\mapsto$ \it accparity\rm (\it data\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.20in}\bf ELSE

\hspace*{0.30in}\bf mem\rm (\it dest\rm )\rm :=\it data

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf MOVI\rm (\it Rn\rm ,\it data\rm ) \rm = \bf PRE

\hspace*{0.20in}\it Rn $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.20in}\rm (\it Rn \rm = \it R1  $\lor$  \it Rn \rm = \it R2\rm )  $\land$

\hspace*{0.20in}\it data $\in$  \it WORD  $\land$

\hspace*{0.20in}\it mem\rm (\it Rn\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it dest \bf WHERE

\hspace*{0.30in}\it dest $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it dest \rm = \it mem\rm (\it Rn\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf mem\rm (\it dest\rm )\rm :=\it data  $\para$

\hspace*{0.30in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf MOVM\rm (\it dest\rm ,\it src\rm ) \rm = \bf PRE

\hspace*{0.20in}\it src $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it dest $\in$  \it MEM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf IF \it dest \rm = \it ACC \bf THEN

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it mem\rm (\it src\rm )\rm ,\it PSW $\mapsto$ \it accparity\rm (\it mem\rm (\it src\rm )\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.20in}\bf ELSE

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it dest $\mapsto$ \it mem\rm (\it src\rm )\rm \}

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf MOVMI\rm (\it dest\rm ,\it Rn\rm ) \rm = \bf PRE

\hspace*{0.20in}\it dest $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it Rn $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\rm (\it Rn \rm = \it R1  $\lor$  \it Rn \rm = \it R0\rm )  $\land$

\hspace*{0.20in}\it mem\rm (\it Rn\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it src \bf WHERE

\hspace*{0.30in}\it src $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it src \rm = \it mem\rm (\it Rn\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it dest \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it mem\rm (\it src\rm )\rm ,\it PSW $\mapsto$ \it accparity\rm (\it mem\rm (\it src\rm )\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it dest $\mapsto$ \it mem\rm (\it src\rm )\rm \}

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf MOVIM\rm (\it Rn\rm ,\it src\rm ) \rm = \bf PRE

\hspace*{0.20in}\it Rn $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.20in}\it src $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\rm (\it Rn \rm = \it R1  $\lor$  \it Rn \rm = \it R2\rm )  $\land$

\hspace*{0.20in}\it mem\rm (\it Rn\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it dest \bf WHERE

\hspace*{0.30in}\it dest $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it dest \rm = \it mem\rm (\it Rn\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it dest \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it mem\rm (\it src\rm )\rm ,\it PSW $\mapsto$ \it accparity\rm (\it mem\rm (\it src\rm )\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it dest $\mapsto$ \it mem\rm (\it src\rm )\rm \}

\hspace*{0.30in}\bf END  $\para$

\hspace*{0.30in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf MOVB\rm (\it dest\rm ,\it src\rm ) \rm = \bf PRE

\hspace*{0.20in}\it dest $\in$  \it BIT\_ADDRESS  $\land$

\hspace*{0.20in}\it src $\in$  \it BIT\_ADDRESS  $\land$

\hspace*{0.20in}\rm (\it dest \rm = \it CY  $\lor$  \it src \rm = \it CY\rm )

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr1\rm ,\it addr2\rm ,\it p1\rm ,\it p2\rm ,\it bit\rm ,\it data \bf WHERE

\hspace*{0.30in}\it addr1 $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it addr2 $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it p1 $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it data $\in$  \it WORD  $\land$

\hspace*{0.30in}\it p2 $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it addr1\rm ,\it p1 \rm = \it dest  $\land$

\hspace*{0.30in}\it addr2\rm ,\it p2 \rm = \it src  $\land$

\hspace*{0.30in}\it bit \rm = \it bitget\rm (\it mem\rm (\it addr2\rm )\rm ,\it p2\rm )  $\land$

\hspace*{0.30in}\it data $\in$  \it WORD  $\land$

\hspace*{0.30in}\it data \rm = \it bitput\rm (\it mem\rm (\it addr1\rm )\rm ,\it p1\rm ,\it bit\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it addr1 \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it data\rm ,\it PSW $\mapsto$ \it accparity\rm (\it data\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it addr1\rm )\rm :=\it data

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}\bf ANL\rm (\it addr\rm ,\it data\rm ) \rm = \bf PRE

\hspace*{0.20in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it data $\in$  \it WORD

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result \rm = \it and\rm (\it mem\rm (\it addr\rm )\rm ,\it data\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it addr \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it accparity\rm (\it result\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it addr\rm )\rm :=\it result

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf ANLI\rm (\it Rn\rm ,\it data\rm ) \rm = \bf PRE

\hspace*{0.20in}\it Rn $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.20in}\rm (\it Rn \rm = \it R1  $\lor$  \it Rn \rm = \it R2\rm )  $\land$

\hspace*{0.20in}\it mem\rm (\it Rn\rm ) $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.20in}\it data $\in$  \it WORD

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it addr \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it addr $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it addr \rm = \it mem\rm (\it Rn\rm )  $\land$

\hspace*{0.30in}\it result \rm = \it and\rm (\it mem\rm (\it addr\rm )\rm ,\it data\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf mem\rm (\it addr\rm )\rm :=\it result

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf ANLM\rm (\it dest\rm ,\it src\rm ) \rm = \bf PRE

\hspace*{0.20in}\it dest $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it src $\in$  \it MEM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result \rm = \it and\rm (\it mem\rm (\it dest\rm )\rm ,\it mem\rm (\it src\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it dest \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it accparity\rm (\it result\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it dest\rm )\rm :=\it result

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf ANLMI\rm (\it dest\rm ,\it Rn\rm ) \rm = \bf PRE

\hspace*{0.20in}\it dest $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it Rn $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.20in}\rm (\it Rn \rm = \it R1  $\lor$  \it Rn \rm = \it R2\rm )  $\land$

\hspace*{0.20in}\it mem\rm (\it Rn\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it src \bf WHERE

\hspace*{0.30in}\it src $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it src \rm = \it mem\rm (\it Rn\rm )  $\land$

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result \rm = \it and\rm (\it mem\rm (\it dest\rm )\rm ,\it mem\rm (\it src\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it dest \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it accparity\rm (\it result\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it dest\rm )\rm :=\it result

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf ANLB\rm (\it ca\rm ,\it bit\rm ) \rm = \bf PRE

\hspace*{0.20in}\it bit $\in$  \it BIT\_ADDRESS  $\land$

\hspace*{0.20in}\it ca $\in$  \it BIT\_ADDRESS  $\land$

\hspace*{0.20in}\it ca \rm = \it CC

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it ps\rm ,\it bt \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it ps $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it addr\rm ,\it ps \rm = \it bit  $\land$

\hspace*{0.30in}\it bt $\in$  \it BIT  $\land$

\hspace*{0.30in}\it bt \rm = \it bitget\rm (\it mem\rm (\it addr\rm )\rm ,\it ps\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it bt \rm = \rm 0 \bf THEN

\hspace*{0.40in}\bf mem\rm (\it PSW\rm )\rm :=\it bitset\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it PSW\rm )\rm :=\it bitclear\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf ANLBC\rm (\it ca\rm ,\it bit\rm ) \rm = \bf PRE

\hspace*{0.20in}\it bit $\in$  \it BIT\_ADDRESS  $\land$

\hspace*{0.20in}\it ca $\in$  \it BIT\_ADDRESS  $\land$

\hspace*{0.20in}\it ca \rm = \it CY

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it ps\rm ,\it bt \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it ps $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it addr\rm ,\it ps \rm = \it bit  $\land$

\hspace*{0.30in}\it bt $\in$  \it BIT  $\land$

\hspace*{0.30in}\it bt \rm = \it bitget\rm (\it mem\rm (\it addr\rm )\rm ,\it ps\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it bt \rm = \rm 1 \bf THEN

\hspace*{0.40in}\bf mem\rm (\it PSW\rm )\rm :=\it bitset\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it PSW\rm )\rm :=\it bitclear\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}\bf ACALL\rm (\it inst\rm ) \rm = \bf PRE

\hspace*{0.20in}\it inst $\in$  \it INSTRUCTION  $\land$

\hspace*{0.20in}\it mem\rm (\it SP\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it sp\_new \bf WHERE

\hspace*{0.30in}\it sp\_new $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it sp\_new \rm = \it SP\_INC\rm (\it mem\rm (\it SP\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it mem\rm (\it sp\_new\rm ) $\mapsto$ \it INSTRUCTION\_NEXT\rm (\it pc\rm )\rm ,\it SP $\mapsto$ \it sp\_new\rm \}  $\para$

\hspace*{0.30in}\it pc\rm :=\it inst

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf RET \rm = \bf PRE

\hspace*{0.20in}\it mem\rm (\it SP\rm ) $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.20in}\it mem\rm (\it mem\rm (\it SP\rm )\rm ) $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it addr \rm = \it mem\rm (\it SP\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it pc\rm :=\it mem\rm (\it addr\rm )  $\para$

\hspace*{0.30in}\bf mem\rm (\it SP\rm )\rm :=\it SP\_DEC\rm (\it mem\rm (\it SP\rm )\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf POP\rm (\it direct\rm ) \rm = \bf PRE

\hspace*{0.20in}\it direct $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it mem\rm (\it SP\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it data \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it addr \rm = \it mem\rm (\it SP\rm )  $\land$

\hspace*{0.30in}\it data $\in$  \it WORD  $\land$

\hspace*{0.30in}\it data \rm = \it mem\rm (\it addr\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it direct \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\rm \{\it ACC $\mapsto$ \it data\rm ,\it SP $\mapsto$ \it SP\_DEC\rm (\it addr\rm )\rm ,\it PSW $\mapsto$ \it accparity\rm (\it data\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\it mem\rm :=\rm \{\it direct $\mapsto$ \it mem\rm (\it addr\rm )\rm ,\it SP $\mapsto$ \it SP\_DEC\rm (\it mem\rm (\it SP\rm )\rm )\rm \}

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf PUSH\rm (\it direct\rm ) \rm = \bf PRE

\hspace*{0.20in}\it direct $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.20in}\it mem\rm (\it SP\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it addr \rm = \it mem\rm (\it SP\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf mem\rm (\it addr\rm )\rm :=\it mem\rm (\it direct\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}

\hspace*{0.10in}\bf JB\rm (\it bit\rm ,\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it bit $\in$  \it BIT\_ADDRESS  $\land$

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it ps \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it ps $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it addr\rm ,\it ps \rm = \it bit

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it pc\rm :=\it bitjump\rm (\it bitget\rm (\it mem\rm (\it addr\rm )\rm ,\it ps\rm )\rm ,\rm 1\rm ,\it pc\rm ,\it jump\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf JBC\rm (\it bit\rm ,\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it bit $\in$  \it BIT\_ADDRESS  $\land$

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it ps\rm ,\it bt\rm ,\it wd \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it ps $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it bt $\in$  \it BIT  $\land$

\hspace*{0.30in}\it wd $\in$  \it WORD  $\land$

\hspace*{0.30in}\it addr\rm ,\it ps \rm = \it bit  $\land$

\hspace*{0.30in}\it bt \rm = \it bitget\rm (\it mem\rm (\it addr\rm )\rm ,\it ps\rm )  $\land$

\hspace*{0.30in}\it wd \rm = \it bitclear\rm (\it mem\rm (\it addr\rm )\rm ,\it ps\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it addr \rm = \it ACC \bf THEN

\hspace*{0.40in}\it pc\rm :=\it bitjump\rm (\it bt\rm ,\rm 1\rm ,\it pc\rm ,\it jump\rm )  $\para$

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it wd\rm ,\it PSW $\mapsto$ \it accparity\rm (\it wd\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\it pc\rm :=\it bitjump\rm (\it bt\rm ,\rm 1\rm ,\it pc\rm ,\it jump\rm )  $\para$

\hspace*{0.40in}\bf mem\rm (\it addr\rm )\rm :=\it wd

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf JC\rm (\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\it pc\rm :=\it bitjump\rm (\it bitget\rm (\it mem\rm (\it PSW\rm )\rm ,\rm 7\rm )\rm ,\rm 1\rm ,\it jump\rm ,\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf JNB\rm (\it bit\rm ,\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it bit $\in$  \it BIT\_ADDRESS  $\land$

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it ps \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it ps $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it addr\rm ,\it ps \rm = \it bit

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it pc\rm :=\it bitjump\rm (\it bitget\rm (\it mem\rm (\it addr\rm )\rm ,\it ps\rm )\rm ,\rm 0\rm ,\it jump\rm ,\it pc\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf JNC\rm (\it bit\rm ,\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it bit $\in$  \it BIT\_ADDRESS  $\land$

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it ps\rm ,\it bt \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it ps $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it bt $\in$  \it BIT  $\land$

\hspace*{0.30in}\it addr\rm ,\it ps \rm = \it bit  $\land$

\hspace*{0.30in}\it bt \rm = \it bitget\rm (\it mem\rm (\it addr\rm )\rm ,\it ps\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it pc\rm :=\it bitjump\rm (\it bt\rm ,\rm 0\rm ,\it jump\rm ,\it pc\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf JNZ\rm (\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf IF \it mem\rm (\it ACC\rm ) $\not =$ \rm 0 \bf THEN

\hspace*{0.30in}\it pc\rm :=\it INSTRUCTION\_IJUMP\rm (\it pc\rm ,\it jump\rm )

\hspace*{0.20in}\bf ELSE

\hspace*{0.30in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf JZ\rm (\it jump\rm ) \rm = \bf PRE

\hspace*{0.20in}\it jump $\in$  \it INSTRUCTION

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf IF \it mem\rm (\it ACC\rm ) \rm = \rm 0 \bf THEN

\hspace*{0.30in}\it pc\rm :=\it INSTRUCTION\_IJUMP\rm (\it pc\rm ,\it jump\rm )

\hspace*{0.20in}\bf ELSE

\hspace*{0.30in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.20in}\bf END

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf CLR\rm (\it bt\rm ) \rm = \bf PRE

\hspace*{0.20in}\it bt $\in$  \it BIT\_ADDRESS

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it ps\rm ,\it data \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it data $\in$  \it WORD  $\land$

\hspace*{0.30in}\it ps $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it addr\rm ,\it ps \rm = \it bt  $\land$

\hspace*{0.30in}\it data \rm = \it bitset\rm (\it mem\rm (\it addr\rm )\rm ,\it ps\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it addr \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it data\rm ,\it PSW $\mapsto$ \it accparity\rm (\it data\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it addr\rm )\rm :=\it data

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf CLRA\rm (\it addr\rm ) \rm = \bf PRE

\hspace*{0.20in}\it addr $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\it addr \rm = \it ACC

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \rm 0\rm ,\it PSW $\mapsto$ \it accparity\rm (\rm 0\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf CPL\rm (\it ca\rm ) \rm = \bf PRE

\hspace*{0.20in}\it ca $\in$  \it BIT\_ADDRESS

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it ps\rm ,\it result \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it ps $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it addr\rm ,\it ps \rm = \it ca  $\land$

\hspace*{0.30in}\it result \rm = \it bitcomplement\rm (\it mem\rm (\it addr\rm )\rm ,\it ps\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it addr \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it accparity\rm (\it result\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it addr\rm )\rm :=\it result

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf CPLA\rm (\it addr\rm ) \rm = \bf PRE

\hspace*{0.20in}\it addr $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\it addr \rm = \it ACC

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result \rm = \it complement\rm (\it mem\rm (\it ACC\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it accparity\rm (\it result\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf SETB\rm (\it bit\rm ) \rm = \bf PRE

\hspace*{0.20in}\it bit $\in$  \it BIT\_ADDRESS

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it ps\rm ,\it data \bf WHERE

\hspace*{0.30in}\it addr $\in$  \it MEM\_ADDR  $\land$

\hspace*{0.30in}\it ps $\in$  \it WORD\_POSITION  $\land$

\hspace*{0.30in}\it addr\rm ,\it ps \rm = \it bit  $\land$

\hspace*{0.30in}\it data \rm = \it bitclear\rm (\it mem\rm (\it addr\rm )\rm ,\it ps\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it addr \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it data\rm ,\it PSW $\mapsto$ \it accparity\rm (\it data\rm ,\it mem\rm (\it PSW\rm )\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it addr\rm )\rm :=\it data

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}

\hspace*{0.10in}\bf ADD\rm (\it ac\rm ,\it src\rm ) \rm = \bf PRE

\hspace*{0.20in}\it ac $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\rm (\it ac \rm = \it AA  $\lor$  \it ac \rm = \it ACC\rm )  $\land$

\hspace*{0.20in}\it src $\in$  \it MEM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it psw\_new \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it psw\_new $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result\rm ,\it psw\_new \rm = \it add\rm (\it mem\rm (\it ACC\rm )\rm ,\it mem\rm (\it src\rm )\rm ,\it mem\rm (\it PSW\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it psw\_new\rm \}

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf ADDI\rm (\it ac\rm ,\it reg\rm ) \rm = \bf PRE

\hspace*{0.20in}\it ac $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\it reg $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\rm (\it ac \rm = \it AA  $\lor$  \it ac \rm = \it ACC\rm )  $\land$

\hspace*{0.20in}\rm (\it reg \rm = \it R0  $\lor$  \it reg \rm = \it R1\rm )  $\land$

\hspace*{0.20in}\it mem\rm (\it reg\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it addr\rm ,\it result\rm ,\it psw\_new \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it psw\_new $\in$  \it WORD  $\land$

\hspace*{0.30in}\it addr \rm = \it mem\rm (\it reg\rm )  $\land$

\hspace*{0.30in}\it result\rm ,\it psw\_new \rm = \it add\rm (\it mem\rm (\it ACC\rm )\rm ,\it mem\rm (\it addr\rm )\rm ,\it mem\rm (\it PSW\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it psw\_new\rm \}

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf ADDD\rm (\it ac\rm ,\it data\rm ) \rm = \bf PRE

\hspace*{0.20in}\it ac $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\rm (\it ac \rm = \it AA  $\lor$  \it ac \rm = \it ACC\rm )  $\land$

\hspace*{0.20in}\it data $\in$  \it WORD

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it psw\_new \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it psw\_new $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result\rm ,\it psw\_new \rm = \it add\rm (\it mem\rm (\it ACC\rm )\rm ,\it data\rm ,\it mem\rm (\it PSW\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it PSW $\mapsto$ \it psw\_new\rm ,\it ACC $\mapsto$ \it result\rm \}

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf SUBB\rm (\it ac\rm ,\it src\rm ) \rm = \bf PRE

\hspace*{0.20in}\it ac $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\rm (\it ac \rm = \it AA  $\lor$  \it ac \rm = \it ACC\rm )  $\land$

\hspace*{0.20in}\it src $\in$  \it MEM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it psw\_new \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it psw\_new $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result\rm ,\it psw\_new \rm = \it substract\rm (\it mem\rm (\it ACC\rm )\rm ,\it mem\rm (\it src\rm )\rm ,\it mem\rm (\it PSW\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it psw\_new\rm \}

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf SUBBI\rm (\it ac\rm ,\it reg\rm ) \rm = \bf PRE

\hspace*{0.20in}\it ac $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\rm (\it ac \rm = \it AA  $\lor$  \it ac \rm = \it ACC\rm )  $\land$

\hspace*{0.20in}\it reg $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\rm (\it reg \rm = \it R0  $\lor$  \it reg \rm = \it R1\rm )  $\land$

\hspace*{0.20in}\it mem\rm (\it reg\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it psw\_new\rm ,\it src \bf WHERE

\hspace*{0.30in}\it src $\in$  \it WORD  $\land$

\hspace*{0.30in}\it src \rm = \it mem\rm (\it reg\rm )  $\land$

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it psw\_new $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result\rm ,\it psw\_new \rm = \it substract\rm (\it mem\rm (\it ACC\rm )\rm ,\it mem\rm (\it src\rm )\rm ,\it mem\rm (\it PSW\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it psw\_new\rm \}

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf SUBBD\rm (\it acc\rm ,\it data\rm ) \rm = \bf PRE

\hspace*{0.20in}\it acc $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\rm (\it acc \rm = \it AA  $\lor$  \it acc \rm = \it ACC\rm )  $\land$

\hspace*{0.20in}\it data $\in$  \it WORD

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it psw\_new \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it psw\_new $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result\rm ,\it psw\_new \rm = \it substract\rm (\it mem\rm (\it AA\rm )\rm ,\it data\rm ,\it mem\rm (\it PSW\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\it mem\rm :=\it mem $\lover$ \rm \{\it ACC $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it psw\_new\rm \}

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf INC\rm (\it src\rm ) \rm = \bf PRE

\hspace*{0.20in}\it src $\in$  \it MEM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it psw\_new \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it psw\_new $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result\rm ,\it psw\_new \rm = \it add\rm (\it mem\rm (\it src\rm )\rm ,\rm 1\rm ,\it mem\rm (\it PSW\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it src \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it src $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it accparity\rm (\it mem\rm (\it PSW\rm )\rm ,\it result\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it src\rm )\rm :=\it result

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf INCI\rm (\it Rn\rm ) \rm = \bf PRE

\hspace*{0.20in}\it Rn $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\rm (\it Rn \rm = \it R0  $\lor$  \it Rn \rm = \it R1\rm )  $\land$

\hspace*{0.20in}\it mem\rm (\it Rn\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it psw\_new\rm ,\it src \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it psw\_new $\in$  \it WORD  $\land$

\hspace*{0.30in}\it src $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it src \rm = \it mem\rm (\it Rn\rm )  $\land$

\hspace*{0.30in}\it result\rm ,\it psw\_new \rm = \it add\rm (\it mem\rm (\it src\rm )\rm ,\rm 1\rm ,\it mem\rm (\it PSW\rm )\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it src \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it src $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it accparity\rm (\it mem\rm (\it PSW\rm )\rm ,\it result\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it src\rm )\rm :=\it result

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf DEC\rm (\it src\rm ) \rm = \bf PRE

\hspace*{0.20in}\it src $\in$  \it MEM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it psw\_new \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it psw\_new $\in$  \it WORD  $\land$

\hspace*{0.30in}\it result\rm ,\it psw\_new \rm = \it substract\rm (\it mem\rm (\it src\rm )\rm ,\rm 1\rm ,\rm 0\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf IF \it src \rm = \it ACC \bf THEN

\hspace*{0.40in}\it mem\rm :=\it mem $\lover$ \rm \{\it src $\mapsto$ \it result\rm ,\it PSW $\mapsto$ \it accparity\rm (\it mem\rm (\it PSW\rm )\rm ,\it result\rm )\rm \}

\hspace*{0.30in}\bf ELSE

\hspace*{0.40in}\bf mem\rm (\it src\rm )\rm :=\it result

\hspace*{0.30in}\bf END

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END\rm ;

\hspace*{0.10in}\bf DECI\rm (\it Rn\rm ) \rm = \bf PRE

\hspace*{0.20in}\it Rn $\in$  \it SFR\_ADDR  $\land$

\hspace*{0.20in}\rm (\it Rn \rm = \it R0  $\lor$  \it Rn \rm = \it R1\rm )  $\land$

\hspace*{0.20in}\it mem\rm (\it Rn\rm ) $\in$  \it RAM\_ADDR

\hspace*{0.10in}\bf THEN

\hspace*{0.20in}\bf ANY \it result\rm ,\it psw\_new\rm ,\it src \bf WHERE

\hspace*{0.30in}\it result $\in$  \it WORD  $\land$

\hspace*{0.30in}\it psw\_new $\in$  \it WORD  $\land$

\hspace*{0.30in}\it src $\in$  \it RAM\_ADDR  $\land$

\hspace*{0.30in}\it src \rm = \it mem\rm (\it Rn\rm )  $\land$

\hspace*{0.30in}\it result\rm ,\it psw\_new \rm = \it substract\rm (\it mem\rm (\it src\rm )\rm ,\rm 1\rm ,\rm 0\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.30in}\bf mem\rm (\it src\rm )\rm :=\it result

\hspace*{0.20in}\bf END  $\para$

\hspace*{0.20in}\it pc\rm :=\it INSTRUCTION\_NEXT\rm (\it pc\rm )

\hspace*{0.10in}\bf END

\bf END

\newpage
\end{sloppypar}\end{document}

### 2008-11-12 trunk/developments/platforms/PIC16C432/PDB/DATA_MEMORY.tex
Arrumando o repositorio

\documentstyle[11pt,supertabular]{article}
\setlength{\topmargin}{13mm}
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\textheight}{225mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm}
\input{psfig}

\begin{document}
\begin{sloppypar}

% DEFINITION DES CARACTERES MATHEMATIQUES B
%------------------------------------------
\def\@setmcodes#1#2#3{{\count0=#1 \count1=#3
	\loop \global\mathcode\count0=\count1 \ifnum \count0<#2
	\advance\count0 by1 \advance\count1 by1 \repeat}}

\@setmcodes{`A}{`Z}{"7441}
\@setmcodes{`a}{`z}{"7461}

\mathcode`\;="8000 % Makes ; active in math mode
{\catcode`\;=\active \gdef;{\semicolon\;}}
\mathchardef\semicolon="003B
%    Nominal distance from top of paper to top of page
\topmargin 0 pt
\textheight 53\baselineskip

%   Left margin on odd-numbered pages
\oddsidemargin  0.15 in
%   Left margin on even-numbered pages
\evensidemargin 0.35 in
%   Width of marginal notes.
\marginparwidth 1 in
%   Note that \oddsidemargin = \evensidemargin
\oddsidemargin 0.25 in
\evensidemargin 0.25 in
\marginparwidth 0.75 in
\textwidth 5.875 in % Width of text line.

\setlength{\parindent}{0pt}
\setlength{\parskip}{0ex}

% DEFINITION DES FONTS
%---------------------
% The AMS extra symbol fonts are loaded.
% Note: sometimes called euxm10
\font\msx=msam10
% Note: sometimes called euym10
\font\msy=msbm10

\newfam\msxfam \textfont\msxfam=\msx
\newfam\msyfam \textfont\msyfam=\msy

\def\famletter#1{\ifcase #1 0\or 1\or 2\or 3\or 4\or 5\or 6\or 7\or
	8\or 9\or A\or B\or C\or D\or E\or F\fi}

\edef\fx{\famletter\msxfam}
\edef\fy{\famletter\msyfam}

\def\bbold{\fam\msyfam \msy}

% SYMBOLES B
%-----------
% makes a quoted expression in mathematical text
\def\token#1{\hbox{`$#1$'}}
% used for error messages in Z specs
\def\report#1{\hbox{`{\tt #1}'}}

% \@myop makes an operator, with a strut to defeat TeX's vertical adjustment.
\def\@myop#1{\mathop{\mathstrut{#1}}\nolimits}

% This underscore doesn't have the little kern --- you get an italic
% correction anyway in math mode.
\def\_{\leavevmode \vbox{\hrule width0.5em}}

% Save \q as \xq for quantifiers q.
\let\xforall=\forall
\let\xexists=\exists
\let\xlambda=\lambda
\let\xmu=\mu

% \p and \f make arrows with 1 and 2 crossings resp.
\def\p#1{\mathrel{\ooalign{\hfil$\mapstochar\mkern 5mu$\hfil\cr$#1$}}}
\def\f#1{\mathrel{\ooalign{\hfil
	$\mapstochar\mkern 3mu\mapstochar\mkern 5mu$\hfil\cr$#1$}}}

\let\mc=\mathchardef

\def	\pow		{\mbox{${\cal P}$}}
\def	\po1		{\mbox{${\cal P}_1$}}
\let	\cross		\times
\def	\lambda		{\@myop{\xlambda}}
\def	\lnot		{\neg\;}
\def	\land		{\mathrel{\wedge}}
\def	\lor		{\mathrel{\vee}}
\let	\implies	\Rightarrow
\let	\iff		\Leftrightarrow
\def	\forall		{\@myop{\xforall}}
\def	\exists		{\@myop{\xexists}}
\def	\semi		{\mathrel{\comp}}
\def	\ssemi		{\mathbin{\rm ;}}
\let	\ensembleVide	\emptyset
\let	\rel		\leftrightarrow
\def	\dom		{\@myop{\sf dom}}
\def	\ran		{\@myop{\sf ran}}
\def	\id		{\@myop{\sf id}}
\def	\comp		{\mathbin{\raise
			0.6ex\hbox{\oalign{\hfil$\scriptscriptstyle
			\rm o$\hfil\cr\hfil$\scriptscriptstyle\rm 9$\hfil}}}}
\def	\para		{\mbox{$\mid\mid$}}
\mc	\dres		"2\fx43
\mc	\rres		"2\fx42
\def	\ndres		{\mathbin{{\dres} \llap{$-$}}}
\def	\nrres		{\mathbin{{\rres}\llap{$-$}}}
\def	\lover		{\mathbin{{\dres} \llap{$-\!\!\!\!-\!$}}}
\def	\rover		{\mathbin{{\rres}\llap{$\!-\!\!\!-$}}}
\let	\fun		\rightarrow
\def	\pfun		{\p\fun}
\def	\pinj		{\p\inj}
\mc	\inj		"3\fx1A
\def	\psurj		{\p\surj}
\mc	\surj		"3\fx10
\def	\bij		{\surj\!\!\!\!\!\!\!\inj}
\def	\nat		{\mbox{${\cal N}$}}
\def	\na1		{\mbox{${\cal N}_1$}}
\def	\num		{\mbox{${\cal Z}$}}
\def	\int		{\mbox{${\cal Z}$}}
\def	\rat		{\mbox{${\cal Q}$}}
\def	\div		{\mathbin{\rm /}}
\def	\mod		{\mathbin{\bf mod}}
\def	\upto		{\mathbin{\ldotp\ldotp}}
\def	\finset		{\mbox{${\cal F}$}}
\def	\finse1		{\mbox{${\cal F}_1$}}
\def	\ffun		{\f\fun}
\def	\finj		{\f\inj}
\def	\seq		{\@myop{\rm seq}}
\def	\cat		{\mathbin{\raise 0.8ex\hbox{$\mathchar"2\fx61$}}}
\def	\sep		{\hspace*{.05in}}

\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{0}

%-------------------%
% Debut du document %
%-------------------%

\vspace*{4mm}
\bf MACHINE

\hspace*{0.20in}\it DATA\_MEMORY

\vspace*{4mm}
\bf SEES

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_ARITHMETICS\rm ,

\hspace*{0.20in}\it BYTE\_DEFINITION\rm ,

\vspace*{4mm}
\hspace*{0.20in}\it POWER2

\vspace*{4mm}
\bf CONSTANTS

\hspace*{0.20in}\it BANK\_WIDTH\rm , \it BANK\rm , \it BANK\_SIZE\rm ,

\hspace*{0.20in}\it LOCATION\_WIDTH\rm , \it LOCATION\rm ,

\hspace*{0.20in}\it ADDRESS\_WIDTH\rm , \it ADDRESS\rm ,

\hspace*{0.20in}\it FILE\rm ,

\hspace*{0.20in}\it make\_address\rm ,

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it INDF\_ADDR\rm , \it TMR0\_ADDR\rm , \it PCL\_ADDR\rm ,

\hspace*{0.20in}\it STATUS\_ADDR\rm , \it FSR\_ADDR\rm ,

\hspace*{0.20in}\it PORTA\_ADDR\rm , \it PORTB\_ADDR\rm ,

\hspace*{0.20in}\it PCLATH\_ADDR\rm , \it INTCON\_ADDR\rm , \it PIR1\_ADDR\rm , \it CMCON\_ADDR\rm ,

\hspace*{0.20in}\it TRISA\_ADDR\rm , \it TRISB\_ADDR\rm ,

\hspace*{0.20in}\it PIE1\_ADDR\rm , \it PCON\_ADDR\rm , \it LININTF\_ADDR\rm , \it VRCON\_ADDR\rm ,

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it UNIMPLEMENTED\_LOCATIONS\rm ,

\vspace*{4mm}
\hspace*{0.20in}\it MAPPED\_REGISTERS\rm ,

\hspace*{0.20in}\it register\_addresses\rm ,

\vspace*{4mm}
\hspace*{0.20in}\it IRP\_POS\rm , \it RP1\_POS\rm , \it RP0\_POS\rm ,

\vspace*{4mm}
\hspace*{0.20in}\it actual\_address

\vspace*{4mm}
\bf PROPERTIES

\hspace*{0.20in}\it BANK\_WIDTH \rm = \rm 2  $\land$

\hspace*{0.20in}\it BANK  $\subseteq$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it BANK \rm = \rm \{ \it vv  $\mid$  \it vv  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it vv\rm ) \rm = \it BANK\_WIDTH \rm \}  $\land$

\hspace*{0.20in}\it LOCATION\_WIDTH \rm = \rm 7  $\land$

\hspace*{0.20in}\it LOCATION  $\subseteq$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it LOCATION \rm = \rm \{ \it vv  $\mid$  \it vv  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it vv\rm ) \rm = \it LOCATION\_WIDTH \rm \}  $\land$

\hspace*{0.20in}\it BANK\_SIZE \rm = \rm 2$^{LOCATION\_WIDTH}$  $\land$

\hspace*{0.20in}\it ADDRESS\_WIDTH \rm = \it BANK\_WIDTH \rm + \it LOCATION\_WIDTH  $\land$

\hspace*{0.20in}\it ADDRESS \rm = \rm 0 $\upto$ \rm (\rm (\rm 2$^{ADDRESS\_WIDTH}$\rm )\rm -\rm 1\rm )  $\land$

\hspace*{0.20in}\it make\_address  $\in$  \it BANK  $\times$  \it LOCATION  $\fun$  \it ADDRESS  $\land$

\hspace*{0.20in}\it make\_address \rm =  $\lambda$ \rm (\it vb\rm , \it vl\rm )\rm . \rm (\it vb  $\in$  \it BANK  $\land$  \it vl  $\in$  \it LOCATION  $\mid$  \it bv\_to\_nat \rm (\it bv\_catenate\rm (\it vb\rm , \it vl\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it FILE \rm = \it ADDRESS  $\fun$  \it BYTE  $\land$

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it INDF\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it INDF\_ADDR \rm = \rm 0  $\land$

\hspace*{0.20in}\it TMR0\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it TMR0\_ADDR \rm = \rm 1  $\land$

\hspace*{0.20in}\it PCL\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PCL\_ADDR \rm = \rm 2  $\land$

\hspace*{0.20in}\it STATUS\_ADDR  $\in$  \it ADDRESS  $\land$  \it STATUS\_ADDR \rm = \rm 3  $\land$

\hspace*{0.20in}\it FSR\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it FSR\_ADDR \rm = \rm 4  $\land$

\hspace*{0.20in}\it PORTA\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PORTA\_ADDR \rm = \rm 5  $\land$

\hspace*{0.20in}\it PORTB\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PORTB\_ADDR \rm = \rm 6  $\land$

\hspace*{0.20in}\it PCLATH\_ADDR  $\in$  \it ADDRESS  $\land$  \it PCLATH\_ADDR \rm = \rm 1\rm 0  $\land$

\hspace*{0.20in}\it INTCON\_ADDR  $\in$  \it ADDRESS  $\land$  \it INTCON\_ADDR \rm = \rm 1\rm 1  $\land$

\hspace*{0.20in}\it PIR1\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PIR1\_ADDR \rm = \rm 1\rm 2  $\land$

\hspace*{0.20in}\it CMCON\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it CMCON\_ADDR \rm = \rm 1\rm 2\rm 7  $\land$

\hspace*{0.20in}\it TRISA\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it TRISA\_ADDR \rm = \rm 1\rm 3\rm 3  $\land$

\hspace*{0.20in}\it TRISB\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it TRISB\_ADDR \rm = \rm 1\rm 3\rm 4  $\land$

\hspace*{0.20in}\it PIE1\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PIE1\_ADDR \rm = \rm 1\rm 4\rm 0  $\land$

\hspace*{0.20in}\it PCON\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PCON\_ADDR \rm = \rm 1\rm 4\rm 2  $\land$

\hspace*{0.20in}\it LININTF\_ADDR  $\in$  \it ADDRESS  $\land$ \it LININTF\_ADDR \rm = \rm 1\rm 4\rm 4  $\land$

\hspace*{0.20in}\it VRCON\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it VRCON\_ADDR \rm = \rm 1\rm 5\rm 9  $\land$

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it UNIMPLEMENTED\_LOCATIONS \rm =

\hspace*{0.40in}\rm 7 $\upto$ \rm 9  $\cup$ \hspace*{0.20in}\rm 1\rm 3 $\upto$ \rm 3\rm 0  $\cup$  \rm 1\rm 3\rm 5 $\upto$ \rm 1\rm 3\rm 7  $\cup$  \rm \{\rm 1\rm 4\rm 1\rm \}  $\cup$  \rm \{\rm 1\rm 4\rm 3\rm \}  $\cup$  \rm 1\rm 4\rm 5 $\upto$ \rm 1\rm 5\rm 8  $\cup$  \rm 1\rm 9\rm 2 $\upto$ \rm 2\rm 4\rm 0  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it MAPPED\_REGISTERS  $\subseteq$  \it ADDRESS  $\land$

\hspace*{0.20in}\it MAPPED\_REGISTERS \rm = \rm \{ \it INDF\_ADDR\rm , \it STATUS\_ADDR\rm , \it FSR\_ADDR\rm , \it PCLATH\_ADDR\rm , \it INTCON\_ADDR \rm \}  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it register\_addresses  $\in$  \it ADDRESS  $\fun$   $ \po1 $  \rm (\it ADDRESS\rm )  $\land$

\hspace*{0.20in} $\forall$ \rm (\it ad\rm )\rm . \rm (\it ad  $\in$  \it ADDRESS  $\implies$

\hspace*{0.40in}\rm (\rm (\it ad  $\not\in$  \it MAPPED\_REGISTERS  $\implies$  \it register\_addresses\rm (\it ad\rm ) \rm = \rm \{\it ad\rm \}\rm )  $\land$

\hspace*{0.45in}\rm (\it ad  $\in$  \it MAPPED\_REGISTERS  $\implies$  \it register\_addresses\rm (\it ad\rm ) \rm = \rm \{\it ad\rm , \it ad \rm + \it BANK\_SIZE\rm \}\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it IRP\_POS  $\in$  \it BYTE\_INDEX  $\land$  \it IRP\_POS \rm = \rm 7  $\land$

\hspace*{0.20in}\it RP1\_POS  $\in$  \it BYTE\_INDEX  $\land$  \it RP1\_POS \rm = \rm 6  $\land$

\hspace*{0.20in}\it RP0\_POS  $\in$  \it BYTE\_INDEX  $\land$  \it RP0\_POS \rm = \rm 5  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it actual\_address  $\in$  \it ADDRESS  $\times$  \it FILE  $\fun$  \it ADDRESS  $\land$

\hspace*{0.20in}\it actual\_address \rm =

\hspace*{0.40in} $\lambda$  \rm (\it ad\rm , \it fi\rm ) \rm . \rm (\it ad  $\in$  \it ADDRESS  $\land$  \it fi  $\in$  \it FILE  $\land$  \it ad  $\in$  \it register\_addresses \rm (\it INDF\_ADDR\rm )  $\mid$

\hspace*{0.60in}\rm (\it fi\rm (\it STATUS\_ADDR\rm )\rm (\it RP0\_POS\rm ) $\times$ \rm 2$^{7}$\rm ) \rm + \rm (\it fi\rm (\it STATUS\_ADDR\rm )\rm (\it RP1\_POS\rm ) $\times$ \rm 2$^{8}$\rm ) \rm + \it bv\_to\_nat\rm (\it fi\rm (\it FSR\_ADDR\rm )\rm )\rm )  $\cup$

\hspace*{0.40in} $\lambda$  \rm (\it ad\rm , \it fi\rm ) \rm . \rm (\it ad  $\in$  \it ADDRESS  $\land$  \it fi  $\in$  \it FILE  $\land$  \it ad  $\not\in$  \it register\_addresses \rm (\it INDF\_ADDR\rm )  $\mid$

\hspace*{0.60in}\rm (\it fi\rm (\it STATUS\_ADDR\rm )\rm (\it IRP\_POS\rm ) $\times$ \rm 2$^{8}$\rm ) \rm + \it bv\_to\_nat \rm (\it fi\rm (\it FSR\_ADDR\rm )\rm )\rm )

\vspace*{4mm}
\bf VARIABLES

\hspace*{0.20in}\it file

\vspace*{4mm}
\bf INVARIANT

\hspace*{0.20in}\it file  $\in$  \it FILE  $\land$

\hspace*{0.20in} $\forall$ \rm (\it ad\rm )\rm .\rm (\it ad  $\in$  \it UNIMPLEMENTED\_LOCATIONS  $\implies$  \it file\rm (\it ad\rm ) \rm = \it BYTE\_ZERO\rm )  $\land$

\hspace*{0.20in} $\forall$ \rm (\it reg\rm )\rm .\rm (\it reg  $\in$  \it MAPPED\_REGISTERS  $\implies$  \it file\rm (\it reg\rm ) \rm = \it file\rm (\it reg\rm +\it BANK\_SIZE\rm )\rm )  $\land$

\hspace*{0.20in}\it file\rm (\it STATUS\_ADDR\rm )\rm (\it IRP\_POS\rm ) \rm = \rm 0  $\land$

\hspace*{0.20in}\it file\rm (\it STATUS\_ADDR\rm )\rm (\it RP1\_POS\rm ) \rm = \rm 0

\vspace*{4mm}
\bf ASSERTIONS

\hspace*{0.20in}\it BANK\_SIZE \rm = \rm 1\rm 2\rm 8\rm ;

\hspace*{0.20in}\it register\_addresses \rm (\it INDF\_ADDR\rm ) \rm = \rm \{\rm 0\rm , \rm 1\rm 2\rm 8\rm \}\rm ;

\hspace*{0.20in}\it register\_addresses \rm (\it STATUS\_ADDR\rm ) \rm = \rm \{\rm 3\rm , \rm 1\rm 3\rm 1\rm \}\rm ;

\hspace*{0.20in}\it register\_addresses \rm (\it FSR\_ADDR\rm ) \rm = \rm \{\rm 4\rm , \rm 1\rm 3\rm 2\rm \}\rm ;

\hspace*{0.20in}\it register\_addresses \rm (\it PCLATH\_ADDR\rm ) \rm = \rm \{\rm 1\rm 0\rm , \rm 1\rm 3\rm 8\rm \}\rm ;

\hspace*{0.20in}\it register\_addresses \rm (\it INTCON\_ADDR\rm ) \rm = \rm \{\rm 1\rm 1\rm , \rm 1\rm 3\rm 9\rm \}\rm ;

\vspace*{4mm}
\hspace*{0.20in}\bf dom\rm (\it file\rm ) \rm = \it ADDRESS\rm ;

\hspace*{0.20in}\bf ran\rm (\it file\rm )  $\subseteq$  \it BYTE

\vspace*{4mm}
\bf INITIALISATION

\hspace*{0.20in}\bf ANY \it random\rm , \it init\_status \bf WHERE

\hspace*{0.40in}\it random  $\in$  \it FILE  $\land$

\hspace*{0.40in}\it init\_status  $\in$  \it BYTE  $\land$

\hspace*{0.40in}\it init\_status \rm (\it RP1\_POS\rm ) \rm = \rm 0

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it file \rm := \it random  $\lover$  \rm \{ \it INDF\_ADDR  $\mapsto$  \it BYTE\_ZERO\rm ,

\hspace*{1.00in}\it STATUS\_ADDR  $\mapsto$  \it init\_status\rm ,

\hspace*{1.00in}\it INDF\_ADDR \rm + \rm 1\rm 2\rm 8  $\mapsto$  \it BYTE\_ZERO\rm ,

\hspace*{1.00in}\it STATUS\_ADDR \rm + \rm 1\rm 2\rm 8  $\mapsto$  \it init\_status \rm \}

\hspace*{0.80in} $\lover$  \rm (\it UNIMPLEMENTED\_LOCATIONS  $\times$  \rm \{ \it BYTE\_ZERO \rm \}\rm )

\hspace*{0.20in}\bf END

\bf END

\newpage
\end{sloppypar}\end{document}

### 2009-04-20 trunk/papers/FM2009/source/paper.tex

\documentclass{llncs}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

%\usepackage{float}

\include{comandos}
\include{SymbolsB_AtelierB}

%
\begin{document}

\title{Formal Verification of Real Models from Platforms Using The B-method }

\author{Val\'{e}rio Medeiros Jr\inst{1}, David D\'{e}harbe\inst{1} , Stephenson Galv\~{a}o \inst{1}}

\institute{Federal University of Rio Grande do Norte, Natal RN 59078-970, Brazil}

\maketitle

\begin{abstract}
This paper describes an approach to model the microcontroller platforms. More specifically, it shows details about
the Z80 model. The model is created using the B method; it applies math and logic concepts to describe the
characteristics from platforms. Therefore, this model can be used in platforms projects to document, build
simulators, verify properties about the model and verify the software in assembly level. The verification in
assembly is utility more important, because this allow the developing software faithful to the algorithm language.
Finally, this paper show some relevant techniques used in models and to describe several details.
\end{abstract}
%
\section{Introduction about hardware models in B}
[One page]
This paper show a model in B from a microcontrooller.
[EXPLAIN THE MOTIVATION]
 [When a error is found at project of plataform the PREJUIZOS are big ]
 [ In critical system, when a error apper in software it can make several PREJUIZOS  ]

[[Other microcontrollers are in building proccess: 8051, Arm..]]
[Some smal details of the model are omitted in order to simplify the explanation.]
[Smal intrudctio about B]
[Structure from paper]

%

\section{B Notation}

[One page]

\subsubsection{ Desing components B }
[half page]
[Explicar como as coias são modeladas]

\section{General definitions for hardware models}

We think that We find it convenient to provide a set of basic definitions to model hardware concepts. These
definitions are grouped in a library that we named $\mathit{HARDWARE\_LIBRARY}$ and arrange through four modules:
$\mathit{BIT\_DEFINITION}$, $\mathit{BIT\_VECTOR\_DEFINITION}$, $\mathit{BYTE\_DEFINITION}$ and
$\mathit{BIT\_VECTOR\_ARITHMETICS}$.

[[Explain better the figure]]

The dependency graph between these modules is depicted in
Figure~\ref{fig:hardware-definition-graph}.

\begin{figure}[h]
\centering
\includegraphics[width=.7\textwidth]{diagramaEstrutural.jpg}
\caption{Dependencies between modules of $\mathit{HARDWARE\_LIBRARY}$ and
platform model.}
\label{fig:hardware-definition-graph}

\end{figure}

Now, We will explain each of these modules and present them in detail.

\subsection{Definitions to represent and manipulate bits}

The entities defined in the module $\mathit{BIT\_DEFINITION}$ are the
type for bits, logical operations on bits (negation, conjunction,
disjunction, exclusive disjunction), as well as a conversion function
from booleans to bits.

First, bits are modeled as the set of integers $\{0, 1\}$:\\

$
\begin{array}{l}
\mathit{BIT} = \mathit{0..1} \\ 
\end{array}
$

The negation is a unary function on bits and it is defined as:

$
\begin{array}{l}
\mathit{bit\_not}  \in  \mathit{BIT}  \fun  \mathit{BIT}  \land \\
\forall ( \mathit{bb}). (\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bb}) =
1-\mathit{bb})\\
\end{array}
$

The module also provides lemmas on negation that may be useful for the
users of the library to develop proofs:

$
\begin{array}{l}
 \mathit{bit\_not}(0) = 1;  \mathit{bit\_not}(1) = 0; \\
\forall (\mathit{bb}).(\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bit\_not}(\mathit{bb})) = \mathit{bb});
\end{array}
$

Conjunction is a unary function on bits and it is defined as:

$
\begin{array}{l}
\mathit{bit\_and} \in \mathit{BIT} \times \mathit{BIT} \fun \mathit{BIT} \land \\
\forall (\mathit{b1}, \mathit{b2}).(\mathit{b1}  \in \mathit{BIT}  \land \mathit{b2} \in \mathit{BIT} \implies \\
\quad ((\mathit{bit\_and}(\mathit{b1}, \mathit{b2}) = 1) \iff (\mathit{b1} = 1)  \land  (\mathit{b2} = 1)))
\end{array}
$

The module provides the following lemmas for conjunction, either:

$
\begin{array}{l}
 \mathit{bit\_and}(0,0) = 0;  \mathit{bit\_and}(0,1) = 0; \\
 \mathit{bit\_and}(1,0) = 0;  \mathit{bit\_and}(1,1) = 1; \\
\forall (\mathit{b1},\mathit{b2}).(\mathit{b1} \in \mathit{BIT} \land \mathit{b2} \in \mathit{BIT} \implies \\
\quad (\mathit{bit\_and}(\mathit{b1}, \mathit{b2}) = \mathit{bit\_and}(\mathit{b2},\mathit{b1}))); \\
\forall (\mathit{b1},\mathit{b2},\mathit{b3}).(\mathit{b1} \in \mathit{BIT} \land  \mathit{b2} \in \mathit{BIT} \land \mathit{b3} \in \mathit{BIT} \implies \\
\quad (\mathit{bit\_and}(\mathit{b1}, \mathit{bit\_and}(\mathit{b2},\mathit{b3})) = \mathit{bit\_and}(\mathit{bit\_and}(\mathit{b1},\mathit{b2}),\mathit{b3}))); \\
\forall (\mathit{b1}).(\mathit{b1} \in \mathit{BIT} \implies (\mathit{bit\_and}(\mathit{b1}, 1) = \mathit{b1})); \\
\forall (\mathit{b1}).(\mathit{b1} \in \mathit{BIT} \implies (\mathit{bit\_and}(\mathit{b1}, 0) = 0));
\end{array}
$

The module provides definitions of $\mathit{bit\_or}$ (disjunction) and $\mathit{bit\_xor}$ (exclusive disjunction),
as well as lemmas on those operators. These are standard and their expression in B is similar as for
$\mathit{bit\_and}$, they are thus omitted.

Finally, the conversion from booleans to bits is simply defined as:

$
\begin{array}{l}
\mathit{bool\_to\_bit} \in \BOOL \fun \mathit{BIT} \land \mathit{bool\_to\_bit} = \{ \TRUE \mapsto 1, \FALSE \mapsto 0 \} \\
\end{array}
$

Observe that all the lemmas that are provided in this module have been
mechanically proved by the theorem prover included with our B
development environment. None of these proofs requires human insight.

\subsection{Representation and manipulation of bit vectors}

Sequences are pre-defined in B, as functions whose the domain is an integer range with lower bound 1 (one). Indices
in bit vectors usually range from 0 (zero) upwards and the model we propose obeys this convention by making an
one-position shift where necessary.  We thus define bit vectors as non-empty sequences of bits, and
$\mathit{BIT\_VECTOR}$ is the set of all such sequences:

$
\begin{array}{l}
\mathit{BIT\_VECTOR} = \seq (\mathit{BIT})
\end{array}
$

The function $\mathit{bv\_size}$ returns the size of a given bit vector. It is basically a wrapper for the
predefined function $\mathbf{size}$ that applies to sequences.

$
\begin{array}{l}
\mathit{bv\_size} \in \mathit{BIT\_VECTOR} \fun \nat_1 \land \\
\mathit{bv\_size} = \lambda bv \bullet (bv \in \mathit{BIT\_VECTOR} \mid \mathbf{size}(bv))
\end{array}
$

We also define two functions $\mathit{bv\_set}$ and $\mathit{bv\_clear}$ that, given a bit vector, and a position of
the bit vector, return the bit vector resulting from setting the corresponding position to 0 or to 1, and a function
$\mathit{bv\_get}$ that, given a bit vector, and a valid position, each one returns the value of the bit at that
position. Only the first definition is shown here:

$
\begin{array}{l}
\mathit{bv\_set} \in \mathit{BIT\_VECTOR} \times \nat \fun \mathit{BIT\_VECTOR} \land \\
\mathit{bv\_set} = \lambda v, n \bullet (v \in \mathit{BIT\_VECTOR} \land 0 \in \nat \land n < \mathit{bv\_size}(v) \mid v \ovr \{ n+1 \mapsto 1 \})
\end{array}
$

The function $bv\_catenate$ takes as parameters two bit vectors $v$ and $w$, and returns the result of the
concatenation of $v$ and $w$, such that $v$ constitutes the most significant part of the result.

% $
% \begin{array}{l}
% \mathit{bv\_catenate} \in \mathit{BIT\_VECTOR} \times \mathit{BIT\_VECTOR} \fun \mathit{BIT\_VECTOR} \land \\
% \mathit{bv\_catenate} = \lambda v, w \bullet (v \in \mathit{ BIT\_VECTOR} \land w \in \mathit{ BIT\_VECTOR}  \mid v 
% %\conc
%   w)
% \end{array}
% $

\hspace*{0.00in} \it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$ \it BIT\_VECTOR $\land$

\hspace*{0.00in} \it bv\_catenate \rm =  $\lambda$  v\rm ,\it w \rm . \rm (\it v  $\in$  \it BIT\_VECTOR $\land$ \it w $\in$  \it BIT\_VECTOR  $\mid$  \it v $\cat$ \it w\rm )

We also define a function $\mathit{bv\_zero}$ that, given a positive
integer $n$, return a bit vector of size $n$, with all bits set to 0.
A similar function, called $\mathit{bv\_one}$, with all bits set to 1
is also defined but not presented here.

$
\begin{array}{l}
\mathit{bv\_zero} \in \nat_1 \fun \mathit{BIT\_VECTOR} \land \\
\mathit{bv\_zero} = \lambda n \bullet (n \in \nat_1 \mid 1..n \times \{0\}) 
\end{array}
$

Additionally, the module provides definitions for the classical
logical combinations of bit vectors: $\mathit{bit\_not}$,
$\mathit{bit\_and}$, $\mathit{bit\_or}$ and $\mathit{bit\_xor}$. Only
the first two are presented here. Observe that the domain of the
binary operators is restricted to pairs of bit vectors of the same
length:

$
\begin{array}{l}
\mathit{bv\_not} \in \mathit{BIT\_VECTOR} \fun \mathit{BIT\_VECTOR} \land \\
\mathit{bv\_not} = \lambda v \bullet (v \in \mathit{BIT\_VECTOR} \mid \\
\quad \lambda i \bullet 0 .. \mathit{bv\_size}(v)-1 \mid \mathit{bit\_not}(v(i))) \land \\
\mathit{bv\_and} \in \mathit{BIT\_VECTOR} \times \mathit{BIT\_VECTOR} \fun \mathit{BIT\_VECTOR} \land \\
\mathit{bv\_and} = \lambda v_1, v_2 \bullet (v_1 \in \mathit{BIT\_VECTOR} \land v_2 \in \mathit{BIT\_VECTOR} \land \\
\quad \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \mid \lambda i \bullet 0 .. \mathit{bv\_size}(v_1)-1 \mid \mathit{bit\_and}(v_1(i), v_2(i)))
\end{array}
$

We provide several lemmas on bit vector operations. These lemmas
express properties on the size of the result of the operations
as well as classical algebraic properties such as associativity
and commutativity. For example:

$
\begin{array}{l}
\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \mathit{bv\_size}(\mathit{bv\_not}(v)) = \mathit{bv\_size}(v)) \\

\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \mathit{bv\_not}(\mathit{bv\_not}(v)) = v) \\

\forall v_1, v_2 \bullet (\{v_1, v_2\} \subseteq \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \implies \\
\quad \mathit{bv\_size}(\mathit{bv\_and}(v_1, v_2)) = \mathit{bv\_size}(v_1)) \\

\forall v_1, v_2 \bullet (\{v_1, v_2\} \subseteq \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \implies \\
\quad \mathit{bv\_size}(\mathit{bv\_and}(v_1, v_2)) = \mathit{bv\_size}(v_2)) \\

\forall v_1, v_2 \bullet (\{v_1, v_2\} \subseteq \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \implies \\
\quad \mathit{bv\_size}(\mathit{bv\_and}(v_1, v_2)) = \mathit{bv\_and}(v_2, v_1)) \\

\forall v_1, v_2, v_3 \bullet (\{v_1, v_2, v_3\} \subseteq \mathit{BIT\_VECTOR}
\land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \land \\
\quad \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_3) \implies \\ 
\quad \mathit{bv\_and}(\mathit{bv\_and}(v_1, v_2), v_3) = \mathit{bv\_and}(v_1,
\mathit{bv\_and}(v_2, v_3))) \\

\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \\
\quad \mathit{bv\_and}(v, \mathit{bv\_zero}(\mathit{bv\_size}(v))) = \mathit{bv\_zero}(\mathit{bv\_size}(v))) \\

\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \\
\quad \mathit{bv\_and}(v, \mathit{bv\_one}(\mathit{bv\_size}(v))) = v) \\

\end{array}
$

\subsection{Modelling bytes}

Bit vectors of length 8 are bytes. They form a common entity in
hardware design. We provide the following definitions:

$
\begin{array}{l}
\mathit{BYTE\_WIDTH} = 8 \land \mathit{BYTE\_INDEX} = 0..(\mathit{BYTE\_WIDTH}-1) \land \\
\mathit{BYTE} \subseteq \mathit{BIT\_VECTOR} \land \mathit{BYTE} = \{ v \in \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v) = \mathit{BYTE\_WIDTH}\} \land \\
\mathit{BYTE\_ZERO} \in \mathit{BYTE} \land \mathit{BYTE} = \mathit{BYTE\_INDEX} \times \{0\} 
\end{array}
$

\subsection{Bit vector arithmetics}

Bit vectors are used to represent and combine numbers: integer ranges
(signed or unsigned) and floating point numbers.

Our library defines a function $\mathit{bv\_to\_nat}$ that maps bit
vectors to natural numbers:

$
\begin{array}{l}
\mathit{bv\_to\_nat} \in \mathit{BIT\_VECTOR} \fun \nat \land \\
\mathit{bv\_to\_nat} = \lambda v \bullet (v \in \mathit{BIT\_VECTOR} \mid \sum i \bullet (i \in \dom(v) \bullet v(i) \times 2^i))
\end{array}
$

We provide the following lemma:

$
\begin{array}{l}
\forall n \bullet (n \in \nat_1 \implies \mathit{bv\_to\_nat}(\mathit{bv\_zero}(n)) = 0)
\end{array}
$

\subsection{Basics data types}

There are data types commonly used in the microcontrollers. They are placed in
the project Types Librarys and define functions to manipulate data on levels of
basic types from hardware. However, in general is needed to detail or redefines
the types and its functions for each platform. Then, the specific concepts are
defined in the project's own platform.

There are six common basics data types, see details in table below.

\begin{tabular}{|p{3.5cm}|c|c|c|}\hline
 $Type\ Name$ & $Range$ & $Physical\ Size $\\ 
\hline
 UCHAR& 0..255 & 1 byte\\ \hline SCHAR& -128..127 & 1 byte\\ \hline BYTE & -- & 1
 byte\\ \hline USHORTINT & 0..65.535 & 2 byte\\ \hline SSHORTINT &
 -32.768..32.767 & 2 byte\\ \hline BV16 & -- & 2 byte\\ \hline
\end{tabular}

There is, for example, the function $\mathit{bv\_to\_nat}$ that is specialized
to $\mathit{byte\_uchar}$. How the set $\mathit{BYTE}$ is a subset of the
$\mathit{BIT\_VECTOR}$, then this function can defined as follows:

$
\begin{array}{l}
\mathit{byte\_uchar} = \lambda (v) \bullet ( v \in BYTE | bv\_nat(v) )
\end{array}
$

Usually, the definitions of the functions of the module $\mathit{TYPES}$ are the
re-used from the basic functions of the hardware library. This provides greater
confidence and facilitates the process of proof, because the prover can use the
defined previous lemma.

A more specific model of $\mathit{byte\_uchar}$ functions is:

$
\begin{array}{l}
\mathit{byte\_uchar} \in \mathit{BYTE}  \fun  \mathit{UCHAR}  \land \\
   \mathit{byte\_uchar} =  \lambda ( v0 )  \bullet  
   ( v0 \in \mathit{BYTE} |  2^{7}* bv\_get( v0 ,7 ) + 2^{6} * bv\_get( v0 ,6 ) + 2^{5}* bv\_get( v0 , 5   )\\ 
   + 2^{4}* bv\_get( v0 , 4 ) + 2^{3}*bv\_get( v0 , 3 ) + 2^{2}* bv\_get( v0 , 2 ) + 2 * bv\_get( v0 , 1 ) +  bv\_get( v0 ,0) )

\end{array}
$

The inverse function is easily defined as $\mathit{uchar\_byte}$.

$
\begin{array}{l}
\mathit{uchar\_byte} \in \mathit{UCHAR}  \fun  \mathit{BYTE}  \land \\
   \mathit{uchar\_byte} = \  ^{-1}(\mathit{byte\_uchar}
   )
\end{array}
$

We also created the following lemmas:

$
\begin{array}{l}
 \forall (val) \bullet (val \in \mathit{UCHAR} |
 \mathit{byte\_uchar}(\mathit{uchar\_byte}(val)) = val) \land\\
 \forall (by) \bullet (by \in \mathit{BYTE} |
 \mathit{uchar\_byte}(\mathit{byte\_uchar}(by)) = by)
\end{array}
$

Similarly, several other functions and lemmas were created for all other
data types.

\section{Description of the B model from Z80}
\label{sec:z80}

The \textit{Z80} is an important microcontroller developed in 1976 by \textit{Zilog}. In this time, It got a great
acceptance because suported all set instruction of \textit{8080}, one of the predecessors of the \textit{Pentium}
from \textit{Intel Corporation}. Then, the Z80 has the set instructions similar to the x86 architecture, it helps 
to build the model of x86 and it provides good features.

The main module includes a instance of memory module and accesses the definitions from: basic data types modules,
$\mathit{ALU}$ and $\mathit{Power2}$ \footnote{The $\mathit{Power2}$ module has the basics definitions to help the
theorems prover about calculus of power.}.

\begin{sloppypar}

\bf MACHINE

\hspace*{0.15in}\it Z80

\bf INCLUDES

\hspace*{0.10in}\it MEMORY

\bf SEES

\hspace*{0.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,

\hspace*{0.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,

\hspace*{0.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,

\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION, \it POWER2
\end{sloppypar}

% Chacteristics of microcontroller
The \textit{Z80} supports 158 differents instructions including all the 78 from
\textit{8080}.The instructions are classified into these categories: load and exchange;
block transfer and search; arithmetic and logical; rotate and shift; bit
manipulation (set, reset, test); jump, call and return; input/output; and basic
cpu control. Each instruction and external action at model is represented by B
operations. By default, all parameters from operations are or pre-defined elements
in the model or integers values in the decimal representation.

The internal registers contain 208 bits of reading/writing memory. It includes two sets of six general purpose
registers which may be used individually as  8-bits registers or as 16-bits register pairs.  The work registers are
represented by variable $\mathit{rgs8}$. The domain of $\mathit{rgs8}$ ($\mathit{id\_regs8}$) is a set formed by
identifiers of registers of 8 bits. These registers can be accessed in pairs, forming 16-bits, resulting in other
set of identifiers of 16-bits registers, named $\mathit{id\_reg16}$. The main work register of Z80 is the
accumulator ($\mathit{rgs8(a0)}$) used for arithmetic, logic, input/output and loading/storing operations.

\begin{sloppypar}
\bf SETS

\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0 \rm , \it f0 \rm , \it f\_0 \rm , \it a\_0 \rm ,

\hspace*{1.0 in}\it b0 \rm , \it c0 \rm , \it b\_0 \rm , \it c\_0 \rm ,

\hspace*{1.00in}\it d0 \rm , \it e0 \rm , \it d\_0 \rm , \it e\_0 \rm ,

\hspace*{1.0in}\it h0 \rm , \it l0 \rm , \it h\_0 \rm , \it l\_0 \} ;

\hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{ \it BC \rm , \it DE \rm , \it HL \rm , \it SP \rm , \it AF \rm \}
\end{sloppypar}

\subsection{Modelling registers and input and output ports} [I need review the english]

The Z80 CPU includes alternative set of accumulator, flag and general registers. The CPU contains a stack
pointer ($\mathit{sp}$), program counter ($\mathit{pc}$), two index registers ($\mathit{ix}$ and $\mathit{iy}$), an
interrupt register ($\mathit{i\_}$), a refresh register ($\mathit{r\_}$), two bits ($\mathit{iff1}$,
$\mathit{iff2}$) used to control the interruptions, a pair of bits to define the interruption mode ($\mathit{im}$)
and the input and output ports ($\mathit{i\_o\_ports}$). Below, its definitions are represented by
\textit{INVARIANT}.

\begin{sloppypar}
\bf INVARIANT

\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$

\hspace*{0.10in}\it pc  $\in$  \it INSTRUCTION  $\land$  \it sp  $\in$  \it BV16  $\land$  \it ix  $\in$  \it BV16  $\land$  \it iy  $\in$  \it BV16  $\land$

\hspace*{0.10in}\it i\_  $\in$  \it BYTE  $\land$  \it r\_\hspace*{0.10in} $\in$  \it BYTE  $\land$

\hspace*{0.10in}\it iff1  $\in$  \it BIT  $\land$ \hspace*{0.10in}\it iff2  $\in$  \it BIT  $\land$

\hspace*{0.10in}\it im \rm : \rm (\it BIT $\times$ \it BIT\rm )  $\land$

\hspace*{0.10in}\it i\_o\_ports  $\in$  \it BYTE  $\fun$  \it BYTE
\end{sloppypar}

\subsection{Flag register} [I need review the english]
Another important element is the ``z'' register ($\mathit{rgs8(z0)}$),  that is used as a flag register. This
register uses only six bits to represent the result status of each instruction. 	
According to the official manual the bits 3 and 5 are not used and the others bits have the follow meaning:
\begin{description}
  \item[$\mathit{bv\_get(rgs8(z0),0)}$] - The Carry bit
  \item[$\mathit{bv\_get(rgs8(z0),1)}$] - The Add/Subtract bit
  \item[$\mathit{bv\_get(rgs8(z0),2)}$] - The Parity or Overflow bit
  \item[$\mathit{bv\_get(rgs8(z0),4)}$] - The Half Carry bit
  \item[$\mathit{bv\_get(rgs8(z0),6)}$] - The Zero bit
  \item[$\mathit{bv\_get(rgs8(z0),7)}$] - The Sign bit 
\end{description}

\subsection{Manipulation data functions from Z80} [I need review the english] There are some specific functions from
Z80 to manipulate the data. The $\mathit{bv\_ireg\_plus\_d}$ is  used to indexed address. It received the
value of register ($\mathit{ix}$ or $\mathit{iy}$) and displacement to return the sum, the result is the memory address dislocated, see
its the definition.

\hspace*{0.0in}\it bv\_ireg\_plus\_d \rm : \rm(\it BV16  $\times$  \it SCHAR  $\fun$  \it BV16\rm )  $\land$

\hspace*{0.0in}\it bv\_ireg\_plus\_d \rm =  $\lambda$  \rm ( \it ix\_iy \rm , \it disloc \rm ) \rm . \rm ( \it ix\_iy  $\in$  \it BV16  $\land$  \it disloc  $\in$  \it SCHAR

\hspace*{0.20in}$\mid$ \it ushort\_bv16 \rm ( \rm (\it bv16\_ushort \rm ( \it ix\_iy \rm ) \rm + \it disloc \rm ) 
$\mod$ \rm 6\rm 5\rm 5\rm 3\rm 6 \rm ) \rm )

Another derived function is  $\mathit{bv\_9ireg\_plus\_d0}$ , this returns the value in the memory  address
returned by $\mathit{bv\_ireg\_plus\_d}$ function ands its definition is similar.

There is a specific function to refresh the flag register, it is named $\mathit{update\_reg\_flag}$. It is typed of
follow mode: \it update\_flag\_reg \rm $\in$ \rm (\it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it
BIT $\times$  \it BIT  $\times$  \it BIT\rm) $\fun$  \rm (\rm \{\it f0\rm \}  $\times$  \it BYTE\rm ).

\it update\_flag\_reg \rm =  $\lambda$  \rm (\it s7 \rm, \it z6 \rm,\it h4 \rm,\it pv2 \rm ,\it n1
\rm ,\it c0 \rm)$\bullet$

\rm ( \it s7 $\in$ \it BIT $\land$ \it z6 $\in$ \it BIT $\land$ \it h4 $\in$ \it
BIT $\land$ \it pv2 $\in$ \it BIT $\land$ \it n1 $\in$ \it BIT $\land$ \it c0 $\in$ \it BIT

\hspace*{0.20in}\rm $\mid$( \it f0  $\mapsto$  \rm \rm [\it c0\rm , \it n1\rm , \it pv2\rm , \rm 1\rm , \it
h4\rm , \rm 1\rm , \it z6\rm , \it s7\rm \rm ]\rm ) \rm )

\subsection{Program, stack and data memory}[I need review the english]

The Z80 uses a unique memory for storing program instructions, data stack and data work. The memory has 16-bits
addressing and each address represent a byte. Thus, the data from the memory module is very simple, as shown below,
but an additional care must be taken to preserve the consistency of memory.

\begin{sloppypar}
\bf INVARIANT \\
\hspace*{0.10in}\it mem  $\in$  \it BV16  $\fun$  \it BYTE

\end{sloppypar}

In general, the instructions can access all memory address, but it is very dangerous. For add more security, is
importante that the program instructions has the access limited by region. Thus, the designer can specify address
regions to restrict the access from instructions. The address regions can be especified how shown below.

$
\begin{array}{l}
\mathit{PROGRAM\_R\_ADR} = 0..16384 \land \mathit{DATA\_R\_ADR} = 16385..49151 \land \\
\mathit{STACK\_R\_ADR} = 49152..65535
\end{array}
$

\begin{property}[Assuring the absence of overlapping of address regions]
 To assure that address regions are well defined, then the designer must to verify the below expression.
\end{property}

\begin{sloppypar}
\hspace*{0.10in}\it PROGRAM\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$  STACK\_R\_ADR $=$ \{\}
\end{sloppypar}

\begin{property}[Preserving the consistency of the memory]
In general, the access to some address regions is dangerous. Then, in each instruction has a specific
pre-condition that verify if the new address memory, that will be updated, is member of its region. For example,
the $\mathit{PUSH}$ program instruction allow write only in the region of stack ($\mathit{STACK\_R\_ADR}$). 
%Thus,if the user try to write in incorrect region then the called of operation cannot be proved.
\end{property}

\subsection{Arithmetic logic unit}

There are many functions defined in the module ALU. In general, these functions take basic definitions to build
new specific functions. The function half8UCHAR is used to get the half part of UCHAR value. It is importante to
know the half carry and it is used in the $\mathit{add8UCHAR}$.

\hspace*{0.0in}

\hspace*{0.0in}\it half8UCHAR  $\in$  \it UCHAR  $\fun$  \it UCHAR  $\land$

\hspace*{0.0in}\it half8UCHAR \rm =  $\lambda$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it UCHAR  $\mid$  \it ww  $\mod$  \it $2^{4}$\rm )

\hspace*{0.0in}

The function $\mathit{add8UCHAR}$ receive a bit carry and two $\mathit{UCHAR}$ values and return respectively the 
sum, the sign bit, the carry bit, the half carry bit and the zero bit. It is typed how follow: \it add8UCHAR \rm :
\rm (\it BIT $\times$ \it UCHAR $\times$ \it UCHAR\rm ) $\fun$ \rm (\it UCHAR $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm ) and its definitions is:

\hspace*{0.20in}

\hspace*{0.0in}\it add8UCHAR\rm = $\lambda$ \rm(\it carry\rm,\it w1\rm \it w2\rm)\rm.\rm

\hspace*{0.0in}(\it carry $\in$  \it BIT  $\land$  \it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it
UCHAR $\mid$

\hspace*{0.40in}\rm(\rm(\rm(\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm ),

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it uchar\_schar\rm (\it w1\rm ) \rm + \it uchar\_schar \rm (\it w2\rm ) $<$ \rm 0\rm ),

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it w1 \rm + \it w2 $>$ \it UCHAR\_MAX\rm )\rm ,

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm ) \rm + \it half8UCHAR\rm ( \it w2\rm )  $\geq$  \it $2^{4}$\rm )\rm,

\hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm )\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )

\hspace*{0.20in}

A similar function to subtract operation is $\mathit{substract8UCHAR}$. There are the same functions for the
$\mathit{SCHAR}$ type, they are respectively $\mathit{add8SCHAR}$ and $\mathit{substract8SCHAR}$, all these
functions are of 8 bits ($\mathit{BYTE}$) and defined similarly. In the same way, the arithmetic functions for 16 bits
($\mathit{BV16}$) are defined.

Other more simply functions also are typed and explained below:

\begin{itemize}
  \item ALU \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and returns its decrement. There is a
  similar function named $\mathit{dec}$.

  \item ALU \it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the pc actual
  value and return its increment.

  \item ALU \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
  byte is zero, otherwise returns 0.

  \item byte \it is\_zero $\in$ BYTE  $\fun$   BIT \rm - It returns 1 if the received byte is zero, otherwise
  returns 0.

  \item byte \it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE \rm - It receives a byte and returns its rotated to
  right. There is a similar function named $\mathit{rotateleft}$.

  \item ALU \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.
\end{itemize}
Every logic functions defined in the $\mathit{BYTE}$ module are included in the $\mathit{ALU}$ module. Then, it was
not necessary rewrite these functions.

\subsection{Interrupts}

% Visão Geral
In the Z80, the interrupts allows that the devices suspend a routine from CPU and start another service routine.
This service routine can exchange data or signals between CPU and external devices. When a routine is finished, then
the CPU come back to the last routine that was interrupted.

For the interrupts, the following things are important:  the interrupt flip-flops ($\mathit{iff1}$ and
$\mathit{iff2}$), the types of interrupts (maskable and non-maskable), the interrupt mode (set with the $\mathit{IM
0}$, $\mathit{IM 1}$, $\mathit{IM 2}$ instructions) and the $\mathit{i\_}$ register.

The $\mathit{iff1}$ and $\mathit{iff2}$ control the maskable interrupts ($\mathit{INT}$). When the $\mathit{iff1}$
is set, the interrupt is enable, otherwise it is disable. The $\mathit{iff2}$ is used only as a tempory storage
place for $\mathit{iff1}$. The $\mathit{iff1}$ is set or reset by instructions $\mathit{EI}$ and $\mathit{DI}$.

The interruptions and the \textit{reset} action change the state of
microcontroller. Theses actions are modeled by B operations and the its main
effects are shown below\footnote{Some definitions: $\mathit{sp\_minus\_two}$ =
$\mathit{dec\_BV16(dec\_BV16(sp))}$, $\mathit{sp\_minus\_one}$=
$\mathit{dec\_BV16(sp)}$, $\mathit{pc\_high}$ is the most significant 8 bits and
$\mathit{pc\_low}$ is the least significant. }.

 \textbf{NMI} - Non-maskable interrupts - The non-maskable cannot be disable
 by the programmer. Then, when a device makes a request, the $sp$ is pushed, the $pc$ receive
 $66H$ (102 in decimal), the $\mathit{iff1}$ is reset , $\mathit{iff2}$ stores
 $\mathit{iff1}$ and refresh register is updated.

% $updateStack(\{ (sp\_minus\_two \fun low\_pc ),(sp\_minus\_one
%    \fun high\_pc ) \}) ||\\  sp:= sp\_minus\_two || pc := 66H || iff1:=0 || 
%    iff2:= iff1 $
\begin{sloppypar}
\bf updateStack\rm (\rm \{ \rm (\it sp\_minus\_two  $\mapsto$  \it pc\_low\rm )\rm ,\rm (\it sp\_minus\_one  $\mapsto$ \it pc\_high \rm ) \rm \}\rm )$\para$

\it sp \rm := \it sp\_minus\_two  $\para$ \it pc \rm := \rm 1\rm 0\rm 2 $\para$ \it iff1\rm :=\rm 0  $\para$  \it iff2\rm := \it iff1 $\para$

\it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\\
\end{sloppypar}

  \textbf{INT} - Maskable Interrupt -  This is usualy reserved to important functions that can be enabled and
  disabled by the programmer. When a maskable interrupt action happens, both $\mathit{iff1}$ and $\mathit{iff2}$ are
  cleared, disabling the interrupts, the $sp$ is pushed, the refresh register is updated  and the other effects
  depend on the interrupt mode.

 \begin{itemize}

  \item The mode 0 is compatible with 8080 and  \it im \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a 
  non-maskable interruption happen , the CPU fetches an instruction of one byte from an external device, usually an
  RST instruction, and the CPU executes it. The instruction code is received from a external device by data bus and
  it is represented by integer parameter called $\mathit{byte\_bus}$ .

   \item The mode 1 is the easiest and \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 1 \rm ). Simply, when a non-maskable
   interruption happens, the program counter receive $38H$ (56 in decimal).

   \item The mode 2 is the most powerfull and  \it im \rm = \rm ( \rm 1  $\mapsto$ \rm 1 \rm ). When a non-maskable
   interruption happens, an indirect call can be made to any address memory. The program counter receives in the
   part most significant the $\mathit{i\_}$ register and, in the  part least siginificant, the $\mathit{byte\_bus}$
   with the last bit reset.
 \end{itemize}

The essential part of maskable interrupt is shown below\footnote{ The
 $\mathit{byte\_bus}$ is parameter from  $\mathit{INT}$ operation }.

\begin{sloppypar}

\hspace*{0.00in}\bf IF \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 0 \rm ) \bf THEN

\hspace*{0.10in}\bf IF\hspace*{0.10in}\it byte\_bus \rm $\in$  \it opcodes\_RST\_instruction

\hspace*{0.10in}\bf THEN\hspace*{1.05in}

\hspace*{0.40in}\it pc \rm := \it byte\_bus \rm - \rm 1\rm 9\rm
9\hspace*{0.10in} $\para$

\hspace*{0.40in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.40in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.40in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.10in}\bf ELSIF \it byte\_bus \rm = \it opcode\_\ldots\_instruction

\hspace*{0.40in}\bf \ldots

\hspace*{0.10in}\bf END

\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm =\hspace*{0.10in}\rm ( \rm
0  $\mapsto$  \rm 1 \rm ) \bf THEN

\hspace*{0.10in}\it pc \rm :=\hspace*{0.10in}\rm 5\rm 6  \hspace*{0.80in}
$\para$

\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm (\it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\hspace*{0.85in}

\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm = \rm ( \rm 1  $\mapsto$ 
\rm 1 \rm ) \bf THEN  \hspace*{0.70in}

\hspace*{0.10in}\it pc \rm := \it bv16\_ushort\rm (\it byte\_bv16\rm ( \it i\_
\rm ,\it bv\_clear\rm (\it rotateleft\rm (\it uchar\_byte\rm (\it byte\_bus\rm )\rm )\rm ,\rm 0\rm )\rm )\rm )$\para$

\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.00in}\bf END\hspace*{0.10in}\\
\end{sloppypar}

\textbf{RESET}  - The reset is other external action that just reset the
registers related to the interruptions.

\begin{sloppypar}
\it iff1 \rm :=\rm 0 $\para$ \it iff2\rm :=\rm 0 $\para$  \it  im\rm := \rm (\rm 0 $\mapsto$ \rm 0\rm )  $\para$ 
\it pc\rm :=\rm 0 $\para$ \it i\_ \rm := [0,0,0,0,0,0,0,0] $\para$

\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm (\it a0  $\mapsto$ \rm [0,0,0,0,0,0,0,0] , \rm (\it
f0 $\mapsto$ \rm [0,0,0,0,0,0,0,0] \rm \} $\para$

\it r\_  \rm := [0,0,0,0,0,0,0,0] $\para$ \it sp \rm := [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0]
\end{sloppypar}

% 
%     Para controlar o registrador I, existem duas intrucoes:
% 
%     - LD A,I (carrega o acumulador com o valor contido em I);
%     - LD I,A (carrega I com o valor contido no acumulador).
% 
%     A instrucao LD A,I possui uma funcao extra, ja' discutida
% em outra aula: ela carrega na `flag' de paridade/sobrecarga o
% valor do IFF ("Interruption Flip-Flop"). Se IFF for 0, entao a
% interrupcao mascaravel esta' desativada. Se IFF for 1, entao
% ela esta' ativada.
% %%%% FIM - http://download.unesp.br/msx/asm/course_4.txt
% 
% 
% 
% 
% Instruções relacionadas com o sistema de interrupção:
%  HALT,EI,DI,M0,M1,M2 
%  
%  LD_A_R, LD_A_I\ldots,  
%  CALL, RET,  RETI,RETN

\subsubsection{I/O Ports} % Can be placed in basic definitions in a paragraph
%Referencia o modeo de modelagem do captura de um evento externo de acordo com
% o proposto em [DAVID_gdc_2007]

[[ Flag registers and the associated functions: negative, half, zero.. ]]

\section{Extern  actions and input and output ports}

The Z80 has an extensive set of input and output instruction and 256 ports for
devices. This model can transfer data blocks and between the I/O devices and any
of the internal registers or memory address.

The $\mathit{IN\_r(C)}$ \footnote{The tools B don't allow use parenteses in identifiers, then the characteres
\{``('',``)''\} are replaced respectively by \{``9'',``0''\} in the real specification.} is represented by the follow B
operation. This instruction is from I/O group and update the flag registers.

\hspace*{0.0in}\bf IN\_r\_(C) \rm ( \it rr \rm ) \rm =

\hspace*{0.0in}\bf PRE \it rr  $\in$  \it id\_reg\_8  $\land$  \it rr  $\not =$  \it f0\hspace*{0.15in}\bf THEN

\hspace*{0.20in}\bf ANY

\hspace*{0.40in}\it negative \rm , \it zero \rm , \it half\_carry \rm , \it pv \rm , \it add\_sub \rm , \it carry

\hspace*{0.20in}\bf WHERE

\hspace*{0.40in}\it negative $\in$ \it BIT $\land$ \it zero $\in$ \it BIT $\land$ \it half\_carry $\in$ \it BIT 
$\land$ \it pv $\in$ \it BIT $\land$

 \hspace*{0.40in}\it add\_sub $\in$ \it BIT $\land$ \it carry $\in$ \it BIT  $\land$

\hspace*{0.40in}\it negative \rm = \it is\_negative \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$

\hspace*{0.40in}\it zero \rm = \it is\_zero \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$

\hspace*{0.40in}\it half\_carry \rm = \rm 0  $\land$

\hspace*{0.40in}\it pv \rm = \it parity\_even \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) $\land$

\hspace*{0.40in}\it add\_sub \rm =\hspace*{0.10in}\rm 0  $\land$

\hspace*{0.40in}\it carry \rm = \it z\_c

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm ( \it rr  $\mapsto$  \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) \rm ,

\hspace*{0.40in}\it update\_flag\_reg\rm (\it negative\rm,\it zero\rm,\it half\_carry\rm,\it pv\rm,\it add\_sub\rm,\it carry)\rm\}$\para$

\hspace*{0.40in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.0in}\bf END \rm

% $
% \begin{array}{l} 
% \mathit{IN\_r\_9C0} ( rr ) = \\
% \quad   \PRE rr \in id\_reg\_8    \THEN\\
% \quad\quad \ANY data\_in, negative , zero , half\_carry , pv , add\_sub , carry\\ 
% \quad\quad \WHERE data\_in \in \mathit{BYTE} \land negative \in \mathit{BIT}\land\\
% \quad\quad\quad carry \in \mathit{BIT} \land half\_carry \in \mathit{BIT} \land zero \in \mathit{BIT} \land \\
% \quad\quad\quad negative = is\_negative (data\_in) \land zero = is\_zero(data\_in ) \land\\
% \quad\quad\quad half\_carry = 0 \land pv =parity\_even\_BYTE ( data\_in )    \land\\
% \quad\quad\quad add\_sub =  0 \land carry = z\_c \\
% \quad\quad  \THEN \\
% \quad\quad\quad i\_o\_ports ( rgs8 ( c0 ) ) := data\_in ||\\
% \quad\quad\quad rgs8 := rgs8 <+ \{ ( rr |-> data\_in ) ,\\
% \quad\quad\quad get\_new\_flag\_register\_SZ\_H\_PvNC ( rgs8 , negative , zero, half\_carry , pv , add\_sub , carry ) \} ||\\
% \quad\quad\quad pc := instruction\_next( pc )\\
% \quad\quad \END\\
% \quad \END\\
% \end{array}
% $

\subsubsection{Instruction set} 
% Talk in general and put a example with midlle
% complex

 \section{Conlusitions}

%  Para a modelagen dessa plataforma, essa modelagen contém linhas e $np $
%  provas não obvias. Uma forma de acelerar o processo de prova, foi paralelizar
%  as provas. Isto foi feito em 4 IMacs e consumiu $ tp $ tempo para verificar
%  automaticamente $np - p $. As demais foi necessário uma certa interação com
%  o provador para sua verificação.
%  
%  
%  Algumas informações do manual do Z80 fornecido pela Zilog são imprecisas,
%  o que compremete um pouco a qualidade do modelo, entretanto alguns pontos
%  foram consultados no documento (Undocument Z80 documented-(Sean Young)) e
%  dúvidas foram sanadas com o uso do simulador(Z80 IDE Simulator). 
%  
%  Para oferecer uma maior garantia do modelo com a implementação do processador
%  algumas atividades foram realizadas. Primeiramente, tentantamos animar e gerar
%  código respectivamente nas ferramentas ProB e AtelierB. Entratanto, o modelo
%  cotém algumas construções complexas que ainda não são bem suportada por elas.
%  Então, foi realizada uma apresentação sobre a descrição do modelo e foi
%  realizada um nova revisão com 3 pessoas.

\paragraph{Notes and Comments.}

% ---- Bibliography ----
%
\begin{thebibliography}{5}

\bibitem {Abrial}
Abrial, J. R. The B Book: Assigning Programs to Meanings. Cambridge University Press, United States of
America, 1 edition, 1996.

% Paper about BHDL

% Paper of ALEMÃO PROB ,java virtual machine

% Outros do artigo de bartira

% O manual do Z80

\bibitem {Bell} 
 D. Bell, I. M. E. J. P. Software Engineering: A Programming Approach. Prentice Hall, New York, 2 edition, 1992.

\bibitem {Dantas}
 Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Proposta e avaliação de uma abordagem de desenvolvimento de
 software fidedigno por construção com o método B. In: SEMISH, Belém, 2008. SBC.

\end{thebibliography}

\end{document}
%
% \bibitem {clar:eke}
% Clarke, F., Ekeland, I.:
% Nonlinear oscillations and
% boundary-value problems for Hamiltonian systems.
% Arch. Rat. Mech. Anal. {\bf 78} (1982) 315--333
% %
% \bibitem {clar:eke:2}
% Clarke, F., Ekeland, I.:
% Solutions p\'{e}riodiques, du
% p\'{e}riode donn\'{e}e, des \'{e}quations hamiltoniennes.
% Note CRAS Paris {\bf 287} (1978) 1013--1015
% %
% \bibitem {mich:tar}
% Michalek, R., Tarantello, G.:
% Subharmonic solutions with prescribed minimal
% period for nonautonomous Hamiltonian systems.
% J. Diff. Eq. {\bf 72} (1988) 28--55
% %
% \bibitem {tar}
% Tarantello, G.:
% Subharmonic solutions for Hamiltonian
% systems via a $\bbbz_{p}$ pseudoindex theory.
% Annali di Matematica Pura (to appear)
% %
% \bibitem {rab}
% Rabinowitz, P.:
% On subharmonic solutions of a Hamiltonian system.
% Comm. Pure Appl. Math. {\bf 33} (1980) 609--633
% \end{thebibliography}
%

### 2009-04-20 developments/platforms/Z80/tex/ALU.tex

\documentclass[11pt]{article}
\setlength{\topmargin}{13mm}
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\textheight}{225mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm}
\usepackage{supertabular}
%\input{psfig}

\begin{document}
\begin{sloppypar}

% DEFINITION DES CARACTERES MATHEMATIQUES B
%------------------------------------------
\def\@setmcodes#1#2#3{{\count0=#1 \count1=#3
	\loop \global\mathcode\count0=\count1 \ifnum \count0<#2
	\advance\count0 by1 \advance\count1 by1 \repeat}}

\@setmcodes{`A}{`Z}{"7441}
\@setmcodes{`a}{`z}{"7461}

\mathcode`\;="8000 % Makes ; active in math mode
{\catcode`\;=\active \gdef;{\semicolon\;}}
\mathchardef\semicolon="003B
%    Nominal distance from top of paper to top of page
\topmargin 0 pt
\textheight 53\baselineskip

%   Left margin on odd-numbered pages
\oddsidemargin  0.15 in
%   Left margin on even-numbered pages
\evensidemargin 0.35 in
%   Width of marginal notes.
\marginparwidth 1 in
%   Note that \oddsidemargin = \evensidemargin
\oddsidemargin 0.25 in
\evensidemargin 0.25 in
\marginparwidth 0.75 in
\textwidth 5.875 in % Width of text line.

\setlength{\parindent}{0pt}
\setlength{\parskip}{0ex}

% DEFINITION DES FONTS
%---------------------
% The AMS extra symbol fonts are loaded.
% Note: sometimes called euxm10
\font\msx=msam10
% Note: sometimes called euym10
\font\msy=msbm10

\newfam\msxfam \textfont\msxfam=\msx
\newfam\msyfam \textfont\msyfam=\msy

\def\famletter#1{\ifcase #1 0\or 1\or 2\or 3\or 4\or 5\or 6\or 7\or
	8\or 9\or A\or B\or C\or D\or E\or F\fi}

\edef\fx{\famletter\msxfam}
\edef\fy{\famletter\msyfam}

\def\bbold{\fam\msyfam \msy}

% SYMBOLES B
%-----------
% makes a quoted expression in mathematical text
\def\token#1{\hbox{`$#1$'}}
% used for error messages in Z specs
\def\report#1{\hbox{`{\tt #1}'}}

% \@myop makes an operator, with a strut to defeat TeX's vertical adjustment.
\def\@myop#1{\mathop{\mathstrut{#1}}\nolimits}

% This underscore doesn't have the little kern --- you get an italic
% correction anyway in math mode.
\def\_{\leavevmode \vbox{\hrule width0.5em}}

% Save \q as \xq for quantifiers q.
\let\xforall=\forall
\let\xexists=\exists
\let\xlambda=\lambda
\let\xmu=\mu

% \p and \f make arrows with 1 and 2 crossings resp.
\def\p#1{\mathrel{\ooalign{\hfil$\mapstochar\mkern 5mu$\hfil\cr$#1$}}}
\def\f#1{\mathrel{\ooalign{\hfil
	$\mapstochar\mkern 3mu\mapstochar\mkern 5mu$\hfil\cr$#1$}}}

\let\mc=\mathchardef

\def	\pow		{\mbox{${\cal P}$}}
\def	\po1		{\mbox{${\cal P}_1$}}
\let	\cross		\times
\def	\lambda		{\@myop{\xlambda}}
\def	\lnot		{\neg\;}
\def	\land		{\mathrel{\wedge}}
\def	\lor		{\mathrel{\vee}}
\let	\implies	\Rightarrow
\let	\iff		\Leftrightarrow
\def	\forall		{\@myop{\xforall}}
\def	\exists		{\@myop{\xexists}}
\def	\semi		{\mathrel{\comp}}
\def	\ssemi		{\mathbin{\rm ;}}
\let	\ensembleVide	\emptyset
\let	\rel		\leftrightarrow
\def	\dom		{\@myop{\sf dom}}
\def	\ran		{\@myop{\sf ran}}
\def	\id		{\@myop{\sf id}}
\def	\comp		{\mathbin{\raise
			0.6ex\hbox{\oalign{\hfil$\scriptscriptstyle
			\rm o$\hfil\cr\hfil$\scriptscriptstyle\rm 9$\hfil}}}}
\def	\para		{\mbox{$\mid\mid$}}
\mc	\dres		"2\fx43
\mc	\rres		"2\fx42
\def	\ndres		{\mathbin{{\dres} \llap{$-$}}}
\def	\nrres		{\mathbin{{\rres}\llap{$-$}}}
\def	\lover		{\mathbin{{\dres} \llap{$-\!\!\!\!-\!$}}}
\def	\rover		{\mathbin{{\rres}\llap{$\!-\!\!\!-$}}}
\let	\fun		\rightarrow
\def	\pfun		{\p\fun}
\def	\pinj		{\p\inj}
\mc	\inj		"3\fx1A
\def	\psurj		{\p\surj}
\mc	\surj		"3\fx10
\def	\bij		{\surj\!\!\!\!\!\!\!\inj}
\def	\nat		{\mbox{${\cal N}$}}
\def	\na1		{\mbox{${\cal N}_1$}}
\def	\num		{\mbox{${\cal Z}$}}
\def	\int		{\mbox{${\cal Z}$}}
\def	\rat		{\mbox{${\cal Q}$}}
\def	\div		{\mathbin{\rm /}}
\def	\mod		{\mathbin{\bf mod}}
\def	\upto		{\mathbin{\ldotp\ldotp}}
\def	\finset		{\mbox{${\cal F}$}}
\def	\finse1		{\mbox{${\cal F}_1$}}
\def	\ffun		{\f\fun}
\def	\finj		{\f\inj}
\def	\seq		{\@myop{\rm seq}}
\def	\cat		{\mathbin{\raise 0.8ex\hbox{$\mathchar"2\fx61$}}}
\def	\sep		{\hspace*{.05in}}

\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{0}

%-------------------%
% Debut du document %
%-------------------%

\bf MACHINE

\hspace*{0.20in}\it ALU

\hspace*{0.20in}

\bf SEES

\hspace*{0.20in}\it BIT\_DEFINITION \rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it BYTE\_DEFINITION\rm ,

\hspace*{0.20in}\it BV16\_DEFINITION\rm ,

\hspace*{0.20in}

\hspace*{0.20in}\it UCHAR\_DEFINITION\rm ,

\hspace*{0.20in}\it SCHAR\_DEFINITION\rm ,

\hspace*{0.20in}\it SSHORT\_DEFINITION\rm ,

\hspace*{0.20in}\it USHORT\_DEFINITION\rm ,

\hspace*{0.20in}\it POWER2

\hspace*{0.20in}

\bf DEFINITIONS

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it pow2\_0 \rm == \rm 1\rm ;

\hspace*{0.20in}\it pow2\_1 \rm == \rm 2\rm ;

\hspace*{0.20in}\it pow2\_2 \rm == \rm 4\rm ;

\hspace*{0.20in}\it pow2\_3 \rm == \rm 8\rm ;

\hspace*{0.20in}\it pow2\_4 \rm == \rm 1\rm 6\rm ;

\hspace*{0.20in}\it pow2\_5 \rm == \rm 3\rm 2\rm ;

\hspace*{0.20in}\it pow2\_6 \rm == \rm 6\rm 4\rm ;

\hspace*{0.20in}\it pow2\_7 \rm == \rm 1\rm 2\rm 8\rm ;

\hspace*{0.20in}\it pow2\_8 \rm == \rm 2\rm 5\rm 6\rm ;

\hspace*{0.20in}\it pow2\_9 \rm == \rm 5\rm 1\rm 2\rm ;

\hspace*{0.20in}\it pow2\_10 \rm == \rm 1\rm 0\rm 2\rm 4\rm ;\hspace*{0.20in}

\hspace*{0.20in}\it pow2\_11 \rm == \rm 2\rm 0\rm 4\rm 8\rm ;\hspace*{0.10in}

\hspace*{0.20in}\it pow2\_12 \rm == \rm 4\rm 0\rm 9\rm 6\rm ;\hspace*{0.15in}

\hspace*{0.20in}\it pow2\_13 \rm == \rm 8\rm 1\rm 9\rm 2\rm ;

\hspace*{0.20in}\it pow2\_14 \rm == \rm 1\rm 6\rm 3\rm 8\rm 4\rm ;\hspace*{0.15in}

\hspace*{0.20in}\it pow2\_15 \rm == \rm 3\rm 2\rm 7\rm 6\rm 8\rm ;

\hspace*{0.20in}\it pow2\_16 \rm == \rm 6\rm 5\rm 5\rm 3\rm 6\hspace*{0.35in}

\hspace*{0.20in}

\hspace*{0.20in}

\bf CONCRETE\_CONSTANTS

\hspace*{0.20in}\it abs\rm ,

\hspace*{0.20in}

\hspace*{0.20in}\it is\_zero16USHORT\rm ,

\hspace*{0.20in}\it is\_negative\rm ,

\hspace*{0.20in}

\hspace*{0.20in}\it half8UCHAR\rm ,

\hspace*{0.20in}\it inc\rm ,\it dec\rm ,

\hspace*{0.20in}\it simple\_add8UCHAR\rm ,

\hspace*{0.20in}\it add8UCHAR \rm , \it substract8UCHAR\rm ,

\hspace*{0.20in}

\hspace*{0.20in}\it add16USHORT\rm ,

\hspace*{0.20in}\it add\_carryUSHORT\rm ,

\hspace*{0.20in}\it add\_halfcarryUSHORT\rm ,

\hspace*{0.20in}\it sub16USHORT\rm ,

\hspace*{0.20in}\it sub\_carryUSHORT\rm ,

\hspace*{0.20in}\it sub\_halfcarryUSHORT\rm ,

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it inc\_BV16\rm ,

\hspace*{0.20in}\it dec\_BV16\rm ,

\hspace*{0.20in}

\hspace*{0.20in}\it update\_refresh\_reg\rm ,

\hspace*{0.20in}

\hspace*{0.20in}\it instruction\_next \rm ,

\hspace*{0.20in}\it instruction\_jump \rm ,

\hspace*{0.20in}

\hspace*{0.20in}\it daa\_function

\hspace*{0.20in}

\bf PROPERTIES

\hspace*{0.20in}

\hspace*{0.20in}\it abs  $\in$ \hspace*{0.10in}\bf INT\hspace*{0.10in} $\fun$  \bf INT  $\land$

\hspace*{0.20in}\it abs \rm =  $\lambda$  \rm (\it nn\rm )\rm .\rm (\it nn  $\in$  \bf INT\hspace*{0.10in} $\land$  \rm 0  $\leq$  \it nn\hspace*{0.10in} $\mid$  \it nn\hspace*{0.10in}\rm )  $\cup$

\hspace*{0.20in} $\lambda$  \rm (\it nn\rm )\rm .\rm (\it nn  $\in$  \bf INT\hspace*{0.10in} $\land$  \rm 0 $>$ \it nn $\mid$  \rm -\it nn\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it is\_zero16USHORT  $\in$  \it USHORT  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it is\_zero16USHORT \rm =  $\lambda$ \rm (\it nat1\rm )\rm .\rm (\it nat1  $\in$  \it USHORT  $\mid$  \it bool\_bit\rm (\bf bool\rm (\it nat1 \rm = \rm 0\hspace*{0.10in}\rm ) \rm ) \rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it is\_negative \rm =  $\lambda$ \rm (\it w1\rm )\rm .\rm (\it w1  $\in$  \it BYTE  $\mid$ \hspace*{0.10in}\it w1\rm (\rm 7\rm )\hspace*{0.10in}\rm ) $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it half8UCHAR  $\in$  \it UCHAR  $\fun$  \it UCHAR  $\land$

\hspace*{0.20in}\it half8UCHAR \rm =  $\lambda$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it UCHAR  $\mid$  \it ww  $\mod$  \it pow2\_4\rm )  $\land$ \hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it inc  $\in$ \hspace*{0.10in}\it BYTE\hspace*{0.10in} $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it inc \rm =  $\lambda$  \rm (\it w1\rm )\rm . \rm (\it w1  $\in$  \it BYTE  $\mid$  \it uchar\_byte\rm (\rm (\rm (\it byte\_uchar\rm (\it w1\rm ) \rm + \rm 1\rm )\hspace*{0.10in} $\mod$  \it pow2\_8 \rm ) \rm ) \rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it dec  $\in$ \hspace*{0.10in}\it BYTE\hspace*{0.10in} $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it dec \rm =  $\lambda$  \rm (\it w1\rm )\rm . \rm (\it w1  $\in$  \it BYTE  $\mid$  \it uchar\_byte\rm ( \rm (\it abs\rm (\it byte\_uchar\rm (\it w1\rm ) \rm - \rm 1 \rm )\rm )  $\mod$  \it pow2\_8 \rm ) \rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it simple\_add8UCHAR\hspace*{0.10in}\rm : \rm (\it UCHAR  $\times$  \it UCHAR\rm )  $\fun$  \it UCHAR  $\land$

\hspace*{0.20in}\it simple\_add8UCHAR \rm =  $\lambda$  \rm (\it w1\rm ,\it w2\rm ) \rm . \rm (\it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it UCHAR  $\mid$  \rm (\it w1 \rm + \it w2\rm )  $\mod$  \it pow2\_8\rm )  $\land$ \hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it add8UCHAR \rm : \rm (\it BIT  $\times$  \it UCHAR  $\times$  \it UCHAR\rm )  $\fun$  \rm (\it UCHAR  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm )  $\land$

\hspace*{0.20in}\it add8UCHAR \rm =  $\lambda$  \rm (\it carry\rm ,\it w1\rm , \it w2\rm )\rm . \rm ( \it carry  $\in$  \it BIT  $\land$  \it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it UCHAR\hspace*{0.20in} $\mid$

\hspace*{0.40in}\rm (\rm ( \rm ( \it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it pow2\_8\rm )\hspace*{0.15in}\rm ,\hspace*{0.10in}

\hspace*{0.60in}\it bool\_bit\rm (\bf bool\rm ( \it carry \rm + \it uchar\_schar\rm (\it w1\rm ) \rm + \it uchar\_schar \rm (\it w2\rm ) $<$ \rm 0\rm )\rm )\hspace*{0.20in}\rm ,

\hspace*{0.60in}\it bool\_bit\rm (\bf bool\rm ( \it carry \rm + \it w1 \rm + \it w2 $>$ \it UCHAR\_MAX\rm )\rm )\rm ,

\hspace*{0.60in}\it bool\_bit\rm (\bf bool\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm ) \rm + \it half8UCHAR\rm ( \it w2\rm )  $\geq$  \it pow2\_4\rm )\rm )\rm ,

\hspace*{0.60in}\it bool\_bit\rm (\bf bool\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it pow2\_8\rm ) \rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )\rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it substract8UCHAR \rm : \rm (\it BIT  $\times$  \it UCHAR  $\times$  \it UCHAR\rm )  $\fun$  \rm (\it UCHAR  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm )  $\land$

\hspace*{0.20in}\it substract8UCHAR \rm =  $\lambda$  \rm (\it carry\rm ,\it w1\rm , \it w2\rm )\rm . \rm ( \it carry  $\in$  \it BIT  $\land$  \it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it UCHAR\hspace*{0.15in} $\mid$

\hspace*{0.40in}\rm ( \rm ( \rm ( \it carry \rm + \it w1 \rm - \it w2 \rm )  $\mod$  \it pow2\_8\rm )\hspace*{0.15in}\rm ,\hspace*{0.10in}

\hspace*{0.60in}\it bool\_bit\rm (\bf bool\rm ( \it carry \rm + \it uchar\_schar\rm (\it w1\rm ) \rm - \it uchar\_schar \rm (\it w2\rm ) $<$ \rm 0\rm )\rm )\hspace*{0.20in}\rm ,

\hspace*{0.60in}\it bool\_bit\rm (\bf bool\rm ( \it carry \rm + \it w1 \rm - \it w2 $>$ \it UCHAR\_MAX\rm )\rm )\rm ,

\hspace*{0.60in}\it bool\_bit\rm (\bf bool\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm ) \rm - \it half8UCHAR\rm ( \it w2\rm )  $\geq$  \it pow2\_4\rm )\rm )\rm ,

\hspace*{0.60in}\it bool\_bit\rm (\bf bool\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm - \it w2 \rm )  $\mod$  \it pow2\_8\rm ) \rm = \rm 0\rm )\rm ) \rm ) \rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it add16USHORT \rm : \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\fun$  \rm (\it USHORT \rm )  $\land$

\hspace*{0.20in}\it add16USHORT \rm =  $\lambda$  \rm (\it b1\rm , \it w1\rm , \it w2\rm )\rm . \rm (\it b1  $\in$  \it BIT  $\land$  \it w1  $\in$  \it USHORT  $\land$  \it w2  $\in$  \it USHORT\hspace*{0.10in} $\mid$ \rm (\it b1\rm +\it w1\rm +\it w2\rm )  $\mod$ \hspace*{0.10in}\it pow2\_16\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it add\_carryUSHORT \rm : \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it add\_carryUSHORT \rm =  $\lambda$  \rm (\it b1\rm ,\it w1\rm , \it w2\rm )\rm . \rm (\it b1  $\in$  \it BIT  $\land$  \it w1  $\in$  \it USHORT  $\land$  \it w2  $\in$  \it USHORT\hspace*{0.10in} $\mid$

\hspace*{0.40in}\it bool\_bit\rm (\bf bool\rm (\rm (\it b1\rm +\it w1\rm +\it w2\rm ) $>$ \it pow2\_16 \rm ) \rm )\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it add\_halfcarryUSHORT \rm : \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it add\_halfcarryUSHORT \rm =  $\lambda$  \rm (\it b1\rm ,\it w1\rm , \it w2\rm )\rm . \rm (\it b1  $\in$  \it BIT  $\land$  \it w1  $\in$  \it USHORT  $\land$  \it w2  $\in$  \it USHORT\hspace*{0.10in} $\mid$

\hspace*{0.40in}\it bool\_bit\rm (\bf bool\rm (\rm ( \it b1 \rm + \rm (\it w1\rm )  $\mod$  \it pow2\_12 \rm + \rm (\it w2\rm )  $\mod$  \it pow2\_12 \rm ) $>$\hspace*{0.10in}\it pow2\_12 \rm )\rm )\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it sub16USHORT \rm : \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\fun$  \rm (\it USHORT \rm )  $\land$

\hspace*{0.20in}\it sub16USHORT \rm =  $\lambda$  \rm (\it b1\rm , \it w1\rm , \it w2\rm )\rm . \rm (\it b1  $\in$  \it BIT  $\land$  \it w1  $\in$  \it USHORT  $\land$  \it w2  $\in$  \it USHORT\hspace*{0.10in} $\mid$ \rm (\it w1\rm -\it w2\rm -\it b1\rm )  $\mod$ \hspace*{0.10in}\it pow2\_16\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it sub\_carryUSHORT \rm : \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it sub\_carryUSHORT \rm =  $\lambda$  \rm (\it b1\rm ,\it w1\rm , \it w2\rm )\rm . \rm (\it b1  $\in$  \it BIT  $\land$  \it w1  $\in$  \it USHORT  $\land$  \it w2  $\in$  \it USHORT\hspace*{0.10in} $\mid$

\hspace*{0.40in}\it bool\_bit\rm (\bf bool\rm (\rm (\it w1\rm -\it w2\rm -\it b1\rm ) $>$ \it pow2\_16 \rm ) \rm )\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it sub\_halfcarryUSHORT \rm : \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it sub\_halfcarryUSHORT \rm =  $\lambda$  \rm (\it b1\rm ,\it w1\rm , \it w2\rm )\rm . \rm (\it b1  $\in$  \it BIT  $\land$  \it w1  $\in$  \it USHORT  $\land$  \it w2  $\in$  \it USHORT\hspace*{0.10in} $\mid$

\hspace*{0.40in}\it bool\_bit\rm (\bf bool\rm (\rm (\hspace*{0.10in}\rm (\it w1\rm )  $\mod$  \it pow2\_12 \rm - \rm (\it w2\rm )  $\mod$  \it pow2\_12 \rm ) \rm -\it b1 $>$\hspace*{0.10in}\it pow2\_12 \rm )\rm )\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it inc\_BV16  $\in$ \hspace*{0.10in}\it BV16\hspace*{0.10in} $\fun$  \it BV16  $\land$

\hspace*{0.20in}\it inc\_BV16 \rm =  $\lambda$  \rm (\it w1\rm )\rm . \rm (\it w1  $\in$  \it BV16  $\mid$  \it ushort\_bv16 \rm (\rm (\it bv16\_ushort\rm (\it w1\rm )\rm +\rm 1 \rm )  $\mod$  \it pow2\_16\rm )\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it dec\_BV16  $\in$ \hspace*{0.10in}\it BV16\hspace*{0.10in} $\fun$  \it BV16  $\land$

\hspace*{0.20in}\it dec\_BV16 \rm =  $\lambda$  \rm (\it w1\rm )\rm . \rm (\it w1  $\in$  \it BV16  $\mid$  \it ushort\_bv16 \rm ( \rm (\it abs\rm (\it bv16\_ushort\rm (\it w1\rm )\rm -\rm 1\rm )\rm )\hspace*{0.10in} $\mod$  \it pow2\_16 \rm )\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it update\_refresh\_reg\hspace*{0.15in} $\in$  \it BYTE  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it update\_refresh\_reg \rm =\hspace*{0.10in} $\lambda$  \rm ( \it v0 \rm ) \rm .\hspace*{0.10in}\rm ( \it v0  $\in$  \it BYTE  $\mid$  \it uchar\_byte \rm (\it pow2\_7 $\times$ \it v0 \rm (\rm 7 \rm ) \rm + \rm (\rm (\hspace*{0.10in}\it pow2\_6  $\times$  \it v0 \rm ( \rm 6 \rm ) \rm + \it pow2\_5  $\times$  \it v0 \rm ( \rm 5 \rm ) \rm + \it pow2\_4  $\times$  \it v0 \rm ( \rm 4 \rm ) \rm +

\hspace*{1.00in}\it pow2\_3  $\times$  \it v0 \rm ( \rm 3 \rm ) \rm +\hspace*{0.15in}\it pow2\_2  $\times$  \it v0 \rm ( \rm 2 \rm ) \rm + \it pow2\_1  $\times$  \it v0 \rm ( \rm 1 \rm ) \rm + \it v0 \rm ( \rm 0 \rm ) \rm )  $\mod$  \it pow2\_7 \rm ) \rm ) \rm )  $\land$ \hspace*{0.15in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it instruction\_next  $\in$  \it USHORT  $\fun$  \it USHORT  $\land$

\hspace*{0.20in}\it instruction\_next \rm =\hspace*{0.10in} $\lambda$  \rm (\it w1\rm )\rm . \rm (\it w1  $\in$  \it USHORT\hspace*{0.10in} $\mid$ \rm (\it w1\rm +\rm 1\rm )  $\mod$ \hspace*{0.10in}\it pow2\_16\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it instruction\_jump \rm :\hspace*{0.10in}\rm (\it USHORT  $\times$  \rm (\rm (\rm -\rm 1\rm 2\rm 6\rm ) $\upto$ \rm 1\rm 2\rm 9\rm ) \rm )  $\fun$  \it USHORT  $\land$

\hspace*{0.20in}\it instruction\_jump \rm =  $\lambda$  \rm ( \it p0 \rm , \it e0 \rm ) \rm . \rm ( \it p0  $\in$  \it USHORT  $\land$  \it e0 \rm : \rm (\rm (\rm -\rm 1\rm 2\rm 6\rm ) $\upto$ \rm 1\rm 2\rm 9\rm )  $\mid$  \rm (\it p0 \rm + \it e0\rm )  $\mod$ \hspace*{0.10in}\it pow2\_16 \rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it daa\_function \rm : \rm ( \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BYTE\rm )  $\fun$  \rm ( \it BYTE  $\times$  \it BIT  $\times$  \it BIT \rm )  $\land$

\hspace*{0.20in} $\forall$ \rm (\it zn\rm , \it c0 \rm , \it h0 \rm , \it value \rm ) \rm .\rm (\it zn  $\in$  \it BIT  $\land$  \it c0  $\in$  \it BIT  $\land$  \it h0  $\in$  \it BIT  $\land$  \it value  $\in$  \it BYTE  $\implies$

\hspace*{0.40in}\rm (\hspace*{0.10in}\rm (\hspace*{0.10in}\it zn \rm = \rm 0  $\land$  \it c0 \rm = \rm 0  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 9  $\land$  \it h0 \rm = \rm 0  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 9  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm =\rm ( \it value \rm , \rm 0 \rm ,\rm 0\hspace*{0.10in}\rm ) \rm )  $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 0  $\land$  \it c0 \rm = \rm 0  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 8  $\land$  \it h0 \rm = \rm 0  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 1\rm 0 $\upto$ \rm 1\rm 5  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 6 \rm )\rm ) \rm , \rm 0 \rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.15in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 6\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm )  $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 0  $\land$  \it c0 \rm = \rm 0  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 9  $\land$  \it h0 \rm = \rm 1  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 3  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 6 \rm )\rm ) \rm , \rm 0 \rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.15in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 6\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm ) $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 0  $\land$  \it c0 \rm = \rm 0  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 1\rm 0 $\upto$ \rm 1\rm 5  $\land$  \it h0 \rm = \rm 0  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 9  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 9\rm 6 \rm )\rm ) \rm , \rm 1\rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.15in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 9\rm 6\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm ) $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 0  $\land$  \it c0 \rm = \rm 0  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 9 $\upto$ \rm 1\rm 5  $\land$  \it h0 \rm = \rm 0  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 1\rm 0 $\upto$ \rm 1\rm 5  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 1\rm 0\rm 2 \rm )\rm ) \rm , \rm 1\rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.15in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 1\rm 0\rm 2\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm ) $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 0  $\land$  \it c0 \rm = \rm 0  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 1\rm 0 $\upto$ \rm 1\rm 5  $\land$  \it h0 \rm = \rm 1  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 3  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 1\rm 0\rm 2 \rm )\rm ) \rm , \rm 1\rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.15in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 1\rm 0\rm 2\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm ) $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 0  $\land$  \it c0 \rm = \rm 1  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 2  $\land$  \it h0 \rm = \rm 0  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 9  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 9\rm 6 \rm )\rm ) \rm , \rm 1\rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.15in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 9\rm 6\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm ) $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 0  $\land$  \it c0 \rm = \rm 1  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 2  $\land$  \it h0 \rm = \rm 0  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 1\rm 0 $\upto$ \rm 1\rm 5  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 1\rm 0\rm 2\rm )\rm ) \rm , \rm 1\rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.15in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 1\rm 0\rm 2\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm ) $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 0  $\land$  \it c0 \rm = \rm 1  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 3  $\land$  \it h0 \rm = \rm 1  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 3  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 1\rm 0\rm 2\rm )\rm ) \rm , \rm 1\rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.15in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 1\rm 0\rm 2\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm ) $\land$

\hspace*{0.60in}

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 1  $\land$  \it c0 \rm = \rm 0  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 9  $\land$  \it h0 \rm = \rm 0  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 9  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm (\hspace*{0.10in}\it value \rm , \rm 0 \rm , \rm 0 \rm ) \rm )  $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 1  $\land$  \it c0 \rm = \rm 0  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 8  $\land$  \it h0 \rm = \rm 1  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 6 $\upto$ \rm 1\rm 5  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 2\rm 5\rm 0\rm )\rm ) \rm , \rm 0\rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.15in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 2\rm 5\rm 0\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm ) $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 1  $\land$  \it c0 \rm = \rm 1  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 7 $\upto$ \rm 1\rm 5  $\land$  \it h0 \rm = \rm 0  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 0 $\upto$ \rm 9  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 1\rm 6\rm 0\rm )\rm ) \rm , \rm 1 \rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.15in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 1\rm 6\rm 0\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm ) $\land$

\hspace*{0.60in}\rm (\hspace*{0.10in}\it zn \rm = \rm 1  $\land$  \it c0 \rm = \rm 1  $\land$  \it get\_upper\_digit\rm (\it value\rm )  $\in$  \rm 6 $\upto$ \rm 7  $\land$  \it h0 \rm = \rm 1  $\land$  \it get\_lower\_digit\rm (\it value\rm )  $\in$  \rm 6 $\upto$ \rm 1\rm 5  $\implies$

\hspace*{0.80in}\it daa\_function\rm (\it zn\rm , \it c0 \rm , \it h0\hspace*{0.10in}\rm , \it value \rm )\rm = \rm ( \it uchar\_byte\rm (\it simple\_add8UCHAR\rm ( \it byte\_uchar\rm ( \it value\rm ) \rm , \rm 1\rm 5\rm 4\rm )\rm ) \rm , \rm 1\rm ,\hspace*{0.10in}\it bool\_bit\rm (\bf bool\rm (\hspace*{0.10in}\it half8UCHAR\rm (\it byte\_uchar\rm (\it value\rm )\rm ) \rm + \it half8UCHAR\rm (\rm 1\rm 5\rm 4\rm )  $\geq$  \it pow2\_4\rm )\rm ) \rm ) \rm )\hspace*{0.10in}\rm ) \rm )

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\bf ASSERTIONS

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it is\_zero16USHORT\rm )\rm = \it USHORT  $\land$

\hspace*{0.20in}\bf ran\rm (\it is\_zero16USHORT\rm )\rm = \it BIT\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it is\_negative\rm )\rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it is\_negative\rm )\rm = \it BIT\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it half8UCHAR\rm )\rm = \it UCHAR  $\land$

\hspace*{0.20in}\bf ran\rm (\it half8UCHAR\rm )\rm = \it UCHAR  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it inc\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it inc\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it dec\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it dec\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it simple\_add8UCHAR\rm ) \rm = \rm (\it UCHAR  $\times$  \it UCHAR\rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it simple\_add8UCHAR\rm ) \rm = \it UCHAR  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm ( \it add8UCHAR \rm ) \rm = \it BIT  $\times$  \it UCHAR  $\times$  \it UCHAR  $\land$

\hspace*{0.20in}\bf ran\rm ( \it add8UCHAR \rm )  $\in$   $\pow$ \rm (\it UCHAR  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm ( \it substract8UCHAR \rm ) \rm = \it BIT $\times$ \it UCHAR  $\times$  \it UCHAR  $\land$

\hspace*{0.20in}\bf ran\rm ( \it substract8UCHAR \rm )  $\in$   $\pow$ \rm (\it UCHAR  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm ( \it add16USHORT \rm ) \rm = \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\land$

\hspace*{0.20in}\bf ran\rm ( \it add16USHORT \rm ) \rm = \rm (\it USHORT \rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it add\_carryUSHORT\rm ) \rm = \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it add\_carryUSHORT\rm ) \rm = \it BIT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it add\_halfcarryUSHORT\rm ) \rm = \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it add\_halfcarryUSHORT\rm ) \rm = \it BIT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it sub16USHORT\rm ) \rm = \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it sub16USHORT\rm ) \rm = \it USHORT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it sub\_carryUSHORT\rm ) \rm = \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it sub\_carryUSHORT\rm ) \rm = \it BIT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it sub\_halfcarryUSHORT\rm ) \rm = \rm ( \it BIT  $\times$  \it USHORT  $\times$  \it USHORT \rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it sub\_halfcarryUSHORT\rm ) \rm = \it BIT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it inc\_BV16\rm ) \rm = \it BV16  $\land$

\hspace*{0.20in}\bf ran\rm (\it inc\_BV16\rm ) \rm = \it BV16  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it dec\_BV16\rm ) \rm = \it BV16  $\land$

\hspace*{0.20in}\bf ran\rm (\it dec\_BV16\rm ) \rm = \it BV16  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it update\_refresh\_reg\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it update\_refresh\_reg\rm ) \rm = \it BYTE  $\land$ \hspace*{0.15in}

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it instruction\_next\rm ) \rm = \it USHORT  $\land$

\hspace*{0.20in}\bf ran\rm (\it instruction\_next\rm ) \rm = \it USHORT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it instruction\_jump\rm ) \rm = \rm (\it USHORT  $\times$  \rm (\rm (\rm -\rm 1\rm 2\rm 6\rm ) $\upto$ \rm 1\rm 2\rm 9\rm ) \rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it instruction\_jump\rm ) \rm = \it USHORT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it daa\_function\rm ) \rm = \rm ( \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BYTE\hspace*{0.10in}\rm ) $\land$

\hspace*{0.20in}\bf ran\rm (\it daa\_function\rm )  $\subseteq$  \rm ( \it BYTE  $\times$  \it BIT  $\times$  \it BIT \rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it abs\rm ) \rm = \bf INT  $\land$  \bf ran\rm (\it abs\rm ) \rm = \bf INT  $\land$

\hspace*{0.20in}

\hspace*{0.20in} \hspace*{0.30in}

\hspace*{0.20in}\rm 0 \rm = \it schar\_sshort \rm ( \rm 0 \rm , \rm 0 \rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it pow2\_0 \rm = \rm 2$^{0}$  $\land$

\hspace*{0.20in}\it pow2\_1 \rm = \rm 2$^{1}$  $\land$

\hspace*{0.20in}\it pow2\_2 \rm = \rm 2$^{2}$  $\land$

\hspace*{0.20in}\it pow2\_3 \rm = \rm 2$^{3}$  $\land$

\hspace*{0.20in}\it pow2\_4 \rm = \rm 2$^{4}$  $\land$

\hspace*{0.20in}\it pow2\_5 \rm = \rm 2$^{5}$  $\land$

\hspace*{0.20in}\it pow2\_6 \rm = \rm 2$^{6}$  $\land$

\hspace*{0.20in}\it pow2\_7 \rm = \rm 2$^{7}$  $\land$

\hspace*{0.20in}\it pow2\_8 \rm = \rm 2$^{8}$  $\land$

\hspace*{0.20in}\it pow2\_9 \rm = \rm 2$^{9}$  $\land$ \hspace*{0.20in}

\hspace*{0.20in}\it pow2\_10 \rm = \rm 2$^{10}$  $\land$

\hspace*{0.20in}\it pow2\_11 \rm = \rm 2$^{11}$  $\land$

\hspace*{0.20in}\it pow2\_12 \rm = \rm 2$^{12}$  $\land$

\hspace*{0.20in}\it pow2\_13 \rm = \rm 2$^{13}$  $\land$

\hspace*{0.20in}\it pow2\_14 \rm = \rm 2$^{14}$  $\land$

\hspace*{0.20in}\it pow2\_15 \rm = \rm 2$^{15}$  $\land$

\hspace*{0.20in}\it pow2\_16 \rm = \rm 2$^{16}$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\bf END

\vspace*{4mm}
\newpage
\end{sloppypar}\end{document}

### 2009-04-20 developments/platforms/Z80/tex/BIT_DEFINITION.tex

\bf MACHINE

\hspace*{0.20in}\it BIT\_DEFINITION

\vspace*{4mm}
\bf CONSTANTS

\hspace*{0.20in}\it BIT\rm ,

\hspace*{0.20in}\it bit\_not\rm ,

\hspace*{0.20in}\it bit\_and\rm ,

\hspace*{0.20in}\it bit\_or\rm ,

\hspace*{0.20in}\it bit\_xor\rm ,

\hspace*{0.20in}\it bool\_bit

\vspace*{4mm}
\bf PROPERTIES

\hspace*{0.20in}\it BIT \rm = \rm 0 $\upto$ \rm 1

\hspace*{0.20in}

\hspace*{0.20in} $\land$  \it bit\_not  $\in$  \it BIT  $\fun$  \it BIT

\hspace*{0.20in} $\land$  \it bit\_not \rm =  $\lambda$ \rm (\it bb\rm )\rm .\rm (\it bb $\in$ \it BIT  $\mid$  \rm 1\rm -\it bb\rm )

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in} $\land$  \it bit\_and  $\in$  \it BIT  $\times$  \it BIT  $\fun$  \it BIT

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\rm (\it bit\_and\rm (\it b1\rm , \it b2\rm ) \rm = \rm 1\rm )  $\iff$  \rm (\it b1 \rm = \rm 1\rm )  $\land$  \rm (\it b2 \rm = \rm 1\rm )\rm )\rm )

\hspace*{0.20in} \hspace*{0.40in}

\hspace*{0.20in}

\hspace*{0.20in} $\land$  \it bit\_or  $\in$  \it BIT  $\times$  \it BIT  $\fun$  \it BIT

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\rm (\it bit\_or\rm (\it b1\rm , \it b2\rm ) \rm = \rm 1\rm )  $\iff$  \rm (\it b1 \rm = \rm 1\rm )  $\lor$  \rm (\it b2 \rm = \rm 1\rm )\rm )\rm )

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in} $\land$  \it bit\_xor  $\in$  \it BIT  $\times$  \it BIT  $\fun$  \it BIT

\hspace*{0.20in} $\land$   $\forall$  \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\rm (\it bit\_xor\rm (\it b1\rm , \it b2\rm ) \rm = \rm 1\rm )  $\iff$  \rm (\it b1  $\not =$  \it b2\rm )\rm )\rm )

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in} $\land$  \it bool\_bit  $\in$  \bf BOOL  $\fun$  \it BIT

\hspace*{0.20in} $\land$  \it bool\_bit \rm = \rm \{ \bf TRUE  $\mapsto$  \rm 1\rm , \bf FALSE  $\mapsto$  \rm 0 \rm \}

\vspace*{4mm}
\bf ASSERTIONS

\hspace*{0.40in}

\hspace*{0.20in} $\forall$  \rm (\it bb\rm )\rm . \rm (\it bb  $\in$  \it BIT  $\implies$  \it bit\_not\rm (\it bb\rm ) \rm = \rm 1\rm -\it bb\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \bf dom\rm (\it bit\_and\rm )\rm =\it BIT $\times$ \it BIT

\hspace*{0.20in} $\land$  \bf ran\rm (\it bit\_not\rm )\rm =\it BIT

\hspace*{0.20in} $\land$  \it bit\_not\rm (\rm 0\rm ) \rm = \rm 1

\hspace*{0.20in} $\land$  \it bit\_not\rm (\rm 1\rm ) \rm = \rm 0

\hspace*{0.20in} $\land$   $\forall$ \rm (\it bb\rm )\rm .\rm (\it bb  $\in$  \it BIT  $\implies$  \it bit\_not\rm (\it bit\_not\rm (\it bb\rm )\rm ) \rm = \it bb\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \bf dom\rm (\it bit\_and\rm )\rm =\it BIT $\times$ \it BIT

\hspace*{0.20in} $\land$  \bf ran\rm (\it bit\_and\rm )\rm =\it BIT

\hspace*{0.20in} $\land$  \it bit\_and\rm (\rm 0\rm ,\rm 0\rm ) \rm = \rm 0

\hspace*{0.20in} $\land$  \it bit\_and\rm (\rm 0\rm ,\rm 1\rm ) \rm = \rm 0

\hspace*{0.20in} $\land$  \it bit\_and\rm (\rm 1\rm ,\rm 0\rm ) \rm = \rm 0

\hspace*{0.20in} $\land$  \it bit\_and\rm (\rm 1\rm ,\rm 1\rm ) \rm = \rm 1

\hspace*{0.20in}

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm ,\it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\it bit\_and\rm (\it b1\rm , \it b2\rm ) \rm = \it bit\_and\rm (\it b2\rm ,\it b1\rm )\rm )\hspace*{0.10in}\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm ,\it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it bit\_and\rm (\it b1\rm , \it b2\rm )\rm =\rm 1  $\implies$  \it bit\_and\rm (\it b2\rm ,\it b1\rm )\rm =\rm 1 \rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm ,\it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it bit\_and\rm (\it b1\rm , \it b2\rm )\rm =\rm 0  $\implies$  \it bit\_and\rm (\it b2\rm ,\it b1\rm )\rm =\rm 0 \rm )

\hspace*{0.20in}

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm ,\it b2\rm ,\it b3\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it b3  $\in$  \it BIT  $\implies$  \rm (\it bit\_and\rm (\it b1\rm , \it bit\_and\rm (\it b2\rm ,\it b3\rm )\rm ) \rm = \it bit\_and\rm (\it bit\_and\rm (\it b1\rm ,\it b2\rm )\rm ,\it b3\rm )\rm )\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\implies$  \rm (\it bit\_and\rm (\it b1\rm , \rm 1\rm ) \rm = \it b1\rm )\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\implies$  \rm (\it bit\_and\rm (\it b1\rm , \rm 0\rm ) \rm = \rm 0\rm )\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \bf dom\rm (\it bit\_or\rm )\rm =\it BIT $\times$ \it BIT

\hspace*{0.20in} $\land$  \bf ran\rm (\it bit\_or\rm )\rm =\it BIT

\hspace*{0.20in} $\land$  \it bit\_or\rm (\rm 0\rm ,\rm 0\rm ) \rm = \rm 0

\hspace*{0.20in} $\land$  \it bit\_or\rm (\rm 0\rm ,\rm 1\rm ) \rm = \rm 1

\hspace*{0.20in} $\land$  \it bit\_or\rm (\rm 1\rm ,\rm 0\rm ) \rm = \rm 1

\hspace*{0.20in} $\land$  \it bit\_or\rm (\rm 1\rm ,\rm 1\rm ) \rm = \rm 1

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\it bit\_or\rm (\it b1\rm , \it b2\rm ) \rm = \it bit\_or\rm (\it b2\rm ,\it b1\rm )\rm )\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it bit\_or\rm (\it b1\rm , \it b2\rm )\rm =\rm 1  $\implies$  \it bit\_or\rm (\it b2\rm ,\it b1\rm )\rm =\rm 1\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it bit\_or\rm (\it b1\rm , \it b2\rm )\rm =\rm 0  $\implies$  \it bit\_or\rm (\it b2\rm ,\it b1\rm )\rm =\rm 0\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it bit\_or\rm (\it b1\rm ,\rm 0\rm )\rm =\rm 1  $\implies$  \it b1\rm =\rm 1\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it bit\_or\rm (\it b1\rm ,\rm 0\rm )\rm =\rm 0  $\implies$  \it b1\rm =\rm 0\rm )

\hspace*{0.20in}

\hspace*{0.25in}

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm ,\it b2\rm ,\it b3\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it b3  $\in$  \it BIT  $\implies$  \rm (\it bit\_or\rm (\it b1\rm , \it bit\_or\rm (\it b2\rm ,\it b3\rm )\rm ) \rm = \it bit\_or\rm (\it bit\_or\rm (\it b1\rm ,\it b2\rm )\rm ,\it b3\rm )\rm )\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm ,\it b2\rm ,\it b3\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it b3  $\in$  \it BIT  $\land$  \it bit\_or\rm (\it b1\rm ,\it b2\rm )\rm =\rm 1  $\implies$  \it bit\_or\rm (\it b1\rm ,\it bit\_or\rm (\it b2\rm ,\it b3\rm )\rm )\rm =\rm 1\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm ,\it b2\rm ,\it b3\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it b3  $\in$  \it BIT  $\land$  \it bit\_or\rm (\it b1\rm ,\it b2\rm )\rm =\rm 1  $\implies$  \it bit\_or\rm (\it b1\rm , \it bit\_or\rm (\it b2\rm ,\it b3\rm )\rm ) \rm = \it bit\_or\rm (\rm 1\rm ,\it b3\rm )\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\implies$  \rm (\it bit\_or\rm (\it b1\rm , \rm 1\rm ) \rm = \rm 1\rm )\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\implies$  \rm (\it bit\_or\rm (\it b1\rm , \rm 0\rm ) \rm = \it b1\rm )\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\implies$  \rm (\it bit\_or\rm (\rm 1\rm ,\it b1\rm ) \rm = \rm 1\rm )\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\implies$  \rm (\it bit\_or\rm (\rm 0\rm ,\it b1\rm ) \rm = \it b1\rm )\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \bf dom\rm (\it bit\_xor\rm )\rm =\it BIT $\times$ \it BIT

\hspace*{0.20in} $\land$  \bf ran\rm (\it bit\_xor\rm )\rm =\it BIT\hspace*{0.10in}

\hspace*{0.20in} $\land$  \it bit\_xor\rm (\rm 0\rm ,\rm 0\rm ) \rm = \rm 0

\hspace*{0.20in} $\land$  \it bit\_xor\rm (\rm 0\rm ,\rm 1\rm ) \rm = \rm 1

\hspace*{0.20in} $\land$  \it bit\_xor\rm (\rm 1\rm ,\rm 0\rm ) \rm = \rm 1

\hspace*{0.20in} $\land$  \it bit\_xor\rm (\rm 1\rm ,\rm 1\rm ) \rm = \rm 0

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\implies$  \rm (\it bit\_xor\rm (\it b1\rm , \it b2\rm ) \rm = \it bit\_xor\rm (\it b2\rm ,\it b1\rm )\rm )\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2 $\in$ \it BIT  $\land$  \it bit\_xor\rm (\it b1\rm , \it b2\rm )\rm =\rm 1  $\implies$  \it bit\_xor\rm (\it b2\rm ,\it b1\rm )\rm =\rm 1\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm , \it b2\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2 $\in$ \it BIT  $\land$  \it bit\_xor\rm (\it b1\rm , \it b2\rm )\rm =\rm 0  $\implies$  \it bit\_xor\rm (\it b2\rm ,\it b1\rm )\rm =\rm 0\rm )

\hspace*{0.20in}

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm ,\it b2\rm ,\it b3\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it b3  $\in$  \it BIT  $\implies$  \rm (\it bit\_xor\rm (\it b1\rm , \it bit\_xor\rm (\it b2\rm ,\it b3\rm )\rm ) \rm = \it bit\_xor\rm (\it bit\_xor\rm (\it b1\rm ,\it b2\rm )\rm ,\it b3\rm )\rm )\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm ,\it b2\rm ,\it b3\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it b3  $\in$  \it BIT  $\land$  \it bit\_xor\rm (\it b1\rm ,\it b2\rm )\rm =\rm 1  $\implies$  \it bit\_xor\rm (\it b1\rm ,\it bit\_xor\rm (\it b2\rm ,\it b3\rm )\rm )\rm =\rm 1\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it b1\rm ,\it b2\rm ,\it b3\rm )\rm .\rm (\it b1  $\in$  \it BIT  $\land$  \it b2  $\in$  \it BIT  $\land$  \it b3  $\in$  \it BIT  $\land$  \it bit\_xor\rm (\it b1\rm ,\it b2\rm )\rm =\rm 1  $\implies$  \it bit\_xor\rm (\it b1\rm , \it bit\_xor\rm (\it b2\rm ,\it b3\rm )\rm ) \rm = \it bit\_xor\rm (\rm 1\rm ,\it b3\rm )\rm )

\hspace*{0.20in}

\hspace*{0.20in} $\land$   $\forall$ \rm (\it bb\rm )\rm .\rm (\it bb  $\in$  \it BIT  $\implies$  \it bit\_xor\rm (\it bb\rm ,\it bb\rm ) \rm = \rm 0\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \bf dom\rm (\it bool\_bit\rm )\rm =\bf BOOL

\hspace*{0.20in} $\land$  \bf ran\rm (\it bit\_xor\rm )\rm =\it BIT

\hspace*{0.20in} $\land$  \it bool\_bit\rm (\bf TRUE\rm ) \rm = \rm 1

\hspace*{0.20in} $\land$  \it bool\_bit\rm (\bf FALSE\rm ) \rm = \rm 0

\vspace*{4mm}
\hspace*{0.20in} $\land$   $\forall$ \rm (\it bb\rm )\rm .\rm (\it bb  $\in$  \it BIT  $\implies$  \it bb \rm = \rm 0  $\lor$  \it bb \rm = \rm 1\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it bb\rm )\rm .\rm (\it bb  $\in$  \it BIT  $\land$   $\neg$  \rm (\it bb \rm = \rm 0\rm )  $\implies$  \it bb \rm = \rm 1\rm )

\hspace*{0.20in} $\land$   $\forall$ \rm (\it bb\rm )\rm .\rm (\it bb  $\in$  \it BIT  $\land$   $\neg$  \rm (\it bb \rm = \rm 1\rm )  $\implies$  \it bb \rm = \rm 0\rm )

\vspace*{4mm}
\bf END\newpage
\end{sloppypar}\end{document}

### 2009-04-20 developments/platforms/Z80/tex/BV16.tex

\bf MACHINE

\hspace*{0.20in}\it BV16\_DEFINITION\hspace*{0.20in}

\bf SEES

\hspace*{0.20in}\it BIT\_DEFINITION\rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it BYTE\_DEFINITION

\hspace*{0.20in}

\bf DEFINITIONS\hspace*{0.20in}

\hspace*{0.20in}\it BV16\_WIDTH \rm == \rm 1\rm 6

\hspace*{0.20in}

\bf CONSTANTS

\hspace*{0.20in}\it BV16\_INDX\rm ,

\hspace*{0.20in}\it PHYS\_BV16\_INDEX\rm ,

\hspace*{0.20in}\it BV16\_ZERO\rm ,

\hspace*{0.20in}\it BV16\rm ,

\hspace*{0.20in}\it bv16\_byte\rm ,

\hspace*{0.20in}\it byte\_bv16\rm ,

\hspace*{0.20in}\it bv16\_bit\_get

\hspace*{0.20in}

\bf PROPERTIES

\vspace*{4mm}
\hspace*{0.20in}\it BV16\_INDX \rm = \rm 1 $\upto$ \rm (\it BV16\_WIDTH\rm )  $\land$

\hspace*{0.20in}\it PHYS\_BV16\_INDEX \rm = \rm 0 $\upto$ \rm (\it BV16\_WIDTH\rm -\rm 1\rm )  $\land$

\hspace*{0.20in}\it BV16 \rm = \rm \{ \it bt  $\mid$  \it bt $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\it BV16\_WIDTH\rm \}  $\land$

\hspace*{0.20in}\it BV16\_ZERO \rm = \it BV16\_INDX $\times$ \rm \{\rm 0\rm \}  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it bv16\_byte $\in$  \it BV16  $\fun$  \it BYTE  $\times$  \it BYTE  $\land$

\hspace*{0.20in}\it bv16\_byte \rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv $\in$ \it BV16  $\mid$  \rm (\rm \{ \rm 1 $\mapsto$ \it bv\rm (\rm 9\rm )\rm ,\rm 2 $\mapsto$ \it bv\rm (\rm 1\rm 0\rm )\rm ,\rm 3 $\mapsto$ \it bv\rm (\rm 1\rm 1\rm )\rm ,\rm 4 $\mapsto$ \it bv\rm (\rm 1\rm 2\rm )\rm ,\rm 5 $\mapsto$ \it bv\rm (\rm 1\rm 3\rm )\rm ,\rm 6 $\mapsto$ \it bv\rm (\rm 1\rm 4\rm )\rm ,\rm 7 $\mapsto$ \it bv\rm (\rm 1\rm 5\rm )\rm ,\rm 8 $\mapsto$ \it bv\rm (\rm 1\rm 6\rm )\rm \}

\hspace*{1.60in}\rm ,\rm \{ \rm 1 $\mapsto$ \it bv\rm (\rm 1\rm )\rm ,\rm 2 $\mapsto$ \it bv\rm (\rm 2\rm )\rm ,\rm 3 $\mapsto$ \it bv\rm (\rm 3\rm )\rm ,\rm 4 $\mapsto$ \it bv\rm (\rm 4\rm )\rm ,\rm 5 $\mapsto$ \it bv\rm (\rm 5\rm )\rm ,\rm 6 $\mapsto$ \it bv\rm (\rm 6\rm )\rm ,\rm 7 $\mapsto$ \it bv\rm (\rm 7\rm )\rm ,\rm 8 $\mapsto$ \it bv\rm (\rm 8\rm )\rm \}\rm ) \rm )  $\land$

\hspace*{0.20in}\it byte\_bv16  $\in$ \hspace*{0.10in}\it BYTE  $\times$  \it BYTE  $\fun$  \it BV16  $\land$ \hspace*{0.80in}

\hspace*{0.25in}\it byte\_bv16 \rm = \it bv16\_byte $^{-1}$   $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv16\_bit\_get\rm : \rm (\it BV16  $\times$  \it PHYS\_BV16\_INDEX \rm ) $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it bv16\_bit\_get \rm =  $\lambda$ \rm (\it bt\rm ,\it ind\rm )\rm .\rm (\it bt $\in$ \it BV16  $\land$  \it ind $\in$ \it PHYS\_BV16\_INDEX  $\mid$  \it bt\rm (\it ind\rm +\rm 1\rm ) \rm )\hspace*{0.20in}

\vspace*{4mm}
\bf ASSERTIONS

\hspace*{0.20in}

\hspace*{0.20in}\it BV16\_ZERO $\in$  \it BV16  $\land$

\hspace*{0.20in}\it BV16 $\subseteq$ \it BIT\_VECTOR  $\land$

\hspace*{0.20in}

\hspace*{0.20in} $\forall$ \rm (\it bt\rm )\rm .\rm (\it bt  $\in$  \it BV16  $\implies$  \it bv\_size\rm (\it bt\rm )\rm =\bf size\rm (\it bt\rm )  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\rm 1\rm 6  $\land$  \it bt  $\in$  \bf seq1\rm (\it BIT\rm )\hspace*{0.10in}\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\bf dom\rm (\it bv16\_byte\rm ) \rm = \it BV16  $\land$

\hspace*{0.20in}\bf ran\rm (\it bv16\_byte\rm ) \rm = \rm (\it BYTE  $\times$  \it BYTE\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\bf dom\rm (\it byte\_bv16\rm ) \rm = \rm (\it BYTE  $\times$  \it BYTE\rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}\bf ran\rm (\it byte\_bv16\rm ) \rm = \it BV16  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it bv16\_bit\_get\rm ) \rm = \rm (\it BV16  $\times$  \it PHYS\_BV16\_INDEX \rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it bv16\_bit\_get\rm ) \rm = \it BIT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\rm \rm [\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm \rm ]  $\in$  \it BV16  $\land$

\hspace*{0.20in}\rm \rm [\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm \rm ]  $\in$  \it BV16

\hspace*{0.20in}

\bf END

\newpage
\end{sloppypar}\end{document}

### 2009-04-20 developments/platforms/Z80/tex/BYTE.tex

\bf MACHINE

\hspace*{0.20in}\it BYTE\_DEFINITION

\hspace*{0.20in}

\bf SEES

\hspace*{0.20in}\it BIT\_DEFINITION\rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it POWER2

\hspace*{0.20in}

\bf DEFINITIONS

\hspace*{0.20in}\it BYTE\_WIDTH \rm == \rm 8 \rm ;

\hspace*{0.20in}

\hspace*{0.20in}\it pow2\_0 \rm == \rm 1\rm ;

\hspace*{0.20in}\it pow2\_1 \rm == \rm 2\rm ;

\hspace*{0.20in}\it pow2\_2 \rm == \rm 4\rm ;

\hspace*{0.20in}\it pow2\_3 \rm == \rm 8\rm ;

\hspace*{0.20in}\it pow2\_4 \rm == \rm 1\rm 6\rm ;

\hspace*{0.20in}\it pow2\_5 \rm == \rm 3\rm 2\rm ;

\hspace*{0.20in}\it pow2\_6 \rm == \rm 6\rm 4\rm ;

\hspace*{0.20in}\it pow2\_7 \rm == \rm 1\rm 2\rm 8\rm ;

\hspace*{0.20in}\it pow2\_8 \rm == \rm 2\rm 5\rm 6\rm ;

\hspace*{0.20in}\it pow2\_9 \rm == \rm 5\rm 1\rm 2\rm ;

\hspace*{0.20in}\it pow2\_10 \rm == \rm 1\rm 0\rm 2\rm 4\rm ;\hspace*{0.20in}

\hspace*{0.20in}\it pow2\_11 \rm == \rm 2\rm 0\rm 4\rm 8\rm ;\hspace*{0.10in}

\hspace*{0.20in}\it pow2\_12 \rm == \rm 4\rm 0\rm 9\rm 6\rm ;\hspace*{0.15in}

\hspace*{0.20in}\it pow2\_13 \rm == \rm 8\rm 1\rm 9\rm 2\rm ;

\hspace*{0.20in}\it pow2\_14 \rm == \rm 1\rm 6\rm 3\rm 8\rm 4\rm ;\hspace*{0.15in}

\hspace*{0.20in}\it pow2\_15 \rm == \rm 3\rm 2\rm 7\rm 6\rm 8\rm ;

\hspace*{0.20in}\it pow2\_16 \rm == \rm 6\rm 5\rm 5\rm 3\rm 6\hspace*{0.35in}

\hspace*{0.20in}

\bf CONSTANTS

\hspace*{0.20in}

\hspace*{0.20in}\it BYTE\rm ,

\hspace*{0.20in}\it BYTE\_INDEX\rm ,

\hspace*{0.20in}\it BYTE\_ZERO\rm ,

\hspace*{0.20in}\it PHYS\_BYTE\_INDEX\rm ,

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it is\_zero\rm ,

\hspace*{0.20in}\it parity\_even\rm ,

\hspace*{0.20in}\it and \rm , \it ior \rm , \it xor \rm ,

\hspace*{0.20in}\it bitclear \rm , \it bitset \rm , \it bitget \rm ,

\hspace*{0.20in}\it complement\rm , \it swap \rm , \it rotateleft \rm , \it rotateright\rm ,\hspace*{0.15in}

\hspace*{0.20in}\it get\_upper\_digit \rm ,

\hspace*{0.20in}\it get\_lower\_digit

\hspace*{0.20in}

\bf PROPERTIES

\hspace*{0.20in}

\hspace*{0.20in}\it BYTE\_INDEX \rm = \rm 1 $\upto$  \rm (\it BYTE\_WIDTH\rm )  $\land$

\hspace*{0.20in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \}\hspace*{0.10in} $\land$

\hspace*{0.20in}\it BYTE\_ZERO  $\in$  \it BYTE  $\land$

\hspace*{0.20in}\it BYTE\_ZERO \rm = \it BYTE\_INDEX  $\times$  \rm \{\rm 0\rm \}  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it is\_zero  $\in$  \it BYTE  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it is\_zero \rm =  $\lambda$ \rm (\it w1\rm )\rm .\rm (\it w1  $\in$  \it BYTE  $\mid$  \it bool\_bit\rm ( \bf bool\rm (\it w1\rm (\rm 1\rm ) \rm + \it w1\rm (\rm 2\rm )\rm + \it w1\rm (\rm 3\rm ) \rm + \it w1\rm (\rm 4\rm ) \rm + \it w1\rm (\rm 5\rm ) \rm + \it w1\rm (\rm 6\rm ) \rm + \it w1\rm (\rm 7\rm ) \rm + \it w1\rm (\rm 8\rm ) \rm = \rm 0\rm ) \rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it parity\_even  $\in$  \it BYTE  $\fun$  \it BIT\hspace*{0.20in} $\land$

\hspace*{0.20in}\it parity\_even \rm =  $\lambda$  \rm (\it bv\rm )\rm . \rm (\it bv  $\in$  \it BYTE  $\mid$  \rm 1\rm -\rm ( \it bv\rm (\rm 1\rm )\rm + \it bv\rm (\rm 2\rm )\rm + \it bv\rm (\rm 3\rm )\rm +\it bv\rm (\rm 4\rm )\rm +\it bv\rm (\rm 5\rm )\rm + \it bv\rm (\rm 6\rm )\rm +\it bv\rm (\rm 7\rm )\rm +\it bv\rm (\rm 8\rm )  $\mod$  \rm 2\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it and \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it and \rm =  $\lambda$ \rm (\it bt1\rm , \it bt2\rm )\rm .\rm (\it bt1 $\in$ \it BYTE  $\land$  \it bt2 $\in$ \it BYTE  $\mid$  \it bv\_and\rm (\it bt1\rm ,\it bt2\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it ior \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it ior \rm =  $\lambda$ \rm (\it bt1\rm , \it bt2\rm )\rm .\rm (\it bt1 $\in$ \it BYTE  $\land$  \it bt2 $\in$ \it BYTE  $\mid$  \it bv\_or\rm (\it bt1\rm ,\it bt2\rm )\rm )\hspace*{0.15in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it xor \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it xor \rm =  $\lambda$ \rm (\it bt1\rm , \it bt2\rm )\rm .\rm (\it bt1  $\in$  \it BYTE  $\land$  \it bt2  $\in$  \it BYTE  $\mid$  \it bv\_xor\rm (\it bt1\rm ,\it bt2\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it bitget \rm : \rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm )  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it bitget \rm =  $\lambda$ \rm (\it bt1\rm , \it ii\rm )\rm .\rm (\it bt1  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it PHYS\_BYTE\_INDEX  $\mid$  \it bt1\rm (\it ii\rm +\rm 1\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it bitset \rm : \rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it bitset \rm =  $\lambda$ \rm (\it bt1\rm , \it ii\rm )\rm .\rm (\it bt1  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it PHYS\_BYTE\_INDEX  $\mid$  \it bv\_set\rm (\it bt1\rm , \it ii\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it bitclear \rm : \rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in} $\forall$  \rm (\it ww\rm , \it ii\rm , \it bb\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it PHYS\_BYTE\_INDEX  $\land$  \it bb  $\in$  \it BIT  $\implies$  \it bitclear\rm (\it ww\rm ,\it ii\rm ) \rm = \it bv\_clear\rm (\it ww\rm , \it ii\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it complement  $\in$  \it BYTE  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it complement \rm =  $\lambda$ \rm (\it bt\rm )\rm .\rm (\it bt  $\in$  \it BYTE  $\mid$  \it bv\_not\rm (\it bt\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it swap  $\in$  \it BYTE  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it swap\rm = $\lambda$ \rm (\it bt\rm )\rm .\rm (\it bt $\in$ \it BYTE  $\mid$ \rm \{\rm 1 $\mapsto$ \it bt\rm (\rm 5\rm )\rm , \rm 2 $\mapsto$ \it bt\rm (\rm 6\rm )\rm , \rm 3 $\mapsto$ \it bt\rm (\rm 7\rm )\rm , \rm 4  $\mapsto$ \it bt\rm (\rm 8\rm )\rm , \rm 5 $\mapsto$ \it bt\rm (\rm 1\rm )\rm , \rm 6 $\mapsto$ \it bt\rm (\rm 2\rm )\rm , \rm 7 $\mapsto$ \it bt\rm (\rm 3\rm )\rm , \rm 8 $\mapsto$ \it bt\rm (\rm 4\rm )\rm \}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it rotateleft  $\in$  \it BYTE  $\fun$  \it BYTE\hspace*{0.10in} $\land$

\hspace*{0.20in}\it rotateleft\hspace*{0.10in}\rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv $\in$ \it BYTE $\mid$  \rm \{\rm 1 $\mapsto$ \it bv\rm (\rm 8\rm )\rm , \rm 2 $\mapsto$ \it bv\rm (\rm 1\rm )\rm , \rm 3 $\mapsto$ \it bv\rm (\rm 2\rm )\rm , \rm 4 $\mapsto$ \it bv\rm (\rm 3\rm )\rm , \rm 5 $\mapsto$ \it bv\rm (\rm 4\rm )\rm ,\rm 6 $\mapsto$ \it bv\rm (\rm 5\rm )\rm ,\rm 7 $\mapsto$ \it bv\rm (\rm 6\rm )\rm ,\rm 8 $\mapsto$ \it bv\rm (\rm 7\rm )\rm \} \rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE\hspace*{0.10in} $\land$

\hspace*{0.20in}\it rotateright\rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv $\in$ \it BYTE $\mid$ \rm \{\rm 1 $\mapsto$ \it bv\rm (\rm 2\rm )\rm , \rm 2 $\mapsto$ \it bv\rm (\rm 3\rm )\rm , \rm 3 $\mapsto$ \it bv\rm (\rm 4\rm )\rm , \rm 4 $\mapsto$ \it bv\rm (\rm 5\rm )\rm ,\rm 5 $\mapsto$ \it bv\rm (\rm 6\rm )\rm , \rm 6 $\mapsto$ \it bv\rm (\rm 7\rm )\rm , \rm 7 $\mapsto$ \it bv\rm (\rm 8\rm )\rm , \rm 8 $\mapsto$ \it bv\rm (\rm 1\rm )\rm \}\rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it get\_upper\_digit \rm = $\lambda$ \rm ( \it by \rm )\rm . \rm ( \it by  $\in$  \it BYTE  $\mid$ \hspace*{0.10in}\it pow2\_3  $\times$  \it by \rm (\rm 8 \rm ) \rm + \it pow2\_2  $\times$  \it by \rm ( \rm 7 \rm ) \rm + \rm 2  $\times$  \it by \rm ( \rm 6 \rm ) \rm +\hspace*{0.10in}\it by \rm ( \rm 5 \rm )\hspace*{0.15in}\rm )  $\land$

\hspace*{0.20in}\it get\_lower\_digit \rm = $\lambda$ \rm ( \it by \rm )\rm . \rm ( \it by  $\in$  \it BYTE  $\mid$ \hspace*{0.10in}\it pow2\_3  $\times$  \it by \rm (\rm 4 \rm ) \rm + \it pow2\_2  $\times$  \it by \rm ( \rm 3 \rm ) \rm + \rm 2  $\times$  \it by \rm ( \rm 2 \rm ) \rm +\hspace*{0.10in}\it by \rm ( \rm 1 \rm )\hspace*{0.15in}\rm )

\hspace*{0.20in}

\hspace*{0.20in}

\bf ASSERTIONS

\vspace*{4mm}
\hspace*{0.20in}\bf size\rm (\it BYTE\_ZERO\rm ) \rm = \rm 8  $\land$

\hspace*{0.20in}\it BYTE  $\subseteq$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it BYTE\_ZERO  $\in$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\bf first\rm (\it BYTE\_ZERO\rm )\rm =\rm 0  $\land$

\hspace*{0.20in}

\hspace*{0.20in} $\forall$ \rm (\it bt\rm )\rm .\rm (\it bt  $\in$  \it BYTE  $\implies$  \it bv\_size\rm (\it bt\rm )\rm =\bf size\rm (\it bt\rm )  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\rm 8  $\land$  \it bt  $\in$  \bf seq1\rm (\it BIT\rm )\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it BYTE\_WIDTH  $\in$   $\nat$   $\land$

\hspace*{0.40in}

\hspace*{0.20in}\bf dom\rm (\it is\_zero\rm )\rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it is\_zero\rm )\rm = \it BIT\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it parity\_even\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it parity\_even\rm ) \rm = \it BIT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it and\rm ) \rm = \it BYTE  $\times$  \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it and\rm ) \rm = \it BYTE\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it ior\rm ) \rm = \it BYTE  $\times$  \it BYTE  $\land$ \hspace*{0.20in}

\hspace*{0.20in}\bf ran\rm (\it ior\rm ) \rm = \it BYTE\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it xor\rm ) \rm = \it BYTE  $\times$  \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it xor\rm ) \rm = \it BYTE\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it bitget\rm )\rm =\rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it bitget\rm )\rm =\it BIT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it bitset\rm )\rm =\rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it bitset\rm )\rm = \it BYTE  $\land$ \hspace*{0.15in}

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it bitclear\rm )\rm =\rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm ) $\land$

\hspace*{0.20in}\bf ran\rm (\it bitclear\rm )\rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it complement\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it complement\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it swap\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it swap\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it rotateleft\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it rotateleft\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it rotateright\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it rotateright\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it get\_upper\_digit \rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it get\_upper\_digit\rm ) \rm = \rm 0 $\upto$ \rm (\it pow2\_4 \rm - \rm 1\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it get\_lower\_digit \rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it get\_lower\_digit\rm ) \rm = \rm 0 $\upto$ \rm (\it pow2\_4 \rm - \rm 1\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\rm \rm [\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm \rm ]  $\in$  \it BYTE  $\land$ \hspace*{0.10in}

\hspace*{0.20in}\rm \rm [\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm \rm ]  $\in$  \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it pow2\_0 \rm = \rm 2$^{0}$  $\land$

\hspace*{0.20in}\it pow2\_1 \rm = \rm 2$^{1}$  $\land$

\hspace*{0.20in}\it pow2\_2 \rm = \rm 2$^{2}$  $\land$

\hspace*{0.20in}\it pow2\_3 \rm = \rm 2$^{3}$  $\land$

\hspace*{0.20in}\it pow2\_4 \rm = \rm 2$^{4}$  $\land$

\hspace*{0.20in}\it pow2\_5 \rm = \rm 2$^{5}$  $\land$

\hspace*{0.20in}\it pow2\_6 \rm = \rm 2$^{6}$  $\land$

\hspace*{0.20in}\it pow2\_7 \rm = \rm 2$^{7}$  $\land$

\hspace*{0.20in}\it pow2\_8 \rm = \rm 2$^{8}$  $\land$

\hspace*{0.20in}\it pow2\_9 \rm = \rm 2$^{9}$  $\land$ \hspace*{0.20in}

\hspace*{0.20in}\it pow2\_10 \rm = \rm 2$^{10}$  $\land$

\hspace*{0.20in}\it pow2\_11 \rm = \rm 2$^{11}$  $\land$

\hspace*{0.20in}\it pow2\_12 \rm = \rm 2$^{12}$  $\land$

\hspace*{0.20in}\it pow2\_13 \rm = \rm 2$^{13}$  $\land$

\hspace*{0.20in}\it pow2\_14 \rm = \rm 2$^{14}$  $\land$

\hspace*{0.20in}\it pow2\_15 \rm = \rm 2$^{15}$  $\land$

\hspace*{0.20in}\it pow2\_16 \rm = \rm 2$^{16}$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\bf END

\newpage
\end{sloppypar}\end{document}

### 2009-04-20 developments/platforms/Z80/tex/MEMORY.tex

\bf MACHINE

\hspace*{0.20in}\it MEMORY

\hspace*{0.20in}

\bf SEES

\hspace*{0.20in}\it ALU\rm ,

\hspace*{0.20in}\it BIT\_DEFINITION \rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it BYTE\_DEFINITION\rm ,

\hspace*{0.20in}\it BV16\_DEFINITION\rm ,

\hspace*{0.20in}

\hspace*{0.20in}\it UCHAR\_DEFINITION\rm ,

\hspace*{0.20in}\it SCHAR\_DEFINITION\rm ,

\hspace*{0.20in}\it SSHORT\_DEFINITION\rm ,

\hspace*{0.20in}\it USHORT\_DEFINITION\rm ,

\hspace*{0.20in}\it POWER2

\bf CONSTANTS

\hspace*{0.20in}\it PROGRAM\_R\_ADR \rm ,

\hspace*{0.20in}\it DATA\_R\_ADR \rm ,

\hspace*{0.20in}\it STACK\_R\_ADR

\hspace*{0.20in}

\bf PROPERTIES

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it PROGRAM\_R\_ADR \rm = \it USHORT  $\land$

\hspace*{0.20in}\it DATA\_R\_ADR \rm = \it USHORT  $\land$

\hspace*{0.20in}\it STACK\_R\_ADR \rm = \it USHORT

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\bf ABSTRACT\_VARIABLES

\hspace*{0.20in}

\hspace*{0.20in}\it mem

\hspace*{0.20in}

\bf INVARIANT

\hspace*{0.20in}

\hspace*{0.20in}\it mem  $\in$  \it BV16  $\fun$  \it BYTE

\hspace*{0.20in}

\hspace*{0.20in}

\bf ASSERTIONS

\hspace*{0.20in}

\hspace*{0.20in}\bf ran\rm (\it mem\rm )  $\subseteq$  \it BYTE  $\land$

\hspace*{0.20in}\bf dom\rm (\it mem\rm ) \rm = \it BV16  $\land$

\hspace*{0.20in} $\forall$ \rm (\it address\rm ,\it value\rm )\rm .\rm ( \it address  $\in$  \it BV16  $\land$  \it value  $\in$  \it BYTE  $\implies$ \hspace*{0.20in}\it mem  $\lover$  \rm \{ \it address  $\mapsto$  \it value \rm \}  $\in$  \it BV16  $\fun$  \it BYTE \rm )

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\bf INITIALISATION

\hspace*{0.20in}\it mem  $:\in$  \it BV16  $\fun$  \it BYTE\hspace*{0.15in}

\hspace*{0.20in}

\bf OPERATIONS

\hspace*{0.20in}

\hspace*{0.20in}\bf updateMem\rm (\it NewValues\rm ) \rm =

\hspace*{0.20in}\bf PRE \it NewValues  $\in$  \it BV16  $\fun$  \it BYTE\hspace*{0.10in} $\land$   $\forall$ \rm (\it elem\rm )\rm .\rm ( \it elem  $\in$  \bf dom\rm (\it NewValues\rm )  $\implies$  \it bv16\_ushort\rm (\it elem\rm )  $\in$  \it PROGRAM\_R\_ADR\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it mem\rm := \it mem  $\lover$  \it NewValues

\hspace*{0.20in}\bf END\rm ;

\hspace*{0.20in}

\hspace*{0.20in}\bf updateAddressMem\rm (\it address\rm , \it value\rm ) \rm =

\hspace*{0.20in}\bf PRE\hspace*{0.15in}\it address  $\in$  \it BV16  $\land$  \it value  $\in$  \it BYTE  $\land$  \it bv16\_ushort\rm (\it address\rm )  $\in$  \it PROGRAM\_R\_ADR

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it mem\rm := \it mem  $\lover$  \rm \{ \it address  $\mapsto$  \it value \rm \}

\hspace*{0.20in}\bf END\rm ;

\hspace*{0.20in}

\hspace*{0.20in}\bf updateStack\rm (\it NewValues\rm ) \rm =

\hspace*{0.20in}\bf PRE \it NewValues  $\in$  \it BV16  $\fun$  \it BYTE\hspace*{0.10in} $\land$ \hspace*{0.10in} $\forall$ \rm (\it elem\rm )\rm .\rm ( \it elem  $\in$  \bf dom\rm (\it NewValues\rm )  $\implies$  \it bv16\_ushort\rm (\it elem\rm )  $\in$  \it STACK\_R\_ADR\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it mem\rm := \it mem  $\lover$  \it NewValues

\hspace*{0.20in}\bf END\rm ;

\hspace*{0.20in}

\hspace*{0.20in}\bf updateAddressStack\rm (\it address\rm , \it value\rm ) \rm =

\hspace*{0.20in}\bf PRE \it address  $\in$  \it BV16  $\land$  \it value  $\in$  \it BYTE  $\land$ \hspace*{0.10in}\it bv16\_ushort\rm (\it address\rm )  $\in$  \it STACK\_R\_ADR\hspace*{0.15in}

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it mem\rm := \it mem  $\lover$  \rm \{ \it address  $\mapsto$  \it value \rm \}

\hspace*{0.20in}\bf END

\hspace*{0.20in}

\bf END\newpage
\end{sloppypar}\end{document}

### 2009-04-20 trunk/developments/platforms/Z80/tex/BIT_VECTOR.tex

\bf MACHINE

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION

\vspace*{4mm}
\bf SEES

\hspace*{0.20in}\it BIT\_DEFINITION

\vspace*{4mm}
\bf CONSTANTS

\hspace*{0.15in}

\hspace*{0.20in}\it BIT\_VECTOR\rm ,

\hspace*{0.20in}\it bv\_catenate\rm ,

\hspace*{0.20in}\it bv\_sub\rm ,

\hspace*{0.20in}\it bv\_zero\rm ,

\hspace*{0.20in}\it bv\_one\rm ,

\hspace*{0.20in}\it bv\_size\rm ,

\hspace*{0.20in}\it bv\_not\rm ,\hspace*{0.20in}

\hspace*{0.20in}\it bv\_and\rm ,\hspace*{0.20in}

\hspace*{0.20in}\it bv\_or\rm ,

\hspace*{0.20in}\it bv\_xor\rm ,

\hspace*{0.20in}\it bv\_get\rm ,

\hspace*{0.20in}\it bv\_set\rm ,

\hspace*{0.20in}\it bv\_clear\rm ,

\hspace*{0.20in}\it bv\_put\rm ,

\hspace*{0.20in}\it phys\_bv\_index

\vspace*{4mm}
\bf PROPERTIES

\hspace*{0.20in}\it BIT\_VECTOR \rm = \bf seq1\rm (\it BIT\rm )

\hspace*{0.15in}

\hspace*{0.20in} $\land$  \it bv\_size  $\in$  \it BIT\_VECTOR  $\fun$   $ \na1 $

\hspace*{0.20in} $\land$  \it bv\_size \rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv  $\in$  \it BIT\_VECTOR  $\mid$  \bf size\rm (\it bv\rm )\rm )

\hspace*{0.20in}

\hspace*{0.20in} $\land$  \it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_catenate \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm ) \rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2 $\in$  \it BIT\_VECTOR  $\mid$  \it v1 $\cat$ \it v2\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it phys\_bv\_index  $\in$  \it BIT\_VECTOR  $\fun$   $\pow$ \rm ( $\nat$ \rm )

\hspace*{0.20in} $\land$  \it phys\_bv\_index \rm =  $\lambda$ \rm (\it v1\rm ) \rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\mid$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )\rm )

\hspace*{0.15in}

\hspace*{0.20in} $\land$  \it bv\_sub  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\times$   $\nat$   $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_sub \rm =  $\lambda$ \rm (\it bv\rm ,\it low\rm ,\it high\rm ) \rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it low $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\land$  \it high $\in$ \it low $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )

\hspace*{1.80in} $\mid$   $\lambda$ \rm (\it i0\rm )\rm .\rm (\it i0 $\in$  \rm 1 $\upto$ \rm (\it high\rm -\it low\rm +\rm 1\rm )  $\mid$  \it bv\rm (\it i0\rm +\it low\rm ) \rm )\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_zero  $\in$   $ \na1 $   $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_zero \rm =  $\lambda$ \rm (\it sz\rm ) \rm . \rm (\it sz  $\in$   $ \na1 $   $\mid$  \rm (\rm 1 $\upto$ \it sz\rm )  $\times$  \rm \{\rm 0\rm \}\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_one  $\in$   $ \na1 $   $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_one \rm =  $\lambda$ \rm (\it sz\rm ) \rm . \rm (\it sz  $\in$   $ \na1 $   $\mid$  \rm (\rm 1 $\upto$ \it sz\rm )  $\times$  \rm \{\rm 1\rm \}\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_not  $\in$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_not \rm =  $\lambda$ \rm (\it v1\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\mid$   $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \it bv\_size\rm (\it v1\rm ) $\mid$ \it bit\_not\rm (\it v1\rm (\it idx\rm ) \rm ) \rm ) \rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_and  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_and \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm ) $\mid$

\hspace*{0.80in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \it bv\_size\rm (\it v1\rm )  $\mid$  \it bit\_and\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_or  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_or \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\mid$

\hspace*{0.80in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \it bv\_size\rm (\it v1\rm )  $\mid$  \it bit\_or\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_xor  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_xor \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\mid$

\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$ \rm 1 $\upto$ \it bv\_size\rm (\it v1\rm )  $\mid$  \it bit\_xor\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )

\hspace*{0.20in}

\hspace*{0.20in} $\land$  \it bv\_clear  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_clear \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\mid$  \it v1  $\lover$  \rm \{ \rm (\it idx\rm +\rm 1\rm )  $\mapsto$  \rm 0 \rm \}\rm )

\hspace*{0.30in}

\hspace*{0.20in} $\land$  \it bv\_set  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_set \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\mid$  \it v1  $\lover$  \rm \{ \rm (\it idx\rm +\rm 1\rm )  $\mapsto$  \rm 1 \rm \}\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_get  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT

\hspace*{0.20in} $\land$  \it bv\_get \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\mid$  \it v1\rm (\it idx\rm +\rm 1\rm )\rm )

\vspace*{8mm}
\hspace*{0.20in} $\land$  \it bv\_put $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\times$ \it BIT $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_put \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm ,\it bit\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\land$  \it bit $\in$ \it BIT  $\mid$  \it v1  $\lover$  \rm \{ \rm (\it idx\rm +\rm 1\rm ) $\mapsto$  \it bit \rm \}\rm )

\hspace*{0.15in}

\hspace*{0.15in}

\vspace*{4mm}
\bf ASSERTIONS

\hspace*{0.20in} $\forall$  \it bv\rm .\rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \it bv\_size\rm (\it bv\_not\rm (\it bv\rm )\rm ) \rm = \it bv\_size\rm (\it bv\rm )\rm )\rm ;

\hspace*{0.55in}

\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm .\rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it indx  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\implies$  \rm (\it bv\_get\rm (\it bv\_not\rm (\it bv\_not\rm (\it bv\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\rm ,\it indx\rm ) \rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\implies$

\hspace*{1.00in}\it bv\_size\rm (\it bv\_catenate\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm ) \rm + \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it low\rm ,\it high\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it low  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\land$  \it high  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\land$  \it low  $\leq$  \it high  $\implies$

\hspace*{1.20in}\it bv\_size\rm (\it bv\_sub\rm (\it bv\rm , \it low\rm , \it high\rm )\rm ) \rm = \it high\rm -\it low\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{1.00in}\it bv\_size\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it indx\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR\hspace*{0.10in} $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it indx $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm ) $\implies$

\hspace*{1.20in}\rm (\it bv\_get\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_and\rm (\it v2\rm , \it v1\rm )\rm ,\it indx\rm )\rm )\rm )\rm ;

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it v3\rm ,\it indx\rm )\rm . \rm (\it v1 $\in$ \it BIT\_VECTOR  $\land$  \it v2 $\in$ \it BIT\_VECTOR  $\land$  \it v3 $\in$ \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm )\rm =\it bv\_size\rm (\it v2\rm )  $\land$  \it bv\_size\rm (\it v1\rm )\rm =\it bv\_size\rm (\it v3\rm )

\hspace*{1.40in} $\land$  \it indx $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm ) $\implies$  \rm (\it bv\_get\rm (\it bv\_and\rm (\it v1\rm , \it bv\_and\rm (\it v2\rm ,\it v3\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_and\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm , \it v3\rm )\rm ,\it indx\rm )\rm ) \rm )\rm ;

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\implies$  \rm (\it bv\_get\rm (\it bv\_and\rm (\it bv\rm , \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm ,\it indx\rm ) \rm =

\hspace*{1.00in}\it bv\_get\rm (\it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ,\it indx\rm ) \rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\implies$  \rm (\it bv\_get\rm (\it bv\_and\rm (\it bv\rm , \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\rm ,\it indx\rm )\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$  \it bv\_size\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm )\rm )\rm ;

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it indx\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\implies$

\hspace*{1.20in}\rm (\it bv\_get\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_or\rm (\it v2\rm , \it v1\rm )\rm ,\it indx\rm )\rm )\rm )\rm ;

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{1.00in}\it bv\_size\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;\hspace*{0.15in}

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it v3\rm ,\it indx\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR $\land$  \it v3  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v3\rm )

\hspace*{1.40in} $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm ) $\implies$  \rm (\it bv\_get\rm (\it bv\_or\rm (\it v1\rm , \it bv\_or\rm (\it v2\rm ,\it v3\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_or\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm , \it v3\rm )\rm ,\it indx\rm )\rm )\rm )\rm ;

\hspace*{0.15in}

\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm . \rm (\it bv $\in$ \it BIT\_VECTOR  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\implies$  \rm (\it bv\_get\rm (\it bv\_or\rm (\it bv\rm , \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ,\it indx\rm )\rm )\rm )\rm ;

\hspace*{0.20in}

\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm . \rm (\it bv $\in$ \it BIT\_VECTOR  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm ) $\implies$  \rm (\it bv\_get\rm (\it bv\_or\rm (\it bv\rm , \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\rm ,\it indx\rm )\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$ \it bv\_size\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm )\rm )\rm ;

\hspace*{0.20in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$ \it bv\_size\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\hspace*{0.20in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it indx\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\implies$

\hspace*{1.20in}\rm (\it bv\_get\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_xor\rm (\it v2\rm , \it v1\rm )\rm ,\it indx\rm )\rm )\rm )\rm ;

\hspace*{1.00in}

\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\implies$  \it bv\_get\rm (\it bv\_xor\rm (\it bv\rm ,\it bv\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ,\it indx\rm )\rm )

\bf END\newpage
\end{sloppypar}\end{document}

### 2009-04-20 trunk/developments/platforms/Z80/tex/BIT_VECTOR_ARI.tex

\bf MACHINE

\hspace*{0.20in}\it BIT\_VECTOR\_ARITHMETICS

\hspace*{0.20in}

\bf SEES

\hspace*{0.20in}\it BIT\_DEFINITION\rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it POWER2

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\bf CONSTANTS

\hspace*{0.20in}\it bv\_nat\rm ,

\hspace*{0.20in}\it bv\_even

\hspace*{0.20in}

\bf PROPERTIES

\hspace*{0.20in}\it bv\_nat  $\in$  \it BIT\_VECTOR  $\fun$   $\nat$   $\land$

\hspace*{0.20in}\it bv\_nat \rm =  $\lambda$  \rm (\it bv\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\mid$  \rm (  $\sum$  \it idx \rm . \rm (\it idx  $\in$  \bf dom\rm (\it bv\rm )  $\mid$  \rm (\rm 2$^{idx}$\rm )  $\times$  \it bv\rm (\it idx\rm )\rm )\rm )\rm ) $\land$

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it bv\_even $\in$ \it BIT\_VECTOR  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it bv\_even\rm = $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv $\in$ \it BIT\_VECTOR $\mid$  \rm 1\rm - \rm (\bf size\rm (\it bv $\rres$ \rm \{\rm 1\rm \}\rm )  $\mod$  \rm 2\rm ) \rm )

\hspace*{0.20in}

\hspace*{0.20in}

\bf ASSERTIONS

\hspace*{0.20in} $\forall$  \it ss\rm . \rm (\it ss  $\in$   $ \na1 $   $\implies$  \rm (\it bv\_nat\rm (\it bv\_zero\rm (\it ss\rm )\rm ) \rm = \rm 0\rm )\rm )  $\land$

\hspace*{0.20in} $\forall$  \it ss\rm . \rm (\it ss  $\in$   $ \na1 $   $\implies$  \rm (\it bv\_even\rm (\it bv\_zero\rm (\it ss\rm )\rm ) \rm = \rm 0\rm )\rm )

\hspace*{0.20in}

\bf END\newpage
\end{sloppypar}\end{document}

### 2009-04-20 papers/FM2009/source/paper.tex

--- 2009-04-20 trunk/papers/FM2009/source/paper.tex
@@
 
-\subsection{Modelling bytes}
+\subsection{Modelling bytes and bit vector of length 16}[I need review the english]
 
@@
 
-$
-\begin{array}{l}
-\mathit{BYTE\_WIDTH} = 8 \land \mathit{BYTE\_INDEX} = 0..(\mathit{BYTE\_WIDTH}-1) \land \\
-\mathit{BYTE} \subseteq \mathit{BIT\_VECTOR} \land \mathit{BYTE} = \{ v \in \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v) = \mathit{BYTE\_WIDTH}\} \land \\
-\mathit{BYTE\_ZERO} \in \mathit{BYTE} \land \mathit{BYTE} = \mathit{BYTE\_INDEX} \times \{0\} 
-\end{array}
-$
+% $
+% \begin{array}{l}
+% \mathit{BYTE\_WIDTH} = 8 \land \mathit{BYTE\_INDEX} = 0..(\mathit{BYTE\_WIDTH}-1) \land \\
+% \mathit{BYTE} \subseteq \mathit{BIT\_VECTOR} \land \mathit{BYTE} = \{ v \in \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v) = \mathit{BYTE\_WIDTH}\} \land \\
+% \mathit{BYTE\_ZERO} \in \mathit{BYTE} \land \mathit{BYTE} = \mathit{BYTE\_INDEX} \times \{0\}
+% \end{array}
+% $
+% Version 2
+
+\hspace*{0.0in}
+
+\hspace*{0.0in}\it BYTE\_WIDTH \rm = 8 $\land$ \it BYTE\_INDEX \rm = 1 $\upto$ \rm (\it BYTE\_WIDTH\rm )$\land$
+
+\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.10in} $\land$
+
+\hspace*{0.0in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \}\hspace*{0.10in} $\land$
+
+\hspace*{0.0in}\it BYTE\_ZERO  $\in$  \it BYTE  $\land$
+
+\hspace*{0.0in}\it BYTE\_ZERO \rm = \it BYTE\_INDEX  $\times$  \rm \{\rm 0\rm \}
+
+\hspace*{0.0in}
+
+The $\mathit{BYTE\_INDEX}$ is domain of byte modeled. it start at 1 to  obey a definition of sequences from B
+method. However, the functions from byte encapsulate the acces and these functions use the
+$\mathit{PHYS\_BYTE\_INDEX}$. The $\mathit{BYTE}$ type is a specializated type from $\mathit{BIT\_VECTOR}$, but it
+has a size limit. Some other more specific definitions are created to help the developer. For example the constant
+$\mathit{BYTE\_ZERO}$ and some below functions.
+
+\begin{itemize}
+  \item \it is\_zero $\in$ BYTE  $\fun$   BIT \rm - It returns 1 if the received byte is zero, otherwise
+  returns 0.
+
+  \item \it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE \rm - It receives a byte and returns its rotated to
+  right. There is a similar function named $\mathit{rotateleft}$.
+
+\end{itemize}
+
+By the very similar way, the type $\mathit{BV16}$ is created.
 
@@
 
-\subsection{Manipulation data functions from Z80} [I need review the english] There are some specific functions from
-Z80 to manipulate the data. The $\mathit{bv\_ireg\_plus\_d}$ is  used to indexed address. It received the
+\begin{property}[Assuring the absence of Overflow]
+ To assure that don´t happen overflow the developer can add this expression ($\mathit{bv\_get(rgs8(z0),0)} \neq 1
+ \land \mathit{bv\_get(rgs8(z0),2) \neq 1}$)in the invariant. By default, the overflow can happen, but many times it
+ can be dangerous. Then, the developer may prohibit its use. This restriction can also become the more dificult to
+ specify.
+\end{property}
+
+\subsection{Manipulation data functions from Z80} [I need to review the english] There are some specific functions
+from Z80 to manipulate the data. The $\mathit{bv\_ireg\_plus\_d}$ is  used to indexed address. It received the
 value of register ($\mathit{ix}$ or $\mathit{iy}$) and displacement to return the sum, the result is the memory address dislocated, see
@@
 \begin{itemize}
-  \item ALU \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and returns its decrement. There is a
+  \item \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and returns its decrement. There is a
   similar function named $\mathit{dec}$.
 
-  \item ALU \it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the pc actual
+  \item \it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the pc actual
   value and return its increment.
 
-  \item ALU \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
+  \item \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
   byte is zero, otherwise returns 0.
 
-  \item byte \it is\_zero $\in$ BYTE  $\fun$   BIT \rm - It returns 1 if the received byte is zero, otherwise
-  returns 0.
+  \item \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.
 
-  \item byte \it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE \rm - It receives a byte and returns its rotated to
-  right. There is a similar function named $\mathit{rotateleft}$.
+\end{itemize}
 
-  \item ALU \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.
-\end{itemize}
-Every logic functions defined in the $\mathit{BYTE}$ module are included in the $\mathit{ALU}$ module. Then, it was
-not necessary rewrite these functions.
+Every logic functions defined in the $\mathit{BYTE}$ and $\mathit{BV16}$ module are included in the $\mathit{ALU}$
+module. Then, it was not necessary rewrite these functions.
 
@@
 
-\subsubsection{I/O Ports} % Can be placed in basic definitions in a paragraph
-%Referencia o modeo de modelagem do captura de um evento externo de acordo com
-% o proposto em [DAVID_gdc_2007]
-
-[[ Flag registers and the associated functions: negative, half, zero.. ]]
-
 \section{Extern  actions and input and output ports}
@@
 
-\hspace*{0.0in}\bf IN\_r\_(C) \rm ( \it rr \rm ) \rm =
+\hspace*{0.0in}\bf IN\_r\_9C0 \rm ( \it rr \rm ) \rm =
 

### 2009-04-21 trunk/papers/FM2009/source/paper.tex
just some changes

@@
 \usepackage{amsfonts}
-
-%\usepackage{float}
 
@@
 %
-\section{Introduction about hardware models in B}
-[One page]
-This paper show a model in B from a microcontrooller.
-[EXPLAIN THE MOTIVATION]
- [When a error is found at project of plataform the PREJUIZOS are big ]
- [ In critical system, when a error apper in software it can make several PREJUIZOS  ]
+\section{Introduction}
+% Explica o contexto geral do artigo
+%  Porque modelar os microcontroladores.
+%    Porque podemos verificar várias propriedades em tempo de projeto, documentar, simulacao através das ferramentas de suporte do método B...
+%    Porque permite verificar software em nível de assemblagem.[citar trabalho de BHDL]
+%  Contextualizacao com grande desafio da verificao formal em nivel de assemblagem[citar grande challenge][citar proposta do ALEMAO do ProB] 
+%  Conclusão da introducao.
+%  Organizacao do paper
 
-[[Other microcontrollers are in building proccess: 8051, Arm..]]
-[Some smal details of the model are omitted in order to simplify the explanation.]
-[Smal intrudctio about B]
-[Structure from paper]
+%  Basear-se no Relatorio técnico e no ERMAC
 
@@
 
-\subsubsection{ Desing components B }
+\subsubsection{Desing components B }
 [half page]
 [Explicar como as coias são modeladas]
-
+% Use the tech report
 \section{General definitions for hardware models}
@@
 
-The module also provides lemmas on negation that may be useful for the
-users of the library to develop proofs:
+The module also provides lemmas on negation that may be useful for the users of the library to develop proofs:
 
@@
 \begin{itemize}
-  \item ALU \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and returns its decrement. There is a
+  \item \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and returns its decrement. There is a
   similar function named $\mathit{dec}$.
 
-  \item ALU \it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the pc actual
+  \item \it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the pc actual
   value and return its increment.
 
-  \item ALU \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
+  \item \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
   byte is zero, otherwise returns 0.
 
-  \item byte \it is\_zero $\in$ BYTE  $\fun$   BIT \rm - It returns 1 if the received byte is zero, otherwise
-  returns 0.
+  \item \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.
 
-  \item byte \it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE \rm - It receives a byte and returns its rotated to
-  right. There is a similar function named $\mathit{rotateleft}$.
+\end{itemize}
 
-  \item ALU \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.
-\end{itemize}
-Every logic functions defined in the $\mathit{BYTE}$ module are included in the $\mathit{ALU}$ module. Then, it was
-not necessary rewrite these functions.
+Every logic functions defined in the $\mathit{BYTE}$ and $\mathit{BV16}$ module are included in the $\mathit{ALU}$
+module. Then, it was not necessary rewrite these functions.
 
-\subsection{Interrupts}
+\subsubsection{Modelling the instructions} 
+[[Show and  LD n A]]
+
+\subsection{Modelling the Interrupts}
 
@@
 
-\subsubsection{I/O Ports} % Can be placed in basic definitions in a paragraph
-%Referencia o modeo de modelagem do captura de um evento externo de acordo com
-% o proposto em [DAVID_gdc_2007]
-
-[[ Flag registers and the associated functions: negative, half, zero.. ]]
-
 \section{Extern actions and input and output ports}
 
-The Z80 has an extensive set of input and output instruction and 256 ports for
+The Z80 has an extensive set of input and output (I/O) instruction and 256 ports for
 devices. This model can transfer data blocks and between the I/O devices and any
@@
 The $\mathit{IN\_r(C)}$ \footnote{The tools B don't allow use parenteses in identifiers, then the characteres
-\{``('',``)''\} are replaced respectively by \{``9'',``0''\} in the real specification.} is represented by the follow B
-operation. This instruction is from I/O group and update the flag registers.
+\{``('',``)''\} are replaced respectively by \{``9'',``0''\} in the real specification.} instruction is represented
+by the follow B operation. This instruction is from I/O group and update the flag registers.
 
-\hspace*{0.0in}\bf IN\_r\_(C) \rm ( \it rr \rm ) \rm =
+\hspace*{0.0in}\bf IN\_r\_9C0 \rm ( \it rr \rm ) \rm =
 
@@
 
-\subsubsection{Instruction set} 
-% Talk in general and put a example with midlle
-% complex
+ \section{Conlusitions}
 
- \section{Conlusitions}
+ [Basear-se no ERMAC]
 
@@
 
-\bibitem {Abrial}
-Abrial, J. R. The B Book: Assigning Programs to Meanings. Cambridge University Press, United States of
-America, 1 edition, 1996.
-
 % Paper about BHDL
@@
 
+\bibitem {Abrial}
+Abrial, J. R. The B Book: Assigning Programs to Meanings. Cambridge University Press, United States of
+America, 1 edition, 1996.
+
+\bibitem {CLEARSY}
+CLEARSY. Atelier B. http://www.atelierb.eu.
+
+\bibitem {Dantas_Semish08}
+ Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Proposta e avalia\c{c}\~{a}o de uma abordagem de desenvolvimento de
+ software fidedigno por constru\c{c}\~{a}o com o m\'{e}todo B. In: SEMISH, Bel\'{e}m, 2008. SBC.
+
+\bibitem {Dantas_Report08}
+Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL.Verified compilation based on the B method: initial
+appraisal.Technical Report, UFRN - DIMAP, 2008.
+
 \bibitem {Bell} 
@@
 
-\bibitem {Dantas}
- Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Proposta e avaliação de uma abordagem de desenvolvimento de
- software fidedigno por construção com o método B. In: SEMISH, Belém, 2008. SBC.
+\bibitem {HOARE}
+HOARE, C. A. R. The verifying compiler, a grand challenge for computing research. In: VMCAI, p. 78-78, 2005.
+
+\bibitem{Medeiros}
+ Val\'{e}rio Medeiros Jr., Stephenson Galv\~{a}o, David D\'{e}harbe. D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Modelagem
+ de Microcontroladores em B. In: ERMAC, Natal, 2008. SBMAC.
+
+\bibitem {WORDSWORTH}
+WORDSWORTH, J. B. Software Engineering with B. Addison Wesley, Boston, 1996.
 

### 2009-04-22 papers/FM2009/source/paper.tex

@@
 
-\usepackage{graphicx}
+\usepackage{graphicx,array}
 \usepackage{amsmath}
@@
 simulators, verify properties about the model and verify the software in assembly level. The verification in
-assembly is utility more important, because this allow the developing software faithful to the algorithm language.
-Finally, this paper show some relevant techniques used in models and to describe several details.
+assembly is the feature more important, because this allow the developing software faithful to the algorithm
+language. Finally, this paper show some relevant techniques used in models and to describe several details.
 \end{abstract}
 %
-\section{Introduction about hardware models in B}
-[One page]
-This paper show a model in B from a microcontrooller.
-[EXPLAIN THE MOTIVATION]
- [When a error is found at project of plataform the PREJUIZOS are big ]
- [ In critical system, when a error apper in software it can make several PREJUIZOS  ]
-
-[[Other microcontrollers are in building proccess: 8051, Arm..]]
-[Some smal details of the model are omitted in order to simplify the explanation.]
-[Smal intrudctio about B]
-[Structure from paper]
-
+\section{Introduction}
+% Explica o contexto geral do artigo
+%  Porque modelar os microcontroladores. \footnote{The interested reader in details from full specification or in
+% other study case is invited to visit the repository from authors at \url{http://b2asm.googlecode.com}.
+%    Porque podemos verificar várias propriedades em tempo de projeto, documentar, simulacao através das ferramentas de suporte do método B...
+%    Porque permite verificar software em nível de assemblagem.[citar trabalho de BHDL]
+%    Porque utilizamos o método B puro  - [citar proposta do ALEMAO do ProB]
+%  Contextualizacao com grande desafio da verificao formal em nivel de assemblagem[citar grande challenge] 
+%  Conclusão da introducao.
+%  Organizacao do paper
+%  Basear-se no Relatorio técnico e no ERMAC
 %
@@
 
-\subsubsection{ Desing components B }
+\subsubsection{Desing components B }
 [half page]
 [Explicar como as coias são modeladas]
+% Use the tech report
 
@@
 The main module includes a instance of memory module and accesses the definitions from: basic data types modules,
-$\mathit{ALU}$ and $\mathit{Power2}$ \footnote{The $\mathit{Power2}$ module has the basics definitions to help the
+\textit{ALU} and \textit{Power2} \footnote{The \textit{Power2} module has the basics definitions to help the
 theorems prover about calculus of power.}.
@@
 
-There are many functions defined in the module ALU. In general, these functions take basic definitions to build
-new specific functions. The function half8UCHAR is used to get the half part of UCHAR value. It is importante to
-know the half carry and it is used in the $\mathit{add8UCHAR}$.
+There are many functions defined in the module \textit{ALU}. In general, these functions take basic definitions to build
+new specific functions. The function $\mathit{half8UCHAR}$ is used to get the half part of $\mathit{UCHAR}$ value.
+It is importante to know the half carry and it is used in the $\mathit{add8UCHAR}$.
 
@@
 
-\hspace*{0.0in}(\it carry $\in$  \it BIT  $\land$  \it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it
-UCHAR $\mid$
+\hspace*{0.0in}(\it carry $\in$  \it BIT  $\land$  \it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it UCHAR
+$\mid$
 
@@
 
-\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it uchar\_schar\rm (\it w1\rm ) \rm + \it uchar\_schar \rm (\it w2\rm ) $<$ \rm 0\rm ),
+\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it uchar\_schar\rm (\it w1\rm ) \rm + \it uchar\_schar \rm (\it
+w2\rm ) $<$ \rm 0\rm ),
 
@@
 
-\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm ) \rm + \it half8UCHAR\rm ( \it w2\rm )  $\geq$  \it $2^{4}$\rm )\rm,
+\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm ) \rm + \it half8UCHAR\rm ( \it
+w2\rm )  $\geq$  \it $2^{4}$\rm )\rm,
 
-\hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm )\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )
+\hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm
+)\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )
 
@@
 $\mathit{SCHAR}$ type, they are respectively $\mathit{add8SCHAR}$ and $\mathit{substract8SCHAR}$, all these
-functions are of 8 bits ($\mathit{BYTE}$) and defined similarly. In the same way, the arithmetic functions for 16 bits
-($\mathit{BV16}$) are defined.
+functions are of 8 bits ($\mathit{BYTE}$) and defined similarly. In the same way, the arithmetic functions for 16
+bits ($\mathit{BV16}$) are defined.
 
@@
 
-Every logic functions defined in the $\mathit{BYTE}$ and $\mathit{BV16}$ module are included in the $\mathit{ALU}$
+Every logic functions are defined in the \textit{BYTE} and \textit{BV16} module are included in the \textit{ALU}
 module. Then, it was not necessary rewrite these functions.
 
-\subsection{Interrupts}
+\subsection{Modelling the instructions}[I need review the english]
 
-% Visão Geral
-In the Z80, the interrupts allows that the devices suspend a routine from CPU and start another service routine.
-This service routine can exchange data or signals between CPU and external devices. When a routine is finished, then
-the CPU come back to the last routine that was interrupted.
+The main module (\textit{Z80}) has two types of operations, a type represent the external actions and the other
+represent the instructions of microcontrollers. The external actions are shown in the section
+~\ref{sec:externalactions} and the other type represents the instructions of microcontroller. A simple example from
+instruction is a $\mathit{LD\_(nn)\_A}$ shown below. Many times to model a instruction is necessary to use the pre-defined functions,
+this facilitates the construction of model. This instruction use the $\mathit{updateAddressMem}$ function from
+\textit{Memory} module and it receives a address memory and its new memory value. After, it increments the program
+counter ($\mathit{pc}$) and update the refresh register ($\mathit{r\_}$).
+
+\hspace*{0.00in}\bf LD\_9nn0\_A \rm ( \it nn \rm ) \rm =
+
+\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} $\land$ \hspace*{0.10in}\it nn\hspace*{0.10in} $\in$  \it DATA\_R\_ADR
+
+\hspace*{0.20in}\bf THEN
+
+\hspace*{0.20in}\bf updateAddressMem \rm ( \it ushort\_bv16 \rm ( \it nn \rm ) \rm , \it rgs8 \rm ( \it a0 \rm )
+\rm )  $\para$
+
+\hspace*{0.20in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
+
+\hspace*{0.00in}\bf END\rm ;
+
+The other instructions have a similar structure.
+
+\subsection{Modelling the external action}
+\label{sec:externalactions}
+
+The external actions change the state of microcontroller and they aren´t instructions, for example, the refreshing
+of I/O port and the interruptions request. The extern action are also model by operations and they are prefixed
+with ``$ext\_$'' and after the name of action. There are just four extern actions: $ext\_update\_io\_ports$,
+$ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The $ext\_update\_io\_ports$ just change the state of I/O port, see.
+
+\hspace*{0.20in}\bf ext\_update\_io\_ports\rm (\it address\rm ,\it value\rm )\rm =
+
+\hspace*{0.20in}\bf PRE \it address  $\in$  \it UCHAR  $\land$ \hspace*{0.10in}\it value  $\in$  \it SCHAR \bf THEN
+
+\hspace*{0.40in}\it io\_ports \rm ( \it uchar\_byte \rm ( \it address \rm ) \rm ) \rm := \it schar\_byte \rm ( \it
+value \rm )
+
+\hspace*{0.20in}\bf END\rm
+
+The others are related to interrupts. The interrupts allows that the devices suspend a routine from CPU and start
+another service routine. This service routine can exchange data or signals between CPU and external devices. When a
+routine is finished, then the CPU come back to the last routine that was interrupted.
 
@@
 
-% $updateStack(\{ (sp\_minus\_two \fun low\_pc ),(sp\_minus\_one
-%    \fun high\_pc ) \}) ||\\  sp:= sp\_minus\_two || pc := 66H || iff1:=0 || 
-%    iff2:= iff1 $
 \begin{sloppypar}
@@
 
-\textbf{RESET}  - The reset is other external action that just reset the
-registers related to the interruptions.
+\textbf{RESET}  - This just reset the registers related to the interruptions.
 
@@
 
-\section{Extern  actions and input and output ports}
+\section{Input and output ports}
 
-The Z80 has an extensive set of input and output instruction and 256 ports for
+The Z80 has an extensive set of input and output (I/O) instruction and 256 ports for
 devices. This model can transfer data blocks and between the I/O devices and any
@@
 The $\mathit{IN\_r(C)}$ \footnote{The tools B don't allow use parenteses in identifiers, then the characteres
-\{``('',``)''\} are replaced respectively by \{``9'',``0''\} in the real specification.} is represented by the follow B
-operation. This instruction is from I/O group and update the flag registers.
+\{``('',``)''\} are replaced respectively by \{``9'',``0''\} in the real specification.} instruction is represented
+by the follow B operation. It is from I/O group, then it reveives a identifier of register
+$\mathit{r}$ and, in this place, it stores the value of $\mathit{C}$ port address. Besides, it increments the
+program counter and updates the flag registers.
 
@@
 
-\subsubsection{Instruction set} 
-% Talk in general and put a example with midlle
-% complex
+\section{Proofs}
+
+[O que as provas garantem?]
+As obrigações de provas geradas pela ferramentas garante a verficação dos tipos de dados e que as
+expressoes são bem formadas \footnote{An expression is called "well-defined" (or unambiguous) if its definition
+assigns it a unique interpretation or value.}. Deste modo, as propriedades citadas anteriomente também são
+verificadas. As propriedades citadas anteriormente oferecem garantias que
+
+[Enfatizar a vantagem]
+The traditional don´t allow verify this things and the properties is flexible. The flexbility allow add
+restrictions about tempo de execuçao and until specifics restrictions about the application of microcontroller.
+
+This flexibility of create propertis allow create conditions to verify if the software to finish in cert instant.
+
+[Estatisticas]
+Types Library
+
+Hardware Library
+
+[Groups]
+\begin{center}
+
+\begin{tabular}{llll}
+\textbf{Group} &  \textbf{Proofs No Obvios} & \textbf{Proofs WD} & × \\ 
+Input and Output & 54 & 296  \\ 
+Logic and Ari & 134 & 413  \\ 
+Manipulation of Bits & 160 & 742  \\ 
+Extern actions & 51 & 84  \\
+General & 140 & 615  \\
+Header & 68 & 169  \\
+ & \textbf{Subtotal} 607 & \textbf{Subtotal} 2319  &  \textbf{Total} 2926 \\
+\end{tabular}
+\end{center}
+
+[Conclusao]
+A primeira vez que realiza as provas é um pouco trabalhoso, porque as definições de tipos e de hardware não
+estavam completas. Então, um trabalho maior foi demandado, por outro lado, foi nece
+
+The first time
+Paralelismo vantagem
 
  \section{Conlusitions}
+
+[]
 
@@
 %  o provador para sua verificação.
-%  
-%  
-%  Algumas informações do manual do Z80 fornecido pela Zilog são imprecisas,
-%  o que compremete um pouco a qualidade do modelo, entretanto alguns pontos
-%  foram consultados no documento (Undocument Z80 documented-(Sean Young)) e
-%  dúvidas foram sanadas com o uso do simulador(Z80 IDE Simulator). 
-%  
+
 %  Para oferecer uma maior garantia do modelo com a implementação do processador
@@
 
-\paragraph{Notes and Comments.}
+\section{Acknowledges}
 
-% ---- Bibliography ----
-%
+This work received supports from CNPq and ANP (National Agency of Oil). Besides, the research Bruno Gomes help in
+the revision.
+
 \begin{thebibliography}{5}
+
+% Paper about BHDL
+
+% Paper of ALEMÃO PROB ,java virtual machine
+
+% O manual do Z80
 
@@
 
-% Paper about BHDL
+\bibitem {CLEARSY}
+CLEARSY. Atelier B. http://www.atelierb.eu.
 
-% Paper of ALEMÃO PROB ,java virtual machine
+\bibitem {Dantas_Semish08}
+ Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Proposta e avalia\c{c}\~{a}o de uma abordagem de desenvolvimento de
+ software fidedigno por constru\c{c}\~{a}o com o m\'{e}todo B. In: SEMISH, Bel\'{e}m, 2008. SBC.
 
-% Outros do artigo de bartira
-
-% O manual do Z80
+\bibitem {Dantas_Report08}
+Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL.Verified compilation based on the B method: initial
+appraisal.Technical Report, UFRN - DIMAP, 2008.
 
@@
 
-\bibitem {Dantas}
- Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Proposta e avaliação de uma abordagem de desenvolvimento de
- software fidedigno por construção com o método B. In: SEMISH, Belém, 2008. SBC.
+\bibitem {HOARE}
+HOARE, C. A. R. The verifying compiler, a grand challenge for computing research. In: VMCAI, p. 78-78, 2005.
+
+\bibitem{Medeiros}
+ Medeiros Jr, Val\'{e}rio; Galv\~{a}o, Stephenson; D\'{e}harbe, David; Modelagem de Microcontroladores em B. In: ERMAC, Natal, 2008. SBMAC.
+
+\bibitem {WORDSWORTH}
+WORDSWORTH, J. B. Software Engineering with B. Addison Wesley, Boston, 1996.
 

### 2009-04-24 papers/FM2009/source/paper.tex
Add "marketing" in the paper :)

@@
 %
+% - Citar o Atelierb
 
@@
 
-\section{Proofs}
-
-[O que as provas garantem?]
-As obrigações de provas geradas pela ferramentas garante a verficação dos tipos de dados e que as
-expressoes são bem formadas \footnote{An expression is called "well-defined" (or unambiguous) if its definition
-assigns it a unique interpretation or value.}. Deste modo, as propriedades citadas anteriomente também são
-verificadas. As propriedades citadas anteriormente oferecem garantias que
-
-[Enfatizar a vantagem]
-The traditional don´t allow verify this things and the properties is flexible. The flexbility allow add
-restrictions about tempo de execuçao and until specifics restrictions about the application of microcontroller.
-
-This flexibility of create propertis allow create conditions to verify if the software to finish in cert instant.
-
-[Estatisticas]
-Types Library
-
-Hardware Library
+\section{Proofs }
+
+[O que as provas garantem?] As obrigações de provas garante a verficação dos tipos de dados,
+das propriedades estabelecidas no modelo e se as expressões são bem formadas \footnote{An expression is called
+"well-defined" (or unambiguous) if its definition assigns it a unique interpretation or value.}. As propriedades
+oferecem garantias adicionais ao projetista do software, pois é capaz de estabelecer uma série de critérios de
+segurança. Além disso, como estas propriedades estão em nível de especificação, então é possível ajusta-las a
+necessidade de um determinado sofwtare sem implicar em mudanças no hardware, o que implica apenas em aumentar ou
+diminuir o número de provas.
+%O desenvolvedor poderia por exemplo aumentar a região de memória necessária para armazenar dados de trabalho.
+
+A seguir tem-se os número de provas não obvias e WD do projeto Z80 separado em grupos. Esta tabela mostra a
+distribuição do número de provas de cada grupo. Um dos grupos que contém mais provas é o de manipulação de bits. Apesar da grande
+quantidade de provas, este grupo foi rapidamente provado, pois ele utiliza vários lemmas providos pela Hardware
+Library, o que acontece também o grupo Logic and Arithmetic e a Types Library.
 
@@
 \begin{center}
-
 \begin{tabular}{llll}
-\textbf{Group} &  \textbf{Proofs No Obvios} & \textbf{Proofs WD} & × \\ 
+\textbf{Group} &  \textbf{Proofs No Obvious} & \textbf{Proofs WD} & × \\ 
 Input and Output & 54 & 296  \\ 
-Logic and Ari & 134 & 413  \\ 
+Logic and Arithmetic & 134 & 413  \\ 
 Manipulation of Bits & 160 & 742  \\ 
@@
 General & 140 & 615  \\
-Header & 68 & 169  \\
- & \textbf{Subtotal} 607 & \textbf{Subtotal} 2319  &  \textbf{Total} 2926 \\
+Properties and Assertions & 68 & 169  \\
+ \textbf{Total} &  607 &  2319  &  %\textbf{Total} 2926 \\
 \end{tabular}
@@
 
-[Conclusao]
-A primeira vez que realiza as provas é um pouco trabalhoso, porque as definições de tipos e de hardware não
-estavam completas. Então, um trabalho maior foi demandado, por outro lado, foi nece
-
-The first time
-Paralelismo vantagem
+[Conclusao] Este projeto tem um grande número provas(2926) e o processo de prova eh uma das etapas que consomem
+maior tempo. Para concluir o processo de prova de todo o microcontrolador foi demando um grande tempo,
+aproximadamente um mês. Isto foi custoso, porque foi necessário elaborar, retificar e provar várias lemmas dos
+projetos de base (Hardware and Types Library). Entretanto, com os projetos básicos mais maduros, torna-se bem mais
+rápida a modelagem e verificação de um novo microprocessador\footnote{Actualy, the microprocessors M8051 and
+PIC16F432 are under construction}.
+
+Com os projetos de base completamente provados, então aproximadamente 89\% das provas não obvias e 40\% da Proofs WD
+foram realizadas automaticamente. As demais provas foram realizadas após o uso de alguns passos de
+prova\footnote{Passos de prova são comandos que indicam um caminho para o provador concluir a prova, esses comandos
+apenas indicam um caminho, ou seja, ele não introduz hipósteses falsas}. Os passos de prova aceleraram bastante o
+processo de prova, pois como existem muitas instruções do microcontrolador com caractéristicas semelhantes e o
+atelierb é capaz de re-usar os passos de prova. Então, poucos passos de prova podem ser úteis para uma grande
+variedade de provas. Um exemplo é o conjunto de 17 passos de prova que rapidamente ajudou a verificar 99\% de provas
+WD. A experiência em adicionar passos de prova também ajudou a reduzir o tempo do processo de prova em mais de
+90\%. Isso aconteceu, porque o uso de passos de prova direciona melhor o provador ao fim da prova, então o processo
+de prova é mais rápido.
+
+Para reduzir ainda mais o custo de prova, um nova funcionalidade provida pelo atelierB foi utizada, a paralelizacao.
+O processo de prova used three computers each one with two cores, o que teoricamente aumentou em seis vezes a
+capacidade de processamento.
+
+Enfim, podemos afirmar que as técnicas citadas foram o que viabilizaram completamente o uso do método B.
+Por este ser um método muito rigoroso, então o seu processo de verificação é torna-se dificil. Por outro, lado as
+provas foram quebradas ( modularizadas )
 
@@
 
-[]
+[][Enfatizar a vantagem] The traditional don´t allow verify this things and the properties is flexible. The
+flexbility allow add restrictions about tempo de execuçao and until specifics restrictions about the application of
+microcontroller.
+
+This flexibility of create propertis allow create conditions to verify if the software to finish in cert instant.
 

### 2009-04-24 trunk/papers/FM2009/source/paper.tex

@@
 simulators, verify properties about the model and verify the software in assembly level. The verification in
-assembly is utility more important, because this allow the developing software faithful to the algorithm language.
-Finally, this paper show some relevant techniques used in models and to describe several details.
+assembly is the feature more important, because this allow the developing software faithful to the algorithm
+language. Finally, this paper show some relevant techniques used in models and to describe several details.
 \end{abstract}
@@
 % Explica o contexto geral do artigo
-%  Porque modelar os microcontroladores.
+%  Porque modelar os microcontroladores. \footnote{The interested reader in details from full specification or in
+% other study case is invited to visit the repository from authors at \url{http://b2asm.googlecode.com}.
 %    Porque podemos verificar várias propriedades em tempo de projeto, documentar, simulacao através das ferramentas de suporte do método B...
 %    Porque permite verificar software em nível de assemblagem.[citar trabalho de BHDL]
-%  Contextualizacao com grande desafio da verificao formal em nivel de assemblagem[citar grande challenge][citar proposta do ALEMAO do ProB] 
+%    Porque utilizamos o método B puro  - [citar proposta do ALEMAO do ProB]
+%  Contextualizacao com grande desafio da verificao formal em nivel de assemblagem[citar grande challenge] 
 %  Conclusão da introducao.
 %  Organizacao do paper
-
 %  Basear-se no Relatorio técnico e no ERMAC
-
-%
+%
+% - Citar o Atelierb
 
@@
 % Use the tech report
+
 \section{General definitions for hardware models}
@@
 
-\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm ) \rm + \it half8UCHAR\rm ( \it w2\rm )  $\geq$  \it $2^{4}$\rm )\rm,
-
-\hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm )\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )
+\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm ) \rm + \it half8UCHAR\rm ( \it
+w2\rm )  $\geq$  \it $2^{4}$\rm )\rm,
+
+\hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm
+)\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )
 
@@
 
-Every logic functions defined in the $\mathit{BYTE}$ and $\mathit{BV16}$ module are included in the $\mathit{ALU}$
+Every logic functions are defined in the \textit{BYTE} and \textit{BV16} module are included in the \textit{ALU}
 module. Then, it was not necessary rewrite these functions.
 
-\subsubsection{Modelling the instructions} 
-[[Show and  LD n A]]
-
-\subsection{Modelling the Interrupts}
-
-% Visão Geral
-In the Z80, the interrupts allows that the devices suspend a routine from CPU and start another service routine.
-This service routine can exchange data or signals between CPU and external devices. When a routine is finished, then
-the CPU come back to the last routine that was interrupted.
+\subsection{Modelling the instructions}[I need review the english]
+
+The main module (\textit{Z80}) has two types of operations, a type represent the external actions and the other
+represent the instructions of microcontrollers. The external actions are shown in the section
+~\ref{sec:externalactions} and the other type represents the instructions of microcontroller. A simple example from
+instruction is a $\mathit{LD\_(nn)\_A}$ shown below. Many times to model a instruction is necessary to use the pre-defined functions,
+this facilitates the construction of model. This instruction use the $\mathit{updateAddressMem}$ function from
+\textit{Memory} module and it receives a address memory and its new memory value. After, it increments the program
+counter ($\mathit{pc}$) and update the refresh register ($\mathit{r\_}$).
+
+\hspace*{0.00in}\bf LD\_9nn0\_A \rm ( \it nn \rm ) \rm =
+
+\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} $\land$ \hspace*{0.10in}\it nn\hspace*{0.10in} $\in$  \it DATA\_R\_ADR
+
+\hspace*{0.20in}\bf THEN
+
+\hspace*{0.20in}\bf updateAddressMem \rm ( \it ushort\_bv16 \rm ( \it nn \rm ) \rm , \it rgs8 \rm ( \it a0 \rm )
+\rm )  $\para$
+
+\hspace*{0.20in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
+
+\hspace*{0.00in}\bf END\rm ;
+
+The other instructions have a similar structure.
+
+\subsection{Modelling the external action}
+\label{sec:externalactions}
+
+The external actions change the state of microcontroller and they aren´t instructions, for example, the refreshing
+of I/O port and the interruptions request. The extern action are also model by operations and they are prefixed
+with ``$ext\_$'' and after the name of action. There are just four extern actions: $ext\_update\_io\_ports$,
+$ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The $ext\_update\_io\_ports$ just change the state of I/O port, see.
+
+\hspace*{0.20in}\bf ext\_update\_io\_ports\rm (\it address\rm ,\it value\rm )\rm =
+
+\hspace*{0.20in}\bf PRE \it address  $\in$  \it UCHAR  $\land$ \hspace*{0.10in}\it value  $\in$  \it SCHAR \bf THEN
+
+\hspace*{0.40in}\it io\_ports \rm ( \it uchar\_byte \rm ( \it address \rm ) \rm ) \rm := \it schar\_byte \rm ( \it
+value \rm )
+
+\hspace*{0.20in}\bf END\rm
+
+The others are related to interrupts. The interrupts allows that the devices suspend a routine from CPU and start
+another service routine. This service routine can exchange data or signals between CPU and external devices. When a
+routine is finished, then the CPU come back to the last routine that was interrupted.
 
@@
 
-\section{Extern actions and input and output ports}
+\section{Input and output ports}
 
@@
 \{``('',``)''\} are replaced respectively by \{``9'',``0''\} in the real specification.} instruction is represented
-by the follow B operation. This instruction is from I/O group and update the flag registers.
+by the follow B operation. It is from I/O group, then it reveives a identifier of register
+$\mathit{r}$ and, in this place, it stores the value of $\mathit{C}$ port address. Besides, it increments the
+program counter and updates the flag registers.
 
@@
 
+\section{Proofs }
+
+[O que as provas garantem?] As obrigações de provas garante a verficação dos tipos de dados,
+das propriedades estabelecidas no modelo e se as expressões são bem formadas \footnote{An expression is called
+"well-defined" (or unambiguous) if its definition assigns it a unique interpretation or value.}. As propriedades
+oferecem garantias adicionais ao projetista do software, pois é capaz de estabelecer uma série de critérios de
+segurança. Além disso, como estas propriedades estão em nível de especificação, então é possível ajusta-las a
+necessidade de um determinado sofwtare sem implicar em mudanças no hardware, o que implica apenas em aumentar ou
+diminuir o número de provas.
+%O desenvolvedor poderia por exemplo aumentar a região de memória necessária para armazenar dados de trabalho.
+
+A seguir tem-se os número de provas não obvias e WD do projeto Z80 separado em grupos. Esta tabela mostra a
+distribuição do número de provas de cada grupo. Um dos grupos que contém mais provas é o de manipulação de bits. Apesar da grande
+quantidade de provas, este grupo foi rapidamente provado, pois ele utiliza vários lemmas providos pela Hardware
+Library, o que acontece também o grupo Logic and Arithmetic e a Types Library.
+
+[Groups]
+\begin{center}
+\begin{tabular}{llll}
+\textbf{Group} &  \textbf{Proofs No Obvious} & \textbf{Proofs WD} & × \\ 
+Input and Output & 54 & 296  \\ 
+Logic and Arithmetic & 134 & 413  \\ 
+Manipulation of Bits & 160 & 742  \\ 
+Extern actions & 51 & 84  \\
+General & 140 & 615  \\
+Properties and Assertions & 68 & 169  \\
+ \textbf{Total} &  607 &  2319  &  %\textbf{Total} 2926 \\
+\end{tabular}
+\end{center}
+
+[Conclusao] Este projeto tem um grande número provas(2926) e o processo de prova eh uma das etapas que consomem
+maior tempo. Para concluir o processo de prova de todo o microcontrolador foi demando um grande tempo,
+aproximadamente um mês. Isto foi custoso, porque foi necessário elaborar, retificar e provar várias lemmas dos
+projetos de base (Hardware and Types Library). Entretanto, com os projetos básicos mais maduros, torna-se bem mais
+rápida a modelagem e verificação de um novo microprocessador\footnote{Actualy, the microprocessors M8051 and
+PIC16F432 are under construction}.
+
+Com os projetos de base completamente provados, então aproximadamente 89\% das provas não obvias e 40\% da Proofs WD
+foram realizadas automaticamente. As demais provas foram realizadas após o uso de alguns passos de
+prova\footnote{Passos de prova são comandos que indicam um caminho para o provador concluir a prova, esses comandos
+apenas indicam um caminho, ou seja, ele não introduz hipósteses falsas}. Os passos de prova aceleraram bastante o
+processo de prova, pois como existem muitas instruções do microcontrolador com caractéristicas semelhantes e o
+atelierb é capaz de re-usar os passos de prova. Então, poucos passos de prova podem ser úteis para uma grande
+variedade de provas. Um exemplo é o conjunto de 17 passos de prova que rapidamente ajudou a verificar 99\% de provas
+WD. A experiência em adicionar passos de prova também ajudou a reduzir o tempo do processo de prova em mais de
+90\%. Isso aconteceu, porque o uso de passos de prova direciona melhor o provador ao fim da prova, então o processo
+de prova é mais rápido.
+
+Para reduzir ainda mais o custo de prova, um nova funcionalidade provida pelo atelierB foi utizada, a paralelizacao.
+O processo de prova used three computers each one with two cores, o que teoricamente aumentou em seis vezes a
+capacidade de processamento.
+
+Enfim, podemos afirmar que as técnicas citadas foram o que viabilizaram completamente o uso do método B.
+Por este ser um método muito rigoroso, então o seu processo de verificação é torna-se dificil. Por outro, lado as
+provas foram quebradas ( modularizadas )
+
  \section{Conlusitions}
 
- [Basear-se no ERMAC]
+[][Enfatizar a vantagem] The traditional don´t allow verify this things and the properties is flexible. The
+flexbility allow add restrictions about tempo de execuçao and until specifics restrictions about the application of
+microcontroller.
+
+This flexibility of create propertis allow create conditions to verify if the software to finish in cert instant.
 
@@
 
-\paragraph{Notes and Comments.}
-
-% ---- Bibliography ----
-%
+\section{Acknowledges}
+
+This work received supports from ANP (National Agency of Oil) and CNPq. Besides, the researcher Bruno Gomes help in
+the revision.
+
 \begin{thebibliography}{5}
@@
 % Paper of ALEMÃO PROB ,java virtual machine
-
-% Outros do artigo de bartira
 
@@
 \bibitem{Medeiros}
- Val\'{e}rio Medeiros Jr., Stephenson Galv\~{a}o, David D\'{e}harbe. D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Modelagem
- de Microcontroladores em B. In: ERMAC, Natal, 2008. SBMAC.
+ Medeiros Jr, Val\'{e}rio; Galv\~{a}o, Stephenson; D\'{e}harbe, David; Modelagem de Microcontroladores em B. In: ERMAC, Natal, 2008. SBMAC.
 

### 2009-04-24 trunk/papers/FM2009/source/paper.tex
.

@@
 
-[O que as provas garantem?] As obrigações de provas garante a verficação dos tipos de dados,
-das propriedades estabelecidas no modelo e se as expressões são bem formadas \footnote{An expression is called
-"well-defined" (or unambiguous) if its definition assigns it a unique interpretation or value.}. As propriedades
-oferecem garantias adicionais ao projetista do software, pois é capaz de estabelecer uma série de critérios de
-segurança. Além disso, como estas propriedades estão em nível de especificação, então é possível ajusta-las a
-necessidade de um determinado sofwtare sem implicar em mudanças no hardware, o que implica apenas em aumentar ou
-diminuir o número de provas.
-%O desenvolvedor poderia por exemplo aumentar a região de memória necessária para armazenar dados de trabalho.
-
+[O que as provas garantem?] As obrigações de provas garante a verficação dos tipos de dados, das propriedades
+estabelecidas no modelo e se as expressões são bem formadas \footnote{An expression is called "well-defined" (or
+unambiguous) if its definition assigns it a unique interpretation or value.}. As propriedades oferecem garantias
+adicionais ao projetista do software, pois é capaz de estabelecer uma série de critérios de segurança. Além disso,
+como estas propriedades estão em nível de especificação, então é possível ajusta-las a necessidade de um determinado
+sofwtare sem implicar em mudanças no hardware, o que implica apenas em aumentar ou diminuir o número de provas.
+% O desenvolvedor poderia por exemplo aumentar a região de memória necessária para armazenar dados de trabalho.
 A seguir tem-se os número de provas não obvias e WD do projeto Z80 separado em grupos. Esta tabela mostra a
-distribuição do número de provas de cada grupo. Um dos grupos que contém mais provas é o de manipulação de bits. Apesar da grande
-quantidade de provas, este grupo foi rapidamente provado, pois ele utiliza vários lemmas providos pela Hardware
-Library, o que acontece também o grupo Logic and Arithmetic e a Types Library.
+distribuição do número de provas de cada grupo. Um dos grupos que contém mais provas é o de manipulação de bits.
+Apesar da grande quantidade de provas, este grupo foi rapidamente provado, pois ele utiliza vários lemmas providos
+pela Hardware Library, o que acontece também com o grupo Logic and Arithmetic e a Types Library.
 
@@
 
-[Conclusao] Este projeto tem um grande número provas(2926) e o processo de prova eh uma das etapas que consomem
+[Conclusao] Este projeto tem um grande número provas e o processo de prova eh uma das etapas que consomem
 maior tempo. Para concluir o processo de prova de todo o microcontrolador foi demando um grande tempo,
@@
 prova\footnote{Passos de prova são comandos que indicam um caminho para o provador concluir a prova, esses comandos
-apenas indicam um caminho, ou seja, ele não introduz hipósteses falsas}. Os passos de prova aceleraram bastante o
+apenas indicam um caminho, ou seja, ele não introduzem hipósteses falsas}. Os passos de prova aceleraram bastante o
 processo de prova, pois como existem muitas instruções do microcontrolador com caractéristicas semelhantes e o
@@
 Para reduzir ainda mais o custo de prova, um nova funcionalidade provida pelo atelierB foi utizada, a paralelizacao.
-O processo de prova used three computers each one with two cores, o que teoricamente aumentou em seis vezes a
+Para isso foram used three computers each one with two cores, o que teoricamente aumentou em seis vezes a
 capacidade de processamento.
 
-Enfim, podemos afirmar que as técnicas citadas foram o que viabilizaram completamente o uso do método B.
-Por este ser um método muito rigoroso, então o seu processo de verificação é torna-se dificil. Por outro, lado as
-provas foram quebradas ( modularizadas )
+A special care was taken to represent the model in way simple and accurate. This care also reduced the number of
+proofs. Pois, o simples fato de otimizar o uso do operador ($\lover$) diminuiu sensívelmente o número de
+obrigações de provas.
+
+Enfim, as técnicas citadas viabilizaram a verificaçao das 2926 obrigacoes de prova. A modularização dos projetos
+definiu lemmas básicos que ajudaram a construir um framework completo e simples. Os passos de prova direcionaram a
+resoluçao mais eficente das provas. A paralelizacao e cuidado na modelagem aceleraram as provas, porque aumentaram
+a capacidade de processameto e reduziram o número de obrigaçoes de provas.
 
@@
 
-[][Enfatizar a vantagem] The traditional don´t allow verify this things and the properties is flexible. The
+[intro]
+A necessidade comercial para produzir \textit{hardware} e \textit{software} de
+alta qualidade é sempre crescente. Métodos formais já têm demonstrado sucesso na
+especificação de \textit{softwares} comerciais e críticos; protocolos padrão e
+projetos de \textit{hardware}.
+
+O suporte fornecido pelo método B é bastante adequado, maduro, extensivo e eficaz
+no processo de especificação formal dessas plataformas. Isso vem sendo
+perceptível, pois até então, todas as características relevantes das plataformas
+foram modeladas de forma apropriada, principalmente as importantes para
+verificação em nível de montagem (\textit{assembly}). Por outro lado, uma das
+qualidades de destaque do método B é relacionada com a boa capacidade de
+verificação automática das ferramentas, pois, na versão estável das modelagens,
+do total de $503$ provas dos módulos principais do \textit{PIC}, \textit{8051} e
+\textit{Z80} apenas $3$ referentes ao \textit{Z80} foram provadas de forma
+semi-automática. Ou seja, desse total de provas menos de $1$\% necessitaram de
+uma simples intervenção do projetista para ser concluída.
+
+Adicionalmente, a aplicação das técnicas modularização e re-uso tornou o processo
+de especificação das plataformas mais padronizado e acelerado. Desse modo, é
+demonstrada a viabilidade dessas especificações, pois com a aquisição de
+maturidade das ferramentas e do projetista, o processo de modelagem ganhou
+agilidade e maior automação na fase de validação do projeto.
+
+Enfim, a modelagem B dessas plataformas possibilitarão a verificação em um dos
+mais altos níveis de certificação de \textit{software}. Não somente, elas podem
+ser importantes no projeto de plataformas para a construção dos simuladores,
+documentação e sua própria verificação formal, permitindo identificar várias
+inconsistências delas em tempo de projeto.
+
+[Enfatizar a vantagem]
+[Flexibilidade]
+The traditional don´t allow verify this things and the properties is flexible. The
 flexbility allow add restrictions about tempo de execuçao and until specifics restrictions about the application of

### 2009-04-25 trunk/papers/FM2009/source/paper.tex
trabalhando a conclusão

@@
 % - Citar o Atelierb
+
+[I need to Review this below text and to answer some questions]
+%  What the objectives from this models ?
+%  The objectives from this models are: supply a documentation that help the
+%  directly in the development and guarantees a series of properties. This
+%  properties are about hardware and software that run in this platform. The
+%  properties verification about software allow to verify software in the
+%  assembly level and It`s provide a high level of security, cause the mains
+%  levels of representation of software are verified, according to
+%  \cite{DANTAS08_SBMF}.
+
+% Why don´t we use a language based in events, how Event-B or CSP? 
+%  Because, we want to prover importants properties about the software and 	
+%  the B-method is more appropriate and don´t lost the capacity of verification
+%  about hardware level.
+
+% What are the advantages to use the B method ? The biggest advantage is the
+% formal verification on assembly level.
+
+% Once established the model of the platform, so I could check something on VHDL?
+% Or generate VHDL?
 
@@
 maior tempo. Para concluir o processo de prova de todo o microcontrolador foi demando um grande tempo,
-aproximadamente um mês. Isto foi custoso, porque foi necessário elaborar, retificar e provar várias lemmas dos
-projetos de base (Hardware and Types Library). Entretanto, com os projetos básicos mais maduros, torna-se bem mais
-rápida a modelagem e verificação de um novo microprocessador\footnote{Actualy, the microprocessors M8051 and
-PIC16F432 are under construction}.
-
-Com os projetos de base completamente provados, então aproximadamente 89\% das provas não obvias e 40\% da Proofs WD
-foram realizadas automaticamente. As demais provas foram realizadas após o uso de alguns passos de
-prova\footnote{Passos de prova são comandos que indicam um caminho para o provador concluir a prova, esses comandos
-apenas indicam um caminho, ou seja, ele não introduzem hipósteses falsas}. Os passos de prova aceleraram bastante o
-processo de prova, pois como existem muitas instruções do microcontrolador com caractéristicas semelhantes e o
-atelierb é capaz de re-usar os passos de prova. Então, poucos passos de prova podem ser úteis para uma grande
-variedade de provas. Um exemplo é o conjunto de 17 passos de prova que rapidamente ajudou a verificar 99\% de provas
-WD. A experiência em adicionar passos de prova também ajudou a reduzir o tempo do processo de prova em mais de
-90\%. Isso aconteceu, porque o uso de passos de prova direciona melhor o provador ao fim da prova, então o processo
-de prova é mais rápido.
-
-Para reduzir ainda mais o custo de prova, um nova funcionalidade provida pelo atelierB foi utizada, a paralelizacao.
-Para isso foram used three computers each one with two cores, o que teoricamente aumentou em seis vezes a
-capacidade de processamento.
+aproximadamente um mês. Isto foi custoso, porque foi necessário elaborar, retificar e provar várias lemmas
+dos projetos de base, Hardware and Types Library. Entretanto, com os projetos básicos mais maduros,
+torna-se bem mais rápida a modelagem e verificação de um novo microprocessador\footnote{Actualy, the
+microprocessors M8051 and PIC16F432 are under construction}.
+
+Com os projetos de base completamente provados, então aproximadamente 89\% das provas não obvias e 55\% da
+Proofs WD foram realizadas automaticamente. As demais provas foram realizadas após o uso de alguns passos
+de prova\footnote{Passos de prova são comandos que indicam um caminho para o provador concluir a prova,
+esses comandos apenas indicam um caminho, ou seja, ele não introduzem hipósteses falsas} associados a
+pattern matching of proof obligations. The user pass associeted to match pattern  aceleraram bastante o
+processo de prova, pois existem muitas instruções do microcontrolador com caractéristicas semelhantes e o
+atelierb é capaz de re-usar os passos de prova. Então, poucos passos de prova podem ser úteis para uma
+grande variedade de provas. Um exemplo é o conjunto de 17 passos de prova que rapidamente ajudou a
+verificar 99\% (2295) das provas WD. A experiência em adicionar passos de prova também ajudou a reduzir o
+tempo do processo de prova em mais de 90\%. Isso aconteceu, porque o uso de passos de prova direciona melhor o
+provador ao fim da prova.
+
+To further reduce cost of proof process, a new feature supported by AtelierB was used, the
+parallelization. To do this were used three computers each one with two cores, the that theoretically
+increased by six times the capacity of processing.
 
 A special care was taken to represent the model in way simple and accurate. This care also reduced the number of
-proofs. Pois, o simples fato de otimizar o uso do operador ($\lover$) diminuiu sensívelmente o número de
-obrigações de provas.
-
-Enfim, as técnicas citadas viabilizaram a verificaçao das 2926 obrigacoes de prova. A modularização dos projetos
-definiu lemmas básicos que ajudaram a construir um framework completo e simples. Os passos de prova direcionaram a
-resoluçao mais eficente das provas. A paralelizacao e cuidado na modelagem aceleraram as provas, porque aumentaram
-a capacidade de processameto e reduziram o número de obrigaçoes de provas.
+proofs. Because, the simple act of optimizing the use the simple act of optimizing the use of
+override operator ($\lover$) fell too the number of proof obligations.
+
+Finally, the techniques mentioned help the verification of all the 2926 proof obligations. The
+modularization of projects difined basic lemmas that help to build a simple and full framework. The
+automatic re-use of commands's proof directed to a resolution more eficiently of proofs. The
+paralelization and the care in modeling accelerate the process's proof, because increase the
+processing capacity and reduced the number of proof obligatiions.
 
  \section{Conlusitions}
-
-[intro]
-A necessidade comercial para produzir \textit{hardware} e \textit{software} de
-alta qualidade é sempre crescente. Métodos formais já têm demonstrado sucesso na
-especificação de \textit{softwares} comerciais e críticos; protocolos padrão e
-projetos de \textit{hardware}.
-
-O suporte fornecido pelo método B é bastante adequado, maduro, extensivo e eficaz
-no processo de especificação formal dessas plataformas. Isso vem sendo
-perceptível, pois até então, todas as características relevantes das plataformas
-foram modeladas de forma apropriada, principalmente as importantes para
-verificação em nível de montagem (\textit{assembly}). Por outro lado, uma das
-qualidades de destaque do método B é relacionada com a boa capacidade de
-verificação automática das ferramentas, pois, na versão estável das modelagens,
-do total de $503$ provas dos módulos principais do \textit{PIC}, \textit{8051} e
-\textit{Z80} apenas $3$ referentes ao \textit{Z80} foram provadas de forma
-semi-automática. Ou seja, desse total de provas menos de $1$\% necessitaram de
-uma simples intervenção do projetista para ser concluída.
+[I need review the english]
+% [intro]
+A necessidade comercial para produzir \textit{hardware} e \textit{software} de alta qualidade é sempre
+crescente. O métodos B já têm demonstrado sucesso na especificação de sistemas comerciais críticos.
+
+O suporte fornecido pelo método B é bastante adequado, maduro, extensivo e
+eficiente no processo de verificação dos modelos das plataformas. Isso vem sendo
+perceptível, pois todas as características relevantes das plataformas foram
+apropriadamente modeladas, principalmente as importantes para verificação em
+nível de montagem (\textit{assembly}).
+
+Por outro lado, uma das qualidades de destaque do método B é relacionada com a
+boa capacidade de verificação automática das ferramentas, pois, na versão estável
+das modelagens, do total de $503$ provas dos módulos principais do \textit{PIC},
+\textit{8051} e \textit{Z80} apenas $3$ referentes ao \textit{Z80} foram provadas
+de forma semi-automática. Ou seja, desse total de provas menos de $1$\%
+necessitaram de uma simples intervenção do projetista para ser concluída.
 

### 2009-04-25 papers/FM2009/source/paper.tex
Parcialment

@@
 
-[Conclusao] Este projeto tem um grande número provas(2926) e o processo de prova eh uma das etapas que consomem
-maior tempo. Para concluir o processo de prova de todo o microcontrolador foi demando um grande tempo,
-aproximadamente um mês. Isto foi custoso, porque foi necessário elaborar, retificar e provar várias lemmas dos
-projetos de base (Hardware and Types Library). Entretanto, com os projetos básicos mais maduros, torna-se bem mais
-rápida a modelagem e verificação de um novo microprocessador\footnote{Actualy, the microprocessors M8051 and
-PIC16F432 are under construction}.
-
-Com os projetos de base completamente provados, então aproximadamente 89\% das provas não obvias e 40\% da Proofs WD
-foram realizadas automaticamente. As demais provas foram realizadas após o uso de alguns passos de
-prova\footnote{Passos de prova são comandos que indicam um caminho para o provador concluir a prova, esses comandos
-apenas indicam um caminho, ou seja, ele não introduz hipósteses falsas}. Os passos de prova aceleraram bastante o
-processo de prova, pois como existem muitas instruções do microcontrolador com caractéristicas semelhantes e o
-atelierb é capaz de re-usar os passos de prova. Então, poucos passos de prova podem ser úteis para uma grande
-variedade de provas. Um exemplo é o conjunto de 17 passos de prova que rapidamente ajudou a verificar 99\% de provas
-WD. A experiência em adicionar passos de prova também ajudou a reduzir o tempo do processo de prova em mais de
-90\%. Isso aconteceu, porque o uso de passos de prova direciona melhor o provador ao fim da prova, então o processo
-de prova é mais rápido.
-
-Para reduzir ainda mais o custo de prova, um nova funcionalidade provida pelo atelierB foi utizada, a paralelizacao.
-O processo de prova used three computers each one with two cores, o que teoricamente aumentou em seis vezes a
-capacidade de processamento.
-
-Enfim, podemos afirmar que as técnicas citadas foram o que viabilizaram completamente o uso do método B.
-Por este ser um método muito rigoroso, então o seu processo de verificação é torna-se dificil. Por outro, lado as
-provas foram quebradas ( modularizadas )
+[Conclusao]This project has many proof obligations and its process of verification is the step of the most time
+consumed. The verification step of all the model spent about a month. This was hard, because was necessary to build,
+correct and prove several expressions from basic projects, Hardware and Types Library. However, when the basic
+projects are ready, the modelling and verification of new plataform \footnote{Actualy, the microprocessors M8051 and
+PIC16F432 are under construction} become more fast and easy. Thus, these basic projects were also built to re-use
+to other platforms.
+
+When the basic projects are ready, then about 89\% proof non obvious and 55\% of Proofs WD was conclused
+automaticly. The others were concluded through the use of commands ...
+Com os projetos de base completamente provados, então aproximadamente 89\% das provas não obvias e 55\% da
+Proofs WD foram realizadas automaticamente. As demais provas foram realizadas após o uso de alguns passos
+de prova\footnote{Passos de prova são comandos que indicam um caminho para o provador concluir a prova,
+esses comandos apenas indicam um caminho, ou seja, ele não introduzem hipósteses falsas} associados a
+pattern matching of proof obligations. The user pass associeted to match pattern  aceleraram bastante o
+processo de prova, pois existem muitas instruções do microcontrolador com caractéristicas semelhantes e o
+atelierb é capaz de re-usar os passos de prova. Então, poucos passos de prova podem ser úteis para uma
+grande variedade de provas. Um exemplo é o conjunto de 17 passos de prova que rapidamente ajudou a
+verificar 99\% (2295) das provas WD. A experiência em adicionar passos de prova também ajudou a reduzir o
+tempo do processo de prova em mais de 90\%. Isso aconteceu, porque o uso de passos de prova direciona melhor o
+provador ao fim da prova.
+
+To further reduce cost of proof process, a new feature supported by AtelierB was used, the
+parallelization. To do this were used three computers each one with two cores, the that theoretically
+increased by six times the capacity of processing.
+
+A special care was taken to represent the model in way simple and accurate. This care also reduced the number of
+proofs. Because, the simple act of optimizing the use the simple act of optimizing the use of
+override operator ($\lover$) fell too the number of proof obligations.
+
+Finally, the techniques mentioned help the verification of all the 2926 proof obligations. The
+modularization of projects difined basic lemmas that help to build a simple and full framework. The
+automatic re-use of commands's proof directed to a resolution more eficiently of proofs. The
+paralelization and the care in modeling accelerate the process's proof, because increase the
+processing capacity and reduced the number of proof obligatiions.
 
  \section{Conlusitions}
-
-[][Enfatizar a vantagem] The traditional don´t allow verify this things and the properties is flexible. The
+[I need review the english]
+% [intro]
+A necessidade comercial para produzir \textit{hardware} e \textit{software} de alta qualidade é sempre
+crescente. O métodos B já têm demonstrado sucesso na especificação de sistemas comerciais críticos.
+
+O suporte fornecido pelo método B é bastante adequado, maduro, extensivo e
+eficiente no processo de verificação dos modelos das plataformas. Isso vem sendo
+perceptível, pois todas as características relevantes das plataformas foram
+apropriadamente modeladas, principalmente as importantes para verificação em
+nível de montagem (\textit{assembly}).
+
+Por outro lado, uma das qualidades de destaque do método B é relacionada com a
+boa capacidade de verificação automática das ferramentas, pois, na versão estável
+das modelagens, do total de $503$ provas dos módulos principais do \textit{PIC},
+\textit{8051} e \textit{Z80} apenas $3$ referentes ao \textit{Z80} foram provadas
+de forma semi-automática. Ou seja, desse total de provas menos de $1$\%
+necessitaram de uma simples intervenção do projetista para ser concluída.
+
+Adicionalmente, a aplicação das técnicas modularização e re-uso tornou o processo
+de especificação das plataformas mais padronizado e acelerado. Desse modo, é
+demonstrada a viabilidade dessas especificações, pois com a aquisição de
+maturidade das ferramentas e do projetista, o processo de modelagem ganhou
+agilidade e maior automação na fase de validação do projeto.
+
+Enfim, a modelagem B dessas plataformas possibilitarão a verificação em um dos
+mais altos níveis de certificação de \textit{software}. Não somente, elas podem
+ser importantes no projeto de plataformas para a construção dos simuladores,
+documentação e sua própria verificação formal, permitindo identificar várias
+inconsistências delas em tempo de projeto.
+
+[Enfatizar a vantagem]
+[Flexibilidade]
+The traditional don´t allow verify this things and the properties is flexible. The
 flexbility allow add restrictions about tempo de execuçao and until specifics restrictions about the application of
@@
 
-This work received supports from CNPq and ANP (National Agency of Oil). Besides, the research Bruno Gomes help in
+This work received supports from ANP (National Agency of Oil) and CNPq. Besides, the researcher Bruno Gomes help in
 the revision.

### 2009-04-25 trunk/papers/FM2009/source/paper.tex
Proofs finished

@@
 
-[Conclusao] Este projeto tem um grande número provas e o processo de prova eh uma das etapas que consomem
-maior tempo. Para concluir o processo de prova de todo o microcontrolador foi demando um grande tempo,
-aproximadamente um mês. Isto foi custoso, porque foi necessário elaborar, retificar e provar várias lemmas
-dos projetos de base, Hardware and Types Library. Entretanto, com os projetos básicos mais maduros,
-torna-se bem mais rápida a modelagem e verificação de um novo microprocessador\footnote{Actualy, the
-microprocessors M8051 and PIC16F432 are under construction}.
-
-Com os projetos de base completamente provados, então aproximadamente 89\% das provas não obvias e 55\% da
-Proofs WD foram realizadas automaticamente. As demais provas foram realizadas após o uso de alguns passos
-de prova\footnote{Passos de prova são comandos que indicam um caminho para o provador concluir a prova,
-esses comandos apenas indicam um caminho, ou seja, ele não introduzem hipósteses falsas} associados a
-pattern matching of proof obligations. The user pass associeted to match pattern  aceleraram bastante o
-processo de prova, pois existem muitas instruções do microcontrolador com caractéristicas semelhantes e o
-atelierb é capaz de re-usar os passos de prova. Então, poucos passos de prova podem ser úteis para uma
-grande variedade de provas. Um exemplo é o conjunto de 17 passos de prova que rapidamente ajudou a
-verificar 99\% (2295) das provas WD. A experiência em adicionar passos de prova também ajudou a reduzir o
-tempo do processo de prova em mais de 90\%. Isso aconteceu, porque o uso de passos de prova direciona melhor o
-provador ao fim da prova.
-
-To further reduce cost of proof process, a new feature supported by AtelierB was used, the
-parallelization. To do this were used three computers each one with two cores, the that theoretically
-increased by six times the capacity of processing.
+% [Conclusao]
+This project has many proof obligations and its process of verification is the step of the most time consumed. The
+verification step of all the model spent about a month. This was hard, because was necessary to build, correct and
+prove several expressions from basic projects, Hardware and Types Library. However, when the basic projects are
+ready, the modelling and verification of new plataform \footnote{Actualy, the microprocessors M8051 and PIC16F432
+are under construction} become more fast and easy. Thus, these basic projects were also built to re-use to other
+platforms. When the basic projects are ready, then about 89\% proof non obvious and 55\% of Proofs WD was conclused
+automaticly.
+
+The others were concluded through the use of proof commands\footnote{The proof commands are step that just indicate
+a walk to prover make the proof, thus, theses commands cannot introduce false hypotheses.} associated to match
+pattern from proof obligations. The match pattern accelerates too much the prove process, because there are many
+similar assembler instructions and the AtelierB can re-use the match pattern and its proof commands associated.
+Therefore, few proof commands can be useful to many obrigações de prova. A good example is a set of 17 proof
+commands that fastly helps to verify 99\% (2295) of proofs WD. To further reduce cost of proof process, a new feature
+supported by AtelierB was used, the parallelization. To do this were used three computers each one with two cores,
+the that theoretically increased by six times the capacity of processing.
 
 A special care was taken to represent the model in way simple and accurate. This care also reduced the number of
-proofs. Because, the simple act of optimizing the use the simple act of optimizing the use of
-override operator ($\lover$) fell too the number of proof obligations.
+proofs. Because, the simple act of optimizing the use of override operator ($\lover$) fell too the number of proof
+obligations.
 

### 2009-04-27 trunk/papers/FM2009/source/paper.tex
Writing the section Related works

@@
 
-\author{Val\'{e}rio Medeiros Jr\inst{1}, David D\'{e}harbe\inst{1} , Stephenson Galv\~{a}o \inst{1}}
+\author{B. Dantas \inst{1}, David D\'{e}harbe\inst{1}, Val\'{e}rio Medeiros Jr\inst{1} }
 
@@
 %
+%\keywords{Fomal Methods, Microprocessor Verification, Hardware}
+
 \section{Introduction}
@@
 
+ \section{Related Works}
+
+% This work had as the main motivation the  verififcation in assembler level and consequently the
+% verification of harware model. The model created express the semantic of functioning of hardware and
+% with this model is possible verify the semantic of hardware components.
+
+There are in the literature of computer science some technics\cite{BHDL_2003,GEMPLUS_99} to model hardware
+and the virtual machines using the B method. Then, the B method has been used successfully to model the
+operational semantic, however The cost modeling was still too expensive and this paper describes some
+techniques to lower the cost modeling.
+
+In general, the researchers are concetrated at model the Java Virtual Machine (JVM), but it is not easy
+to specify and become more difficult to write good lemmas, to normalize the functions and to verify the
+properties. Then, this technique specify a thing more simple to get better results.
+
+This strategy allow modularize the project in more two levels (Hardware an Types)
+Como nosso projeto é modularizado e formado por dois projetos básicos, então possível geração de código
+para descrição de hardware é simples, pois trabalhamos no nível de descrições de hardware e tipos de
+dados e suas funções visam ter um baixo acoplamento e alta coesão.
+
+ [Java Bytecode 	]
+ - Gplus -Preferimos dar um passo mais curto, ao invés de partir diretamente a máquina virtual, porque
+ preferimos fundamentar bem os projetos básicos, para assim facilitar o processo de prova.
+ Além disso, ter resultados concretos mais rapidamente e ao mesmo tempo aproximar-se da modelagem do
+ processadores modernos atuais.
+
+ [Uma das grandes vantagens é a evolução suporte ao método B]
+
+ A other advantage is the use of news and advanceds funcionalities of B method. The first is
+ the paralelization of prove procces and  the second is instensive use of math patterns associated with proof commands.
+
+ [AMM]
+Verificação formal já vem sendo aplicada com sucesso na verificação \ldots 
+Porém os custos eram altos, pois se passou X tempo para modelar \ldots e
+
+A modelagem do microcontrolador foi construida por um unico designer and consumiu aproximadamente six
+month to study the plattaform Z80, model, adjust to simplifi and prove.
+
  \section{Conlusitions}
+
+ % Trab. Futuros
+ %- Refinar o modelo para gerar através dos  código e apartir deste código criar um
+ %interpretador/simulador
+ %- Construir compilador
+
+%  \begin{itemize}
+%    Citar ambuiguidades encontradas no manual e eventuais erros encontrados.
+%   \item Utilidade o processo de desenvolvimento do hardware:
+%   \subitem Ajuda na remoção de eventuais ambiguidades providas de uma descrição informal.
+%   \subitem Verificação de propriedades: sobre a manipulação e tipo dos dados.
+%   \subitem Documentação formalizada matemáticamente
+%   \subitem Possibilidade da geracao de código VHDL apartir do modelo B.
+%   \subitem Colaboracao para construção de um simulador/interpretador
+% \end{itemize}
+
 [I need review the english]
@@
 
+[ O que foi o diferencial ? qual grande vantagem]
+[Eficiência da técnica/método B]
 Por outro lado, uma das qualidades de destaque do método B é relacionada com a
 boa capacidade de verificação automática das ferramentas, pois, na versão estável
-das modelagens, do total de $503$ provas dos módulos principais do \textit{PIC},
+das modelagens, do total das $503$ provas dos módulos principais do \textit{PIC},
 \textit{8051} e \textit{Z80} apenas $3$ referentes ao \textit{Z80} foram provadas
@@
 
+[maior vantagem]
 Enfim, a modelagem B dessas plataformas possibilitarão a verificação em um dos
@@
 This flexibility of create propertis allow create conditions to verify if the software to finish in cert instant.
+
+No contexto da verificação em nível de assembly:
+Exemplificar com propriedade de restrição temporal.
+Exp. é possível verificar se o software satisfazer restrições temporais e para isso utiliza
 
@@
 
-% Paper of ALEMÃO PROB ,java virtual machine
-
 % O manual do Z80
@@
 America, 1 edition, 1996.
+
+\bibitem {BHDL_2003}    
+   Aljer, P. Devienne, S. Tison and J-L. Boulanger andG. Mariano. Bhdl: Circuit Design in B. A. In
+    ACSD, Third International Conference on Application of Concurrency to System Design, pages 241-242, 2003.
+
+\bibitem {GEMPLUS_99}
+Casset L.; Lanet J. L. A Formal Specification of the Java Bytecode Semantics using the B
+method.Technical Report, Gemplus. 1999.
 

### 2009-04-27 papers/FM2009/source/paper.tex

@@
 
-[Conclusao]This project has many proof obligations and its process of verification is the step of the most time
-consumed. The verification step of all the model spent about a month. This was hard, because was necessary to build,
-correct and prove several expressions from basic projects, Hardware and Types Library. However, when the basic
-projects are ready, the modelling and verification of new plataform \footnote{Actualy, the microprocessors M8051 and
-PIC16F432 are under construction} become more fast and easy. Thus, these basic projects were also built to re-use
-to other platforms.
-
-When the basic projects are ready, then about 89\% proof non obvious and 55\% of Proofs WD was conclused
-automaticly. The others were concluded through the use of commands ...
-Com os projetos de base completamente provados, então aproximadamente 89\% das provas não obvias e 55\% da
-Proofs WD foram realizadas automaticamente. As demais provas foram realizadas após o uso de alguns passos
-de prova\footnote{Passos de prova são comandos que indicam um caminho para o provador concluir a prova,
-esses comandos apenas indicam um caminho, ou seja, ele não introduzem hipósteses falsas} associados a
-pattern matching of proof obligations. The user pass associeted to match pattern  aceleraram bastante o
-processo de prova, pois existem muitas instruções do microcontrolador com caractéristicas semelhantes e o
-atelierb é capaz de re-usar os passos de prova. Então, poucos passos de prova podem ser úteis para uma
-grande variedade de provas. Um exemplo é o conjunto de 17 passos de prova que rapidamente ajudou a
-verificar 99\% (2295) das provas WD. A experiência em adicionar passos de prova também ajudou a reduzir o
-tempo do processo de prova em mais de 90\%. Isso aconteceu, porque o uso de passos de prova direciona melhor o
-provador ao fim da prova.
-
-To further reduce cost of proof process, a new feature supported by AtelierB was used, the
-parallelization. To do this were used three computers each one with two cores, the that theoretically
-increased by six times the capacity of processing.
+% [Conclusao]
+This project has many proof obligations and its process of verification is the step of the most time consumed. The
+verification step of all the model spent about a month. This was hard, because was necessary to build, correct and
+prove several expressions from basic projects, Hardware and Types Library. However, when the basic projects are
+ready, the modelling and verification of new plataform \footnote{Actualy, the microprocessors M8051 and PIC16F432
+are under construction} become more fast and easy. Thus, these basic projects were also built to re-use to other
+platforms. When the basic projects are ready, then about 89\% proof non obvious and 55\% of Proofs WD was conclused
+automaticly.
+
+The others were concluded through the use of proof commands\footnote{The proof commands are step that just indicate
+a walk to prover make the proof, thus, theses commands cannot introduce false hypotheses.} associated to match
+pattern from proof obligations. The match pattern accelerates too much the prove process, because there are many
+similar assembler instructions and the AtelierB can re-use the match pattern and its proof commands associated.
+Therefore, few proof commands can be useful to many obrigações de prova. A good example is a set of 17 proof
+commands that fastly helps to verify 99\% (2295) of proofs WD. To further reduce cost of proof process, a new feature
+supported by AtelierB was used, the parallelization. To do this were used three computers each one with two cores,
+the that theoretically increased by six times the capacity of processing.
 
 A special care was taken to represent the model in way simple and accurate. This care also reduced the number of
-proofs. Because, the simple act of optimizing the use the simple act of optimizing the use of
-override operator ($\lover$) fell too the number of proof obligations.
+proofs. Because, the simple act of optimizing the use of override operator ($\lover$) fell too the number of proof
+obligations.
 
@@
 
+ \section{Related Works}
+
+% This work had as the main motivation the  verififcation in assembler level and consequently the
+% verification of harware model. The model created express the semantic of functioning of hardware and
+% with this model is possible verify the semantic of hardware components.
+
+There are in the literature of computer science some technics\cite{BHDL_2003,GEMPLUS_99} to model hardware
+and the virtual machines using the B method. Then, the B method has been used successfully to model the
+operational semantic, however The cost modeling was still too expensive and this paper describes some
+techniques to lower the cost modeling. In general, the researchers are concetrated at model the Java
+Virtual Machine (JVM), but it is not easy to specify  it, because it very complex, and become more
+difficult to write good lemmas, to normalize the functions and to verify the properties. Then, this
+technique specify a model more simple to get better results. Besides, the objectives are to get concret
+results fastly and, in the same time, stay near to model of actual processors.
+% Re-uso  - modularization
+The most of related works are specific to one plataform and its basic definitions cannot be re-used at in
+other platforms. The used strategy created a good logic separation about the concepts of modelling. Thus,
+the basic projects, the Hardware and Types Library, have functions with a low coupling and high
+cohesion, that promove the re-use. Soon, these re-use better too the code generation and prove
+process.{{Pare aqui \ssssssss }}
+
+ [Uma das grandes vantagens é a evolução suporte ao método B]
+
+ A other advantage of this work is the use of news and advanceds funcionalities of B method. The first is
+ the paralelization of prove procces and  the second is instensive use of math patterns associated with proof
+ commands. In the
+
+ [AMM]
+Verificação formal já vem sendo aplicada com sucesso na verificação \ldots Porém os custos eram altos,
+pois se passou X tempo para modelar \ldots e
+
+A modelagem do microcontrolador foi construida por um unico designer and consumiu aproximadamente six
+month to study the plattaform Z80, model, adjust to simplifi and prove.
+
  \section{Conlusitions}
+
+ % Trab. Futuros
+ %- Refinar o modelo para gerar através dos  código e apartir deste código criar um
+ %interpretador/simulador
+ %- Construir compilador
+
+%  \begin{itemize}
+%    Citar ambuiguidades encontradas no manual e eventuais erros encontrados.
+%   \item Utilidade o processo de desenvolvimento do hardware:
+%   \subitem Ajuda na remoção de eventuais ambiguidades providas de uma descrição informal.
+%   \subitem Verificação de propriedades: sobre a manipulação e tipo dos dados.
+%   \subitem Documentação formalizada matemáticamente
+%   \subitem Possibilidade da geracao de código VHDL apartir do modelo B.
+%   \subitem Colaboracao para construção de um simulador/interpretador
+% \end{itemize}
+
 [I need review the english]

### 2009-04-27 papers/FM2009/source/paper.tex
almost finishing the related works

@@
 
-There are in the literature of computer science some technics\cite{BHDL_2003,GEMPLUS_99} to model hardware
-and the virtual machines using the B method. Then, the B method has been used successfully to model the
-operational semantic, however The cost modeling was still too expensive and this paper describes some
-techniques to lower the cost modeling. In general, the researchers are concetrated at model the Java
-Virtual Machine (JVM), but it is not easy to specify  it, because it very complex, and become more
-difficult to write good lemmas, to normalize the functions and to verify the properties. Then, this
-technique specify a model more simple to get better results. Besides, the objectives are to get concret
-results fastly and, in the same time, stay near to model of actual processors.
+There are in the literature of computer science some technics\cite{BHDL_2003,GEMPLUS_99} to model hardware and the
+virtual machines using the B method. Then, the B method has been used successfully to model the operational
+semantic, however the cost of modeling was still expensive and this paper describes some techniques to lower the
+cost modeling.
+
+In general, the researchers are concetrated at model the Java Virtual Machine (JVM), but it is not easy to specify 
+it, because it very big and complex, and become more difficult to write good lemmas, to normalize the functions and
+to verify its properties. Then, this work specify a model more simple to get better results fastly, in the same
+time, stay near to model of actual processors.
 % Re-uso  - modularization
-The most of related works are specific to one plataform and its basic definitions cannot be re-used at in
-other platforms. The used strategy created a good logic separation about the concepts of modelling. Thus,
-the basic projects, the Hardware and Types Library, have functions with a low coupling and high
-cohesion, that promove the re-use. Soon, these re-use better too the code generation and prove
-process.{{Pare aqui \ssssssss }}
-
- [Uma das grandes vantagens é a evolução suporte ao método B]
-
- A other advantage of this work is the use of news and advanceds funcionalities of B method. The first is
- the paralelization of prove procces and  the second is instensive use of math patterns associated with proof
- commands. In the
+The most of related works are specific to one plataform\cite{GEMPLUS_99,ammm} and its basic definitions cannot be
+re-used at in other platforms. The used strategy created a good logic separation about the concepts of modelling.
+Thus, the basic projects, the Hardware and Types Library, have functions with a low coupling and high cohesion, that
+promove the re-use. Soon, these re-use better too the code generation, the confiability and prove process.
+
+ Others advantages of this work are the use of news and advanceds funcionalities of B method. Normaly, the related
+ works don`t enjoy this funcionalities, because these are news and required advanced knowleadge about tools for B
+ method. These advantages are the paralelization of prove proccess and the instensive use of math patterns
+ associated with proof commands.
 
  [AMM]
-Verificação formal já vem sendo aplicada com sucesso na verificação \ldots Porém os custos eram altos,
-pois se passou X tempo para modelar \ldots e
+Verificação formal já vem sendo aplicada com sucesso na verificação \ldots Porém os custos eram altos, pois se
+passou X tempo para modelar \ldots e
 
@@
 month to study the plattaform Z80, model, adjust to simplifi and prove.
+
+Quais as desvantagens/problemas com relaçao aos outros?
+-The B method is not good to verify all kindy of cararacteristics.
+-The our results are not mature, b 
+-Actually, we don`t define 
+-The B method can verify many diferentes properties of model, but some properties cannot be easy to verify and
+model.
+ -Some times, the B method is very dificult to model, eventually a other metodology can verify
 

### 2009-04-28 papers/FM2009/source/paper.tex
Related works

@@
 
-In general, the researchers are concetrated at model the Java Virtual Machine (JVM), but it is not easy to specify 
-it, because it very big and complex, and become more difficult to write good lemmas, to normalize the functions and
-to verify its properties. Then, this work specify a model more simple to get better results fastly, in the same
-time, stay near to model of actual processors.
+In general, the researchers are concetrated at model the Java Virtual Machine (JVM), but it is not easy to
+specify it, because it very big and complex, and become more difficult to write good lemmas, to normalize
+the functions and to verify its properties. Then, this work specify a model more simple to get better
+results fastly, in the same time, stay near to model of actual processors.
 % Re-uso  - modularization
-The most of related works are specific to one plataform\cite{GEMPLUS_99,ammm} and its basic definitions cannot be
-re-used at in other platforms. The used strategy created a good logic separation about the concepts of modelling.
-Thus, the basic projects, the Hardware and Types Library, have functions with a low coupling and high cohesion, that
-promove the re-use. Soon, these re-use better too the code generation, the confiability and prove process.
-
- Others advantages of this work are the use of news and advanceds funcionalities of B method. Normaly, the related
- works don`t enjoy this funcionalities, because these are news and required advanced knowleadge about tools for B
- method. These advantages are the paralelization of prove proccess and the instensive use of math patterns
- associated with proof commands.
+The most of related works are specific to one plataform\cite{GEMPLUS_99,ammm} and its basic definitions
+cannot be re-used at in other platforms. The used strategy created a good logic separation about the
+concepts of modelling. Thus, the basic projects, the Hardware and Types Library, have functions with a low
+coupling and high cohesion, that promove the re-use. Soon, these re-use better too the code generation,
+the confiability and prove process.
+
+ Others advantages of this work are the aplication of news and advanceds funcionalities of B method.
+ These advantages are the paralelization of prove proccess and the instensive use of math patterns associated
+ with proof commands. Normaly, the related works don`t enjoy this funcionalities, because these are news
+ and required advanced knowleadge about tools for B method.
 
  [AMM]
-Verificação formal já vem sendo aplicada com sucesso na verificação \ldots Porém os custos eram altos, pois se
-passou X tempo para modelar \ldots e
+ Adici \erro [ Adicionar aqui a comparação enviada ao e-mail de David]
+
+Verificação formal já vem sendo aplicada com sucesso na verificação \ldots Porém os custos eram altos,
+pois se passou X tempo para modelar \ldots e
 
@@
 
- \section{Conlusitions}
+ \section{Conclusitions}
 

### 2009-04-28 papers/FM2009/source/paper.tex

@@
 \section{Introduction}
+
+The B method allows the construction of safety computacional systems by proofs verification and generation
+that guarantees the compatibility among modelling steps and that the system desirable state allways be
+true.Therefore, a initial abstract model of the system is defined and refined until the implementation
+level. B still allows the code generation from implementation, but this fase isn't verified.
+
+In\cite{Dantas_Report08} is presented a approach to complement the B method until the codify level:
+extension of the B method until the assembly level. In this approach, the system construct beggining from
+the abstract level and it is formally refined until the target plataform assembly. Therefore, the first
+step of this approach is the target plataform modelling, the microprocessor, in B. Using the
+responsability division mechanism provide by B basic modules was developted to be used as elementary
+library to constructing generic microprocessor, in which bit data type, bit arrays and others are
+specified together with its operations\cite{Medeiros}. The resulting processor B machine  will be used as
+library in which its operations are used in the modelling of the systems. The use of this library occurs
+at the B implementation level of the system.
+
+The processor modelling can be used in plataforms project documentation, simulators construction,
+consistency verification and software verification at assembly level. The verification at assembly level
+is more interesting utility because allows the developtment of trustworth software by construction, that
+is extremaly desirable in embedded critical software.
+
+In this paper we introduced the modelling of Z80 processor\cite{Z80}, since elementary library until
+assembly instructions. The paper is structured as follows. Section\ref{B Notation} provides a general introduction
+to the B method.Section\ref{General definitions for hardware models} presents the elementary library of
+microprocessors basic concepts. Section\ref{Description of the B model from Z80} presents the Z80 B
+modelling. Section\ref{} TERMINAR
+
 % Explica o contexto geral do artigo
@@
 
-[O que as provas garantem?] As obrigações de provas garante a verficação dos tipos de dados, das propriedades
+[O que as provas garantem?]
+ As obrigações de provas garante a verficação dos tipos de dados, das propriedades
 estabelecidas no modelo e se as expressões são bem formadas \footnote{An expression is called "well-defined" (or
@@
 como estas propriedades estão em nível de especificação, então é possível ajusta-las a necessidade de um determinado
-sofwtare sem implicar em mudanças no hardware, o que implica apenas em aumentar ou diminuir o número de provas.
+sofwtare sem acarretar em mudanças no hardware, o que implica apenas em aumentar ou diminuir o número de
+provas.
 % O desenvolvedor poderia por exemplo aumentar a região de memória necessária para armazenar dados de trabalho.
@@
 General & 140 & 615  \\
-Properties and Assertions & 68 & 169  \\
+Initialization, Properties and Assertions & 68 & 169  \\
  \textbf{Total} &  607 &  2319  &  %\textbf{Total} 2926 \\
@@
 The others were concluded through the use of proof commands\footnote{The proof commands are step that just indicate
-a walk to prover make the proof, thus, theses commands cannot introduce false hypotheses.} associated to match
-pattern from proof obligations. The match pattern accelerates too much the prove process, because there are many
+a walk to prover make the proof, thus theses commands cannot introduce false hypotheses.} associated to
+match pattern from proof obligations. The match pattern accelerates too much the prove process, because there are many
 similar assembler instructions and the AtelierB can re-use the match pattern and its proof commands associated.
@@
 
-% This work had as the main motivation the  verififcation in assembler level and consequently the
-% verification of harware model. The model created express the semantic of functioning of hardware and
-% with this model is possible verify the semantic of hardware components.
+ [ I need review the english]
 
@@
 semantic, however the cost of modeling was still expensive and this paper describes some techniques to lower the
-cost modeling.
+cost modeling. Although, the modeling of the Z80 is not yet represented some small aspects.
+
+%The objective of researchers
 
@@
 specify it, because it very big and complex, and become more difficult to write good lemmas, to normalize
-the functions and to verify its properties. Then, this work specify a model more simple to get better
-results fastly, in the same time, stay near to model of actual processors.
+the functions and to verify its properties. Then, this work specify a model more simple to get fastly
+better results , in the same time, stay near to model of actual processors.
+
 % Re-uso  - modularization
+
 The most of related works are specific to one plataform\cite{GEMPLUS_99,ammm} and its basic definitions
@@
 
- Others advantages of this work are the aplication of news and advanceds funcionalities of B method.
- These advantages are the paralelization of prove proccess and the instensive use of math patterns associated
+% advantages
+
+Others advantages of this work are the aplication of news and advanceds funcionalities of B method. These
+advantages are the paralelization of prove proccess and the instensive use of math patterns associated
 with proof commands. Normaly, the related works don`t enjoy this funcionalities, because these are news
@@
 
- [AMM]
- Adici \erro [ Adicionar aqui a comparação enviada ao e-mail de David]
-
-Verificação formal já vem sendo aplicada com sucesso na verificação \ldots Porém os custos eram altos,
-pois se passou X tempo para modelar \ldots e
-
-A modelagem do microcontrolador foi construida por um unico designer and consumiu aproximadamente six
-month to study the plattaform Z80, model, adjust to simplifi and prove.
-
-Quais as desvantagens/problemas com relaçao aos outros?
--The B method is not good to verify all kindy of cararacteristics.
--The our results are not mature, b 
--Actually, we don`t define 
--The B method can verify many diferentes properties of model, but some properties cannot be easy to verify and
-model.
- -Some times, the B method is very dificult to model, eventually a other metodology can verify
+% main problems
+
+This work had as the main motivation the  verififcation in assembler level and consequently the
+verification of harware model. Thus, the some aspects about hardware aren't modeled. For example, the time
+of execution of instruction, the pipeline and others questions not directly related with the semantic of
+software. However, there are several other specialized techinques to verify this questions. The care to
+simplify the model also helps in the prove process.
 
  \section{Conclusitions}
-
- % Trab. Futuros
- %- Refinar o modelo para gerar através dos  código e apartir deste código criar um
- %interpretador/simulador
- %- Construir compilador
-
-%  \begin{itemize}
-%    Citar ambuiguidades encontradas no manual e eventuais erros encontrados.
-%   \item Utilidade o processo de desenvolvimento do hardware:
-%   \subitem Ajuda na remoção de eventuais ambiguidades providas de uma descrição informal.
-%   \subitem Verificação de propriedades: sobre a manipulação e tipo dos dados.
-%   \subitem Documentação formalizada matemáticamente
-%   \subitem Possibilidade da geracao de código VHDL apartir do modelo B.
-%   \subitem Colaboracao para construção de um simulador/interpretador
-% \end{itemize}
-
 [I need review the english]
-% [intro]
-A necessidade comercial para produzir \textit{hardware} e \textit{software} de alta qualidade é sempre
-crescente. O métodos B já têm demonstrado sucesso na especificação de sistemas comerciais críticos.
-
-O suporte fornecido pelo método B é bastante adequado, maduro, extensivo e
-eficiente no processo de verificação dos modelos das plataformas. Isso vem sendo
-perceptível, pois todas as características relevantes das plataformas foram
-apropriadamente modeladas, principalmente as importantes para verificação em
-nível de montagem (\textit{assembly}).
-
-[ O que foi o diferencial ? qual grande vantagem]
-[Eficiência da técnica/método B]
-Por outro lado, uma das qualidades de destaque do método B é relacionada com a
-boa capacidade de verificação automática das ferramentas, pois, na versão estável
-das modelagens, do total das $503$ provas dos módulos principais do \textit{PIC},
-\textit{8051} e \textit{Z80} apenas $3$ referentes ao \textit{Z80} foram provadas
-de forma semi-automática. Ou seja, desse total de provas menos de $1$\%
-necessitaram de uma simples intervenção do projetista para ser concluída.
-
-Adicionalmente, a aplicação das técnicas modularização e re-uso tornou o processo
-de especificação das plataformas mais padronizado e acelerado. Desse modo, é
-demonstrada a viabilidade dessas especificações, pois com a aquisição de
-maturidade das ferramentas e do projetista, o processo de modelagem ganhou
-agilidade e maior automação na fase de validação do projeto.
-
-[maior vantagem]
-Enfim, a modelagem B dessas plataformas possibilitarão a verificação em um dos
-mais altos níveis de certificação de \textit{software}. Não somente, elas podem
-ser importantes no projeto de plataformas para a construção dos simuladores,
-documentação e sua própria verificação formal, permitindo identificar várias
-inconsistências delas em tempo de projeto.
-
-[Enfatizar a vantagem]
-[Flexibilidade]
-The traditional don´t allow verify this things and the properties is flexible. The
-flexbility allow add restrictions about tempo de execuçao and until specifics restrictions about the application of
-microcontroller.
-
-This flexibility of create propertis allow create conditions to verify if the software to finish in cert instant.
-
-No contexto da verificação em nível de assembly:
-Exemplificar com propriedade de restrição temporal.
-Exp. é possível verificar se o software satisfazer restrições temporais e para isso utiliza
-
-%  Para a modelagen dessa plataforma, essa modelagen contém linhas e $np $
-%  provas não obvias. Uma forma de acelerar o processo de prova, foi paralelizar
-%  as provas. Isto foi feito em 4 IMacs e consumiu $ tp $ tempo para verificar
-%  automaticamente $np - p $. As demais foi necessário uma certa interação com
-%  o provador para sua verificação.
-
-%  Para oferecer uma maior garantia do modelo com a implementação do processador
-%  algumas atividades foram realizadas. Primeiramente, tentantamos animar e gerar
-%  código respectivamente nas ferramentas ProB e AtelierB. Entratanto, o modelo
-%  cotém algumas construções complexas que ainda não são bem suportada por elas.
-%  Então, foi realizada uma apresentação sobre a descrição do modelo e foi
-%  realizada um nova revisão com 3 pessoas.
+% - Intro
+% - Importancia do trabalho  -
+% - Qual a grande vantagem - A automação
+% - Trabalhos futuros  [O que almejamos citar o AAA] -> In the future, this works like  -> %- Refinar o  modelo  para gerar através dos  código e apartir deste código criar um  interpretador/simulador
+
+%Intro
+This work shown a approach to model formally a microprocessor using the B method. The applied approach
+provide interesting results and had a great automatization in the prove proccess. Besides, this research
+intend extend the works to verification in the assembly level.
+
+This research is much interesting to design microcontroller. The formal model allow founnd some erros at
+the oficial manual\cite{Zilog} and ambiguities. The Z80's B model can replace the documentation used by
+assembler programmers, because the formal model represents the instructions effects and the B-method has a
+easy language, that remembers the pascal language. Besides, the formal model restricts the defintions to
+correct typing, use expressions well-defined and allow verify properties on the model.
+
+The Atelierb supports code generation to programming languages and latex documents, this feature also
+helps to build an interpreter or simulator through the formal model. In the future, this functionality can
+be changed to genarate code of hardware description. Actually, the news and advanced features automate the
+process even more.
+
+%[Automação]
+Actually, the mains attractions of B method to model microcontroolers are: the support the paralelization
+of prove proccess and matching patterns associated to proof commands.
+% Por outro lado, uma das qualidades de destaque do método B é relacionada com a
+% boa capacidade de verificação automática das ferramentas, pois, na versão estável
+% das modelagens, do total das $503$ provas dos módulos principais do \textit{PIC},
+% \textit{8051} e \textit{Z80} apenas $3$ referentes ao \textit{Z80} foram provadas
+% de forma semi-automática. Ou seja, desse total de provas menos de $1$\%
+% necessitaram de uma simples intervenção do projetista para ser concluída.
+
+% Adicionalmente, a aplicação das técnicas modularização e re-uso tornou o processo
+% de especificação das plataformas mais padronizado e acelerado. Desse modo, é
+% demonstrada a viabilidade dessas especificações, pois com a aquisição de
+% maturidade das ferramentas e do projetista, o processo de modelagem ganhou
+% agilidade e maior automação na fase de validação do projeto.
+% 
+% [maior vantagem]
+% Enfim, a modelagem B dessas plataformas possibilitarão a verificação em um dos
+% mais altos níveis de certificação de \textit{software}. Não somente, elas podem
+% ser importantes no projeto de plataformas para a construção dos simuladores,
+% documentação e sua própria verificação formal, permitindo identificar várias
+% inconsistências delas em tempo de projeto.
+
+[Trabalhos futuros ]
+ [O que almejamos citar o AAA]
+  In the future, this works like  -> %- Refinar o  modelo  para gerar através dos  código e apartir deste código criar um  interpretador/simulador
+% Exp. é possível verificar se o software satisfazer restrições temporais e para isso utiliza
+
+[Fechamento das ideias]
+% Desta forma, o suporte fornecido pelo método B é bastante adequado, maduro, extensivo e 
+% eficiente no processo de verificação dos modelos das plataformas. Isso vem sendo
+% perceptível, pois todas as características relevantes das plataformas foram
+%  apropriadamente modeladas, principalmente as importantes para verificação em
+%  nível de montagem (\textit{assembly}).
 
@@
 
-This work received supports from ANP (National Agency of Oil) and CNPq. Besides, the researcher Bruno Gomes help in
-the revision.
+This work received supports from ANP (National Agency of Oil) and CNPq. 
+%Besides, the researcher Bruno Gomes help in the revision.
 

### 2009-06-20 papers/SBMF2009/source/paper.tex
Initial import.

\documentclass{llncs}

\usepackage{graphicx,array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{epsfig}
\usepackage{float}
\usepackage{supertabular} % used in B Symbols

\include{comandos}
\include{SymbolsB_AtelierB}

%
\begin{document}

\title{A Formal Model of Microcontroller Instruction Set in B}

\author{Val\'{e}rio Medeiros Jr\inst{1}, David D\'{e}harbe\inst{1}}

\institute{Federal University of Rio Grande do Norte, Natal RN 59078-970, Brazil}

\maketitle

\begin{abstract}
This paper describes an approach to model the microcontroller platforms. More specifically, it shows details about
the Z80 model. The model has been developed using the B method; which applies math and logic concepts to describe
characteristics from platforms. Therefore, this modelling can be used in platform projects to document, build
simulators, verify properties about the model and verify at assembly level software. The verification
at assembly is the feature more important, because this allows to develop faithful software to the algorithm
language. Thus, the platform model was built for this purpose. Finally, this paper shows some relevant techniques
used in models and to describe some details of our approach.
\end{abstract}
%
%\keywords{Fomal Methods, microcontroller Verification, Hardware}

\section{Introduction}

This work describes briefly the modelling of microcontroller Z80~\cite{Z80_manual} and some techniques used to build
it. This modelling\footnote{The interested reader in more details is invited to visit our repository
at: http://code.google.com/p/b2asm} use the B method~\cite{Abrial}, because it has a great maturity and advanced
support. Besides, the modelling of microcontroller can extend the level of verification of B method, as it will be
explained.

The B method supports the construction of safety systems model by verification of proofs that guarantees its
correctness. So, an initial abstract model of the system is defined and refined until the implementation model. The
B tools  still have the code generator to programming language, but this feature is not verified. In~\cite{Dantas_SBMF08} is presented an approach to extend the verification of B method until the assembler language.
In this approach, the construction of systems begins from the abstract level and it is formally refined until the
target assembly platform.

Therefore, the first step of this approach is to build the platform modelling, of the microcontroller, in B. Using the
responsibility division mechanism provided by B, basic modules was developed to be used as elementary library to
help the constructing of microcontroller model. The elementary library has many definitions about common concepts used
in the microcontrollers. This library is used by Z80 model and other two model microcontrollers that
are under construction.

The microcontroller modelling can be used in platform projects documentation, simulators construction, consistency
verification and software verification at assembly level. The verification at assembly level is more interesting
utility because it allows the development of trustworthy software by construction, that is extremely desirable in
embedded critical software.

In this paper we introduced the modelling of Z80 microcontroller, since elementary library until assembly
instructions. The paper is structured as follows. Section~\ref{sec:B_method} provides a general introduction to the
B method. Section~\ref{sec:models} presents the elementary libraries and of microcontrollers basic concepts.
Section~\ref{sec:z80} presents the Z80 B modelling. Section ~\ref{sec:Proofs} explain a little about the proofs of
models. The next section (\ref{sec:relatedworks}) discusses the related works. Finally, the last section has the
conclusions.

\section{B Method}
\label{sec:B_method}

The B method for software development~\cite{Abrial} is based on the B Abstract Machine Notation (AMN) and
the use of formally proved refinements up to a specification sufficiently concrete that programming     
code can be automatically generated from it. Its mathematical basis consists of first order logic,      integer arithmetic and set theory, and its corresponding constructs are very similar to those of the Z notation

 A B specification is structured in modules. A module defines a set of valid states, including a set of initial
 states, and operations that may provoke a transition between states. The design process starts with a module with a
 so-called functional model of the system under development. In this initial modelling stage, the B method requires
 that the user proves that, in a machine, all the initial states are valid, and that operations do not define
 transitions from valid states to invalid states.

 Essentially, a B module contains two main parts: a header and the available operations, in the
 figure~\ref{fig:maqB} has a intuitive example. The clause $\mathit{MACHINE}$ has the name of module.  The following two are used to access an external module and create an instance of an external module. The
 $\mathit{VARIABLES}$  declares the variable names. The following declares its types and the restrictions. The
 $\mathit{INITIALIZATION} $  defines the its initial value. The last define the operations that remember procedures
 from programming language.

\begin{figure}
%\begin{small}
$$
\begin{array}{l}
\MACHINE \quad  \mathit{micro} \quad \quad \quad \quad \quad  \quad \quad \quad \quad \INITIALISATION  \mathit{pc}
:= 0\\
\SEES \quad \mathit{TYPES}, \mathit{ALU} \quad \quad \quad \quad \quad \quad \quad \quad \OPERATIONS\\
\INCLUDES    \quad \mathit{MEMORY} \quad \quad \quad \quad \quad \quad \quad \mathit{JMP} \mathit{( jump )} =
\\
\VARIABLES \quad    \mathit{pc}  \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \PRE \mathit{jump}
\in\mathit{INSTRUCTION}\\ 
 \INVARIANT \quad \mathit{pc} \in \mathit{INSTRUCTION}  \quad \ \THEN \mathit{pc} :=
\mathit{jump}\\  
\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad
\END \quad  \ldots \\
% \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad
%\quad  \ldots
\end{array}
$$
% \begin{array}{l}
% \MACHINE \quad \quad \quad \quad \quad \quad  \quad \quad \quad \quad \INITIALISATION \\
% \quad   \mathit{micro} \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad   \mathit{pc} := 0\\
% \SEES \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad   \OPERATIONS\\
% \quad \mathit{TYPES}, \mathit{ALU}\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad \\
% \INCLUDES                \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad   \mathit{JMP} \mathit{( jump )} = \\
% \quad   \mathit{MEMORY}   \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \PRE\\
% \VARIABLES      \quad \quad \quad \quad \quad \quad \quad \quad \quad  \quad \quad  \mathit{jump} \in
% \mathit{INSTRUCTION}\\ \quad   \mathit{pc}    \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad\quad \quad \THEN\\
% \INVARIANT   \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \mathit{pc} := \mathit{jump}\\
% \quad \mathit{pc} \in \mathit{INSTRUCTION} \ \ \quad \quad \quad \quad \quad \quad \END\\
% \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad
% \quad \quad  \ldots
% \end{array}

%\end{small}
\caption{A part of B simplified module}
\label{fig:maqB}
\end{figure}

\section{Overview of Plataform's B Models}
\label{sec:models}

We have been developed a reusable set of basic definitions to model hardware concepts and data types
concepts. These definitions are grouped in two projects of libraries and other project to model the
platform. Thus, our workspace is arranged through three projects: hardware library, types library and
the project of specific platform, in this case the Z80.

The dependency graph between these projects is depicted in figure~\ref{fig:hardware-definition-graph}
and the next sections will explain each one of them.

\begin{figure}[h] \centering
\includegraphics[width=.92\textwidth]{diagramaEstrutural_vertical.jpg}
 \caption{Dependency graph between the projects that compose the platform model of Z80.}
\label{fig:hardware-definition-graph}
\end{figure}

\subsection{Definitions to Represent and Manipulate Bits}

The entities defined in the module $\mathit{BIT\_DEFINITION}$ are the
type for bits, logical operations on bits (negation, conjunction,
disjunction, exclusive disjunction), as well as a conversion function
from booleans to bits.

First, bits are modelled as the set of integers: $\mathit{BIT} = \mathit{0..1}$. The negation is an unary
function on bits and it is defined as:

$
\begin{array}{l}
\mathit{bit\_not}  \in  \mathit{BIT}  \fun  \mathit{BIT}  \land \forall ( \mathit{bb}). (\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bb}) =
1-\mathit{bb})\\
\end{array}
$

The module also provides lemmas on negation that may be useful for the users of the library to develop proofs:

$
\begin{array}{l}
%  \mathit{bit\_not}(0) = 1;  \mathit{bit\_not}(1) = 0; \\
\forall (\mathit{bb}).(\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bit\_not}(\mathit{bb})) = \mathit{bb})
\end{array}
$

Conjunction is an unary function on bits and it is defined as:

$
\begin{array}{l}
\mathit{bit\_and} \in \mathit{BIT} \times \mathit{BIT} \fun \mathit{BIT} \land \\
\forall (\mathit{b1}, \mathit{b2}).(\mathit{b1}  \in \mathit{BIT}  \land \mathit{b2} \in \mathit{BIT} \implies \\
\quad ((\mathit{bit\_and}(\mathit{b1}, \mathit{b2}) = 1) \iff (\mathit{b1} = 1)  \land  (\mathit{b2} = 1)))
\end{array}
$

The module provides the following lemmas for conjunction, either:

$
\begin{array}{l}
%  \mathit{bit\_and}(0,0) = 0;  \mathit{bit\_and}(0,1) = 0; \\
%  \mathit{bit\_and}(1,0) = 0;  \mathit{bit\_and}(1,1) = 1; \\
\forall (\mathit{b1},\mathit{b2}).(\mathit{b1} \in \mathit{BIT} \land \mathit{b2} \in \mathit{BIT} \implies \\
\quad (\mathit{bit\_and}(\mathit{b1}, \mathit{b2}) = \mathit{bit\_and}(\mathit{b2},\mathit{b1})))\land \\
\forall (\mathit{b1},\mathit{b2},\mathit{b3}).(\mathit{b1} \in \mathit{BIT} \land  \mathit{b2} \in \mathit{BIT} \land \mathit{b3} \in \mathit{BIT} \implies \\
\quad (\mathit{bit\_and}(\mathit{b1}, \mathit{bit\_and}(\mathit{b2},\mathit{b3})) = \mathit{bit\_and}(\mathit{bit\_and}(\mathit{b1},\mathit{b2}),\mathit{b3})))\\
% \forall (\mathit{b1}).(\mathit{b1} \in \mathit{BIT} \implies (\mathit{bit\_and}(\mathit{b1}, 1) = \mathit{b1})); \\
% \forall (\mathit{b1}).(\mathit{b1} \in \mathit{BIT} \implies (\mathit{bit\_and}(\mathit{b1}, 0) = 0));
\end{array}
$

The module provides definitions of $\mathit{bit\_or}$ (disjunction) and $\mathit{bit\_xor}$ (exclusive disjunction),
as well as lemmas on those operators. These are standard and their expression in B is similar as for
$\mathit{bit\_and}$, they are thus omitted.

Finally, the conversion from booleans to bits is simply defined as:

$
\begin{array}{l}
\mathit{bool\_to\_bit} \in \BOOL \fun \mathit{BIT} \land \mathit{bool\_to\_bit} = \{ \TRUE \mapsto 1, \FALSE \mapsto 0 \} \\
\end{array}
$

Observe that all the lemmas that are provided in this module have been
mechanically proved by the theorem prover included with our B
development environment. None of these proofs requires human insight.

\subsection{Representation and Manipulation of Bit Vectors}

Sequences are pre-defined in B, as functions whose the domain is an integer range with lower bound 1 (one). Indices
in bit vectors usually range from 0 (zero) upwards and the model we propose obeys this convention by making an
one-position shift where necessary. This shift is important to use the predefined functions of sequences. We
thus define bit vectors as non-empty sequences of bits, and $\mathit{BIT\_VECTOR}$ is the set of all such sequences:
$
\begin{array}{l}
\mathit{BIT\_VECTOR} = \seq (\mathit{BIT})
\end{array}
$

The function $\mathit{bv\_size}$ returns the size of a given bit vector. It is basically a wrapper for the
predefined function $\mathbf{size}$ that applies to sequences.

$
\begin{array}{l}
\mathit{bv\_size} \in \mathit{BIT\_VECTOR} \fun \nat_1 \land \\
\mathit{bv\_size} = \lambda bv . (bv \in \mathit{BIT\_VECTOR} \mid \mathbf{size}(bv))
\end{array}
$

We also define two functions $\mathit{bv\_set}$ and $\mathit{bv\_clear}$ that, given a bit vector, and a position of
the bit vector, return the bit vector resulting from setting the corresponding position to 0 or to 1, and a function
$\mathit{bv\_get}$ that, given a bit vector, and a valid position, each one returns the value of the bit at that
position. Only the first definition is shown here:

$
\begin{array}{l}
\mathit{bv\_set} \in \mathit{BIT\_VECTOR} \times \nat \fun \mathit{BIT\_VECTOR} \land \mathit{bv\_set} =\\
\lambda v, n . (v \in \mathit{BIT\_VECTOR} \land n \in \nat \land n <\mathit{bv\_size}(v)
\mid v \lover \{ n+1 \mapsto 1 \})
\end{array}
$

Additionally, the module provides definitions for the classical
logical combinations of bit vectors: $\mathit{bit\_not}$,
$\mathit{bit\_and}$, $\mathit{bit\_or}$ and $\mathit{bit\_xor}$. Only
the first two are presented here. Observe that the domain of the
binary operators is restricted to pairs of bit vectors of the same
length:

$
\begin{array}{l}
\mathit{bv\_not} \in \mathit{BIT\_VECTOR} \fun \mathit{BIT\_VECTOR} \land \\
\mathit{bv\_not} = \lambda v . (v \in \mathit{BIT\_VECTOR} \mid \quad \lambda i . (1 .. \mathit{bv\_size}(v)) \mid \mathit{bit\_not}(v(i))) \land \\
\mathit{bv\_and} \in \mathit{BIT\_VECTOR} \times \mathit{BIT\_VECTOR} \fun \mathit{BIT\_VECTOR} \land \\
\mathit{bv\_and} = \lambda v_1, v_2 . (v_1 \in \mathit{BIT\_VECTOR} \land v_2 \in \mathit{BIT\_VECTOR} \land \\
\quad \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \mid \lambda i . (1 .. \mathit{bv\_size}(v_1)) \mid
\mathit{bit\_and}(v_1(i), v_2(i)))
\end{array}
$

We provide several lemmas on bit vector operations. These lemmas
express properties on the size of the result of the operations
as well as classical algebraic properties such as associativity
and commutativity.

\subsection{Modelling Bytes and Bit Vector of Length 16}

Bit vectors of length 8 are bytes. They form a common entity in
hardware design. We provide the following definitions:

\hspace*{0.0in}\it BYTE\_WIDTH \rm = 8 $\land$ \it BYTE\_INDEX \rm = 1 $\upto$ \rm  BYTE\_WIDTH\rm $\land$

\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.10in} $\land$

\hspace*{0.0in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \}\hspace*{0.10in} $\land$

\hspace*{0.0in}\it BYTE\_ZERO  $\in$  \it BYTE  $\land$ \it BYTE\_ZERO \rm = \it BYTE\_INDEX  $\times$  \rm \{\rm 0\rm \}

The $\mathit{BYTE\_INDEX}$ is domain of byte modelled. It starts at 1 to obey a definition of sequences from B
method. However, the functions from byte encapsulate the access and these functions use the
$\mathit{PHYS\_BYTE\_INDEX}$. The $\mathit{BYTE}$ type is a specialized type from $\mathit{BIT\_VECTOR}$, but it has
a size limit. Some others more specific definitions are created to help the developer. The type $\mathit{BV16}$ is
created for bit vector of length 16 in a similar way.

\subsection{Bit Vector Arithmetics}

Bit vectors are used to represent and combine numbers: integer ranges (signed or unsigned). Then, our library
defines functions to manipulate this date, for example, the function $\mathit{bv\_to\_nat}$ that maps bit
vectors to natural numbers:

$
\begin{array}{l}
\mathit{bv\_to\_nat} \in \mathit{BIT\_VECTOR} \fun \nat \land \\
\mathit{bv\_to\_nat} = \lambda v . (v \in \mathit{BIT\_VECTOR} \mid \sum i . (i \in \dom(v) . v(i) \times 2^i))
\end{array}
$

An  associated lemma is: $\forall n . (n \in \nat_1 \implies \mathit{bv\_to\_nat}(\mathit{nat\_to\_bv}(n)) = n)$

\subsection{Basics Data Types}

The instruction set of microcontrollers usually have common data types. These types are placed in the types
library. Each type module has functions to manipulate and convert its data. There are six common basics data types
represented by modules, see details in table~\ref{tab:types}.

\begin{table}
\caption{Descriptions of basic data types}
\label{tab:types}

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
 $Type\ Name$ & UCHAR & SCHAR & USHORTINT & SSHORTINT  & BYTE & BV16 \\\hline
 $Range$ & 0..255 & -128..127 & 0..65.535  & -32.768..32.767  & -- & --\\ \hline
 $Physical\ Size $ & 1 byte & 1 byte & 2 bytes & 2 bytes &  1 bytes & 2 bytes \\ \hline
\end{tabular}
\end{center} 
\end{table}

 Usually, for each type module just needs to detail or redefines the pre-defined concepts. For
 example, the function $\mathit{bv\_to\_nat}$ from bit vector arithmetics that is specialized to
 $\mathit{byte\_uchar}$. The set $\mathit{BYTE}$ is a subset of the
$\mathit{BIT\_VECTOR}$, then this function can defined as follows:

$
\begin{array}{l}
\mathit{byte\_uchar} \in \mathit{BYTE} \fun \nat \land \\
\mathit{byte\_uchar} = \lambda (v) . ( v \in BYTE | bv\_to\_nat(v) )
\end{array}
$

The definitions of the library types reuse the basic definitions from the hardware library. This provides greater
confidence and facilitates the prove process, because the prover can reuse the
previous defined lemma.

The inverse function is easily defined as $\mathit{uchar\_byte}$.

$
\begin{array}{l}
\mathit{uchar\_byte} \in \mathit{UCHAR}  \fun  \mathit{BYTE}  \land \\
   \mathit{uchar\_byte} = \ (\mathit{byte\_uchar}) ^{-1}
\end{array}
$

% We also created the following lemmas:
%  
% $
% \begin{array}{l}
%  \forall (val) . (val \in \mathit{UCHAR} |
%  \mathit{byte\_uchar}(\mathit{uchar\_byte}(val)) = val) \land\\
%  \forall (by) . (by \in \mathit{BYTE} |
%  \mathit{uchar\_byte}(\mathit{byte\_uchar}(by)) = by)
% \end{array}
% $

Similarly, several other functions and lemmas were created for all other
data types.

\section{Description of the Z80 B model }
\label{sec:z80}

The \textit{Z80} is an important CISC microcontroller developed by \textit{Zilog}~\cite{Z80_manual}. It supports
158 different instructions. These instructions are classified into these categories: load and exchange; block transfer
and search; arithmetic and logical; rotate and shift; bit manipulation (set, reset, test); jump, call and return; input/output; and basic cpu control.

The main module includes an instance of memory module and accesses the definitions from: basic data types modules
and \textit{ALU}.

\begin{sloppypar}

\bf MACHINE

\hspace*{0.15in}\it Z80

\bf INCLUDES

\hspace*{0.10in}\it MEMORY

\bf SEES

\hspace*{0.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,

\hspace*{0.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,

\hspace*{0.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,

\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION
\end{sloppypar}

% Chacteristics of microcontroller
Each instruction is represented by B operations in the module Z80. By default, all parameters from
operations are or predefined elements in the model or integers values in the decimal representation.
The internal registers contain 208 bits of reading/writing memory. It includes two sets of six general purpose
registers which may be used individually as 8-bits registers or as 16-bits register pairs.  The work registers are
represented by variable $\mathit{rgs8}$. The domain of $\mathit{rgs8}$ ($\mathit{id\_regs8}$) is a set formed by
identifiers of registers of 8 bits. These registers can be accessed in pairs, forming 16-bits, resulting in other
set of identifiers of 16-bits registers, named $\mathit{id\_reg16}$. The main work register of Z80 is the
accumulator ($\mathit{rgs8(a0)}$) used for arithmetic, logic, input/output and loading/storing operations.
\subsection{Modelling registers, input and output ports and instructions}

The Z80 has many register of diferrents types and many instructions. It includes alternative set of accumulator,
flag and general registers. The CPU contains a stack pointer ($\mathit{sp}$), program counter ($\mathit{pc}$), two
index registers ($\mathit{ix}$ and $\mathit{iy}$), an interrupt register ($\mathit{i\_}$), a refresh register
($\mathit{r\_}$), two bits ($\mathit{iff1}$, $\mathit{iff2}$) used to control the interruptions, a pair of bits to
define the interruption mode ($\mathit{im}$) and the input and output ports ($\mathit{i\_o\_ports}$). Below, its
definitions are specified in the \textit{INVARIANT}.

\begin{sloppypar}
\bf INVARIANT

\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ pc  $\in$  \it INSTRUCTION  $\land$

\hspace*{0.10in}\it  sp  $\in$  \it BV16  $\land$  \it ix  $\in$  \it BV16  $\land$  \it iy  $\in$  \it BV16  $\land$

\hspace*{0.10in}\it i\_  $\in$  \it BYTE  $\land$  \it r\_ $\in$  \it BYTE  $\land$ iff1  $\in$  \it BIT
$\land$ \it iff2  $\in$  \it BIT  $\land$

\hspace*{0.10in}\it im \rm : \rm (\it BIT $\times$ \it BIT\rm )  $\land$ i\_o\_ports  $\in$  \it BYTE 
$\fun$  \it BYTE
\end{sloppypar}

 A simple example of instruction is a $\mathit{LD\_n\_A}$, as shown below.
Many times, to model an instruction is necessary to use the predefined functions, these help the construction of
model. This instruction use the $\mathit{updateAddressMem}$ function from \textit{Memory} module and it receives an
address memory and its new memory value. Finally it increments the program counter ($\mathit{pc}$) and update the
refresh register ($\mathit{r\_}$).

\hspace*{0.00in}\bf LD\_n\_A \rm ( \it nn \rm ) \rm =

\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} %$\land$  \it nn $\in$\it DATA\_R\_ADR

\hspace*{0.20in}\bf THEN

\hspace*{0.20in}\bf updateAddressMem \rm ( \it ushort\_to\_bv16 \rm ( \it nn \rm ) \rm , \it rgs8 \rm ( \it a0 \rm )
\rm )  $\para$

\hspace*{0.20in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it
update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.00in}\bf END\rm

The microcontroller model can specify several security properties. For example, the last operation could have a
restriction to write only in a defined region of memory.

\section{Proofs}
\label{sec:Proofs}
% [intro/scope] [O que as provas garantem? citar que dificeis, que é necessário buscar técnicas eficientes
% para realizar as provas ]

 The proof obligations allow to verify the data types, important system properties and if the expressions are
 well-defined (WD)\footnote{An expression is called ``well-defined'' (or unambiguous) if its definition assigns it a
 unique interpretation or value.}. The properties provide additional guarantees, because it can set many safety
 rules. However, the model can be very difficult to prove.

% [Conclusao] This project has many proof obligations and
The building of model was a hard and difficult task, because we needed to build, prove, fix and re-prove
several predicates of the libraries.

However, few proof commands\footnote{The proof commands are step that just indicate a walk to prover make the proof,
thus theses commands cannot introduce false hypotheses.} can be useful to prove the majority of the proof
obligations. Because, there are many similar assembler instructions. A good example is a set of 17 proof commands
that quickly aided the verification of 99\% (2295) of proofs WD. Finally, the techniques mentioned  help
the verification of all the 2926 proof obligations.

 \section{Related Works}
 \label{sec:relatedworks}

% [ I need review the english]

There are in the literature of computer science some approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and
the virtual machines using the B method. Then, the B method has been used successfully to model the operational
semantic. However the cost of modelling was still expensive and this paper quoted some techniques to lower the cost
of modelling. Although, the modelling of the Z80 do not have modelled some small aspects.
% The objective of researchers

In general, the researchers~\cite{GEMPLUS_99} are concentrated at model the Java Virtual Machines (JVM), but it is
not easy to specify, because it is very big and complex. Then, this work specify a model more simple to get better
results, at the same time, the model stay near to model of actual microcontrollers. The main motivation of our
research is the verification of assembler level and consequently the verification of hardware model. Thus, the some
aspects are not modelled. For example, the time of execution of instruction, the pipeline and others questions not
directly related with the semantic software. However, there are many other specialized techniques to verify these
questions.

 \section{Conclusions}
\label{sec:conclusions} 
This work has shown an approach to the formal modelling of microcontrollers using the B method. This approach
provide interesting results and had a great automatization in the prove process. For instance, it is interesting to
say that: during the building process of the Z80 model we found some errors and ambiguities in the official
manual\cite{Z80_manual}. The platform model can replace (or improve) the documentation used by assembler
programmers. Because the formal model represents the instructions effects and the B method has an easy notation,
similar to pascal language.  Besides, the formal model restricts the definitions to correct typing, uses expressions
well-defined and allows verifying properties of the model.

% This research is much interesting to
% design microcontroller. Because, the building of formal model found some erros and ambiguities at the oficial
% manual~\cite{Z80_manual}. The Z80's B model can replace the documentation used by assembler programmers, because the
% formal model represents the instructions effects and the B-method has a easy language, that remembers the pascal
% language. Besides, the formal model restricts the defintions to correct typing, use expressions well-defined and
% allow to verify properties on the model.

% [Trabalhos futuros ]
Future works comprise the development of a real study case in the oil area and the project of a compiler to aid the
approach.
%  The next steps of the work comprises the development of a real study case in the oil area and the project of a
% compiler to aid the automatization of verification at the assembly level

% The next works are the development of a real study case in the oil area and after the project of a compiler to aid
% the automatization of verification at the assembler level.

 \paragraph{Acknowledges:}
This work received supports from ANP (National Agency of Oil) and CNPq.
% Besides, the researcher Bruno Gomes help in the revision.

\begin{thebibliography}{5}

\bibitem {Abrial}
Abrial, J. R. The B Book: Assigning Programs to Meanings. Cambridge University Press, United States of
America, 1 edition, 1996.

\bibitem {BHDL_2003}    
Aljer, P. Devienne, S. Tison and J-L. Boulanger andG. Mariano. Bhdl: Circuit Design in B. A. In
ACSD, Third International Conference on Application of Concurrency to System Design, pages 241-242, 2003.

\bibitem {GEMPLUS_99}
Casset L.; Lanet J. L. A Formal Specification of the Java Bytecode Semantics using the B
method.Technical Report, Gemplus. 1999.

% \bibitem {CLEARSY}
% Clearsy. Atelier B. http://www.atelierb.eu.

\bibitem {Dantas_SBMF08}
Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; Et al. Applying the B Method to Take on the Grand Challenge of
Verified Compilation. In: SBMF, Savaldor, 2008. SBC.

\bibitem {HOARE}
Hoare, C. A. R. The verifying compiler, a grand challenge for computing research. In: VMCAI, p. 78-78, 2005.

\bibitem{Z80_manual}
Zilog. Z80 Family CPU User Manual. www.zilog.com/docs/z80/um0080.pdf

\end{thebibliography}

\end{document}
%
% \bibitem {clar:eke}
% Clarke, F., Ekeland, I.:
% Nonlinear oscillations and
% boundary-value problems for Hamiltonian systems.
% Arch. Rat. Mech. Anal. {\bf 78} (1982) 315--333
% %
% \bibitem {clar:eke:2}
% Clarke, F., Ekeland, I.:
% Solutions p\'{e}riodiques, du
% p\'{e}riode donn\'{e}e, des \'{e}quations hamiltoniennes.
% Note CRAS Paris {\bf 287} (1978) 1013--1015
% %
% \bibitem {mich:tar}
% Michalek, R., Tarantello, G.:
% Subharmonic solutions with prescribed minimal
% period for nonautonomous Hamiltonian systems.
% J. Diff. Eq. {\bf 72} (1988) 28--55
% %
% \bibitem {tar}
% Tarantello, G.:
% Subharmonic solutions for Hamiltonian
% systems via a $\bbbz_{p}$ pseudoindex theory.
% Annali di Matematica Pura (to appear)
% %
% \bibitem {rab}
% Rabinowitz, P.:
% On subharmonic solutions of a Hamiltonian system.
% Comm. Pure Appl. Math. {\bf 33} (1980) 609--633
% \end{thebibliography}
%

### 2009-06-26 papers/SBMF2009/source/paper.tex

@@
-\documentclass{llncs}
+\documentclass[a4paper]{llncs}
 
@@
 
-\title{A Formal Model of Microcontroller Instruction Set in B}
+\title{Formal Modelling of a Microcontroller Instruction Set in B}
 
@@
 \begin{abstract}
-This paper describes an approach to model the microcontroller platforms. More specifically, it shows details about
-the Z80 model. The model has been developed using the B method; which applies math and logic concepts to describe
-characteristics from platforms. Therefore, this modelling can be used in platform projects to document, build
-simulators, verify properties about the model and verify at assembly level software. The verification
-at assembly is the feature more important, because this allows to develop faithful software to the algorithm
-language. Thus, the platform model was built for this purpose. Finally, this paper shows some relevant techniques
-used in models and to describe some details of our approach.
+  This paper describes an approach to model the functional aspects of
+  the instruction set of microcontroller platforms using the notation
+  of the B method. The paper presents specifically the case of the Z80 
+  platform. This work is a contribution towards the extension of the
+  B method to handle developments up to assembly level code.
 \end{abstract}
@@
 
-This work describes briefly the modelling of microcontroller Z80~\cite{Z80_manual} and some techniques used to build
-it. This modelling\footnote{The interested reader in more details is invited to visit our repository
-at: http://code.google.com/p/b2asm} use the B method~\cite{Abrial}, because it has a great maturity and advanced
-support. Besides, the modelling of microcontroller can extend the level of verification of B method, as it will be
-explained.
+The B method~\cite{Abrial} supports the construction of safety systems
+model by verification of proofs that guarantees its correctness. So,
+an initial abstract model of the system requirements is defined and
+then it is refined until the implementation model. Development
+environments based on the B method also include source code generators
+for several programming languages, but the result of this translation
+cannot be compared by formal means. \cite{Dantas_SBMF08} presented
+recently an approach to extend the scope of the B method up to the
+assembly level language language. One key component of this approach
+is to build, within the framework of the B method, formal modelos of
+the instruction set of such assembly languages.
 
-The B method supports the construction of safety systems model by verification of proofs that guarantees its
-correctness. So, an initial abstract model of the system is defined and refined until the implementation model. The
-B tools  still have the code generator to programming language, but this feature is not verified.
-In~\cite{Dantas_SBMF08} is presented an approach to extend the verification of B method until the assembler language.
-In this approach, the construction of systems begins from the abstract level and it is formally refined until the
-target assembly platform.
+This work gives an overview of the formal modelling of the instruction
+set of the Z80 microcontroller~\cite{Z80_manual}\footnote{The
+  interested reader in more details is invited to visit our repository
+  at: http://code.google.com/p/b2asm.}. Using the responsibility
+division mechanism provided by B, auxiliary libraries of basic modules
+were developed as part of the construction of microcontroller
+model. Such library has many definitions about common concepts used in
+the microcontrollers; besides the Z80 model, it is used by two other
+microcontrollers models that are under way.
 
-Therefore, the first step of this approach is to build the platform modelling, of the microcontroller, in B. Using the
-responsibility division mechanism provided by B, basic modules was developed to be used as elementary library to
-help the constructing of microcontroller model. The elementary library has many definitions about common concepts used
-in the microcontrollers. This library is used by Z80 model and other two model microcontrollers that
-are under construction.
+Other possible uses of a formal model of a micro-controller
+instruction set include documentation, the construction of simulators,
+and be possibly the starting point of a verification effort for the
+actual implementation of a Z80 design. Moreover the model of the
+instruction set could be instrumented with non-functional aspects,
+such as the number of cycles it takes to execute an instruction, to
+prove lower and upper bounds on the execution time of a routine. The
+goal of this project, though, is to provide a basis for the generation
+of software artifacts at the assembly level that are amenable to 
+refinement verification within the B method.
 
-The microcontroller modelling can be used in platform projects documentation, simulators construction, consistency
-verification and software verification at assembly level. The verification at assembly level is more interesting
-utility because it allows the development of trustworthy software by construction, that is extremely desirable in
-embedded critical software.
-
-In this paper we introduced the modelling of Z80 microcontroller, since elementary library until assembly
-instructions. The paper is structured as follows. Section~\ref{sec:B_method} provides a general introduction to the
-B method. Section~\ref{sec:models} presents the elementary libraries and of microcontrollers basic concepts.
-Section~\ref{sec:z80} presents the Z80 B modelling. Section ~\ref{sec:Proofs} explain a little about the proofs of
-models. The next section (\ref{sec:relatedworks}) discusses the related works. Finally, the last section has the
-conclusions.
+This paper is focused on the presentation of the Z80 model, including
+elementary libraries to describe hardware aspects. The paper is
+structured as follows. Section~\ref{sec:B_method} provides a short
+introduction to the B method. Section~\ref{sec:models} presents the
+elementary libraries and the modelling of some elements common to
+microcontrollers.  Section~\ref{sec:z80} presents the B model of the
+Z80 instruction set. Section ~\ref{sec:Proofs} provides some
+information on the proof effort needed to analyze the presented
+models. Related work is discussed in
+Section~\ref{sec:relatedworks}. Finally, the last section is devoted
+to the conclusions.
 
@@
 
-The B method for software development~\cite{Abrial} is based on the B Abstract Machine Notation (AMN) and
-the use of formally proved refinements up to a specification sufficiently concrete that programming     
-code can be automatically generated from it.
-Its mathematical basis consists of first order logic,      integer arithmetic and set theory, and its corresponding constructs are very similar to those of the Z notation
+The B method for software development~\cite{Abrial} is based on the B
+Abstract Machine Notation (AMN) and the use of formally proved
+refinements up to a specification sufficiently concrete that
+programming code can be automatically generated from it. Its
+mathematical basis consists of first order logic, integer arithmetic
+and set theory, and its corresponding constructs are very similar to
+those of the Z notation
 
- A B specification is structured in modules. A module defines a set of valid states, including a set of initial
- states, and operations that may provoke a transition between states. The design process starts with a module with a
- so-called functional model of the system under development. In this initial modelling stage, the B method requires
- that the user proves that, in a machine, all the initial states are valid, and that operations do not define
- transitions from valid states to invalid states.
+A B specification is structured in modules. A module defines a set of
+valid states, including a set of initial states, and operations that
+may provoke a transition between states. The design process starts
+with a module with a so-called functional model of the system under
+development. In this initial modelling stage, the B method requires
+that the user proves that, in a machine, all the initial states are
+valid, and that operations do not define transitions from valid states
+to invalid states.
 
- Essentially, a B module contains two main parts: a header and the available operations, in the
- figure~\ref{fig:maqB} has a intuitive example.
-The clause $\mathit{MACHINE}$ has the name of module.
-The following two are used to access an external module and create an instance of an external module.
-The
- $\mathit{VARIABLES}$  declares the variable names. The following declares its types and the restrictions. The
- $\mathit{INITIALIZATION} $  defines the its initial value. The last define the operations that remember procedures
- from programming language.
+Essentially, a B module contains two main parts: a header and the
+available operations. Figure~\ref{fig:maqB} has a very basic
+example. The clause $\mathit{MACHINE}$ has the name of module.  The
+following two clauses respectively reference external modules and
+create an instance of an external module. The $\mathit{VARIABLES}$
+declares the name of the variables that compose the state of the
+machine. Next, the $\mathit{INVARIANT}$ clause defines the type and
+other restrictions on the variables. The $\mathit{INITIALIZATION} $
+sepecifies the initial states. Finally, operations correspond to
+the transitions between states of the machine.
 
 \begin{figure}
-%\begin{small}
+  % \begin{small}
   $$
-\begin{array}{l}
-\MACHINE \quad  \mathit{micro} \quad \quad \quad \quad \quad  \quad \quad \quad \quad \INITIALISATION  \mathit{pc}
-:= 0\\
-\SEES \quad \mathit{TYPES}, \mathit{ALU} \quad \quad \quad \quad \quad \quad \quad \quad \OPERATIONS\\
-\INCLUDES    \quad \mathit{MEMORY} \quad \quad \quad \quad \quad \quad \quad \mathit{JMP} \mathit{( jump )} =
-\\
-\VARIABLES \quad    \mathit{pc}  \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \PRE \mathit{jump}
-\in\mathit{INSTRUCTION}\\ 
- \INVARIANT \quad \mathit{pc} \in \mathit{INSTRUCTION}  \quad \ \THEN \mathit{pc} :=
-\mathit{jump}\\  
-\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad
-\END \quad  \ldots \\
-% \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad
-%\quad  \ldots
+  \begin{array}{lcl}
+    \begin{array}[t]{l}
+      \MACHINE \quad  \mathit{micro} \\
+      \SEES \quad \mathit{TYPES}, \mathit{ALU} \\
+      \INCLUDES  \quad \mathit{MEMORY} \\
+      \VARIABLES \quad    \mathit{pc}  \\
+      \INVARIANT \quad \mathit{pc} \in \mathit{INSTRUCTION} 
+    \end{array}
+    & \hspace*{1cm} &
+    \begin{array}[t]{l}
+      \INITIALISATION  \mathit{pc} := 0 \\
+      \OPERATIONS\\
+      \mathit{JMP} \mathit{( jump )} = \\
+      \quad \PRE \mathit{jump} \in \mathit{INSTRUCTION}\\ 
+      \quad \THEN \mathit{pc} := \mathit{jump}\\  
+      \quad \END \\
+      \END
+    \end{array}
   \end{array}
   $$
-% \begin{array}{l}
-% \MACHINE \quad \quad \quad \quad \quad \quad  \quad \quad \quad \quad \INITIALISATION \\
-% \quad   \mathit{micro} \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad   \mathit{pc} := 0\\
-% \SEES \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad   \OPERATIONS\\
-% \quad \mathit{TYPES}, \mathit{ALU}\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad \\
-% \INCLUDES                \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad   \mathit{JMP} \mathit{( jump )} = \\
-% \quad   \mathit{MEMORY}   \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \PRE\\
-% \VARIABLES      \quad \quad \quad \quad \quad \quad \quad \quad \quad  \quad \quad  \mathit{jump} \in
-% \mathit{INSTRUCTION}\\ \quad   \mathit{pc}    \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad\quad \quad \THEN\\
-% \INVARIANT   \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \mathit{pc} := \mathit{jump}\\
-% \quad \mathit{pc} \in \mathit{INSTRUCTION} \ \ \quad \quad \quad \quad \quad \quad \END\\
-% \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad
-% \quad \quad  \ldots
-% \end{array}
-
-%\end{small}
-\caption{A part of B simplified module}
+\caption{A very basic B machine.}
 \label{fig:maqB}
@@
 
-\section{Overview of Plataform's B Models}
+\section{Model structure and basic components}
 \label{sec:models}
 
-We have been developed a reusable set of basic definitions to model hardware concepts and data types
-concepts. These definitions are grouped in two projects of libraries and other project to model the
-platform. Thus, our workspace is arranged through three projects: hardware library, types library and
-the project of specific platform, in this case the Z80.
-
-The dependency graph between these projects is depicted in figure~\ref{fig:hardware-definition-graph}
-and the next sections will explain each one of them.
+We have been developed a reusable set of basic definitions to model
+hardware concepts and data types concepts. These definitions are
+grouped into two separate development projects and are available as
+libraries. A third project is devoted to the higher-level aspects of
+the Z80 model. Thus, the workspace is composed of: a hardware library,
+a types library and a project for the specific platform, in this case
+the Z80. The corresponding dependency diagram is depicted in
+Figure~\ref{fig:hardware-definition-graph}; information specific to
+each project is presented in the following.
 
@@
 \includegraphics[width=.92\textwidth]{diagramaEstrutural_vertical.jpg}
- \caption{Dependency graph between the projects that compose the platform model of Z80.}
+ \caption{Dependency diagram of the Z80 model.}
 \label{fig:hardware-definition-graph}
@@
 
-First, bits are modelled as the set of integers: $\mathit{BIT} = \mathit{0..1}$. The negation is an unary
-function on bits and it is defined as:
+First, bits are modelled as a set of integers: $\mathit{BIT} =
+\mathit{0..1}$. The negation is an unary function on bits and it is
+defined as:
 
@@
 
-The module also provides lemmas on negation that may be useful for the users of the library to develop proofs:
+The module also provides lemmas on negation that may be useful for the
+users of the library to develop proofs:
 
@@
 
-The module provides definitions of $\mathit{bit\_or}$ (disjunction) and $\mathit{bit\_xor}$ (exclusive disjunction),
-as well as lemmas on those operators. These are standard and their expression in B is similar as for
-$\mathit{bit\_and}$, they are thus omitted.
+The module provides definitions of $\mathit{bit\_or}$ (disjunction)
+and $\mathit{bit\_xor}$ (exclusive disjunction), as well as lemmas on
+those operators. These are standard and their expression in B is
+similar as for $\mathit{bit\_and}$, they are thus omitted.
 
@@
 
-Sequences are pre-defined in B, as functions whose the domain is an integer range with lower bound 1 (one). Indices
-in bit vectors usually range from 0 (zero) upwards and the model we propose obeys this convention by making an
-one-position shift where necessary. This shift is important to use the predefined functions of sequences. We
-thus define bit vectors as non-empty sequences of bits, and $\mathit{BIT\_VECTOR}$ is the set of all such sequences:
-$
-\begin{array}{l}
-\mathit{BIT\_VECTOR} = \seq (\mathit{BIT})
-\end{array}
-$
+Sequences are pre-defined in B, as functions whose the domain is an
+integer range with lower bound 1 (one). Indices in bit vectors usually
+range from 0 (zero) upwards and the model we propose obeys this
+convention by making an one-position shift where necessary. This shift
+is important to use the predefined functions of sequences. We thus
+define bit vectors as non-empty sequences of bits, and
+$\mathit{BIT\_VECTOR}$ is the set of all such sequences: 
+$\mathit{BIT\_VECTOR} = \seq (\mathit{BIT})$.
 
@@
 
-The \textit{Z80} is an important CISC microcontroller developed by \textit{Zilog}~\cite{Z80_manual}. It supports
-158 different instructions. These instructions are classified into these categories: load and exchange; block transfer
-and search; arithmetic and logical; rotate and shift; bit manipulation (set, reset, test); jump, call and return; input/output; and basic cpu control.
+The \textit{Z80} is a CISC microcontroller developed by
+\textit{Zilog}~\cite{Z80_manual}. It supports 158 different
+instructions. These instructions are classified into these categories:
+load and exchange; block transfer and search; arithmetic and logical;
+rotate and shift; bit manipulation (set, reset, test); jump, call and
+return; input/output; and basic cpu control.
 
@@
 
-The Z80 has many register of diferrents types and many instructions. It includes alternative set of accumulator,
+The Z80 has many register of differents types and many instructions. It includes alternative set of accumulator,
 flag and general registers. The CPU contains a stack pointer ($\mathit{sp}$), program counter ($\mathit{pc}$), two
@@
 \label{sec:Proofs}
-% [intro/scope] [O que as provas garantem? citar que dificeis, que é necessário buscar técnicas eficientes
+% [O que as provas garantem? citar que dificeis, que é necessário buscar técnicas eficientes
 % para realizar as provas ]
@@
 
-% [ I need review the english]
-
-There are in the literature of computer science some approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and
-the virtual machines using the B method. Then, the B method has been used successfully to model the operational
-semantic. However the cost of modelling was still expensive and this paper quoted some techniques to lower the cost
-of modelling. Although, the modelling of the Z80 do not have modelled some small aspects.
+There are in the literature of computer science some
+approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and the
+virtual machines using the B method. Then, the B method has been used
+successfully to model the operational semantic. However the cost of
+modelling was still expensive and this paper quoted some techniques to
+lower the cost of modelling. Although, the modelling of the Z80 do not
+have modelled some small aspects.
 % The objective of researchers
 
-In general, the researchers~\cite{GEMPLUS_99} are concentrated at model the Java Virtual Machines (JVM), but it is
-not easy to specify, because it is very big and complex. Then, this work specify a model more simple to get better
-results, at the same time, the model stay near to model of actual microcontrollers. The main motivation of our
-research is the verification of assembler level and consequently the verification of hardware model. Thus, the some
-aspects are not modelled. For example, the time of execution of instruction, the pipeline and others questions not
-directly related with the semantic software. However, there are many other specialized techniques to verify these
-questions.
+In general, the researchers~\cite{GEMPLUS_99} are concentrated at
+model the Java Virtual Machines (JVM), but it is not easy to specify,
+because it is very big and complex. Then, this work specify a model
+more simple to get better results, at the same time, the model stay
+near to model of actual microcontrollers. The main motivation of our
+research is the verification of assembler level and consequently the
+verification of hardware model. Thus, the some aspects are not
+modelled. For example, the time of execution of instruction, the
+pipeline and others questions not directly related with the semantic
+software. However, there are many other specialized techniques to
+verify these questions.
 
@@
 \label{sec:conclusions}
-This work has shown an approach to the formal modelling of microcontrollers using the B method. This approach
-provide interesting results and had a great automatization in the prove process. For instance, it is interesting to
-say that: during the building process of the Z80 model we found some errors and ambiguities in the official
-manual\cite{Z80_manual}. The platform model can replace (or improve) the documentation used by assembler
-programmers. Because the formal model represents the instructions effects and the B method has an easy notation,
-similar to pascal language.  Besides, the formal model restricts the definitions to correct typing, uses expressions
-well-defined and allows verifying properties of the model.
+
+This work has shown an approach to the formal modelling of the
+instruction set of microcontrollers using the B method. During the
+construction of this model, several ambiguities and errors were
+encountered in the during the official reference~\cite{Z80_manual}. As
+the B notation has a syntax that is not too distant from that of
+imperative programming languages, such model could be used to improve
+the documentation used by assembler programmers. Besides, the formal
+notation used is analyzed by software that guarantees the
+correctness of typing, the well-definedness of expressions, in addition
+to safety properties of the micro-controller state.
 
@@
 % [Trabalhos futuros ]
-Future works comprise the development of a real study case in the oil area and the project of a compiler to aid the
-approach.
+Future works comprise the development of software with the B method
+from functional specification to assembly level, using the Z80 model
+presented in this paper. The mechanic compilation from B algorithmic 
+constructs to assembly platform is also envisioned.
+
 %  The next steps of the work comprises the development of a real study case in the oil area and the project of a
@@
 \paragraph{Acknowledges:}
-This work received supports from ANP (National Agency of Oil) and CNPq.
-% Besides, the researcher Bruno Gomes help in the revision.
+This work received support from ANP (\textit{Agência Nacional do
+  Petróleo, Gás Natural e Biocombustíveis}) and CNPq.
 

### 2009-06-26 trunk/papers/SBMF2009/source/paper.tex
Initial import.

--- 2009-06-20 papers/SBMF2009/source/paper.tex
@@
 
-This work describes briefly the modelling of microcontroller Z80~\cite{Z80_manual} and some techniques used to build
-it. This modelling\footnote{The interested reader in more details is invited to visit our repository
-at: http://code.google.com/p/b2asm} use the B method~\cite{Abrial}, because it has a great maturity and advanced
-support. Besides, the modelling of microcontroller can extend the level of verification of B method, as it will be
-explained.
+The B method~\cite{Abrial} supports the construction of safety systems
+model by verification of proofs that guarantees its correctness. So,
+an initial abstract model of the system requirements is defined and
+then it is refined until the implementation model. Development
+environments based on the B method also include source code generators
+for several programming languages, but the result of this translation
+cannot be compared by formal means. \cite{Dantas_SBMF08} presented
+recently an approach to extend the scope of the B method up to the
+assembly level language language. One key component of this approach
+is to build, within the framework of the B method, formal modelos of
+the instruction set of such assembly languages.
 
-The B method supports the construction of safety systems model by verification of proofs that guarantees its
-correctness. So, an initial abstract model of the system is defined and refined until the implementation model. The
-B tools  still have the code generator to programming language, but this feature is not verified.
-In~\cite{Dantas_SBMF08} is presented an approach to extend the verification of B method until the assembler language.
-In this approach, the construction of systems begins from the abstract level and it is formally refined until the
-target assembly platform.
+This work gives an overview of the formal modelling of the instruction
+set of the Z80 microcontroller~\cite{Z80_manual}\footnote{The
+  interested reader in more details is invited to visit our repository
+  at: http://code.google.com/p/b2asm.}. Using the responsibility
+division mechanism provided by B, auxiliary libraries of basic modules
+were developed as part of the construction of microcontroller
+model. Such library has many definitions about common concepts used in
+the microcontrollers; besides the Z80 model, it is used by two other
+microcontrollers models that are under way.
 
-Therefore, the first step of this approach is to build the platform modelling, of the microcontroller, in B. Using the
-responsibility division mechanism provided by B, basic modules was developed to be used as elementary library to
-help the constructing of microcontroller model. The elementary library has many definitions about common concepts used
-in the microcontrollers. This library is used by Z80 model and other two model microcontrollers that
-are under construction.
+Other possible uses of a formal model of a micro-controller
+instruction set include documentation, the construction of simulators,
+and be possibly the starting point of a verification effort for the
+actual implementation of a Z80 design. Moreover the model of the
+instruction set could be instrumented with non-functional aspects,
+such as the number of cycles it takes to execute an instruction, to
+prove lower and upper bounds on the execution time of a routine. The
+goal of this project, though, is to provide a basis for the generation
+of software artifacts at the assembly level that are amenable to 
+refinement verification within the B method.
 
-The microcontroller modelling can be used in platform projects documentation, simulators construction, consistency
-verification and software verification at assembly level. The verification at assembly level is more interesting
-utility because it allows the development of trustworthy software by construction, that is extremely desirable in
-embedded critical software.
+This paper is focused on the presentation of the Z80 model, including
+elementary libraries to describe hardware aspects. The paper is
+structured as follows. Section~\ref{sec:B_method} provides a short
+introduction to the B method. Section~\ref{sec:models} presents the
+elementary libraries and the modelling of some elements common to
+microcontrollers.  Section~\ref{sec:z80} presents the B model of the
+Z80 instruction set. Section ~\ref{sec:Proofs} provides some
+information on the proof effort needed to analyze the presented
+models. Related work is discussed in
+Section~\ref{sec:relatedworks}. Finally, the last section is devoted
+to the conclusions.
 
-In this paper we introduced the modelling of Z80 microcontroller, since elementary library until assembly
-instructions. The paper is structured as follows. Section~\ref{sec:B_method} provides a general introduction to the
-B method. Section~\ref{sec:models} presents the elementary libraries and of microcontrollers basic concepts.
-Section~\ref{sec:z80} presents the Z80 B modelling. Section ~\ref{sec:Proofs} explain a little about the proofs of
-models. The next section (\ref{sec:relatedworks}) discusses the related works. Finally, the last section has the
-conclusions.
-
-\section{B Method}
+\section{Introduction to the B Method}
 \label{sec:B_method}
 
-The B method for software development~\cite{Abrial} is based on the B Abstract Machine Notation (AMN) and
-the use of formally proved refinements up to a specification sufficiently concrete that programming     
-code can be automatically generated from it.
-Its mathematical basis consists of first order logic,      integer arithmetic and set theory, and its corresponding constructs are very similar to those of the Z notation
+The B method for software development~\cite{Abrial} is based on the B
+Abstract Machine Notation (AMN) and the use of formally proved
+refinements up to a specification sufficiently concrete that
+programming code can be automatically generated from it. Its
+mathematical basis consists of first order logic, integer arithmetic
+and set theory, and its corresponding constructs are very similar to
+those of the Z notation
 
- A B specification is structured in modules. A module defines a set of valid states, including a set of initial
- states, and operations that may provoke a transition between states. The design process starts with a module with a
- so-called functional model of the system under development. In this initial modelling stage, the B method requires
- that the user proves that, in a machine, all the initial states are valid, and that operations do not define
- transitions from valid states to invalid states.
+A B specification is structured in modules. A module defines a set of
+valid states, including a set of initial states, and operations that
+may provoke a transition between states. The design process starts
+with a module with a so-called functional model of the system under
+development. In this initial modelling stage, the B method requires
+that the user proves that, in a machine, all the initial states are
+valid, and that operations do not define transitions from valid states
+to invalid states.
 
- Essentially, a B module contains two main parts: a header and the available operations, in the
- figure~\ref{fig:maqB} has a intuitive example.
-The clause $\mathit{MACHINE}$ has the name of module.
-The following two are used to access an external module and create an instance of an external module.
-The
- $\mathit{VARIABLES}$  declares the variable names. The following declares its types and the restrictions. The
- $\mathit{INITIALIZATION} $  defines the its initial value. The last define the operations that remember procedures
- from programming language.
+Essentially, a B module contains two main parts: a header and the
+available operations. Figure~\ref{fig:maqB} has a very basic
+example. The clause $\mathit{MACHINE}$ has the name of module.  The
+following two clauses respectively reference external modules and
+create an instance of an external module. The $\mathit{VARIABLES}$
+declares the name of the variables that compose the state of the
+machine. Next, the $\mathit{INVARIANT}$ clause defines the type and
+other restrictions on the variables. The $\mathit{INITIALIZATION} $
+sepecifies the initial states. Finally, operations correspond to
+the transitions between states of the machine.
 
 \begin{figure}
-%\begin{small}
+  % \begin{small}
   $$
-\begin{array}{l}
-\MACHINE \quad  \mathit{micro} \quad \quad \quad \quad \quad  \quad \quad \quad \quad \INITIALISATION  \mathit{pc}
-:= 0\\
-\SEES \quad \mathit{TYPES}, \mathit{ALU} \quad \quad \quad \quad \quad \quad \quad \quad \OPERATIONS\\
-\INCLUDES    \quad \mathit{MEMORY} \quad \quad \quad \quad \quad \quad \quad \mathit{JMP} \mathit{( jump )} =
-\\
-\VARIABLES \quad    \mathit{pc}  \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \PRE \mathit{jump}
-\in\mathit{INSTRUCTION}\\ 
- \INVARIANT \quad \mathit{pc} \in \mathit{INSTRUCTION}  \quad \ \THEN \mathit{pc} :=
-\mathit{jump}\\  
-\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad
-\END \quad  \ldots \\
-% \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad
-%\quad  \ldots
+  \begin{array}{lcl}
+    \begin{array}[t]{l}
+      \MACHINE \quad  \mathit{micro} \\
+      \SEES \quad \mathit{TYPES}, \mathit{ALU} \\
+      \INCLUDES  \quad \mathit{MEMORY} \\
+      \VARIABLES \quad    \mathit{pc}  \\
+      \INVARIANT \quad \mathit{pc} \in \mathit{INSTRUCTION} 
+    \end{array}
+    & \hspace*{1cm} &
+    \begin{array}[t]{l}
+      \INITIALISATION  \mathit{pc} := 0 \\
+      \OPERATIONS\\
+      \mathit{JMP} \mathit{( jump )} = \\
+      \quad \PRE \mathit{jump} \in \mathit{INSTRUCTION}\\ 
+      \quad \THEN \mathit{pc} := \mathit{jump}\\  
+      \quad \END \\
+      \END
+    \end{array}
   \end{array}
   $$
-% \begin{array}{l}
-% \MACHINE \quad \quad \quad \quad \quad \quad  \quad \quad \quad \quad \INITIALISATION \\
-% \quad   \mathit{micro} \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad   \mathit{pc} := 0\\
-% \SEES \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad   \OPERATIONS\\
-% \quad \mathit{TYPES}, \mathit{ALU}\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad \\
-% \INCLUDES                \quad \quad \quad \quad \quad \quad \quad \quad \quad   \quad   \mathit{JMP} \mathit{( jump )} = \\
-% \quad   \mathit{MEMORY}   \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \PRE\\
-% \VARIABLES      \quad \quad \quad \quad \quad \quad \quad \quad \quad  \quad \quad  \mathit{jump} \in
-% \mathit{INSTRUCTION}\\ \quad   \mathit{pc}    \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad\quad \quad \THEN\\
-% \INVARIANT   \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad   \mathit{pc} := \mathit{jump}\\
-% \quad \mathit{pc} \in \mathit{INSTRUCTION} \ \ \quad \quad \quad \quad \quad \quad \END\\
-% \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad
-% \quad \quad  \ldots
-% \end{array}
-
-%\end{small}
-\caption{A part of B simplified module}
+\caption{A very basic B machine.}
 \label{fig:maqB}
@@
 
-\subsection{Definitions to Represent and Manipulate Bits}
+\subsection{Bit representation and manipulation}
 
@@
 
-\subsection{Modelling Bytes and Bit Vector of Length 16}
+\subsection{Modelling Bytes and Bit Vectors of Length 16}
 
@@
 
-The $\mathit{BYTE\_INDEX}$ is domain of byte modelled. It starts at 1 to obey a definition of sequences from B
-method. However, the functions from byte encapsulate the access and these functions use the
-$\mathit{PHYS\_BYTE\_INDEX}$. The $\mathit{BYTE}$ type is a specialized type from $\mathit{BIT\_VECTOR}$, but it has
-a size limit. Some others more specific definitions are created to help the developer. The type $\mathit{BV16}$ is
-created for bit vector of length 16 in a similar way.
+The $\mathit{BYTE\_INDEX}$ is the domain of the functions modelling
+bytes. It starts at 1 to obey a definition of sequences from B
+method. However, it is common in hardware architectures to start
+indexing from zero. The definition $\mathit{PHYS\_BYTE\_INDEX}$ is
+used to provide functionalities obeying this convention. The
+$\mathit{BYTE}$ type is a specialized type from
+$\mathit{BIT\_VECTOR}$, but it has a size limit. Other specific
+definitions are provided to facilitate further modelling: the type
+$\mathit{BV16}$ is created for bit vector of length 16 in a similar
+way.
 
@@
 
-Bit vectors are used to represent and combine numbers: integer ranges (signed or unsigned). Then, our library
-defines functions to manipulate this date, for example, the function $\mathit{bv\_to\_nat}$ that maps bit
-vectors to natural numbers:
+Bit vectors are used to represent and combine numbers: integer ranges
+(signed or unsigned). Therefore, our library includes functions to
+manipulate such data, for example, the function $\mathit{bv\_to\_nat}$
+that maps bit vectors to natural numbers:
 
@@
 
-The instruction set of microcontrollers usually have common data types. These types are placed in the types
-library. Each type module has functions to manipulate and convert its data. There are six common basics data types
-represented by modules, see details in table~\ref{tab:types}.
+The instruction set of microcontrollers usually have common data
+types. These types are placed in the types library. Each type module
+has functions to manipulate and convert its data. There are six common
+basics data types represented by modules, see details in
+table~\ref{tab:types}.
 
@@
 
- Usually, for each type module just needs to detail or redefines the pre-defined concepts. For
- example, the function $\mathit{bv\_to\_nat}$ from bit vector arithmetics that is specialized to
- $\mathit{byte\_uchar}$. The set $\mathit{BYTE}$ is a subset of the
-$\mathit{BIT\_VECTOR}$, then this function can defined as follows:
+Usually, each type module just needs to instantiate concepts that were
+already defined in the hardware modelling library.  For example, the
+function $\mathit{bv\_to\_nat}$ from bit vector arithmetics is
+specialized to $\mathit{byte\_uchar}$. As the set $\mathit{BYTE}$ is a
+subset of the $\mathit{BIT\_VECTOR}$, this function can defined
+as follows:
 
@@
 
-The definitions of the library types reuse the basic definitions from the hardware library. This provides greater
-confidence and facilitates the prove process, because the prover can reuse the
-previous defined lemma.
+The definitions of the library types reuse the basic definitions from
+the hardware library. This provides greater confidence and facilitates
+the proof process, because the prover can reuse the previously defined
+lemma.
 
-The inverse function is easily defined as $\mathit{uchar\_byte}$.
+The inverse function $\mathit{uchar\_byte}$ is easily defined:
 
@@
 
-The \textit{Z80} is an important CISC microcontroller developed by \textit{Zilog}~\cite{Z80_manual}. It supports
-158 different instructions. These instructions are classified into these categories: load and exchange; block transfer
-and search; arithmetic and logical; rotate and shift; bit manipulation (set, reset, test); jump, call and return; input/output; and basic cpu control.
+The \textit{Z80} is a CISC microcontroller developed by
+\textit{Zilog}~\cite{Z80_manual}. It supports 158 different
+instructions. These instructions are classified into these categories:
+load and exchange; block transfer and search; arithmetic and logical;
+rotate and shift; bit manipulation (set, reset, test); jump, call and
+return; input/output; and basic cpu control.
 
-The main module includes an instance of memory module and accesses the definitions from: basic data types modules
-and \textit{ALU}.
+The main module includes an instance of the memory module and accesses the
+definitions from basic data types modules and the \textit{ALU} module.
 
@@
 % Chacteristics of microcontroller
-Each instruction is represented by B operations in the module Z80. By default, all parameters from
-operations are or predefined elements in the model or integers values in the decimal representation.
-The internal registers contain 208 bits of reading/writing memory. It includes two sets of six general purpose
-registers which may be used individually as 8-bits registers or as 16-bits register pairs.  The work registers are
-represented by variable $\mathit{rgs8}$. The domain of $\mathit{rgs8}$ ($\mathit{id\_regs8}$) is a set formed by
-identifiers of registers of 8 bits. These registers can be accessed in pairs, forming 16-bits, resulting in other
-set of identifiers of 16-bits registers, named $\mathit{id\_reg16}$. The main work register of Z80 is the
-accumulator ($\mathit{rgs8(a0)}$) used for arithmetic, logic, input/output and loading/storing operations.
+Each instruction is represented by a B operation in the module Z80. By
+default, all parameters from operations are either predefined elements
+in the model or integers values in the decimal representation.  The
+internal registers contain 208 bits of reading/writing memory. It
+includes two sets of six general purpose registers which may be used
+individually as 8-bits registers or as 16-bits register pairs.  The
+working registers are represented by variable $\mathit{rgs8}$. The
+domain of $\mathit{rgs8}$ ($\mathit{id\_regs8}$) is a set formed by
+identifiers of registers of 8 bits. These registers can be accessed in
+pairs, forming 16-bits, resulting in another set of identifiers of
+16-bits registers, named $\mathit{id\_reg16}$. The main working
+register of Z80 is the accumulator ($\mathit{rgs8(a0)}$) used for
+arithmetic, logic, input/output and loading/storing operations.
+
 \subsection{Modelling registers, input and output ports and instructions}
 
-The Z80 has many register of diferrents types and many instructions. It includes alternative set of accumulator,
-flag and general registers. The CPU contains a stack pointer ($\mathit{sp}$), program counter ($\mathit{pc}$), two
-index registers ($\mathit{ix}$ and $\mathit{iy}$), an interrupt register ($\mathit{i\_}$), a refresh register
-($\mathit{r\_}$), two bits ($\mathit{iff1}$, $\mathit{iff2}$) used to control the interruptions, a pair of bits to
-define the interruption mode ($\mathit{im}$) and the input and output ports ($\mathit{i\_o\_ports}$). Below, its
-definitions are specified in the \textit{INVARIANT}.
+The Z80 has different types of registers and instructions. The CPU
+contains general-purpose registers ($\mathit{id\_reg\_8}$), a stack
+pointer ($\mathit{sp}$), program counter ($\mathit{pc}$), two index
+registers ($\mathit{ix}$ and $\mathit{iy}$), an interrupt register
+($\mathit{i\_}$), a refresh register ($\mathit{r\_}$), two bits
+($\mathit{iff1}$, $\mathit{iff2}$) used to control the interruptions,
+a pair of bits to define the interruption mode ($\mathit{im}$) and the
+input and output ports ($\mathit{i\_o\_ports}$). Below, part of the
+corresponding definitions are replicated from the \textit{INVARIANT}:
 
@@
 \bf INVARIANT
-
 \hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ pc  $\in$  \it INSTRUCTION  $\land$
@@
 
- A simple example of instruction is a $\mathit{LD\_n\_A}$, as shown below.
-Many times, to model an instruction is necessary to use the predefined functions, these help the construction of
-model. This instruction use the $\mathit{updateAddressMem}$ function from \textit{Memory} module and it receives an
-address memory and its new memory value. Finally it increments the program counter ($\mathit{pc}$) and update the
-refresh register ($\mathit{r\_}$).
+A simple example of instruction is a $\mathit{LD\_n\_A}$, as shown
+below.  Many times, to model an instruction is necessary to use the
+predefined functions, these help the construction of model. This
+instruction use the $\mathit{updateAddressMem}$ function from
+\textit{Memory} module and it receives an address memory and its new
+memory value. Finally it increments the program counter
+($\mathit{pc}$) and update the refresh register ($\mathit{r\_}$).
 
@@
 \label{sec:Proofs}
-% [intro/scope] [O que as provas garantem? citar que dificeis, que é necessário buscar técnicas eficientes
+% [O que as provas garantem? citar que dificeis, que é necessário buscar técnicas eficientes
 % para realizar as provas ]
 
- The proof obligations allow to verify the data types, important system properties and if the expressions are
- well-defined (WD)\footnote{An expression is called ``well-defined'' (or unambiguous) if its definition assigns it a
- unique interpretation or value.}. The properties provide additional guarantees, because it can set many safety
- rules. However, the model can be very difficult to prove.
+The proof obligations allow to verify the data types, important system
+properties and if the expressions are well-defined (WD)\footnote{An
+  expression is called ``well-defined'' (or unambiguous) if its
+  definition assigns it a unique interpretation or value.}. The
+properties provide additional guarantees, because it can set many
+safety rules. However, the model can be very difficult to prove.
 
 % [Conclusao] This project has many proof obligations and
-The building of model was a hard and difficult task, because we needed to build, prove, fix and re-prove
-several predicates of the libraries.
+The construction of then model was a difficult task, as several
+iterations were needed to provide the good library definitions 
+as well as to optimize the definition of the functionality of
+the microprocessor instructions by grouping common data
+manipulation into auxiliary functions.
 
-However, few proof commands\footnote{The proof commands are step that just indicate a walk to prover make the proof,
-thus theses commands cannot introduce false hypotheses.} can be useful to prove the majority of the proof
-obligations. Because, there are many similar assembler instructions. A good example is a set of 17 proof commands
-that quickly aided the verification of 99\% (2295) of proofs WD. Finally, the techniques mentioned  help
-the verification of all the 2926 proof obligations.
+However, few proof commands\footnote{The proof commands are steps that
+  direct the prover to find the proof, and cannot introduce false
+  hypothesis.} need to be used to prove most proof obligations. As
+there are many similar assembly instructions, some human-directed
+proofs, when replayed, could discharge other proof obligations. A good
+example is a set of 17 proof commands that quickly aided the
+verification of 99\% (2295) of proofs WD. Finally, all of the 2926
+proof obligations were proved using the tool support of the
+development environment.
 
@@
 
-% [ I need review the english]
+There are in the literature of computer science some
+approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and the
+virtual machines using the B method. Then, the B method has been used
+successfully to model the operational semantic. However the cost of
+modelling was still expensive and this paper quoted some techniques to
+lower the cost of modelling. 
+% Although, the modelling of the Z80 do not
+%have modelled some small aspects.
 
-There are in the literature of computer science some approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and
-the virtual machines using the B method. Then, the B method has been used successfully to model the operational
-semantic. However the cost of modelling was still expensive and this paper quoted some techniques to lower the cost
-of modelling. Although, the modelling of the Z80 do not have modelled some small aspects.
 % The objective of researchers
 
-In general, the researchers~\cite{GEMPLUS_99} are concentrated at model the Java Virtual Machines (JVM), but it is
-not easy to specify, because it is very big and complex. Then, this work specify a model more simple to get better
-results, at the same time, the model stay near to model of actual microcontrollers. The main motivation of our
-research is the verification of assembler level and consequently the verification of hardware model. Thus, the some
-aspects are not modelled. For example, the time of execution of instruction, the pipeline and others questions not
-directly related with the semantic software. However, there are many other specialized techniques to verify these
-questions.
+In general, the researchers employing the B method have focused on
+more abstract level of description of software.  Considering low-level
+aspect, there has been previous work on modelling the Java Virtual
+Machine\cite{GEMPLUS_99}.
+
+The main motivation of our research is the development of verified
+software up to the assembly level, which requires specifying the
+semantics of the underlying hardware. Thus, some aspects were not
+modelled in our work such as the execution time of the instructions.
+Also we did not the micro-architecture of the hardware as the scope of
+our work does not include hardware verification. However, there are
+many other specialized techniques to verify these questions.
 
@@
 \paragraph{Acknowledges:}
-This work received supports from ANP (National Agency of Oil) and CNPq.
-% Besides, the researcher Bruno Gomes help in the revision.
+This work received support from ANP (\textit{Ag\^{e}ncia Nacional do
+  Petr\'{o}leo, G\'{a}s Natural e Biocombust\'{\i}veis}) and CNPq
+(\textit{Conselho Nacional de Desenvolvimento Cient\'{\i}fico e
+  Tecnol\'{o}gico}).
 

### 2009-06-27 trunk/papers/SBMF2009/source/paper.tex
Mais algumas retificações dos revisores realizadas.

@@
-\documentclass[a4paper]{llncs}
+\documentclass{llncs}
 
@@
 \title{Formal Modelling of a Microcontroller Instruction Set in B}
+
+\hyphenation{se-pa-ra-te}
 
@@
 environments based on the B method also include source code generators
-for several programming languages, but the result of this translation
-cannot be compared by formal means. \cite{Dantas_SBMF08} presented
+for programming languages, but the result of this translation
+cannot be compared by formal means. The paper \cite{Dantas_SBMF08} presented
 recently an approach to extend the scope of the B method up to the
-assembly level language language. One key component of this approach
-is to build, within the framework of the B method, formal modelos of
+assembly level language. One key component of this approach
+is to build, within the framework of the B method, formal models of
 the instruction set of such assembly languages.
@@
 
-Other possible uses of a formal model of a micro-controller
+Other possible uses of a formal model of a microcontroller
 instruction set include documentation, the construction of simulators,
@@
 
-The B method for software development~\cite{Abrial} is based on the B
-Abstract Machine Notation (AMN) and the use of formally proved
-refinements up to a specification sufficiently concrete that
-programming code can be automatically generated from it. Its
-mathematical basis consists of first order logic, integer arithmetic
-and set theory, and its corresponding constructs are very similar to
-those of the Z notation
+The B method for software development~\cite{Abrial} is based on the B Abstract
+Machine Notation (AMN) and the use of formally proved refinements up to a
+specification sufficiently concrete that programming code can be automatically
+generated from it. Its mathematical basis consists of first order logic, integer
+arithmetic and set theory, and its corresponding constructs are similar to those
+of the Z notation.
 
@@
 development. In this initial modelling stage, the B method requires
-that the user proves that, in a machine, all the initial states are
+that the user proves that, in a machine, all the its initial states are
 valid, and that operations do not define transitions from valid states
@@
 libraries. A third project is devoted to the higher-level aspects of
-the Z80 model. Thus, the workspace is composed of: a hardware library,
+the platform. Thus, the workspace is composed of: a hardware library,
 a types library and a project for the specific platform, in this case
@@
 \begin{figure}[h] \centering
-\includegraphics[width=.92\textwidth]{diagramaEstrutural_vertical.jpg}
+\includegraphics[width=.92\textwidth]{diagramaEstrutural_vertical2.png}
  \caption{Dependency diagram of the Z80 model.}
@@
 
-\subsection{Bit representation and manipulation}
+\subsection{Bit Representation and Manipulation}
 
@@
 
-The $\mathit{BYTE\_INDEX}$ is the domain of the functions modelling
-bytes. It starts at 1 to obey a definition of sequences from B
-method. However, it is common in hardware architectures to start
-indexing from zero. The definition $\mathit{PHYS\_BYTE\_INDEX}$ is
-used to provide functionalities obeying this convention. The
-$\mathit{BYTE}$ type is a specialized type from
-$\mathit{BIT\_VECTOR}$, but it has a size limit. Other specific
-definitions are provided to facilitate further modelling: the type
-$\mathit{BV16}$ is created for bit vector of length 16 in a similar
-way.
+The $\mathit{BYTE\_INDEX}$ is the domain of the functions modelling bytes. It
+starts at 1 to obey a definition of sequences from B. However, it is common in
+hardware architectures to start indexing from zero. The definition
+$\mathit{PHYS\_BYTE\_INDEX}$ is used to provide functionalities obeying this
+convention. The $\mathit{BYTE}$ type is a specialized type from
+$\mathit{BIT\_VECTOR}$, but it has a size limit. Other specific definitions are
+provided to facilitate further modelling: the type $\mathit{BV16}$ is created for
+bit vector of length 16 in a similar way.
 
@@
 The \textit{Z80} is a CISC microcontroller developed by
-\textit{Zilog}~\cite{Z80_manual}. It supports 158 different
-instructions. These instructions are classified into these categories:
-load and exchange; block transfer and search; arithmetic and logical;
-rotate and shift; bit manipulation (set, reset, test); jump, call and
-return; input/output; and basic cpu control.
+\textit{Zilog}~\cite{Z80_manual}. It supports 158 different instructions and all
+of them were specified. These instructions are classified into these categories:
+load and exchange; block transfer and search; arithmetic and logical; rotate and
+shift; bit manipulation; jump, call and return; input/output; and basic cpu
+control.
 
@@
 
-\subsection{Modelling registers, input and output ports and instructions}
+\subsection{Modelling Registers, Input and Output Ports and Instructions}
 
@@
 input and output ports ($\mathit{i\_o\_ports}$). Below, part of the
-corresponding definitions are replicated from the \textit{INVARIANT}:
+corresponding definitions are replicated from the $\INVARIANT$:
 
 \begin{sloppypar}
-\bf INVARIANT
+%\bf INVARIANT
 \hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ pc  $\in$  \it INSTRUCTION  $\land$
@@
 
-The microcontroller model can specify several security properties. For example, the last operation could have a
-restriction to write only in a defined region of memory.
+The microcontroller model can specify security properties. For example, the
+last operation could have a restriction to write only in a defined region of memory.
 
@@
 
-However, few proof commands\footnote{The proof commands are steps that
-  direct the prover to find the proof, and cannot introduce false
-  hypothesis.} need to be used to prove most proof obligations. As
-there are many similar assembly instructions, some human-directed
-proofs, when replayed, could discharge other proof obligations. A good
-example is a set of 17 proof commands that quickly aided the
-verification of 99\% (2295) of proofs WD. Finally, all of the 2926
-proof obligations were proved using the tool support of the
-development environment.
+However, few proof commands\footnote{The proof commands are steps that direct the
+prover to find the proof, and cannot introduce false hypothesis.} need to be used
+to prove most proof obligations. As there are many similar assembly instructions,
+some human-directed proofs, when replayed, could discharge other proof
+obligations. A good example is a set of 17 proof commands that quickly aided the
+verification of 99\% (2295) of proofs WD. A new feature of B environment also
+was used, the parallelization. This used four computers each one with two cores and
+it increased many times the capacity of data processing. Finally, all of the
+2926 proof obligations were proved using the tool support of the development
+environment.
 
@@
 There are in the literature of computer science some
-approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and the
-virtual machines using the B method. Then, the B method has been used
-successfully to model the operational semantic. However the cost of
-modelling was still expensive and this paper quoted some techniques to
-lower the cost of modelling. 
-% Although, the modelling of the Z80 do not
-%have modelled some small aspects.
+approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and the virtual machines
+using the B method. Then, the B method has been used successfully to model the
+operational semantic. However the cost of modelling was still expensive and this
+paper quoted some techniques to lower the cost of modelling.
+% Although, the modelling of the Z80 do not have modelled some small aspects.
 
@@
 aspect, there has been previous work on modelling the Java Virtual
-Machine\cite{GEMPLUS_99}.
+Machine~\cite{GEMPLUS_99}.
 
@@
 instruction set of microcontrollers using the B method. During the
-construction of this model, several ambiguities and errors were
-encountered in the during the official reference~\cite{Z80_manual}. As
+construction of this model, some ambiguities and errors were
+encountered in the official reference~\cite{Z80_manual}. As
 the B notation has a syntax that is not too distant from that of
@@
 correctness of typing, the well-definedness of expressions, in addition
-to safety properties of the micro-controller state.
+to safety properties of the microcontroller state.
 
@@
 \bibitem {BHDL_2003}    
-Aljer, P. Devienne, S. Tison and J-L. Boulanger andG. Mariano. Bhdl: Circuit Design in B. A. In
+Aljer, P. Devienne; S. Tison  J-L. Boulanger and G. Mariano. Bhdl: Circuit Design in B. A. In
 ACSD, Third International Conference on Application of Concurrency to System Design, pages 241-242, 2003.
@@
 \bibitem {Dantas_SBMF08}
-Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; Et al. Applying the B Method to Take on the Grand Challenge of
+Dantas, B; D\'{e}harbe, D.; Galv\~{a}o, S. L.; Moreira, A. M. and
+Medeiros Jr, V. G.. Applying the B Method to Take on the Grand Challenge of
 Verified Compilation. In: SBMF, Savaldor, 2008. SBC.
@@
 \bibitem {HOARE}
-Hoare, C. A. R. The verifying compiler, a grand challenge for computing research. In: VMCAI, p. 78-78, 2005.
+Hoare, C. A. R. The verifying compiler, a grand challenge for computing research.
+In: VMCAI, p. 78-78, 2005.
 
 \bibitem{Z80_manual}
-Zilog. Z80 Family CPU User Manual. www.zilog.com/docs/z80/um0080.pdf
+Zilog. Z80 Family CPU User Manual. http://www.zilog.com/docs/z80/um0080.pdf
 

### 2009-06-27 papers/SBMF2009/source/observacoes_revisores.tex
algumas observações.

---------------------------------------------

Paper: 25 Title: A Formal Model of Microcontroller Instruction Set in B

-------------------- review 1 --------------------

PAPER: 25 TITLE: A Formal Model of Microcontroller Instruction Set in B

OVERALL RATING: 1 (weak accept) ----------------------- REVIEW
--------------------

The paper describes a B model of the Z80 microcontroller. The major goal of the
project is to be able to verify software at the assembly level.

The project, its current achievements and its scope seem to be adequate to the
thesis & dissertation workshop.

Suggestions:
 - The paper needs an extensive and careful review of the English
language. 
[Res: Vou levar p/ um outro revisor]
 - The authors must look at other previous work on formal verification
of microcontrollers, processors, architectures and assembly code (regardless the
formal language used and the processor verified). For instance, the work by
Anthony Fox in the verification of ARM can be a good starting point. His
technical reports cite many other related work (Sawada and Hunt, Tahar and Kumar,
Windley and Coe, etc.)
[Res: Vou levar p/ um outro revisor]
 - Sections 6 and 7 contain several claims that must be better explained. "this paper quoted some
 techniques to lower the cost of modelling" (which ones?) "This work specify a model more simple to get better
results, at the same time, the model stay near to model of actual
microcontrollers." (why is your model simpler? what makes it simpler and still
closer to the microcontrollers in comparison to previous work?) "B method has an
easy notation, similar to the Pascal language" (I don't think B and Pascal are
similar) 
[Res: V]
- "we found some errors and ambiguities in the official manual" This is
an excellent contribution that the paper overlooks. - In reference 4, do not use
"Et al."

-------------------- review 2 --------------------

PAPER: 25 TITLE: A Formal Model of Microcontroller Instruction Set in B

OVERALL RATING: 2 (accept) ----------------------- REVIEW --------------------

The paper describes an interesting encoding of the Z80 microcontroller model
using the B method. This encoding is useful for proving a number of properties
about the microcontroller. They found some errors and ambiguities in the
microcontroller official manual. This is a very nice result and shows the
importance of formally specifying and proving things using a theorem prover.

However, the paper is not well-written. Run the spell checker and rewrite some
sentences. In the abstract and introduction, you should describe more about the
context and show the importance of formalizing the Z80 model.
[Res: Não encontrei erros reais no spell checker, mas vou levar o trabalho para um revisor!]
[O contexto é a documentação das instruções/especificação e a principal importancia é a verificação em]
[nível de assembly. Tudo isso é esclarecido de forma suscinta devido ao espaço. ]
Additionally, in the conclusions, discuss more and describe the experience of formalizing the
model and proving properties of it. It would be nice if you give more details
about your dissertation (schedule, the main goal,…).
[Res: A experiencia sobre o processo de prova é descrito suscintamente na seção Proofs]
[Vou enfatizar que os trabalhos futuros será elaborado na dissertação]

(page 1) “… two model microcontrollers that are under construction …” Do you
reuse some parts of the model? And proofs? Evolving specifications is not an easy
task since some previous proofs may be proved again, as you mention in Section 5
(you need to reprove some lemmas). A simple renaming may have an impact on
several proofs. Does the B tool have some support in this scenario? Do you have
some guidelines in order to minimize the impact during evolution?
[Res - Vou adicionar: Esses projetos estão estruturado de um modo que pode facilmente]
[ ser reusado por um plataforma diferente ]

(page 3) “… we have been developed a reusable set of basic definitions …” A
well-structured specification will allow you to improve reusability (and also
reuse some proofs). Do you structure your specification in a way that can
minimize the impact?
[Res: Praticamente não deve existir impacto nas bibliotecas, quando uma nova plataforma é reusa isso. ]
[ A resposta anterior já trata isso!]
Does the B prover allow the user to define proof strategies (a user defined proof
command that is created based on other ones)?
[Res: O provador tem estratégia de descrever comandos que podem ser aplicados a várias oP diferente]
[,desde que tenha forma semelhantes. Inf] Proving theorems in a theorem prover is not an easy task, 
and your experience can be very useful. This is a good topic to explore more and discuss your experience.
[Res- Vou adicionar: Sim! Foi utilizado quatro computadores dual-core para realizar as provas  ]

Discuss whether the lemmas you proved were difficult. Does the B theorem prover
help you to prove most of them automatically? Show more details about the
automation level of the prover. Do you have some experience in other theorem
provers (Z-Eves, PVS, Isabelle,…)? In case you have, can you compare them to the
B prover?
[Res - Vou add: Sem usar comandos de prova, em geral 70 foi automatica  ,]
[Não tenho experiência em outros provadores]

(page 6) “… It supports 158 instructions …” Do you specify all of them?
[Res - Adicionar: YES! :D]

(page 8) “… Thus, some aspects are not modeled …” What is the impact of not
considering some aspects in the results of your proofs? You must give more
details about it. Moreover, you should precisely state all aspects that you do
not consider.
[Res - Adicionar: Por exemplo, tempo e detales architecture, porque eles não são relevantes para
verificação assembly]
What are the benefits of your approach for a system using this microcontroller?
[Res: O texto esclarece que é mais fácil que tentar inicialmente em um processo real]

(page 8) “… we found some errors and ambiguities in the official manual …”
Describe some kinds of errors. This is a very nice result and you must give much
more details.
[Res: O paper só tem oito páginas :( ]

It is important to mention some details about what are you going to do in your
dissertation, mention the main goal of it, and the results that you are aiming
at. It would be nice to precisely state the problem you are tackling.
[Res Adicionar: Vou retificar que os trabalhos futuros serão feitos na minha dissertação]

Minor comments (page 2) 
“… the Z notation …” -> missing (.) 
(page 2) 
“… Section 2 explain …” 
(page 2) “… the figure 1 …”  [NÃO ENTENDI ESSE COMENTÁRIO] 
(page 2) “… the its initial …” 
(page 6) “…from operations are or predefined…”

In Section 3, it is better to first explain the microcontroller and then what you
have done. In the beginning of each section, it is important to relate it with
the previous and following sections.
[Res: Como o espaço é pequeno e o modelo do microcontroller usa os conceitos das biblioteca,]
[Então, eu explico primeiro as blibliotecas que fiz, depois o modelo do microcontroller.]

(page 5) “… The type BV16 is created for bit vector of length 16 …” Where is this
type used? For each type declaration, give an intuition specifying which part of
the microcontroller it is representing.
[Res - Adicionar: Vou adicionar o módulo BV16 e BYTES na figura 1. Não entendi bem ]
[Mas todos os tipos citado são utilizados pelo microcontrolador.]

### 2009-06-29 trunk/papers/SBMF2009/source/paper.tex
Versão re-re-visada.

@@
 % [Conclusao] This project has many proof obligations and
-The construction of then model was a difficult task, as several
-iterations were needed to provide the good library definitions 
-as well as to optimize the definition of the functionality of
-the microprocessor instructions by grouping common data
-manipulation into auxiliary functions.
+% The construction of then model was a difficult task, as several
+% iterations were needed to provide the good library definitions 
+% as well as to optimize the definition of the functionality of
+% the microprocessor instructions by grouping common data
+% manipulation into auxiliary functions.
+
+Several iterations were needed to provide the good library definitions as well as
+to fine-tune the model of the microprocessor instructions by factoring common
+functionalities into auxiliary definitions.
 
@@
 obligations. A good example is a set of 17 proof commands that quickly aided the
-verification of 99\% (2295) of proofs WD. A new feature of B environment also
-was used, the parallelization. This used four computers each one with two cores and
-it increased many times the capacity of data processing. Finally, all of the
+verification of 99\% (2295) of proofs WD. We also set up a proving environment
+consisting of networked computers to take advantage of the distribution
+facilities now provided in the B development environment. Finally, all of the
 2926 proof obligations were proved using the tool support of the development

### 2009-06-29 trunk/papers/SBMF2009/source/paper.tex
Novamente substitui microprocessor por microcontroller

@@
 Several iterations were needed to provide the good library definitions as well as
-to fine-tune the model of the microprocessor instructions by factoring common
+to fine-tune the model of the microcontroller instructions by factoring common
 functionalities into auxiliary definitions.

### 2009-07-01 trunk/papers/SBMF2009/source/paper.tex
Final submetido

@@
 The B method~\cite{Abrial} supports the construction of safety systems
-model by verification of proofs that guarantees its correctness. So,
+models by verification of proofs that guarantees its correctness. So,
 an initial abstract model of the system requirements is defined and
@@
 
-A B specification is structured in modules. A module defines a set of
-valid states, including a set of initial states, and operations that
-may provoke a transition between states. The design process starts
-with a module with a so-called functional model of the system under
-development. In this initial modelling stage, the B method requires
-that the user proves that, in a machine, all the its initial states are
-valid, and that operations do not define transitions from valid states
-to invalid states.
+A B specification is structured in modules. A module defines a set of valid
+states, including a set of initial states, and operations that may provoke a
+transition between states. The design process starts with a module with a
+so-called functional model of the system under development. In this initial
+modelling stage, the B method requires that the user proves that, in a machine,
+all the its initial states are valid, and that operations do not define
+transitions from valid states to invalid states.
 
-Essentially, a B module contains two main parts: a header and the
-available operations. Figure~\ref{fig:maqB} has a very basic
-example. The clause $\mathit{MACHINE}$ has the name of module.  The
-following two clauses respectively reference external modules and
-create an instance of an external module. The $\mathit{VARIABLES}$
-declares the name of the variables that compose the state of the
-machine. Next, the $\mathit{INVARIANT}$ clause defines the type and
-other restrictions on the variables. The $\mathit{INITIALIZATION} $
-sepecifies the initial states. Finally, operations correspond to
-the transitions between states of the machine.
+Essentially, a B module contains two main parts: a header and the available
+operations. Figure~\ref{fig:maqB} has a very basic example. The clause
+$\mathit{MACHINE}$ has the name of module.  The next two clauses respectively
+reference external modules and create an instance of an external module. The
+$\mathit{VARIABLES}$ clauses declares the name of the variables that compose the
+state of the machine. Next, the $\mathit{INVARIANT}$ clause defines the type and
+other restrictions on the variables. The $\mathit{INITIALIZATION} $ sepecifies
+the initial states. Finally, operations correspond to the transitions between
+states of the machine.
 
@@
 \begin{figure}[h] \centering
-\includegraphics[width=.92\textwidth]{diagramaEstrutural_vertical2.png}
+\includegraphics[width=.92\textwidth]{diagramaEstrutural.png}
  \caption{Dependency diagram of the Z80 model.}
@@
   definition assigns it a unique interpretation or value.}. The
-properties provide additional guarantees, because it can set many
+properties provide additional guarantees, because they can set many
 safety rules. However, the model can be very difficult to prove.
@@
 obligations. A good example is a set of 17 proof commands that quickly aided the
-verification of 99\% (2295) of proofs WD. We also set up a proving environment
+verification of 99\% (2295) of WD proofs. We also set up a proving environment
 consisting of networked computers to take advantage of the distribution
@@
 approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and the virtual machines
-using the B method. Then, the B method has been used successfully to model the
-operational semantic. However the cost of modelling was still expensive and this
+using the B method. Then, in both works the B method has been used successfully
+to model the operational semantic. However the cost of modelling was still expensive and this
 paper quoted some techniques to lower the cost of modelling.
@@
 modelled in our work such as the execution time of the instructions.
-Also we did not the micro-architecture of the hardware as the scope of
+Also we did not consider the microarchitecture of the hardware as the scope of
 our work does not include hardware verification. However, there are
@@
 
-This work has shown an approach to the formal modelling of the
-instruction set of microcontrollers using the B method. During the
-construction of this model, some ambiguities and errors were
-encountered in the official reference~\cite{Z80_manual}. As
-the B notation has a syntax that is not too distant from that of
-imperative programming languages, such model could be used to improve
-the documentation used by assembler programmers. Besides, the formal
-notation used is analyzed by software that guarantees the
-correctness of typing, the well-definedness of expressions, in addition
-to safety properties of the microcontroller state.
+This work has shown an approach to the formal modelling of the instruction set of
+microcontrollers using the B method. During the construction of this model, some
+ambiguities and errors were encountered in the official reference for Z80
+microcontroller~\cite{Z80_manual}. As the B notation has a syntax that is not too
+distant from that of imperative programming languages, such model could be used
+to improve the documentation used by assembler programmers. Besides, the formal
+notation used is analyzed by software that guarantees the correctness of typing,
+the well-definedness of expressions, in addition to safety properties of the
+microcontroller state.
 

### 2009-07-01 papers/SBMF2009/source/paper.tex
Final submetido

@@
 
-\section{B Method}
+\section{Introduction to the B Method}
 \label{sec:B_method}
 
-The B method for software development~\cite{Abrial} is based on the B
-Abstract Machine Notation (AMN) and the use of formally proved
-refinements up to a specification sufficiently concrete that
-programming code can be automatically generated from it. Its
-mathematical basis consists of first order logic, integer arithmetic
-and set theory, and its corresponding constructs are very similar to
-those of the Z notation
+The B method for software development~\cite{Abrial} is based on the B Abstract
+Machine Notation (AMN) and the use of formally proved refinements up to a
+specification sufficiently concrete that programming code can be automatically
+generated from it. Its mathematical basis consists of first order logic, integer
+arithmetic and set theory, and its corresponding constructs are similar to those
+of the Z notation.
 
-A B specification is structured in modules. A module defines a set of
-valid states, including a set of initial states, and operations that
-may provoke a transition between states. The design process starts
-with a module with a so-called functional model of the system under
-development. In this initial modelling stage, the B method requires
-that the user proves that, in a machine, all the initial states are
-valid, and that operations do not define transitions from valid states
-to invalid states.
+A B specification is structured in modules. A module defines a set of valid
+states, including a set of initial states, and operations that may provoke a
+transition between states. The design process starts with a module with a
+so-called functional model of the system under development. In this initial
+modelling stage, the B method requires that the user proves that, in a machine,
+all the its initial states are valid, and that operations do not define
+transitions from valid states to invalid states.
 
-Essentially, a B module contains two main parts: a header and the
-available operations. Figure~\ref{fig:maqB} has a very basic
-example. The clause $\mathit{MACHINE}$ has the name of module.  The
-following two clauses respectively reference external modules and
-create an instance of an external module. The $\mathit{VARIABLES}$
-declares the name of the variables that compose the state of the
-machine. Next, the $\mathit{INVARIANT}$ clause defines the type and
-other restrictions on the variables. The $\mathit{INITIALIZATION} $
-sepecifies the initial states. Finally, operations correspond to
-the transitions between states of the machine.
+Essentially, a B module contains two main parts: a header and the available
+operations. Figure~\ref{fig:maqB} has a very basic example. The clause
+$\mathit{MACHINE}$ has the name of module.  The next two clauses respectively
+reference external modules and create an instance of an external module. The
+$\mathit{VARIABLES}$ clauses declares the name of the variables that compose the
+state of the machine. Next, the $\mathit{INVARIANT}$ clause defines the type and
+other restrictions on the variables. The $\mathit{INITIALIZATION} $ specifies
+the initial states. Finally, operations correspond to the transitions between
+states of the machine.
 
@@
 \begin{figure}[h] \centering
-\includegraphics[width=.92\textwidth]{diagramaEstrutural_vertical.jpg}
+\includegraphics[width=.92\textwidth]{diagramaEstrutural.png}
  \caption{Dependency diagram of the Z80 model.}
@@
 
-\subsection{Definitions to Represent and Manipulate Bits}
+\subsection{Bit Representation and Manipulation}
 
@@
 
-The $\mathit{BYTE\_INDEX}$ is domain of byte modelled. It starts at 1 to obey a definition of sequences from B
-method. However, the functions from byte encapsulate the access and these functions use the
-$\mathit{PHYS\_BYTE\_INDEX}$. The $\mathit{BYTE}$ type is a specialized type from $\mathit{BIT\_VECTOR}$, but it has
-a size limit. Some others more specific definitions are created to help the developer. The type $\mathit{BV16}$ is
-created for bit vector of length 16 in a similar way.
+The $\mathit{BYTE\_INDEX}$ is the domain of the functions modelling bytes. It
+starts at 1 to obey a definition of sequences from B. However, it is common in
+hardware architectures to start indexing from zero. The definition
+$\mathit{PHYS\_BYTE\_INDEX}$ is used to provide functionalities obeying this
+convention. The $\mathit{BYTE}$ type is a specialized type from
+$\mathit{BIT\_VECTOR}$, but it has a size limit. Other specific definitions are
+provided to facilitate further modelling: the type $\mathit{BV16}$ is created for
+bit vector of length 16 in a similar way.
 
@@
 The \textit{Z80} is a CISC microcontroller developed by
-\textit{Zilog}~\cite{Z80_manual}. It supports 158 different
-instructions. These instructions are classified into these categories:
-load and exchange; block transfer and search; arithmetic and logical;
-rotate and shift; bit manipulation (set, reset, test); jump, call and
-return; input/output; and basic cpu control.
+\textit{Zilog}~\cite{Z80_manual}. It supports 158 different instructions and all
+of them were specified. These instructions are classified into these categories:
+load and exchange; block transfer and search; arithmetic and logical; rotate and
+shift; bit manipulation; jump, call and return; input/output; and basic cpu
+control.
 
-The main module includes an instance of memory module and accesses the definitions from: basic data types modules
-and \textit{ALU}.
+The main module includes an instance of the memory module and accesses the
+definitions from basic data types modules and the \textit{ALU} module.
 
@@
 % Chacteristics of microcontroller
-Each instruction is represented by B operations in the module Z80. By default, all parameters from
-operations are or predefined elements in the model or integers values in the decimal representation.
-The internal registers contain 208 bits of reading/writing memory. It includes two sets of six general purpose
-registers which may be used individually as 8-bits registers or as 16-bits register pairs.  The work registers are
-represented by variable $\mathit{rgs8}$. The domain of $\mathit{rgs8}$ ($\mathit{id\_regs8}$) is a set formed by
-identifiers of registers of 8 bits. These registers can be accessed in pairs, forming 16-bits, resulting in other
-set of identifiers of 16-bits registers, named $\mathit{id\_reg16}$. The main work register of Z80 is the
-accumulator ($\mathit{rgs8(a0)}$) used for arithmetic, logic, input/output and loading/storing operations.
-\subsection{Modelling registers, input and output ports and instructions}
+Each instruction is represented by a B operation in the module Z80. By
+default, all parameters from operations are either predefined elements
+in the model or integers values in the decimal representation.  The
+internal registers contain 208 bits of reading/writing memory. It
+includes two sets of six general purpose registers which may be used
+individually as 8-bits registers or as 16-bits register pairs.  The
+working registers are represented by variable $\mathit{rgs8}$. The
+domain of $\mathit{rgs8}$ ($\mathit{id\_regs8}$) is a set formed by
+identifiers of registers of 8 bits. These registers can be accessed in
+pairs, forming 16-bits, resulting in another set of identifiers of
+16-bits registers, named $\mathit{id\_reg16}$. The main working
+register of Z80 is the accumulator ($\mathit{rgs8(a0)}$) used for
+arithmetic, logic, input/output and loading/storing operations.
 
-The Z80 has many register of differents types and many instructions. It includes alternative set of accumulator,
-flag and general registers. The CPU contains a stack pointer ($\mathit{sp}$), program counter ($\mathit{pc}$), two
-index registers ($\mathit{ix}$ and $\mathit{iy}$), an interrupt register ($\mathit{i\_}$), a refresh register
-($\mathit{r\_}$), two bits ($\mathit{iff1}$, $\mathit{iff2}$) used to control the interruptions, a pair of bits to
-define the interruption mode ($\mathit{im}$) and the input and output ports ($\mathit{i\_o\_ports}$). Below, its
-definitions are specified in the \textit{INVARIANT}.
+\subsection{Modelling Registers, Input and Output Ports and Instructions}
+
+The Z80 has different types of registers and instructions. The CPU
+contains general-purpose registers ($\mathit{id\_reg\_8}$), a stack
+pointer ($\mathit{sp}$), program counter ($\mathit{pc}$), two index
+registers ($\mathit{ix}$ and $\mathit{iy}$), an interrupt register
+($\mathit{i\_}$), a refresh register ($\mathit{r\_}$), two bits
+($\mathit{iff1}$, $\mathit{iff2}$) used to control the interruptions,
+a pair of bits to define the interruption mode ($\mathit{im}$) and the
+input and output ports ($\mathit{i\_o\_ports}$). Below, part of the
+corresponding definitions are replicated from the $\INVARIANT$:
 
 \begin{sloppypar}
-\bf INVARIANT
-
+%\bf INVARIANT
 \hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ pc  $\in$  \it INSTRUCTION  $\land$
@@
 
- The proof obligations allow to verify the data types, important system properties and if the expressions are
- well-defined (WD)\footnote{An expression is called ``well-defined'' (or unambiguous) if its definition assigns it a
- unique interpretation or value.}. The properties provide additional guarantees, because it can set many safety
- rules. However, the model can be very difficult to prove.
+The proof obligations allow to verify the data types, important system
+properties and if the expressions are well-defined (WD)\footnote{An
+  expression is called ``well-defined'' (or unambiguous) if its
+  definition assigns it a unique interpretation or value.}. The
+properties provide additional guarantees, because they can set many
+safety rules. However, the model can be very difficult to prove.
 
 % [Conclusao] This project has many proof obligations and
-The building of model was a hard and difficult task, because we needed to build, prove, fix and re-prove
-several predicates of the libraries.
+% The construction of then model was a difficult task, as several
+% iterations were needed to provide the good library definitions 
+% as well as to optimize the definition of the functionality of
+% the microprocessor instructions by grouping common data
+% manipulation into auxiliary functions.
 
-However, few proof commands\footnote{The proof commands are step that just indicate a walk to prover make the proof,
-thus theses commands cannot introduce false hypotheses.} can be useful to prove the majority of the proof
-obligations. Because, there are many similar assembler instructions. A good example is a set of 17 proof commands
-that quickly aided the verification of 99\% (2295) of proofs WD. Finally, the techniques mentioned  help
-the verification of all the 2926 proof obligations.
+Several iterations were needed to provide the good library definitions as well as
+to fine-tune the model of the microcontroller instructions by factoring common
+functionalities into auxiliary definitions.
+
+However, few proof commands\footnote{The proof commands are steps that direct the
+prover to find the proof, and cannot introduce false hypothesis.} need to be used
+to prove most proof obligations. As there are many similar assembly instructions,
+some human-directed proofs, when replayed, could discharge other proof
+obligations. A good example is a set of 17 proof commands that quickly aided the
+verification of 99\% (2295) of WD proofs. We also set up a proving environment
+consisting of networked computers to take advantage of the distribution
+facilities now provided in the B development environment. Finally, all of the
+2926 proof obligations were proved using the tool support of the development
+environment.
 
@@
 There are in the literature of computer science some
-approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and the
-virtual machines using the B method. Then, the B method has been used
-successfully to model the operational semantic. However the cost of
-modelling was still expensive and this paper quoted some techniques to
-lower the cost of modelling. Although, the modelling of the Z80 do not
-have modelled some small aspects.
+approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and the virtual machines
+using the B method. Then, in both works the B method has been used successfully
+to model the operational semantic. However the cost of modelling was still expensive and this
+paper quoted some techniques to lower the cost of modelling.
+% Although, the modelling of the Z80 do not have modelled some small aspects.
+
 % The objective of researchers
 
-In general, the researchers~\cite{GEMPLUS_99} are concentrated at
-model the Java Virtual Machines (JVM), but it is not easy to specify,
-because it is very big and complex. Then, this work specify a model
-more simple to get better results, at the same time, the model stay
-near to model of actual microcontrollers. The main motivation of our
-research is the verification of assembler level and consequently the
-verification of hardware model. Thus, the some aspects are not
-modelled. For example, the time of execution of instruction, the
-pipeline and others questions not directly related with the semantic
-software. However, there are many other specialized techniques to
-verify these questions.
+In general, the researchers employing the B method have focused on
+more abstract level of description of software.  Considering low-level
+aspect, there has been previous work on modelling the Java Virtual
+Machine~\cite{GEMPLUS_99}.
+
+The main motivation of our research is the development of verified
+software up to the assembly level, which requires specifying the
+semantics of the underlying hardware. Thus, some aspects were not
+modelled in our work such as the execution time of the instructions.
+Also we did not consider the microarchitecture of the hardware as the scope of
+our work does not include hardware verification. However, there are
+many other specialized techniques to verify these questions.
 
@@
 
-This work has shown an approach to the formal modelling of the
-instruction set of microcontrollers using the B method. During the
-construction of this model, several ambiguities and errors were
-encountered in the during the official reference~\cite{Z80_manual}. As
-the B notation has a syntax that is not too distant from that of
-imperative programming languages, such model could be used to improve
-the documentation used by assembler programmers. Besides, the formal
-notation used is analyzed by software that guarantees the
-correctness of typing, the well-definedness of expressions, in addition
-to safety properties of the micro-controller state.
+This work has shown an approach to the formal modelling of the instruction set of
+microcontrollers using the B method. During the construction of this model, some
+ambiguities and errors were encountered in the official reference for Z80
+microcontroller~\cite{Z80_manual}. As the B notation has a syntax that is not too
+distant from that of imperative programming languages, such model could be used
+to improve the documentation used by assembler programmers. Besides, the formal
+notation used is analyzed by software that guarantees the correctness of typing,
+the well-definedness of expressions, in addition to safety properties of the
+microcontroller state.
 
@@
 \paragraph{Acknowledges:}
-This work received support from ANP (\textit{Agência Nacional do
-  Petróleo, Gás Natural e Biocombustíveis}) and CNPq.
+This work received support from ANP (\textit{Ag\^{e}ncia Nacional do
+  Petr\'{o}leo, G\'{a}s Natural e Biocombust\'{\i}veis}) and CNPq
+(\textit{Conselho Nacional de Desenvolvimento Cient\'{\i}fico e
+  Tecnol\'{o}gico}).
 

### 2009-08-06 papers/SBMF2009/Presentation/content/1_introduction.tex

\section{Introduction}

\begin{frame}
\frametitle{Overview}

\begin{itemize}[<+->]
  \item The B method supports the construction of safety systems
  \begin{itemize}
    \item Until the implementation model
    \item But the result of this translation is not guaranteed by formal means
    \item Then the transformation process can introduce small bugs\\
       \includegraphics[width=.7\textwidth]{figures/passos_de_desenvolvimento_tradicional.png}
  \end{itemize}
  \item The last year a paper [Dantas, 2008] proposed a approach to extend the formal verification
  \item One key of this approach is the formal model of the instruction set  of execution platform % of such assembly languages
\end{itemize}
% TO-DO Ajustar a figura

	\note{ ................... }
\end{frame}

\begin{frame}
\frametitle{Overview}

\begin{itemize}[<+->]
  \item The B method supports the construction of safety systems
  \begin{itemize}
    \item Until the implementation model
    \item But the result of this translation is not guaranteed by formal means
    \item Then the transformation process can introduce small bugs\\
       \includegraphics[width=.7\textwidth]{figures/passos_de_desenvolvimento_extendido.png}
  \end{itemize}
  \item The last year a paper [Dantas, 2008] proposed a approach to extend the formal verification
  \item One key of this approach is the formal model of the instruction set  of execution platform % of such assembly languages
\end{itemize}
% TO-DO Ajustar a figura

	\note{ ................... }
\end{frame}

\begin{frame}
\frametitle{Modelling}

\begin{itemize}[<+->]
  \item The modelling is builded using some libraries developed
  \item The libraries has common concepts that can be used to others platforms too
  \item Utilities of this Formal Model:
  \begin{itemize}
    \item Documentation
    \item Simulation
    \item The formal verification until the assembly level
    \item A possible support to verify the Z80 design

  \end{itemize}
\end{itemize}

	\note{ ................... }
\end{frame}

\begin{frame}
\frametitle{Objective}

\begin{itemize}[<+->]
  \item The main actual objective is represent the instruction set
    \begin{itemize}
    \item The aspects no-functional are not represented: logic circuits, pipeline, data bus, \ldots 
  \end{itemize}
  \item Because there are many efficient ways to verify hardware

\end{itemize}
%

	\note{ ................... }
\end{frame}

### 2009-08-06 papers/SBMF2009/Presentation/content/3_Modelling.tex

\section{Modelling the Instruction Set in B}
\begin{frame}
  \frametitle{Modelling  Microcontrollers}

  \begin{itemize}[<+->]
  \item Relations 
  \begin{itemize}
  	\item States of platform  $\Leftrightarrow$ States of modell
		\begin{itemize}
		\item Registers and Memory $\Leftrightarrow$  Variables of model
		\end{itemize}
	\item Assembly Instructions $\Leftrightarrow$ B operations
        \begin{itemize}
        \item Changes of states of platform $\Leftrightarrow$ B substitutions
        \end{itemize}
  \end{itemize}

 \end{itemize}
\end{frame}

\begin{frame}
 \frametitle{Structure of Projects B}
 \begin{center}\includegraphics[width=1.\textwidth]{figures/diagramaEstrutural_vertical.png}
 \end{center}

 \end{frame}

\begin{frame}
 \frametitle{Project Hardware Library}

  	  	\begin{itemize}

  	  		\item Definition of \textit{Bit}
  			{\small
 			$
 			\begin{array}{l}
 			\mathit{BIT} = 0..1 \\
 			\end{array}
 			$
 			}

 			\item The function $\mathit{bit\_not}$ is a unary function:
			{\small
			$
			\begin{array}{l}
			\mathit{bit\_not}  \in  \mathit{BIT}  \fun  \mathit{BIT}  \land \\
			\forall (\mathit{bb}).(\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bb}) = 1-\mathit{bb})\\
			\end{array}
			$
			}

			\item Some utils lemmas to development proofs

			{\small
			$
			\begin{array}{l}
			\mathit{bit\_not}(0) = 1; 	\mathit{bit\_not}(1) = 0; \\
			\forall (\mathit{bb}).(\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bit\_not}(\mathit{bb})) = \mathit{bb});
			\end{array}
			$
			}
			\item The conjuction is a binary function defined as:
			{\small
			$
			\begin{array}{l}
			\mathit{bit\_and} \in \mathit{BIT} \times \mathit{BIT} \fun \mathit{BIT} \land \\
			\forall (\mathit{b1}, \mathit{b2}).(\mathit{b1}  \in \mathit{BIT}  \land \mathit{b2} \in \mathit{BIT} \implies \\
			\quad ((\mathit{bit\_and}(\mathit{b1}, \mathit{b2}) = 1) \iff (\mathit{b1} =1)  \land  (\mathit{b2} = 1))) \land \\
			 bit\_and(0,0) = 0	\land bit\_and(0,1) = 0	\land\\
			 bit\_and(1,0) = 0	\land bit\_and(1,1) = 1 \\
			\end{array}
			$
			}

  		\end{itemize}

\end{frame}

\begin{frame}
 \frametitle{Projeto \textit{Hardware Library}}

 	\begin{itemize}
 	  		\item The module \textit{BIT\_DEFINITION} has:

 	  		\begin{itemize}
 	  		\item Definiions of 	$\mathit{bit\_or},\mathit{bit\_xor}$ and
 	  		$bool\_to\_bit$.
 	  		\item Lemmas to associativity and commutativity of functions.
           \end{itemize}

			\item  The module \textit{BIT\_VECTOR\_DEFINITION} has

           \begin{itemize}
	 	  		\item Definition of bit vectors:
				{\small
				$
				\begin{array}{l}
				\mathit{BIT\_VECTOR} = \seq (\mathit{BIT}) % Implementação antiga, mas vale
				\end{array}
				$
				}
	 	  		\item Definition of functions:
	 	  		 {\small
				$
				\begin{array}{l}
				\mathit{bv\_size} \in \mathit{BIT\_VECTOR} \fun \nat_1 \land \\
				\mathit{bv\_size} = \lambda bv \bullet (bv \in \mathit{BIT\_VECTOR} \mid \mathbf{size}(bv))
				\end{array}
				$
				}
				\item It has others functions: \textit{bv\_catenate}, \textit{bv\_sub},
				\textit{bv\_not}, \textit{bv\_and}, \textit{bv\_or}, \textit{bv\_xor},
				\textit{bv\_set}, \textit{bv\_clear} e \ldots
			\begin{itemize}
	                  \item This functions using the definitions of  \textit{BIT\_DEFINITION}
	              	\end{itemize}
           \end{itemize}

           \item The module \textit{BIT\_VECTOR\_ARITHMETICS} define basics arithmetic functions
           \item The module \textit{BYTE} e \textit{BV16\_DEFINITION} define specialized functions to bit vectors of size 8 and 16
 	\end{itemize}

\end{frame}

\begin{frame}
 \frametitle{Types Library Project}
	  	\begin{itemize}
	        \item Define the common types and functions on platforms of 8 and 16 bits
	        \item For example:
	        \begin{itemize}
	          \item Function: \hspace*{0.0in}\it uchar\_schar  $\in$  \it UCHAR  $\fun$  \it SCHAR
	%  		  \hspace*{0.0in}\it uchar\_schar\rm = $\lambda$ \rm (\it v1\rm )\rm .\rm (\it v1 $\in$ \it UCHAR
	%  			$\land$  \it v1 $\leq$ \it SCHAR\_MAX $\mid$  \it v1\rm )  $\land$\\
	%  			\hspace*{0.0in}\it uchar\_schar\rm = $\lambda$ \rm (\it v1\rm )\rm .\rm (\it v1 $\in$ \it UCHAR
	%  			$\land$   $\neg$ \rm (\it v1 $\leq$ \it SCHAR\_MAX\rm )  $\mid$  \it v1\rm -\it UCHAR\_MAX \rm + \rm 1
	%  			\rm )

	          \item Lemma: $\forall(x).( x \in UCHAR \Rightarrow uchar\_byte(byte\_uchar(x)) = x)$
	        \end{itemize}
	      \end{itemize}

	\begin{table}[h]
	\begin{center}
	{\scriptsize
	\begin{tabular}{|c|c|c|c|c|c|c|}
	\hline
	 $Nome$&$\mathit{UCHAR}$&$\mathit{SCHAR}$&$\mathit{USHORTINT}$&$\mathit{SSHORTINT}$&$\mathit{BYTE}$&$\mathit{BV16}$\\
	 \hline $Faixa$&0..255&-128..127&0..65.535&-32.768..32.767&--&--\\ \hline
	 $Tamanho$ & 1 byte & 1 byte & 2 bytes & 2 bytes &  1 bytes & 2 bytes \\ \hline
	\end{tabular}
	}
	\end{center}
	\caption{Description of integer data types}
	\label{tab:types}
	\end{table}
\end{frame}

\begin{frame}
 \frametitle{Platform project(Z80)}
  \begin{itemize}
    \item Z80 is a microcontroller  of 8 bits with addressing of 16 bits
    \item Supports:
    \begin{itemize}
      \item 158 instructions (included all of 8080 microcontroller )
      \item Several addressing modes and interruption types
      \item 256 input and output ports (8 bits)
      \item Some operations with 16 bits
    \end{itemize}

      \item Header of modelling B
  \end{itemize}

	 {\scriptsize
	 \begin{sloppypar}
	\hspace*{1.0in}\bf MACHINE\\
	\hspace*{1.15in}\it Z80\\
	\hspace*{1.0in}\bf INCLUDES\\
	\hspace*{1.10in}\it MEMORY\\
	\hspace*{1.0in}\bf SEES\\
	\hspace*{1.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,\\
	\hspace*{1.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,\\
	\hspace*{1.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,\\
	\hspace*{1.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION\\
	\end{sloppypar}
	}

\end{frame}

\begin{frame}
\frametitle{Modelling Registers, input/output ports and instructions from Z80}

 \begin{itemize}
   \item The invariant:

	{\scriptsize
	\begin{sloppypar}
	%\bf INVARIANT\\
	\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ pc  $\in$  \it INSTRUCTION
	$\land$\\
	\hspace*{0.10in}\it  sp  $\in$  \it BV16  $\land$  \it ix  $\in$  \it BV16  $\land$  \it iy  $\in$  \it
	BV16  $\land$\\
	\hspace*{0.10in}\it i\_  $\in$  \it BYTE  $\land$  \it r\_ $\in$  \it BYTE  $\land$ iff1  $\in$  \it BIT
	$\land$ \it iff2  $\in$  \it BIT  $\land$ \\
	\hspace*{0.10in}\it im \rm $\in$ \rm (\it BIT $\times$ \it BIT\rm )  $\land$ \it i\_o\_ports  $\in$
	\it BYTE $\fun$  \it BYTE\\
	\end{sloppypar}
	}
	\item A simple instruction
	{\scriptsize
	\begin{sloppypar}
	\hspace*{0.10in}\bf LD\_n\_A \rm ( \it nn \rm ) \rm =\\
	\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in}\\ %$\land$  \it nn $\in$\it DATA\_R\_ADR
	\hspace*{0.20in}\bf THEN\\
	\hspace*{0.20in}\bf updateAddressMem \rm ( \it ushort\_to\_bv16 \rm ( \it nn \rm ) \rm , \it rgs8 \rm ( \it a0 \rm )
	\rm )  $\para$\\
	\hspace*{0.20in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it
	update\_refresh\_reg\rm (\it r\_\rm )\\
	\hspace*{0.10in}\bf END\rm
	\end{sloppypar}

	}

	\end{itemize}
% \end{itemize}
% 	\begin{center}
% 	{\small
% 	$$
% 	\begin{array}{l}
% 	\mathit{BTFSS} (\mathit{ff}, \mathit{bb}) =\\
% 	\quad \PRE \mathit{ff} : \mathit{REGISTER} \land  \mathit{bb} : \mathit{BYTE\_INDEX} \\
% 	\quad \THEN \\
% 	\quad \quad	\IF \mathit{bitget}(\mathit{mem}(\mathit{ff}), \mathit{bb}) = 1 \THEN\\
% 	\quad\quad\quad	   \mathit{pc} :=\mathit{instruction\_next}(\mathit{instruction\_next}(\mathit{pc}))\\
% 	\quad\quad	\ELSE\\
% 	\quad\quad\quad	  \mathit{pc} := \mathit{instruction\_next}(\mathit{pc})\\
% 	\quad\quad	\END\\
% 	\quad \END
% 	\end{array}
% 	$$
% 	}\\
% 	Instrução de controle do \textit{PIC16C432}
%
% 	\end{center}

\end{frame}

### 2009-08-06 trunk/papers/SBMF2009/Presentation/content/5_Proofs.tex

\section{Prove Process}

\begin{frame}
\frametitle{Prove Process of Projects}

\begin{itemize}
   \item Statistics:\\
    \begin{center}{\footnotesize
  	 \begin{tabular}{|c|c|c|c|}
		\hline
		 \textsl{Projects} &  \textsl{POs}	& \textsl{POs WD} &	\textsl{Total}\\
		\hline
		\textit{Power} & 	21 &	4 &	25\\
		\hline
		\textit{Types Library} &	22 &	68 &	90\\
		\hline
		\textit{Hardware Library} &	108 &	324 &	432\\
		\hline
		\textit{Z80} &	631	 & 2470 & 	3101\\
		\hline
		   &    &  & 		\textbf{3648}\\
		\hline
	 \end{tabular}
   }
	\end{center}

   \item  Approximately 25\% are not solved automatically
	\item  Approximately 18 days are needed to make almost all the proofs using a monolithic computer
		\begin{itemize}
			\item But You can reduce this time if you use several computer to prove
		\end{itemize}
	\item  Sometimes was needed  replay the proofs when were needed news fixes
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{A example of user pass}

\begin{itemize}
   \item Specialized:
   \item \ldots
   \item Generally:
\end{itemize}

\end{frame}

### 2009-08-06 papers/SBMF2009/Presentation/content/3_Modelling.tex
The presentation

@@
 \section{Modelling the Instruction Set in B}
-\begin{frame}
-  \frametitle{Modelling  Microcontrollers}
-
-  \begin{itemize}[<+->]
-  \item Relations 
-  \begin{itemize}
-  	\item States of platform  $\Leftrightarrow$ States of modell
-		\begin{itemize}
-		\item Registers and Memory $\Leftrightarrow$  Variables of model
-		\end{itemize}
-	\item Assembly Instructions $\Leftrightarrow$ B operations
-        \begin{itemize}
-        \item Changes of states of platform $\Leftrightarrow$ B substitutions
-        \end{itemize}
-  \end{itemize}
-
- \end{itemize}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Modelling  Microcontrollers}  
+% 
+%   \begin{itemize}[<+->]
+%   \item Relations 
+%   \begin{itemize}
+%   	\item States of platform  $\Leftrightarrow$ States of modell
+% 		\begin{itemize}
+% 		\item Registers and Memory $\Leftrightarrow$  Variables of model
+% 		\end{itemize}
+% 	\item Assembly Instructions $\Leftrightarrow$ B operations
+%         \begin{itemize}
+%         \item Changes of states of platform $\Leftrightarrow$ B substitutions
+%         \end{itemize}
+%   \end{itemize}
+%     
+%  \end{itemize}
+% \end{frame}
 
 \begin{frame}
- \frametitle{Structure of Projects B}
+ \frametitle{Structure of B Projects}
  \begin{center}\includegraphics[width=1.\textwidth]{figures/diagramaEstrutural_vertical.png}
@@
 \begin{frame}
- \frametitle{Project Hardware Library}
+ \frametitle{Hardware Library Project}
 
@@
 \begin{frame}
- \frametitle{Projeto \textit{Hardware Library}}
+ \frametitle{Hardware Library Project}
 
@@
            \item The module \textit{BIT\_VECTOR\_ARITHMETICS} define basics arithmetic functions
-           \item The module \textit{BYTE} e \textit{BV16\_DEFINITION} define specialized functions to bit vectors of size 8 and 16
+           \item The module \textit{BYTE} and \textit{BV16\_DEFINITION} define specialized functions to
+           bit vectors of size 8 and 16
  	\end{itemize}
@@
 \begin{frame}
- \frametitle{Platform project(Z80)}
+ \frametitle{Platform Project(Z80)}
   \begin{itemize}
@@
 
-      \item Header of modelling B
+  \item Modelling 
+  \begin{itemize}
+  	\item States of platform  $\Leftrightarrow$ States of modell
+		\begin{itemize}
+		\item Registers and Memory $\Leftrightarrow$  Variables of model
+		\end{itemize}
+	\item Assembly Instructions $\Leftrightarrow$ B operations
+        \begin{itemize}
+        \item Changes of states of platform $\Leftrightarrow$ B substitutions
         \end{itemize}
 
-	 {\scriptsize
-	 \begin{sloppypar}
-	\hspace*{1.0in}\bf MACHINE\\
-	\hspace*{1.15in}\it Z80\\
-	\hspace*{1.0in}\bf INCLUDES\\
-	\hspace*{1.10in}\it MEMORY\\
-	\hspace*{1.0in}\bf SEES\\
-	\hspace*{1.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,\\
-	\hspace*{1.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,\\
-	\hspace*{1.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,\\
-	\hspace*{1.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION\\
-	\end{sloppypar}
-	}
+ \end{itemize}
+
+  \end{itemize}
 
@@
 \begin{frame}
-\frametitle{Modelling Registers, input/output ports and instructions from Z80}
-
- \begin{itemize}
-   \item The invariant:
+\frametitle{Modelling Registers, Input/Output Ports and Instructions from Z80}
+%  \begin{itemize}
+%  \item Header of Modelling B
+%  \end{itemize}  
+	 {\scriptsize
+	 \begin{sloppypar}
+		\hspace*{.0in}\bf MACHINE\\
+		\hspace*{.15in}\it Z80\\
+		\hspace*{.0in}\bf INCLUDES\\
+		\hspace*{.10in}\it MEMORY\\
+		\hspace*{.0in}\bf SEES\\ \ldots
+	% 	\hspace*{1.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,\\
+	% 	\hspace*{1.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,\\
+	% 	\hspace*{1.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,\\
+	% 	\hspace*{1.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION\\
+		\end{sloppypar}
+	}
+% \begin{itemize}  
+ %  \item The invariant:
 
@@
 	\begin{sloppypar}
-	%\bf INVARIANT\\
+	\bf INVARIANT\\
 	\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ pc  $\in$  \it INSTRUCTION
@@
 	\it BYTE $\fun$  \it BYTE\\
+	\ldots
 	\end{sloppypar}
+
 	}
-	\item A simple instruction
+
+%	\item A simple instruction
 	{\scriptsize
 	\begin{sloppypar}
+    \bf OPERATIONS\\
 	\hspace*{0.10in}\bf LD\_n\_A \rm ( \it nn \rm ) \rm =\\
@@
 	\hspace*{0.10in}\bf END\rm
+	\ldots
 	\end{sloppypar}
@@
 
-	\end{itemize}
+	%\end{itemize}
 % \end{itemize}

### 2009-08-06 papers/SBMF2009/Presentation/content/7_Conclusions.tex
The presentation

\section{Conclusions}

 \begin{frame}
 \frametitle{Considerations about the work}

 \begin{itemize}
   \item It can be used to: \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 
    \includegraphics[width=.2\textwidth]{figures/processador_generico.jpg}
 	\begin{itemize} 
 	  \item Identify  errors on the official reference
 	  \item Important to  assembly programmers
 	  \item Formal verification until assembly language \\
 	   \includegraphics[height=.25\textheight]{figures/dutos.png} \ \ \ \ \ \ \ \ \ \ 
        \includegraphics[height=.25\textheight]{figures/trem_bala2.png} 
 	\end{itemize} 
 \end{itemize}

 \end{frame}

\begin{frame}
  \frametitle{Sugestions/Doubts/Critics }
  	\begin{center}
    {\huge ?}\\
    \includegraphics[width=.5\textwidth]{figures/rodin.png}\\
    \large{ If you prefer, you can send an e-mail to 
    %\textcolor{blue!70!black}{\textit{valerio}}@ppgsc.ufrn.br or
    \textcolor{blue!70!black}{\textit{valerio.jr}}@gmail.com 
    }
	\end{center}

\end{frame}

### 2009-08-06 trunk/papers/SBMF2009/Presentation/content/1_introduction.tex
The presentation

--- 2009-08-06 papers/SBMF2009/Presentation/content/1_introduction.tex
@@
 \begin{frame}
-\frametitle{Overview}
+\frametitle{Initial Overview}
 
@@
     \item Until the implementation model
-    \item But the result of this translation is not guaranteed by formal means
-    \item Then the transformation process can introduce small bugs\\
-       \includegraphics[width=.7\textwidth]{figures/passos_de_desenvolvimento_tradicional.png}
+    \item But the result of this translation is not guaranteed by formal means\\
+    \begin{center}
+       \includegraphics[height=.5\textheight]{figures/b-method-actual_new.pdf}
+    \end{center}
+       \begin{itemize} \item   \small{The transformation process can introduce small bugs}
        \end{itemize}
+  \end{itemize}
+\end{itemize}
+
+	\note{ ................... }
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Extended Overview}
+
+\begin{itemize}[<+->]
+  \item The B method supports the construction of safety systems
+  \begin{itemize}
     \item The last year a paper [Dantas, 2008] proposed a approach to extend the formal verification
     \item One key of this approach is the formal model of the instruction set  of execution platform % of such assembly languages
+      \\ \begin{center} \includegraphics[height=.5\textheight]{figures/b-method-ideal_new.pdf} \end{center}
+  \end{itemize}
+
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Overview}
+\frametitle{Modelling Assembly Instruction Set in B}
 
 \begin{itemize}[<+->]
-  \item The B method supports the construction of safety systems
-  \begin{itemize}
-    \item Until the implementation model
-    \item But the result of this translation is not guaranteed by formal means
-    \item Then the transformation process can introduce small bugs\\
-       \includegraphics[width=.7\textwidth]{figures/passos_de_desenvolvimento_extendido.png}
-  \end{itemize}
-  \item The last year a paper [Dantas, 2008] proposed a approach to extend the formal verification
-  \item One key of this approach is the formal model of the instruction set  of execution platform % of such assembly languages
-\end{itemize}
-% TO-DO Ajustar a figura
-
-	\note{ ................... }
-\end{frame}
-
-\begin{frame}
-\frametitle{Modelling}
-
-\begin{itemize}[<+->]
-  \item The modelling is builded using some libraries developed
-  \item The libraries has common concepts that can be used to others platforms too
-  \item Utilities of this Formal Model:
+  \item The modelling is builded using some developed libraries
+  \item The libraries have common concepts that can be used to others platforms of 8 or 16 bits
+  \item Utilities of this formal model:
   \begin{itemize}
@@
 \begin{itemize}[<+->]
-  \item The main actual objective is represent the instruction set
+  \item The main actual objective is represent the assembly instruction set
     \begin{itemize}
@@
   \end{itemize}
-  \item Because there are many efficient ways to verify hardware
+  \item Moreover there are many efficient ways to verify hardware
 

### 2009-08-07 trunk/papers/SBMF2009/Presentation/content/5_Proofs.tex
version almost conclused

@@
-\section{Prove Process}
+\section{Proof Process}
 
 \begin{frame}
-\frametitle{Prove Process of Projects}
+\frametitle{Proof Process of Projects}
 
@@
 	\item Approximately 18 days are needed to make almost all the proofs using a monolithic computer
-		\begin{itemize}
-			\item But You can reduce this time if you use several computer to prove
-		\end{itemize}
-	\item  Sometimes was needed  replay the proofs when were needed news fixes
+	\item Some proofs needed to be replayed when we make changes to the specification
+	\item The proof's time can be reduced using advanced features
+
 \end{itemize}
@@
 \begin{frame}
-\frametitle{A example of user pass}
+\frametitle{Proof Process - Advanced Features}
 
 	\begin{itemize}
-   \item Specialized:
-   \item \ldots
-   \item Generally:
+	\item Parallel theorem prover with 3 dual core computers 
+	\item Two simple examples of user pass from theorem prover
+
+ 		\begin{itemize}
+ 		   \item Specific:\\
+ 		  	 $Pattern( x , y : dom(bitget) )$ \\ $\& ff(0) \& dd \& eh(dom(bitget)) \& ss \& pr;$\\
+ 		    \ldots
+ 		   \item General:\\
+ 		   	$Pattern( x : dom(y) )$ \\ $\& ff(0) \& dd \& eh(dom(y)) \& ss \& pr;$
+         \end{itemize}
+
     \end{itemize}

### 2009-08-07 papers/SBMF2009/Presentation/content/1_introduction.tex
version almost conclused

@@
     \item Until the implementation model
-    \item But the result of this translation is not guaranteed by formal means
-    \item Then the transformation process can introduce small bugs\\
-       \includegraphics[width=.7\textwidth]{figures/passos_de_desenvolvimento_tradicional.png}
+    \item But part of this translation is not guaranteed by formal means\\
+    \begin{center}
+       \includegraphics[height=.5\textheight]{figures/b-method-actual_new.pdf}
+    \end{center}
+       \begin{itemize} \item   \small{The transformation process can introduce small bugs}
        \end{itemize}
-  \item The last year a paper [Dantas, 2008] proposed a approach to extend the formal verification
+  \end{itemize}
+\end{itemize}
+
+	\note{ ................... }
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Extended Overview}
+
+\begin{itemize}[<+->]
+    \item We proposed [Dantas, 2008]  a approach to extend the formal verification
     \item One key of this approach is the formal model of the instruction set  of execution platform % of such assembly languages
+      \\ \begin{center} \includegraphics[height=.5\textheight]{figures/b-method-ideal_new.pdf} \end{center}
+
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Overview}
+\frametitle{Modelling Assembly Instruction Set in B}
 
 \begin{itemize}[<+->]
-  \item The B method supports the construction of safety systems
-  \begin{itemize}
-    \item Until the implementation model
-    \item But the result of this translation is not guaranteed by formal means
-    \item Then the transformation process can introduce small bugs\\
-       \includegraphics[width=.7\textwidth]{figures/passos_de_desenvolvimento_extendido.png}
-  \end{itemize}
-  \item The last year a paper [Dantas, 2008] proposed a approach to extend the formal verification
-  \item One key of this approach is the formal model of the instruction set  of execution platform % of such assembly languages
-\end{itemize}
-% TO-DO Ajustar a figura
-
-	\note{ ................... }
-\end{frame}
-
-\begin{frame}
-\frametitle{Modelling}
-
-\begin{itemize}[<+->]
-  \item The modelling is builded using some libraries developed
-  \item The libraries has common concepts that can be used to others platforms too
-  \item Utilities of this Formal Model:
+  \item We developed libraries of hardware models%.The modelling is builded using some  developed libraries
+  \item The libraries have common concepts that can be used to others platforms of 8 or 16 bits
+  \item Utilities of this formal model:
   \begin{itemize}
@@
 
-\begin{frame}
-\frametitle{Objective}
-
-\begin{itemize}[<+->]
-  \item The main actual objective is represent the instruction set
-    \begin{itemize}
-    \item The aspects no-functional are not represented: logic circuits, pipeline, data bus, \ldots 
-  \end{itemize}
-  \item Because there are many efficient ways to verify hardware
-
-\end{itemize}
 % 
-
-	\note{ ................... }
-\end{frame}
+% \begin{frame}
+% \frametitle{Objective}  
+% 
+% \begin{itemize}[<+->]
+%   \item The main actual objective is represent the assembly instruction set
+%     \begin{itemize}
+%     \item The aspects no-functional are not represented: logic circuits, pipeline, data bus, \ldots 
+%   \end{itemize}
+%   \item Moreover there are many efficient ways to verify hardware
+% 
+% \end{itemize}
+% % 
+% 
+% 	\note{ ................... }
+% \end{frame}
+%

### 2009-08-14 papers/SBMF2009/Presentation/content/1_introduction.tex

@@
   \begin{itemize}
-    \item Until the implementation model
+    \item From design model
+    \item To implementation model
     \item But part of this translation is not guaranteed by formal means\\
@@
 
-\begin{itemize}[<+->]
-    \item We proposed [Dantas, 2008]  a approach to extend the formal verification
-    \item One key of this approach is the formal model of the instruction set  of execution platform % of such assembly languages
+\begin{itemize}%[<+->]
+    \item We proposed a approach [Dantas, 2008] to extend the formal
+    verification
+    \item One key element of this approach is the formal model of the
+    instruction set of execution platform % of such assembly languages
      \\ \begin{center}     \includegraphics[height=.5\textheight]{figures/b-method-ideal_new.pdf} \end{center}
-
 \end{itemize}
-% TO-DO Ajustar a figura
-
-	\note{ ................... }
+	\note{...................}
 \end{frame}
@@
     \item A possible support to verify the Z80 design
+    	\item Reference for verification of Z80 hardware implementations
 

### 2009-08-14 papers/SBMF2009/Presentation/content/7_Conclusions.tex
version almost conclused

@@
  \begin{frame}
- \frametitle{Considerations about the work}
+ \frametitle{Conclusions about the Work}
 
  \begin{itemize}
-   \item It can be used to: \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 
+   \item Contributions: \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
     \includegraphics[width=.2\textwidth]{figures/processador_generico.jpg}
  	\begin{itemize} 
- 	  \item Identify  errors on the official reference
- 	  \item Important to  assembly programmers
- 	  \item Formal verification until assembly language \\
- 	   \includegraphics[height=.25\textheight]{figures/dutos.png} \ \ \ \ \ \ \ \ \ \ 
-        \includegraphics[height=.25\textheight]{figures/trem_bala2.png} 
+ 	  \item Identify  errors in the official reference
+ 	  \item Help assembly programmers
+ 	  \item Formal verification until assembly language 
      \end{itemize} 
+     \item Future work is the verified development of a petroleum production test system\\ 
+     \begin{center}   
+ 	   \includegraphics[height=.25\textheight]{figures/dutos.png}
+	 \end{center}%\ \ \ \ \ \ \ \ \ \ 
+     %   \includegraphics[height=.25\textheight]{figures/trem_bala2.png}
+
  \end{itemize}
@@
     %\textcolor{blue!70!black}{\textit{valerio}}@ppgsc.ufrn.br or
-    \textcolor{blue!70!black}{\textit{valerio.jr}}@gmail.com 
+    \textcolor{blue!70!black}{\textit{valerio.jr}}@gmail.com \\
+    Project available at http://code.google.com/p/b2asm
     }

### 2009-08-16 papers/SBMF2009/Presentation/content/1_introduction.tex
Praticamente final

@@
 \begin{itemize}[<+->]
-  \item We developed libraries of hardware models%.The modelling is builded using some  developed libraries
+  \item We developed libraries with definitions of hardware and integer types
+	%.The modelling is builded using some  developed libraries
   \item The libraries have common concepts that can be used to others platforms of 8 or 16 bits
@@
     \item Simulation
-    \item The formal verification until the assembly level
-    \item A possible support to verify the Z80 design
     \item Reference for verification of Z80 hardware implementations 
+    \item The formal verification to the assembly level
+    %\item A possible support to verify the Z80 design
 

### 2010-04-05 papers/Article2010/source/paper.tex
Initial

--- 2009-04-27 papers/FM2009/source/paper.tex
@@
 \documentclass{llncs}
+%\documentclass{article}
 
@@
 \usepackage{amsfonts}
+\usepackage{epsfig}
+\usepackage{float}
+%\usepackage{supertabular} % used in B Symbols
 
@@
 
-\title{Formal Verification of Real Models from Platforms Using The B-method }
-
-\author{B. Dantas \inst{1}, David D\'{e}harbe\inst{1}, Val\'{e}rio Medeiros Jr\inst{1} }
-
-\institute{Federal University of Rio Grande do Norte, Natal RN 59078-970, Brazil}
+\title{Formal Construction of Microcontroller Instruction Set Model Using B}
+
+\hyphenation{se-pa-ra-te}
+
+\author{ David D\'{e}harbe\inst{1}, Val\'{e}rio Medeiros Jr\inst{2} }
+\institute{Federal University of Rio Grande do Norte, Natal RN 59078-970, Brazil\inst{1} and\\  
+Federal Institute of Education, Science and Technology of Rio Grande do Norte, Natal RN
+59015-000, Brazil\inst{2}}
 
@@
 \begin{abstract}
-This paper describes an approach to model the microcontroller platforms. More specifically, it shows details about
-the Z80 model. The model is created using the B method; it applies math and logic concepts to describe the
-characteristics from platforms. Therefore, this model can be used in platforms projects to document, build
-simulators, verify properties about the model and verify the software in assembly level. The verification in
-assembly is the feature more important, because this allow the developing software faithful to the algorithm
-language. Finally, this paper show some relevant techniques used in models and to describe several details.
+  [To redo]
+  This paper describes an approach to model the functional aspects of
+  the instruction set of microcontroller platforms using the notation
+  of the B method. The paper presents specifically the case of the Z80 
+  platform. This work is a contribution towards the extension of the
+  B method to handle developments up to assembly level code.
 \end{abstract}
 %
-%\keywords{Fomal Methods, Microprocessor Verification, Hardware}
+%\keywords{Fomal Methods, microcontroller Verification, Hardware}
 
 \section{Introduction}
-% Explica o contexto geral do artigo
-%  Porque modelar os microcontroladores. \footnote{The interested reader in details from full specification or in
-% other study case is invited to visit the repository from authors at \url{http://b2asm.googlecode.com}.
-%    Porque podemos verificar várias propriedades em tempo de projeto, documentar, simulacao através das ferramentas de suporte do método B...
-%    Porque permite verificar software em nível de assemblagem.[citar trabalho de BHDL]
-%    Porque utilizamos o método B puro  - [citar proposta do ALEMAO do ProB]
-%  Contextualizacao com grande desafio da verificao formal em nivel de assemblagem[citar grande challenge] 
-%  Conclusão da introducao.
-%  Organizacao do paper
-%  Basear-se no Relatorio técnico e no ERMAC
-%
-% - Citar o Atelierb
-
-[I need to Review this below text and to answer some questions]
-%  What the objectives from this models ?
-%  The objectives from this models are: supply a documentation that help the
-%  directly in the development and guarantees a series of properties. This
-%  properties are about hardware and software that run in this platform. The
-%  properties verification about software allow to verify software in the
-%  assembly level and It`s provide a high level of security, cause the mains
-%  levels of representation of software are verified, according to
-%  \cite{DANTAS08_SBMF}.
-
-% Why don´t we use a language based in events, how Event-B or CSP? 
-%  Because, we want to prover importants properties about the software and 	
-%  the B-method is more appropriate and don´t lost the capacity of verification
-%  about hardware level.
-
-% What are the advantages to use the B method ? The biggest advantage is the
-% formal verification on assembly level.
-
-% Once established the model of the platform, so I could check something on VHDL?
-% Or generate VHDL?
-
-\section{B Notation}
-
-[One page]
-
-\subsubsection{Desing components B }
-[half page]
-[Explicar como as coias são modeladas]
-% Use the tech report
-
-\section{General definitions for hardware models}
-
-We think that We find it convenient to provide a set of basic definitions to model hardware concepts. These
-definitions are grouped in a library that we named $\mathit{HARDWARE\_LIBRARY}$ and arrange through four modules:
-$\mathit{BIT\_DEFINITION}$, $\mathit{BIT\_VECTOR\_DEFINITION}$, $\mathit{BYTE\_DEFINITION}$ and
-$\mathit{BIT\_VECTOR\_ARITHMETICS}$.
-
-[[Explain better the figure]]
-
-The dependency graph between these modules is depicted in
-Figure~\ref{fig:hardware-definition-graph}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=.7\textwidth]{diagramaEstrutural.jpg}
-\caption{Dependencies between modules of $\mathit{HARDWARE\_LIBRARY}$ and
-platform model.}
+[To update]
+The B method~\cite{Abrial} supports the construction of safety systems
+models by verification of proofs that guarantees its correctness. So,
+an initial abstract model of the system requirements is defined and
+then it is refined until the implementation model. Development
+environments based on the B method also include source code generators
+for programming languages, but the result of this translation
+cannot be compared by formal means. The paper \cite{Dantas_SBMF08} presented
+recently an approach to extend the scope of the B method up to the
+assembly level language. One key component of this approach
+is to build, within the framework of the B method, formal models of
+the instruction set of such assembly languages.
+
+This work gives an overview of the formal modelling of the instruction
+set of the Z80 microcontroller~\cite{Z80_manual}\footnote{The
+  interested reader in more details is invited to visit our repository
+  at: http://code.google.com/p/b2asm.}. Using the responsibility
+division mechanism provided by B, auxiliary libraries of basic modules
+were developed as part of the construction of microcontroller
+model. Such library has many definitions about common concepts used in
+the microcontrollers; besides the Z80 model, it is used by two other
+microcontrollers models that are under way.
+
+Other possible uses of a formal model of a microcontroller
+instruction set include documentation, the construction of simulators,
+and be possibly the starting point of a verification effort for the
+actual implementation of a Z80 design. Moreover the model of the
+instruction set could be instrumented with non-functional aspects,
+such as the number of cycles it takes to execute an instruction, to
+prove lower and upper bounds on the execution time of a routine. The
+goal of this project, though, is to provide a basis for the generation
+of software artifacts at the assembly level that are amenable to 
+refinement verification within the B method.
+
+This paper is focused on the presentation of the Z80 model, including
+elementary libraries to describe hardware aspects. The paper is
+structured as follows. Section~\ref{sec:B_method} provides a short
+introduction to the B method. Section~\ref{sec:models} presents the
+elementary libraries and the modelling of some elements common to
+microcontrollers.  Section~\ref{sec:z80} presents the B model of the
+Z80 instruction set. Section ~\ref{sec:Proofs} provides some
+information on the proof effort needed to analyze the presented
+models. Related work is discussed in
+Section~\ref{sec:relatedworks}. Finally, the last section is devoted
+to the conclusions.
+
+\section{Introduction to the B Method}
+\label{sec:B_method}
+
+The B method for software development~\cite{Abrial} is based on the B Abstract
+Machine Notation (AMN) and the use of formally proved refinements up to a
+specification sufficiently concrete that programming code can be automatically
+generated from it. Its mathematical basis consists of first order logic, integer
+arithmetic and set theory, and its corresponding constructs are similar to those
+of the Z notation.
+
+A B specification is structured in modules. A module defines a set of valid
+states, including a set of initial states, and operations that may provoke a
+transition between states. The design process starts with a module with a
+so-called functional model of the system under development. In this initial
+modelling stage, the B method requires that the user proves that, in a machine,
+all the its initial states are valid, and that operations do not define
+transitions from valid states to invalid states.
+
+Essentially, a B module contains two main parts: a header and the available
+operations. Figure~\ref{fig:maqB} has a very basic example. The clause
+$\mathit{MACHINE}$ has the name of module.  The next two clauses respectively
+reference external modules and create an instance of an external module. The
+$\mathit{VARIABLES}$ clauses declares the name of the variables that compose the
+state of the machine. Next, the $\mathit{INVARIANT}$ clause defines the type and
+other restrictions on the variables. The $\mathit{INITIALIZATION} $ specifies
+the initial states. Finally, operations correspond to the transitions between
+states of the machine.
+
+\begin{figure}
+  % \begin{small}
+  $$
+  \begin{array}{lcl}
+    \begin{array}[t]{l}
+      \MACHINE \quad  \mathit{micro} \\
+      \SEES \quad \mathit{TYPES}, \mathit{ALU} \\
+      \INCLUDES  \quad \mathit{MEMORY} \\
+      \VARIABLES \quad    \mathit{pc}  \\
+      \INVARIANT \quad \mathit{pc} \in \mathit{INSTRUCTION} 
+    \end{array}
+    & \hspace*{1cm} &
+    \begin{array}[t]{l}
+      \INITIALISATION  \mathit{pc} := 0 \\
+      \OPERATIONS\\
+      \mathit{JMP} \mathit{( jump )} = \\
+      \quad \PRE \mathit{jump} \in \mathit{INSTRUCTION}\\ 
+      \quad \THEN \mathit{pc} := \mathit{jump}\\  
+      \quad \END \\
+      \END
+    \end{array}
+  \end{array}
+  $$
+\caption{A very basic B machine.}
+\label{fig:maqB}
+\end{figure}
+
+\section{Model structure and basic components}
+\label{sec:models}
+
+We have been developed a reusable set of basic definitions to model hardware concepts and data types
+concepts. These definitions are grouped into two separate development projects and are available as
+libraries. A third project is devoted to the higher-level aspects of the platform. Thus, the workspace is
+composed of: a power library\footnote{The power library has the basics definitions to help the theorems
+prover about simple calculus of power, especially power of two. It has constants, that represents all
+results of power function needed, and the definition of power function.}, a hardware library, a types
+library and a project for the specific platform, in this case the Z80. The corresponding dependency
+diagram is depicted in Figure~\ref{fig:hardware-definition-graph}; information specific to each project is
+presented in the following.
+
+\begin{figure}[h] \centering
+\includegraphics[width=.92\textwidth]{diagramaEstrutural_vertical.png}
+ \caption{Dependency diagram of the Z80 model.}
 \label{fig:hardware-definition-graph}
-
 \end{figure}
 
-Now, We will explain each of these modules and present them in detail.
-
-\subsection{Definitions to represent and manipulate bits}
+\subsection{Bit Representation and Manipulation}
 
@@
 
-First, bits are modeled as the set of integers $\{0, 1\}$:\\
-
-$
-\begin{array}{l}
-\mathit{BIT} = \mathit{0..1} \\ 
-\end{array}
-$
-
-The negation is a unary function on bits and it is defined as:
-
-$
-\begin{array}{l}
-\mathit{bit\_not}  \in  \mathit{BIT}  \fun  \mathit{BIT}  \land \\
-\forall ( \mathit{bb}). (\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bb}) =
+First, bits are modelled as a set of integers: $\mathit{BIT} =
+\mathit{0..1}$. The negation is an unary function on bits and it is
+defined as:
+
+$
+\begin{array}{l}
+\mathit{bit\_not}  \in  \mathit{BIT}  \fun  \mathit{BIT}  \land \forall ( \mathit{bb}). (\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bb}) =
 1-\mathit{bb})\\
@@
 
-The module also provides lemmas on negation that may be useful for the users of the library to develop proofs:
-
-$
-\begin{array}{l}
- \mathit{bit\_not}(0) = 1;  \mathit{bit\_not}(1) = 0; \\
-\forall (\mathit{bb}).(\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bit\_not}(\mathit{bb})) = \mathit{bb});
-\end{array}
-$
-
-Conjunction is a unary function on bits and it is defined as:
+The module also provides lemmas on negation that may be useful for the
+users of the library to develop proofs:
+
+$
+\begin{array}{l}
+%  \mathit{bit\_not}(0) = 1;  \mathit{bit\_not}(1) = 0; \\
+\forall (\mathit{bb}).(\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bit\_not}(\mathit{bb})) = \mathit{bb})
+\end{array}
+$
+
+Conjunction is an unary function on bits and it is defined as:
 
@@
 \begin{array}{l}
- \mathit{bit\_and}(0,0) = 0;  \mathit{bit\_and}(0,1) = 0; \\
- \mathit{bit\_and}(1,0) = 0;  \mathit{bit\_and}(1,1) = 1; \\
+%  \mathit{bit\_and}(0,0) = 0;  \mathit{bit\_and}(0,1) = 0; \\
+%  \mathit{bit\_and}(1,0) = 0;  \mathit{bit\_and}(1,1) = 1; \\
 \forall (\mathit{b1},\mathit{b2}).(\mathit{b1} \in \mathit{BIT} \land \mathit{b2} \in \mathit{BIT} \implies \\
-\quad (\mathit{bit\_and}(\mathit{b1}, \mathit{b2}) = \mathit{bit\_and}(\mathit{b2},\mathit{b1}))); \\
+\quad (\mathit{bit\_and}(\mathit{b1}, \mathit{b2}) = \mathit{bit\_and}(\mathit{b2},\mathit{b1})))\land \\
 \forall (\mathit{b1},\mathit{b2},\mathit{b3}).(\mathit{b1} \in \mathit{BIT} \land  \mathit{b2} \in \mathit{BIT} \land \mathit{b3} \in \mathit{BIT} \implies \\
-\quad (\mathit{bit\_and}(\mathit{b1}, \mathit{bit\_and}(\mathit{b2},\mathit{b3})) = \mathit{bit\_and}(\mathit{bit\_and}(\mathit{b1},\mathit{b2}),\mathit{b3}))); \\
-\forall (\mathit{b1}).(\mathit{b1} \in \mathit{BIT} \implies (\mathit{bit\_and}(\mathit{b1}, 1) = \mathit{b1})); \\
-\forall (\mathit{b1}).(\mathit{b1} \in \mathit{BIT} \implies (\mathit{bit\_and}(\mathit{b1}, 0) = 0));
-\end{array}
-$
-
-The module provides definitions of $\mathit{bit\_or}$ (disjunction) and $\mathit{bit\_xor}$ (exclusive disjunction),
-as well as lemmas on those operators. These are standard and their expression in B is similar as for
-$\mathit{bit\_and}$, they are thus omitted.
+\quad (\mathit{bit\_and}(\mathit{b1}, \mathit{bit\_and}(\mathit{b2},\mathit{b3})) = \mathit{bit\_and}(\mathit{bit\_and}(\mathit{b1},\mathit{b2}),\mathit{b3})))\\
+% \forall (\mathit{b1}).(\mathit{b1} \in \mathit{BIT} \implies (\mathit{bit\_and}(\mathit{b1}, 1) = \mathit{b1})); \\
+% \forall (\mathit{b1}).(\mathit{b1} \in \mathit{BIT} \implies (\mathit{bit\_and}(\mathit{b1}, 0) = 0));
+\end{array}
+$
+
+The module provides definitions of $\mathit{bit\_or}$ (disjunction)
+and $\mathit{bit\_xor}$ (exclusive disjunction), as well as lemmas on
+those operators. These are standard and their expression in B is
+similar as for $\mathit{bit\_and}$, they are thus omitted.
 
@@
 
-\subsection{Representation and manipulation of bit vectors}
-
-Sequences are pre-defined in B, as functions whose the domain is an integer range with lower bound 1 (one). Indices
-in bit vectors usually range from 0 (zero) upwards and the model we propose obeys this convention by making an
-one-position shift where necessary.  We thus define bit vectors as non-empty sequences of bits, and
+\subsection{Representation and Manipulation of Bit Vectors}
+
+Sequences are pre-defined in B, as functions whose the domain is an
+integer range with lower bound 1 (one). Indices in bit vectors usually
+range from 0 (zero) upwards and the model we propose obeys this
+convention by making an one-position shift where necessary. This shift
+is important to use the predefined functions of sequences. We thus
+define bit vectors as non-empty sequences of bits, and
 $\mathit{BIT\_VECTOR}$ is the set of all such sequences:
-
-$
-\begin{array}{l}
-\mathit{BIT\_VECTOR} = \seq (\mathit{BIT})
-\end{array}
-$
+$\mathit{BIT\_VECTOR} = \seq1(\mathit{BIT})$.
 
@@
 \mathit{bv\_size} \in \mathit{BIT\_VECTOR} \fun \nat_1 \land \\
-\mathit{bv\_size} = \lambda bv \bullet (bv \in \mathit{BIT\_VECTOR} \mid \mathbf{size}(bv))
-\end{array}
-$
-
-We also define two functions $\mathit{bv\_set}$ and $\mathit{bv\_clear}$ that, given a bit vector, and a position of
-the bit vector, return the bit vector resulting from setting the corresponding position to 0 or to 1, and a function
-$\mathit{bv\_get}$ that, given a bit vector, and a valid position, each one returns the value of the bit at that
-position. Only the first definition is shown here:
-
-$
-\begin{array}{l}
-\mathit{bv\_set} \in \mathit{BIT\_VECTOR} \times \nat \fun \mathit{BIT\_VECTOR} \land \\
-\mathit{bv\_set} = \lambda v, n \bullet (v \in \mathit{BIT\_VECTOR} \land 0 \in \nat \land n < \mathit{bv\_size}(v) \mid v \ovr \{ n+1 \mapsto 1 \})
+\mathit{bv\_size} = \lambda bv . (bv \in \mathit{BIT\_VECTOR} \mid \mathbf{size}(bv))
+\end{array}
+$
+
+We also define two functions $\mathit{bv\_set}$ and $\mathit{bv\_clear}$ that, given a bit vector, and a
+position of the bit vector, return the bit vector resulting from setting the corresponding position to 0
+or to 1, and a function $\mathit{bv\_get}$ that, given a bit vector, and a valid position, each one
+returns the value of the bit at that position. Only the first definition is shown here:
+
+$
+\begin{array}{l}
+\mathit{bv\_set} \in \mathit{BIT\_VECTOR} \times \nat \fun \mathit{BIT\_VECTOR} \land \mathit{bv\_set} =\\
+\lambda v, n . (v \in \mathit{BIT\_VECTOR} \land n \in \nat \land n <\mathit{bv\_size}(v)
+\mid v \lover \{ n+1 \mapsto 1 \})
 \end{array}
@@
 
-\hspace*{0.00in} \it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$ \it BIT\_VECTOR $\land$
-
-\hspace*{0.00in} \it bv\_catenate \rm =  $\lambda$  v\rm ,\it w \rm . \rm (\it v  $\in$  \it BIT\_VECTOR $\land$ \it w $\in$  \it BIT\_VECTOR  $\mid$  \it v $\cat$ \it w\rm )
+\hspace*{0.00in} \it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$ \it
+BIT\_VECTOR $\land$\\
+\hspace*{0.00in} \it bv\_catenate \rm =  $\lambda$  v\rm ,\it w \rm . \rm (\it v 
+$\in$  \it BIT\_VECTOR $\land$ \it w $\in$  \it BIT\_VECTOR  $\mid$  \it v\^ \it w\rm ) %$\cat$
 
@@
 \mathit{bv\_not} \in \mathit{BIT\_VECTOR} \fun \mathit{BIT\_VECTOR} \land \\
-\mathit{bv\_not} = \lambda v \bullet (v \in \mathit{BIT\_VECTOR} \mid \\
-\quad \lambda i \bullet 0 .. \mathit{bv\_size}(v)-1 \mid \mathit{bit\_not}(v(i))) \land \\
+\mathit{bv\_not} = \lambda v . (v \in \mathit{BIT\_VECTOR} \mid \quad \lambda i . (1 .. \mathit{bv\_size}(v)) \mid \mathit{bit\_not}(v(i))) \land \\
 \mathit{bv\_and} \in \mathit{BIT\_VECTOR} \times \mathit{BIT\_VECTOR} \fun \mathit{BIT\_VECTOR} \land \\
-\mathit{bv\_and} = \lambda v_1, v_2 \bullet (v_1 \in \mathit{BIT\_VECTOR} \land v_2 \in \mathit{BIT\_VECTOR} \land \\
-\quad \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \mid \lambda i \bullet 0 .. \mathit{bv\_size}(v_1)-1 \mid \mathit{bit\_and}(v_1(i), v_2(i)))
+\mathit{bv\_and} = \lambda v_1, v_2 . (v_1 \in \mathit{BIT\_VECTOR} \land v_2 \in \mathit{BIT\_VECTOR} \land \\
+\quad \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \mid \lambda i . (1 .. \mathit{bv\_size}(v_1)) \mid
+\mathit{bit\_and}(v_1(i), v_2(i)))
 \end{array}
@@
 as well as classical algebraic properties such as associativity
-and commutativity. For example:
-
-$
-\begin{array}{l}
-\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \mathit{bv\_size}(\mathit{bv\_not}(v)) = \mathit{bv\_size}(v)) \\
-
-\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \mathit{bv\_not}(\mathit{bv\_not}(v)) = v) \\
-
-\forall v_1, v_2 \bullet (\{v_1, v_2\} \subseteq \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \implies \\
-\quad \mathit{bv\_size}(\mathit{bv\_and}(v_1, v_2)) = \mathit{bv\_size}(v_1)) \\
-
-\forall v_1, v_2 \bullet (\{v_1, v_2\} \subseteq \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \implies \\
-\quad \mathit{bv\_size}(\mathit{bv\_and}(v_1, v_2)) = \mathit{bv\_size}(v_2)) \\
-
-\forall v_1, v_2 \bullet (\{v_1, v_2\} \subseteq \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \implies \\
-\quad \mathit{bv\_size}(\mathit{bv\_and}(v_1, v_2)) = \mathit{bv\_and}(v_2, v_1)) \\
-
-\forall v_1, v_2, v_3 \bullet (\{v_1, v_2, v_3\} \subseteq \mathit{BIT\_VECTOR}
-\land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \land \\
-\quad \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_3) \implies \\ 
-\quad \mathit{bv\_and}(\mathit{bv\_and}(v_1, v_2), v_3) = \mathit{bv\_and}(v_1,
-\mathit{bv\_and}(v_2, v_3))) \\
-
-\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \\
-\quad \mathit{bv\_and}(v, \mathit{bv\_zero}(\mathit{bv\_size}(v))) = \mathit{bv\_zero}(\mathit{bv\_size}(v))) \\
-
-\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \\
-\quad \mathit{bv\_and}(v, \mathit{bv\_one}(\mathit{bv\_size}(v))) = v) \\
-
-\end{array}
-$
-
-\subsection{Modelling bytes and bit vector of length 16}[I need review the english]
+and commutativity.
+
+\subsection{Modelling Bytes and Bit Vectors of Length 16}
 
@@
 
+\hspace*{0.0in}\it BYTE\_WIDTH \rm = 8 $\land$ \it BYTE\_INDEX \rm = 1 $\upto$ \rm  BYTE\_WIDTH\rm $\land$
+
+\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.10in} $\land$
+
+\hspace*{0.0in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \}\hspace*{0.10in} $\land$
+
+\hspace*{0.0in}\it BYTE\_ZERO  $\in$  \it BYTE  $\land$ \it BYTE\_ZERO \rm = \it BYTE\_INDEX  $\times$  \rm \{\rm 0\rm \}
+
+The $\mathit{BYTE\_INDEX}$ is the domain of the functions modelling bytes. It starts at 1 to obey a
+definition of sequences from B. However, it is common in hardware architectures to start indexing from
+zero. The definition $\mathit{PHYS\_BYTE\_INDEX}$ is used to provide functionalities obeying this
+convention. The $\mathit{BYTE}$ type is a specialized type from $\mathit{BIT\_VECTOR}$, but it has a size
+limit. Other specific definitions are provided to facilitate further modelling: the type $\mathit{BV16}$
+is created for bit vector of length 16 in a similar way.
+
+\subsection{Bit Vector Arithmetics}
+
+Bit vectors are used to represent and combine numbers: integer ranges (signed or unsigned). Therefore, our
+library includes functions to manipulate such data, for example, the function $\mathit{bv\_to\_nat}$ that
+maps bit vectors to natural numbers:
+
+$
+\begin{array}{l}
+\mathit{bv\_to\_nat} \in \mathit{BIT\_VECTOR} \fun \nat \land \\
+\mathit{bv\_to\_nat} = \lambda v . (v \in \mathit{BIT\_VECTOR} \mid \sum i . (i \in \dom(v) . v(i)
+\times 2^{i-1}))
+\end{array}
+$
+
+An  associated lemma is: $\forall n . (n \in \nat_1 \implies \mathit{bv\_to\_nat}(\mathit{nat\_to\_bv}(n)) = n)$
+
+\subsection{Basics Data Types}
+
+The instruction set of microcontrollers usually have common data types. These types are placed in the
+types library. Each type module has functions to manipulate and convert its data. There are six common
+basics data types represented by modules, see details in table~\ref{tab:types}.
+
+\begin{table}
+\caption{Descriptions of basic data types}
+\label{tab:types}
+
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\hline
+ $Type\ Name$ & UCHAR & SCHAR & USHORTINT & SSHORTINT  & BYTE & BV16 \\\hline
+ $Range$ & 0..255 & -128..127 & 0..65.535  & -32.768..32.767  & -- & --\\ \hline
+ $Physical\ Size $ & 1 byte & 1 byte & 2 bytes & 2 bytes &  1 bytes & 2 bytes \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+Usually, each type module just needs to instantiate concepts that were already defined in the hardware
+modelling library.  For example, the function $\mathit{bv\_to\_nat}$ from bit vector arithmetics is
+specialized to $\mathit{byte\_uchar}$. As the set $\mathit{BYTE}$ is a subset of the
+$\mathit{BIT\_VECTOR}$, this function can defined as follows:
+
+$
+\begin{array}{l}
+\mathit{byte\_uchar} \in \mathit{BYTE} \fun \nat \land \\
+\mathit{byte\_uchar} = \lambda (v) . ( v \in BYTE | bv\_to\_nat(v) )
+\end{array}
+$
+
+The definitions of the library types reuse the basic definitions from the hardware library. This provides
+greater confidence and facilitates the proof process, because the prover can reuse the previously defined
+lemma.
+
+The inverse function $\mathit{uchar\_byte}$ is easily defined:
+
+$
+\begin{array}{l}
+\mathit{uchar\_byte} \in \mathit{UCHAR}  \fun  \mathit{BYTE}  \land \\
+   \mathit{uchar\_byte} = \ (\mathit{byte\_uchar}) ^{-1}
+\end{array}
+$
+
+% We also created the following lemmas:
+%
 % $
 % \begin{array}{l}
-% \mathit{BYTE\_WIDTH} = 8 \land \mathit{BYTE\_INDEX} = 0..(\mathit{BYTE\_WIDTH}-1) \land \\
-% \mathit{BYTE} \subseteq \mathit{BIT\_VECTOR} \land \mathit{BYTE} = \{ v \in \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v) = \mathit{BYTE\_WIDTH}\} \land \\
-% \mathit{BYTE\_ZERO} \in \mathit{BYTE} \land \mathit{BYTE} = \mathit{BYTE\_INDEX} \times \{0\}
+%  \forall (val) . (val \in \mathit{UCHAR} |
+%  \mathit{byte\_uchar}(\mathit{uchar\_byte}(val)) = val) \land\\
+%  \forall (by) . (by \in \mathit{BYTE} |
+%  \mathit{uchar\_byte}(\mathit{byte\_uchar}(by)) = by)
 % \end{array}
 % $
-% Version 2
-
-\hspace*{0.0in}
-
-\hspace*{0.0in}\it BYTE\_WIDTH \rm = 8 $\land$ \it BYTE\_INDEX \rm = 1 $\upto$ \rm (\it BYTE\_WIDTH\rm )$\land$
-
-\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.10in} $\land$
-
-\hspace*{0.0in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \}\hspace*{0.10in} $\land$
-
-\hspace*{0.0in}\it BYTE\_ZERO  $\in$  \it BYTE  $\land$
-
-\hspace*{0.0in}\it BYTE\_ZERO \rm = \it BYTE\_INDEX  $\times$  \rm \{\rm 0\rm \}
-
-\hspace*{0.0in}
-
-The $\mathit{BYTE\_INDEX}$ is domain of byte modeled. it start at 1 to  obey a definition of sequences from B
-method. However, the functions from byte encapsulate the acces and these functions use the
-$\mathit{PHYS\_BYTE\_INDEX}$. The $\mathit{BYTE}$ type is a specializated type from $\mathit{BIT\_VECTOR}$, but it
-has a size limit. Some other more specific definitions are created to help the developer. For example the constant
-$\mathit{BYTE\_ZERO}$ and some below functions.
-
-\begin{itemize}
-  \item \it is\_zero $\in$ BYTE  $\fun$   BIT \rm - It returns 1 if the received byte is zero, otherwise
-  returns 0.
-
-  \item \it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE \rm - It receives a byte and returns its rotated to
-  right. There is a similar function named $\mathit{rotateleft}$.
-
-\end{itemize}
-
-By the very similar way, the type $\mathit{BV16}$ is created.
-
-\subsection{Bit vector arithmetics}
-
-Bit vectors are used to represent and combine numbers: integer ranges
-(signed or unsigned) and floating point numbers.
-
-Our library defines a function $\mathit{bv\_to\_nat}$ that maps bit
-vectors to natural numbers:
-
-$
-\begin{array}{l}
-\mathit{bv\_to\_nat} \in \mathit{BIT\_VECTOR} \fun \nat \land \\
-\mathit{bv\_to\_nat} = \lambda v \bullet (v \in \mathit{BIT\_VECTOR} \mid \sum i \bullet (i \in \dom(v) \bullet v(i) \times 2^i))
-\end{array}
-$
-
-We provide the following lemma:
-
-$
-\begin{array}{l}
-\forall n \bullet (n \in \nat_1 \implies \mathit{bv\_to\_nat}(\mathit{bv\_zero}(n)) = 0)
-\end{array}
-$
-
-\subsection{Basics data types}
-
-There are data types commonly used in the microcontrollers. They are placed in
-the project Types Librarys and define functions to manipulate data on levels of
-basic types from hardware. However, in general is needed to detail or redefines
-the types and its functions for each platform. Then, the specific concepts are
-defined in the project's own platform.
-
-There are six common basics data types, see details in table below.
-
-\begin{tabular}{|p{3.5cm}|c|c|c|}\hline
- $Type\ Name$ & $Range$ & $Physical\ Size $\\ 
-\hline
- UCHAR& 0..255 & 1 byte\\ \hline SCHAR& -128..127 & 1 byte\\ \hline BYTE & -- & 1
- byte\\ \hline USHORTINT & 0..65.535 & 2 byte\\ \hline SSHORTINT &
- -32.768..32.767 & 2 byte\\ \hline BV16 & -- & 2 byte\\ \hline
-\end{tabular}
-
-There is, for example, the function $\mathit{bv\_to\_nat}$ that is specialized
-to $\mathit{byte\_uchar}$. How the set $\mathit{BYTE}$ is a subset of the
-$\mathit{BIT\_VECTOR}$, then this function can defined as follows:
-
-$
-\begin{array}{l}
-\mathit{byte\_uchar} = \lambda (v) \bullet ( v \in BYTE | bv\_nat(v) )
-\end{array}
-$
-
-Usually, the definitions of the functions of the module $\mathit{TYPES}$ are the
-re-used from the basic functions of the hardware library. This provides greater
-confidence and facilitates the process of proof, because the prover can use the
-defined previous lemma.
-
-A more specific model of $\mathit{byte\_uchar}$ functions is:
-
-$
-\begin{array}{l}
-\mathit{byte\_uchar} \in \mathit{BYTE}  \fun  \mathit{UCHAR}  \land \\
-   \mathit{byte\_uchar} =  \lambda ( v0 )  \bullet  
-   ( v0 \in \mathit{BYTE} |  2^{7}* bv\_get( v0 ,7 ) + 2^{6} * bv\_get( v0 ,6 ) + 2^{5}* bv\_get( v0 , 5   )\\ 
-   + 2^{4}* bv\_get( v0 , 4 ) + 2^{3}*bv\_get( v0 , 3 ) + 2^{2}* bv\_get( v0 , 2 ) + 2 * bv\_get( v0 , 1 ) +  bv\_get( v0 ,0) )
-
-\end{array}
-$
-
-The inverse function is easily defined as $\mathit{uchar\_byte}$.
-
-$
-\begin{array}{l}
-\mathit{uchar\_byte} \in \mathit{UCHAR}  \fun  \mathit{BYTE}  \land \\
-   \mathit{uchar\_byte} = \  ^{-1}(\mathit{byte\_uchar}
-   )
-\end{array}
-$
-
-We also created the following lemmas:
-
-$
-\begin{array}{l}
- \forall (val) \bullet (val \in \mathit{UCHAR} |
- \mathit{byte\_uchar}(\mathit{uchar\_byte}(val)) = val) \land\\
- \forall (by) \bullet (by \in \mathit{BYTE} |
- \mathit{uchar\_byte}(\mathit{byte\_uchar}(by)) = by)
-\end{array}
-$
-
-Similarly, several other functions and lemmas were created for all other
-data types.
-
-\section{Description of the B model from Z80}
+
+Similarly, several other functions and lemmas were created for all other data types.
+
+\section{Description of the Z80 B model }
 \label{sec:z80}
 
-The \textit{Z80} is an important microcontroller developed in 1976 by \textit{Zilog}. In this time, It got a great
-acceptance because suported all set instruction of \textit{8080}, one of the predecessors of the \textit{Pentium}
-from \textit{Intel Corporation}. Then, the Z80 has the set instructions similar to the x86 architecture, it helps 
-to build the model of x86 and it provides good features.
-
-The main module includes a instance of memory module and accesses the definitions from: basic data types modules,
-\textit{ALU} and \textit{Power2} \footnote{The \textit{Power2} module has the basics definitions to help the
-theorems prover about calculus of power.}.
+The \textit{Z80} is a CISC microcontroller developed by \textit{Zilog}~\cite{Z80_manual}. It supports 158
+different instructions, including all the 78 from Intel 8080 microprocessor, and all of them were
+specified. These instructions are classified into these categories: load and exchange; block transfer and
+search; arithmetic and logical; rotate and shift; bit manipulation; jump, call and return; input/output;
+and basic cpu control.
+
+The main module includes an instance of the memory module and accesses the definitions from basic data
+types modules and the \textit{ALU} module.
 
@@
 
-\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION, \it POWER2
+\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION
 \end{sloppypar}
@@
 % Chacteristics of microcontroller
-The \textit{Z80} supports 158 differents instructions including all the 78 from
-\textit{8080}.The instructions are classified into these categories: load and exchange;
-block transfer and search; arithmetic and logical; rotate and shift; bit
-manipulation (set, reset, test); jump, call and return; input/output; and basic
-cpu control. Each instruction and external action at model is represented by B
-operations. By default, all parameters from operations are or pre-defined elements
-in the model or integers values in the decimal representation.
-
-The internal registers contain 208 bits of reading/writing memory. It includes two sets of six general purpose
-registers which may be used individually as  8-bits registers or as 16-bits register pairs.  The work registers are
-represented by variable $\mathit{rgs8}$. The domain of $\mathit{rgs8}$ ($\mathit{id\_regs8}$) is a set formed by
-identifiers of registers of 8 bits. These registers can be accessed in pairs, forming 16-bits, resulting in other
-set of identifiers of 16-bits registers, named $\mathit{id\_reg16}$. The main work register of Z80 is the
-accumulator ($\mathit{rgs8(a0)}$) used for arithmetic, logic, input/output and loading/storing operations.
+ The internal registers contain 176 bits of reading/writing memory that are represented by
+identifiers used as parameters in the instructions . It includes two sets of six general purpose
+registers which may be used individually as 8-bits registers or as 16-bits register pairs.  The working registers
+are represented by variable $\mathit{rgs8}$. The domain of $\mathit{rgs8}$ ($\mathit{id\_regs8}$) is a set
+formed by identifiers of registers of 8 bits. These registers can be accessed in pairs, forming 16-bits,
+resulting in another set of identifiers of 16-bits registers, named $\mathit{id\_reg16}$.
 
@@
 \end{sloppypar}
+
+The main working register of Z80 is the accumulator ($\mathit{rgs8(a0)}$) used for arithmetic/logic,
+input/output and loading/storing operations.
+
+% \subsection{Modelling Registers, Input and Output Ports and Instructions}
+% 
+% The Z80 has different types of registers and instructions. The CPU
+% contains general-purpose registers ($\mathit{id\_reg\_8}$), a stack
+% pointer ($\mathit{sp}$), program counter ($\mathit{pc}$), two index
+% registers ($\mathit{ix}$ and $\mathit{iy}$), an interrupt register
+% ($\mathit{i\_}$), a refresh register ($\mathit{r\_}$), two bits
+% ($\mathit{iff1}$, $\mathit{iff2}$) used to control the interruptions,
+% a pair of bits to define the interruption mode ($\mathit{im}$) and the
+% input and output ports ($\mathit{i\_o\_ports}$). Below, part of the
+% corresponding definitions are replicated from the $\INVARIANT$:
+% 
+% \begin{sloppypar}
+% %\bf INVARIANT
+% \hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ pc  $\in$  \it INSTRUCTION  $\land$
+% 
+% \hspace*{0.10in}\it  sp  $\in$  \it BV16  $\land$  \it ix  $\in$  \it BV16  $\land$  \it iy  $\in$  \it BV16  $\land$
+% 
+% \hspace*{0.10in}\it i\_  $\in$  \it BYTE  $\land$  \it r\_ $\in$  \it BYTE  $\land$ iff1  $\in$  \it BIT
+% $\land$ \it iff2  $\in$  \it BIT  $\land$
+% 
+% \hspace*{0.10in}\it im \rm $\in$ \rm (\it BIT $\times$ \it BIT\rm )  $\land$ i\_o\_ports  $\in$  \it BYTE
+% $\fun$  \it BYTE
+% \end{sloppypar}
+% 
+% A simple example of instruction is a $\mathit{LD\_n\_A}$, as shown
+% below.  Many times, to model an instruction is necessary to use the
+% predefined functions, these help the construction of model. This
+% instruction use the $\mathit{updateAddressMem}$ function from
+% \textit{Memory} module and it receives an address memory and its new
+% memory value. Finally it increments the program counter
+% ($\mathit{pc}$) and update the refresh register ($\mathit{r\_}$).
+% 
+% \hspace*{0.00in}\bf LD\_n\_A \rm ( \it nn \rm ) \rm =
+% 
+% \hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} %$\land$  \it nn $\in$\it DATA\_R\_ADR
+% 
+% \hspace*{0.20in}\bf THEN
+% 
+% \hspace*{0.20in}\bf updateAddressMem \rm ( \it ushort\_to\_bv16 \rm ( \it nn \rm ) \rm , \it rgs8 \rm ( \it a0 \rm )
+% \rm )  $\para$
+% 
+% \hspace*{0.20in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it
+% update\_refresh\_reg\rm (\it r\_\rm )
+% 
+% \hspace*{0.00in}\bf END\rm
+% 
+% The microcontroller model can specify security properties. For example, the
+% last operation could have a restriction to write only in a defined region of memory.
 
@@
 \end{sloppypar}
+
+%
 
@@
 \begin{property}[Assuring the absence of Overflow]
- To assure that don´t happen overflow the developer can add this expression ($\mathit{bv\_get(rgs8(z0),0)} \neq 1
- \land \mathit{bv\_get(rgs8(z0),2) \neq 1}$)in the invariant. By default, the overflow can happen, but many times it
- can be dangerous. Then, the developer may prohibit its use. This restriction can also become the more dificult to
+ To assure that don¥t happen overflow the developer can add this expression ($\mathit{bv\_get(rgs8(z0),0)} \neq 1
+ \land \mathit{bv\_get(rgs8(z0),2) \neq 1}$) in the invariant. By default, the overflow can happen, but
+ many times it can be dangerous. Then, the developer may prohibit its use. This restriction can also become the more dificult to
  specify.
@@
 
-The main module (\textit{Z80}) has two types of operations, a type represent the external actions and the other
-represent the instructions of microcontrollers. The external actions are shown in the section
-~\ref{sec:externalactions} and the other type represents the instructions of microcontroller. A simple example from
-instruction is a $\mathit{LD\_(nn)\_A}$ shown below. Many times to model a instruction is necessary to use the pre-defined functions,
-this facilitates the construction of model. This instruction use the $\mathit{updateAddressMem}$ function from
-\textit{Memory} module and it receives a address memory and its new memory value. After, it increments the program
-counter ($\mathit{pc}$) and update the refresh register ($\mathit{r\_}$).
+Each instruction is represented by a B operation in the module Z80. The main module (\textit{Z80}) has two
+types of operations, a type represents the external actions and the other represents the instructions of
+microcontrollers. The external actions are shown in the section ~\ref{sec:externalactions}.  A simple
+example from instruction is a $\mathit{LD\_(nn)\_A}$ shown below. Many times to model a instruction is
+necessary to use the pre-defined functions, this facilitates the construction of model. By default, all
+parameters from operations are either predefined elements in the model or integers values in the decimal
+representation. This instruction use the $\mathit{updateAddressMem}$ function from \textit{Memory} module
+and it receives a address memory and its new memory value. After, it increments the program counter
+($\mathit{pc}$) and update the refresh register ($\mathit{r\_}$).
 
@@
 
-\subsection{Modelling the external action}
-\label{sec:externalactions}
-
-The external actions change the state of microcontroller and they aren´t instructions, for example, the refreshing
-of I/O port and the interruptions request. The extern action are also model by operations and they are prefixed
-with ``$ext\_$'' and after the name of action. There are just four extern actions: $ext\_update\_io\_ports$,
-$ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The $ext\_update\_io\_ports$ just change the state of I/O port, see.
-
-\hspace*{0.20in}\bf ext\_update\_io\_ports\rm (\it address\rm ,\it value\rm )\rm =
-
-\hspace*{0.20in}\bf PRE \it address  $\in$  \it UCHAR  $\land$ \hspace*{0.10in}\it value  $\in$  \it SCHAR \bf THEN
-
-\hspace*{0.40in}\it io\_ports \rm ( \it uchar\_byte \rm ( \it address \rm ) \rm ) \rm := \it schar\_byte \rm ( \it
-value \rm )
-
-\hspace*{0.20in}\bf END\rm
-
-The others are related to interrupts. The interrupts allows that the devices suspend a routine from CPU and start
-another service routine. This service routine can exchange data or signals between CPU and external devices. When a
-routine is finished, then the CPU come back to the last routine that was interrupted.
-
-For the interrupts, the following things are important:  the interrupt flip-flops ($\mathit{iff1}$ and
-$\mathit{iff2}$), the types of interrupts (maskable and non-maskable), the interrupt mode (set with the $\mathit{IM
-0}$, $\mathit{IM 1}$, $\mathit{IM 2}$ instructions) and the $\mathit{i\_}$ register.
-
-The $\mathit{iff1}$ and $\mathit{iff2}$ control the maskable interrupts ($\mathit{INT}$). When the $\mathit{iff1}$
-is set, the interrupt is enable, otherwise it is disable. The $\mathit{iff2}$ is used only as a tempory storage
-place for $\mathit{iff1}$. The $\mathit{iff1}$ is set or reset by instructions $\mathit{EI}$ and $\mathit{DI}$.
-
-The interruptions and the \textit{reset} action change the state of
-microcontroller. Theses actions are modeled by B operations and the its main
-effects are shown below\footnote{Some definitions: $\mathit{sp\_minus\_two}$ =
-$\mathit{dec\_BV16(dec\_BV16(sp))}$, $\mathit{sp\_minus\_one}$=
-$\mathit{dec\_BV16(sp)}$, $\mathit{pc\_high}$ is the most significant 8 bits and
-$\mathit{pc\_low}$ is the least significant. }.
-
- \textbf{NMI} - Non-maskable interrupts - The non-maskable cannot be disable
- by the programmer. Then, when a device makes a request, the $sp$ is pushed, the $pc$ receive
- $66H$ (102 in decimal), the $\mathit{iff1}$ is reset , $\mathit{iff2}$ stores
- $\mathit{iff1}$ and refresh register is updated.
-
-\begin{sloppypar}
-\bf updateStack\rm (\rm \{ \rm (\it sp\_minus\_two  $\mapsto$  \it pc\_low\rm )\rm ,\rm (\it sp\_minus\_one  $\mapsto$ \it pc\_high \rm ) \rm \}\rm )$\para$
-
-\it sp \rm := \it sp\_minus\_two  $\para$ \it pc \rm := \rm 1\rm 0\rm 2 $\para$ \it iff1\rm :=\rm 0  $\para$  \it iff2\rm := \it iff1 $\para$
-
-\it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\\
-\end{sloppypar}
-
-  \textbf{INT} - Maskable Interrupt -  This is usualy reserved to important functions that can be enabled and
-  disabled by the programmer. When a maskable interrupt action happens, both $\mathit{iff1}$ and $\mathit{iff2}$ are
-  cleared, disabling the interrupts, the $sp$ is pushed, the refresh register is updated  and the other effects
-  depend on the interrupt mode.
-
- \begin{itemize}
-
-  \item The mode 0 is compatible with 8080 and  \it im \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a 
-  non-maskable interruption happen , the CPU fetches an instruction of one byte from an external device, usually an
-  RST instruction, and the CPU executes it. The instruction code is received from a external device by data bus and
-  it is represented by integer parameter called $\mathit{byte\_bus}$ .
-
-  \item The mode 1 is the easiest and \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 1 \rm ). Simply, when a non-maskable
-  interruption happens, the program counter receive $38H$ (56 in decimal).
-
-  \item The mode 2 is the most powerfull and  \it im \rm = \rm ( \rm 1  $\mapsto$ \rm 1 \rm ). When a non-maskable
-  interruption happens, an indirect call can be made to any address memory. The program counter receives in the
-  part most significant the $\mathit{i\_}$ register and, in the  part least siginificant, the $\mathit{byte\_bus}$
-  with the last bit reset.
- \end{itemize}
-
-The essential part of maskable interrupt is shown below\footnote{ The
- $\mathit{byte\_bus}$ is parameter from  $\mathit{INT}$ operation }.
-
-\begin{sloppypar}
-
-\hspace*{0.00in}\bf IF \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 0 \rm ) \bf THEN
-
-\hspace*{0.10in}\bf IF\hspace*{0.10in}\it byte\_bus \rm $\in$  \it opcodes\_RST\_instruction
-
-\hspace*{0.10in}\bf THEN\hspace*{1.05in}
-
-\hspace*{0.40in}\it pc \rm := \it byte\_bus \rm - \rm 1\rm 9\rm
-9\hspace*{0.10in} $\para$
-
-\hspace*{0.40in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
-sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
-
-\hspace*{0.40in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
-\rm \} \rm )  $\para$
-
-\hspace*{0.40in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
-r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
-
-\hspace*{0.10in}\bf ELSIF \it byte\_bus \rm = \it opcode\_\ldots\_instruction
-
-\hspace*{0.40in}\bf \ldots
-
-\hspace*{0.10in}\bf END
-
-\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm =\hspace*{0.10in}\rm ( \rm
-0  $\mapsto$  \rm 1 \rm ) \bf THEN
-
-\hspace*{0.10in}\it pc \rm :=\hspace*{0.10in}\rm 5\rm 6  \hspace*{0.80in}
-$\para$
-
-\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm (\it
-sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
-
-\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
-\rm \} \rm )  $\para$
-
-\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
-r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\hspace*{0.85in}
-
-\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm = \rm ( \rm 1  $\mapsto$ 
-\rm 1 \rm ) \bf THEN  \hspace*{0.70in}
-
-\hspace*{0.10in}\it pc \rm := \it bv16\_ushort\rm (\it byte\_bv16\rm ( \it i\_
-\rm ,\it bv\_clear\rm (\it rotateleft\rm (\it uchar\_byte\rm (\it byte\_bus\rm )\rm )\rm ,\rm 0\rm )\rm )\rm )$\para$
-
-\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
-sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
-
-\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
-\rm \} \rm )  $\para$
-
-\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
-r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
-
-\hspace*{0.00in}\bf END\hspace*{0.10in}\\
-\end{sloppypar}
-
-\textbf{RESET}  - This just reset the registers related to the interruptions.
-
-\begin{sloppypar}
-\it iff1 \rm :=\rm 0 $\para$ \it iff2\rm :=\rm 0 $\para$  \it  im\rm := \rm (\rm 0 $\mapsto$ \rm 0\rm )  $\para$ 
-\it pc\rm :=\rm 0 $\para$ \it i\_ \rm := [0,0,0,0,0,0,0,0] $\para$
-
-\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm (\it a0  $\mapsto$ \rm [0,0,0,0,0,0,0,0] , \rm (\it
-f0 $\mapsto$ \rm [0,0,0,0,0,0,0,0] \rm \} $\para$
-
-\it r\_  \rm := [0,0,0,0,0,0,0,0] $\para$ \it sp \rm := [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0]
-\end{sloppypar}
-
-% 
-%     Para controlar o registrador I, existem duas intrucoes:
-% 
-%     - LD A,I (carrega o acumulador com o valor contido em I);
-%     - LD I,A (carrega I com o valor contido no acumulador).
-% 
-%     A instrucao LD A,I possui uma funcao extra, ja' discutida
-% em outra aula: ela carrega na `flag' de paridade/sobrecarga o
-% valor do IFF ("Interruption Flip-Flop"). Se IFF for 0, entao a
-% interrupcao mascaravel esta' desativada. Se IFF for 1, entao
-% ela esta' ativada.
-% %%%% FIM - http://download.unesp.br/msx/asm/course_4.txt
-% 
-% 
-% 
-% 
-% Instruções relacionadas com o sistema de interrupção:
-%  HALT,EI,DI,M0,M1,M2 
-%  
-%  LD_A_R, LD_A_I\ldots,  
-%  CALL, RET,  RETI,RETN
-
-\section{Input and output ports}
+\subsection{Modelling the Input/Output Instructions}
 
@@
 
-\section{Proofs }
-
-[O que as provas garantem?] As obrigações de provas garante a verficação dos tipos de dados, das propriedades
-estabelecidas no modelo e se as expressões são bem formadas \footnote{An expression is called "well-defined" (or
-unambiguous) if its definition assigns it a unique interpretation or value.}. As propriedades oferecem garantias
-adicionais ao projetista do software, pois é capaz de estabelecer uma série de critérios de segurança. Além disso,
-como estas propriedades estão em nível de especificação, então é possível ajusta-las a necessidade de um determinado
-sofwtare sem implicar em mudanças no hardware, o que implica apenas em aumentar ou diminuir o número de provas.
-% O desenvolvedor poderia por exemplo aumentar a região de memória necessária para armazenar dados de trabalho.
-A seguir tem-se os número de provas não obvias e WD do projeto Z80 separado em grupos. Esta tabela mostra a
-distribuição do número de provas de cada grupo. Um dos grupos que contém mais provas é o de manipulação de bits.
-Apesar da grande quantidade de provas, este grupo foi rapidamente provado, pois ele utiliza vários lemmas providos
-pela Hardware Library, o que acontece também com o grupo Logic and Arithmetic e a Types Library.
-
-[Groups]
+\subsection{Modelling the external action}
+\label{sec:externalactions}
+
+The external actions change the state of microcontroller and they aren¥t instructions, for example, the refreshing
+of I/O port and the interruptions request. The extern action are also model by operations and they are prefixed
+with ``$ext\_$'' and after the name of action. There are just four extern actions: $ext\_update\_io\_ports$,
+$ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The $ext\_update\_io\_ports$ just change the state of I/O port, see.
+
+\hspace*{0.20in}\bf ext\_update\_io\_ports\rm (\it address\rm ,\it value\rm )\rm =
+
+\hspace*{0.20in}\bf PRE \it address  $\in$  \it UCHAR  $\land$ \hspace*{0.10in}\it value  $\in$  \it SCHAR \bf THEN
+
+\hspace*{0.40in}\it io\_ports \rm ( \it uchar\_byte \rm ( \it address \rm ) \rm ) \rm := \it schar\_byte \rm ( \it
+value \rm )
+
+\hspace*{0.20in}\bf END\rm
+
+The others are related to interrupts. The interrupts allows that the devices suspend a routine from CPU and start
+another service routine. This service routine can exchange data or signals between CPU and external devices. When a
+routine is finished, then the CPU come back to the last routine that was interrupted.
+
+For the interrupts, the following things are important:  the interrupt flip-flops ($\mathit{iff1}$ and
+$\mathit{iff2}$), the types of interrupts (maskable and non-maskable), the interrupt mode (set with the $\mathit{IM
+0}$, $\mathit{IM 1}$, $\mathit{IM 2}$ instructions) and the $\mathit{i\_}$ register.
+
+The $\mathit{iff1}$ and $\mathit{iff2}$ control the maskable interrupts ($\mathit{INT}$). When the $\mathit{iff1}$
+is set, the interrupt is enable, otherwise it is disable. The $\mathit{iff2}$ is used only as a tempory storage
+place for $\mathit{iff1}$. The $\mathit{iff1}$ is set or reset by instructions $\mathit{EI}$ and $\mathit{DI}$.
+
+The interruptions and the \textit{reset} action change the state of
+microcontroller. Theses actions are modeled by B operations and the its main
+effects are shown below\footnote{Some definitions: $\mathit{sp\_minus\_two}$ =
+$\mathit{dec\_BV16(dec\_BV16(sp))}$, $\mathit{sp\_minus\_one}$=
+$\mathit{dec\_BV16(sp)}$, $\mathit{pc\_high}$ is the most significant 8 bits and
+$\mathit{pc\_low}$ is the least significant. }.
+
+ \textbf{NMI} - Non-maskable interrupts - The non-maskable cannot be disable
+ by the programmer. Then, when a device makes a request, the $sp$ is pushed, the $pc$ receive
+ $66H$ (102 in decimal), the $\mathit{iff1}$ is reset , $\mathit{iff2}$ stores
+ $\mathit{iff1}$ and refresh register is updated.
+
+\begin{sloppypar}
+\bf updateStack\rm (\rm \{ \rm (\it sp\_minus\_two  $\mapsto$  \it pc\_low\rm )\rm ,\rm (\it sp\_minus\_one  $\mapsto$ \it pc\_high \rm ) \rm \}\rm )$\para$
+
+\it sp \rm := \it sp\_minus\_two  $\para$ \it pc \rm := \rm 1\rm 0\rm 2 $\para$ \it iff1\rm :=\rm 0  $\para$  \it iff2\rm := \it iff1 $\para$
+
+\it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\\
+\end{sloppypar}
+
+  \textbf{INT} - Maskable Interrupt -  This is usualy reserved to important functions that can be enabled and
+  disabled by the programmer. When a maskable interrupt action happens, both $\mathit{iff1}$ and $\mathit{iff2}$ are
+  cleared, disabling the interrupts, the $sp$ is pushed, the refresh register is updated  and the other effects
+  depend on the interrupt mode.
+
+ \begin{itemize}
+
+  \item The mode 0 is compatible with 8080 and  \it im \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a 
+  non-maskable interruption happen , the CPU fetches an instruction of one byte from an external device, usually an
+  RST instruction, and the CPU executes it. The instruction code is received from a external device by data bus and
+  it is represented by integer parameter called $\mathit{byte\_bus}$ .
+
+  \item The mode 1 is the easiest and \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 1 \rm ). Simply, when a non-maskable
+  interruption happens, the program counter receive $38H$ (56 in decimal).
+
+  \item The mode 2 is the most powerfull and  \it im \rm = \rm ( \rm 1  $\mapsto$ \rm 1 \rm ). When a non-maskable
+  interruption happens, an indirect call can be made to any address memory. The program counter receives in the
+  part most significant the $\mathit{i\_}$ register and, in the  part least siginificant, the $\mathit{byte\_bus}$
+  with the last bit reset.
+ \end{itemize}
+
+The essential part of maskable interrupt is shown below\footnote{ The
+ $\mathit{byte\_bus}$ is parameter from  $\mathit{INT}$ operation }.
+
+\begin{sloppypar}
+
+\hspace*{0.00in}\bf IF \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 0 \rm ) \bf THEN
+
+\hspace*{0.10in}\bf IF\hspace*{0.10in}\it byte\_bus \rm $\in$  \it opcodes\_RST\_instruction
+
+\hspace*{0.10in}\bf THEN\hspace*{1.05in}
+
+\hspace*{0.40in}\it pc \rm := \it byte\_bus \rm - \rm 1\rm 9\rm
+9\hspace*{0.10in} $\para$
+
+\hspace*{0.40in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
+sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
+
+\hspace*{0.40in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
+\rm \} \rm )  $\para$
+
+\hspace*{0.40in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
+r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
+
+\hspace*{0.10in}\bf ELSIF \it byte\_bus \rm = \it opcode\_\ldots\_instruction
+
+\hspace*{0.40in}\bf \ldots
+
+\hspace*{0.10in}\bf END
+
+\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm =\hspace*{0.10in}\rm ( \rm
+0  $\mapsto$  \rm 1 \rm ) \bf THEN
+
+\hspace*{0.10in}\it pc \rm :=\hspace*{0.10in}\rm 5\rm 6  \hspace*{0.80in}
+$\para$
+
+\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm (\it
+sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
+
+\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
+\rm \} \rm )  $\para$
+
+\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
+r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\hspace*{0.85in}
+
+\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm = \rm ( \rm 1  $\mapsto$ 
+\rm 1 \rm ) \bf THEN  \hspace*{0.70in}
+
+\hspace*{0.10in}\it pc \rm := \it bv16\_ushort\rm (\it byte\_bv16\rm ( \it i\_
+\rm ,\it bv\_clear\rm (\it rotateleft\rm (\it uchar\_byte\rm (\it byte\_bus\rm )\rm )\rm ,\rm 0\rm )\rm )\rm )$\para$
+
+\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
+sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
+
+\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
+\rm \} \rm )  $\para$
+
+\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
+r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
+
+\hspace*{0.00in}\bf END\hspace*{0.10in}\\
+\end{sloppypar}
+
+\textbf{RESET}  - This just reset the registers related to the interruptions.
+
+\begin{sloppypar}
+\it iff1 \rm :=\rm 0 $\para$ \it iff2\rm :=\rm 0 $\para$  \it  im\rm := \rm (\rm 0 $\mapsto$ \rm 0\rm )  $\para$ 
+\it pc\rm :=\rm 0 $\para$ \it i\_ \rm := [0,0,0,0,0,0,0,0] $\para$
+
+\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm (\it a0  $\mapsto$ \rm [0,0,0,0,0,0,0,0] , \rm (\it
+f0 $\mapsto$ \rm [0,0,0,0,0,0,0,0] \rm \} $\para$
+
+\it r\_  \rm := [0,0,0,0,0,0,0,0] $\para$ \it sp \rm := [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0]
+\end{sloppypar}
+
+% 
+%     Para controlar o registrador I, existem duas intrucoes:
+% 
+%     - LD A,I (carrega o acumulador com o valor contido em I);
+%     - LD I,A (carrega I com o valor contido no acumulador).
+% 
+%     A instrucao LD A,I possui uma funcao extra, ja discutida
+% em outra aula: ela carrega na flag de paridade/sobrecarga o
+% valor do IFF (Interruption Flip-Flop). Se IFF for 0, entao a
+% interrupcao mascaravel esta' desativada. Se IFF for 1, entao
+% ela esta ativada.
+% %%%% FIM - http://download.unesp.br/msx/asm/course_4.txt
+% 
+% 
+% 
+% 
+% Instrucoes relacionadas com o sistema de interrupcaoo:
+%  HALT,EI,DI,M0,M1,M2 
+%  
+%  LD_A_R, LD_A_I\ldots,  
+%  CALL, RET,  RETI,RETN
+
+ [To create a better link (safety rules -> Proofs)]
+The microcontroller model can specify security properties. For example, the
+last operation could have a restriction to write only in a defined region of memory.
+
+\section{Proofs}
+\label{sec:Proofs}
+[To get new informations from dissertation]
+% [O que as provas garantem? citar que dificeis, que é necessário buscar técnicas eficientes
+% para realizar as provas ]
+
+The proof obligations allow to verify the data types, important system
+properties and if the expressions are well-defined (WD)\footnote{An
+  expression is called ``well-defined'' (or unambiguous) if its
+  definition assigns it a unique interpretation or value.}. The
+properties provide additional guarantees, because they can set many
+safety rules. However, the model can be very difficult to prove.
+Some statiscts are shown below:
+%[Groups]
 \begin{center}
 \begin{tabular}{llll}
-\textbf{Group} &  \textbf{Proofs No Obvious} & \textbf{Proofs WD} & × \\ 
+\textbf{Group} &  \textbf{Proofs No Obvious} & \textbf{Proofs WD} &   \\ 
 Input and Output & 54 & 296  \\ 
@@
 
-% [Conclusao]
-This project has many proof obligations and its process of verification is the step of the most time consumed. The
-verification step of all the model spent about a month. This was hard, because was necessary to build, correct and
-prove several expressions from basic projects, Hardware and Types Library. However, when the basic projects are
-ready, the modelling and verification of new plataform \footnote{Actualy, the microprocessors M8051 and PIC16F432
-are under construction} become more fast and easy. Thus, these basic projects were also built to re-use to other
-platforms. When the basic projects are ready, then about 89\% proof non obvious and 55\% of Proofs WD was conclused
-automaticly.
-
-The others were concluded through the use of proof commands\footnote{The proof commands are step that just indicate
-a walk to prover make the proof, thus, theses commands cannot introduce false hypotheses.} associated to match
-pattern from proof obligations. The match pattern accelerates too much the prove process, because there are many
-similar assembler instructions and the AtelierB can re-use the match pattern and its proof commands associated.
-Therefore, few proof commands can be useful to many obrigações de prova. A good example is a set of 17 proof
-commands that fastly helps to verify 99\% (2295) of proofs WD. To further reduce cost of proof process, a new feature
-supported by AtelierB was used, the parallelization. To do this were used three computers each one with two cores,
-the that theoretically increased by six times the capacity of processing.
-
-A special care was taken to represent the model in way simple and accurate. This care also reduced the number of
-proofs. Because, the simple act of optimizing the use of override operator ($\lover$) fell too the number of proof
-obligations.
-
-Finally, the techniques mentioned help the verification of all the 2926 proof obligations. The
-modularization of projects difined basic lemmas that help to build a simple and full framework. The
-automatic re-use of commands's proof directed to a resolution more eficiently of proofs. The
-paralelization and the care in modeling accelerate the process's proof, because increase the
-processing capacity and reduced the number of proof obligatiions.
+% [Conclusao] This project has many proof obligations and
+% The construction of then model was a difficult task, as several
+% iterations were needed to provide the good library definitions
+% as well as to optimize the definition of the functionality of
+% the microprocessor instructions by grouping common data
+% manipulation into auxiliary functions.
+
+Several iterations were needed to provide the good library definitions as well as
+to fine-tune the model of the microcontroller instructions by factoring common
+functionalities into auxiliary definitions.
+
+However, few proof commands\footnote{The proof commands are steps that direct the
+prover to find the proof, and cannot introduce false hypothesis.} need to be used
+to prove most proof obligations. As there are many similar assembly instructions,
+some human-directed proofs, when replayed, could discharge other proof
+obligations. A good example is a set of 17 proof commands that quickly aided the
+verification of 99\% (2295) of WD proofs. We also set up a proving environment
+consisting of networked computers to take advantage of the distribution
+facilities now provided in the B development environment. Finally, all of the
+2926 proof obligations were proved using the tool support of the development
+environment.
 
 \section{Related Works}
-
-% This work had as the main motivation the  verififcation in assembler level and consequently the
-% verification of harware model. The model created express the semantic of functioning of hardware and
-% with this model is possible verify the semantic of hardware components.
-
-There are in the literature of computer science some technics\cite{BHDL_2003,GEMPLUS_99} to model hardware and the
-virtual machines using the B method. Then, the B method has been used successfully to model the operational
-semantic, however the cost of modeling was still expensive and this paper describes some techniques to lower the
-cost modeling.
-
-In general, the researchers are concetrated at model the Java Virtual Machine (JVM), but it is not easy to specify 
-it, because it very big and complex, and become more difficult to write good lemmas, to normalize the functions and
-to verify its properties. Then, this work specify a model more simple to get better results fastly, in the same
-time, stay near to model of actual processors.
-% Re-uso  - modularization
-The most of related works are specific to one plataform\cite{GEMPLUS_99,ammm} and its basic definitions cannot be
-re-used at in other platforms. The used strategy created a good logic separation about the concepts of modelling.
-Thus, the basic projects, the Hardware and Types Library, have functions with a low coupling and high cohesion, that
-promove the re-use. Soon, these re-use better too the code generation, the confiability and prove process.
-
- Others advantages of this work are the use of news and advanceds funcionalities of B method. Normaly, the related
- works don`t enjoy this funcionalities, because these are news and required advanced knowleadge about tools for B
- method. These advantages are the paralelization of prove proccess and the instensive use of math patterns
- associated with proof commands.
-
- [AMM]
-Verificação formal já vem sendo aplicada com sucesso na verificação \ldots Porém os custos eram altos, pois se
-passou X tempo para modelar \ldots e
-
-A modelagem do microcontrolador foi construida por um unico designer and consumiu aproximadamente six
-month to study the plattaform Z80, model, adjust to simplifi and prove.
-
-Quais as desvantagens/problemas com relaçao aos outros?
--The B method is not good to verify all kindy of cararacteristics.
--The our results are not mature, b 
--Actually, we don`t define 
--The B method can verify many diferentes properties of model, but some properties cannot be easy to verify and
-model.
- -Some times, the B method is very dificult to model, eventually a other metodology can verify
-
- \section{Conlusitions}
-
- % Trab. Futuros
- %- Refinar o modelo para gerar através dos  código e apartir deste código criar um
- %interpretador/simulador
- %- Construir compilador
-
-%  \begin{itemize}
-%    Citar ambuiguidades encontradas no manual e eventuais erros encontrados.
-%   \item Utilidade o processo de desenvolvimento do hardware:
-%   \subitem Ajuda na remoção de eventuais ambiguidades providas de uma descrição informal.
-%   \subitem Verificação de propriedades: sobre a manipulação e tipo dos dados.
-%   \subitem Documentação formalizada matemáticamente
-%   \subitem Possibilidade da geracao de código VHDL apartir do modelo B.
-%   \subitem Colaboracao para construção de um simulador/interpretador
-% \end{itemize}
-
-[I need review the english]
-% [intro]
-A necessidade comercial para produzir \textit{hardware} e \textit{software} de alta qualidade é sempre
-crescente. O métodos B já têm demonstrado sucesso na especificação de sistemas comerciais críticos.
-
-O suporte fornecido pelo método B é bastante adequado, maduro, extensivo e
-eficiente no processo de verificação dos modelos das plataformas. Isso vem sendo
-perceptível, pois todas as características relevantes das plataformas foram
-apropriadamente modeladas, principalmente as importantes para verificação em
-nível de montagem (\textit{assembly}).
-
-[ O que foi o diferencial ? qual grande vantagem]
-[Eficiência da técnica/método B]
-Por outro lado, uma das qualidades de destaque do método B é relacionada com a
-boa capacidade de verificação automática das ferramentas, pois, na versão estável
-das modelagens, do total das $503$ provas dos módulos principais do \textit{PIC},
-\textit{8051} e \textit{Z80} apenas $3$ referentes ao \textit{Z80} foram provadas
-de forma semi-automática. Ou seja, desse total de provas menos de $1$\%
-necessitaram de uma simples intervenção do projetista para ser concluída.
-
-Adicionalmente, a aplicação das técnicas modularização e re-uso tornou o processo
-de especificação das plataformas mais padronizado e acelerado. Desse modo, é
-demonstrada a viabilidade dessas especificações, pois com a aquisição de
-maturidade das ferramentas e do projetista, o processo de modelagem ganhou
-agilidade e maior automação na fase de validação do projeto.
-
-[maior vantagem]
-Enfim, a modelagem B dessas plataformas possibilitarão a verificação em um dos
-mais altos níveis de certificação de \textit{software}. Não somente, elas podem
-ser importantes no projeto de plataformas para a construção dos simuladores,
-documentação e sua própria verificação formal, permitindo identificar várias
-inconsistências delas em tempo de projeto.
-
-[Enfatizar a vantagem]
-[Flexibilidade]
-The traditional don´t allow verify this things and the properties is flexible. The
-flexbility allow add restrictions about tempo de execuçao and until specifics restrictions about the application of
-microcontroller.
-
-This flexibility of create propertis allow create conditions to verify if the software to finish in cert instant.
-
-No contexto da verificação em nível de assembly:
-Exemplificar com propriedade de restrição temporal.
-Exp. é possível verificar se o software satisfazer restrições temporais e para isso utiliza
-
-%  Para a modelagen dessa plataforma, essa modelagen contém linhas e $np $
-%  provas não obvias. Uma forma de acelerar o processo de prova, foi paralelizar
-%  as provas. Isto foi feito em 4 IMacs e consumiu $ tp $ tempo para verificar
-%  automaticamente $np - p $. As demais foi necessário uma certa interação com
-%  o provador para sua verificação.
-
-%  Para oferecer uma maior garantia do modelo com a implementação do processador
-%  algumas atividades foram realizadas. Primeiramente, tentantamos animar e gerar
-%  código respectivamente nas ferramentas ProB e AtelierB. Entratanto, o modelo
-%  cotém algumas construções complexas que ainda não são bem suportada por elas.
-%  Então, foi realizada uma apresentação sobre a descrição do modelo e foi
-%  realizada um nova revisão com 3 pessoas.
-
-\section{Acknowledges}
-
-This work received supports from ANP (National Agency of Oil) and CNPq. Besides, the researcher Bruno Gomes help in
-the revision.
-
+\label{sec:relatedworks}
+
+There are in the literature of computer science some
+approaches~\cite{BHDL_2003,GEMPLUS_99} to model hardware and the virtual machines
+using the B method. Then, in both works the B method has been used successfully
+to model the operational semantic. However the cost of modelling was still expensive and this
+paper quoted some techniques to lower the cost of modelling.
+% Although, the modelling of the Z80 do not have modelled some small aspects.
+
+% The objective of researchers
+
+In general, the researchers employing the B method have focused on
+more abstract level of description of software.  Considering low-level
+aspect, there has been previous work on modelling the Java Virtual
+Machine~\cite{GEMPLUS_99}.
+
+The main motivation of our research is the development of verified
+software up to the assembly level, which requires specifying the
+semantics of the underlying hardware. Thus, some aspects were not
+modelled in our work such as the execution time of the instructions.
+Also we did not consider the microarchitecture of the hardware as the scope of
+our work does not include hardware verification. However, there are
+many other specialized techniques to verify these questions.
+
+\section{Conclusions}
+\label{sec:conclusions}
+
+This work has shown an approach to the formal modelling of the instruction set of
+microcontrollers using the B method. During the construction of this model, some
+ambiguities and errors were encountered in the official reference for Z80
+microcontroller~\cite{Z80_manual}. As the B notation has a syntax that is not too
+distant from that of imperative programming languages, such model could be used
+to improve the documentation used by assembler programmers. Besides, the formal
+notation used is analyzed by software that guarantees the correctness of typing,
+the well-definedness of expressions, in addition to safety properties of the
+microcontroller state.
+
+% This research is much interesting to
+% design microcontroller. Because, the building of formal model found some erros and ambiguities at the oficial
+% manual~\cite{Z80_manual}. The Z80's B model can replace the documentation used by assembler programmers, because the
+% formal model represents the instructions effects and the B-method has a easy language, that remembers the pascal
+% language. Besides, the formal model restricts the defintions to correct typing, use expressions well-defined and
+% allow to verify properties on the model.
+
+% [Trabalhos futuros ]
+Future works comprise the development of software with the B method
+from functional specification to assembly level, using the Z80 model
+presented in this paper. The mechanic compilation from B algorithmic 
+constructs to assembly platform is also envisioned.
+
+% The next steps of the work comprises the development of a real study case in the oil area and the
+% project of a compiler to aid the automatization of verification at the assembly level
+
+% The next works are the development of a real study case in the oil area and after the project of a
+% compiler to aid the automatization of verification at the assembler level.
+
+\paragraph{Acknowledges:}
+This work was partially supported by INES (www.ines.org.br), funded by CNPq grant \ldots 
+%573964/2008-4 and by CNPq grants 553597/2008-6, 550946/2007-1, and 620132/2008-6..
+
+%\bibliographystyle{splncs}
+%\bibliography{paper}
+[To change the format; to add: thesis suggested by David, Article from ABZ 2010 suggested by David  ]
 \begin{thebibliography}{5}
-
-% Paper about BHDL
-
-% O manual do Z80
 
@@
 \bibitem {BHDL_2003}
-   Aljer, P. Devienne, S. Tison and J-L. Boulanger andG. Mariano. Bhdl: Circuit Design in B. A. In
+Aljer, P. Devienne; S. Tison  J-L. Boulanger and G. Mariano. Bhdl: Circuit Design in B. A. In
 ACSD, Third International Conference on Application of Concurrency to System Design, pages 241-242, 2003.
@@
 \bibitem {CLEARSY}
-CLEARSY. Atelier B. http://www.atelierb.eu.
-
-\bibitem {Dantas_Semish08}
-Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Proposta e avalia\c{c}\~{a}o de uma abordagem de desenvolvimento de
- software fidedigno por constru\c{c}\~{a}o com o m\'{e}todo B. In: SEMISH, Bel\'{e}m, 2008. SBC.
-
-\bibitem {Dantas_Report08}
-Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL.Verified compilation based on the B method: initial
-appraisal.Technical Report, UFRN - DIMAP, 2008.
-
-\bibitem {Bell} 
- D. Bell, I. M. E. J. P. Software Engineering: A Programming Approach. Prentice Hall, New York, 2 edition, 1992.
+ Clearsy. Atelier B. http://www.atelierb.eu.
+
+\bibitem {Dantas_SBMF08}
+Dantas, B; D\'{e}harbe, D.; Galv\~{a}o, S. L.; Moreira, A. M. and
+Medeiros Jr, V. G.. Applying the B Method to Take on the Grand Challenge of
+Verified Compilation. In: SBMF, Savaldor, 2008. SBC.
 
 \bibitem {HOARE}
-HOARE, C. A. R. The verifying compiler, a grand challenge for computing research. In: VMCAI, p. 78-78, 2005.
-
-\bibitem{Medeiros}
- Medeiros Jr, Val\'{e}rio; Galv\~{a}o, Stephenson; D\'{e}harbe, David; Modelagem de Microcontroladores em B. In: ERMAC, Natal, 2008. SBMAC.
-
-\bibitem {WORDSWORTH}
-WORDSWORTH, J. B. Software Engineering with B. Addison Wesley, Boston, 1996.
+Hoare, C. A. R. The verifying compiler, a grand challenge for computing research.
+In: VMCAI, p. 78-78, 2005.
+
+\bibitem{Z80_manual}
+Zilog. Z80 Family CPU User Manual. http://www.zilog.com/docs/z80/um0080.pdf
 

### 2012-01-08 papers/ArticleZ80/source/paper.tex

\documentclass[a4paper]{llncs}

%\documentclass[11pt]{report} % use larger type;

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\usepackage{graphicx,array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{epsfig}
\usepackage{float}
%\usepackage{supertabular} % used in B Symbols
\geometry{a4paper} % or letterpaper (US) or a5paper or....

\usepackage[english]{babel} %\usepackage{supertabular} % used in B Symbols
\usepackage{hyperref} % add URL in bibliography
\usepackage{url} % allow add URL in document

\include{comandos}
\include{SymbolsB_AtelierB}

% Allow use multiples footnote references 
\newcommand{\footnoteremember}[2]{
  \footnote{#2}
  \newcounter{#1}
  \setcounter{#1}{\value{footnote}}
}
\newcommand{\footnoterecall}[1]{
  \footnotemark[\value{#1}]
}

%
\begin{document}

\title{Experiences in Modelling a Microcontroller Instruction Set Using B}
%
%\hyphenation{se-pa-ra-te}
%
%
%
\author{Val\'{e}rio Medeiros Jr\inst{1}, David D\'{e}harbe\inst{2}}

\institute{ Federal Institute of Education, Science and Technology of Rio Grande do Norte, Natal RN 59015-000, Brazil
           \and 
		    Federal University of Rio Grande do Norte, Natal RN 59078-970, Brazil}

\maketitle

\begin{abstract}

  This paper describes an approach to model the functional aspects of the instruction set of
  microcontroller platforms and several details about the representation of
  elements from microcontrollers.  Several models were developed using the notation of the B method. 
  They are used to develop a formally verified software up to the assembly level and allow the
  simulation of models. This simulation is able to guarantee the consistency between the execution
  of a software model and a real  execution of software, it has wide variety of usages in industry
  and academia. This paper presents specifically the case of the Z80 platform and quote a theoretic 
  case study important in tanks of the petroleum industry. This work is a contribution towards the
  extension of the B method to handle developments up to the assembly level code.

%  \paragraph[Keywords]{
   \textbf{Keywords}: Embedded Software, Simulation, B method and Verification%}
\end{abstract}

%\keywords{ Validation, Verification, Microcontroller, Hardware}
%
%\setcounter{secnumdepth}{1}
%\setcounter{chapter}{1}
%\setcounter{section}{1}

\section{Introduction}

The B method~\cite{Abrial} supports the construction of safety systems
models by generating proof obligation that must be verified to guarantee its correctness. So,
an initial abstract model of the system requirements is defined and
then it is refined up to the implementation model. Development
environments based on the B method also include source code generators
for programming languages, but the result of this translation
cannot be compared by formal means. \cite{Dantas_SBMF08} proposed
an approach to extend the scope of the B method up to the
assembly level language. One key component of this approach
is to build, within the framework of the B method, formal models of
the instruction set of such assembly languages.

This work presents a formal modell of the instruction set of the Z80 
microcontroller~\cite{Z80_manual}. This microcontroller
was chosen because of several factors: it contains essential and common
concepts in microcontrollers and until microprocessors; it has extensive
documentation available; and it was widely used and is still in use. 
A first Z80 modell presented in \cite{Valerio_SBMF09} was verified completely,
but it uses constructs that are not well supported to animate. 
This modelling was changed to support animation in ProB\cite{proB}.
Basically, the new modelling\footnote{The interested reader in more
details is invited to visit our repository at: http://code.google.com/p/b2asm/.}
changed elements represented by infinite sets and adjusted the implications
of this change. Using the responsibility division mechanisms provided by B, auxiliary
libraries of basic modules were developed as part of the construction of microcontroller
model. Such library has many definitions about common concepts used in
the microcontrollers;besides the new Z80 model, a theoretic case study in 
petroleum production test system was developed using Z80 assembly language
to analyse the code simulation and verification process.

The formal model of instruction set provides many benefits. The animation of 
Z80 model allows to simulate the execution assembly programs including
support for interrupts and input and output instructions. Other possible uses
of a formal model of a microcontroller instruction set include documentation,
the construction of simulators, and can be possibly the starting point of a 
verification effort for the actual implementation of a Z80 design. Moreover, 
the model of the instruction set could be instrumented with non-functional
aspects, such as the number of cycles it takes to execute an instruction, to
prove lower and upper bounds on the execution time of a routine.
Two aspects of these formalization are of particular importance to us:
provide a more solid documentation artifact for microcontrollers, 
build a reference model for a formal compilation approach in B.

\subsection{Problems in formalizing instruction set}

In general, the manuals of microcontrollers can be more suited for
developers in assembly language. Many times, each instruction of microcontrollers
is shown in its official manual using textual description, math description, examples,
encoding, number of cycles and other information.Some times, the instruction descriptions
have an own notation, some informations are organized in different sections of document 
and the textual description are not standardized. This does not allow the interpretation by parser.
% b) The case of Z80
For example in official manual of the microcontroller Z80 \cite{Z80_manual}, if the users
want to know the action of an instruction and its effects on the flag register then he needs
to read an informal textual description. So, this description can add
some ambiguities or even mistakes. The official manual has also several problems that were 
identified over time~\cite{UndocumentedZ80}. The Z80 is a microcontroller that  has been
tested and used for many years. This intensive use and test facilitated to find several
errors in the description of the instructions in its manual. Several of these problems
have been reported in a technical report \cite{UndocumentedZ80}. Some problems are inaccurate
information, for example the description of the action of an instruction on the flag
register, partial information and distributed on different pages of manual.

% in the Z80 manual happens was that,

%A simple example of error is in some official documentation saying that a group of instructions leaves the \textit{CF} flag unaffected and the \textit{NF}
%flag is not always set, but may also be reset \cite{UndocumentedZ80}.
%% \begin{itemize} \item Some input and output instructionsThe
%% INI/INIR/IND/INDR/OUTI/OUTD/OTIR/OTDR instructions do affect the CF flag (some
%% official documentation says they leave it unaffected, important!) and the NF
%% flag is not always set but may also be reset (see 4.3 for exact operation)
%% \cite{UndocumentedZ80}. Este erro he descrito na phegina (274 do documento e
%% 294 do pdf) do manual oficial do Z80 (Z80_CPU_adotado.pdf) Citar apenas esse
%% exemplo abaixo! \item
%%when a determined interruption (NMI) is accepted, the bit IFF1 is not   
%%copied to bit IFF2~\cite{UndocumentedZ80}. 
%%
%This example can cause serious problems for programmers.
%% Este erro he descrito na pagina (22 do documento e 26do pdf) do manual
%% official do Z80 (Z80_CPU_adotado.pdf) Only IFF1 is reset.
%This error and others 7 different errors were previously identified by~\cite{UndocumentedZ80}. So to help the 
%construction of Z80 B model, other references~\cite{UndocumentedZ80,Simulator_z80}
%were also consulted.

A solution to avoid ambiguity and inconsistencies is to specify formally the
assembly instruction set, this creates a pattern of representation and validates
properties of instructions. The formal model also restricts the definitions to
correct typing and uses only well-defined expressions. Furthermore, the developer
can use with B method different levels of abstraction and add more specific details
in refined models.

%Finally, the main goal of this project, though, is to provide a basis for the
%generation of software artefacts at the assembly level, those are amenable to 
%refinement verification within the B method.

%\subparagraph{Report structure}

This work is focused on the presentation of modelling of assembly instruction
set, including elementary libraries to describe hardware aspects; and the
modelling and simulation of a case study of an embedded verified software that
can be used in tanks of the petroleum industry.

This paper is structured as it follows. %Section~\ref{sec:B_method} provides a short introduction to the B method. 
Section~\ref{sec:models} presents the elementary libraries and
the modelling of some elements common to microcontrollers.  Section~\ref{sec:z80}
presents the B model of the Z80 instruction set. %Section ~\ref{sec:Verification} provides some information on the proof effort needed to analyse the presented models.
Section ~\ref{sec:studycase} explains the building and verifying assembly-level refinements. Related works are discussed in
section~\ref{sec:relatedworks}. Finally, the last section is devoted to the conclusions.
%
%\section{Introduction to the B Method}
%\label{sec:B_method}
%
%The B method for software development~\cite{Abrial} is based on the B Abstract
%Machine Notation (AMN) for the formal specification of functional requirements and the use of formally proved refinements up to a
%specification sufficiently concrete that programming code can be automatically
%generated from it. Its mathematical basis consists of first order logic, integer
%arithmetic and set theory, and its corresponding constructs are similar to those
%of the Z notation~\cite{Spivey96}.
%
%A B specification is structured in modules. A module defines a set of valid
%states, including a set of initial states, and operations that may provoke a
%transition among states. The design process starts with a module with a
%so-called functional model of the system under development. In this initial
%modelling stage, the B method requires that the user proves that, in a machine,
%all its initial states are valid, and operations do not define
%transitions from valid states to invalid states.
%
%
%Essentially, a B module contains two main parts: a header and the available
%operations. Figure~\ref{fig:maqB} has a very basic example. The clause
%\MACHINE  has the name of module.  The next two clauses, respectively,
%reference external modules and create an instance of an external module.  The
%\VARIABLES clause declares the name of the variables that compose the
%state of the machine. Next, the \INVARIANT clause defines the valid states,
%specifying  the type and other restrictions on the variables. The \INITIALISATION specifies
%the initial states. Finally, operations correspond to the transitions among
%states of the machine.
%
%
%\begin{figure}[h]
%  % \begin{small}
%\begin{center}
%\begin{sloppypar}
%
%  $\begin{array}{lcl}
%    \begin{array}[t]{l}
%      \MACHINE \quad  \textit{micro} \\
%      \SEES \quad \textit{TYPES}, \textit{ALU} \\
%      \INCLUDES  \quad \textit{MEMORY} \\
%      \VARIABLES \quad    \textit{pc}  \\
%      \INVARIANT\\ 
%      \quad \textit{pc} \in \textit{INSTRUCTION} 
%    \end{array}
%    & \hspace*{0cm} &
%    \begin{array}[t]{l}
%      \INITIALISATION  \textit{pc} := 0 \\
%      \OPERATIONS\\
%      \textit{JMP} \textit{( jump )} = \\
%      \quad \PRE \textit{jump} \in \textit{INSTRUCTION}\\ 
%      \quad \THEN \textit{pc} := \textit{jump}\\  
%      \quad \END \\
%      \END
%    \end{array}
%  \end{array}
%  $
% \end{sloppypar} 
%\end{center}
%\caption{A very basic B machine.}
%\label{fig:maqB}
%\end{figure}

\section{Model structure and basic components}
\label{sec:models}
%Em cada subsecaoo descrever os principais lemas que foram provados.
%Sugestheo de David:
% a) structure
% b) formalization of bit-level concepts
% c) formalization of bit vectors
% d) formalization of byte and word-level concepts
% e) formalization of integer ranges

We have developed a reusable set of basic definitions to model hardware concepts and data types
concepts. The models contains definitions to represent the: registers, interruptions, input and output
ports, memory and instructions. These definitions are grouped into separated development projects
and are available as libraries.

Thus the workspace is composed of modules group and libraries: 
\begin{itemize}

\item \textbf{Power}: it has the basics definitions to help the theorems
prover about simple calculus of power, especially power of two. It contains the definition of power function
and it has constants, that represent all required results of power function.

\item \textbf{Hardware Library}: it has the definition of bit vectors with size 8 and 16, basic functions to manipulate bit vectors and important assertions. This is presented in sections \ref{subsec:HardwareLibrary1}, \ref{subsec:HardwareLibrary2} and \ref{subsec:HardwareLibrary3}.

 \item \textbf{Types Library}: it has the definition of naturals and integers represented with  8 bits and 16 bits, basic conversion functions between the types and bit vectors and important assertions. This is presented in section \ref{subsec:Types}.

 \item \textbf{Platform}: it has the definition functions of arithmetic logic unit,  memory unit, registers and assembly instructions. This is presented in section \ref{sec:z80}.

\end{itemize}

 The corresponding dependency
diagram is depicted in Figure~\ref{fig:hardware-definition-graph}; information specific to each project is
presented in the following.

\begin{figure}[h]
\centering
\includegraphics[width=.77\textwidth]{diagramaEstrutural_vertical_ProB.png}
 \caption{Dependency diagram of the Z80 model.}
\label{fig:hardware-definition-graph}
\end{figure}

\subsection{Bit representation and manipulation}
\label{subsec:HardwareLibrary1}

The entities defined in the module $\textit{BIT\_DEFINITION}$ are the
type for bits, logical operations on bits (negation, conjunction,
disjunction, exclusive disjunction), as well as a conversion function
from Boolean to bit.

First, bits are modelled as a set of integers: $\textit{BIT} =
\textit{0..1}$. The negation is an unary function on bits and it is
defined as:

$
\begin{array}{l}
\textit{bit\_not}  \in  \textit{BIT}  \fun  \textit{BIT}  \land 
\textit{bit\_not} \rm = \rm \{ \rm 0  \mapsto  \rm 1 , \rm 1  \mapsto  \rm 0 \rm \}
\end{array}
$
%\forall ( \textit{bb}). (\textit{bb} \in \textit{BIT} \implies {bit\_not}(\textit{bb}) =1-\textit{bb})\\

The module also provides lemmas on negation that may be useful for the
users of the library to develop proofs:

$
\begin{array}{l}
%  \textit{bit\_not}(0) = 1;  \textit{bit\_not}(1) = 0; \\
\forall (\textit{bb}).(\textit{bb} \in \textit{BIT} \implies \textit{bit\_not}(\textit{bit\_not}(\textit{bb})) = \textit{bb})
\end{array}
$

Conjunction is a binary function on bits and it is defined as:

$
\begin{array}{l}
\textit{bit\_and} \in \textit{BIT} \times \textit{BIT} \fun \textit{BIT} \land \\
\forall (\textit{b1}, \textit{b2}).(\textit{b1}  \in \textit{BIT}  \land \textit{b2} \in \textit{BIT} \implies \\
\quad ((\textit{bit\_and}(\textit{b1}, \textit{b2}) = 1) \iff (\textit{b1} = 1)  \land  (\textit{b2} = 1)))
\end{array}
$

The module provides the following lemmas for commutativity and associativity:

$
\begin{array}{l}
%  \textit{bit\_and}(0,0) = 0;  \textit{bit\_and}(0,1) = 0; \\
%  \textit{bit\_and}(1,0) = 0;  \textit{bit\_and}(1,1) = 1; \\
\forall (\textit{b1},\textit{b2}).(\textit{b1} \in \textit{BIT} \land \textit{b2} \in \textit{BIT} \implies \\
\quad (\textit{bit\_and}(\textit{b1}, \textit{b2}) = \textit{bit\_and}(\textit{b2},\textit{b1})))\land \\
\forall (\textit{b1},\textit{b2},\textit{b3}).(\textit{b1} \in \textit{BIT} \land  \textit{b2} \in \textit{BIT} \land \textit{b3} \in \textit{BIT} \implies \\
\quad (\textit{bit\_and}(\textit{b1}, \textit{bit\_and}(\textit{b2},\textit{b3})) = \textit{bit\_and}(\textit{bit\_and}(\textit{b1},\textit{b2}),\textit{b3})))\\
% \forall (\textit{b1}).(\textit{b1} \in \textit{BIT} \implies (\textit{bit\_and}(\textit{b1}, 1) = \textit{b1})); \\
% \forall (\textit{b1}).(\textit{b1} \in \textit{BIT} \implies (\textit{bit\_and}(\textit{b1}, 0) = 0));
\end{array}
$

The module provides definitions of $\textit{bit\_or}$ (disjunction)
and $\textit{bit\_xor}$ (exclusive disjunction), as well as lemmas on
those operators. These are standard and their expression in B is
similar as for $\textit{bit\_and}$, they are thus omitted from this paper.

Finally, the conversion from Boolean to bit is simply defined as:

$
\begin{array}{l}
\textit{bool\_to\_bit} \in \BOOL \fun \textit{BIT} \land \textit{bool\_to\_bit} = \{ \TRUE \mapsto 1, \FALSE \mapsto 0 \} \\
\end{array}
$

Observe that all the lemmas that are provided in this module have been
mechanically proved by the theorem prover included with our B
development environment. None of these proofs requires human insight.

\subsection{Representation and manipulation of bit vectors}
\label{subsec:HardwareLibrary2}

Sequences are pre-defined in B, as functions whose the domain is an
integer range with lower bound 1 (one). Indices in bit vectors usually
range from 0 (zero) upwards and the model we propose obeys this
convention by making an one-position shift where necessary. This shift
is important to use the predefined functions of sequences. We thus
define bit vectors as non-empty sequences of bits, and
$\textit{BIT\_VECTOR}$ is the set of all such sequences:
$\textit{BIT\_VECTOR} = \seq1(\textit{BIT})$. The $\textit{BIT\_VECTOR}$ is an infinite set and it hinders the animation in ProB. 
To avoid this obstruction, the specialized types (\textit{BYTE} and \textit{BV16}) cannot reference directly \textit{BIT\_VECTOR}.

% CAN COMMENT TO REDUCE SPACE
%The function $\textit{bv\_size}$ returns the size of a given bit vector. It is basically a wrapper for the
%predefined function $\textbf{size}$ that applies to sequences.
%
%$
%\begin{array}{l}
%\textit{bv\_size} \in \textit{BIT\_VECTOR} \fun \nat_1 \land \\
%\textit{bv\_size} = \lambda bv . (bv \in \textit{BIT\_VECTOR} \mid \textbf{size}(bv))
%\end{array}
%$

We also define two functions $\textit{bv\_set}$ and $\textit{bv\_clear}$ that, given a bit vector, and a
position of the bit vector, return the bit vector resulting from setting the corresponding position to 0
or to 1, and a function $\textit{bv\_get}$ that, given a bit vector, and a valid position, each one
returns the value of the bit at that position. Only the first definition is shown here:

$
\begin{array}{l}
\textit{bv\_set} \in \textit{BIT\_VECTOR} \times \nat \fun \textit{BIT\_VECTOR} \land \textit{bv\_set} =\\
\lambda v, n . (v \in \textit{BIT\_VECTOR} \land n \in \nat \land n <\textit{size}(v)
\mid v \lover \{ n+1 \mapsto 1 \})
\end{array}
$

The function $bv\_catenate$ takes as parameters two bit vectors $v$ and $w$, and returns the result of the
concatenation of $v$ and $w$, such that $v$ constitutes the most significant part of the result.

% $
% \begin{array}{l}
% \textit{bv\_catenate} \in \textit{BIT\_VECTOR} \times \textit{BIT\_VECTOR} \fun \textit{BIT\_VECTOR} \land \\
% \textit{bv\_catenate} = \lambda v, w \bullet (v \in \textit{ BIT\_VECTOR} \land w \in \textit{ BIT\_VECTOR}  \mid v 
% %\conc
%   w)
% \end{array}
% $

\hspace*{0.00in} \it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$ \it
BIT\_VECTOR $\land$\\
\hspace*{0.21in} \it bv\_catenate \rm =  $\lambda$  v\rm ,\it w \rm . \rm (\it v 
$\in$  \it BIT\_VECTOR $\land$ \it w $\in$  \it BIT\_VECTOR  $\mid$  \it v\^ \it w\rm ) %$\cat$

%COMMENT TO REDUCE SPACE
%We also define a function $\textit{bv\_zero}$ that, given a positive
%integer $n$, return a bit vector of size $n$, with all bits set to 0.
%A similar function, called $\textit{bv\_one}$, with all bits set to 1
%is also defined but not presented here.
%
%$
%\begin{array}{l}
%\textit{bv\_zero} \in \nat_1 \fun \textit{BIT\_VECTOR} \land \\
%\textit{bv\_zero} = \lambda n . (n \in \nat_1 \mid 1..n \times \{0\}) 
%\end{array}
%$
%

Additionally, the module provides definitions for the classical
logical combinations of bit vectors: $\textit{bit\_not}$,
$\textit{bit\_and}$, $\textit{bit\_or}$ and $\textit{bit\_xor}$. Only
the first two are presented here. Observe that the domain of the
binary operators is restricted to pairs of bit vectors of the same
length:

$
\begin{array}{l}
\textit{bv\_not} \in \textit{BIT\_VECTOR} \fun \textit{BIT\_VECTOR} \land \\
\textit{bv\_not} = \lambda v . (v \in \textit{BIT\_VECTOR} \mid \quad \lambda i . (1 .. \textit{size}(v)) \mid \textit{bit\_not}(v(i))) \land \\
\textit{bv\_and} \in \textit{BIT\_VECTOR} \times \textit{BIT\_VECTOR} \fun \textit{BIT\_VECTOR} \land \\
\textit{bv\_and} = \lambda v_1, v_2 . (v_1 \in \textit{BIT\_VECTOR} \land v_2 \in \textit{BIT\_VECTOR} \land \\
\textit{size}(v_1) = \textit{size}(v_2) \mid \lambda i . (1 .. \textit{size}(v_1)) \mid
\textit{bit\_and}(v_1(i), v_2(i)))
\end{array}
$

We provide several lemmas on bit vector operations. These lemmas
express properties on the size of the result of the operations
as well as classical algebraic properties such as associativity
and commutativity.

\subsection{Modelling bytes and bit vectors of length 16}
\label{subsec:HardwareLibrary3}

Bit vectors of length 8 are bytes. They are a common entity in
hardware design. We provide the following definitions:

\hspace*{0.0in}\it BYTE\_WIDTH \rm = 8 $\land$ \it BYTE\_INDEX \rm = 1 $\upto$ \rm  BYTE\_WIDTH\rm  \hspace*{0.05in} $\land$

\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.10in} $\land$

\hspace*{0.0in}\it BYTE \rm =\rm (\it BYTE\_INDEX  $\fun$  \it BIT\rm )\hspace*{0.10in} $\land$  \bf card\rm (\it BYTE\rm )\rm = $2^{8}$  $\land$

\hspace*{0.0in}$\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BYTE  $\implies$  \bf size\rm (\it b1\rm )\rm =\it BYTE\_WIDTH  $\land$  \it b1  $\in$  \bf seq1\rm (\it BIT\rm )\rm )

%\hspace*{0.0in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \} $\land$

%\hspace*{0.0in}\it BYTE\_ZERO  $\in$  \it BYTE  $\land$ \it BYTE\_ZERO \rm = \it BYTE\_INDEX  $\times$  \rm \{\rm 0\rm \}

The $\textit{BYTE\_INDEX}$ is the domain of the functions modelling bytes. It starts at 1 to obey a
definition of sequences from B. However, it is common in hardware architectures to start indexing from
zero. The definition $\textit{PHYS\_BYTE\_INDEX}$ is used to provide functionalities obeying this
convention. The $\textit{BYTE}$ type is as a specialized type of $\textit{BIT\_VECTOR}$, but it has a size
limit. Other specific definitions are provided to facilitate further modelling: the type $\textit{BV16}$
is created for bit vector of length 16 in a similar way.

\subsection{Bit vector arithmetic}
\label{subsec:Types}

Bit vectors are used to represent and combine numbers: integer ranges (signed or unsigned). Therefore, our
library includes functions to manipulate such data, for example, the function $\textit{bv\_to\_nat}$ that
maps bit vectors to natural numbers:

$
\begin{array}{l}
\textit{bv\_to\_nat} \in \textit{BIT\_VECTOR} \fun \nat  \land \\
\textit{bv\_to\_nat} = \lambda v .  (v \in \textit{BIT\_VECTOR} \mid  \sum i . (i \in \dom(v) . v(i)
\times 2^{i-1}))
\end{array}
$

An  associated lemma is: $\forall n . (n \in \nat_1 \implies \textit{bv\_to\_nat}(\textit{nat\_to\_bv}(n)) = n)$

\subsection{Basics data types}

The instruction set architecture usually have common bit vector type and integer type.
In table~\ref{tab:types}, the first three are placed in the hardware library and the last four types are placed in the integer types library. 
Each type module has functions to manipulate and convert its data. There are seven common basic data types represented by modules.

\begin{table}
\caption{Details of basic data types}
\label{tab:types}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
 $Type\ Name$ & $Range$ & $Physical\ Size $  & $Associated\ Proof\ Obligations$ \\\hline
 BIT & 0..1 & 1 bit &   118 \\\hline %49 + 69 = 118 \\\hline
 BYTE & -- & 1 bytes &  153\\\hline%17 + 136 = 153\\\hline
 BV16 & -- & 2 bytes &  75 \\ \hline%5 + 70 = 75 \\ \hline
 UCHAR & 0..255 & 1 byte &  30 \\\hline %4 + 26 = 30 \\\hline
 SCHAR & -128..127 & 1 byte & 26 \\\hline %2 + 24 = 26 \\\hline
 USHORT & 0..65.535 & 2 byte & 62 \\\hline % 4 + 58 = 62 \\\hline
 SSHORT & -32.768..32.767 & 2 byte & 58\\\hline %  2 + 56 = 58\\\hline
%  &  & Total & 522\\\hline %  2 + 56 = 58\\\hline
\end{tabular}
\end{center}
\end{table}

Usually, each type module just needs to instantiate concepts that were already defined in the hardware
modelling library.  For example, the function $\textit{bv\_to\_nat}$ from bit vector arithmetic is
specialized to $\textit{byte\_uchar}$. As the set $\textit{BYTE}$ is a subset of the
$\textit{BIT\_VECTOR}$, this function can defined as follows:

$
\begin{array}{l}
\textit{byte\_uchar} \in \textit{BYTE} \fun \nat \land \\
\textit{byte\_uchar} = \lambda (v) . ( v \in BYTE | bv\_to\_nat(v) )
\end{array}
$

The definitions of the library types reuse the basic definitions from the hardware library. This provides
greater confidence and facilitates the proof process, because the prover can reuse the previously defined
lemma.

%The inverse function $\textit{uchar\_byte}$ is easily defined:
%
%$
%\begin{array}{l}
%\textit{uchar\_byte} \in \textit{UCHAR}  \fun  \textit{BYTE}  \land \\
%   \textit{uchar\_byte} = \ (\textit{byte\_uchar}) ^{-1}
%\end{array}
%$

 We also created the following lemmas:

 $
 \begin{array}{l}
  \forall (val) . (val \in \textit{UCHAR} |
  \textit{byte\_uchar}(\textit{uchar\_byte}(val)) = val) \land\\
  \forall (by) . (by \in \textit{BYTE} |
  \textit{uchar\_byte}(\textit{byte\_uchar}(by)) = by)
 \end{array}
 $

Similarly, several other general functions and lemmas were created for all other data types.
%[NEW PARAGRAPH-TO-REVIEW]
%Similarly, several other general functions and lemmas were created for all other data types.
%Some times, we need create arithmetic and logic functions that are more specific for a
%determined computer architecture. These functions are developed in the module ALU 
%(arithmetic  logic unit). The following shows some examples of specific functions used in
%microcontroller model of Z80. For example, the function $\textit{instruction_next}$
%receives a $\textit{BYTE}$ and returns a $\textit{BYTE}$ that contains a sequence of bit:
%a signal bit,

Nearly 50\% of all these proof obligations shown in the table~\ref{tab:types}
are verified almost automatically. Although many others proof obligations are difficult to verify, because these
check the type of functions with non-linear arithmetic is a difficult task.

\subsection{Verification Process in Hardware Library and Types Library} 
\label{sec:VerificationHardwareLibrary}

\input{VerificationLibraryHardware}

\section{A B model of the Z80 instruction set}
\label{sec:z80}

The \textit{Z80} is a CISC microcontroller developed by
\textit{Zilog}~\cite{Z80_manual}. The Z80 is composed by different elements and a
simplified internal organization is shown in the Figure \ref{fig:DiagramBlock}.
This figure has some important elements of Z80 CPU: ALU, registers of 8 and 16
bits and input/output ports.

\begin{figure}[h] \centering
\includegraphics[width=0.70\textwidth]{images/Architecture.png}
\caption{Simplified internal organization of Z80 CPU.}
\label{fig:DiagramBlock}
\end{figure}

The Z80 has 158 different instructions, including all the 78 from Intel 8080
microprocessor, and all of them were formally specified in B. These instructions are classified
into these categories: load and exchange; block transfer and search; arithmetic
and logical; rotate and shift; bit manipulation; jump, call and return;
input/output; and basic cpu control. Each category of instruction has different
elements of specification.

This section shows the elements that make up the different types of instructions.
The main elements are specified in the microcontroller module Z80 and parts of it
are presented below.  Basically, the state of microcontroller is formed by data 
from program counter, ports, registers, etc. The transitions between the states 
are actioned by execution of instruction or an external action. Groups of registers
are represented by the variables of the specification states that appear in clause
\textit{VARIABLES}. The declaration of valid states of variables is
in the \textit{INVARIANT} clause and the initial state is defined in the
\textit{INITIALISATION} clause. The assembly instructions are defined by the
clause \textit{OPERATIONS} and, in general, each instruction has three actions: update 
the program counter, update the flag register and its main effect.
%Several functions were created to Z80.
General functions that can be used in others microcontroller models are defined in
the modules of data types. Specific functions of the microcontroller are defined in the ALU (arithmetic logic unit) module.

The main module includes an instance of the memory module and accesses the definitions from basic data
types modules and the \textit{ALU} module.

\begin{sloppypar}

\bf MACHINE

\hspace*{0.15in}\it Z80

\bf INCLUDES

\hspace*{0.10in}\it MEMORY

\bf SEES

\hspace*{0.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,

\hspace*{0.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,

\hspace*{0.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,

\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION
\end{sloppypar}

\subsection{Modelling registers}

The Z80 CPU includes alternative set of accumulator, flag and general registers. The CPU contains a stack
pointer ($\textit{sp}$), program counter ($\textit{pc}$), two index registers ($\textit{ix}$ and $\textit{iy}$), an
interrupt register ($\textit{i\_}$), a refresh register ($\textit{r\_}$), two bits ($\textit{iff1}$,
$\textit{iff2}$) used to control the interruptions, a pair of bits to define the interruption mode ($\textit{im}$)
and the input and output ports ($\textit{i\_o\_ports}$). Below, its definitions are represented by
\textit{INVARIANT}.

\begin{sloppypar}
\bf INVARIANT

\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$

\hspace*{0.10in}\it pc  $\in$  \it INSTRUCTION  $\land$  \it sp  $\in$  \it BV16  $\land$  \it ix  $\in$  \it BV16  $\land$  \it iy  $\in$  \it BV16  $\land$

\hspace*{0.10in}\it i\_  $\in$  \it BYTE  $\land$  \it r\_ $\in$  \it BYTE  $\land$

\hspace*{0.10in}\it iff1  $\in$  \it BIT  $\land$ \hspace*{0.10in}\it iff2  $\in$  \it BIT  $\land$

\hspace*{0.10in}\it im $\in$ (\it BIT $\times$ \it BIT\rm )  $\land$

\hspace*{0.10in}\it i\_o\_ports  $\in$  \it BYTE  $\fun$  \it BYTE
\end{sloppypar}

%

% Features of microcontroller
  The internal registers contain 176 bits of read/write memory that are represented by
identifiers used as parameters in the instructions . It includes two sets of six general purpose
registers which may be used individually as 8-bit registers or as 16-bit register pairs.  The working
registers are represented by variable $\textit{rgs8}$. The domain of $\textit{rgs8}$
($\textit{id\_regs8}$) is a set formed by identifiers of registers of 8 bits. These registers are called
main register set (\textit{a0,b0,c0,d0,e0,f0,h0,l0}) and alternate register set
(\textit{a\_0,b\_0,c\_0,d\_0,e\_0,f\_0,h\_0,l\_0}). These registers can be accessed in pairs, forming
16-bits, resulting in another set of identifiers of 16-bits registers, named $\textit{id\_reg16}$.

\begin{sloppypar}

	\bf SETS\\
	\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0 \rm , \it f0 \rm , \it f\_0 \rm , \it a\_0 \rm ,\\
	\hspace*{0.93in}\it b0 \rm , \it c0 \rm , \it b\_0 \rm , \it c\_0\rm,\\
	\hspace*{0.93in}\it d0 \rm , \it e0 \rm , \it d\_0 \rm , \it e\_0\rm,\\
	\hspace*{0.93in}\it h0 \rm , \it l0 \rm , \it h\_0 \rm , \it l\_0\};\\
	\hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{ \it BC \rm , \it DE \rm , \it HL \rm , \it SP \rm , \it AF \rm \}

\end{sloppypar}

The main working register of Z80 is the accumulator ($\textit{rgs8(a0)}$) used for arithmetic/logic,
input/output and loading/storing operations.

\subsubsection{Flag register:}
another important element is the ``f'' register ($\textit{rgs8(f0)}$),  that is
used as a flag register. This register uses only six bits to represent the execution result status of each instruction.
According to the official manual the bits 3 and 5 are not used and the others bits have the follow meaning: $\textit{bv\_get(rgs8(f0),0)}$ - the carry bit; $\textit{bv\_get(rgs8(f0),1)}$ - the add/subtract bit; $\textit{bv\_get(rgs8(f0),2)}$ - the parity or overflow bit; $\textit{bv\_get(rgs8(f0),4)}$ - the half carry bit; 
$\textit{bv\_get(rgs8(f0),6)}$ - the zero bit; and $\textit{bv\_get(rgs8(f0),7)}$ - the sign bit.
%\begin{description}
%  \item[$\textit{bv\_get(rgs8(f0),0)}$] - The Carry bit
%  \item[$\textit{bv\_get(rgs8(f0),1)}$] - The Add/Subtract bit
%  \item[$\textit{bv\_get(rgs8(f0),2)}$] - The Parity or Overflow bit
%  \item[$\textit{bv\_get(rgs8(f0),4)}$] - The Half Carry bit
%  \item[$\textit{bv\_get(rgs8(f0),6)}$] - The Zero bit
%  \item[$\textit{bv\_get(rgs8(f0),7)}$] - The Sign bit 
%\end{description}
These bits also can be used to specify security properties for microcontroller programs.

\textbf{Assuring the absence of overflow:}
 \emph{To assure that an overflow does not happen, the developer can add this
 expression (\textit{bv\_get(rgs8(f0),0)} $\neq$ 1 $\land$
 \textit{bv\_get(rgs8(f0),2) $\neq$ 1}) in the invariant.
% When a overflow happen,it can be dangerous. So the developer may restrict its use. However, this restriction can also
% become more difficult to verify the model.
}

\subsection{Manipulation data functions from Z80}

 There are some specific functions from Z80 to manipulate the data. In addressing mode, the function
$\textit{bv\_ireg\_plus\_d}$ receives the value of a register ($\textit{ix}$ or $\textit{iy}$) 
used as a base to point to a region in memory from which data is to be stored or retrieved.
%and the displacement to return the sum, the result is the dislocated address memory, 
 An additional byte is included in indexed instructions to specify an offset from this base. This displacement is specified as a two's complement signed integer. This function is defined as:

\hspace*{0.0in}\it bv\_ireg\_plus\_d \rm : \rm(\it BV16  $\times$  \it SCHAR  $\fun$  \it BV16\rm )  $\land$

\hspace*{0.0in}\it bv\_ireg\_plus\_d \rm =  $\lambda$  \rm ( \it ix\_iy \rm , \it offset \rm ) \rm . \rm ( \it ix\_iy  $\in$  \it BV16  $\land$  \it offset  $\in$  \it SCHAR

\hspace*{0.20in}$\mid$ \it ushort\_bv16 \rm ( \rm (\it bv16\_ushort \rm ( \it ix\_iy \rm ) \rm + \it offset \rm ) 
\textbf{\textit{mod}} $2^{16}$ ) \rm )

Another derived function is  $\textit{bv\_(ireg\_plus\_d)}$, this returns the value in the memory  address
returned by $\textit{bv\_ireg\_plus\_d}$ function and its definition is similar.

There is a specific function to refresh the flag register, it is named $\textit{update\_reg\_flag}$. It is typed and defined as follow:\it update\_flag\_reg \rm $\in$ \rm (\it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it
BIT $\times$  \it BIT  $\times$  \it BIT\rm) $\fun$  \rm (\rm \{\it f0\rm \}  $\times$  \it BYTE\rm ).

\it update\_flag\_reg \rm =  $\lambda$  \rm (\it s7 \rm, \it z6 \rm,\it h4 \rm,\it pv2 \rm ,\it n1
\rm ,\it c0 \rm) . 
\rm ( \it s7 $\in$ \it BIT $\land$ \it z6 $\in$ \it BIT $\land$ \it h4 $\in$ \it BIT $\land$ \it pv2 $\in$ \it BIT $\land$ \it n1 $\in$ \it BIT $\land$ \it c0 $\in$ \it BIT \hspace*{0.20in}\rm $\mid$( \it f0  $\mapsto$  \rm \rm [\it c0\rm , \it n1\rm , \it pv2\rm , \rm 1\rm , \it h4\rm , \rm 1\rm , \it z6\rm , \it s7\rm \rm ]\rm ) \rm )

\subsection{Program, stack and data memory}

The Z80 uses a unique memory for storing program instructions, data stack and
general-purpose data. The memory has 16-bit addresses and each address holds a byte.
Thus, the data from the memory module is very simple:

\begin{sloppypar}
\bf INVARIANT \\
\hspace*{0.10in}\it mem  $\in$  \it BV16  $\fun$  \it BYTE

\end{sloppypar}

In general, the instructions can access all memory address, but it is dangerous. The user may mistakenly
access and change data in memory. For added  security, it is important that the program instructions has
limited access by region. Thus the designer can specify address regions to restrict the access from
instructions. The address regions can be specified using constants
($\textit{PROGRAM\_R\_ADR,DATA\_R\_ADR,STACK\_R\_ADR}$), these define respectively a range of restricted
address for: programs instructions, general purpose data and data stack.

$
\begin{array}{l}
\textit{PROGRAM\_R\_ADR} = 0..16384 \land\\
\textit{DATA\_R\_ADR} = 16385..49151 \land\\
\textit{STACK\_R\_ADR} = 49152..65535
\end{array}
$

\textbf{Assuring the absence of overlapping of address regions:}
 \emph{To assure that address regions are well defined, then the designer must to verify the expression:}
\begin{sloppypar}
\hspace*{0.10in}\it PROGRAM\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$  STACK\_R\_ADR $=$ \{\}
\end{sloppypar}

\textbf{Preserving the consistency of the memory:} \emph{In general, the access to
some address regions is dangerous. Then, each instruction has a specific pre-condition that verify
if the new address memory, that will be updated, is member of its region. For example, the $\textit{PUSH}$
program instruction allows write only in the region of stack ($\textit{STACK\_R\_ADR}$).}

\subsection{Arithmetic logic unit}

There are many functions in the module \textit{ALU}. In general, the definition of these 
functions use basic definitions or previously defined functions. For example, the function
$\textit{half8UCHAR}$ is used to get the half part of $\textit{UCHAR}$ value.
It is important to know the half carry and it is used in the function $\textit{add8UCHAR}$.

\hspace*{0.0in}

\hspace*{0.0in}\it half8UCHAR  $\in$  \it UCHAR  $\fun$  \it UCHAR  $\land$

\hspace*{0.0in}\it half8UCHAR \rm =  $\lambda$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it UCHAR  $\mid$  \it ww  $\mod$  \it $2^{4}$\rm )

\hspace*{0.0in}

The function $\textit{add8UCHAR}$ receives a bit carry and two $\textit{UCHAR}$ values and returns respectively the 
sum, the sign bit, the carry bit, the half carry bit and the zero bit. It is typed as follows: \it add8UCHAR \rm :
\rm (\it BIT $\times$ \it UCHAR $\times$ \it UCHAR\rm ) $\fun$ \rm (\it UCHAR $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm ) and its definition is:

\hspace*{0.0in}\it add8UCHAR\rm = $\lambda$ \rm(\it carry\rm,\it w1\rm, \it w2\rm)\rm.\rm

\hspace*{0.0in}(\it carry $\in$  \it BIT  $\land$  \it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it UCHAR
$\mid$

\hspace*{0.40in}\rm(\rm(\rm(\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm ),

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it uchar\_schar\rm (\it w1\rm ) \rm + \it uchar\_schar \rm (\it
w2\rm ) $<$ \rm 0\rm ),

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it w1 \rm + \it w2 $>$ \it UCHAR\_MAX\rm )\rm ,

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm ) \rm + \it half8UCHAR\rm ( \it
w2\rm )  $\geq$  \it $2^{4}$\rm )\rm,

\hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm
)\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )

\hspace*{0.20in}

A related function to subtract operation is $\textit{subtract8UCHAR}$. There are the same functions for the
$\textit{SCHAR}$ type, they are respectively $\textit{add8SCHAR}$ and $\textit{subtract8SCHAR}$, all these
functions are of 8 bits ($\textit{BYTE}$) and defined similarly. In the same way, the arithmetic functions for 16
bits ($\textit{BV16}$) are defined.
The module ALU has several others functions, for example:%but only some simplest functions also are explained below:
%\begin{itemize}
  %\item \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and
  %returns its increment. There is a similar function named $\textit{dec}$.
%  \item 
\it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the  
  actual value from program counter register (\textit{pc}) and returns its increment;
% \item 
and \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns the most significant bit,
 in other words, the signal bit.
  % \item \it is\_zero  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
 % byte is zero, otherwise returns 0.
  %\item \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.
%\end{itemize}
As the logic functions that are defined in the \textit{BYTE} and \textit{BV16}
module are included in the \textit{ALU} module, they can be seen and used
directly in the \textit{ALU} and \textit{Z80} modules.

\subsection{Modelling the actions and instructions}

Each instruction is represented by a B operation in the module Z80. The main
module (\textit{Z80}) has three categories of operations: a category represents the 
instructions of microcontrollers, a second category represents the input and output of data 
(shown in ~\ref{sec:modellingIO}) and the last category represents the external
actions (shown in~\ref{sec:externalactions}). A simple example of instruction category is
$\textit{LD\_(nn)\_A}$\footnoteremember{myfootnote}{The tools B does not allow
to use parentheses in identifiers, and the characters \{``('',``)''\} are
replaced respectively by \{``9'',``0''\} in the actual specification.} shown below. The
pre-defined functions are necessary many times to model the instructions, these
functions facilitate the construction of instruction set model. By default, all
parameters in operations are either predefined elements in the model or
integers values in decimal representation. This instruction use the
$\textit{updateAddressMem}$ operation from \textit{Memory} module and it receives
a address memory and its new memory value. It also increments the program
counter ($\textit{pc}$) and update the refresh register ($\textit{r\_}$).
The other instructions have a similar structure.

\hspace*{0.00in}\bf LD\_9nn0\_A \rm ( \it nn \rm ) \rm =

\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} $\land$ \hspace*{0.10in}\it nn\hspace*{0.10in} $\in$  \it DATA\_R\_ADR

\hspace*{0.20in}\bf THEN

\hspace*{0.20in}\bf updateAddressMem \rm ( \it ushort\_bv16 \rm ( \it nn \rm ) \rm , \it rgs8 \rm ( \it a0 \rm )
\rm )  $\para$

\hspace*{0.20in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.00in}\bf END\rm

\subsection{Modelling the input/output instructions}
\label{sec:modellingIO}
The Z80 has an extensive set of input and output (I/O) instructions and 256 ports for
devices. This model can transfer data blocks between the I/O devices and any
of the internal registers or memory address.

The $\textit{IN\_r(C)}$\footnoterecall{myfootnote} instruction is represented
by the following B operation. It receives an register identifier ``$\textit{rr}$'' and, it stores the value from ``rr'' in the $\textit{C}$ port address. Besides, it increments the program counter and updates the flag registers.

\hspace*{0.0in}\bf IN\_r\_9C0 \rm ( \it rr \rm ) \rm =

\hspace*{0.0in}\bf PRE \it rr  $\in$  \it id\_reg\_8  $\land$  \it rr  $\not =$  \it f0\hspace*{0.15in}\bf THEN

\hspace*{0.20in}\bf ANY

\hspace*{0.40in}\it negative \rm , \it zero \rm , \it half\_carry \rm , \it pv \rm , \it add\_sub \rm , \it carry

\hspace*{0.20in}\bf WHERE

\hspace*{0.40in}\it negative $\in$ \it BIT $\land$ \it zero $\in$ \it BIT $\land$ \it half\_carry $\in$ \it BIT 
$\land$ \it pv $\in$ \it BIT $\land$

 \hspace*{0.40in}\it add\_sub $\in$ \it BIT $\land$ \it carry $\in$ \it BIT  $\land$

\hspace*{0.40in}\it negative \rm = \it is\_negative \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$

\hspace*{0.40in}\it zero \rm = \it is\_zero \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$

\hspace*{0.40in}\it half\_carry \rm = \rm 0  $\land$

\hspace*{0.40in}\it pv \rm = \it parity\_even \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) $\land$

\hspace*{0.40in}\it add\_sub \rm =\hspace*{0.10in}\rm 0  $\land$

\hspace*{0.40in}\it carry \rm = \it z\_c

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm ( \it rr  $\mapsto$  \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) \rm ,

\hspace*{0.40in}\it update\_flag\_reg\rm (\it negative\rm,\it zero\rm,\it half\_carry\rm,\it pv\rm,\it add\_sub\rm,\it carry)\rm\}$\para$

\hspace*{0.40in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.0in}\bf END \rm

%The main I/O instructions work similarly, for example: $\textit{OUT(n),A}$  or $\textit{OUT (C ), r}$.

% $
% \begin{array}{l} 
% \textit{IN\_r\_9C0} ( rr ) = \\
% \quad   \PRE rr \in id\_reg\_8    \THEN\\
% \quad\quad \ANY data\_in, negative , zero , half\_carry , pv , add\_sub , carry\\ 
% \quad\quad \WHERE data\_in \in \textit{BYTE} \land negative \in \textit{BIT}\land\\
% \quad\quad\quad carry \in \textit{BIT} \land half\_carry \in \textit{BIT} \land zero \in \textit{BIT} \land \\
% \quad\quad\quad negative = is\_negative (data\_in) \land zero = is\_zero(data\_in ) \land\\
% \quad\quad\quad half\_carry = 0 \land pv =parity\_even\_BYTE ( data\_in )    \land\\
% \quad\quad\quad add\_sub =  0 \land carry = z\_c \\
% \quad\quad  \THEN \\
% \quad\quad\quad i\_o\_ports ( rgs8 ( c0 ) ) := data\_in ||\\
% \quad\quad\quad rgs8 := rgs8 <+ \{ ( rr |-> data\_in ) ,\\
% \quad\quad\quad get\_new\_flag\_register\_SZ\_H\_PvNC ( rgs8 , negative , zero, half\_carry , pv , add\_sub , carry ) \} ||\\
% \quad\quad\quad pc := instruction\_next( pc )\\
% \quad\quad \END\\
% \quad \END\\
% \end{array}
% $

\subsection{Modelling the external actions}
\label{sec:externalactions}

The external actions change the state of the microcontroller, for example, refreshing the I/O ports
and interruptions request. The external actions are also modelled by operations and they are named with the
prefix ``$ext\_$'' and followed by the name of action. There are four external
actions: $ext\_update\_io\_ports$, $ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The
$ext\_update\_io\_ports$ just change the state of I/O port, see.

\hspace*{0.20in}\bf ext\_update\_io\_ports\rm (\it address\rm ,\it value\rm )\rm =

\hspace*{0.20in}\bf PRE \it address  $\in$  \it UCHAR  $\land$ \hspace*{0.10in}\it value  $\in$  \it SCHAR \bf THEN

\hspace*{0.40in}\it io\_ports \rm ( \it uchar\_byte \rm ( \it address \rm ) \rm ) \rm := \it schar\_byte \rm ( \it
value \rm )

\hspace*{0.20in}\bf END\rm

The other external actions are related to interruptions. The interruptions allow
that the devices suspend a routine from CPU and start another service routine.
This service routine can exchange data or signals between CPU and external
devices. When a routine is finished, then the CPU comes back to the last routine
that was interrupted.

For the interrupts, the following elements are important:  the interrupt flip-flops
($\textit{iff1}$ and $\textit{iff2}$), the types of interrupts (maskable and
non-maskable), the interrupt mode (set with the $\textit{IM 0}$, $\textit{IM 1}$,
$\textit{IM 2}$ instructions) and the $\textit{i\_}$ register.

Flip-flops $\textit{iff1}$ and $\textit{iff2}$ control the maskable interrupts
($\textit{INT}$). When $\textit{iff1}$ is set, the interrupt is enabled,
otherwise it is disabled; $\textit{iff2}$ is used only as backup for $\textit{iff1}$. The
instructions $\textit{EI}$ and $\textit{DI}$ respectively enable and disable the
maskable interruptions, setting  $\textit{iff1}$ to 1 and 0.

The interruptions and the \textit{reset} action can change the state of
program counter. Theses actions are modelled by B operations and its main
effects are shown below\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$ holds the value of stack pointer minus 2,
%=$\textit{dec\_BV16(dec\_BV16(sp))}$
 $\textit{sp\_minus\_one}$ is the value of stack pointer minus 1,
%=$\textit{dec\_BV16(sp)}$
$\textit{pc\_high}$ holds the most significant 8 bits and
$\textit{pc\_low}$ holds the least significant 8 bits.}.

 \textbf{NMI} - Non-maskable interrupts cannot be disabled
 by the programmer. Then, when a device makes a request, $sp$ is pushed,
 $pc$ receives $66H$ (102 in decimal), $\textit{iff1}$ is reset , $\textit{iff2}$ stores
 $\textit{iff1}$ and the refresh register is updated.

\begin{sloppypar}
\bf updateStack\rm (\rm \{ \rm (\it sp\_minus\_two  $\mapsto$  \it pc\_low\rm )\rm ,\rm (\it sp\_minus\_one  $\mapsto$ \it pc\_high \rm ) \rm \}\rm )$\para$

\it sp \rm := \it sp\_minus\_two  $\para$ \it pc \rm := \rm 1\rm 0\rm 2 $\para$ \it iff1\rm :=\rm 0  $\para$  \it iff2\rm := \it iff1 $\para$

\it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\\
\end{sloppypar}

  \textbf{INT} - Maskable Interrupt is usually reserved for important functions that can be enabled and
  disabled by the programmer. When a maskable interrupt action happens, both $\textit{iff1}$ and $\textit{iff2}$ are
  cleared, disabling the interrupts, $sp$ is pushed, the refresh register is updated and the other effects
  depend on the interrupt mode register (\it im\rm).

 \begin{itemize}

  \item The mode 0 is compatible with 8080 and this mode is selected when \it im \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a 
  non-maskable interrupt happens, the CPU fetches an instruction of one byte
  from an external device, usually an RST instruction, and the CPU executes it.
  The instruction code is received from an external device by data bus and it is represented by integer parameter called $\textit{byte\_bus}$ .

  \item The mode 1 is the simplest and this mode is selected when \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 1
  \rm ). Simply, when a non-maskable interruption happens, the program counter
  receives $38H$ (56 in decimal).

  \item The mode 2 is the most flexible and this mode is selected when \it im \rm = \rm ( \rm 1  $\mapsto$
  \rm 1 \rm ). When a non-maskable interruption happens, an indirect call can be
  made to any address memory. The program counter receives a bit vector of size
  16 composed with two parts: the most significant part of the $\textit{i\_}$
  register and the least significant part of the $\textit{byte\_bus}$ with the
  last bit cleared.

 \end{itemize}

The essential part of maskable interrupt is shown below, where $\textit{byte\_bus}$ is a parameter of the $\textit{INT}$ operation:

\begin{sloppypar}

\hspace*{0.00in}\bf IF \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 0 \rm ) \bf THEN

\hspace*{0.10in}\bf IF\hspace*{0.10in}\it byte\_bus \rm $\in$  \it opcodes\_RST\_instruction

\hspace*{0.10in}\bf THEN\hspace*{1.05in}

\hspace*{0.40in}\it pc \rm := \it byte\_bus \rm - \rm 1\rm 9\rm
9\hspace*{0.10in} $\para$

\hspace*{0.40in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.40in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.40in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.10in}\bf ELSIF \it byte\_bus \rm = \it opcode\_\ldots\_instruction

\hspace*{0.40in}\bf \ldots

\hspace*{0.10in}\bf END

\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm =\hspace*{0.10in}\rm ( \rm
0  $\mapsto$  \rm 1 \rm ) \bf THEN

\hspace*{0.10in}\it pc \rm :=\hspace*{0.10in}\rm 5\rm 6  \hspace*{0.80in}
$\para$

\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm (\it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\hspace*{0.85in}

\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm = \rm ( \rm 1  $\mapsto$ 
\rm 1 \rm ) \bf THEN  \hspace*{0.70in}

\hspace*{0.10in}\it pc \rm := \it bv16\_ushort\rm (\it byte\_bv16\rm ( \it i\_
\rm ,\it bv\_clear\rm (\it rotateleft\rm (\it uchar\_byte\rm (\it byte\_bus\rm )\rm )\rm ,\rm 0\rm )\rm )\rm )$\para$

\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.00in}\bf END\hspace*{0.10in}\\
\end{sloppypar}

\textbf{RESET}  - This just resets the registers related to the interruptions.

\begin{sloppypar}
\it iff1 \rm :=\rm 0 $\para$ \it iff2\rm :=\rm 0 $\para$  \it  im\rm := \rm (\rm 0 $\mapsto$ \rm 0\rm )  $\para$ 
\it pc\rm :=\rm 0 $\para$ \it i\_ \rm := [0,0,0,0,0,0,0,0] $\para$

\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm (\it a0  $\mapsto$ \rm [0,0,0,0,0,0,0,0] , \rm (\it
f0 $\mapsto$ \rm [0,0,0,0,0,0,0,0] \rm \} $\para$

\it r\_  \rm := [0,0,0,0,0,0,0,0] $\para$ \it sp \rm := [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0]\\
\end{sloppypar}

% The animation

The modelling of instruction set, interruptions and input/output ports become the animation interesting.
The animation in ProB version is not so fast and practical as a simulation, for example in \cite{Simulator_z80},
because ProB works manipulating logic and math concepts that can have infinite sets and non determinism, then ProB requires more processing, specially when it works analysing  complex expressions or very large expressions.
However, a user of ProB with this modelling can: execute manually a sequence of Z80's instructions that represent a program and view the visited states of ports, memory and registers, it is important to debug the effect action on the state space computed; create a tracing of the states; analyse constants, functions and expressions; and search possible violations.

%\subsection{Considerations about the Z80 model}

The Z80 formal model provides many benefits, because of the verification of
generated proof obligations guarantees: the correct use of data types, the developed
security properties  and the well-defined of all the expressions. %\footnote{An expression is called ``well-defined'' (or unambiguous) if its definition assigns it a unique interpretation or value.}.
Furthermore, the designer has a big flexibility to create new and specific security properties, this is very useful
to adjust the verification in accordance with the requirements. Moreover, this
example of model could replace or improve the used documentation for users and
assembly programmers. Z80 model is also useful to develop verified software 
up to the assembly level. There are many benefits when a formal model is
developed, but to verify the model is not an easy task, the first version of Z80 model have already verified completely,but the new Z80 model with support to animation in ProB has not been verified completely,
because the changes in Z80 model are recent and the verification time is long.

A case study using the first version of Z80 model was also created and verified
since the first abstract B model up to B assembly model. This case study is an
object of the pilot project developed to analyse a petroleum production test
system in each oil field. Its modelling was developed in according to \cite{LAUT_SERGIO}
and its B assembly model can be executed manually in ProB. The modelling of
instruction set and the B assembly model allow to simulate the execution of microcontroller
and analyse several important properties.

\section{Related works}
\label{sec:relatedworks}

There are in the literature several
approaches~\cite{BHDL_2003,DBLP:journals/entcs/EvansG08,DBLP:conf/fmco/Leuschel08,DBLP:conf/asm/Wright08}
to model hardware and the virtual machines using B. In these works, the B
has been used successfully to model the operational semantic. However the cost of modelling is still expensive.

These recent works \cite{Marc20113,springerlink:Yuan2011} have a similar approach using Event B. 
However, our work use the B method, that seems more appropriated to software development, because it has
implementable language defined, called B0, and tools to convert the models to a programming language.

The main motivation of our research is the development of verified
software up to the assembly level, which requires specifying the
semantics of the underlying hardware. Thus some aspects were not
modelled in our work such as the execution time of the instructions.
Also we did not consider the micro architecture of the hardware as the scope of
our work does not include hardware verification.

\section{Conclusions}
\label{sec:conclusions}

% Um pequeno resumo do trabalho
This work has shown an approach to the formal modelling of the instruction set of
microcontroller using the B method. During the construction of this model, some
problems were found in the official reference for Z80
microcontroller~\cite{Z80_manual}. The designer fixed the found problems in
documentation and also developed a case study: a verified embedded software. This
case study was interesting to analyse the technique of formal verification up to
assembly level.

% TODO: Adaptar para apenas fundamentar melhor o trabalho já desenvolvido. Citar o artigo SBMF e a dissertação. 
% Os meus trabalhos já desenvolvidos/ resultados atuais.
The next works quoted are directly related to the objective this paper.
The first work \cite{DantasSemish2008} shows the developed metodology to verify
software up to assembly level using B. A second work\cite{Dantas_SBMF08}
presents more details about the verification approach and a small example
software verified up to assembly level in three different platforms. 
The last work \cite{Valerio_SBMF09} shows a general view of the first version of Z80
model and techniques used in verification process.

%Link com o parágrafo anterior 
These works created a important step to software verification up to assembly
language, they show the actuals difficulties and suggest improvements in the
B tools.

% TODO
% [Trabalhos futuros ] LLVM:Existe também a pretensão de especificar as

Future works comprise the development of software with the B method from
functional specification to assembly level, using the Z80 model presented in this
work. The mechanic compilation from B algorithmic constructs to assembly platform
is also envisioned. Finally, another important activity is develop a formal model of a platform used in LLVM Compiler~\cite{DBLP:conf/cgo/LattnerA04}.

%\paragraph{Acknowledges:}
%This work was partially supported by INES (www.ines.org.br), funded by CNPq
%grant \ldots
%573964/2008-4 and by CNPq grants 553597/2008-6, 550946/2007-1, and 620132/2008-6..

%\bibliographystyle{splncs}

\bibliographystyle{plain}
\bibliography{paper}
% [To change the format; to add: thesis suggested by David, Article from ABZ 2010 suggested by David  ]
% \begin{thebibliography}{5}
%

\end{document}

%
% \bibitem {clar:eke}
% Clarke, F., Ekeland, I.:
% Nonlinear oscillations and
% boundary-value problems for Hamiltonian systems.
% Arch. Rat. Mech. Anal. {\bf 78} (1982) 315--333
% %
% \bibitem {clar:eke:2}
% Clarke, F., Ekeland, I.:
% Solutions p\'{e}riodiques, du
% p\'{e}riode donn\'{e}e, des \'{e}quations hamiltoniennes.
% Note CRAS Paris {\bf 287} (1978) 1013--1015
% %
% \bibitem {mich:tar}
% Michalek, R., Tarantello, G.:
% Subharmonic solutions with prescribed minimal
% period for nonautonomous Hamiltonian systems.
% J. Diff. Eq. {\bf 72} (1988) 28--55
% %
% \bibitem {tar}
% Tarantello, G.:
% Subharmonic solutions for Hamiltonian
% systems via a $\bbbz_{p}$ pseudoindex theory.
% Annali di Matematica Pura (to appear)
% %
% \bibitem {rab}
% Rabinowitz, P.:
% On subharmonic solutions of a Hamiltonian system.
% Comm. Pure Appl. Math. {\bf 33} (1980) 609--633
% \end{thebibliography}
%

### 2012-01-09 trunk/papers/ArticleZ80/source/paper.tex
Change the related works (I think that it can be better).

--- 2012-01-08 papers/ArticleZ80/source/paper.tex
@@
 
-There are in the literature several
-approaches~\cite{BHDL_2003,DBLP:journals/entcs/EvansG08,DBLP:conf/fmco/Leuschel08,DBLP:conf/asm/Wright08}
-to model hardware and the virtual machines using B. In these works, the B
-has been used successfully to model the operational semantic. However the cost of modelling is still expensive.
-
-These recent works \cite{Marc20113,springerlink:Yuan2011} have a similar approach using Event B. 
-However, our work use the B method, that seems more appropriated to software development, because it has
-implementable language defined, called B0, and tools to convert the models to a programming language.
-
-The main motivation of our research is the development of verified
-software up to the assembly level, which requires specifying the
-semantics of the underlying hardware. Thus some aspects were not
-modelled in our work such as the execution time of the instructions.
-Also we did not consider the micro architecture of the hardware as the scope of
-our work does not include hardware verification.
+There are in the literature several similar approaches to model hardware and instruction set using B.
+The paper \cite{springerlink:Yuan2011} reports a method to specify, design and construct sound and complete instruction set model by stepwise refinement and formal proof using the formal method Event-B.
+It also discuss about desirable properties of instruction set model.
+However, our work \cite{Valerio_SBMF09} and \cite{Subotic2010} use the B method, that seems more appropriated to software development, because it has implementable language defined,
+called B0, and tools to convert the models to a programming language.
+In \cite{Marc20113} is present an related experience on using B in the design of secure micro-controllers.
+It tries to the feasibility of such a technique for high confidence trustful devices.
+The work \cite{Subotic2010} describes a similar approach with MIPS CPU architecture and it shows how the CPU is formally specified and implemented using the B method.
+
+% 
+%There are in the literature several
+%approaches~\cite{BHDL_2003,DBLP:journals/entcs/EvansG08,DBLP:conf/fmco/Leuschel08,DBLP:conf/asm/Wright08}
+%to model hardware and the virtual machines using B. In these works, the B
+%has been used successfully to model the operational semantic. However the cost of modelling is still expensive. 
+%
+%These recent works \cite{Marc20113,springerlink:Yuan2011} have a similar approach using Event B. 
+%However, our work use the B method, that seems more appropriated to software development, because it has
+%implementable language defined, called B0, and tools to convert the models to a programming language.
+%
+%
+%The main motivation of our research is the development of verified
+%software up to the assembly level, which requires specifying the
+%semantics of the underlying hardware. Thus some aspects were not
+%modelled in our work such as the execution time of the instructions.
+%Also we did not consider the micro architecture of the hardware as the scope of
+%our work does not include hardware verification.
 

### 2012-01-17 trunk/papers/ArticleZ80/source/paper.tex
Several updates.

@@
 %
-%\hyphenation{se-pa-ra-te}
+\hyphenation{se-pa-ra-te}
+\hyphenation{in-struc-tion}
+\hyphenation{re-gis-ter}
+\hyphenation{re-gis-ters}
 %
@@
 
-  This paper describes an approach to model the functional aspects of the instruction set of
-  microcontroller platforms and several details about the representation of
-  elements from microcontrollers.  Several models were developed using the notation of the B method. 
-  They are used to develop a formally verified software up to the assembly level and allow the
-  simulation of models. This simulation is able to guarantee the consistency between the execution
-  of a software model and a real  execution of software, it has wide variety of usages in industry
-  and academia. This paper presents specifically the case of the Z80 platform and quote a theoretic 
-  case study important in tanks of the petroleum industry. This work is a contribution towards the
-  extension of the B method to handle developments up to the assembly level code.
+  This paper describes an approach to model the functional aspects of
+  the instruction set of microcontrollers and several details about
+  the representation of microcontrollers.  Different models were
+  developed using the notation of the B method. This paper presents
+  specifically a B-based formal model of the Z80 instruction set.
 
 %  \paragraph[Keywords]{
-   \textbf{Keywords}: Embedded Software, Simulation, B method and Verification%}
+%   \textbf{Keywords}: Embedded Software, Simulation, B method and Verification%}
 \end{abstract}
@@
 
-The B method~\cite{Abrial} supports the construction of safety systems
-models by generating proof obligation that must be verified to guarantee its correctness. So,
-an initial abstract model of the system requirements is defined and
-then it is refined up to the implementation model. Development
-environments based on the B method also include source code generators
-for programming languages, but the result of this translation
-cannot be compared by formal means. \cite{Dantas_SBMF08} proposed
-an approach to extend the scope of the B method up to the
-assembly level language. One key component of this approach
-is to build, within the framework of the B method, formal models of
-the instruction set of such assembly languages.
+The B method~\cite{Abrial} supports the formal development of software
+from a specification of functional requirements down to an imperative
+implementation of those requirements, amenable to synthesis to a
+programming language such as C or Ada. However, the synthesis step is
+not amenable to proof of correctness, and~\cite{Dantas_SBMF08} proposed
+an approach to extend the scope of the B method up to the assembly
+level language. One key component of this approach is to build, within
+the framework of the B method, formal models of the instruction set of
+such assembly languages.
 
@@
 \subsection{Problems in formalizing instruction set}
-
+% Compact version 
 In general, the manuals of microcontrollers can be more suited for
@@
 is shown in its official manual using textual description, math description, examples,
-encoding, number of cycles and other information.Some times, the instruction descriptions
+encoding, number of cycles and other information. Some times, the instruction descriptions
 have an own notation, some informations are organized in different sections of document 
 and the textual description are not standardized. This does not allow the interpretation by parser.
-% b) The case of Z80
-For example in official manual of the microcontroller Z80 \cite{Z80_manual}, if the users
-want to know the action of an instruction and its effects on the flag register then he needs
-to read an informal textual description. So, this description can add
-some ambiguities or even mistakes. The official manual has also several problems that were 
-identified over time~\cite{UndocumentedZ80}. The Z80 is a microcontroller that  has been
-tested and used for many years. This intensive use and test facilitated to find several
-errors in the description of the instructions in its manual. Several of these problems
-have been reported in a technical report \cite{UndocumentedZ80}. Some problems are inaccurate
-information, for example the description of the action of an instruction on the flag
-register, partial information and distributed on different pages of manual.
-
-% in the Z80 manual happens was that,
-
-%A simple example of error is in some official documentation saying that a group of instructions leaves the \textit{CF} flag unaffected and the \textit{NF}
-%flag is not always set, but may also be reset \cite{UndocumentedZ80}.
-%% \begin{itemize} \item Some input and output instructionsThe
-%% INI/INIR/IND/INDR/OUTI/OUTD/OTIR/OTDR instructions do affect the CF flag (some
-%% official documentation says they leave it unaffected, important!) and the NF
-%% flag is not always set but may also be reset (see 4.3 for exact operation)
-%% \cite{UndocumentedZ80}. Este erro he descrito na phegina (274 do documento e
-%% 294 do pdf) do manual oficial do Z80 (Z80_CPU_adotado.pdf) Citar apenas esse
-%% exemplo abaixo! \item
-%%when a determined interruption (NMI) is accepted, the bit IFF1 is not   
-%%copied to bit IFF2~\cite{UndocumentedZ80}. 
-%%
-%This example can cause serious problems for programmers.
-%% Este erro he descrito na pagina (22 do documento e 26do pdf) do manual
-%% official do Z80 (Z80_CPU_adotado.pdf) Only IFF1 is reset.
-%This error and others 7 different errors were previously identified by~\cite{UndocumentedZ80}. So to help the 
-%construction of Z80 B model, other references~\cite{UndocumentedZ80,Simulator_z80}
-%were also consulted.
+The manuals also have semi-formal and informal descriptions, these descriptions may make mistakes.
+The official manual of the microcontroller Z80 has also several problems that were identified
+over time~\cite{UndocumentedZ80}.
 
@@
 
+%In general, the manuals of microcontrollers can be more suited for
+%developers in assembly language. Many times, each instruction of microcontrollers
+%is shown in its official manual using textual description, math description, examples,
+%encoding, number of cycles and other information.Some times, the instruction descriptions
+%have an own notation, some informations are organized in different sections of document 
+%and the textual description are not standardized. This does not allow the interpretation by parser.
+%% b) The case of Z80
+%For example in official manual of the microcontroller Z80 \cite{Z80_manual}, if the users
+%want to know the action of an instruction and its effects on the flag register then he needs
+%to read an informal textual description. So, this description can add
+%some ambiguities or even mistakes. The official manual has also several problems that were 
+%identified over time~\cite{UndocumentedZ80}. The Z80 is a microcontroller that  has been
+%tested and used for many years. This intensive use and test facilitated to find several
+%errors in the description of the instructions in its manual. Several of these problems
+%have been reported in a technical report \cite{UndocumentedZ80}. Some problems are inaccurate
+%information, for example the description of the action of an instruction on the flag
+%register, partial information and distributed on different pages of manual.
+%
+%
+%% in the Z80 manual happens was that,
+%
+%%A simple example of error is in some official documentation saying that a group of instructions leaves the \textit{CF} flag unaffected and the \textit{NF}
+%%flag is not always set, but may also be reset \cite{UndocumentedZ80}.
+%%% \begin{itemize} \item Some input and output instructionsThe
+%%% INI/INIR/IND/INDR/OUTI/OUTD/OTIR/OTDR instructions do affect the CF flag (some
+%%% official documentation says they leave it unaffected, important!) and the NF
+%%% flag is not always set but may also be reset (see 4.3 for exact operation)
+%%% \cite{UndocumentedZ80}. Este erro he descrito na phegina (274 do documento e
+%%% 294 do pdf) do manual oficial do Z80 (Z80_CPU_adotado.pdf) Citar apenas esse
+%%% exemplo abaixo! \item
+%%%when a determined interruption (NMI) is accepted, the bit IFF1 is not   
+%%%copied to bit IFF2~\cite{UndocumentedZ80}. 
+%%%
+%%This example can cause serious problems for programmers.
+%%% Este erro he descrito na pagina (22 do documento e 26do pdf) do manual
+%%% official do Z80 (Z80_CPU_adotado.pdf) Only IFF1 is reset.
+%%This error and others 7 different errors were previously identified by~\cite{UndocumentedZ80}. So to help the 
+%%construction of Z80 B model, other references~\cite{UndocumentedZ80,Simulator_z80}
+%%were also consulted.
+%
+%A solution to avoid ambiguity and inconsistencies is to specify formally the
+%assembly instruction set, this creates a pattern of representation and validates
+%properties of instructions. The formal model also restricts the definitions to
+%correct typing and uses only well-defined expressions. Furthermore, the developer
+%can use with B method different levels of abstraction and add more specific details
+%in refined models.
+% 
 %Finally, the main goal of this project, though, is to provide a basis for the
@@
 %\subparagraph{Report structure}
-
-This work is focused on the presentation of modelling of assembly instruction
-set, including elementary libraries to describe hardware aspects; and the
-modelling and simulation of a case study of an embedded verified software that
-can be used in tanks of the petroleum industry.
+%
+%This work is focused on the presentation of modelling of assembly instruction
+%set, including elementary libraries to describe hardware aspects; and the
+%modelling and simulation of a case study of an embedded verified software that
+%can be used in tanks of the petroleum industry.
 
@@
 presents the B model of the Z80 instruction set. %Section ~\ref{sec:Verification} provides some information on the proof effort needed to analyse the presented models.
-Section ~\ref{sec:studycase} explains the building and verifying assembly-level refinements. Related works are discussed in
-section~\ref{sec:relatedworks}. Finally, the last section is devoted to the conclusions.
+%Section ~\ref{sec:studycase} explains the building and verifying assembly-level refinements. 
+Related works are discussed in section~\ref{sec:relatedworks}. Finally, the last section is devoted to the conclusions.
 %
@@
 
-\hspace*{0.0in}\it BYTE\_WIDTH \rm = 8 $\land$ \it BYTE\_INDEX \rm = 1 $\upto$ \rm  BYTE\_WIDTH\rm  \hspace*{0.05in} $\land$
-
-\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.10in} $\land$
-
-\hspace*{0.0in}\it BYTE \rm =\rm (\it BYTE\_INDEX  $\fun$  \it BIT\rm )\hspace*{0.10in} $\land$  \bf card\rm (\it BYTE\rm )\rm = $2^{8}$  $\land$
+\hspace*{0.0in}\it BYTE\_WIDTH \rm = 8 $\land$ \it BYTE\_INDEX \rm = 1 $\upto$ \rm  BYTE\_WIDTH\rm  \hspace*{0.03in} $\land$
+
+\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.03in} $\land$ \hspace*{0.0in}\it BYTE \rm =\rm (\it BYTE\_INDEX  $\fun$  \it BIT\rm )\hspace*{0.0in} $\land$
 
@@
 Nearly 50\% of all these proof obligations shown in the table~\ref{tab:types} 
-are verified almost automatically. Although many others proof obligations are difficult to verify, because these
-check the type of functions with non-linear arithmetic is a difficult task.
-
-\subsection{Verification Process in Hardware Library and Types Library} 
+are verified almost automatically. Although many others proof obligations are difficult to verify,
+because they are functions of type conversion and contain arithmetic expressions difficult 
+to manipulate with the interactive prover of AtelierB.
+
+\subsection{Verification in Hardware Library and Types Library} 
 \label{sec:VerificationHardwareLibrary}
@@
 
+	\begin{center}
 	\begin{sloppypar}
 
-\bf MACHINE
-
-\hspace*{0.15in}\it Z80
-
-\bf INCLUDES
-
+	  \begin{tabbing}
+	    \begin{array}[t]{l}
+		\bf MACHINE\\
+		\hspace*{0.15in}\it Z80\\
+		\bf INCLUDES\\
 		\hspace*{0.10in}\it MEMORY
-
-\bf SEES
-
-\hspace*{0.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,
-
-\hspace*{0.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,
-
-\hspace*{0.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,
-
-\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION
+	    \end{array}
+	     \hspace*{1cm} 
+	    \begin{array}[t]{l}
+	    \bf SEES\\
+		\hspace*{0.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,\\
+		\hspace*{0.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,\\
+		\hspace*{0.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,\\
+		\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION\\
+	    \end{array}
+	  \end{tabbing}
+
 	 \end{sloppypar} 
+	\end{center}
 
@@
 $\textit{iff2}$) used to control the interruptions, a pair of bits to define the interruption mode ($\textit{im}$)
-and the input and output ports ($\textit{i\_o\_ports}$). Below, its definitions are represented by
+and the input and output ports ($\textit{i\_o\_ports}$). These definitions are represented by
 \textit{INVARIANT}.
@@
 
-\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$
-
-\hspace*{0.10in}\it pc  $\in$  \it INSTRUCTION  $\land$  \it sp  $\in$  \it BV16  $\land$  \it ix  $\in$  \it BV16  $\land$  \it iy  $\in$  \it BV16  $\land$
-
-\hspace*{0.10in}\it i\_  $\in$  \it BYTE  $\land$  \it r\_ $\in$  \it BYTE  $\land$
-
-\hspace*{0.10in}\it iff1  $\in$  \it BIT  $\land$ \hspace*{0.10in}\it iff2  $\in$  \it BIT  $\land$
-
-\hspace*{0.10in}\it im $\in$ (\it BIT $\times$ \it BIT\rm )  $\land$
-
-\hspace*{0.10in}\it i\_o\_ports  $\in$  \it BYTE  $\fun$  \it BYTE
+\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ \it pc  $\in$  \it INSTRUCTION  $\land$  \it sp  $\in$  \it BV16 $\land$  \it ix  $\in$  \it BV16  $\land$
+
+\hspace*{0.10in}\it iy  $\in$  \it BV16  $\land$ \it i\_  $\in$  \it BYTE  $\land$  \it r\_ $\in$  \it BYTE  $\land$ \it iff1  $\in$  \it BIT  $\land$ \it iff2  $\in$  \it BIT  $\land$
+
+\hspace*{0.10in}\it im $\in$ (\it BIT $\times$ \it BIT\rm )  $\land$  \it i\_o\_ports  $\in$  \it BYTE  $\fun$  \it BYTE
 \end{sloppypar}
@@
 (\textit{a\_0,b\_0,c\_0,d\_0,e\_0,f\_0,h\_0,l\_0}). These registers can be accessed in pairs, forming
-16-bits, resulting in another set of identifiers of 16-bits registers, named $\textit{id\_reg16}$.
+16-bits, resulting in another set of identifiers of 16-bits registers, named $\textit{id\_reg16}$. These sets are represented below.
 
@@
 	\bf SETS\\
-	\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0 \rm , \it f0 \rm , \it f\_0 \rm , \it a\_0 \rm ,\\
-	\hspace*{0.93in}\it b0 \rm , \it c0 \rm , \it b\_0 \rm , \it c\_0\rm,\\
-	\hspace*{0.93in}\it d0 \rm , \it e0 \rm , \it d\_0 \rm , \it e\_0\rm,\\
-	\hspace*{0.93in}\it h0 \rm , \it l0 \rm , \it h\_0 \rm , \it l\_0\};\\
+	\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0 \rm , \it f0 \rm ,\it b0 \rm , \it c0 \rm ,\it d0 \rm , \it e0 \rm ,\it h0 \rm , \it l0 \rm ,\it a\_0 \rm , \it f\_0 \rm ,\it b\_0 \rm , \it c\_0 \rm ,\it d\_0 \rm , \it e\_0 \rm ,\it h\_0 \rm , \it l\_0 \rm \};\\
 	\hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{ \it BC \rm , \it DE \rm , \it HL \rm , \it SP \rm , \it AF \rm \}
@@
 
-The main working register of Z80 is the accumulator ($\textit{rgs8(a0)}$) used for arithmetic/logic,
-input/output and loading/storing operations.
-
-\subsubsection{Flag register:}
-another important element is the ``f'' register ($\textit{rgs8(f0)}$),  that is
-used as a flag register. This register uses only six bits to represent the execution result status of each instruction.
+%\begin{sloppypar}
+%
+%	\bf SETS\\
+%	\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0 \rm , \it f0 \rm , \it f\_0 \rm , \it a\_0 \rm ,\\
+%	\hspace*{0.93in}\it b0 \rm , \it c0 \rm , \it b\_0 \rm , \it c\_0\rm,\\
+%	\hspace*{0.93in}\it d0 \rm , \it e0 \rm , \it d\_0 \rm , \it e\_0\rm,\\
+%	\hspace*{0.93in}\it h0 \rm , \it l0 \rm , \it h\_0 \rm , \it l\_0\};\\
+%	\hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{ \it BC \rm , \it DE \rm , \it HL \rm , \it SP \rm , \it AF \rm \}
+%\end{sloppypar}
+
+The main working register of Z80 is the accumulator ($\textit{rgs8(a0)}$) used for arithmetic/logic, input/output and loading/storing operations.
+Another important element is the ``f'' register ($\textit{rgs8(f0)}$),  that is
+used as a \textbf{flag register}. This register uses only six bits to represent the execution result status of each instruction.
 According to the official manual the bits 3 and 5 are not used and the others bits have the follow meaning: $\textit{bv\_get(rgs8(f0),0)}$ - the carry bit; $\textit{bv\_get(rgs8(f0),1)}$ - the add/subtract bit; $\textit{bv\_get(rgs8(f0),2)}$ - the parity or overflow bit; $\textit{bv\_get(rgs8(f0),4)}$ - the half carry bit;
@@
 BIT $\times$  \it BIT  $\times$  \it BIT\rm) $\fun$  \rm (\rm \{\it f0\rm \}  $\times$  \it BYTE\rm ).
-
-\it update\_flag\_reg \rm =  $\lambda$  \rm (\it s7 \rm, \it z6 \rm,\it h4 \rm,\it pv2 \rm ,\it n1
+The definition is: \it update\_flag\_reg \rm =  $\lambda$  \rm (\it s7 \rm, \it z6 \rm,\it h4 \rm,\it pv2 \rm ,\it n1
 \rm ,\it c0 \rm) . 
-\rm ( \it s7 $\in$ \it BIT $\land$ \it z6 $\in$ \it BIT $\land$ \it h4 $\in$ \it BIT $\land$ \it pv2 $\in$ \it BIT $\land$ \it n1 $\in$ \it BIT $\land$ \it c0 $\in$ \it BIT \hspace*{0.20in}\rm $\mid$( \it f0  $\mapsto$  \rm \rm [\it c0\rm , \it n1\rm , \it pv2\rm , \rm 1\rm , \it h4\rm , \rm 1\rm , \it z6\rm , \it s7\rm \rm ]\rm ) \rm )
+\rm ( \it s7 $\in$ \it BIT $\land$ \it z6 $\in$ \it BIT $\land$ \it h4 $\in$ \it BIT $\land$ \it pv2 $\in$ \it BIT $\land$ \it n1 $\in$ \it BIT $\land$ \it c0 $\in$ \it BIT \rm $\mid$( \it f0  $\mapsto$ \rm [\it c0\rm , \it n1\rm , \it pv2\rm , \rm 1\rm , \it h4\rm , \rm 1\rm , \it z6\rm , \it s7\rm \rm ]\rm ) \rm )
 
@@
 general-purpose data. The memory has 16-bit addresses and each address holds a byte.
-Thus, the data from the memory module is very simple:
-
-\begin{sloppypar}
-\bf INVARIANT \\
-\hspace*{0.10in}\it mem  $\in$  \it BV16  $\fun$  \it BYTE
-
-\end{sloppypar}
+Thus, the memory is very simple: \it mem  $\in$  \it BV16  $\fun$  \it BYTE\rm.
 
@@
 access and change data in memory. For added  security, it is important that the program instructions has
-limited access by region. Thus the designer can specify address regions to restrict the access from
+limited access by region. Thus the designer can specify address regions in a refinement model to restrict the access from
 instructions. The address regions can be specified using constants
@@
 address for: programs instructions, general purpose data and data stack.
-
-$
-\begin{array}{l}
-\textit{PROGRAM\_R\_ADR} = 0..16384 \land\\
-\textit{DATA\_R\_ADR} = 16385..49151 \land\\
-\textit{STACK\_R\_ADR} = 49152..65535
-\end{array}
-$
+%
+%$
+%\begin{array}{l}
+%\textit{PROGRAM\_R\_ADR} = 0..16384 \land\\
+%\textit{DATA\_R\_ADR} = 16385..49151 \land\\
+%\textit{STACK\_R\_ADR} = 49152..65535
+%\end{array}
+%$
 
@@
  \emph{To assure that address regions are well defined, then the designer must to verify the expression:}
-\begin{sloppypar}
-\hspace*{0.10in}\it PROGRAM\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$  STACK\_R\_ADR $=$ \{\}
-\end{sloppypar}
+%\begin{sloppypar}
+%\hspace*{0.10in}
+\it PROGRAM\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ STACK\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ PROGRAM\_R\_ADR $\cap$  STACK\_R\_ADR  $=$ \{\} 
+%\end{sloppypar}
 
@@
 
-\hspace*{0.0in}
-
 \hspace*{0.0in}\it half8UCHAR  $\in$  \it UCHAR  $\fun$  \it UCHAR  $\land$
@@
 \hspace*{0.0in}\it half8UCHAR \rm =  $\lambda$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it UCHAR  $\mid$  \it ww  $\mod$  \it $2^{4}$\rm )
-
-\hspace*{0.0in}
 
@@
 
-\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm ) \rm + \it half8UCHAR\rm ( \it
-w2\rm )  $\geq$  \it $2^{4}$\rm )\rm,
+\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm) \rm + \it half8UCHAR\rm (\it w2\rm)  $\geq$  \it $2^{4}$\rm )\rm,
 
@@
 In \cite{Marc20113} is present an related experience on using B in the design of secure micro-controllers.
-It tries to the feasibility of such a technique for high confidence trustful devices.
+It tries to show the feasibility of such a technique for high confidence trustful devices.
 The work \cite{Subotic2010} describes a similar approach with MIPS CPU architecture and it shows how the CPU is formally specified and implemented using the B method.
@@
 is also envisioned. Finally, another important activity is develop a formal model of a platform used in LLVM Compiler~\cite{DBLP:conf/cgo/LattnerA04}.
+
+\subsubsection{Acknowledgements:} This experiment could not have succeeded without the interesting work of Michael Leuschel who developed several improvements in the ProB and provided important aid to enable the animation of the Z80 modelling.
+
+% This work was partially supported by INES (www.ines.org.br),
+%funded by CNPq grant 573964/2008-4 and by CNPq grants 553597/2008-6, 550946/2007-
+%1, and 620132/2008-6
 

### 2012-01-19 papers/ArticleZ80/source/paper.tex

@@
 
-The B method~\cite{Abrial} supports the construction of safety systems
-models by generating proof obligation that must be verified to guarantee its correctness. So,
-an initial abstract model of the system requirements is defined and
-then it is refined up to the implementation model. Development
-environments based on the B method also include source code generators
-for programming languages, but the result of this translation
-cannot be compared by formal means. \cite{Dantas_SBMF08} proposed
-an approach to extend the scope of the B method up to the
-assembly level language. One key component of this approach
-is to build, within the framework of the B method, formal models of
-the instruction set of such assembly languages.
-
-This work presents a formal modell of the instruction set of the Z80 
-microcontroller~\cite{Z80_manual}. This microcontroller
-was chosen because of several factors: it contains essential and common
-concepts in microcontrollers and until microprocessors; it has extensive
-documentation available; and it was widely used and is still in use. 
-A first Z80 modell presented in \cite{Valerio_SBMF09} was verified completely,
-but it uses constructs that are not well supported to animate. 
-This modelling was changed to support animation in ProB\cite{proB}.
-Basically, the new modelling\footnote{The interested reader in more
-details is invited to visit our repository at: http://code.google.com/p/b2asm/.}
-changed elements represented by infinite sets and adjusted the implications
-of this change. Using the responsibility division mechanisms provided by B, auxiliary
-libraries of basic modules were developed as part of the construction of microcontroller
-model. Such library has many definitions about common concepts used in
-the microcontrollers;besides the new Z80 model, a theoretic case study in 
-petroleum production test system was developed using Z80 assembly language
-to analyse the code simulation and verification process.
-
-The formal model of instruction set provides many benefits. The animation of 
-Z80 model allows to simulate the execution assembly programs including
-support for interrupts and input and output instructions. Other possible uses
-of a formal model of a microcontroller instruction set include documentation,
-the construction of simulators, and can be possibly the starting point of a 
-verification effort for the actual implementation of a Z80 design. Moreover, 
-the model of the instruction set could be instrumented with non-functional
-aspects, such as the number of cycles it takes to execute an instruction, to
-prove lower and upper bounds on the execution time of a routine.
-Two aspects of these formalization are of particular importance to us:
-provide a more solid documentation artifact for microcontrollers, 
-build a reference model for a formal compilation approach in B.
+The B method~\cite{Abrial} supports the formal development of software
+from a specification of functional requirements down to an imperative
+implementation of those requirements, amenable to synthesis to a
+programming language such as C or Ada. However, the synthesis step is
+not amenable to proof of correctness, and~\cite{Dantas_SBMF08} proposed
+an approach to extend the scope of the B method up to the assembly
+level language. One key component of this approach is to build, within
+the framework of the B method, formal models of the instruction set of
+such assembly languages.
+
+This work presents a formal model of the instruction set of the Z80
+microcontroller~\cite{Z80_manual}. This microcontroller was chosen
+because of several factors: it contains essential and common concepts
+of microcontrollers and even microprocessors; it has extensive
+documentation available; and it has been widely used.  A first,
+partial, Z80 model presented in \cite{Valerio_SBMF09} was fully
+verified, but it uses constructs that make it impossible to animate.
+
+The new model supports animation in ProB\cite{proB}.  Essentially, the
+new model\footnote{The interested reader in more details is invited to
+  visit our repository at: http://code.google.com/p/b2asm/.}  no
+longer contains definitions derived from infinite sets. Using B
+constructs for modularity, auxiliary libraries of basic modules were
+developed as part of the construction of microcontroller model. Such
+library has many definitions of concepts common in microcontrollers.
+Besides the new Z80 model, a petroleum production test system was
+developed using Z80 assembly language to analyse the code simulation
+and verification process.
+
+Providing the formal model of an instruction set yields many
+benefits. The animation of Z80 model allows to simulate the execution
+assembly programs including support for interrupts and input and
+output instructions. Other possible uses of a formal model of a
+microcontroller instruction set include documentation, the
+construction of simulators, and can be possibly the starting point of
+a verification effort for the actual implementation of a Z80
+design. Moreover, the model of the instruction set could be
+instrumented with non-functional aspects, such as the number of cycles
+it takes to execute an instruction, to prove lower and upper bounds on
+the execution time of a routine.  Two aspects of these formalization
+are of particular importance to us: provide a more solid documentation
+artifact for microcontrollers, build a reference model for a formal
+compilation approach in B.
 
 \subsection{Problems in formalizing instruction set}
-
-In general, the manuals of microcontrollers can be more suited for
-developers in assembly language. Many times, each instruction of microcontrollers
-is shown in its official manual using textual description, math description, examples,
-encoding, number of cycles and other information.Some times, the instruction descriptions
-have an own notation, some informations are organized in different sections of document 
-and the textual description are not standardized. This does not allow the interpretation by parser.
-% b) The case of Z80
-For example in official manual of the microcontroller Z80 \cite{Z80_manual}, if the users
-want to know the action of an instruction and its effects on the flag register then he needs
-to read an informal textual description. So, this description can add
-some ambiguities or even mistakes. The official manual has also several problems that were 
-identified over time~\cite{UndocumentedZ80}. The Z80 is a microcontroller that  has been
-tested and used for many years. This intensive use and test facilitated to find several
-errors in the description of the instructions in its manual. Several of these problems
-have been reported in a technical report \cite{UndocumentedZ80}. Some problems are inaccurate
-information, for example the description of the action of an instruction on the flag
-register, partial information and distributed on different pages of manual.
-
-% in the Z80 manual happens was that,
-
-%A simple example of error is in some official documentation saying that a group of instructions leaves the \textit{CF} flag unaffected and the \textit{NF}
-%flag is not always set, but may also be reset \cite{UndocumentedZ80}.
-%% \begin{itemize} \item Some input and output instructionsThe
-%% INI/INIR/IND/INDR/OUTI/OUTD/OTIR/OTDR instructions do affect the CF flag (some
-%% official documentation says they leave it unaffected, important!) and the NF
-%% flag is not always set but may also be reset (see 4.3 for exact operation)
-%% \cite{UndocumentedZ80}. Este erro he descrito na phegina (274 do documento e
-%% 294 do pdf) do manual oficial do Z80 (Z80_CPU_adotado.pdf) Citar apenas esse
-%% exemplo abaixo! \item
-%%when a determined interruption (NMI) is accepted, the bit IFF1 is not   
-%%copied to bit IFF2~\cite{UndocumentedZ80}. 
-%%
-%This example can cause serious problems for programmers.
-%% Este erro he descrito na pagina (22 do documento e 26do pdf) do manual
-%% official do Z80 (Z80_CPU_adotado.pdf) Only IFF1 is reset.
-%This error and others 7 different errors were previously identified by~\cite{UndocumentedZ80}. So to help the 
-%construction of Z80 B model, other references~\cite{UndocumentedZ80,Simulator_z80}
-%were also consulted.
-
-A solution to avoid ambiguity and inconsistencies is to specify formally the
-assembly instruction set, this creates a pattern of representation and validates
-properties of instructions. The formal model also restricts the definitions to
-correct typing and uses only well-defined expressions. Furthermore, the developer
-can use with B method different levels of abstraction and add more specific details
+% Compact version
+
+In general, the manuals of microcontrollers can be more suited for \marginpar{DD: o título da sub-seção não condiz com seu conteúdo.
+Não cita nenhum problema na formalização.
+Não vejo muito bem o que ele traz de novo que já não foi dito anteriormente.}
+developers in assembly language. Many times, each instruction of
+microcontrollers is shown in its official manual using textual
+description, math description, examples, encoding, number of cycles
+and other information. Sometimes, the instruction descriptions have
+an own notation, some informations are organized in different sections
+of document and the textual description are not standardized. This
+does not allow the interpretation by parser.  The manuals also have
+semi-formal and informal descriptions, these descriptions may make
+mistakes.  The official manual of the microcontroller Z80 has also
+several problems that were identified over
+time~\cite{UndocumentedZ80}.
+
+A solution to avoid ambiguity and inconsistencies is to specify
+formally the assembly instruction set, this creates a pattern of
+representation and validates properties of instructions. The formal
+model also restricts the definitions to correct typing and uses only
+well-defined expressions. Furthermore, the developer can use with B
+method different levels of abstraction and add more specific details
 in refined models.
 
+%In general, the manuals of microcontrollers can be more suited for
+%developers in assembly language. Many times, each instruction of microcontrollers
+%is shown in its official manual using textual description, math description, examples,
+%encoding, number of cycles and other information.Some times, the instruction descriptions
+%have an own notation, some informations are organized in different sections of document 
+%and the textual description are not standardized. This does not allow the interpretation by parser.
+%% b) The case of Z80
+%For example in official manual of the microcontroller Z80 \cite{Z80_manual}, if the users
+%want to know the action of an instruction and its effects on the flag register then he needs
+%to read an informal textual description. So, this description can add
+%some ambiguities or even mistakes. The official manual has also several problems that were 
+%identified over time~\cite{UndocumentedZ80}. The Z80 is a microcontroller that  has been
+%tested and used for many years. This intensive use and test facilitated to find several
+%errors in the description of the instructions in its manual. Several of these problems
+%have been reported in a technical report \cite{UndocumentedZ80}. Some problems are inaccurate
+%information, for example the description of the action of an instruction on the flag
+%register, partial information and distributed on different pages of manual.
+%
+%
+%% in the Z80 manual happens was that,
+%
+%%A simple example of error is in some official documentation saying that a group of instructions leaves the \textit{CF} flag unaffected and the \textit{NF}
+%%flag is not always set, but may also be reset \cite{UndocumentedZ80}.
+%%% \begin{itemize} \item Some input and output instructionsThe
+%%% INI/INIR/IND/INDR/OUTI/OUTD/OTIR/OTDR instructions do affect the CF flag (some
+%%% official documentation says they leave it unaffected, important!) and the NF
+%%% flag is not always set but may also be reset (see 4.3 for exact operation)
+%%% \cite{UndocumentedZ80}. Este erro he descrito na phegina (274 do documento e
+%%% 294 do pdf) do manual oficial do Z80 (Z80_CPU_adotado.pdf) Citar apenas esse
+%%% exemplo abaixo! \item
+%%%when a determined interruption (NMI) is accepted, the bit IFF1 is not   
+%%%copied to bit IFF2~\cite{UndocumentedZ80}. 
+%%%
+%%This example can cause serious problems for programmers.
+%%% Este erro he descrito na pagina (22 do documento e 26do pdf) do manual
+%%% official do Z80 (Z80_CPU_adotado.pdf) Only IFF1 is reset.
+%%This error and others 7 different errors were previously identified by~\cite{UndocumentedZ80}. So to help the 
+%%construction of Z80 B model, other references~\cite{UndocumentedZ80,Simulator_z80}
+%%were also consulted.
+%
+%A solution to avoid ambiguity and inconsistencies is to specify formally the
+%assembly instruction set, this creates a pattern of representation and validates
+%properties of instructions. The formal model also restricts the definitions to
+%correct typing and uses only well-defined expressions. Furthermore, the developer
+%can use with B method different levels of abstraction and add more specific details
+%in refined models.
+% 
 %Finally, the main goal of this project, though, is to provide a basis for the
@@
 %\subparagraph{Report structure}
-
-This work is focused on the presentation of modelling of assembly instruction
-set, including elementary libraries to describe hardware aspects; and the
-modelling and simulation of a case study of an embedded verified software that
-can be used in tanks of the petroleum industry.
-
-This paper is structured as it follows. %Section~\ref{sec:B_method} provides a short introduction to the B method. 
-Section~\ref{sec:models} presents the elementary libraries and
-the modelling of some elements common to microcontrollers.  Section~\ref{sec:z80}
-presents the B model of the Z80 instruction set. %Section ~\ref{sec:Verification} provides some information on the proof effort needed to analyse the presented models.
-Section ~\ref{sec:studycase} explains the building and verifying assembly-level refinements. Related works are discussed in
-section~\ref{sec:relatedworks}. Finally, the last section is devoted to the conclusions.
+%
+%This work is focused on the presentation of modelling of assembly instruction
+%set, including elementary libraries to describe hardware aspects; and the
+%modelling and simulation of a case study of an embedded verified software that
+%can be used in tanks of the petroleum industry.
+
+This paper is structured as it follows.  Section~\ref{sec:models}
+presents the elementary libraries and the modeling of some elements
+common to microcontrollers.  Section~\ref{sec:z80} presents the B
+model of the Z80 instruction set.  Related works are discussed in
+section~\ref{sec:relatedworks}. Finally, the last section is devoted
+to the conclusions.
+
 %
@@
 
-First, bits are modelled as a set of integers: $\textit{BIT} =
+First, bits are modeled as a set of integers: $\textit{BIT} =
 \textit{0..1}$. The negation is an unary function on bits and it is
@@
 
-\subsection{Modelling bytes and bit vectors of length 16}
+\subsection{Modeling bytes and bit vectors of length 16}
 \label{subsec:HardwareLibrary3}
@@
 
-The $\textit{BYTE\_INDEX}$ is the domain of the functions modelling bytes. It starts at 1 to obey a
+The $\textit{BYTE\_INDEX}$ is the domain of the functions modeling bytes. It starts at 1 to obey a
 definition of sequences from B. However, it is common in hardware architectures to start indexing from
@@
 convention. The $\textit{BYTE}$ type is as a specialized type of $\textit{BIT\_VECTOR}$, but it has a size
-limit. Other specific definitions are provided to facilitate further modelling: the type $\textit{BV16}$
+limit. Other specific definitions are provided to facilitate further modeling: the type $\textit{BV16}$
 is created for bit vector of length 16 in a similar way.
@@
 Usually, each type module just needs to instantiate concepts that were already defined in the hardware
-modelling library.  For example, the function $\textit{bv\_to\_nat}$ from bit vector arithmetic is
+modeling library.  For example, the function $\textit{bv\_to\_nat}$ from bit vector arithmetic is
 specialized to $\textit{byte\_uchar}$. As the set $\textit{BYTE}$ is a subset of the
@@
 
+	\begin{center}
 	\begin{sloppypar}
 
-\bf MACHINE
-
-\hspace*{0.15in}\it Z80
-
-\bf INCLUDES
-
+	  \begin{tabbing}
+	    \begin{array}[t]{l}
+		\bf MACHINE\\
+		\hspace*{0.15in}\it Z80\\
+		\bf INCLUDES\\
 		\hspace*{0.10in}\it MEMORY
-
-\bf SEES
-
-\hspace*{0.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,
-
-\hspace*{0.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,
-
-\hspace*{0.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,
-
-\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION
+	    \end{array}
+	     \hspace*{1cm} 
+	    \begin{array}[t]{l}
+	    \bf SEES\\
+		\hspace*{0.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,\\
+		\hspace*{0.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,\\
+		\hspace*{0.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,\\
+		\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION\\
+	    \end{array}
+	  \end{tabbing}
+
 	 \end{sloppypar} 
-
-\subsection{Modelling registers}
+	\end{center}
+
+\subsection{Modeling registers}
 
@@
 
-\subsection{Modelling the actions and instructions}
+\subsection{Modeling the actions and instructions}
 
@@
 instructions of microcontrollers, a second category represents the input and output of data 
-(shown in ~\ref{sec:modellingIO}) and the last category represents the external
+(shown in ~\ref{sec:modelingIO}) and the last category represents the external
 actions (shown in~\ref{sec:externalactions}). A simple example of instruction category is
@@
 
-\subsection{Modelling the input/output instructions}
-\label{sec:modellingIO}
+\subsection{Modeling the input/output instructions}
+\label{sec:modelingIO}
 The Z80 has an extensive set of input and output (I/O) instructions and 256 ports for
@@
 
-\subsection{Modelling the external actions}
+\subsection{Modeling the external actions}
 \label{sec:externalactions}
@@
 The external actions change the state of the microcontroller, for example, refreshing the I/O ports
-and interruptions request. The external actions are also modelled by operations and they are named with the
+and interruptions request. The external actions are also modeled by operations and they are named with the
 prefix ``$ext\_$'' and followed by the name of action. There are four external
@@
 The interruptions and the \textit{reset} action can change the state of
-program counter. Theses actions are modelled by B operations and its main
+program counter. Theses actions are modeled by B operations and its main
 effects are shown below\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$ holds the value of stack pointer minus 2,
@@
 
-The modelling of instruction set, interruptions and input/output ports become the animation interesting.
+The modeling of instruction set, interruptions and input/output ports become the animation interesting.
 The animation in ProB version is not so fast and practical as a simulation, for example in \cite{Simulator_z80},
 because ProB works manipulating logic and math concepts that can have infinite sets and non determinism, then ProB requires more processing, specially when it works analysing  complex expressions or very large expressions.
-However, a user of ProB with this modelling can: execute manually a sequence of Z80's instructions that represent a program and view the visited states of ports, memory and registers, it is important to debug the effect action on the state space computed; create a tracing of the states; analyse constants, functions and expressions; and search possible violations.
+However, a user of ProB with this modeling can: execute manually a sequence of Z80's instructions that represent a program and view the visited states of ports, memory and registers, it is important to debug the effect action on the state space computed; create a tracing of the states; analyse constants, functions and expressions; and search possible violations.
 
@@
 object of the pilot project developed to analyse a petroleum production test
-system in each oil field. Its modelling was developed in according to \cite{LAUT_SERGIO}
-and its B assembly model can be executed manually in ProB. The modelling of
+system in each oil field. Its modeling was developed in according to \cite{LAUT_SERGIO}
+and its B assembly model can be executed manually in ProB. The modeling of
 instruction set and the B assembly model allow to simulate the execution of microcontroller
@@
 
-There are in the literature several
-approaches~\cite{BHDL_2003,DBLP:journals/entcs/EvansG08,DBLP:conf/fmco/Leuschel08,DBLP:conf/asm/Wright08}
-to model hardware and the virtual machines using B. In these works, the B
-has been used successfully to model the operational semantic. However the cost of modelling is still expensive.
-
-These recent works \cite{Marc20113,springerlink:Yuan2011} have a similar approach using Event B. 
-However, our work use the B method, that seems more appropriated to software development, because it has
-implementable language defined, called B0, and tools to convert the models to a programming language.
-
-The main motivation of our research is the development of verified
-software up to the assembly level, which requires specifying the
-semantics of the underlying hardware. Thus some aspects were not
-modelled in our work such as the execution time of the instructions.
-Also we did not consider the micro architecture of the hardware as the scope of
-our work does not include hardware verification.
+There are in the literature several similar approaches to model hardware and instruction set using B.
+The paper \cite{springerlink:Yuan2011} reports a method to specify, design and construct sound and complete instruction set model by stepwise refinement and formal proof using the formal method Event-B.
+It also discuss about desirable properties of instruction set model.
+However, our work \cite{Valerio_SBMF09} and \cite{Subotic2010} use the B method, that seems more appropriated to software development, because it has implementable language defined,
+called B0, and tools to convert the models to a programming language.
+In \cite{Marc20113} is present an related experience on using B in the design of secure micro-controllers.
+It tries to show the feasibility of such a technique for high confidence trustful devices.
+The work \cite{Subotic2010} describes a similar approach with MIPS CPU architecture and it shows how the CPU is formally specified and implemented using the B method.
+
+% 
+%There are in the literature several
+%approaches~\cite{BHDL_2003,DBLP:journals/entcs/EvansG08,DBLP:conf/fmco/Leuschel08,DBLP:conf/asm/Wright08}
+%to model hardware and the virtual machines using B. In these works, the B
+%has been used successfully to model the operational semantic. However the cost of modeling is still expensive. 
+%
+%These recent works \cite{Marc20113,springerlink:Yuan2011} have a similar approach using Event B. 
+%However, our work use the B method, that seems more appropriated to software development, because it has
+%implementable language defined, called B0, and tools to convert the models to a programming language.
+%
+%
+%The main motivation of our research is the development of verified
+%software up to the assembly level, which requires specifying the
+%semantics of the underlying hardware. Thus some aspects were not
+%modeled in our work such as the execution time of the instructions.
+%Also we did not consider the micro architecture of the hardware as the scope of
+%our work does not include hardware verification.
 
@@
 % Um pequeno resumo do trabalho
-This work has shown an approach to the formal modelling of the instruction set of
+This work has shown an approach to the formal modeling of the instruction set of
 microcontroller using the B method. During the construction of this model, some
@@
 is also envisioned. Finally, another important activity is develop a formal model of a platform used in LLVM Compiler~\cite{DBLP:conf/cgo/LattnerA04}.
+
+\subsubsection{Acknowledgements:} This experiment could not have succeeded without the interesting work of Michael Leuschel who developed several improvements in the ProB and provided important aid to enable the animation of the Z80 modeling.
+
+% This work was partially supported by INES (www.ines.org.br),
+%funded by CNPq grant 573964/2008-4 and by CNPq grants 553597/2008-6, 550946/2007-
+%1, and 620132/2008-6
 

### 2012-01-19 trunk/papers/ArticleZ80/source/paper.tex
Abstract updated, (Please, David look the new abstract).

@@
 
-  This paper describes an approach to model the functional aspects of
-  the instruction set of microcontrollers and several details about
-  the representation of microcontrollers.  Different models were
-  developed using the notation of the B method. This paper presents
-  specifically a B-based formal model of the Z80 instruction set.
-
-%  \paragraph[Keywords]{
-%   \textbf{Keywords}: Embedded Software, Simulation, B method and Verification%}
+The rigorous development of an embedded software in a microcontroller requires 
+a specification of its instruction set and the software requirements specification. 
+The actual industrial practice is to report the instruction set informally using natural language
+or with added semi-formal notation. This practice may leave space for errors and inconsistencies \cite{springerlink:Yuan2011}.
+This paper presents specifically a B formal model of the Z80 instruction set. 
+This model  can be  used on documentation, simulation, test and  a model-driven approach
+to design and develop software from the functional specification level down to assembly.
+%Using this model-driven approach and the instruction set model, a small case study to
+%analyse a petroleum production test system in oil field was developed.
+%  
+% \paragraph[Keywords]{ \textbf{Keywords}: Embedded Software, Simulation, B
+% method and Verification%}
 \end{abstract}
@@
 
-This work presents a formal modell of the instruction set of the Z80 
-microcontroller~\cite{Z80_manual}. This microcontroller
-was chosen because of several factors: it contains essential and common
-concepts in microcontrollers and until microprocessors; it has extensive
-documentation available; and it was widely used and is still in use. 
-A first Z80 modell presented in \cite{Valerio_SBMF09} was verified completely,
-but it uses constructs that are not well supported to animate. 
-This modelling was changed to support animation in ProB\cite{proB}.
-Basically, the new modelling\footnote{The interested reader in more
-details is invited to visit our repository at: http://code.google.com/p/b2asm/.}
-changed elements represented by infinite sets and adjusted the implications
-of this change. Using the responsibility division mechanisms provided by B, auxiliary
-libraries of basic modules were developed as part of the construction of microcontroller
-model. Such library has many definitions about common concepts used in
-the microcontrollers;besides the new Z80 model, a theoretic case study in 
-petroleum production test system was developed using Z80 assembly language
-to analyse the code simulation and verification process.
-
-The formal model of instruction set provides many benefits. The animation of 
-Z80 model allows to simulate the execution assembly programs including
-support for interrupts and input and output instructions. Other possible uses
-of a formal model of a microcontroller instruction set include documentation,
-the construction of simulators, and can be possibly the starting point of a 
-verification effort for the actual implementation of a Z80 design. Moreover, 
-the model of the instruction set could be instrumented with non-functional
-aspects, such as the number of cycles it takes to execute an instruction, to
-prove lower and upper bounds on the execution time of a routine.
-Two aspects of these formalization are of particular importance to us:
-provide a more solid documentation artifact for microcontrollers, 
-build a reference model for a formal compilation approach in B.
+This work presents a formal model of the instruction set of the Z80
+microcontroller~\cite{Z80_manual}. This microcontroller was chosen
+because of several factors: it contains essential and common concepts
+of microcontrollers and even microprocessors; it has extensive
+documentation available; and it has been widely used.  A first,
+partial, Z80 model presented in \cite{Valerio_SBMF09} was fully
+verified, but it uses constructs that make it impossible to animate.
+
+The new model supports animation in ProB\cite{proB}.  Essentially, the
+new model\footnote{The interested reader in more details is invited to
+  visit our repository at: http://code.google.com/p/b2asm/.}  no
+longer contains definitions derived from infinite sets. Using B
+constructs for modularity, auxiliary libraries of basic modules were
+developed as part of the construction of microcontroller model. Such
+library has many definitions of concepts common in microcontrollers.
+Besides the new Z80 model, a petroleum production test system was
+developed using Z80 assembly language to analyse the code simulation
+and verification process.
+
+Providing the formal model of an instruction set yields many
+benefits. The animation of Z80 model allows to simulate the execution
+assembly programs including support for interrupts and input and
+output instructions. Other possible uses of a formal model of a
+microcontroller instruction set include documentation, the
+construction of simulators, and can be possibly the starting point of
+a verification effort for the actual implementation of a Z80
+design. Moreover, the model of the instruction set could be
+instrumented with non-functional aspects, such as the number of cycles
+it takes to execute an instruction, to prove lower and upper bounds on
+the execution time of a routine.  Two aspects of these formalization
+are of particular importance to us: provide a more solid documentation
+artifact for microcontrollers, build a reference model for a formal
+compilation approach in B.
 
@@
 % Compact version
-In general, the manuals of microcontrollers can be more suited for
-developers in assembly language. Many times, each instruction of microcontrollers
-is shown in its official manual using textual description, math description, examples,
-encoding, number of cycles and other information. Some times, the instruction descriptions
-have an own notation, some informations are organized in different sections of document 
-and the textual description are not standardized. This does not allow the interpretation by parser.
-The manuals also have semi-formal and informal descriptions, these descriptions may make mistakes.
-The official manual of the microcontroller Z80 has also several problems that were identified
-over time~\cite{UndocumentedZ80}.
-
-	A solution to avoid ambiguity and inconsistencies is to specify formally the
-assembly instruction set, this creates a pattern of representation and validates
-properties of instructions. The formal model also restricts the definitions to
-correct typing and uses only well-defined expressions. Furthermore, the developer
-can use with B method different levels of abstraction and add more specific details
+
+In general, the manuals of microcontrollers can be more suited for \marginpar{DD: o título da sub-seção não condiz com seu conteúdo.
+Não cita nenhum problema na formalização.
+Não vejo muito bem o que ele traz de novo que já não foi dito anteriormente.}
+developers in assembly language. Many times, each instruction of
+microcontrollers is shown in its official manual using textual
+description, math description, examples, encoding, number of cycles
+and other information. Sometimes, the instruction descriptions have
+an own notation, some informations are organized in different sections
+of document and the textual description are not standardized. This
+does not allow the interpretation by parser.  The manuals also have
+semi-formal and informal descriptions, these descriptions may make
+mistakes.  The official manual of the microcontroller Z80 has also
+several problems that were identified over
+time~\cite{UndocumentedZ80}.
+
+A solution to avoid ambiguity and inconsistencies is to specify
+formally the assembly instruction set, this creates a pattern of
+representation and validates properties of instructions. The formal
+model also restricts the definitions to correct typing and uses only
+well-defined expressions. Furthermore, the developer can use with B
+method different levels of abstraction and add more specific details
 in refined models.
@@
 
-This paper is structured as it follows. %Section~\ref{sec:B_method} provides a short introduction to the B method. 
-Section~\ref{sec:models} presents the elementary libraries and
-the modelling of some elements common to microcontrollers.  Section~\ref{sec:z80}
-presents the B model of the Z80 instruction set. %Section ~\ref{sec:Verification} provides some information on the proof effort needed to analyse the presented models.
-%Section ~\ref{sec:studycase} explains the building and verifying assembly-level refinements. 
-Related works are discussed in section~\ref{sec:relatedworks}. Finally, the last section is devoted to the conclusions.
+This paper is structured as it follows.  Section~\ref{sec:models}
+presents the elementary libraries and the modeling of some elements
+common to microcontrollers.  Section~\ref{sec:z80} presents the B
+model of the Z80 instruction set.  Related works are discussed in
+section~\ref{sec:relatedworks}. Finally, the last section is devoted
+to the conclusions.
+
 %
@@
 
-\subsection{Modelling registers}
+\subsection{Modeling registers}
 
@@
 
-\hspace*{0.40in}\it zero \rm = \it is\_zero \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$
-
-\hspace*{0.40in}\it half\_carry \rm = \rm 0  $\land$
+\hspace*{0.40in}\it zero \rm = \it is\_zero \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$ \it half\_carry \rm = \rm 0  $\land$
 
 \hspace*{0.40in}\it pv \rm = \it parity\_even \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) $\land$
-
-\hspace*{0.40in}\it add\_sub \rm =\hspace*{0.10in}\rm 0  $\land$
-
-\hspace*{0.40in}\it carry \rm = \it z\_c
+\it add\_sub \rm =\hspace*{0.10in}\rm 0  $\land$ 
+\it carry \rm = \it z\_c
 
@@
 %to model hardware and the virtual machines using B. In these works, the B
-%has been used successfully to model the operational semantic. However the cost of modelling is still expensive. 
+%has been used successfully to model the operational semantic. However the cost of modeling is still expensive. 
 %
@@
 %semantics of the underlying hardware. Thus some aspects were not
-%modelled in our work such as the execution time of the instructions.
+%modeled in our work such as the execution time of the instructions.
 %Also we did not consider the micro architecture of the hardware as the scope of
@@
 model and techniques used in verification process.
-
 %Link com o parágrafo anterior 
@@
 
-\subsubsection{Acknowledgements:} This experiment could not have succeeded without the interesting work of Michael Leuschel who developed several improvements in the ProB and provided important aid to enable the animation of the Z80 modelling.
+\subsubsection{Acknowledgements:} This experiment could not have succeeded without the interesting work of Michael Leuschel who developed several improvements in the ProB and provided important aid to enable the animation of the Z80 modeling.
 

### 2012-03-15 trunk/papers/B2012-work/javacardAPIpaper.tex
inicial

\documentclass{entcs} 
\usepackage{entcsmacro}
\usepackage{graphicx}
\sloppy
% The following is enclosed to allow easy detection of differences in
% ascii coding.
% Upper-case    A B C D E F G H I J K L M N O P Q R S T U V W X Y Z
% Lower-case    a b c d e f g h i j k l m n o p q r s t u v w x y z
% Digits        0 1 2 3 4 5 6 7 8 9
% Exclamation   !           Double quote "          Hash (number) #
% Dollar        $           Percent      %          Ampersand     &
% Acute accent  '           Left paren   (          Right paren   )
% Asterisk      *           Plus         +          Comma         ,
% Minus         -           Point        .          Solidus       /
% Colon         :           Semicolon    ;          Less than     <
% Equals        =3D           Greater than >          Question mark ?
% At            @           Left bracket [          Backslash     \
% Right bracket ]           Circumflex   ^          Underscore    _
% Grave accent  `           Left brace   {          Vertical bar  |
% Right brace   }           Tilde        ~

% A couple of exemplary definitions:

\newcommand{\Nat}{{\mathbb N}}
\newcommand{\Real}{{\mathbb R}}

\def\lastname{Martins Moreira}

\begin{document}
\begin{frontmatter}
  \title{Modeling the Java Card API (or Interface?) in B} \author{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{Simone - Bruno - David \thanksref{coemail}}
  \address{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais?} \thanks[anamaria]{Email:
    \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
        anamaria@dimap.ufrn.br}}} \thanks[coemail]{Email:
    \href{mailto:couserid@codept.coinst.coedu} {\texttt{\normalshape
        couserid@codept.coinst.coedu}}}

% AMM: Precisa ser arrumado: The ENTCS macro package provides two alternatives to listing authors
% names and addresses.  These are described in detail in the file
% \texttt{instraut.dvi}. Basically, listing each author and his or her
% address in turn, is the simplest method.  But, if there are several
% authors and two or more share the same address (but not all authors
% are at this address), then the method of listing authors first, and
% then the addresses, and of referencing addresses to authors should be
% used.

\begin{abstract} 
  The B method development process starts from an abstract model of a
  set of services that are to be developed, and includes formally
  verified refinements, which introduce implementation details. At a
  final refinement step these implementation details depend on the
  language for which code is to be generated. To rigorously develop
  Java Card Applications with the B method, then, it is necessary to
  model in B the constructs of Java Card, so that the gap between the
  B formally verified implementation model and the Java Card generated
  code is as small as it can be. In the context of the BSmart project,
  we have developed the specification of the Java Card Application
  Programming Interface (Java Card API), to be used in the development
  of Java Card applications with B.
\end{abstract}
\begin{keyword}
  Java Card, smart cards, B method, formal methods
\end{keyword}
\end{frontmatter}
\section{Introduction}\label{intro}

In the context of the BSmart project, we are interested in the use of
B to develop Java Card~\cite{chen:2000} applications.  Java Card is a
restricted and optimized version of Java which allows memory and
processor constrained devices, such as smart cards, to store and run
small applications.  Java Card applications have characteristics
(small size and need for security) which make then good candidates for
formal development with the B method. They also present some
regularity of structure that induced our group into the definition of
a method and a tool which specialize B for development of such
applications, providing greater support for the developer than pure B
tools \cite{atelierB, proB}. They are the BSmart method and tool
presented in \cite{mainpapers}.

Previous works have also dealt with B for Java Card, \cite{BOM,
  JBTools e cia}, but none has, at our knowledge, specified Java Card
in B so that the formal development includes Java Card aspects. In the
first versions of BSmart \cite{}, these aspects were included by the code
ganerator, and in JBTools \cite{}, the generated code needed to be edited by
the developer to include these aspects.

These works lacked then the ability to formally verify the Java Card
dependent aspects of the application. To be able to change this, a B
specification of these aspects is needed, and this contribution is
what we present in this paper: the B specification of the Java Card
API, with...  This specification can be used with the BSmart tool,
independently, in any B tool, or even as a complement to the official
Java Card documentation \cite{}.

In the following...

\section{Smart Cards and Java Card}\label{sec:javacard}

AMM: TRECHOS EXTRAIDOS E JA UM POUCO ALTERADOS DO ABZ 2010. DEPENDENDO
DO QUE VAMOS MOSTRAR NA ESPECIFICAÇAO PODEMOS DEIXAR MAIS OU MENOS
DETALHES AQUI. NO FINAL, PRECISAMOS VERIFICAR QUE NAO COPIAMOS DEMAIS
O QUE ESTAVA LA.

A smart card application is distributed between on-card and off-card
components.  The server application on the card side (called applet)
provides the application services. The off-card client (called host
application) resides in a computer or electronic terminal.

The information exchange between the host and card applications is
made through a a half-duplexed low level protocol, named
\emph{Application Protocol Data Unit} (APDU).  The ISO 7816-4 standard
specifies two kinds of APDU's, which are the command and the response
APDU (Figure~\ref{fig:APDUs}). Both specify data packets. A command APDU
is sent by the host, requiring some applet service and a response APDU is
sent by the applet, responding to the host request with the result of the
service processing.

\begin{figure}[!ht]
\centering
%\includegraphics[scale=0.3]{figs/comandoAPDU.jpg}
%\includegraphics[scale=0.3]{figs/respostaAPDU.jpg}
\caption{Command and response APDUs. Source:~\cite{ortiz1intro:2003}}
\label{fig:APDUs}
\end{figure}

The main component of the Java Card platform (Figure~\ref{fig:jcre})
is its runtime environment (JCRE), composed of a Java Card Virtual
Machine (JCVM), an API, and, usually, system and
industry-specific classes~\cite{chen:2000}. The JCRE acts as a small
operating system, being responsible for the control of the application
lifetime, security and resource management.

\begin{figure}[h]
\centering
%\includegraphics[width=0.35\textwidth]{figs/jcplatform.png}
\caption{Java Card platform components. Adapted from:~\cite{ortiz1intro:2003}}
\label{fig:jcre}
\end{figure}

\paragraph{Java Card applets}\label{sec:applets}

A Java Card applet is a class that inherits the
\emph{javacard.framewok.Applet} class of the Java Card API and is
implemented upon the Java Card subset of Java.  During the applet
conversion for card installation, a verification phase is performed to
check conformance of the classes with Java Card restrictions.

The current usual Java Card specification (2.2.x) allows two kinds of
applets.  The older, and most commonly used, kind of applet manipulates
directly the APDU protocol packages while the newer one abstracts from
the lower level protocol using Remote Method Invocation (RMI). In this
paper we will call the lower level applet \emph{APDU applet} and the
higher level one, \emph{RMI applet}. RMI introduces a layer of
abstraction above the APDU protocol, and, due to this fact, it is usually
less efficient than APDU applets.

\section{B specification of the Java Card API}

\section{Using the model in BSmart}

Uma discussão a ser feita quanto a essa parte:

ainda não publicamos nada em revista relacionado ao BSmart - como já
tem muitas publicações pontuais, acho que uma possibilidade de
publicar em revista seria fazer algo muito completo: método,
ferramenta, host e cartão, kit, estudo(s) de caso, verificações de
correção... A dúvida então é: será que falar aqui do passport vai
prejudicar uma publicação mais importante? Vamos tentar essa
publicação mais importante ou não? Eu acho que podemos usar o passport aqui só para mostrar um exemplo concreto de uso da especificação da API, sem focar na parte da avaliação do BSmart. Concordam?

\section{Conclusions}

\bibliographystyle{plain}
\bibliography{bib}

\end{document}

### 2012-03-15 trunk/papers/B2012-work/javacardAPIpaper.tex
retirada da discussao

@@
 
-Uma discussão a ser feita quanto a essa parte:
-
-ainda não publicamos nada em revista relacionado ao BSmart - como já
-tem muitas publicações pontuais, acho que uma possibilidade de
-publicar em revista seria fazer algo muito completo: método,
-ferramenta, host e cartão, kit, estudo(s) de caso, verificações de
-correção... A dúvida então é: será que falar aqui do passport vai
-prejudicar uma publicação mais importante? Vamos tentar essa
-publicação mais importante ou não? Eu acho que podemos usar o passport aqui só para mostrar um exemplo concreto de uso da especificação da API, sem focar na parte da avaliação do BSmart. Concordam?
-
 \section{Conclusions}

### 2012-03-16 papers/B2012-work/javacardAPIpaper.tex

--- 2012-03-15 trunk/papers/B2012-work/javacardAPIpaper.tex
@@
 \begin{frontmatter}
-  \title{Modeling the Java Card API (or Interface?) in B} \author{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
-  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{Simone - Bruno - David \thanksref{coemail}}
-  \address{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais?} \thanks[anamaria]{Email:
+%  \title{Modeling the Java Card API (or Interface?) in B} \author{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
+%  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{S. de Oliveira Santos \thanksref{simone}}
+%  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{B. Emerson Gurgel Gomes \thanksref{bruno}}
+%    \address{IFRN, Currais Novos, RN, Brazil} \author{D. Boris Paul D\'{e}harbe \thanksref{david}}
+%      \address{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais?} \thanks[anamaria]{Email:
+%    \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
+%        anamaria@dimap.ufrn.br}}} 
+%      \thanks[simone]{Email:
+%    \href{mailto:simone82@ppgsc.ufrn.br} {\texttt{\normalshape
+%        simone82@ppgsc.ufrn.br}}}
+%      \thanks[bruno]{Email:
+%    \href{mailto:bruno.gurgel@ifrn.edu.br} {\texttt{\normalshape
+%        bruno.gurgel@ifrn.edu.br}}}
+%       \thanks[david]{Email:
+%    \href{mailto:david@dimap.ufrn.br} {\texttt{\normalshape
+%        david@dimap.ufrn.br}}}
+
+  \title{Modeling the Java Card API (or Interface?) in B} \author[ufrn]{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
+  \author[ufrn]{S. de Oliveira Santos \thanksref{simone}}
+  \author[ifrn]{B. Emerson Gurgel Gomes \thanksref{bruno}}
+  \author[ufrn]{D. Boris Paul D\'{e}harbe \thanksref{david}}
+  \address[ufrn]{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais? REUNI?}  \address[ifrn]{IFRN, Currais Novos, RN, Brazil} 
+ \thanks[anamaria]{Email:
     \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
-        anamaria@dimap.ufrn.br}}} \thanks[coemail]{Email:
-    \href{mailto:couserid@codept.coinst.coedu} {\texttt{\normalshape
-        couserid@codept.coinst.coedu}}}
+        anamaria@dimap.ufrn.br}}} 
+      \thanks[simone]{Email:
+    \href{mailto:simone82@ppgsc.ufrn.br} {\texttt{\normalshape
+        simone82@ppgsc.ufrn.br}}}
+      \thanks[bruno]{Email:
+    \href{mailto:bruno.gurgel@ifrn.edu.br} {\texttt{\normalshape
+        bruno.gurgel@ifrn.edu.br}}}
+       \thanks[david]{Email:
+    \href{mailto:david@dimap.ufrn.br} {\texttt{\normalshape
+        david@dimap.ufrn.br}}}
 
@@
 applications, providing greater support for the developer than pure B
-tools \cite{atelierB, proB}. They are the BSmart method and tool
-presented in \cite{mainpapers}.
+tools \cite{atelierB,proB}. They are the BSmart method and tool
+presented in \cite{Gomes10}.
+%colocar o de 2008 também
 
-Previous works have also dealt with B for Java Card, \cite{BOM,
-  JBTools e cia}, but none has, at our knowledge, specified Java Card
+Previous works have also dealt with B for Java Card, 
+\cite{BOM,JBtools,e cia}, but none has, at our knowledge, specified Java Card
 in B so that the formal development includes Java Card aspects. In the
-first versions of BSmart \cite{}, these aspects were included by the code
-ganerator, and in JBTools \cite{}, the generated code needed to be edited by
+first versions of BSmart \cite{Gomes10}, these aspects were included by the code
+ganerator, and in JBTools \cite{JBtools}, the generated code needed to be edited by
 the developer to include these aspects.
@@
 independently, in any B tool, or even as a complement to the official
-Java Card documentation \cite{}.
+Java Card documentation \cite{oracle}.
 

### 2012-03-16 papers/B2012-work/javacardAPIpaper.tex
mais um pouquinho

@@
 
+Among the different components of the Java Card platform, the Java Card API is the one without which no refinement into Java Card code could be done.
+Suppose, for instance, that we had the B abstract machine of fig \ref{maquina}, which models ...
+Of course, this is a very simple example, but it is enough to show the importance of having the model of the Java Card API modeled in B.
+As all data in it is implementable in C, for instance, and it is a deterministic specification, B tools can directly generate code for it in C, and the resulting code would be very similar to the original B machine.
+However,  Java Card code corresponding to it would look like the one presented in fig.
+\ref{impl java card}.
+In a rigorous development process, we don't want the code generator to carry out all these non-verified transformations in the code.
+So, we need to be able to refine the original B machine into something like the B refinement of fig.
+\ref{B ref usando a API}.
+Then, code generation would be straightforward, as it should be.
+This is then the reason why it needed to be the first Java Card component to be modeled in B.
+(AMM: maybe also cite related work ?
+WHich are the formalization works that did not focus on the API?)
+
+The Java Card API, in its version 2.2.2, is defined in 14 packages, containing 93 classes and interfaces, as shown in table \ref{incluir a tabela da dissert da simone}.
+Their interfaces have all been modeled in B, and the resulting models are available at LINK.
+Differently from  JML and OCL, though, B is not object oriented, and some modeling solutions had to be found to cope with this lack of resources.
+We present in the following the main characteristics and restrictions of our models:
+
+\begin{itemize}
+
+\item Only the abstract models corresponding to each class or interface have been developed.
+Two reasons led to this choice: (1) code is not to be generated from the specification of the API, as we are only creating an abstract layer to existing implementations; and (2)  in B, only the abstract machines are seen by modules that build upon others, through the SEES, USES, IMPORTS, INCLUDES and EXTENDS clauses.
+
+\item Conditions on method inputs which lead to exception raising in the Java Card platform specification are coded into pre-conditions in B.
+(em que casos isso não é feito - NULL? outros?)
+
+\item Method outputs are modeled in B as taking any value of the correct type (e em que casos temos um pouco mais).
+
+\end{itemize}
+
 \section{Using the model in BSmart}

### 2012-03-23 trunk/papers/B2012-work/javacardAPIpaper.tex
alguns pequenos ajustes e tarefas para Simone.

@@
 \begin{frontmatter}
-  \title{Modeling the Java Card API (or Interface?) in B} \author{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
-  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{Simone - Bruno - David \thanksref{coemail}}
-  \address{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais?} \thanks[anamaria]{Email:
+%  \title{Modeling the Java Card API (or Interface?) in B} \author{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
+%  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{S. de Oliveira Santos \thanksref{simone}}
+%  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{B. Emerson Gurgel Gomes \thanksref{bruno}}
+%    \address{IFRN, Currais Novos, RN, Brazil} \author{D. Boris Paul D\'{e}harbe \thanksref{david}}
+%      \address{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais?} \thanks[anamaria]{Email:
+%    \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
+%        anamaria@dimap.ufrn.br}}} 
+%      \thanks[simone]{Email:
+%    \href{mailto:simone82@ppgsc.ufrn.br} {\texttt{\normalshape
+%        simone82@ppgsc.ufrn.br}}}
+%      \thanks[bruno]{Email:
+%    \href{mailto:bruno.gurgel@ifrn.edu.br} {\texttt{\normalshape
+%        bruno.gurgel@ifrn.edu.br}}}
+%       \thanks[david]{Email:
+%    \href{mailto:david@dimap.ufrn.br} {\texttt{\normalshape
+%        david@dimap.ufrn.br}}}
+
+  \title{Modeling the Java Card API (or Interface?) in B} \author[ufrn]{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
+  \author[ufrn]{S. de Oliveira Santos \thanksref{simone}}
+  \author[ifrn]{B. Gurgel Gomes \thanksref{bruno}}
+  \author[ufrn]{D.  D\'{e}harbe \thanksref{david}}
+  \address[ufrn]{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais? REUNI?}  \address[ifrn]{IFRN, Currais Novos, RN, Brazil} 
+ \thanks[anamaria]{Email:
     \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
-        anamaria@dimap.ufrn.br}}} \thanks[coemail]{Email:
-    \href{mailto:couserid@codept.coinst.coedu} {\texttt{\normalshape
-        couserid@codept.coinst.coedu}}}
-
-% AMM: Precisa ser arrumado: The ENTCS macro package provides two alternatives to listing authors
-% names and addresses.  These are described in detail in the file
-% \texttt{instraut.dvi}. Basically, listing each author and his or her
-% address in turn, is the simplest method.  But, if there are several
-% authors and two or more share the same address (but not all authors
-% are at this address), then the method of listing authors first, and
-% then the addresses, and of referencing addresses to authors should be
-% used.
+        anamaria@dimap.ufrn.br}}} 
+      \thanks[simone]{Email:
+    \href{mailto:simone82@ppgsc.ufrn.br} {\texttt{\normalshape
+        simone82@ppgsc.ufrn.br}}}
+      \thanks[bruno]{Email:
+    \href{mailto:bruno.gurgel@ifrn.edu.br} {\texttt{\normalshape
+        bruno.gurgel@ifrn.edu.br}}}
+       \thanks[david]{Email:
+    \href{mailto:david@dimap.ufrn.br} {\texttt{\normalshape
+        david@dimap.ufrn.br}}}
 
@@
 applications, providing greater support for the developer than pure B
-tools \cite{atelierB, proB}. They are the BSmart method and tool
-presented in \cite{mainpapers}.
+tools \cite{atelierB,proB}. They are the BSmart method and tool,
+presented in \cite{BSmart08,Gomes10}.
 
-Previous works have also dealt with B for Java Card, \cite{BOM,
-  JBTools e cia}, but none has, at our knowledge, specified Java Card
+Previous works have also dealt with B for  Smart Cards, 
+\cite{Bert:2003,JBtools,tatibouet:lncs,requetBtoC}, but none has, at our knowledge, specified Java Card
 in B so that the formal development includes Java Card aspects. In the
-first versions of BSmart \cite{}, these aspects were included by the code
-ganerator, and in JBTools \cite{}, the generated code needed to be edited by
+first versions of BSmart \cite{Gomes10}, these aspects were included by the code
+ganerator, and in JBTools \cite{JBtools}, the generated code needed to be edited by
 the developer to include these aspects.
 
-These works lacked then the ability to formally verify the Java Card
-dependent aspects of the application. To be able to change this, a B
-specification of these aspects is needed, and this contribution is
-what we present in this paper: the B specification of the Java Card
-API, with...  This specification can be used with the BSmart tool,
-independently, in any B tool, or even as a complement to the official
-Java Card documentation \cite{}.
+These works lacked then the ability to formally verify the Java Card dependent aspects of the application. To be able to change this, a B
+specification of these aspects is needed, and this contribution is what we present in this paper: the B specification of the Java Card
+API, with its primitive types and all classes of the 14 packages which define the API. This specification can be used with the BSmart tool,
+independently, in any B tool, or even as a complement to the official Java Card documentation \cite{oracle}.
+
+SIMONE -> verifique que está certo o que falei acima sobre a especificação.
 
@@
 
+Among the different components of the Java Card platform, the Java Card API is the one without which no refinement into Java Card code could be done.
+Suppose, for instance, that we had the B abstract machine of fig \ref{maquina}, which models ...
+Of course, this is a very simple example, but it is enough to show the importance of having the model of the Java Card API modeled in B.
+As all data in it is implementable in C, for instance, and it is a deterministic specification, B tools can directly generate code for it in C, and the resulting code would be very similar to the original B machine.
+However,  Java Card code corresponding to it would look like the one presented in fig.
+\ref{impl java card}.
+In a rigorous development process, we don't want the code generator to carry out all these non-verified transformations in the code.
+So, we need to be able to refine the original B machine into something such as the B refinement of fig.
+\ref{B ref usando a API}.
+Then, code generation would be straightforward, as it should be.
+This is then the reason why it needed to be the first Java Card component to be modeled in B.
+(AMM: maybe also cite related work ?
+WHich are the formalization works that did not focus on the API?)
+
+SIMONE -> por favor proponha um bem pequeno exemplo para colocarmos aqui. A idéia é que mostre que ser B0 não basta para gerar o código Java Card.
+
+The Java Card API, in its version 2.2.2, is defined in 14 packages, containing 93 classes and interfaces, as shown in table \ref{incluir a tabela da dissert da simone}.
+Their interfaces have all been modeled in B, and the resulting models are available at LINK.
+Differently from  JML and OCL, though, B is not object oriented, and some modeling solutions had to be found to cope with this lack of resources.
+We present in the following the main characteristics and restrictions of our models:
+
+SIMONE-> coloque a tabela, por favor. o LINK pode ser substituído pela página do google code.
+
+\begin{itemize}
+
+\item Only the abstract models corresponding to each class or interface have been developed.
+Two reasons led to this choice: (1) code is not to be generated from the specification of the API, as we are only creating an abstract layer to existing implementations; and (2)  in B, only the abstract machines are seen by modules that build upon others, through the SEES, USES, IMPORTS, INCLUDES and EXTENDS clauses.
+
+\item Conditions on method inputs which lead to exception raising in the Java Card platform specification are coded into pre-conditions in B.
+(em que casos isso não é feito - NULL? outros?)
+
+\item Method outputs are modeled in B as taking any value of the correct type (e em que casos temos um pouco mais).
+
+\end{itemize}
+
 \section{Using the model in BSmart}

### 2012-03-26 papers/B2012-work/javacardAPIpaper.tex

@@
   \author[ufrn]{S. de Oliveira Santos \thanksref{simone}}
-  \author[ifrn]{B. Emerson Gurgel Gomes \thanksref{bruno}}
-  \author[ufrn]{D. Boris Paul D\'{e}harbe \thanksref{david}}
+  \author[ifrn]{B. Gurgel Gomes \thanksref{bruno}}
+  \author[ufrn]{D.  D\'{e}harbe \thanksref{david}}
   \address[ufrn]{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais? REUNI?}  \address[ifrn]{IFRN, Currais Novos, RN, Brazil} 
@@
         david@dimap.ufrn.br}}}
-
-% AMM: Precisa ser arrumado: The ENTCS macro package provides two alternatives to listing authors
-% names and addresses.  These are described in detail in the file
-% \texttt{instraut.dvi}. Basically, listing each author and his or her
-% address in turn, is the simplest method.  But, if there are several
-% authors and two or more share the same address (but not all authors
-% are at this address), then the method of listing authors first, and
-% then the addresses, and of referencing addresses to authors should be
-% used.
 
@@
 applications, providing greater support for the developer than pure B
-tools \cite{atelierB,proB}. They are the BSmart method and tool
-presented in \cite{Gomes10}.
-%colocar o de 2008 também
+tools \cite{atelierB,proB}. They are the BSmart method and tool,
+presented in \cite{BSmart08,Gomes10}.
 
-Previous works have also dealt with B for Java Card, 
-\cite{BOM,JBtools,e cia}, but none has, at our knowledge, specified Java Card
+Previous works have also dealt with B for Smart Cards, 
+\cite{Bert:2003,JBtools,tatibouet:lncs,requetBtoC}, but none has, at our knowledge, specified Java Card
 in B so that the formal development includes Java Card aspects. In the
 first versions of BSmart \cite{Gomes10}, these aspects were included by the code
-ganerator, and in JBTools \cite{JBtools}, the generated code needed to be edited by
+generator, and in JBTools \cite{JBtools}, the generated code needed to be edited by
 the developer to include these aspects.
 
-These works lacked then the ability to formally verify the Java Card
-dependent aspects of the application. To be able to change this, a B
-specification of these aspects is needed, and this contribution is
-what we present in this paper: the B specification of the Java Card
-API, with...  This specification can be used with the BSmart tool,
-independently, in any B tool, or even as a complement to the official
-Java Card documentation \cite{oracle}.
+These works lacked then the ability to formally verify the Java Card dependent aspects of the application. To be able to change this, a B
+specification of these aspects is needed, and this contribution is what we present in this paper: the B specification of the Java Card
+API, with its primitive types and all classes of the 14 packages which define the API. This specification can be used with the BSmart tool,
+independently, in any B tool, or even as a complement to the official Java Card documentation \cite{oracle}.
+
+SIMONE -> verifique que está certo o que falei acima sobre a especificação.
+COMENT SIMONE -> Acho que está muito bom, não tenho o que acrescentar.
 
@@
 Among the different components of the Java Card platform, the Java Card API is the one without which no refinement into Java Card code could be done.
-Suppose, for instance, that we had the B abstract machine of fig \ref{maquina}, which models ...
+Suppose, for instance, that we had the B abstract machine of Fig.
+\ref{fig:Bmachine}, which models a bus card with a variable that controls balance and a operation to add credits to the balance.
 Of course, this is a very simple example, but it is enough to show the importance of having the model of the Java Card API modeled in B.
 As all data in it is implementable in C, for instance, and it is a deterministic specification, B tools can directly generate code for it in C, and the resulting code would be very similar to the original B machine.
-However,  Java Card code corresponding to it would look like the one presented in fig.
-\ref{impl java card}.
+However,  Java Card code corresponding to it would look like the one presented in Fig.
+\ref{fig:jcapplet}.
 In a rigorous development process, we don't want the code generator to carry out all these non-verified transformations in the code.
-So, we need to be able to refine the original B machine into something like the B refinement of fig.
+So, we need to be able to refine the original B machine into something such as the B refinement of fig.
 \ref{B ref usando a API}.
@@
 
-The Java Card API, in its version 2.2.2, is defined in 14 packages, containing 93 classes and interfaces, as shown in table \ref{incluir a tabela da dissert da simone}.
+SIMONE -> por favor proponha um bem pequeno exemplo para colocarmos aqui. A idéia é que mostre que ser B0 não basta para gerar o código Java Card.
+COMENT SIMONE -> coloquei duas figuras, veja o que acha do exemplo, é o mesmo que foi usado na dissertação, ele é simples e pequeno.
+Coloquei a máquina abstrata e o modelo do applet Java Card (veja o que acha do tamanho das figuras), estou vendo com o Bruno como a implementação desta máquina fica com a API, porque o modelo dela que fizemos foi sem o uso da API, ela possui apenas o uso dos tipos primitivos e as operações de soma e comparação deles.
+O exemplo que nós usamos na validação do trabalho que tinha o uso da API foi um refinamento da máquina applet.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.8\textwidth]{figs/machineB.png}
+\caption{Example B machine.}
+\label{fig:Bmachine}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.8\textwidth]{figs/appletJC.png}
+\caption{Example Java Card applet.}
+\label{fig:jcapplet}
+\end{figure}
+
+The Java Card API, in its version 2.2.2, is defined in 14 packages, containing 93 classes and interfaces, as shown in Table \ref{tab:packages}.
 Their interfaces have all been modeled in B, and the resulting models are available at LINK.
@@
 We present in the following the main characteristics and restrictions of our models:
+
+\begin{table}[h]
+\begin{center}
+\caption{Packages of the Java Card API (2.2.2)}
+\begin{tabular}{|l|c||l|c|}
+ \hline
+\textbf{Package} & \textbf{classes/} & \textbf{Package} & \textbf{classes/} \\ 
+				 & \textbf{package}  &					& \textbf{package} \\ \hline \hline
+java.io & 1 & javacardx.biometry & 5 \\ \hline
+java.lang & 12 & javacardx.crypto & 2 \\ \hline
+java.rmi & 2 & javacardx.external & 3 \\ \hline
+javacard.framework & 19 & javacardx.framework.math & 3 \\ \hline
+javacard.framework.service & 8 & javacardx.framework.tlv & 7 \\ \hline
+javacard.security & 27 & javacardx.framework.util & 2 \\ \hline
+javacardx.apdu & 1 & javacardx.framework.util.intx & 1 \\
+ \hline
+\end{tabular}
+\label{tab:packages}
+\end{center}
+\end{table}
+
+SIMONE-> coloque a tabela, por favor. o LINK pode ser substituído pela página do google code.
+COMENT SIMONE -> A tabela está no formato que estava lá na dissertação (muito grande?), tem que ver se tem algum modelo padrão a ser seguido para as tabelas, se tiver é so me passar que eu faço os ajustes.
+Sobre o link do google code, eu não sei qual o link público lá.
 

### 2012-03-26 trunk/papers/B2012-work/javacardAPIpaper.tex

@@
 small applications.  Java Card applications have characteristics
-(small size and need for security) which make then good candidates for
+(small size and need for security) which make them good candidates for
 formal development with the B method. They also present some
@@
 first versions of BSmart \cite{Gomes10}, these aspects were included by the code
-ganerator, and in JBTools \cite{JBtools}, the generated code needed to be edited by
+generator, and in JBTools \cite{JBtools}, the generated code needed to be edited by
 the developer to include these aspects.
@@
 independently, in any B tool, or even as a complement to the official Java Card documentation \cite{oracle}.
-
-SIMONE -> verifique que está certo o que falei acima sobre a especificação.
 
@@
 Among the different components of the Java Card platform, the Java Card API is the one without which no refinement into Java Card code could be done.
-Suppose, for instance, that we had the B abstract machine of fig \ref{maquina}, which models ...
+Suppose, for instance, that we had the B abstract machine of Fig.
+\ref{fig:Bmachine}, which models a bus card with a variable that controls balance and a operation to add credits to the balance.
 Of course, this is a very simple example, but it is enough to show the importance of having the model of the Java Card API modeled in B.
 As all data in it is implementable in C, for instance, and it is a deterministic specification, B tools can directly generate code for it in C, and the resulting code would be very similar to the original B machine.
-However,  Java Card code corresponding to it would look like the one presented in fig.
-\ref{impl java card}.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.8\textwidth]{figs/machineB.png}
+\caption{Example B machine.}
+\label{fig:Bmachine}
+\end{figure}
+
+However,  Java Card code corresponding to it would look like the one presented in Fig.
+\ref{fig:jcapplet}.
 In a rigorous development process, we don't want the code generator to carry out all these non-verified transformations in the code.
@@
 
-SIMONE -> por favor proponha um bem pequeno exemplo para colocarmos aqui. A idéia é que mostre que ser B0 não basta para gerar o código Java Card.
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.8\textwidth]{figs/appletJC.png}
+\caption{Example Java Card applet.}
+\label{fig:jcapplet}
+\end{figure}
 
-The Java Card API, in its version 2.2.2, is defined in 14 packages, containing 93 classes and interfaces, as shown in table \ref{incluir a tabela da dissert da simone}.
-Their interfaces have all been modeled in B, and the resulting models are available at LINK.
+COMENT SIMONE -> coloquei duas figuras, veja o que acha do exemplo, é o mesmo que foi usado na dissertação, ele é simples e pequeno.
+Coloquei a máquina abstrata e o modelo do applet Java Card (veja o que acha do tamanho das figuras), estou vendo com o Bruno como a implementação desta máquina fica com a API, porque o modelo dela que fizemos foi sem o uso da API, ela possui apenas o uso dos tipos primitivos e as operações de soma e comparação deles.
+O exemplo que nós usamos na validação do trabalho que tinha o uso da API foi um refinamento da máquina applet.
+
+PARA SIMONE -> Por favor, coloque os exemplos sob a forma de listing. Fica mais fácil de editar, controlar o tamanho e conteúdo e fica mais legível.
+Quanto ao código javacard, basta o da operação addcredit. Aliás, o nome não está batendo. Não verifiquei o resto. Por favor verifique.
+
+The Java Card API, in its version 2.2.2, is defined in 14 packages, containing 93 classes and interfaces, as shown in Table \ref{tab:packages}.
+Their interfaces have all been modeled in B, and the resulting models are available at the KitSmart project page\footnote{https://code.google.com/p/kitsmart/}.
 Differently from  JML and OCL, though, B is not object oriented, and some modeling solutions had to be found to cope with this lack of resources.
@@
 
-SIMONE-> coloque a tabela, por favor. o LINK pode ser substituído pela página do google code.
+\begin{table}[h]
+\begin{center}
+\caption{Packages of the Java Card API (2.2.2) and the number of classes in each package.}
+\begin{tabular}{|l|c||l|c|}
+ \hline
+
+\textbf{\small{Package}} & \textbf{\small{classes}} & \textbf{\small{Package}} & \textbf{\small{classes}}\\ \hline \hline
+\small{java.io} & \small{1} & \small{javacardx.biometry} & \small{5} \\ \hline
+\small{java.lang} & \small{12} & \small{javacardx.crypto }& \small{2} \\ \hline
+\small{java.rmi} & \small{2} & \small{javacardx.external} & \small{3} \\ \hline
+\small{javacard.framework} & \small{19} & \small{javacardx.framework.math} & \small{3} \\ \hline
+\small{javacard.framework.service} & \small{8} & \small{javacardx.framework.tlv} & \small{7} \\ \hline
+\small{javacard.security} & \small{27} & \small{javacardx.framework.util} & \small{2} \\ \hline
+\small{javacardx.apdu} & \small{1} & \small{javacardx.framework.util.intx} & \small{1} \\
+
+ \hline
+\end{tabular}
+\label{tab:packages}
+\end{center}
+\end{table}
 

### 2012-03-27 trunk/papers/B2012-work/javacardAPIpaper.tex

@@
 
-  \title{Modeling the Java Card API (or Interface?) in B} \author[ufrn]{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
+  \title{Modeling the Java Card API (or Interface?) in B}
+  \author[ufrn]{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
   \author[ufrn]{S. de Oliveira Santos \thanksref{simone}}
-  \author[ifrn]{B. Gurgel Gomes \thanksref{bruno}}
-  \author[ufrn]{D.  D\'{e}harbe \thanksref{david}}
-  \address[ufrn]{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais? REUNI?}  \address[ifrn]{IFRN, Currais Novos, RN, Brazil} 
- \thanks[anamaria]{Email:
-    \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
-        anamaria@dimap.ufrn.br}}} 
-      \thanks[simone]{Email:
-    \href{mailto:simone82@ppgsc.ufrn.br} {\texttt{\normalshape
-        simone82@ppgsc.ufrn.br}}}
-      \thanks[bruno]{Email:
-    \href{mailto:bruno.gurgel@ifrn.edu.br} {\texttt{\normalshape
-        bruno.gurgel@ifrn.edu.br}}}
-       \thanks[david]{Email:
-    \href{mailto:david@dimap.ufrn.br} {\texttt{\normalshape
-        david@dimap.ufrn.br}}}
+  \author[ifrn]{B. Gurgel Gomes \thanksref{bruno}} \author[ufrn]{D.
+    D\'{e}harbe \thanksref{david}} \address[ufrn]{DIMAp-UFRN, Natal,
+    RN, Brazil} \thanks[ALL]{This work is partly supported by CAPES
+    and CNPq grants 306033/2009-7, 560014/2010-4 and 573964/2008-4
+    (National Institute of Science and Technology for Software
+    Engineering---INES, \texttt{www.ines.org.br}).}
+\address[ifrn]{IFRN, Currais Novos, RN, Brazil}
+\thanks[anamaria]{Email: \href{mailto:anamaria@dimap.ufrn.br}
+  {\texttt{\normalshape anamaria@dimap.ufrn.br}}}
+\thanks[simone]{Email: \href{mailto:simone82@ppgsc.ufrn.br}
+  {\texttt{\normalshape simone82@ppgsc.ufrn.br}}}
+\thanks[bruno]{Email: \href{mailto:bruno.gurgel@ifrn.edu.br}
+  {\texttt{\normalshape bruno.gurgel@ifrn.edu.br}}}
+\thanks[david]{Email: \href{mailto:david@dimap.ufrn.br}
+  {\texttt{\normalshape david@dimap.ufrn.br}}}
 
@@
   The B method development process starts from an abstract model of a
-  set of services that are to be developed, and includes formally
+  set of services that are to be developed, and proceeds with formally
   verified refinements, which introduce implementation details. At a
@@
   B formally verified implementation model and the Java Card generated
-  code is as small as it can be. In the context of the BSmart project,
+  code be as narrow as possible. In the context of the BSmart project,
   we have developed the specification of the Java Card Application
@@
 formal development with the B method. They also present some
-regularity of structure that induced our group into the definition of
-a method and a tool which specialize B for development of such
+regularity of structure that led our group to define a method and
+implement a tool which specialize B for development of such
 applications, providing greater support for the developer than pure B
@@
 Previous works have also dealt with B for Smart Cards,
-\cite{Bert:2003,JBtools,tatibouet:lncs,requetBtoC}, but none has, at our knowledge, specified Java Card
-in B so that the formal development includes Java Card aspects. In the
-first versions of BSmart \cite{Gomes10}, these aspects were included by the code
-generator, and in JBTools \cite{JBtools}, the generated code needed to be edited by
+\cite{Bert:2003,JBtools,tatibouet:lncs,requetBtoC}, but none has, at
+our knowledge, specified Java Card in B so that the formal development
+includes Java Card aspects. In the first versions of BSmart
+\cite{Gomes10}, these aspects were included by the code generator, and
+in JBTools \cite{JBtools}, the generated code needed to be edited by
 the developer to include these aspects.
 
-These works lacked then the ability to formally verify the Java Card dependent aspects of the application. To be able to change this, a B
-specification of these aspects is needed, and this contribution is what we present in this paper: the B specification of the Java Card
-API, with its primitive types and all classes of the 14 packages which define the API. This specification can be used with the BSmart tool,
-independently, in any B tool, or even as a complement to the official Java Card documentation \cite{oracle}.
+These works lacked then the ability to formally verify the Java Card
+dependent aspects of the application. To be able to change this, a B
+specification of these aspects is needed, and this is the contribution
+of this paper: the B specification of the Java Card API, with its
+primitive types and all the classes of the 14 packages composing the
+API. This specification can be used with the BSmart tool, 
+independently, in any B tool, or even as a complement to the official
+Java Card documentation \cite{oracle}.
 
@@
 
-Among the different components of the Java Card platform, the Java Card API is the one without which no refinement into Java Card code could be done.
-Suppose, for instance, that we had the B abstract machine of Fig.
-\ref{fig:Bmachine}, which models a bus card with a variable that controls balance and a operation to add credits to the balance.
-Of course, this is a very simple example, but it is enough to show the importance of having the model of the Java Card API modeled in B.
-As all data in it is implementable in C, for instance, and it is a deterministic specification, B tools can directly generate code for it in C, and the resulting code would be very similar to the original B machine.
+Among the different components of the Java Card platform, the Java
+Card API is the one without which no refinement into Java Card code
+could be done.  Suppose, for instance, that we had the B abstract
+machine of Fig.~\ref{fig:Bmachine}, which models a bus card with a
+variable that controls balance and a operation to add credits to the
+balance. This is simple example provides enough evidence of the
+importance of having a B model of the Java Card API.  As all data in
+it is implementable in C, for instance, and it is a deterministic
+specification, B tools can directly generate code for it in C, and the
+resulting code would be very similar to the original B machine.
 
@@
 
-However,  Java Card code corresponding to it would look like the one presented in Fig.
-\ref{fig:jcapplet}.
-In a rigorous development process, we don't want the code generator to carry out all these non-verified transformations in the code.
-So, we need to be able to refine the original B machine into something such as the B refinement of fig.
-\ref{B ref usando a API}.
-Then, code generation would be straightforward, as it should be.
-This is then the reason why it needed to be the first Java Card component to be modeled in B.
-(AMM: maybe also cite related work ?
-WHich are the formalization works that did not focus on the API?)
+However, Java Card code corresponding to it would look like the one
+presented in Fig.~\ref{fig:jcapplet}. In a rigorous development
+process, we do not want the code generator to carry out all these
+non-verified transformations in the code. So, we need to be able to
+refine the original B machine into something such as the B refinement
+of Fig.~\ref{B ref usando a API}. Then, code generation would be
+straightforward (as it should be). This is then the reason why it
+needed to be the first Java Card component to be modeled in B. (AMM:
+maybe also cite related work ? Which are the formalization works that
+did not focus on the API?)
 
@@
 
-PARA SIMONE -> Por favor, coloque os exemplos sob a forma de listing. Fica mais fácil de editar, controlar o tamanho e conteúdo e fica mais legível.
-Quanto ao código javacard, basta o da operação addcredit. Aliás, o nome não está batendo. Não verifiquei o resto. Por favor verifique.
+[[[ PARA SIMONE -> Por favor, coloque os exemplos sob a forma de listing. Fica mais fácil de editar, controlar o tamanho e conteúdo e fica mais legível.
+Quanto ao código javacard, basta o da operação addcredit. Aliás, o nome não está batendo. Não verifiquei o resto. Por favor verifique. ]]]
 
-The Java Card API, in its version 2.2.2, is defined in 14 packages, containing 93 classes and interfaces, as shown in Table \ref{tab:packages}.
-Their interfaces have all been modeled in B, and the resulting models are available at the KitSmart project page\footnote{https://code.google.com/p/kitsmart/}.
-Differently from  JML and OCL, though, B is not object oriented, and some modeling solutions had to be found to cope with this lack of resources.
-We present in the following the main characteristics and restrictions of our models:
+The Java Card API, in its version 2.2.2, is composed of 14 packages,
+containing 93 classes and interfaces, as shown in
+Table~\ref{tab:packages}. Their interfaces have all been modeled in B,
+and the resulting models are available at the KitSmart project
+page\footnote{https://code.google.com/p/kitsmart/}. Differently from
+JML and OCL, though, B is not object oriented, and some modeling
+solutions had to be found to cope with this lack of resources.  We
+present in the following the main characteristics and restrictions of
+our models:
 
@@
 
-\item Only the abstract models corresponding to each class or interface have been developed.
-Two reasons led to this choice: (1) code is not to be generated from the specification of the API, as we are only creating an abstract layer to existing implementations; and (2)  in B, only the abstract machines are seen by modules that build upon others, through the SEES, USES, IMPORTS, INCLUDES and EXTENDS clauses.
+\item Only the abstract models corresponding to each class or
+  interface have been developed. Two reasons led to this choice: (1)
+  code is not to be generated from the specification of the API, as we
+  are only creating an abstract layer to existing implementations; and
+  (2) in B, only the abstract machines are seen by modules that build
+  upon others, through the SEES, USES, IMPORTS, INCLUDES and EXTENDS
+  clauses.
 
-\item Conditions on method inputs which lead to exception raising in the Java Card platform specification are coded into pre-conditions in B.
-(em que casos isso não é feito - NULL? outros?)
+\item Conditions on method inputs which lead to exception raising in
+  the Java Card platform specification are coded into pre-conditions
+  in B.
 
-\item Method outputs are modeled in B as taking any value of the correct type (e em que casos temos um pouco mais).
+  [[[ AMM (em que casos isso não é feito - NULL? outros?) ]]]
+
+\item Method outputs are modeled in B as taking any value of the
+  correct type
+
+  [[[ AMM (e em que casos temos um pouco mais). ]]]
 

### 2012-03-27 trunk/papers/B2012-work/javacardAPIpaper.tex
some text on exceptions

@@
 machine of Fig.~\ref{fig:Bmachine}, which models a bus card with a
-variable that controls balance and a operation to add credits to the
-balance. This is simple example provides enough evidence of the
+variable that keeps track of the current credit the user has (\texttt{balance}) and an operation to add credits to this
+balance (\texttt{add\_credit}!!! verificar). This simple example provides enough evidence of the
 importance of having a B model of the Java Card API.  As all data in
@@
 
-[[[ PARA SIMONE -> Por favor, coloque os exemplos sob a forma de listing. Fica mais fácil de editar, controlar o tamanho e conteúdo e fica mais legível.
-Quanto ao código javacard, basta o da operação addcredit. Aliás, o nome não está batendo. Não verifiquei o resto. Por favor verifique. ]]]
+[[[ PARA SIMONE -> Por favor, coloque os exemplos sob a forma de listing. Fica mais facil de editar, controlar o tamanho e conteudo e fica mais legivel.
+Quanto ao codigo javacard, basta o da operacao addcredit. Alias, o nome não esta batendo. Nao verifiquei o resto. Por favor verifique. ]]]
+
+[[[ PARA SIMONE -> tambem precisamos com urgencia daquela anelise sobre o que muda da 2.2.2 para a 3 ]]]
 
@@
 
+As previously stated,  we chose to model exception raising conditions as operation pre-conditions in the B models.
+For instance, if the Java Card specification states that a method should throw an exception in case a parameter array is empty, the pre-condition of the corresponding B operation will require that this parameter is not empty.
+The B method then requires that every call to that operation satisfies the operation's pre-condition.
+As a consequence, code that is formally generated using the B method and the Java Card API specification should never raise this specific exception, as it is statically proved that all calls to this operation are made with non-empty parameter arrays.
+Not all situations leading to exceptional conditions are modeled, however, as some of them depend on the Java Card Runtime Environment (JCRE) and its state, which has not yet been modeled, and some require a more detailed specification of the contents of some data and other internal behavior.
+
 \section{Using the model in BSmart}
 
-\section{Conclusions}
+\section{Conclusions and future work}
+
+With the current level of abstraction of the model and the fact that the Java Card Runtime Environment (JCRE) has not been modeled, some
+conditions which may lead to exception raising are not yet modeled.
+Being able to model them would make developed applications even more sure, as more static verification could be carried out.
+Continuation of this work then will enlarge the scope of the modeling to cope statically with a larger set of situations.
 

### 2012-03-27 papers/B2012-work/javacardAPIpaper.tex

--- 2012-03-16 papers/B2012-work/javacardAPIpaper.tex
@@
 \usepackage{graphicx}
+
+\usepackage{listings}
+% Define B sintaxe
+\lstdefinelanguage{B}  {
+keywords={MACHINE, REFINEMENT, REFINES, IMPLEMENTATION, SEES, IMPORTS, SETS, INCLUDES, EXTENDS, CONSTANTS, CONCRETE_CONSTANTS, CONCRETE_VARIABLES, PROPERTIES, DEFINITIONS, VARIABLES, ABSTRACT_VARIABLES, INITIALISATION, INVARIANT, VALUES, OPERATIONS, BEGIN, PRE, THEN, END, NAT, NAT1, BOOL, MAXINT, IF, ELSE, ELSIF, ANY, WHERE, CHOICE, OR, EITHER, SELECT, VAR, IN},
+sensitive=true,% 
+alsoletter={\$},% 
+comment=[l]{\#},% 
+string=[b]",% 
+string=[b]'% 
+}
+
+% Set listing styles
+\lstset{ %
+%language=Pseudo,                % the language of the code
+basicstyle=\footnotesize\ttfamily,            % the size of the fonts that are used for the code
+extendedchars=true,
+numbers=left,                   % where to put the line-numbers
+numberstyle=\tiny,      % the size of the fonts that are used for the line-numbers 
+%stepnumber=2,                   % the step between two line-numbers. If it's 1, each line will be numbered
+%numbersep=5pt,                  % how far the line-numbers are from the code
+%backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
+%showspaces=false,               % show spaces adding particular underscores
+%showstringspaces=false,         % underline spaces within strings
+%showtabs=false,                 % show tabs within strings adding particular underscores
+frame=b,   %trBL,                   % adds a frame around the code
+stringstyle=\color{white}\ttfamily,
+%linewidth=0.8\textwidth, % width of the line
+tabsize=2,                      % sets default tabsize to 2 spaces
+columns=fixed,
+basewidth=.6em,
+xleftmargin=20pt,
+xrightmargin=5pt,
+%captionpos=b,                   % sets the caption-position to bottom
+breaklines=true,                % sets automatic line breaking
+showspaces=false,           % Leerzeichen anzeigen ?
+showtabs=false,             % Tabs anzeigen ?
+%backgroundcolor=\color{lightgray},
+keywordstyle=\color{red},
+showstringspaces=false
+%breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
+%title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
+%escapeinside={\%*}{*)},         % if you want to add a comment within your code
+%keywords={}            % if you want to add more keywords to the set
+}
+
 \sloppy
@@
 
-  \title{Modeling the Java Card API (or Interface?) in B} \author[ufrn]{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
+  \title{Modeling the Java Card API (or Interface?) in B}
+  \author[ufrn]{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
   \author[ufrn]{S. de Oliveira Santos \thanksref{simone}}
-  \author[ifrn]{B. Emerson Gurgel Gomes \thanksref{bruno}}
-  \author[ufrn]{D. Boris Paul D\'{e}harbe \thanksref{david}}
-  \address[ufrn]{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais? REUNI?}  \address[ifrn]{IFRN, Currais Novos, RN, Brazil} 
- \thanks[anamaria]{Email:
-    \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
-        anamaria@dimap.ufrn.br}}} 
-      \thanks[simone]{Email:
-    \href{mailto:simone82@ppgsc.ufrn.br} {\texttt{\normalshape
-        simone82@ppgsc.ufrn.br}}}
-      \thanks[bruno]{Email:
-    \href{mailto:bruno.gurgel@ifrn.edu.br} {\texttt{\normalshape
-        bruno.gurgel@ifrn.edu.br}}}
-       \thanks[david]{Email:
-    \href{mailto:david@dimap.ufrn.br} {\texttt{\normalshape
-        david@dimap.ufrn.br}}}
-
-% AMM: Precisa ser arrumado: The ENTCS macro package provides two alternatives to listing authors
-% names and addresses.  These are described in detail in the file
-% \texttt{instraut.dvi}. Basically, listing each author and his or her
-% address in turn, is the simplest method.  But, if there are several
-% authors and two or more share the same address (but not all authors
-% are at this address), then the method of listing authors first, and
-% then the addresses, and of referencing addresses to authors should be
-% used.
+  \author[ifrn]{B. Gurgel Gomes \thanksref{bruno}} \author[ufrn]{D.
+    D\'{e}harbe \thanksref{david}} \address[ufrn]{DIMAp-UFRN, Natal,
+    RN, Brazil} \thanks[ALL]{This work is partly supported by CAPES
+    and CNPq grants 306033/2009-7, 560014/2010-4 and 573964/2008-4
+    (National Institute of Science and Technology for Software
+    Engineering---INES, \texttt{www.ines.org.br}).}
+\address[ifrn]{IFRN, Currais Novos, RN, Brazil}
+\thanks[anamaria]{Email: \href{mailto:anamaria@dimap.ufrn.br}
+  {\texttt{\normalshape anamaria@dimap.ufrn.br}}}
+\thanks[simone]{Email: \href{mailto:simone82@ppgsc.ufrn.br}
+  {\texttt{\normalshape simone82@ppgsc.ufrn.br}}}
+\thanks[bruno]{Email: \href{mailto:bruno.gurgel@ifrn.edu.br}
+  {\texttt{\normalshape bruno.gurgel@ifrn.edu.br}}}
+\thanks[david]{Email: \href{mailto:david@dimap.ufrn.br}
+  {\texttt{\normalshape david@dimap.ufrn.br}}}
 
@@
 small applications.  Java Card applications have characteristics
-(small size and need for security) which make then good candidates for
+(small size and need for security) which make them good candidates for
 formal development with the B method. They also present some
-regularity of structure that induced our group into the definition of
-a method and a tool which specialize B for development of such
+regularity of structure that led our group to define a method and
+implement a tool which specialize B for development of such
 applications, providing greater support for the developer than pure B
-tools \cite{atelierB,proB}. They are the BSmart method and tool
-presented in \cite{Gomes10}.
-%colocar o de 2008 também
+tools \cite{atelierB,proB}. They are the BSmart method and tool,
+presented in \cite{BSmart08,Gomes10}.
 
-Previous works have also dealt with B for Java Card, 
-\cite{BOM,JBtools,e cia}, but none has, at our knowledge, specified Java Card
-in B so that the formal development includes Java Card aspects. In the
-first versions of BSmart \cite{Gomes10}, these aspects were included by the code
-ganerator, and in JBTools \cite{JBtools}, the generated code needed to be edited by
+Previous works have also dealt with B for Smart Cards,
+\cite{Bert:2003,JBtools,tatibouet:lncs,requetBtoC}, but none has, at
+our knowledge, specified Java Card in B so that the formal development
+includes Java Card aspects. In the first versions of BSmart
+\cite{Gomes10}, these aspects were included by the code generator, and
+in JBTools \cite{JBtools}, the generated code needed to be edited by
 the developer to include these aspects.
@@
 dependent aspects of the application. To be able to change this, a B
-specification of these aspects is needed, and this contribution is
-what we present in this paper: the B specification of the Java Card
-API, with...  This specification can be used with the BSmart tool,
+specification of these aspects is needed, and this is the contribution
+of this paper: the B specification of the Java Card API, with its
+primitive types and all the classes of the 14 packages composing the
+API. This specification can be used with the BSmart tool, 
 independently, in any B tool, or even as a complement to the official
@@
 
-Among the different components of the Java Card platform, the Java Card API is the one without which no refinement into Java Card code could be done.
-Suppose, for instance, that we had the B abstract machine of fig \ref{maquina}, which models ...
-Of course, this is a very simple example, but it is enough to show the importance of having the model of the Java Card API modeled in B.
-As all data in it is implementable in C, for instance, and it is a deterministic specification, B tools can directly generate code for it in C, and the resulting code would be very similar to the original B machine.
-However,  Java Card code corresponding to it would look like the one presented in fig.
-\ref{impl java card}.
-In a rigorous development process, we don't want the code generator to carry out all these non-verified transformations in the code.
-So, we need to be able to refine the original B machine into something like the B refinement of fig.
-\ref{B ref usando a API}.
-Then, code generation would be straightforward, as it should be.
-This is then the reason why it needed to be the first Java Card component to be modeled in B.
-(AMM: maybe also cite related work ?
-WHich are the formalization works that did not focus on the API?)
+Among the different components of the Java Card platform, the Java
+Card API is the one without which no refinement into Java Card code
+could be done.  Suppose, for instance, that we had the B abstract
+machine of Fig.~\ref{lis:buscard}, which models a bus card with a
+variable that keeps track of the current credit the user has (\texttt{balance}) and an operation to add credits to this
+balance (\texttt{add\_credit}!!! verificar). This simple example provides enough evidence of the
+importance of having a B model of the Java Card API.  As all data in
+it is implementable in C, for instance, and it is a deterministic
+specification, B tools can directly generate code for it in C, and the
+resulting code would be very similar to the original B machine.
 
-The Java Card API, in its version 2.2.2, is defined in 14 packages, containing 93 classes and interfaces, as shown in table \ref{incluir a tabela da dissert da simone}.
-Their interfaces have all been modeled in B, and the resulting models are available at LINK.
-Differently from  JML and OCL, though, B is not object oriented, and some modeling solutions had to be found to cope with this lack of resources.
-We present in the following the main characteristics and restrictions of our models:
+\lstset{language=B, caption={B machine Buscard},label=lis:buscard}
+\begin{lstlisting}
+MACHINE
+	Buscard
+CONSTANTS
+	maximum
+PROPERTIES
+	maximum : INTEGER & maximum = 100
+VARIABLES
+	balance
+INVARIANT
+	balance : INTEGER & balance <= maximum
+INITIALISATION
+	balance := 0
+OPERATIONS
+	add(credit) =
+	PRE
+		credit : INTEGER & credit > 0  & balance + credit <= maximum
+	THEN
+		balance := balance + credit
+	END
+END
+\end{lstlisting}
+
+However, Java Card code corresponding to it would look like the one
+presented in Fig.~\ref{lis:add}. In a rigorous development
+process, we do not want the code generator to carry out all these
+non-verified transformations in the code. So, we need to be able to
+refine the original B machine into something such as the B refinement
+of Fig.~\ref{B ref usando a API}. Then, code generation would be
+straightforward (as it should be). This is then the reason why it
+needed to be the first Java Card component to be modeled in B. (AMM:
+maybe also cite related work ? Which are the formalization works that
+did not focus on the API?)
+
+\lstset{language=Java, caption={Method add in applet Java Card},label=lis:add}
+\begin{lstlisting}
+public void add(APDU apdu) {
+	byte[] buffer = apdu.getBuffer();
+	byte credit = (byte) Util.makeShort((byte) buffer[ISO7816.OFFSET_P1], (byte)buffer[ISO7816.OFFSET_P2]);
+	if (!(credit >= 0)) {
+		ISOException.throwIt(EXCEPTIONS.negative_credit);
+	}
+	else if (!(balance + credit <= MAXIMUM)) {
+		ISOException.throwIt(EXCEPTIONS.balance_exceeded);
+	}
+	else {
+		balance = (balance + credit);
+	}
+}
+\end{lstlisting}
+
+[[[ PARA SIMONE -> tambem precisamos com urgencia daquela anelise sobre o que muda da 2.2.2 para a 3 ]]]
+
+The Java Card API, in its version 2.2.2, is composed of 14 packages,
+containing 93 classes and interfaces, as shown in
+Table~\ref{tab:packages}. Their interfaces have all been modeled in B,
+and the resulting models are available at the KitSmart project
+page\footnote{https://code.google.com/p/kitsmart/}. Differently from
+JML and OCL, though, B is not object oriented, and some modeling
+solutions had to be found to cope with this lack of resources.  We
+present in the following the main characteristics and restrictions of
+our models:
+
+\begin{table}[h]
+\begin{center}
+\caption{Packages of the Java Card API (2.2.2) and the number of classes in each package.}
+\begin{tabular}{|l|c||l|c|}
+ \hline
+
+\textbf{\small{Package}} & \textbf{\small{classes}} & \textbf{\small{Package}} & \textbf{\small{classes}}\\ \hline \hline
+\small{java.io} & \small{1} & \small{javacardx.biometry} & \small{5} \\ \hline
+\small{java.lang} & \small{12} & \small{javacardx.crypto }& \small{2} \\ \hline
+\small{java.rmi} & \small{2} & \small{javacardx.external} & \small{3} \\ \hline
+\small{javacard.framework} & \small{19} & \small{javacardx.framework.math} & \small{3} \\ \hline
+\small{javacard.framework.service} & \small{8} & \small{javacardx.framework.tlv} & \small{7} \\ \hline
+\small{javacard.security} & \small{27} & \small{javacardx.framework.util} & \small{2} \\ \hline
+\small{javacardx.apdu} & \small{1} & \small{javacardx.framework.util.intx} & \small{1} \\
+
+ \hline
+\end{tabular}
+\label{tab:packages}
+\end{center}
+\end{table}
 

### 2012-03-28 trunk/papers/B2012-work/javacardAPIpaper.tex
um pouco mais sobre a modelagem.

@@
 could be done.  Suppose, for instance, that we had the B abstract
-machine of Fig.~\ref{fig:Bmachine}, which models a bus card with a
+machine of Fig.~\ref{lis:buscard}, which models a bus card with a
 variable that keeps track of the current credit the user has (\texttt{balance}) and an operation to add credits to this
-balance (\texttt{add\_credit}!!! verificar). This simple example provides enough evidence of the
+balance (\texttt{add}). This simple example provides enough evidence of the
 importance of having a B model of the Java Card API.  As all data in
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{figs/machineB.png}
-\caption{Example B machine.}
-\label{fig:Bmachine}
-\end{figure}
+\lstset{language=B, caption={B machine Buscard},label=lis:buscard}
+\begin{lstlisting}
+MACHINE
+	Buscard
+CONSTANTS
+	maximum
+PROPERTIES
+	maximum : INTEGER & maximum = 100
+VARIABLES
+	balance
+INVARIANT
+	balance : INTEGER & balance <= maximum
+INITIALISATION
+	balance := 0
+OPERATIONS
+	add(credit) =
+	PRE
+		credit : INTEGER & credit > 0  & balance + credit <= maximum
+	THEN
+		balance := balance + credit
+	END
+END
+\end{lstlisting}
 
-However, Java Card code corresponding to it would look like the one
-presented in Fig.~\ref{fig:jcapplet}. In a rigorous development
+However, Java Card code corresponding to the operation \texttt{add} of this B specification would look like the one
+presented in Fig.~\ref{lis:addjavacard}. In a rigorous development
 process, we do not want the code generator to carry out all these
@@
 refine the original B machine into something such as the B refinement
-of Fig.~\ref{B ref usando a API}. Then, code generation would be
-straightforward (as it should be). This is then the reason why it
-needed to be the first Java Card component to be modeled in B. (AMM:
-maybe also cite related work ? Which are the formalization works that
-did not focus on the API?)
+of Fig.~\ref{lis:addimplementation}. Then, code generation would be
+straightforward (as it should be).
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{figs/appletJC.png}
-\caption{Example Java Card applet.}
-\label{fig:jcapplet}
-\end{figure}
+\lstset{language=Java, caption={Method add in applet Java Card},label=lis:addjavacard}
+\begin{lstlisting}
+public void add(APDU apdu) {
+	byte[] buffer = apdu.getBuffer();
+	byte credit = (byte) Util.makeShort((byte) buffer[ISO7816.OFFSET_P1], (byte)buffer[ISO7816.OFFSET_P2]);
+	if (!(credit >= 0)) {
+		ISOException.throwIt(EXCEPTIONS.negative_credit);
+	}
+	else if (!(balance + credit <= MAXIMUM)) {
+		ISOException.throwIt(EXCEPTIONS.balance_exceeded);
+	}
+	else {
+		balance = (balance + credit);
+	}
+}
+\end{lstlisting}
 
-COMENT SIMONE -> coloquei duas figuras, veja o que acha do exemplo, é o mesmo que foi usado na dissertação, ele é simples e pequeno.
-Coloquei a máquina abstrata e o modelo do applet Java Card (veja o que acha do tamanho das figuras), estou vendo com o Bruno como a implementação desta máquina fica com a API, porque o modelo dela que fizemos foi sem o uso da API, ela possui apenas o uso dos tipos primitivos e as operações de soma e comparação deles.
-O exemplo que nós usamos na validação do trabalho que tinha o uso da API foi um refinamento da máquina applet.
+\lstset{language=B, caption={B machine Buscard},label=lis:addimplementation}
+\begin{lstlisting}
+	add(apdu) =
+	BEGIN
+		buffer <- - getBuffer();
+		credit <-- cast ( );
+		IF 
+			not(credit >= 0)      !!! byte version
+		THEN
+			ISOException.throwIt(EXCEPTIONS.negative_credit)
+		ELSIF 
+			not(balance + credit <= MAXIMUM)     !!! byte version
+		THEN
+			ISOException.throwIt(EXCEPTIONS.balance_exceeded)
+		ELSE
+		        balance := sum_byte (balance, credit)  
+		END
+	END
 
-[[[ PARA SIMONE -> Por favor, coloque os exemplos sob a forma de listing. Fica mais facil de editar, controlar o tamanho e conteudo e fica mais legivel.
-Quanto ao codigo javacard, basta o da operacao addcredit. Alias, o nome não esta batendo. Nao verifiquei o resto. Por favor verifique. ]]]
+	[[[[BRUNO e SIMONE: isso esta todo errado - coloquei aqui mais ou menos o que eu acho que deve ser a cara da implementacao B que gera o codigo java card.
+Voces devem ter algum exemplo para recuperar a estrutura certa, nao?]]]]
 
-[[[ PARA SIMONE -> tambem precisamos com urgencia daquela anelise sobre o que muda da 2.2.2 para a 3 ]]]
+\end{lstlisting}
+
+We can identify two main kinds of transformation that had to be carried out in order to obtain the code  of operation \texttt{add} in figs.
+\ref{lis:addjavacard} and \ref{lis:addimplementation} from the abstract specification in fig.
+\ref{lis:buscard}:
+
+\begin{enumerate}
+\item the transition into a defensive style of programming by the inclusion of explicit tests to check that the input conditions (pre conditions in the  B specification) are satisfied before executing the corresponding action;
+
+\item the inclusion of Java Card specific aspects, such as the change of interface of \texttt{add} due to the need of passing data via the APDU protocol and the manipulation of the APDU buffer by methods offered the Java Card API.
+
+\end{enumerate}
+
+The transition into a defensive style of programming is part of the BSmart method, and implemented by the BSmart Full Function component.
+This component generates a B refinement of the original specification where pre-conditions are transformed into conditionals (IF-THEN-ELSE-END constructs), with the indication of the need to throw the corresponding exception when conditions are not met by input data and states.
+The resulting B model can then be formally verified as usual in the B method.
+
+The change of interface is also taken care of by the BSmart method, so that the development can be done without need to extend the B method.
+More details on this [[[????
+(we haven't published this part but there is the technical report )]]].
+
+But all of these transformations rely, at the bottom, on the need to reference Java Card API types, classes and methods on the B level.
+This is then the reason why the Java Card API needed to be the first Java Card component to be modeled in B.
+
+[[[(PARA SIMONE e BRUNO: maybe also cite related work ? Which are the formalization works that did not focus on the API?)]]]
+
+[[[ PARA SIMONE -> tambem precisamos com urgencia daquela analise sobre o que muda da 2.2.2 para a 3 ]]]
 
@@
 
+\item Each machine corresponding to a Java Card class defines an abstract set with the same name (distinction on character cases, only).
+For instance, the machine Throwable declares the abstract set THROWABLE.
+This set can then be used to reference the type corresponding to the class.
+
+\item The hierarchy of the Java Card API objects is partially modeled by abstract set inclusion.
+For instance, the class Throwable extends the class Object, in Java Card.
+In our model, we also have that the machine Throwable EXTENDS the machine Object, but the meaning of a machine extension in B is different from the meaning of a class extension in Java Card.
+To show that all throwables are indeed objects, a property stating that  THROWABLE is a subset of OBJECT is included in the Throwable machine.
+
 \end{itemize}
@@
 Not all situations leading to exceptional conditions are modeled, however, as some of them depend on the Java Card Runtime Environment (JCRE) and its state, which has not yet been modeled, and some require a more detailed specification of the contents of some data and other internal behavior.
+
+[[[ incluir aqui parte da discussao sobre as opcoes de modelagem  com o argumento pela escolha dos subsets para a tipagem ]]] 
+As our focus was on the interfaces of the Java Card API methods,
 

### 2012-03-29 trunk/papers/B2012-work/javacardAPIpaper.tex

@@
 
-\lstset{language=B, caption={B machine Buscard},label=lis:addimplementation}
+\lstset{language=B, caption={B implementation},label=lis:addimplementation}
 \begin{lstlisting}
-	add(apdu) =
-	BEGIN
-		buffer <- - getBuffer();
-		credit <-- cast ( );
-		IF 
-			not(credit >= 0)      !!! byte version
+add (apdu) =
+	VAR 
+		bf, cr,  apdu_state, res, curr_bal
+	IN
+		bf          <-- apdu.getBuffer;
+		apdu_state  <-- apdu.getCurrentState;
+		curr_bal <-- bal.get_num;
+
+		IF (apdu_state = STATE_INITIAL) 
 		THEN
-			ISOException.throwIt(EXCEPTIONS.negative_credit)
-		ELSIF 
-			not(balance + credit <= MAXIMUM)     !!! byte version
+			res <-- apdu.setIncomingAndReceive
+		ELSIF (OFFSET_CDATA >= 0 &
+			sum_short(OFFSET_CDATA, 2) <= size(bf) & 
+			bf /= [])
 		THEN 
-			ISOException.throwIt(EXCEPTIONS.balance_exceeded)
-		ELSE
-		        balance := sum_byte (balance, credit)  
+			cr  <-- util.getShort(bf, OFFSET_CDATA);
+
+			IF sum_short (curr_bal, cr) <= MAX_BALANCE & 
+				sum_short (curr_bal, cr) >= 0 
+			THEN 
+				bal.add (cr)
 			END
 		END
+END;
+\end{lstlisting}
 
-	[[[[BRUNO e SIMONE: isso esta todo errado - coloquei aqui mais ou menos o que eu acho que deve ser a cara da implementacao B que gera o codigo java card.
+%\lstset{language=B, caption={B machine Buscard},label=lis:addimplementation}
+%\begin{lstlisting}
+%	add(apdu) =
+%	BEGIN
+%		buffer <- - getBuffer();
+%		credit <-- cast ( );
+%		IF 
+%			not(credit >= 0)      !!! byte version
+%		THEN
+%			ISOException.throwIt(EXCEPTIONS.negative_credit)
+%		ELSIF 
+%			not(balance + credit <= MAXIMUM)     !!! byte version
+%		THEN
+%			ISOException.throwIt(EXCEPTIONS.balance_exceeded)
+%		ELSE
+%		        balance := sum_byte (balance, credit)  
+%		END
+%	END
+%	
+%	[[[[BRUNO e SIMONE: isso esta todo errado - coloquei aqui mais ou menos o que eu acho que deve ser a cara da implementacao B que gera o codigo java card.
 Voces devem ter algum exemplo para recuperar a estrutura certa, nao?]]]]
-
-\end{lstlisting}
+%	
+%\end{lstlisting}
 

### 2012-03-29 trunk/papers/B2012-work/javacardAPIpaper.tex
incluido texto sobre overloading

@@
 
-%\lstset{language=B, caption={B machine Buscard},label=lis:addimplementation}
-%\begin{lstlisting}
-%	add(apdu) =
-%	BEGIN
-%		buffer <- - getBuffer();
-%		credit <-- cast ( );
-%		IF 
-%			not(credit >= 0)      !!! byte version
-%		THEN
-%			ISOException.throwIt(EXCEPTIONS.negative_credit)
-%		ELSIF 
-%			not(balance + credit <= MAXIMUM)     !!! byte version
-%		THEN
-%			ISOException.throwIt(EXCEPTIONS.balance_exceeded)
-%		ELSE
-%		        balance := sum_byte (balance, credit)  
-%		END
-%	END
-%	
-%	[[[[BRUNO e SIMONE: isso esta todo errado - coloquei aqui mais ou menos o que eu acho que deve ser a cara da implementacao B que gera o codigo java card.
-Voces devem ter algum exemplo para recuperar a estrutura certa, nao?]]]]
-%	
-%\end{lstlisting}
-
 We can identify two main kinds of transformation that had to be carried out in order to obtain the code  of operation \texttt{add} in figs.
@@
 The transition into a defensive style of programming is part of the BSmart method, and implemented by the BSmart Full Function component.
-This component generates a B refinement of the original specification where pre-conditions are transformed into conditionals (IF-THEN-ELSE-END constructs), with the indication of the need to throw the corresponding exception when conditions are not met by input data and states.
+This component generates a B refinement of the original specification where pre-conditions are transformed into conditionals (IF-THEN-ELSE-END constructs), with the indication of the need to throw the corresponding exception when the specified conditions are not met by input data and states.
 The resulting B model can then be formally verified as usual in the B method.
@@
 
-But all of these transformations rely, at the bottom, on the need to reference Java Card API types, classes and methods on the B level.
+But all of these transformations rely on the need to reference Java Card API types, classes and methods on the B level.
 This is then the reason why the Java Card API needed to be the first Java Card component to be modeled in B.
@@
 
-The Java Card API, in its version 2.2.2, is composed of 14 packages,
-containing 93 classes and interfaces, as shown in
-Table~\ref{tab:packages}. Their interfaces have all been modeled in B,
-and the resulting models are available at the KitSmart project
-page\footnote{https://code.google.com/p/kitsmart/}. Differently from
-JML and OCL, though, B is not object oriented, and some modeling
-solutions had to be found to cope with this lack of resources.  We
-present in the following the main characteristics and restrictions of
-our models:
+The Java Card API, in its version 2.2.2\footnote{Java Card 2.2.2, although not the most recent version of the Java Card platform, is still widely used, and can easily be extended to Java Card 3 Classic Edition,  a backward compatible evolution of version 2.2.2., including some bug fixes, clarifications against the Java Card v2.2.2 specifications and new security algorithms (Java Card 3 Platform Release Notes, available at http://java.sun.com/javacard/3.0.1/specs.jsp).}, is composed of 14 packages, containing 93 classes and interfaces, as shown in
+Table~\ref{tab:packages}. Their interfaces have all been modeled in B, and the resulting models are available at the KitSmart project
+page\footnote{https://code.google.com/p/kitsmart/}. 
+Differently from JML and OCL, though, B is not object oriented, and some modeling solutions had to be found to cope with this lack of resources.
+We present in the following the main characteristics and some restrictions of our models:
 
@@
 
-\item Conditions on method inputs which lead to exception raising in
-  the Java Card platform specification are coded into pre-conditions
+\item Conditions on method inputs which may lead to exception raising in the Java Card platform specification are coded into pre-conditions
   in B.
 
-  [[[ AMM (em que casos isso não é feito - NULL? outros?) ]]]
-
-\item Method outputs are modeled in B as taking any value of the
-  correct type
-
-  [[[ AMM (e em que casos temos um pouco mais). ]]]
+ \item Method outputs are modeled in B as taking any value of the correct type, or, when possible in the current abstraction level, any value of a more restricted set of specified return values.
 
@@
 
+\item Overloading is not allowed in B and occurs frequently in Java Card.
+To model overloaded methods in B an extension to overloaded method names was included.
+We then have equals, equals\_1, equals\_2, etc.
 \end{itemize}
@@
 Not all situations leading to exceptional conditions are modeled, however, as some of them depend on the Java Card Runtime Environment (JCRE) and its state, which has not yet been modeled, and some require a more detailed specification of the contents of some data and other internal behavior.
+In these cases, the  exception raising conditions are only left as comments to the B specification, and can be used as additional documentation or to be considered in future work.
+
+One delicate issue to be dealt with both by the user of the API specification and the tools designed to work with it, is the use of overloaded methods in an application development.
+When overloaded operations are used in an application development, the user must explicitly choose which version of the overloaded operation is to be used.
+Then, verifications will be carried out considering this specific version of the operation.
+When code is generated, either the code generator will remove the name extension automatically, if designed for it, or the user will have to edit the generated code manually.
+In any case, it is important that the actual method called in the Java Card code is the one with which verifications have been carried out.
+In case of overriding, in particular, it will the user's responsibility to choose in B the one which is on scope at this specific point in Java, unless a extra verification step is included in the code generator that checks that the operation used in B is the one in scope in Java and registers an error in case it is not.
 

### 2012-03-29 trunk/papers/B2012-work/javacardAPIpaper.tex
codigos atualizados

@@
 PROPERTIES
-	maximum : INTEGER & maximum = 100
+	maximum : INTEGER & MAX_BALANCE  = 100
 VARIABLES
@@
 INVARIANT
-	balance : INTEGER & balance <= maximum
+	balance : INTEGER & balance <= MAX_BALANCE 
 INITIALISATION
@@
 	PRE
-		credit : INTEGER & credit > 0  & balance + credit <= maximum
+		credit : INTEGER & credit > 0  & balance + credit <= MAX_BALANCE 
 	THEN
@@
 \begin{lstlisting}
-public void add(APDU apdu) {
-	byte[] buffer = apdu.getBuffer();
-	byte credit = (byte) Util.makeShort((byte) buffer[ISO7816.OFFSET_P1], (byte)buffer[ISO7816.OFFSET_P2]);
-	if (!(credit >= 0)) {
-		ISOException.throwIt(EXCEPTIONS.negative_credit);
+public void add (APDU apdu) { 
+	byte[] bf;  short cr; byte apdu_state; short res;
+
+	bf = apdu.getBuffer();
+	apdu_state = apdu.getCurrentState();
+
+	if (apdu_state == STATE_INITIAL) {
+		res = apdu.setIncomingAndReceive();
+	} else if (ISO7816.OFFSET_CDATA >= 0 && 
+			(short) (ISO7816.OFFSET_CDATA + 2) <= bf.length && 
+			bf != null) {
+
+		cr = Util.getShort (bf, ISO7816.OFFSET_CDATA);
+
+		if (!((short) (balance + cr) <= MAX_BALANCE)) {
+			ISOException.throwIt(EX_BALANCE_EXCEEDED);
+		} else if (!((short) (balance + cr) >= 0)) {
+			ISOException.throwIt(EX_NEGATIVE_BALANCE);
+		} else if (! (cr > 0)) {
+			ISOException.throwIt(EX_NEGATIVE_CREDIT);
+		} else {
+			balance = (short) (balance + cr);
 		}
-	else if (!(balance + credit <= MAXIMUM)) {
-		ISOException.throwIt(EXCEPTIONS.balance_exceeded);
-	}
-	else {
-		balance = (balance + credit);
 	} 
@@
 VAR 
-		bf, cr,  apdu_state, res, curr_bal
+	bf, cr, apdu_state, res
 IN
@@
     	apdu_state  <-- apdu.getCurrentState;
-		curr_bal <-- bal.get_num;
 
@@
 	THEN 
-			cr  <-- util.getShort(bf, OFFSET_CDATA);
+		cr  <-- Util.getShort(bf, OFFSET_CDATA);
 
-			IF sum_short (curr_bal, cr) <= MAX_BALANCE & 
-				sum_short (curr_bal, cr) >= 0 
+		IF not (sum_short (balance, cr) <= MAX_BALANCE)
 		THEN 
-				bal.add (cr)
+			ISOException.throwIt(EX_BALANCE_EXCEEDED)
+		ELSIF not (sum_short (balance, cr) >= 0)
+		THEN ISOException.throwIt(EX_NEGATIVE_BALANCE)
+		ELSIF not (cr > 0)
+		THEN ISOException.throwIt(EX_NEGATIVE_CREDIT) 
+		ELSE  
+			balance := sum_short(balance, cr)
 		END
 	END
-END;
+END
 \end{lstlisting}

### 2012-03-29 trunk/papers/B2012-work/javacardAPIpaper.tex

@@
 conversion for card installation, a verification phase is performed to
-check conformance of the classes with Java Card restrictions.
+check conformance of the classes to Java Card restrictions.
 
 The current usual Java Card specification (2.2.x) allows two kinds of
-applets.  The older, and most commonly used, kind of applet manipulates
-directly the APDU protocol packages while the newer one abstracts from
-the lower level protocol using Remote Method Invocation (RMI). In this
-paper we will call the lower level applet \emph{APDU applet} and the
-higher level one, \emph{RMI applet}. RMI introduces a layer of
-abstraction above the APDU protocol, and, due to this fact, it is usually
+applets.  The older, and most commonly used, kind of applet
+manipulates directly the APDU protocol packages while the newer one
+abstracts from the lower level protocol using Remote Method Invocation
+(RMI). In this paper we will call the lower level applet \emph{APDU
+  applet} and the higher level one, \emph{RMI applet}. RMI introduces
+a layer of abstraction above the APDU protocol, and hence is usually
 less efficient than APDU applets.
@@
 machine of Fig.~\ref{lis:buscard}, which models a bus card with a
-variable that keeps track of the current credit the user has (\texttt{balance}) and an operation to add credits to this
-balance (\texttt{add}). This simple example provides enough evidence of the
+variable that keeps track of the current credit the user has
+(\texttt{balance}) and an operation to add credits to this balance
+(\texttt{add}). This simple example provides enough evidence of the
 importance of having a B model of the Java Card API.  As all data in
-it is implementable in C, for instance, and it is a deterministic
+it is implementable in C, for instance, and as it is a deterministic
 specification, B tools can directly generate code for it in C, and the
@@
 	PRE
-		credit : INTEGER & credit > 0  & balance + credit <= MAX_BALANCE 
+		credit : INTEGER & credit > 0  & 
+                balance + credit <= MAX_BALANCE 
 	THEN
@@
 
-However, Java Card code corresponding to the operation \texttt{add} of this B specification would look like the one
-presented in Fig.~\ref{lis:addjavacard}. In a rigorous development
-process, we do not want the code generator to carry out all these
-non-verified transformations in the code. So, we need to be able to
-refine the original B machine into something such as the B refinement
-of Fig.~\ref{lis:addimplementation}. Then, code generation would be
+However, Java Card code corresponding to the operation \texttt{add} of
+this B specification would look like the one presented in
+Fig.~\ref{lis:addjavacard}. In a rigorous development process, we do
+not want the code generator to carry out all these non-verified
+transformations in the code. So, we need to be able to refine the
+original B machine into something such as the B refinement of
+Fig.~\ref{lis:addimplementation}. Then, code generation would be
 straightforward (as it should be).
@@
 
-We can identify two main kinds of transformation that had to be carried out in order to obtain the code  of operation \texttt{add} in figs.
-\ref{lis:addjavacard} and \ref{lis:addimplementation} from the abstract specification in fig.
-\ref{lis:buscard}:
+We can identify two main kinds of transformation that have to be
+carried out in order to obtain the code of operation \texttt{add} in
+figs.~\ref{lis:addjavacard} and~\ref{lis:addimplementation} from the
+abstract specification in fig.~\ref{lis:buscard}:
 
 \begin{enumerate}
-\item the transition into a defensive style of programming by the inclusion of explicit tests to check that the input conditions (pre conditions in the  B specification) are satisfied before executing the corresponding action;
+\item the transition into a defensive style of programming by the
+  inclusion of explicit tests to check that the input conditions (pre
+  conditions in the B specification) are satisfied before executing
+  the corresponding action;
 
-\item the inclusion of Java Card specific aspects, such as the change of interface of \texttt{add} due to the need of passing data via the APDU protocol and the manipulation of the APDU buffer by methods offered the Java Card API.
+\item the inclusion of Java Card specific aspects, such as changing
+  the signature of \texttt{add} so that data is transmitted via the
+  APDU protocol and the APDU buffer is manipulated by methods offered
+  the Java Card API.
 
@@
 
-The transition into a defensive style of programming is part of the BSmart method, and implemented by the BSmart Full Function component.
-This component generates a B refinement of the original specification where pre-conditions are transformed into conditionals (IF-THEN-ELSE-END constructs), with the indication of the need to throw the corresponding exception when the specified conditions are not met by input data and states.
-The resulting B model can then be formally verified as usual in the B method.
+The transition to a defensive style of programming is part of the
+BSmart method, and implemented by the BSmart Full Function
+component. This component generates a B refinement of the original
+specification where pre-conditions are transformed into conditionals
+(IF-THEN-ELSE-END constructs), with the indication of the need to
+throw the corresponding exception when the specified conditions are
+not met by input data and states.  The resulting B model can then be
+formally verified as usual in the B method.
 
-The change of interface is also taken care of by the BSmart method, so that the development can be done without need to extend the B method.
-More details on this [[[????
-(we haven't published this part but there is the technical report )]]].
+The change of interface is also taken care of by the BSmart method, so
+that the development can be done without need to extend the B
+method, as detailed in~\cite{Gomes10}.
 
-But all of these transformations rely on the need to reference Java Card API types, classes and methods on the B level.
-This is then the reason why the Java Card API needed to be the first Java Card component to be modeled in B.
+But all of these transformations rely on the need to reference Java
+Card API types, classes and methods on the B level.  This is then the
+reason why the Java Card API needed to be the first Java Card
+component to be modeled in B.
 
@@
 
-The Java Card API, in its version 2.2.2\footnote{Java Card 2.2.2, although not the most recent version of the Java Card platform, is still widely used, and can easily be extended to Java Card 3 Classic Edition,  a backward compatible evolution of version 2.2.2., including some bug fixes, clarifications against the Java Card v2.2.2 specifications and new security algorithms (Java Card 3 Platform Release Notes, available at http://java.sun.com/javacard/3.0.1/specs.jsp).}, is composed of 14 packages, containing 93 classes and interfaces, as shown in
-Table~\ref{tab:packages}. Their interfaces have all been modeled in B, and the resulting models are available at the KitSmart project
-page\footnote{https://code.google.com/p/kitsmart/}. 
-Differently from JML and OCL, though, B is not object oriented, and some modeling solutions had to be found to cope with this lack of resources.
-We present in the following the main characteristics and some restrictions of our models:
+The Java Card API, in its version 2.2.2\footnote{Java Card 2.2.2,
+  although not the most recent version of the Java Card platform, is
+  still widely used, and can easily be extended to Java Card 3 Classic
+  Edition, a backward compatible evolution of version 2.2.2.,
+  including some bug fixes, clarifications against the Java Card
+  v2.2.2 specifications and new security algorithms (Java Card 3
+  Platform Release Notes, available at
+  http://java.sun.com/javacard/3.0.1/specs.jsp).}, is composed of 14
+packages, containing 93 classes and interfaces, as shown in
+Table~\ref{tab:packages}. Their interfaces have all been modeled in B,
+and the resulting models are available at the KitSmart project
+page\footnote{https://code.google.com/p/kitsmart/}.  Differently from
+JML and OCL, though, B is not object oriented, and some modeling
+solutions had to be found to cope with this lack of resources.  We
+present in the following the main characteristics and some
+restrictions of our models:
 
@@
 
-\item Conditions on method inputs which may lead to exception raising in the Java Card platform specification are coded into pre-conditions
-  in B.
+\item Conditions on method inputs which may lead to exception raising
+  in the Java Card platform specification are coded into
+  pre-conditions in B.
 
- \item Method outputs are modeled in B as taking any value of the correct type, or, when possible in the current abstraction level, any value of a more restricted set of specified return values.
+\item Method outputs are modeled in B as taking any value of the
+  correct type, or, when possible in the current abstraction level,
+  any value of a more restricted set of specified return values.
 
-\item Each machine corresponding to a Java Card class defines an abstract set with the same name (distinction on character cases, only).
-For instance, the machine Throwable declares the abstract set THROWABLE.
-This set can then be used to reference the type corresponding to the class.
+\item Each machine corresponding to a Java Card class defines an
+  abstract set with the same name (distinction on character cases,
+  only). For instance, the machine Throwable declares the abstract set
+  THROWABLE. This set can then be used to reference the type
+  corresponding to the class.
 
-\item The hierarchy of the Java Card API objects is partially modeled by abstract set inclusion.
-For instance, the class Throwable extends the class Object, in Java Card.
-In our model, we also have that the machine Throwable EXTENDS the machine Object, but the meaning of a machine extension in B is different from the meaning of a class extension in Java Card.
-To show that all throwables are indeed objects, a property stating that THROWABLE is a subset of OBJECT is included in the Throwable machine.
+\item The hierarchy of the Java Card API objects is partially modeled
+  by abstract set inclusion. For instance, the class Throwable extends
+  the class Object, in Java Card. In our model, we also have that the
+  machine Throwable EXTENDS the machine Object, but the meaning of a
+  machine extension in B is different from the meaning of a class
+  extension in Java Card. To show that all throwables are indeed
+  objects, a property stating that THROWABLE is a subset of OBJECT is
+  included in the Throwable machine.
 
-\item Overloading is not allowed in B and occurs frequently in Java Card.
-To model overloaded methods in B an extension to overloaded method names was included.
-We then have equals, equals\_1, equals\_2, etc.
+\item Overloading is not allowed in B and occurs frequently in Java
+  Card. To model overloaded methods in B an extension to overloaded
+  method names was included. We then have equals, equals\_1,
+  equals\_2, etc.
 \end{itemize}
 
-As previously stated,  we chose to model exception raising conditions as operation pre-conditions in the B models.
-For instance, if the Java Card specification states that a method should throw an exception in case a parameter array is empty, the pre-condition of the corresponding B operation will require that this parameter is not empty.
-The B method then requires that every call to that operation satisfies the operation's pre-condition.
-As a consequence, code that is formally generated using the B method and the Java Card API specification should never raise this specific exception, as it is statically proved that all calls to this operation are made with non-empty parameter arrays.
-Not all situations leading to exceptional conditions are modeled, however, as some of them depend on the Java Card Runtime Environment (JCRE) and its state, which has not yet been modeled, and some require a more detailed specification of the contents of some data and other internal behavior.
-In these cases, the  exception raising conditions are only left as comments to the B specification, and can be used as additional documentation or to be considered in future work.
+As previously stated, we chose to model exception raising conditions
+as operation pre-conditions in the B models. For instance, if the Java
+Card specification states that a method should throw an exception in
+case a parameter array is empty, the pre-condition of the
+corresponding B operation will require that this parameter is not
+empty. The B method then requires that every call to that operation
+satisfies the operation's pre-condition. As a consequence, code that
+is formally generated using the B method and the Java Card API
+specification should never raise this specific exception, as it is
+statically proved that all calls to this operation are made with
+non-empty parameter arrays. Not all situations leading to exceptional
+conditions are modeled, however, as some of them depend on the Java
+Card Runtime Environment (JCRE) and its state, which has not yet been
+modeled, and some require a more detailed specification of the
+contents of some data and other internal behavior. In these cases, the
+exception raising conditions are only left as comments to the B
+specification, and can be used as additional documentation or to be
+considered in future work.
 
-One delicate issue to be dealt with both by the user of the API specification and the tools designed to work with it, is the use of overloaded methods in an application development.
-When overloaded operations are used in an application development, the user must explicitly choose which version of the overloaded operation is to be used.
-Then, verifications will be carried out considering this specific version of the operation.
-When code is generated, either the code generator will remove the name extension automatically, if designed for it, or the user will have to edit the generated code manually.
-In any case, it is important that the actual method called in the Java Card code is the one with which verifications have been carried out.
-In case of overriding, in particular, it will the user's responsibility to choose in B the one which is on scope at this specific point in Java, unless a extra verification step is included in the code generator that checks that the operation used in B is the one in scope in Java and registers an error in case it is not.
+One delicate issue to be dealt with both by the user of the API
+specification and the tools designed to work with it, is the use of
+overloaded methods in an application development. When overloaded
+operations are used in an application development, the user must
+explicitly choose which version of the overloaded operation is to be
+used. Then, verifications will be carried out considering this
+specific version of the operation. When code is generated, either the
+code generator will remove the name extension automatically, if
+designed for it, or the user will have to edit the generated code
+manually. In any case, it is important that the actual method called
+in the Java Card code is the one with which verifications have been
+carried out.  In case of overriding, in particular, it will be either
+the user's responsibility to choose the B one which is in scope at
+this specific point in Java, or an extra verification step will be
+included in the code generator to check that the operation used in
+B is the one in scope in Java and registers an error in case it is
+not.
 
@@
 
-With the current level of abstraction of the model and the fact that the Java Card Runtime Environment (JCRE) has not been modeled, some
-conditions which may lead to exception raising are not yet modeled.
-Being able to model them would make developed applications even more sure, as more static verification could be carried out.
-Continuation of this work then will enlarge the scope of the modeling to cope statically with a larger set of situations.
+With the current level of abstraction of the model and the fact that
+the Java Card Runtime Environment (JCRE) has not been modeled, some
+conditions which may lead to exception raising are not yet
+modeled. Being able to model them would make developed applications
+even more sure, as more static verification could be carried
+out. Continuation of this work then will enlarge the scope of the
+modeling to cope statically with a larger set of situations.
 

### 2012-03-30 trunk/papers/B2012-work/javacardAPIpaper.tex

@@
   Platform Release Notes, available at
-  http://java.sun.com/javacard/3.0.1/specs.jsp).}, is composed of 14
+  http://www.oracle.com/technetwork/java/javame/javacard/download/overview/index.html).}, is composed of 14
 packages, containing 93 classes and interfaces, as shown in

### 2012-03-30 papers/B2012-work/javacardAPIpaper.tex

@@
 machine of Fig.~\ref{lis:buscard}, which models a bus card with a
-variable that keeps track of the current credit the user has (\texttt{balance}) and an operation to add credits to this
-balance (\texttt{add\_credit}!!! verificar). This simple example provides enough evidence of the
+variable that keeps track of the current credit the user has
+(\texttt{balance}) and an operation to add credits to this balance
+(\texttt{add}). This simple example provides enough evidence of the
 importance of having a B model of the Java Card API.  As all data in
-it is implementable in C, for instance, and it is a deterministic
+it is implementable in C, for instance, and as it is a deterministic
 specification, B tools can directly generate code for it in C, and the
@@
 CONSTANTS
-	maximum
+	MAX_BALANCE
 PROPERTIES
-	maximum : INTEGER & maximum = 100
+	MAX_BALANCE : INTEGER & MAX_BALANCE  = 100
 VARIABLES
@@
 	PRE
-		credit : INTEGER & credit > 0  & balance + credit <= maximum
+		 credit : INTEGER & credit > 0  & credit <= MAX_BALANCE &
+                balance + credit <= MAX_BALANCE 
 	THEN
@@
 
-However, Java Card code corresponding to it would look like the one
-presented in Fig.~\ref{lis:add}. In a rigorous development
-process, we do not want the code generator to carry out all these
-non-verified transformations in the code. So, we need to be able to
-refine the original B machine into something such as the B refinement
-of Fig.~\ref{B ref usando a API}. Then, code generation would be
-straightforward (as it should be). This is then the reason why it
-needed to be the first Java Card component to be modeled in B. (AMM:
-maybe also cite related work ? Which are the formalization works that
-did not focus on the API?)
+However, Java Card code corresponding to the operation \texttt{add} of
+this B specification would look like the one presented in
+Fig.~\ref{lis:addjavacard}. In a rigorous development process, we do
+not want the code generator to carry out all these non-verified
+transformations in the code. So, we need to be able to refine the
+original B machine into something such as the B refinement of
+Fig.~\ref{lis:addimplementation}. Then, code generation would be
+straightforward (as it should be).
 
-\lstset{language=Java, caption={Method add in applet Java Card},label=lis:add}
+\lstset{language=Java, caption={Method add in applet Java Card},label=lis:addjavacard}
 \begin{lstlisting}
-public void add(APDU apdu) {
-	byte[] buffer = apdu.getBuffer();
-	byte credit = (byte) Util.makeShort((byte) buffer[ISO7816.OFFSET_P1], (byte)buffer[ISO7816.OFFSET_P2]);
-	if (!(credit >= 0)) {
-		ISOException.throwIt(EXCEPTIONS.negative_credit);
+public void add (APDU apdu) { 
+	byte[] bf;  byte cr; byte apdu_state; short res;
+
+	bf = apdu.getBuffer();
+	apdu_state = apdu.getCurrentState();
+
+	if (apdu_state == STATE_INITIAL) {
+		res = apdu.setIncomingAndReceive();
+	} else if (ISO7816.OFFSET_CDATA >= 0 && 
+			(short) (ISO7816.OFFSET_CDATA + 2) <= bf.length && 
+			bf != null) {
+
+		cr = Util.getShort (bf, ISO7816.OFFSET_CDATA);
+
+		if (!(cr <= MAX_BALANCE)) {
+			ISOException.throwIt(EX_CREDIT_EXCEEDED);
+		} else if (!((short) (balance + cr) <= MAX_BALANCE)) {
+			ISOException.throwIt(EX_BALANCE_EXCEEDED);
+		} else if (!((short) (balance + cr) >= 0)) {
+			ISOException.throwIt(EX_NEGATIVE_BALANCE);
+		} else if (! (cr > 0)) {
+			ISOException.throwIt(EX_NEGATIVE_CREDIT);
+		} else {
+			balance = (short) (balance + cr);
 		}
-	else if (!(balance + credit <= MAXIMUM)) {
-		ISOException.throwIt(EXCEPTIONS.balance_exceeded);
-	}
-	else {
-		balance = (balance + credit);
 	} 
@@
 
-[[[ PARA SIMONE -> tambem precisamos com urgencia daquela anelise sobre o que muda da 2.2.2 para a 3 ]]]
+\lstset{language=B, caption={B implementation},label=lis:addimplementation}
+\begin{lstlisting}
+add (apdu) = 
+VAR 
+	bf, cr, apdu_state, res
+IN
+	bf  <-- apdu.getBuffer;		
+    	apdu_state  <-- apdu.getCurrentState;
 
-The Java Card API, in its version 2.2.2, is composed of 14 packages,
-containing 93 classes and interfaces, as shown in
+	IF (apdu_state = STATE_INITIAL) 
+	THEN
+		res <-- apdu.setIncomingAndReceive                
+	ELSIF (OFFSET_CDATA >= 0 &
+		sum_short(OFFSET_CDATA, 2) <= size(bf) & 
+		bf /= [])
+	THEN 
+		cr  <-- Util.getShort(bf, OFFSET_CDATA);
+
+		IF not (cr <= MAX_BALANCE)
+		THEN 
+			ISOException.throwIt(EX_CREDIT_EXCEEDED)
+		ELSIF not (sum_short (balance, cr) <= MAX_BALANCE)
+		THEN 
+			ISOException.throwIt(EX_BALANCE_EXCEEDED)
+		ELSIF not (sum_short (balance, cr) >= 0)
+		THEN ISOException.throwIt(EX_NEGATIVE_BALANCE)
+		ELSIF not (cr > 0)
+		THEN ISOException.throwIt(EX_NEGATIVE_CREDIT) 
+		ELSE  
+			balance := sum_short(balance, cr)
+		END
+	END
+END
+\end{lstlisting}
+
+We can identify two main kinds of transformation that have to be
+carried out in order to obtain the code of operation \texttt{add} in
+figs.~\ref{lis:addjavacard} and~\ref{lis:addimplementation} from the
+abstract specification in fig.~\ref{lis:buscard}:
+
+\begin{enumerate}
+\item the transition into a defensive style of programming by the
+  inclusion of explicit tests to check that the input conditions (pre
+  conditions in the B specification) are satisfied before executing
+  the corresponding action;
+
+\item the inclusion of Java Card specific aspects, such as changing
+  the signature of \texttt{add} so that data is transmitted via the
+  APDU protocol and the APDU buffer is manipulated by methods offered
+  the Java Card API.
+
+\end{enumerate}
+
+The transition to a defensive style of programming is part of the
+BSmart method, and implemented by the BSmart Full Function
+component. This component generates a B refinement of the original
+specification where pre-conditions are transformed into conditionals
+(IF-THEN-ELSE-END constructs), with the indication of the need to
+throw the corresponding exception when the specified conditions are
+not met by input data and states.  The resulting B model can then be
+formally verified as usual in the B method.
+
+The change of interface is also taken care of by the BSmart method, so
+that the development can be done without need to extend the B
+method, as detailed in~\cite{Gomes10}.
+
+But all of these transformations rely on the need to reference Java
+Card API types, classes and methods on the B level.  This is then the
+reason why the Java Card API needed to be the first Java Card
+component to be modeled in B.
+
+[[[(PARA SIMONE e BRUNO: maybe also cite related work ? Which are the formalization works that did not focus on the API?)]]]
+
+[[[ PARA SIMONE -> tambem precisamos com urgencia daquela analise sobre o que muda da 2.2.2 para a 3 ]]]
+
+The Java Card API, in its version 2.2.2\footnote{Java Card 2.2.2,
+  although not the most recent version of the Java Card platform, is
+  still widely used, and can easily be extended to Java Card 3 Classic
+  Edition, a backward compatible evolution of version 2.2.2.,
+  including some bug fixes, clarifications against the Java Card
+  v2.2.2 specifications and new security algorithms (Java Card 3
+  Platform Release Notes, available at
+  http://www.oracle.com/technetwork/java/javame/javacard/download/overview/index.html).}, is composed of 14
+packages, containing 93 classes and interfaces, as shown in
 Table~\ref{tab:packages}. Their interfaces have all been modeled in B,
@@
 solutions had to be found to cope with this lack of resources.  We
-present in the following the main characteristics and restrictions of
-our models:
+present in the following the main characteristics and some
+restrictions of our models:
 
@@
 
-\item Conditions on method inputs which lead to exception raising in
-  the Java Card platform specification are coded into pre-conditions
-  in B.
-
-  [[[ AMM (em que casos isso não é feito - NULL? outros?) ]]]
+\item Conditions on method inputs which may lead to exception raising
+  in the Java Card platform specification are coded into
+  pre-conditions in B.
 
 \item Method outputs are modeled in B as taking any value of the
-  correct type
+  correct type, or, when possible in the current abstraction level,
+  any value of a more restricted set of specified return values.
 
-  [[[ AMM (e em que casos temos um pouco mais). ]]]
+\item Each machine corresponding to a Java Card class defines an
+  abstract set with the same name (distinction on character cases,
+  only). For instance, the machine Throwable declares the abstract set
+  THROWABLE. This set can then be used to reference the type
+  corresponding to the class.
 
+\item The hierarchy of the Java Card API objects is partially modeled
+  by abstract set inclusion. For instance, the class Throwable extends
+  the class Object, in Java Card. In our model, we also have that the
+  machine Throwable EXTENDS the machine Object, but the meaning of a
+  machine extension in B is different from the meaning of a class
+  extension in Java Card. To show that all throwables are indeed
+  objects, a property stating that THROWABLE is a subset of OBJECT is
+  included in the Throwable machine.
+
+\item Overloading is not allowed in B and occurs frequently in Java
+  Card. To model overloaded methods in B an extension to overloaded
+  method names was included. We then have equals, equals\_1,
+  equals\_2, etc.
 \end{itemize}
 
-As previously stated,  we chose to model exception raising conditions as operation pre-conditions in the B models.
-For instance, if the Java Card specification states that a method should throw an exception in case a parameter array is empty, the pre-condition of the corresponding B operation will require that this parameter is not empty.
-The B method then requires that every call to that operation satisfies the operation's pre-condition.
-As a consequence, code that is formally generated using the B method and the Java Card API specification should never raise this specific exception, as it is statically proved that all calls to this operation are made with non-empty parameter arrays.
-Not all situations leading to exceptional conditions are modeled, however, as some of them depend on the Java Card Runtime Environment (JCRE) and its state, which has not yet been modeled, and some require a more detailed specification of the contents of some data and other internal behavior.
+As previously stated, we chose to model exception raising conditions
+as operation pre-conditions in the B models. For instance, if the Java
+Card specification states that a method should throw an exception in
+case a parameter array is empty, the pre-condition of the
+corresponding B operation will require that this parameter is not
+empty. The B method then requires that every call to that operation
+satisfies the operation's pre-condition. As a consequence, code that
+is formally generated using the B method and the Java Card API
+specification should never raise this specific exception, as it is
+statically proved that all calls to this operation are made with
+non-empty parameter arrays. Not all situations leading to exceptional
+conditions are modeled, however, as some of them depend on the Java
+Card Runtime Environment (JCRE) and its state, which has not yet been
+modeled, and some require a more detailed specification of the
+contents of some data and other internal behavior. In these cases, the
+exception raising conditions are only left as comments to the B
+specification, and can be used as additional documentation or to be
+considered in future work.
+
+One delicate issue to be dealt with both by the user of the API
+specification and the tools designed to work with it, is the use of
+overloaded methods in an application development. When overloaded
+operations are used in an application development, the user must
+explicitly choose which version of the overloaded operation is to be
+used. Then, verifications will be carried out considering this
+specific version of the operation. When code is generated, either the
+code generator will remove the name extension automatically, if
+designed for it, or the user will have to edit the generated code
+manually. In any case, it is important that the actual method called
+in the Java Card code is the one with which verifications have been
+carried out.  In case of overriding, in particular, it will be either
+the user's responsibility to choose the B one which is in scope at
+this specific point in Java, or an extra verification step will be
+included in the code generator to check that the operation used in
+B is the one in scope in Java and registers an error in case it is
+not.
+
+[[[ incluir aqui parte da discussao sobre as opcoes de modelagem  com o argumento pela escolha dos subsets para a tipagem ]]] 
+As our focus was on the interfaces of the Java Card API methods,
 

### 2012-04-02 trunk/papers/B2012-work/javacardAPIpaper.tex
mais texto sobre a modelagem e ajustes nas referencias.

@@
 \begin{frontmatter}
-%  \title{Modeling the Java Card API (or Interface?) in B} \author{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
-%  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{S. de Oliveira Santos \thanksref{simone}}
-%  \address{DIMAp-UFRN, Natal, RN, Brazil} \author{B. Emerson Gurgel Gomes \thanksref{bruno}}
-%    \address{IFRN, Currais Novos, RN, Brazil} \author{D. Boris Paul D\'{e}harbe \thanksref{david}}
-%      \address{DIMAp-UFRN, Natal, RN, Brazil} \thanks[ALL]{INES-CAPES-projeto GD?... o que mais?} \thanks[anamaria]{Email:
-%    \href{mailto:anamaria@dimap.ufrn.br} {\texttt{\normalshape
-%        anamaria@dimap.ufrn.br}}} 
-%      \thanks[simone]{Email:
-%    \href{mailto:simone82@ppgsc.ufrn.br} {\texttt{\normalshape
-%        simone82@ppgsc.ufrn.br}}}
-%      \thanks[bruno]{Email:
-%    \href{mailto:bruno.gurgel@ifrn.edu.br} {\texttt{\normalshape
-%        bruno.gurgel@ifrn.edu.br}}}
-%       \thanks[david]{Email:
-%    \href{mailto:david@dimap.ufrn.br} {\texttt{\normalshape
-%        david@dimap.ufrn.br}}}
 
-  \title{Modeling the Java Card API (or Interface?) in B}
+  \title{Modeling the Java Card API in B}
   \author[ufrn]{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
@@
 (\texttt{add}). This simple example provides enough evidence of the
-importance of having a B model of the Java Card API.  As all data in
-it is implementable in C, for instance, and as it is a deterministic
-specification, B tools can directly generate code for it in C, and the
+importance of having a B model of the Java Card API.  B tools can directly 
+generate code for it in C, for instance, because it  is a deterministic specification and all data in
+it is implementable in C.  The
 resulting code would be very similar to the original B machine.
@@
 CONSTANTS
-	maximum
+	MAX_BALANCE
 PROPERTIES
-	maximum : INTEGER & MAX_BALANCE  = 100
+	MAX_BALANCE : INTEGER & MAX_BALANCE  = 100
 VARIABLES
@@
 	PRE
-		credit : INTEGER & credit > 0  & 
+		 credit : INTEGER & 
+		 credit > 0  & credit <= MAX_BALANCE &
                 balance + credit <= MAX_BALANCE 
@@
 public void add (APDU apdu) { 
-	byte[] bf;  short cr; byte apdu_state; short res;
+	byte[] bf;  byte cr; byte apdu_state; short res;
 
@@
 
-		if (!((short) (balance + cr) <= MAX_BALANCE)) {
+		if (!(cr <= MAX_BALANCE)) {
+			ISOException.throwIt(EX_CREDIT_EXCEEDED);
+		} else if (!((short) (balance + cr) <= MAX_BALANCE)) {
 			ISOException.throwIt(EX_BALANCE_EXCEEDED);
-		} else if (!((short) (balance + cr) >= 0)) {
-			ISOException.throwIt(EX_NEGATIVE_BALANCE);
 		} else if (! (cr > 0)) {
@@
 
-		IF not (sum_short (balance, cr) <= MAX_BALANCE)
+		IF not (cr <= MAX_BALANCE)
+		THEN 
+			ISOException.throwIt(EX_CREDIT_EXCEEDED)
+		ELSIF not (sum_short (balance, cr) <= MAX_BALANCE)
 		THEN 
 			ISOException.throwIt(EX_BALANCE_EXCEEDED)
-		ELSIF not (sum_short (balance, cr) >= 0)
-		THEN ISOException.throwIt(EX_NEGATIVE_BALANCE)
 		ELSIF not (cr > 0)
-		THEN ISOException.throwIt(EX_NEGATIVE_CREDIT) 
+		THEN 
+			ISOException.throwIt(EX_NEGATIVE_CREDIT) 
 		ELSE  
@@
 that the development can be done without need to extend the B
-method, as detailed in~\cite{Gomes10}.
+method, as presented in~\cite{Gomes10}.
 
@@
 reason why the Java Card API needed to be the first Java Card
-component to be modeled in B.
-
-[[[(PARA SIMONE e BRUNO: maybe also cite related work ? Which are the formalization works that did not focus on the API?)]]]
-
-[[[ PARA SIMONE -> tambem precisamos com urgencia daquela analise sobre o que muda da 2.2.2 para a 3 ]]]
+component to be modeled in B. This is a common choice, as we can see in previous work on the formalization of 
+Java Card in JML \cite{pollBerg01} and OCL \cite{Larsson:2004}.
 
@@
 
+  \item In the current version of the specification, states are not modeled. 
+  The B machines corresponding to Java Card classes or interfaces are stateless machines.
+
 \item Each machine corresponding to a Java Card class defines an
@@
   by abstract set inclusion. For instance, the class Throwable extends
-  the class Object, in Java Card. In our model, we also have that the
-  machine Throwable EXTENDS the machine Object, but the meaning of a
-  machine extension in B is different from the meaning of a class
-  extension in Java Card. To show that all throwables are indeed
+  the class Object, in Java Card. In our model, we  have that the
+  machine Throwable SEES the machine Object, and too show that all throwables are indeed
   objects, a property stating that THROWABLE is a subset of OBJECT is
-  included in the Throwable machine.
+  included, associated to the Throwable machine.
 
@@
 exception raising conditions are only left as comments to the B
-specification, and can be used as additional documentation or to be
+specification, and can either be used as additional documentation or be
 considered in future work.
@@
 operations are used in an application development, the user must
-explicitly choose which version of the overloaded operation is to be
+explicitly choose which version (in B) of the overloaded operation is to be
 used. Then, verifications will be carried out considering this
@@
 
-[[[ incluir aqui parte da discussao sobre as opcoes de modelagem  com o argumento pela escolha dos subsets para a tipagem ]]] 
-As our focus was on the interfaces of the Java Card API methods,
+Finally, the choices related to class derived types and return values were consequences of our focus on the interfaces of the Java Card API methods.
+The declaration of an abstract set for each class allows to model in B that a parameter of a particular class is expected or is to be returned by a method.
+
+\lstset{language=B, caption={B machine Buscard},label=lis:object}
+\begin{lstlisting}
+MACHINE Object
+SEES  ObjectSet
+OPERATIONS
+ 	result <-- equals(obj) =
+	PRE obj : OBJECT   /* declared in the ObjectSet machine */
+	THEN
+		result :: {TRUE, FALSE}
+	END
+END
+
+MACHINE Throwable
+/* This class extends java.lang.Object */
+SEES
+    Object,
+    ThrowableSet
+END
+
+MACHINE ObjectSet    
+SETS
+     OBJECT   
+END
+
+MACHINE ThrowableSet
+SEES 
+    ObjectSet    
+CONSTANTS  THROWABLE
+PROPERTIES THROWABLE <: OBJECT
+END
+\end{lstlisting}
 

### 2012-04-02 trunk/papers/B2012-work/javacardAPIpaper.tex
secao 4 - adicionado texto

@@
 
-\section{B specification of the Java Card API}
+\section{B specification of the Java Card API}\label{sec:api-spec}
 
@@
 
+The models are used in the different phases of the 
+BSmart method development. Initially, the application 
+abstract specification uses the models of basic Java types, 
+since Java Card specific aspects are ideally related only
+in subsequent refinements.
+
+The BSmart Full Function refinement introduces the modelling of exceptions, 
+as stated in section~\ref{sec:api-spec}. This is performed through  the inclusion
+of \emph{ISOException} machine, which receives, in its \emph{throwIt} operation, 
+a short value corresponding to the \emph{status} code that must be returned to 
+the client when the conditions are not satisfied.
+
+In the B implementation, the whole set of the API can be related. Usually, besides
+ISOException (or other exception model), in a typical Java Card development 
+are imported \emph{Apdu} and \emph{Util}  modules. The fist defines operations
+to start the process of receiving (from host) and sending (to host) information 
+through the \emph{apdu buffer}. The preconditions of these sending/receiving 
+operations aid the verification process, since the specifier must to observe 
+the correct order in operation calling. The \emph{Util} component contains useful 
+operations to effectively get and put data in the buffer. When translating to 
+\emph{Java Card} the operation calling are translated to the corresponding 
+\emph{Java Card} method.
+
 \section{Conclusions and future work}

### 2012-04-02 trunk/papers/B2012-work/javacardAPIpaper.tex
mudancas david

@@
 
-AMM: TRECHOS EXTRAIDOS E JA UM POUCO ALTERADOS DO ABZ 2010. DEPENDENDO
-DO QUE VAMOS MOSTRAR NA ESPECIFICAÇAO PODEMOS DEIXAR MAIS OU MENOS
-DETALHES AQUI. NO FINAL, PRECISAMOS VERIFICAR QUE NAO COPIAMOS DEMAIS
-O QUE ESTAVA LA.
-
 A smart card application is distributed between on-card and off-card
@@
 provides the application services. The off-card client (called host
-application) resides in a computer or electronic terminal.
-
-The information exchange between the host and card applications is
+application) resides in a computer or electronic terminal. The information 
+exchange between the host and card applications is
 made through a a half-duplexed low level protocol, named
-\emph{Application Protocol Data Unit} (APDU).  The ISO 7816-4 standard
-specifies two kinds of APDU's, which are the command and the response
-APDU (Figure~\ref{fig:APDUs}). Both specify data packets. A command APDU
-is sent by the host, requiring some applet service and a response APDU is
-sent by the applet, responding to the host request with the result of the
-service processing.
-
-\begin{figure}[!ht]
-\centering
-%\includegraphics[scale=0.3]{figs/comandoAPDU.jpg}
-%\includegraphics[scale=0.3]{figs/respostaAPDU.jpg}
-\caption{Command and response APDUs. Source:~\cite{ortiz1intro:2003}}
-\label{fig:APDUs}
-\end{figure}
+\emph{Application Protocol Data Unit} (APDU).
 
@@
 
-\begin{figure}[h]
-\centering
-%\includegraphics[width=0.35\textwidth]{figs/jcplatform.png}
-\caption{Java Card platform components. Adapted from:~\cite{ortiz1intro:2003}}
-\label{fig:jcre}
-\end{figure}
+The Java Card API provides the means to deal with the APDU protocol,
 
@@
 
-\lstset{language=B, caption={B machine Buscard},label=lis:object}
+\lstset{language=B, caption={Basic API Definitions},label=lis:object}
 \begin{lstlisting}
@@
 
-The models are used in the different phases of the 
+The models of the Java Card API are used in the different phases of the 
 BSmart method development. Initially, the application 
@@
 
-In the B implementation, the whole set of the API can be related. Usually, besides
-ISOException (or other exception model), in a typical Java Card development 
-are imported \emph{Apdu} and \emph{Util}  modules. The fist defines operations
-to start the process of receiving (from host) and sending (to host) information 
-through the \emph{apdu buffer}. The preconditions of these sending/receiving 
-operations aid the verification process, since the specifier must to observe 
-the correct order in operation calling. The \emph{Util} component contains useful 
-operations to effectively get and put data in the buffer. When translating to 
-\emph{Java Card} the operation calling are translated to the corresponding 
-\emph{Java Card} method.
+In the B implementation, the whole API can be referenced. Usually,
+besides ISOException (or another exception model), a typical Java Card
+development imports the \emph{Apdu} and \emph{Util} modules. The first
+defines operations to start the process of receiving (from host) and
+sending (to host) information through the \emph{apdu buffer}. The
+preconditions of these sending/receiving operations aid the
+verification process, since the specifier must observe the correct
+order of operation calls. The \emph{Util} component contains useful
+operations to effectively get and put data in the buffer. When
+translating to \emph{Java Card} the operation calls are translated
+to the corresponding \emph{Java Card} method.
 

### 2012-04-02 trunk/papers/B2012-work/javacardAPIpaper.tex
quase - arrumada a apresentacao de java card, quebrada a secao 3, incluida explicacao do modelo de Object e cia e seu uso.

@@
 
-The main component of the Java Card platform (Figure~\ref{fig:jcre})
+The main component of the Java Card platform
 is its runtime environment (JCRE), composed of a Java Card Virtual
-Machine (JCVM), an API, and, usually, system and
+Machine (JCVM), an Application Programming Interface (API), and, usually, system and
 industry-specific classes~\cite{chen:2000}. The JCRE acts as a small
@@
 
-The Java Card API provides the means to deal with the APDU protocol,
+The Java Card API is the part of the Java Card platform that a developer of any Java Card application must use.
+It defines the constituents of an applet, provides the means to deal with the APDU protocol and communicate data and commands between the host application and the card, and defines some basic services for security and cryptography implementation.
 
-\paragraph{Java Card applets}\label{sec:applets}
-
-A Java Card applet is a class that inherits the
-\emph{javacard.framewok.Applet} class of the Java Card API and is
-implemented upon the Java Card subset of Java.  During the applet
+In a Java Card application, the applet is the part which defines the card provided services. 
+An applet inherits the \emph{javacard.framewok.Applet} class of the Java Card API and is
+implemented upon the Java Card subset of Java to provide these services.  During the applet
 conversion for card installation, a verification phase is performed to
@@
 
-The current usual Java Card specification (2.2.x) allows two kinds of
-applets.  The older, and most commonly used, kind of applet
-manipulates directly the APDU protocol packages while the newer one
-abstracts from the lower level protocol using Remote Method Invocation
-(RMI). In this paper we will call the lower level applet \emph{APDU
-  applet} and the higher level one, \emph{RMI applet}. RMI introduces
-a layer of abstraction above the APDU protocol, and hence is usually
-less efficient than APDU applets.
-
 \section{B specification of the Java Card API}\label{sec:api-spec}
 
+\paragraph{Motivation}
 Among the different components of the Java Card platform, the Java
@@
 could be done.  Suppose, for instance, that we had the B abstract
-machine of Fig.~\ref{lis:buscard}, which models a bus card with a
+machine of Listing~\ref{lis:buscard}, which models a bus card with a
 variable that keeps track of the current credit the user has
@@
 this B specification would look like the one presented in
-Fig.~\ref{lis:addjavacard}. In a rigorous development process, we do
+Listing~\ref{lis:addjavacard}. In a rigorous development process, we do
 not want the code generator to carry out all these non-verified
@@
 original B machine into something such as the B refinement of
-Fig.~\ref{lis:addimplementation}. Then, code generation would be
+Listing~\ref{lis:addimplementation}. Then, code generation would be
 straightforward (as it should be).
@@
 figs.~\ref{lis:addjavacard} and~\ref{lis:addimplementation} from the
-abstract specification in fig.~\ref{lis:buscard}:
+abstract specification in Listing~\ref{lis:buscard}:
 
@@
 
+\paragraph{The model}
 The Java Card API, in its version 2.2.2\footnote{Java Card 2.2.2,
@@
   the class Object, in Java Card. In our model, we  have that the
-  machine Throwable SEES the machine Object, and too show that all throwables are indeed
+  machine Throwable SEES the machine Object, and to show that all throwables are indeed
   objects, a property stating that THROWABLE is a subset of OBJECT is
@@
 
+A small piece of specification with some of the characteristics described above is presented in Listing \ref{lis:object}.
+It presents two \texttt{java.lang} classes: \texttt{Object} and \texttt{Throwable}.
+\texttt{Object} is  superclass to almost all API classes, with its only method  \texttt{equals} that checks if an object (from which a reference is received as argument) is the same as the current one.
+\texttt{Throwable} extends \texttt{Object}, but does not define any other method nor redefines \texttt{equals}.
+The inheritance hierarchy is partially simulated in B by the corresponding set inclusion.
+These abstract sets are then used to refer to the corresponding class as a type as can be seen in Listing \ref{lis:class-types}.
+
+\lstset{language=B, caption={Basic API Definitions},label=lis:object}
+\begin{lstlisting}
+MACHINE Object
+SEES  ObjectSet
+OPERATIONS
+ 	result <-- equals(obj) =
+	PRE obj : OBJECT   /* declared in the ObjectSet machine */
+	THEN
+		result :: {TRUE, FALSE}
+	END
+END
+
+MACHINE Throwable
+/* This class extends java.lang.Object */
+SEES
+    Object,
+    ThrowableSet
+END
+
+MACHINE ObjectSet
+SETS
+     OBJECT   
+END
+
+MACHINE ThrowableSet
+SEES 
+    ObjectSet    
+CONSTANTS  THROWABLE
+PROPERTIES THROWABLE <: OBJECT
+END
+
+\end{lstlisting}
+
+\lstset{language=B, caption={Using classes as types},label=lis:class-types}
+\begin{lstlisting}
+op(anObject,otherObject) = 
+	PRE	
+		anObject : OBJECT & otherObject : THROWABLE
+...
+\end{lstlisting}
+
+\paragraph{Discussions on the model}
 As previously stated, we chose to model exception raising conditions
@@
 The declaration of an abstract set for each class allows to model in B that a parameter of a particular class is expected or is to be returned by a method.
-
-\lstset{language=B, caption={Basic API Definitions},label=lis:object}
-\begin{lstlisting}
-MACHINE Object
-SEES  ObjectSet
-OPERATIONS
- 	result <-- equals(obj) =
-	PRE obj : OBJECT   /* declared in the ObjectSet machine */
-	THEN
-		result :: {TRUE, FALSE}
-	END
-END
-
-MACHINE Throwable
-/* This class extends java.lang.Object */
-SEES
-    Object,
-    ThrowableSet
-END
-
-MACHINE ObjectSet
-SETS
-     OBJECT   
-END
-
-MACHINE ThrowableSet
-SEES 
-    ObjectSet    
-CONSTANTS  THROWABLE
-PROPERTIES THROWABLE <: OBJECT
-END
-\end{lstlisting}
 

### 2012-04-02 papers/B2012-work/javacardAPIpaper.tex
versao que vou submeter agora.

@@
 
-In the following...
+In the following we present some important aspects of the Java Card platform, in section \ref{sec:javacard}, and we introduce the B model of the Java Card API,  in section \ref{sec:api-spec}.
+We conclude with some notes on the use of the proposed model, in section \ref{sec:using-the-API}, and some conclusions and future work (in section \ref{sec:conc}).
 
 \section{Smart Cards and Java Card}\label{sec:javacard}
-
-AMM: TRECHOS EXTRAIDOS E JA UM POUCO ALTERADOS DO ABZ 2010. DEPENDENDO
-DO QUE VAMOS MOSTRAR NA ESPECIFICAÇAO PODEMOS DEIXAR MAIS OU MENOS
-DETALHES AQUI. NO FINAL, PRECISAMOS VERIFICAR QUE NAO COPIAMOS DEMAIS
-O QUE ESTAVA LA.
 
@@
 provides the application services. The off-card client (called host
-application) resides in a computer or electronic terminal.
+application) resides in a computer or electronic terminal. The information 
+exchange between the host and card applications is
+made through a a half-duplexed low level protocol, named
+\emph{Application Protocol Data Unit} (APDU).
 
-The information exchange between the host and card applications is
-made through a a half-duplexed low level protocol, named
-\emph{Application Protocol Data Unit} (APDU).  The ISO 7816-4 standard
-specifies two kinds of APDU's, which are the command and the response
-APDU (Figure~\ref{fig:APDUs}). Both specify data packets. A command APDU
-is sent by the host, requiring some applet service and a response APDU is
-sent by the applet, responding to the host request with the result of the
-service processing.
-
-\begin{figure}[!ht]
-\centering
-%\includegraphics[scale=0.3]{figs/comandoAPDU.jpg}
-%\includegraphics[scale=0.3]{figs/respostaAPDU.jpg}
-\caption{Command and response APDUs. Source:~\cite{ortiz1intro:2003}}
-\label{fig:APDUs}
-\end{figure}
-
-The main component of the Java Card platform (Figure~\ref{fig:jcre})
+The main component of the Java Card platform
 is its runtime environment (JCRE), composed of a Java Card Virtual
-Machine (JCVM), an API, and, usually, system and
+Machine (JCVM), an Application Programming Interface (API), and, usually, system and
 industry-specific classes~\cite{chen:2000}. The JCRE acts as a small
@@
 
-\begin{figure}[h]
-\centering
-%\includegraphics[width=0.35\textwidth]{figs/jcplatform.png}
-\caption{Java Card platform components. Adapted from:~\cite{ortiz1intro:2003}}
-\label{fig:jcre}
-\end{figure}
+The Java Card API is the part of the Java Card platform that a developer of any Java Card application must use.
+It defines the constituents of an applet, provides the means to deal with the APDU protocol and communicate data and commands between the host application and the card, and defines some basic services for security and cryptography implementation.
 
-\paragraph{Java Card applets}\label{sec:applets}
-
-A Java Card applet is a class that inherits the
-\emph{javacard.framewok.Applet} class of the Java Card API and is
-implemented upon the Java Card subset of Java.  During the applet
+In a Java Card application, the applet is the part which defines the card provided services. 
+An applet inherits the \emph{javacard.framewok.Applet} class of the Java Card API and is
+implemented upon the Java Card subset of Java to provide these services.  During the applet
 conversion for card installation, a verification phase is performed to
@@
 
-The current usual Java Card specification (2.2.x) allows two kinds of
-applets.  The older, and most commonly used, kind of applet
-manipulates directly the APDU protocol packages while the newer one
-abstracts from the lower level protocol using Remote Method Invocation
-(RMI). In this paper we will call the lower level applet \emph{APDU
-  applet} and the higher level one, \emph{RMI applet}. RMI introduces
-a layer of abstraction above the APDU protocol, and hence is usually
-less efficient than APDU applets.
+\section{B specification of the Java Card API}\label{sec:api-spec}
 
-\section{B specification of the Java Card API}
-
+\paragraph{Motivation}
 Among the different components of the Java Card platform, the Java
@@
 	PRE
-		 credit : INTEGER & credit > 0  & credit <= MAX_BALANCE &
+		 credit : INTEGER & 
+		 credit > 0  & credit <= MAX_BALANCE &
                 balance + credit <= MAX_BALANCE 
@@
 			ISOException.throwIt(EX_BALANCE_EXCEEDED);
-		} else if (!((short) (balance + cr) >= 0)) {
-			ISOException.throwIt(EX_NEGATIVE_BALANCE);
 		} else if (! (cr > 0)) {
@@
 			ISOException.throwIt(EX_BALANCE_EXCEEDED)
-		ELSIF not (sum_short (balance, cr) >= 0)
-		THEN ISOException.throwIt(EX_NEGATIVE_BALANCE)
 		ELSIF not (cr > 0)
-		THEN ISOException.throwIt(EX_NEGATIVE_CREDIT) 
+		THEN 
+			ISOException.throwIt(EX_NEGATIVE_CREDIT) 
 		ELSE  
@@
 reason why the Java Card API needed to be the first Java Card
-component to be modeled in B.
+component to be modeled in B. This is a common choice, as we can see in previous work on the formalization of 
+Java Card in JML \cite{pollBerg01} and OCL \cite{Larsson:2004}.
 
-[[[(PARA SIMONE e BRUNO: maybe also cite related work ? Which are the formalization works that did not focus on the API?)]]]
-
-[[[ PARA SIMONE -> tambem precisamos com urgencia daquela analise sobre o que muda da 2.2.2 para a 3 ]]]
-
+\paragraph{The model}
 The Java Card API, in its version 2.2.2\footnote{Java Card 2.2.2,
@@
   by abstract set inclusion. For instance, the class Throwable extends
-  the class Object, in Java Card. In our model, we also have that the
-  machine Throwable EXTENDS the machine Object, but the meaning of a
-  machine extension in B is different from the meaning of a class
-  extension in Java Card. To show that all throwables are indeed
+  the class Object, in Java Card. In our model, we  have that the
+  machine Throwable SEES the machine Object, and to show that all throwables are indeed
   objects, a property stating that THROWABLE is a subset of OBJECT is
-  included in the Throwable machine.
+  included, associated to the Throwable machine.
 
@@
 
-[[[ incluir aqui parte da discussao sobre as opcoes de modelagem  com o argumento pela escolha dos subsets para a tipagem ]]] 
-As our focus was on the interfaces of the Java Card API methods,
+Finally, the choices related to class derived types and return values were consequences of our focus on the interfaces of the Java Card API methods.
+The declaration of an abstract set for each class allows to model in B that a parameter of a particular class is expected or is to be returned by a method.
 
-\section{Using the model in BSmart}
+\section{Using the model in BSmart}\label{sec:using-the-API}
 
-\section{Conclusions and future work}
+The models of the Java Card API are used in the different phases of the 
+BSmart method development. Initially, the application 
+abstract specification uses the models of basic Java types, 
+since Java Card specific aspects are ideally related only
+in subsequent refinements.
 
-With the current level of abstraction of the model and the fact that
-the Java Card Runtime Environment (JCRE) has not been modeled, some
-conditions which may lead to exception raising are not yet
-modeled. Being able to model them would make developed applications
-even more sure, as more static verification could be carried
-out. Continuation of this work then will enlarge the scope of the
-modeling to cope statically with a larger set of situations.
+The BSmart Full Function refinement introduces the modelling of exceptions, 
+as stated in section~\ref{sec:api-spec}. This is performed through  the inclusion
+of \emph{ISOException} machine, which receives, in its \emph{throwIt} operation, 
+a short value corresponding to the \emph{status} code that must be returned to 
+the client when the conditions are not satisfied.
+
+In the B implementation, the whole API can be referenced. Usually,
+besides ISOException (or another exception model), a typical Java Card
+development imports the \emph{Apdu} and \emph{Util} modules. The first
+defines operations to start the process of receiving (from host) and
+sending (to host) information through the \emph{apdu buffer}. The
+preconditions of these sending/receiving operations aid the
+verification process, since the specifier must observe the correct
+order of operation calls. The \emph{Util} component contains useful
+operations to effectively get and put data in the buffer. When
+translating to \emph{Java Card} the operation calls are translated
+to the corresponding \emph{Java Card} method.
+
+As a case study on the use of this API on the formal development of Java Card applications, part of the electronic passport specification\footnote{Available at http://www.icao.int.} is being modeled and developed in BSmart, showing that it is feasible to have a Java Card-like B refinement  such as the one presented in Listing \ref{lis:addimplementation}.
+
+\section{Conclusions and future work}\label{sec:conc}
+
+The B model described in this paper has as intent to be of use to Java Card application developers, either as additional (formal) documentation to the official one, or, better, as an imported library to be used in the formal development of applications with the B method.
+In section \ref{sec:api-spec} some characteristics and limitations of the work have been presented.
+Current work on using the API model to formally develop Java Card applications with B shows promising results.
+
+Among the limitations of this work, some may be dealt with by adapted tools and some will actually require some special attention from the developer, even with tool support, such as method overloading and overriding .
+These are the characteristics that are related to the fact that B is not object oriented and does not include exception raising and handling mechanisms.
+Some work on this last item can be seen on \cite{BR2003}, but it has not been implemented.
+
+With the current level of abstraction of the model and the fact that the Java Card Runtime Environment (JCRE) has not been modeled, different conditions which may lead to exception raising are not yet modeled.
+Being able to model them would make developed applications even more sure, as more static verification could be carried
+out. Continuation of this work then will enlarge the scope of the modeling to cope statically with a larger set of situations.
 

### 2012-04-20 trunk/papers/B2012-work/javacardAPIpaper.tex

@@
-\documentclass{entcs} 
-\usepackage{entcsmacro}
+\documentclass{llncs} 
 \usepackage{graphicx}
@@
 %backgroundcolor=\color{lightgray},
-keywordstyle=\color{red},
+%keywordstyle=\color{red},
 showstringspaces=false
@@
 
-\def\lastname{Martins Moreira}
+\begin{document}
 
-\begin{document}
-\begin{frontmatter}
+\mainmatter
 
-  \title{Modeling the Java Card API in B}
-  \author[ufrn]{A. Martins Moreira\thanksref{ALL}\thanksref{anamaria}}
-  \author[ufrn]{S. de Oliveira Santos \thanksref{simone}}
-  \author[ifrn]{B. Gurgel Gomes \thanksref{bruno}} \author[ufrn]{D.
-    D\'{e}harbe \thanksref{david}} \address[ufrn]{DIMAp-UFRN, Natal,
-    RN, Brazil} \thanks[ALL]{This work is partly supported by CAPES
+\title{Modeling the Java Card API in B \thanks{This work is partly supported by CAPES
  and CNPq grants 306033/2009-7, 560014/2010-4 and 573964/2008-4
-    (National Institute of Science and Technology for Software
-    Engineering---INES, \texttt{www.ines.org.br}).}
-\address[ifrn]{IFRN, Currais Novos, RN, Brazil}
-\thanks[anamaria]{Email: \href{mailto:anamaria@dimap.ufrn.br}
-  {\texttt{\normalshape anamaria@dimap.ufrn.br}}}
-\thanks[simone]{Email: \href{mailto:simone82@ppgsc.ufrn.br}
-  {\texttt{\normalshape simone82@ppgsc.ufrn.br}}}
-\thanks[bruno]{Email: \href{mailto:bruno.gurgel@ifrn.edu.br}
-  {\texttt{\normalshape bruno.gurgel@ifrn.edu.br}}}
-\thanks[david]{Email: \href{mailto:david@dimap.ufrn.br}
-  {\texttt{\normalshape david@dimap.ufrn.br}}}
+ (National Institute of Science and Technology for Software Engineering---INES, \texttt{www.ines.org.br}).}}
+
+\author{A. Martins Moreira\ \and S. de Oliveira Santos \and B. Gurgel Gomes \and D. D\'{e}harbe}
+\institute{Federal University of Rio Grande do Norte (UFRN); Natal; Brazil}
+%email{\{anamaria,david\}@dimap.ufrn.br, simone82@ppgsc.ufrn.br,bruno.gurgel@ifrn.edu.br}
+
+\maketitle
 
@@
 \end{abstract}
-\begin{keyword}
-  Java Card, smart cards, B method, formal methods
-\end{keyword}
-\end{frontmatter}
+
 \section{Introduction}\label{intro}
@@
 In the context of the BSmart project, we are interested in the use of
-B to develop Java Card~\cite{chen:2000} applications.  Java Card is a
+B~\cite{BBook} to develop Java Card~\cite{chen:2000} applications.  Java Card is a
 restricted and optimized version of Java which allows memory and
@@
 regularity of structure that led our group to define a method and
-implement a tool which specialize B for development of such
+implement a tool which specialize B for the development of such
 applications, providing greater support for the developer than pure B
@@
 in JBTools \cite{JBtools}, the generated code needed to be edited by
-the developer to include these aspects.
-
-These works lacked then the ability to formally verify the Java Card
+the developer to include these aspects. These works lacked then the ability to formally verify the Java Card
 dependent aspects of the application. To be able to change this, a B
@@
 
-In the following...
+In the following we present some important aspects of the Java Card platform, in section \ref{sec:javacard}, and we introduce the B model of the Java Card API,  in section \ref{sec:api-spec}.
+We conclude with some notes on the use of the proposed model, in section \ref{sec:using-the-API}, and some conclusions and future work (in section \ref{sec:conc}).
 
@@
 exchange between the host and card applications is
-made through a a half-duplexed low level protocol, named
+made through a  low level protocol, named
 \emph{Application Protocol Data Unit} (APDU).
@@
 \begin{lstlisting}
-MACHINE
-	Buscard
-CONSTANTS
-	MAX_BALANCE
-PROPERTIES
-	MAX_BALANCE : INTEGER & MAX_BALANCE  = 100
-VARIABLES
-	balance
-INVARIANT
-	balance : INTEGER & balance <= MAX_BALANCE 
-INITIALISATION
-	balance := 0
+MACHINE	Buscard
+CONSTANTS	MAX_BALANCE
+PROPERTIES	MAX_BALANCE : INTEGER & MAX_BALANCE  = 100
+VARIABLES	balance
+INVARIANT	balance : INTEGER & balance <= MAX_BALANCE 
+INITIALISATION balance := 0
 OPERATIONS
 	add(credit) =
-	PRE
-		 credit : INTEGER & 
+	PRE 	credit : INTEGER &  
 		credit > 0  & credit <= MAX_BALANCE & 
 		balance + credit <= MAX_BALANCE 
-	THEN
-		balance := balance + credit
+	THEN balance := balance + credit
 	END
@@
 	byte[] bf;  byte cr; byte apdu_state; short res;			
-
 	bf = apdu.getBuffer();
 	apdu_state = apdu.getCurrentState();
-
 	if (apdu_state == STATE_INITIAL) { 
-		res = apdu.setIncomingAndReceive();
-	} else if (ISO7816.OFFSET_CDATA >= 0 && 
-			(short) (ISO7816.OFFSET_CDATA + 2) <= bf.length && 
-			bf != null) {
-
+			res = apdu.setIncomingAndReceive();} 
+	else if ( ISO7816.OFFSET_CDATA >= 0 && bf != null &&
+		(short) (ISO7816.OFFSET_CDATA + 2) <= bf.length) {
 			cr = Util.getShort (bf, ISO7816.OFFSET_CDATA);
-
 		if (!(cr <= MAX_BALANCE)) {	
@@
 			ISOException.throwIt(EX_NEGATIVE_CREDIT);
-		} else {
-			balance = (short) (balance + cr);
-		}
-	} 
-}
+		} else {balance = (short) (balance + cr);}}}
 \end{lstlisting}
@@
 add (apdu) = 
-VAR 
-	bf, cr, apdu_state, res
+VAR 	bf, cr, apdu_state, res
 IN
@@
     	apdu_state  <-- apdu.getCurrentState;
-
 	IF (apdu_state = STATE_INITIAL) 
-	THEN
-		res <-- apdu.setIncomingAndReceive                
+	THEN res <-- apdu.setIncomingAndReceive                
 	ELSIF (OFFSET_CDATA >= 0 &
@@
 		cr  <-- Util.getShort(bf, OFFSET_CDATA);
-
 		IF not (cr <= MAX_BALANCE)
-		THEN 
-			ISOException.throwIt(EX_CREDIT_EXCEEDED)
+		THEN ISOException.throwIt(EX_CREDIT_EXCEEDED)
 		ELSIF not (sum_short (balance, cr) <= MAX_BALANCE)
-		THEN 
-			ISOException.throwIt(EX_BALANCE_EXCEEDED)
+		THEN ISOException.throwIt(EX_BALANCE_EXCEEDED)
 		ELSIF not (cr > 0)
-		THEN 
-			ISOException.throwIt(EX_NEGATIVE_CREDIT) 
-		ELSE  
-			balance := sum_short(balance, cr)
+		THEN ISOException.throwIt(EX_NEGATIVE_CREDIT) 
+		ELSE  balance := sum_short(balance, cr)
 		END
@@
 The transition to a defensive style of programming is part of the
-BSmart method, and implemented by the BSmart Full Function
-component. This component generates a B refinement of the original
+BSmart method and tool, which generates a B refinement of the original
 specification where pre-conditions are transformed into conditionals
@@
   http://www.oracle.com/technetwork/java/javame/javacard/download/overview/index.html).}, is composed of 14
-packages, containing 93 classes and interfaces, as shown in
-Table~\ref{tab:packages}. Their interfaces have all been modeled in B,
+packages, containing 93 classes and interfaces. Their interfaces have all been modeled in B,
 and the resulting models are available at the KitSmart project
@@
 restrictions of our models:
-
-\begin{table}[h]
-\begin{center}
-\caption{Packages of the Java Card API (2.2.2) and the number of classes in each package.}
-\begin{tabular}{|l|c||l|c|}
- \hline
-
-\textbf{\small{Package}} & \textbf{\small{classes}} & \textbf{\small{Package}} & \textbf{\small{classes}}\\ \hline \hline
-\small{java.io} & \small{1} & \small{javacardx.biometry} & \small{5} \\ \hline
-\small{java.lang} & \small{12} & \small{javacardx.crypto }& \small{2} \\ \hline
-\small{java.rmi} & \small{2} & \small{javacardx.external} & \small{3} \\ \hline
-\small{javacard.framework} & \small{19} & \small{javacardx.framework.math} & \small{3} \\ \hline
-\small{javacard.framework.service} & \small{8} & \small{javacardx.framework.tlv} & \small{7} \\ \hline
-\small{javacard.security} & \small{27} & \small{javacardx.framework.util} & \small{2} \\ \hline
-\small{javacardx.apdu} & \small{1} & \small{javacardx.framework.util.intx} & \small{1} \\
-
- \hline
-\end{tabular}
-\label{tab:packages}
-\end{center}
-\end{table}
+%
+%\begin{table}[h]
+%\begin{center}
+%\caption{Packages of the Java Card API (2.2.2) and the number of classes in each package.}
+%\begin{tabular}{|l|c||l|c|}
+% \hline
+% 
+%\textbf{\small{Package}} & \textbf{\small{classes}} & \textbf{\small{Package}} & \textbf{\small{classes}}\\ \hline \hline
+%\small{java.io} & \small{1} & \small{javacardx.biometry} & \small{5} \\ \hline
+%\small{java.lang} & \small{12} & \small{javacardx.crypto }& \small{2} \\ \hline
+%\small{java.rmi} & \small{2} & \small{javacardx.external} & \small{3} \\ \hline
+%\small{javacard.framework} & \small{19} & \small{javacardx.framework.math} & \small{3} \\ \hline
+%\small{javacard.framework.service} & \small{8} & \small{javacardx.framework.tlv} & \small{7} \\ \hline
+%\small{javacard.security} & \small{27} & \small{javacardx.framework.util} & \small{2} \\ \hline
+%\small{javacardx.apdu} & \small{1} & \small{javacardx.framework.util.intx} & \small{1} \\
+%
+% \hline
+%\end{tabular}
+%\label{tab:packages}
+%\end{center}
+%\end{table}
 
@@
   objects, a property stating that THROWABLE is a subset of OBJECT is
-  included, associated to the Throwable machine.
+  included.
 
@@
   Card. To model overloaded methods in B an extension to overloaded
-  method names was included. We then have equals, equals\_1,
-  equals\_2, etc.
+  method names was included.
+
   \end{itemize}
@@
 It presents two \texttt{java.lang} classes: \texttt{Object} and \texttt{Throwable}.
-\texttt{Object} is  superclass to almost all API classes, with its only method  \texttt{equals} that checks if an object (from which a reference is received as argument) is the same as the current one.
+\texttt{Object} is  superclass to  all API classes, with its only method  \texttt{equals} that checks if an object (from which a reference is received as argument) is the same as the current one.
 \texttt{Throwable} extends \texttt{Object}, but does not define any other method nor redefines \texttt{equals}.
@@
 MACHINE Object
-SEES  ObjectSet
+SEES  javaLangClasses 
 OPERATIONS
  	result <-- equals(obj) =
-	PRE obj : OBJECT   /* declared in the ObjectSet machine */
-	THEN
-		result :: {TRUE, FALSE}
-	END
-END
+	   PRE obj : OBJECT  /* declared in javaLangClasses */
+	   THEN result :: {TRUE, FALSE}
 
-MACHINE Throwable
-/* This class extends java.lang.Object */
+MACHINE Throwable  /* extends java.lang.Object */
 SEES
-    Object,
-    ThrowableSet
-END
+    Object,  javaLangClasses
 
-MACHINE ObjectSet
-SETS
-     OBJECT   
-END
-
-MACHINE ThrowableSet
-SEES 
-    ObjectSet    
-CONSTANTS  THROWABLE
-PROPERTIES THROWABLE <: OBJECT
-END
-
+MACHINE javaLangClasses
+SETS  OBJECT   
+CONSTANTS  THROWABLE , ...			
+PROPERTIES  THROWABLE <: OBJECT & ...
 \end{lstlisting}
@@
 
-\section{Using the model in BSmart}
+\section{Using the model in BSmart}\label{sec:using-the-API}
 
@@
 defines operations to start the process of receiving (from host) and
-sending (to host) information through the \emph{apdu buffer}. The
-preconditions of these sending/receiving operations aid the
-verification process, since the specifier must observe the correct
-order of operation calls. The \emph{Util} component contains useful
-operations to effectively get and put data in the buffer. When
+sending (to host) information through the \emph{apdu buffer}.  
+The \emph{Util} component contains useful
+operations to get and put data in the buffer. When
 translating to \emph{Java Card} the operation calls are translated
@@
 
-\section{Conclusions and future work}
+As a case study on the use of this API on the formal development of Java Card applications, part of the electronic passport specification\footnote{Available at http://www.icao.int.} is being modeled and developed in BSmart, showing that it is feasible to have a Java Card-like B refinement  such as the one presented in Listing \ref{lis:addimplementation}.
 
-With the current level of abstraction of the model and the fact that
-the Java Card Runtime Environment (JCRE) has not been modeled, some
-conditions which may lead to exception raising are not yet
-modeled. Being able to model them would make developed applications
-even more sure, as more static verification could be carried
-out. Continuation of this work then will enlarge the scope of the
-modeling to cope statically with a larger set of situations.
+\section{Conclusions and future work}\label{sec:conc}
+
+The B model described in this paper has as intent to be of use to Java Card application developers, either as additional (formal) documentation to the official one, or, better, as an imported library to be used in the formal development of applications with the B method.
+In section \ref{sec:api-spec} some characteristics and limitations of the work have been presented.
+Current work on using the API model to formally develop Java Card applications with B shows promising results.
+
+Among the limitations of this work, some may be dealt with by adapted tools and some will actually require some special attention from the developer, even with tool support, such as method overloading and overriding .
+These are the characteristics that are related to the fact that B is not object oriented and does not include exception raising and handling mechanisms.
+Some work on this last item can be seen on \cite{BR2003}, but it has not been implemented.
+
+With the current level of abstraction of the model and the fact that the Java Card Runtime Environment (JCRE) has not been modeled, different conditions which may lead to exception raising are not yet modeled.
+Being able to model them would make developed applications even more sure, as more static verification could be carried
+out. Continuation of this work then will enlarge the scope of the modeling to cope statically with a larger set of situations.
 

### 2012-04-20 trunk/papers/B2012-work/javacardAPIpaper.tex
correcoes de david

@@
 regularity of structure that led our group to define a method and
-implement a tool which specialize B for the development of such
-applications, providing greater support for the developer than pure B
+implement a tool which specializes B for the development of such
+applications, providing more focussed support for the developer than pure B
 tools \cite{atelierB,proB}. They are the BSmart method and tool,
@@
 Previous works have also dealt with B for Smart Cards,
-\cite{Bert:2003,JBtools,tatibouet:lncs,requetBtoC}, but none has, at
+\cite{Bert:2003,JBtools,tatibouet:lncs,requetBtoC}, but none has, to
 our knowledge, specified Java Card in B so that the formal development
 includes Java Card aspects. In the first versions of BSmart
-\cite{Gomes10}, these aspects were included by the code generator, and
-in JBTools \cite{JBtools}, the generated code needed to be edited by
+\cite{Gomes10}, these aspects were included by the code generator. In 
+contrast, with the JBTools~\cite{JBtools} approach, the generated code needed to be edited by
 the developer to include these aspects. These works lacked then the ability to formally verify the Java Card
-dependent aspects of the application. To be able to change this, a B
+dependent aspects of the application. To change this, a B
 specification of these aspects is needed, and this is the contribution
@@
 
-In the following we present some important aspects of the Java Card platform, in section \ref{sec:javacard}, and we introduce the B model of the Java Card API,  in section \ref{sec:api-spec}.
+In the following, we present some important aspects of the Java Card platform, in section \ref{sec:javacard}, and we introduce the B model of the Java Card API,  in section \ref{sec:api-spec}.
 We conclude with some notes on the use of the proposed model, in section \ref{sec:using-the-API}, and some conclusions and future work (in section \ref{sec:conc}).
@@
 The Java Card API is the part of the Java Card platform that a developer of any Java Card application must use.
-It defines the constituents of an applet, provides the means to deal with the APDU protocol and communicate data and commands between the host application and the card, and defines some basic services for security and cryptography implementation.
+It defines the constituents of an applet, provides the means to deal with the APDU protocol and to communicate data and commands between the host application and the card.
+It also defines some basic security services such as encryption.
 
@@
 (\texttt{add}). This simple example provides enough evidence of the
-importance of having a B model of the Java Card API.  B tools can directly 
-generate code for it in C, for instance, because it  is a deterministic specification and all data in
-it is implementable in C.  The
-resulting code would be very similar to the original B machine.
+importance of having a B model of the Java Card API.  B tools can
+directly generate code for it in C, for instance, because it is a
+deterministic specification and all data in it is implementable in C.
+The resulting code would be very similar to the original B machine.
 
@@
 The change of interface is also taken care of by the BSmart method, so
-that the development can be done without need to extend the B
+that the development can be done without extending the B
 method, as presented in~\cite{Gomes10}.
@@
 reason why the Java Card API needed to be the first Java Card
-component to be modeled in B. This is a common choice, as we can see in previous work on the formalization of 
-Java Card in JML \cite{pollBerg01} and OCL \cite{Larsson:2004}.
+component to be modeled in B. This is a common choice, as we can see
+in previous work on the formalization of Java Card in JML
+\cite{pollBerg01} and OCL \cite{Larsson:2004}.
 
@@
   code is not to be generated from the specification of the API, as we
-  are only creating an abstract layer to existing implementations; and
+  are only creating a model fo existing implementations; and
   (2) in B, only the abstract machines are seen by modules that build
@@
 
-  \item In the current version of the specification, states are not modeled. 
-  The B machines corresponding to Java Card classes or interfaces are stateless machines.
+\item In the current version of the specification, states are not
+  modeled.  The B machines corresponding to Java Card classes or
+  interfaces are stateless.
 
@@
   the class Object, in Java Card. In our model, we  have that the
-  machine Throwable SEES the machine Object, and to show that all throwables are indeed
+  machine Throwable SEES the machine Object, and to constrain that all throwables are indeed
   objects, a property stating that THROWABLE is a subset of OBJECT is
@@
 It presents two \texttt{java.lang} classes: \texttt{Object} and \texttt{Throwable}.
-\texttt{Object} is  superclass to  all API classes, with its only method  \texttt{equals} that checks if an object (from which a reference is received as argument) is the same as the current one.
+\texttt{Object} is a superclass to  all API classes, with a single method  \texttt{equals} that checks if an object (from which a reference is received as argument) is the same as the current one.
 \texttt{Throwable} extends \texttt{Object}, but does not define any other method nor redefines \texttt{equals}.
@@
 
-One delicate issue to be dealt with both by the user of the API
+One subtle issue to be dealt with both by the user of the API
 specification and the tools designed to work with it, is the use of
@@
 carried out.  In case of overriding, in particular, it will be either
-the user's responsibility to choose the B one which is in scope at
+the user's responsibility to choose the B operation in scope at
 this specific point in Java, or an extra verification step will be
@@
 
-Finally, the choices related to class derived types and return values were consequences of our focus on the interfaces of the Java Card API methods.
-The declaration of an abstract set for each class allows to model in B that a parameter of a particular class is expected or is to be returned by a method.
+Finally, the choices related to class derived types and return values
+were consequences of our focus on the interfaces of the Java Card API
+methods. The declaration of an abstract set for each class allows to
+model in B that a parameter of a particular class is expected or is to
+be returned by a method.
 
@@
 Among the limitations of this work, some may be dealt with by adapted tools and some will actually require some special attention from the developer, even with tool support, such as method overloading and overriding .
-These are the characteristics that are related to the fact that B is not object oriented and does not include exception raising and handling mechanisms.
+These are consequence of B not being object oriented and not providing exception raising and handling mechanisms.
 Some work on this last item can be seen on \cite{BR2003}, but it has not been implemented.
@@
 With the current level of abstraction of the model and the fact that the Java Card Runtime Environment (JCRE) has not been modeled, different conditions which may lead to exception raising are not yet modeled.
-Being able to model them would make developed applications even more sure, as more static verification could be carried
+Extending the model to include them would make developed applications more reliable, as more static verification could be carried
 out. Continuation of this work then will enlarge the scope of the modeling to cope statically with a larger set of situations.

### 2012-05-15 papers/ArticleZ80/source/paper.tex
ajustes

@@
 %\usepackage{supertabular} % used in B Symbols
-\geometry{a4paper} % or letterpaper (US) or a5paper or....
+%\geometry{a4paper} % or letterpaper (US) or a5paper or....
 
 \usepackage[english]{babel} %\usepackage{supertabular} % used in B Symbols
-\usepackage{hyperref} % add URL in bibliography
+%\usepackage{hyperref} % add URL in bibliography
 \usepackage{url} % allow add URL in document
@@
 \include{SymbolsB_AtelierB}
+
+%TODO: Verificar possíveis diferenças do FORMATO LNCS
 
@@
 
-\title{Experiences in Modelling a Microcontroller Instruction Set Using B}
+\title{Experience in Modeling a Microcontroller Instruction Set Using B}
 %
@@
 \hyphenation{re-gis-ters}
+
 %
@@
 
-\institute{ Federal Institute of Education, Science and Technology of Rio Grande do Norte, Natal RN 59015-000, Brazil
+\institute{Federal Institute of Education, Science and Technology of Rio Grande do Norte, Natal, RN, Brazil
   \and 
-		    Federal University of Rio Grande do Norte, Natal RN 59078-970, Brazil}
+  Federal University of Rio Grande do Norte, Natal, RN, Brazil}
 
@@
 
-  This paper describes an approach to model the functional aspects of
-  the instruction set of microcontrollers and several details about
-  the representation of microcontrollers.  Different models were
-  developed using the notation of the B method. This paper presents
-  specifically a B-based formal model of the Z80 instruction set.
-
-%  \paragraph[Keywords]{
-%   \textbf{Keywords}: Embedded Software, Simulation, B method and Verification%}
+  This paper presents a formal model, developed in B, of the Z80
+  instruction set. The formal model includes different libraries that
+  specify entities commonly used in microcontrolers and
+  microprocessors. Not only this model can be used for documentation
+  purposes, but also to simulate and test the assembly code. It is also
+  a fundamental component to extend the reach of the B method from the
+  specification of the functional requirements of a system to an assembly-level
+  implementation thereof.
+% Using this model-driven approach and the instruction set model, a small case
+% study to analyse a petroleum production test system in oil field was developed.
+%  \paragraph[Keywords]{ \textbf{Keywords}: Embedded Software, Simulation, B
+% method and Verification%}
 \end{abstract}
-
-%\keywords{ Validation, Verification, Microcontroller, Hardware}
-%
-%\setcounter{secnumdepth}{1}
-%\setcounter{chapter}{1}
-%\setcounter{section}{1}
 
@@
 of microcontrollers and even microprocessors; it has extensive
-documentation available; and it has been widely used.  A first,
+available documentation; and it has been widely used.  A first,
 partial, Z80 model presented in \cite{Valerio_SBMF09} was fully
-verified, but it uses constructs that make it impossible to animate.
-
-The new model supports animation in ProB\cite{proB}.  Essentially, the
+verified, but it used constructs that made it impossible to animate.
+
+The new model supports animation in ProB~\cite{proB}.  Essentially, the
 new model\footnote{The interested reader in more details is invited to
@@
 constructs for modularity, auxiliary libraries of basic modules were
-developed as part of the construction of microcontroller model. Such
-library has many definitions of concepts common in microcontrollers.
+developed as part of the construction of the microcontroller model. Such
+a library has many definitions of concepts common to microcontrollers.
 Besides the new Z80 model, a petroleum production test system was
-developed using Z80 assembly language to analyse the code simulation
-and verification process.
-
-Providing the formal model of an instruction set yields many
-benefits. The animation of Z80 model allows to simulate the execution
-assembly programs including support for interrupts and input and
-output instructions. Other possible uses of a formal model of a
-microcontroller instruction set include documentation, the
-construction of simulators, and can be possibly the starting point of
+developed using the Z80 assembly language to analyse the code simulation
+and the verification process.
+
+The formal model of an instruction set has several
+applications. Animation allows to simulate the execution of assembly
+programs including support for interrupts, input and output
+instructions. Other possible uses include documentation, the
+construction of simulators, and can possibly be the starting point of
 a verification effort for the actual implementation of a Z80
@@
 it takes to execute an instruction, to prove lower and upper bounds on
-the execution time of a routine.  Two aspects of these formalization
+the execution time of a routine.  Two aspects of such formalizations
 are of particular importance to us: provide a more solid documentation
-artifact for microcontrollers, build a reference model for a formal
+artifact for microcontrollers and build a reference model for a formal
 compilation approach in B.
 
-\subsection{Problems in formalizing instruction set}
-% Compact version
-
-In general, the manuals of microcontrollers can be more suited for \marginpar{DD: o título da sub-seção não condiz com seu conteúdo.
-Não cita nenhum problema na formalização.
-Não vejo muito bem o que ele traz de novo que já não foi dito anteriormente.}
-developers in assembly language. Many times, each instruction of
-microcontrollers is shown in its official manual using textual
-description, math description, examples, encoding, number of cycles
-and other information. Sometimes, the instruction descriptions have
-an own notation, some informations are organized in different sections
-of document and the textual description are not standardized. This
-does not allow the interpretation by parser.  The manuals also have
-semi-formal and informal descriptions, these descriptions may make
-mistakes.  The official manual of the microcontroller Z80 has also
-several problems that were identified over
-time~\cite{UndocumentedZ80}.
-
-A solution to avoid ambiguity and inconsistencies is to specify
-formally the assembly instruction set, this creates a pattern of
-representation and validates properties of instructions. The formal
-model also restricts the definitions to correct typing and uses only
-well-defined expressions. Furthermore, the developer can use with B
-method different levels of abstraction and add more specific details
-in refined models.
+It is often the case that the official documentation of an instruction
+uses textual description, mathematical formulas, examples,
+etc. Sometimes, the instruction descriptions have a specific notation,
+some informations are organized in different sections of a document and
+the textual descriptions are not standardized.  The manuals also have
+semi-formal and informal descriptions, these descriptions may leave
+errors, inconsistency and ambiguity.  For example, in the official
+manual of the microcontroller Z80 \cite{Z80_manual}, the semantics of
+an instruction (in particular its effects on the flag register) are
+described informally, and this description is scattered in
+different pages.  Furthermore, the official manual of the
+microcontroller Z80 has also several problems that were identified
+over time~\cite{UndocumentedZ80}.
+
+A solution to avoid ambiguity and inconsistencies is to formally
+specify the assembly instruction set, making it possible to prove
+properties of the instructions. The formal model also restricts the
+definitions to correct typing and uses only well-defined
+expressions. Furthermore, the developer may use the refinement
+capabilities of the B method to specify at different levels of
+abstraction and add more specific details or useful restrictions in
+refined models.
 
@@
 
-This paper is structured as it follows.  Section~\ref{sec:models}
+This paper is structured as follows.  Section~\ref{sec:models}
 presents the elementary libraries and the modeling of some elements
 common to microcontrollers.  Section~\ref{sec:z80} presents the B
-model of the Z80 instruction set.  Related works are discussed in
+model of the Z80 instruction set.  Related work are discussed in
 section~\ref{sec:relatedworks}. Finally, the last section is devoted
@@
 
-We have developed a reusable set of basic definitions to model hardware concepts and data types
-concepts. The models contains definitions to represent the: registers, interruptions, input and output
-ports, memory and instructions. These definitions are grouped into separated development projects
-and are available as libraries.
-
-Thus the workspace is composed of modules group and libraries: 
-\begin{itemize}
-
-\item \textbf{Power}: it has the basics definitions to help the theorems
-prover about simple calculus of power, especially power of two. It contains the definition of power function
-and it has constants, that represent all required results of power function.
-
-\item \textbf{Hardware Library}: it has the definition of bit vectors with size 8 and 16, basic functions to manipulate bit vectors and important assertions.
-This is presented in sections \ref{subsec:HardwareLibrary1}, \ref{subsec:HardwareLibrary2} and \ref{subsec:HardwareLibrary3}.
-
- \item \textbf{Types Library}: it has the definition of naturals and integers represented with  8 bits and 16 bits, basic conversion functions between the types and bit vectors and important assertions.
-This is presented in section \ref{subsec:Types}.
-
- \item \textbf{Platform}: it has the definition functions of arithmetic logic unit,  memory unit, registers and assembly instructions. This is presented in section \ref{sec:z80}.
-
-\end{itemize}
-
- The corresponding dependency
-diagram is depicted in Figure~\ref{fig:hardware-definition-graph}; information specific to each project is
-presented in the following.
+We have developed a reusable set of basic definitions to model
+hardware concepts and data types concepts. The models contain
+definitions to represent: registers, interruptions, input and
+output ports, memory and instructions. These definitions are grouped
+into separated development projects and are available as libraries.
+Thus the workspace is composed of the following: 
+\begin{description}
+
+\item[Power] contains the standard definition of exponentiation, as
+  well as some basic properties, e.g. values of different powers of
+  two. Given this definition, these properties have been proved
+  interactively with the Atelier-B prover.
+
+\item[Hardware Library] defines bit vectors with size 8 and 16, basic
+  functions to manipulate bit vectors and important
+  assertions. Details are given in sections
+  \ref{subsec:HardwareLibrary1}, \ref{subsec:HardwareLibrary2} and
+  \ref{subsec:HardwareLibrary3}.
+
+\item[Types Library] defines sets of naturals and integers represented
+  with 8 bits and 16 bits, basic conversion functions between those
+  types and bit vectors, and important assertions. It is presented in
+  section \ref{subsec:Types}.
+
+\item[Platform] defines functions of the arithmetic logic unit, memory
+  unit, registers and assembly instructions. Section \ref{sec:z80}
+  provides additional details.
+
+\end{description}
+
+The dependencies between those libraries are depicted in
+Figure~\ref{fig:hardware-definition-graph}.
 
@@
 type for bits, logical operations on bits (negation, conjunction,
-disjunction, exclusive disjunction), as well as a conversion function
-from Boolean to bit.
-
-First, bits are modeled as a set of integers: $\textit{BIT} =
+disjunction, exclusive disjunction), and a conversion function from
+Booleans to bits.
+
+We model bits as a set of integers: $\textit{BIT} =
 \textit{0..1}$. The negation is an unary function on bits and it is
 defined as:
-
-$
+$$
 \begin{array}{l}
@@
 \end{array}
-$
+$$
 %\forall ( \textit{bb}). (\textit{bb} \in \textit{BIT} \implies {bit\_not}(\textit{bb}) =1-\textit{bb})\\
@@
 users of the library to develop proofs:
-
-$
+$$
 \begin{array}{l}
@@
 \end{array}
-$
-
+$$
 Conjunction is a binary function on bits and it is defined as:
-
-$
+$$
 \begin{array}{l}
@@
 \end{array}
-$
-
+$$
 The module provides the following lemmas for commutativity and associativity:
-
-$
+$$
 \begin{array}{l}
@@
 \end{array}
-$
-
+$$
 The module provides definitions of $\textit{bit\_or}$ (disjunction)
@@
 those operators. These are standard and their expression in B is
-similar as for $\textit{bit\_and}$, they are thus omitted from this paper.
+similar to $\textit{bit\_and}$.
 
 Finally, the conversion from Boolean to bit is simply defined as:
-
-$
+$$
 \begin{array}{l}
@@
 \end{array}
-$
-
-Observe that all the lemmas that are provided in this module have been
-mechanically proved by the theorem prover included with our B
-development environment. None of these proofs requires human insight.
+$$
+All the lemmas that are provided in this module have been
+automatically proved by the prover of Atelier-B.
 
@@
 
-Sequences are pre-defined in B, as functions whose the domain is an
-integer range with lower bound 1 (one). Indices in bit vectors usually
-range from 0 (zero) upwards and the model we propose obeys this
-convention by making an one-position shift where necessary. This shift
-is important to use the predefined functions of sequences. We thus
-define bit vectors as non-empty sequences of bits, and
-$\textit{BIT\_VECTOR}$ is the set of all such sequences:
-$\textit{BIT\_VECTOR} = \seq1(\textit{BIT})$. The $\textit{BIT\_VECTOR}$ is an infinite set and it hinders the animation in ProB. 
-To avoid this obstruction, the specialized types (\textit{BYTE} and \textit{BV16}) cannot reference directly \textit{BIT\_VECTOR}.
+Sequences are predefined objects in B and are natural candidates to
+represent bit vectors. In B, sequences are functions whose domain
+is an integer range with lower bound 1 (one). Indices in bit vectors
+usually range from 0 (zero) upwards and the model we propose obeys
+this convention by making an offset where necessary, so that
+predefined functions and proof rules may be used. We define bit
+vectors as non-empty sequences of bits: $\textit{BIT\_VECTOR} =
+\seq1 (\textit{BIT})$. As $\textit{BIT\_VECTOR}$ is an infinite set,
+this definition hinders the animation in ProB.  As a workaround, we
+defined specialized types \textit{BYTE} and \textit{BV16} without
+reference to \textit{BIT\_VECTOR}.
 
@@
 
-We also define two functions $\textit{bv\_set}$ and $\textit{bv\_clear}$ that, given a bit vector, and a
-position of the bit vector, return the bit vector resulting from setting the corresponding position to 0
-or to 1, and a function $\textit{bv\_get}$ that, given a bit vector, and a valid position, each one
-returns the value of the bit at that position. Only the first definition is shown here:
-
-$
+We also define two functions $\textit{bv\_set}$ and
+$\textit{bv\_clear}$ that, given a bit vector, and a position in a
+bit vector, returns the bit vector resulting from setting the
+corresponding position to 1 and 0, and a function $\textit{bv\_get}$
+that, given a bit vector and a valid position, returns the value of
+the bit at that position. Only the first definition is shown here:
+$$
 \begin{array}{l}
-\textit{bv\_set} \in \textit{BIT\_VECTOR} \times \nat \fun \textit{BIT\_VECTOR} \land \textit{bv\_set} =\\
+\textit{bv\_set} \in \textit{BIT\_VECTOR} \times \nat \fun \textit{BIT\_VECTOR} \land \\ \textit{bv\_set} =
 \lambda v, n . (v \in \textit{BIT\_VECTOR} \land n \in \nat \land n <\textit{size}(v)
@@
 \end{array}
-$
-
-The function $bv\_catenate$ takes as parameters two bit vectors $v$ and $w$, and returns the result of the
-concatenation of $v$ and $w$, such that $v$ constitutes the most significant part of the result.
+$$
+
+Function $bv\_catenate$ takes as parameters two bit vectors $v$ and
+$w$, and returns the concatenation of $v$ and $w$, such
+that $v$ constitutes the most significant part of the result.
 
@@
 $\textit{bit\_and}$, $\textit{bit\_or}$ and $\textit{bit\_xor}$. Only
-the first two are presented here. Observe that the domain of the
-binary operators is restricted to pairs of bit vectors of the same
-length:
-
-$
+the first two are presented here. The domain of the binary operators
+is restricted to pairs of bit vectors of the same length:
+$$
 \begin{array}{l}
@@
 \end{array}
-$
-
+$$
 We provide several lemmas on bit vector operations. These lemmas
@@
 
-Bit vectors of length 8 are bytes. They are a common entity in
-hardware design. We provide the following definitions:
+Bytes, i.e. bit vectors of length 8, are a common entity in hardware
+design. We provide the following definitions:
 
@@
 
-The $\textit{BYTE\_INDEX}$ is the domain of the functions modeling bytes. It starts at 1 to obey a
-definition of sequences from B. However, it is common in hardware architectures to start indexing from
-zero. The definition $\textit{PHYS\_BYTE\_INDEX}$ is used to provide functionalities obeying this
-convention. The $\textit{BYTE}$ type is as a specialized type of $\textit{BIT\_VECTOR}$, but it has a size
-limit. Other specific definitions are provided to facilitate further modeling: the type $\textit{BV16}$
-is created for bit vector of length 16 in a similar way.
+$\textit{BYTE\_INDEX}$ is the domain of the functions modeling
+bytes. It starts at 1 to be compatible with sequences. However, since
+the hardware community starts indexing from zero, definition
+$\textit{PHYS\_BYTE\_INDEX}$ provides functionalities obeying this
+convention. Type $\textit{BYTE}$ is as a specialized type of
+$\textit{BIT\_VECTOR}$, with a size constraint. Other specific
+definitions are provided to facilitate further modeling: the type
+$\textit{BV16}$ is created for bit vectors of length 16 in a similar
+way.
 
@@
 
-Bit vectors are used to represent and combine numbers: integer ranges (signed or unsigned). Therefore, our
-library includes functions to manipulate such data, for example, the function $\textit{bv\_to\_nat}$ that
+Bit vectors are used to represent and combine numbers: integer ranges
+(signed or unsigned). Therefore, our library includes functions to
+manipulate such data, for example, the function $\textit{bv\_to\_nat}$
 maps bit vectors to natural numbers:
-
-$
+$$
 \begin{array}{l}
@@
 \end{array}
-$
-
-An  associated lemma is: $\forall n . (n \in \nat_1 \implies \textit{bv\_to\_nat}(\textit{nat\_to\_bv}(n)) = n)$
-
-\subsection{Basics data types}
-
-The instruction set architecture usually have common bit vector type and integer type.
-In table~\ref{tab:types}, the first three are placed in the hardware library and the last four types are placed in the integer types library. 
-Each type module has functions to manipulate and convert its data. There are seven common basic data types represented by modules.
-
-\begin{table}
+$$
+
+An important lemma associated to this definition is: 
+$$\forall n . (n \in \nat_1 \implies \textit{bv\_to\_nat}(\textit{nat\_to\_bv}(n)) = n).$$
+
+\subsection{Basic data types}
+
+The definition of instruction sets usually have bit vector and integer
+types as listed in table~\ref{tab:types}. For each type, a number of
+functions are defined, including conversion between different types.
+
+\begin{table}[h]
 \caption{Details of basic data types}
@@
 \begin{center}
-\begin{tabular}{|c|c|c|c|}
+\begin{tabular}{|c|c|c|c|c|}
 \hline
- $Type\ Name$ & $Range$ & $Physical\ Size $  & $Associated\ Proof\ Obligations$ \\\hline
- BIT & 0..1 & 1 bit &   118 \\\hline %49 + 69 = 118 \\\hline
- BYTE & -- & 1 bytes &  153\\\hline%17 + 136 = 153\\\hline
- BV16 & -- & 2 bytes &  75 \\ \hline%5 + 70 = 75 \\ \hline
- UCHAR & 0..255 & 1 byte &  30 \\\hline %4 + 26 = 30 \\\hline
- SCHAR & -128..127 & 1 byte & 26 \\\hline %2 + 24 = 26 \\\hline
- USHORT & 0..65.535 & 2 byte & 62 \\\hline % 4 + 58 = 62 \\\hline
- SSHORT & -32.768..32.767 & 2 byte & 58\\\hline %  2 + 56 = 58\\\hline
-%  &  & Total & 522\\\hline %  2 + 56 = 58\\\hline
+ \textit{Type\ Name} & \textit{Range} & \textit{Physical Size}  & \textit{Num. POs} & \textit{Library} \\\hline
+ BIT & 0..1 & 1 bit &   118 & Hardware \\\hline 
+ BYTE & -- & 1 bytes &  154 & Hardware \\\hline
+ BV16 & -- & 2 bytes &  75 & Hardware \\ \hline
+ UCHAR & 0..255 & 1 byte &  30 & Types \\\hline
+ SCHAR & -128..127 & 1 byte & 30 & Types \\\hline 
+ USHORT & 0..65535 & 2 byte & 62 & Types \\\hline
+ SSHORT & -32768..32767 & 2 bytes & 62 & Types \\\hline 
 \end{tabular}
@@
 
-Usually, each type module just needs to instantiate concepts that were already defined in the hardware
-modeling library.  For example, the function $\textit{bv\_to\_nat}$ from bit vector arithmetic is
-specialized to $\textit{byte\_uchar}$. As the set $\textit{BYTE}$ is a subset of the
-$\textit{BIT\_VECTOR}$, this function can defined as follows:
-
-$
+Usually, each type module just needs to specialize definitions that
+are given in the hardware modeling library.  For example, the function
+$\textit{bv\_to\_nat}$ from bit vector arithmetic is specialized to
+$\textit{byte\_uchar}$. As the set $\textit{BYTE}$ is a subset of the
+$\textit{BIT\_VECTOR}$, this function can be defined as follows:
+$$
 \begin{array}{l}
@@
 \end{array}
-$
-
-The definitions of the library types reuse the basic definitions from the hardware library. This provides
-greater confidence and facilitates the proof process, because the prover can reuse the previously defined
-lemma.
+$$
+
+The definitions of the library types reuse the basic definitions from
+the hardware library. This provides greater confidence and facilitates
+the proof process, because the prover can reuse the previously defined
+lemmas.
 
@@
 
- We also created the following lemmas:
-
- $
+We also created lemmas on pairs of dual conversion functions, such as:
+$$
  \begin{array}{l}
@@
  \end{array}
- $
-
-Similarly, several other general functions and lemmas were created for all other data types.
+$$
+
 %[NEW PARAGRAPH-TO-REVIEW]
@@
 
-Nearly 50\% of all these proof obligations shown in the table~\ref{tab:types} 
-are verified almost automatically. Although many others proof obligations are difficult to verify,
-because they are functions of type conversion and contain arithmetic expressions difficult 
-to manipulate with the interactive prover of AtelierB.
-
-\subsection{Verification in Hardware Library and Types Library} 
+Nearly 50\% of all these proof obligations shown in the
+Table~\ref{tab:types} are verified automatically or with few proof
+commands\footnote{The proof commands are steps that help the prover to find the proof.} .
+However many others proof obligations are difficult to verify, as they involve functions
+that contain arithmetic expressions that are difficult to manipulate with the interactive
+prover of AtelierB. However, all theses 531 proof obligations were verified completely.
+
+\subsection{Verification of hardware library and types library} 
 \label{sec:VerificationHardwareLibrary}
 
-\input{VerificationLibraryHardware}
+This section presents the verification of the aforementioned modules
+using AtelierB provers.  By default, AtelierB~\cite{atelierB} does not
+generate proof obligations to check the type of functions defined in
+the $PROPERTIES$ clause of a machine. For example, the definition $inc
+\in \{0,1\} \fun \{0,1\} \land inc = \lambda(x).( x \in \{0,1\} | x +
+1)$, does not create a proof obligation to check the consistency of
+the definition. However, this definition is inconsistent since
+$inc(1)=2$ and $ 2 \not\in \{0,1\}$.
+
+To avoid such inconsistencies, we state separately the required
+properties in the \textit{ASSERTIONS\/} clause: this causes
+corresponding proof obligations to be generated and verified. However,
+we have not been able to build the corresponding proofs using the 
+rule base included in the free distribution of AtelierB. So we
+formalized new proof rules to check these properties in AtelierB and
+used them in the interactive prover of AtelierB.
+
+The hardware library has many conversion functions with
+arithmetic,  for example, Bit\_vector to Natural, Natural to
+Bit\_vector, Integer to Bit\_vector and Bit\_vector to Integer. These
+functions are organized in modules and each module contains auxiliary
+functions for a data type. The modules are represented by gray
+rectangles in Figure \ref{DiagramTypesRules}.
+
+\begin{figure}[he]
+\centering
+\includegraphics[width=3.in]{images/Diagram_Types_and_Rules.pdf}
+\caption{Conversion functions}
+\label{DiagramTypesRules}
+\end{figure}
+
+The conversion functions need to satisfy two properties: bijectivity
+and equality to the reverse of the dual conversion function. Three
+such proof obligations were verified automatically with AtelierB: (\it
+uchar\_schar\rm \ $\in$ \textit{UCHAR} $\bij$ \textit{SCHAR}), (\it
+schar\_uchar\rm \ $\in$ \textit{SCHAR} $\bij$ \textit{UCHAR}) and (\it
+byte\_bv16\rm \ $=$ \it bv16\_byte\rm $^{-1}$).
+%\begin{enumerate}
+%
+%\item (\it uchar\_schar\rm \ $\in$ \textit{UCHAR} $\bij$ \textit{SCHAR})
+%\item (\it schar\_uchar\rm \ $\in$ \textit{SCHAR} $\bij$ \textit{UCHAR})
+%\item (\it byte\_bv16\rm \ $=$ \it bv16\_byte\rm $^{-1}$) 
+%
+%\end{enumerate}
+For the remaining proof obligations, we again had to create new proof
+rules. There are three kinds of proof rules, that we illustrate by giving
+one example each:
+
+%Four conditions are needed to be a bijective function:
+%
+%\begin{description}
+%\item $func$ is a function:\\
+%$\forall$ \rm (\it xx\rm ,\it yy\rm ,\it zz\rm )\rm .\rm ( \it xx  $\in$  \bf dom\rm (\it func\rm )  $\land$  \it yy $\in$  \bf ran\rm (\it func\rm )  $\land$ \\ \it zz $\in$  \bf ran\rm (\it func\rm )  $\land$ \rm ( \it xx $\mapsto$  \it yy\rm )  $\in$  \it func  $\land$  \rm ( \it xx $\mapsto$  \it zz\rm )  $\in$  \it func  $\implies$  \it yy\rm =\it zz \rm )
+%
+%\item $func$ is an injective function:\\
+%$\forall$ \rm (\it vv\rm ,\it xx\rm )\rm .\rm ( \it vv $\in$  \bf dom\rm (\it func\rm )  $\land$  \it xx  $\in$  \bf dom\rm (\it func\rm )  $\implies$  \it func\rm (\it vv\rm )\rm =\it func\rm (\it xx\rm ) \rm )
+%
+%\item $func$ is a total function:\\
+%$\forall$ \it xx\rm .\rm ( \it xx  $\in$  \bf dom\rm (\it func\rm )  $\implies$   $\exists$ \it yy\rm .\rm ( \it yy  $\in$  \bf ran\rm (\it func\rm )  $\land$  \it func\rm (\it xx\rm )\rm =\it yy \rm )\rm )
+%
+%\item $func$ is surjective function:\\
+% $\forall$ \it yy\rm .\rm ( \it yy $\in$  \bf ran\rm (\it func\rm ) $\implies$   $\exists$ \it xx\rm .\rm ( \it xx  $\in$  \bf dom\rm (\it func\rm ) $\land$  \it func\rm (\it xx\rm )\rm =\it yy \rm )\rm )
+%
+%
+%\end{description} 
+%
+%Here, two conversion functions between bit vectors and the natural are presented.
+These functions are finite and convert from \textit{BYTE} to \textit{UCHAR} and vice versa.\textit{UCHAR} is a range between 0 and 255 and \textit{BYTE} is a bit vector with eight elements enumerated starting from 1.
+We need to prove that $byte\_uchar$  and $uchar\_byte$  are bijective functions and  $byte\_uchar = uchar\_byte^{-1}$.
+These functions are defined as follows:
+%\begin{description}
+%
+%
+%\item\it byte\_uchar \rm = $\lambda$  \rm ( \it v0 \rm ) \rm .
+\rm ( \it v0  $\in$  \it BYTE  $\mid$ \hspace*{0.10in}\it 2$^{7}$ $\times$ \it v0\rm (\rm 8\rm )\rm +\it 2$^{6}$  $\times$ \it v0\rm (\rm 7\rm )\rm + \it 2$^{5}$ $\times$ \it v0\rm (\rm 6\rm )\rm +\it 2$^{4}$ $\times$ \it v0\rm (\rm 5\rm ) \rm + \it 2$^{3}$  $\times$  \it v0\rm (\rm 4 \rm ) \rm + \it 2$^{2}$  $\times$ \it v0\rm (\rm 3\rm ) \rm + \it 2$^{1}$  $\times$  \it v0\rm (\rm 2\rm ) \rm + \it 2$^{0}$  $\times$  \it v0\rm (\rm 1\rm )\rm)
+%
+%\item\it uchar\_byte \rm = $\lambda$  \rm ( \it v0 \rm ) \rm .
+\rm ( \it v0  $\in$  \it UCHAR $\mid$ \rm [\rm (\it v0   $\mod$ \it 2$^{1}$\rm ) $\div$ \it 2$^{0}$\rm , \rm (\it v0  $\mod$  \it 2$^{2}$\rm ) $\div$ \it 2$^{1}$\rm , \rm (\it v0  $\mod$  \it 2$^{3}$\rm ) $\div$ \it 2$^{2}$\rm , \rm (\it v0  $\mod$  \it 2$^{4}$\rm ) $\div$ \it 2$^{3}$\rm , \rm (\it v0  $\mod$  \it 2$^{5}$\rm ) $\div$ \it 2$^{4}$\rm , \rm(\it v0  $\mod$  \it 2$^{6}$\rm ) $\div$ \it 2$^{5}$\rm ,\rm (\it v0  $\mod$  \it 2$^{7}$\rm ) $\div$ \it 2$^{6}$\rm , \rm (\it v0  $\mod$  \it 2$^{8}$\rm ) $\div$ \it 2$^{7}$\rm \rm ] \rm)
+%
+%\end{description}
+%
+%The equivalent definitions using B sequence notation and set notation are:\\
+%\it uchar\_byte \rm = \rm \{ \rm 0 $\mapsto$ \rm \rm [\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm \rm ]\rm, \ldots  , \rm 2\rm 5\rm 5 $\mapsto$ \rm \rm [\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm \rm ]\rm \}  \\
+%\it byte\_uchar \rm =  \rm \{\rm \rm [\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm \rm ] $\mapsto$ \rm 0\rm, \ldots , \rm \rm [\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm \rm ] $\mapsto$ \rm 2\rm 5\rm 5 \rm \} \rm
+%
+%\paragraph{B Rules:}
+%These rules represent that  $byte\_uchar$  and $uchar\_byte$  are bijective functions and  $byte\_uchar = uchar\_byte^{-1}$. These rules are used on interactive prover of Atelierb.
+
+\begin{enumerate}
+
+\item For the proof obligation (\it byte\_uchar\rm \ $\in$ \textit{BYTE} $\bij$
+\textit{UCHAR}), the proof rule is:\\
+$\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}  $\mid$  \rm 1\rm 2\rm 8 $\times$ \it v0\rm (\rm 8\rm )\rm +\rm 6\rm 4 $\times$ \it v0\rm (\rm 7\rm )\rm +\rm 3\rm 2 $\times$ \it v0\rm (\rm 6\rm )\rm +\rm 1\rm 6 $\times$ \it v0\rm (\rm 5\rm )\rm +\rm 8 $\times$ \it v0\rm (\rm 4\rm )\rm +\rm 4 $\times$ \it v0\rm (\rm 3\rm )\rm +\rm 2 $\times$ \it v0\rm (\rm 2\rm )\rm +\rm 1 $\times$ \it v0\rm (\rm 1\rm )\rm ) \rm $\in$ \rm (\rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}\rm )  $\bij$  \rm 0 $\upto$ \rm 2\rm 5\rm 5 \\
+
+\item For the proof obligation (\it uchar\_byte\rm \ $\in$ \textit{UCHAR} $\bij$
+\textit{BYTE}), the proof rule is:\\ $\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 0$\upto$255  $\mid$  \rm \rm  [(\it v0  $\mod$  \rm 2)$\div$1,(\it v0  $\mod$  \rm 4)$\div$\rm 2\rm,(\it v0  $\mod$  \rm 8)$\div$\rm 4\rm,(\it v0  $\mod$  \rm 16)$\div$\rm 8\rm,(\it v0  $\mod$  \rm 32)$\div$\rm 1\rm 6\rm,(\it v0  $\mod$  \rm 6\rm 4)$\div$\rm 3\rm 2\rm,(\it v0  $\mod$  \rm 1\rm 2\rm 8)$\div$\rm 6\rm 4\rm,(\it v0  $\mod$  \rm 2\rm 5\rm 6)$\div$\rm 1\rm 2\rm 8\rm \rm ]\rm ) \rm   $\in$  \rm 0 $\upto$ \rm 2\rm 5\rm 5   $\bij$ \rm (\rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}\rm )\\
+
+\item For the proof obligation (\it byte\_uchar\rm  =  \it
+uchar\_byte\rm$^{-1}$), the proof rule is:\\ $\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}  $\mid$  \rm 1\rm 2\rm 8 $\times$ \it v0\rm (\rm 8\rm )\rm +\rm 6\rm 4 $\times$ \it v0\rm (\rm 7\rm )\rm +\rm 3\rm 2 $\times$ \it v0\rm (\rm 6\rm )\rm +\rm 1\rm 6 $\times$ \it v0\rm (\rm 5\rm )\rm +\rm 8 $\times$ \it v0\rm (\rm 4\rm )\rm +\rm 4 $\times$ \it v0\rm (\rm 3\rm )\rm +\rm 2 $\times$ \it v0\rm (\rm 2\rm )\rm +\rm 1 $\times$ \it v0\rm (\rm 1\rm )\rm )=($\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 0$\upto$255  $\mid$  \rm \rm  [(\it v0  $\mod$  \rm 2)$\div$1,(\it v0  $\mod$  \rm 4)$\div$\rm 2\rm,(\it v0  $\mod$ \rm 8) $\div$\rm 4\rm,(\it v0  $\mod$  \rm 16)$\div$\rm 8\rm,(\it v0  $\mod$  \rm 32)$\div$\rm 1\rm 6\rm,(\it v0  $\mod$ \rm 6\rm 4)$\div$\rm 3\rm 2\rm,(\it v0  $\mod$  \rm 1\rm 2\rm 8)$\div$\rm 6\rm 4\rm,(\it v0  $\mod$  \rm 2\rm 5\rm 6)$\div$\rm 1\rm 2\rm 8\rm \rm ]\rm )\rm)$^{-1}$
+\end{enumerate}
+
+Eight proof obligations require that conversion functions are finite. %: \it
+% byte\_uchar $\in$ $\finset$\rm (\it byte\_uchar\rm ); \it uchar\_byte  $\in$
+% $\finset$\rm (\it uchar\_byte\rm ); \it ushort\_bv16 $ \in \finset$\rm (\it
+% ushort\_bv16\rm ); \it bv16\_ushort $ \in \finset$\rm (\it bv16\_ushort\rm ); \it
+% byte\_schar $ \in \finset$\rm (\it byte\_schar\rm ); \it schar\_byte $ \in
+% \finset$\rm (\it schar\_byte\rm ); \it sshort\_bv16 $ \in \finset$\rm (\it
+% sshort\_bv16\rm ); and \it bv16\_sshort $ \in \finset$\rm (\it bv16\_sshort\rm).
+ These proof obligations were verified using the expression evaluator of ProB.
+Finally, we have included rules related to binary arithmetic as developed
+in~\cite{Leibnizens}. These rules are used in the interactive theorem prover with
+binary arithmetic to prove important properties~\cite{James2010,Dasgupta2006}.
 
@@
 The \textit{Z80} is a CISC microcontroller developed by
-\textit{Zilog}~\cite{Z80_manual}. The Z80 is composed by different elements and a
-simplified internal organization is shown in the Figure \ref{fig:DiagramBlock}.
-This figure has some important elements of Z80 CPU: ALU, registers of 8 and 16
-bits and input/output ports.
+\textit{Zilog}~\cite{Z80_manual}. The Z80 is composed by different
+elements and a simplified internal organization is shown in the Figure
+\ref{fig:DiagramBlock}.  This figure has some important elements of
+Z80 CPU: ALU, registers of 8 and 16 bits and input/output ports.
 
@@
 
-The Z80 has 158 different instructions, including all the 78 from Intel 8080
+The Z80 has 158 different instructions, including all the 78 from the Intel 8080
 microprocessor, and all of them were formally specified in B. These instructions are classified
@@
 %Several functions were created to Z80.
-General functions that can be used in others microcontroller models are defined in
+General functions that can be used in other microcontroller models are defined in
 the modules of data types. Specific functions of the microcontroller are defined in the ALU (arithmetic logic unit) module.
@@
 
-%
-
 % Features of microcontroller
-  The internal registers contain 176 bits of read/write memory that are represented by
+  The internal registers contains 176 bits of read/write memory that are represented by
 identifiers used as parameters in the instructions . It includes two sets of six general purpose
@@
 registers are represented by variable $\textit{rgs8}$. The domain of $\textit{rgs8}$
-($\textit{id\_regs8}$) is a set formed by identifiers of registers of 8 bits. These registers are called
+($\textit{id\_reg8}$) is a set formed by identifiers of registers of 8 bits. These registers are called
 main register set (\textit{a0,b0,c0,d0,e0,f0,h0,l0}) and alternate register set
@@
 	\bf SETS\\
-	\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0 \rm , \it f0 \rm ,\it b0 \rm , \it c0 \rm ,\it d0 \rm , \it e0 \rm ,\it h0 \rm , \it l0 \rm ,\it a\_0 \rm , \it f\_0 \rm ,\it b\_0 \rm , \it c\_0 \rm ,\it d\_0 \rm , \it e\_0 \rm ,\it h\_0 \rm , \it l\_0 \rm \};\\
-	\hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{ \it BC \rm , \it DE \rm , \it HL \rm , \it SP \rm , \it AF \rm \}
+\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0\rm, \it f0\rm, \it b0\rm, \it
+c0\rm, \it d0\rm, \it e0\rm, \it h0\rm, \it l0\rm, \it a\_0\rm,  \it f\_0\rm,
+\it b\_0\rm, \it c\_0\rm, \it d\_0\rm, \it e\_0\rm, \it h\_0\rm, \it l\_0 \rm
+\};\\ \hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{  \it AF \rm, \it BC\rm, \it
+DE\rm, \it HL\rm, \it SP\rm\}
 
@@
 used as a \textbf{flag register}. This register uses only six bits to represent the execution result status of each instruction.
-According to the official manual the bits 3 and 5 are not used and the others bits have the follow meaning: $\textit{bv\_get(rgs8(f0),0)}$ - the carry bit; $\textit{bv\_get(rgs8(f0),1)}$ - the add/subtract bit; $\textit{bv\_get(rgs8(f0),2)}$ - the parity or overflow bit; $\textit{bv\_get(rgs8(f0),4)}$ - the half carry bit;
-$\textit{bv\_get(rgs8(f0),6)}$ - the zero bit; and $\textit{bv\_get(rgs8(f0),7)}$ - the sign bit.
-%\begin{description}
-%  \item[$\textit{bv\_get(rgs8(f0),0)}$] - The Carry bit
-%  \item[$\textit{bv\_get(rgs8(f0),1)}$] - The Add/Subtract bit
-%  \item[$\textit{bv\_get(rgs8(f0),2)}$] - The Parity or Overflow bit
-%  \item[$\textit{bv\_get(rgs8(f0),4)}$] - The Half Carry bit
-%  \item[$\textit{bv\_get(rgs8(f0),6)}$] - The Zero bit
-%  \item[$\textit{bv\_get(rgs8(f0),7)}$] - The Sign bit 
-%\end{description}
-These bits also can be used to specify security properties for microcontroller programs.
+According to the official manual the bits 3 and 5 are not used and the others bits have the follow meaning: %$\textit{bv\_get(rgs8(f0),0)}$ - the carry bit; $\textit{bv\_get(rgs8(f0),1)}$ - the add/subtract bit; $\textit{bv\_get(rgs8(f0),2)}$ - the parity or overflow bit; $\textit{bv\_get(rgs8(f0),4)}$ - the half carry bit; $\textit{bv\_get(rgs8(f0),6)}$ - the zero bit; and $\textit{bv\_get(rgs8(f0),7)}$ - the sign bit.
+\begin{description}
+  \item[$\textit{bv\_get(rgs8(f0),0)}$] - The Carry bit.
+  \item[$\textit{bv\_get(rgs8(f0),1)}$] - The Add/Subtract bit.
+  \item[$\textit{bv\_get(rgs8(f0),2)}$] - The Parity or Overflow bit.
+  \item[$\textit{bv\_get(rgs8(f0),4)}$] - The Half Carry bit.
+  \item[$\textit{bv\_get(rgs8(f0),6)}$] - The Zero bit.
+  \item[$\textit{bv\_get(rgs8(f0),7)}$] - The Sign bit.
+\end{description}
+These bits can also be used to specify security properties for microcontroller programs.
 
@@
 %and the displacement to return the sum, the result is the dislocated address memory, 
- An additional byte is included in indexed instructions to specify an offset from this base. This displacement is specified as a two's complement signed integer. This function is defined as:
+ An additional byte is included in indexed instructions to specify an offset from this base. This displacement is specified as a two's complement signed integer. This function is defined as:\\
 
@@
 \hspace*{0.20in}$\mid$ \it ushort\_bv16 \rm ( \rm (\it bv16\_ushort \rm ( \it ix\_iy \rm ) \rm + \it offset \rm ) 
-\textbf{\textit{mod}} $2^{16}$ ) \rm )
-
-Another derived function is  $\textit{bv\_(ireg\_plus\_d)}$, this returns the value in the memory  address
-returned by $\textit{bv\_ireg\_plus\_d}$ function and its definition is similar.
+\textbf{\textit{mod}} $2^{16}$ ) \rm )\\
+
+Another derived function is  $\textit{bv\_(ireg\_plus\_d)}$, which returns the value in the memory  address
+returned by the $\textit{bv\_ireg\_plus\_d}$ function and its definition is similar.
 
@@
 It is typed and defined as follow:\it update\_flag\_reg \rm $\in$ \rm (\it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it
-BIT $\times$  \it BIT  $\times$  \it BIT\rm) $\fun$  \rm (\rm \{\it f0\rm \}  $\times$  \it BYTE\rm ).
-The definition is: \it update\_flag\_reg \rm =  $\lambda$  \rm (\it s7 \rm, \it z6 \rm,\it h4 \rm,\it pv2 \rm ,\it n1
+BIT $\times$  \it BIT  $\times$  \it BIT\rm) $\fun$  \rm (\rm \{\it f0\rm \}  $\times$  \it BYTE\rm ).  This function is defined as:\\
+
+ \it update\_flag\_reg \rm =  $\lambda$  \rm (\it s7 \rm, \it z6 \rm,\it h4 \rm,\it pv2 \rm ,\it n1
 \rm ,\it c0 \rm) . 
@@
 
-In general, the instructions can access all memory address, but it is dangerous. The user may mistakenly
+In general, the instructions can access all memory addresses, but it is dangerous. The user may mistakenly
 access and change data in memory. For added  security, it is important that the program instructions has
@@
 \textbf{Assuring the absence of overlapping of address regions:}
- \emph{To assure that address regions are well defined, then the designer must to verify the expression:}
+ \emph{To assure that address regions are well defined, then the designer must verify the expression:}
 %\begin{sloppypar}
 %\hspace*{0.10in}
-\it PROGRAM\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ STACK\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ PROGRAM\_R\_ADR $\cap$  STACK\_R\_ADR  $=$ \{\} 
+\it PROGRAM\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ STACK\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ PROGRAM\_R\_ADR $\cap$  STACK\_R\_ADR  $=$ \{\} \rm
 %\end{sloppypar}
 
-\textbf{Preserving the consistency of the memory:} \emph{In general, the access to
-some address regions is dangerous. Then, each instruction has a specific pre-condition that verify
-if the new address memory, that will be updated, is member of its region. For example, the $\textit{PUSH}$
-program instruction allows write only in the region of stack ($\textit{STACK\_R\_ADR}$).}
+\textbf{Preserving the memory consistency:} \emph{In general, the access to
+some address regions is dangerous. Then, each instruction has a specific pre-condition that verifies
+if the new address memory, that will be updated, is a member of its region. For example, the $\textit{PUSH}$
+program instruction permits write only in the region of stack ($\textit{STACK\_R\_ADR}$\rm).}
 
@@
 $\textit{half8UCHAR}$ is used to get the half part of $\textit{UCHAR}$ value.
-It is important to know the half carry and it is used in the function $\textit{add8UCHAR}$.
+It is important to know the half carry and it is used in the function $\textit{add8UCHAR}$. \\
 
@@
 
-\hspace*{0.0in}\it half8UCHAR \rm =  $\lambda$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it UCHAR  $\mid$  \it ww  $\mod$  \it $2^{4}$\rm )
+\hspace*{0.0in}\it half8UCHAR \rm =  $\lambda$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it UCHAR  $\mid$  \it ww  $\mod$  \it $2^{4}$\rm )\\
 
@@
 sum, the sign bit, the carry bit, the half carry bit and the zero bit. It is typed as follows: \it add8UCHAR \rm :
-\rm (\it BIT $\times$ \it UCHAR $\times$ \it UCHAR\rm ) $\fun$ \rm (\it UCHAR $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm ) and its definition is:
+\rm (\it BIT $\times$ \it UCHAR $\times$ \it UCHAR\rm ) $\fun$ \rm (\it UCHAR $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm ) and its definition is:\\
 
@@
 \hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm
-)\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )
-
-\hspace*{0.20in}
+)\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )\\
 
@@
 
-Each instruction is represented by a B operation in the module Z80. The main
+Each instruction is represented by a B operation in Z80 module. The main
 module (\textit{Z80}) has three categories of operations: a category represents the 
@@
 (shown in ~\ref{sec:modelingIO}) and the last category represents the external
-actions (shown in~\ref{sec:externalactions}). A simple example of instruction category is
-$\textit{LD\_(nn)\_A}$\footnoteremember{myfootnote}{The tools B does not allow
+actions (shown in~\ref{sec:externalactions}). A simple example of instruction is
+$\textit{LD\_(nn)\_A}$\footnoteremember{myfootnote}{The B tools does not allow
 to use parentheses in identifiers, and the characters \{``('',``)''\} are
@@
 
-\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} $\land$ \hspace*{0.10in}\it nn\hspace*{0.10in} $\in$  \it DATA\_R\_ADR
-
-\hspace*{0.20in}\bf THEN
+\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} $\land$ \hspace*{0.10in}\it nn\hspace*{0.10in} $\in$  \it DATA\_R\_ADR \hspace*{0.10in}\bf THEN
 
@@
 by the following B operation.
-It receives an register identifier ``$\textit{rr}$'' and, it stores the value from ``rr'' in the $\textit{C}$ port address.
+It receives a register identifier ``$\textit{rr}$'' and, it stores the value from ``rr'' in the $\textit{C}$ port address.
 Besides, it increments the program counter and updates the flag registers.
@@
 
-\hspace*{0.20in}\bf ANY
-
-\hspace*{0.40in}\it negative \rm , \it zero \rm , \it half\_carry \rm , \it pv \rm , \it add\_sub \rm , \it carry
+\hspace*{0.20in}\bf ANY \hspace*{0.10in}\it negative \rm , \it zero \rm , \it half\_carry \rm , \it pv \rm , \it add\_sub \rm , \it carry
 
@@
 \hspace*{0.40in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
+
+\hspace*{0.20in}\bf END
 
@@
 The external actions change the state of the microcontroller, for example, refreshing the I/O ports
-and interruptions request. The external actions are also modeled by operations and they are named with the
+and interruption request. The external actions are also modeled by operations and they are named with the
 prefix ``$ext\_$'' and followed by the name of action. There are four external
-actions: $ext\_update\_io\_ports$, $ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The
-$ext\_update\_io\_ports$ just change the state of I/O port, see.
+actions: $ext\_update\_io\_ports$, $ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The instruction
+$ext\_update\_io\_ports$ just changes the state of I/O port, see.
 
@@
 
-The other external actions are related to interruptions. The interruptions allow
-that the devices suspend a routine from CPU and start another service routine.
+The other external actions are related to interruptions. Interruptions allow
+devices to suspend a routine from CPU and start another service routine.
 This service routine can exchange data or signals between CPU and external
@@
 program counter. Theses actions are modeled by B operations and its main
-effects are shown below\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$ holds the value of stack pointer minus 2,
+effects are presented here\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$ holds the value of stack pointer minus 2,
 %=$\textit{dec\_BV16(dec\_BV16(sp))}$
@@
  by the programmer. Then, when a device makes a request, $sp$ is pushed,
- $pc$ receives $66H$ (102 in decimal), $\textit{iff1}$ is reset , $\textit{iff2}$ stores
+ $pc$ receives $66H$ (102 in decimal), $\textit{iff1}$ is reset, $\textit{iff2}$ stores
  $\textit{iff1}$ and the refresh register is updated.
@@
 
-  \item The mode 0 is compatible with 8080 and this mode is selected when \it im \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a 
-  non-maskable interrupt happens, the CPU fetches an instruction of one byte
+  \item The mode 0 is compatible with 8080 and this mode is selected when \it im \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ).
+When a  non-maskable interrupt happens, the CPU fetches an instruction of one byte
   from an external device, usually an RST instruction, and the CPU executes it.
-  The instruction code is received from an external device by data bus and it is represented by integer parameter called $\textit{byte\_bus}$ .
+  The instruction code is received from an external device by data bus and it is represented by integer parameter called $\textit{byte\_bus}$.
 
@@
 
-The modeling of instruction set, interruptions and input/output ports become the animation interesting.
-The animation in ProB version is not so fast and practical as a simulation, for example in \cite{Simulator_z80},
-because ProB works manipulating logic and math concepts that can have infinite sets and non determinism, then ProB requires more processing, specially when it works analysing  complex expressions or very large expressions.
-However, a user of ProB with this modeling can: execute manually a sequence of Z80's instructions that represent a program and view the visited states of ports, memory and registers, it is important to debug the effect action on the state space computed; create a tracing of the states; analyse constants, functions and expressions; and search possible violations.
-
-%\subsection{Considerations about the Z80 model}
+The modeling of instruction set, interruptions and input/output ports become the
+animation interesting. The animation in ProB is not so fast and practical as a
+simulation, for example in \cite{Simulator_z80}, because ProB works manipulating
+logic and math concepts that can have infinite sets and non determinism, then
+ProB requires more processing, specially when it works analysing  complex
+expressions or very large expressions. However, a user of ProB with this modeling
+can: execute manually a sequence of Z80's instructions that represent a program
+and view the visited states of ports, memory and registers. It is important to
+debug the effect action on the state space computed; create a tracing of the
+states; analyse constants, functions and expressions; and search possible
+violations.
+
+\subsection{Verification process}% ESTOU TRABALHANDO AQUI
+
+% parágrafo inicial
+This project has many proof obligations and its construction of then model was a difficult task, as several
+% iterations were needed to provide the good library definitions
+% as well as to optimize the definition of the functionality of
+% the microprocessor instructions by grouping common data
+% manipulation into auxiliary functions.
+
+% [O que as provas garantem? citar que dificeis, que é necessário buscar técnicas eficientes
+% para realizar as provas ]
+
+The proof obligations allow to verify the data types, important system
+properties and if the expressions are well-defined (WD)\footnote{An
+  expression is called ``well-defined'' (or unambiguous) if its
+  definition assigns it a unique interpretation or value.}. The
+properties provide additional guarantees, because they can set many
+safety rules. However, the model can be very difficult to prove.
+
+% Conclusão: toque final
+Several iterations were needed to provide the good library definitions as well as
+to fine-tune the model of the microcontroller instructions by factoring common
+functionalities into auxiliary definitions.
+
+%OK -> como as regras foram utilizadas ?  Respondido na seção das regras -
+
+%-> Quanto tempo foi gasto nessa verificação ? a minha dissertação diz que aproximadamente dois meses. Utilizando as regras este tempo caiu p/ 5 dias (linkar com este parágrafo)
+
+%-> Aproximadamente quanto tempo de processamento elas reduziram ?
+% Porque o modelo não foi verificado completamente ? Isto é viavel ?
+ %There are many benefits when a formal model is
+%developed, but to verify the model is not an easy task, the first version of Z80 model have already verified %completely,but the new Z80 model with support to animation in ProB has not been verified completely,
+%because the changes in Z80 model are recent and the verification time is long.
+
+TODO:ENGLISH REVIEW
+
+%Decomposição do módulo Z80
+The Z80 module was decomposed in 8 modules of instructions, it was very helpful to decrease the time verification, because the AtelierB can start the simultaneously 8 instances of provers reducing the time verification by up to 8 times, when it uses a processor with 8 cores.
+
+ %-> O que significa cada coluna e linha da tabela ?  	
+The follow table~\ref{tab:completestatistcs} contains the results of verification process, it shows only the number of no obvious proof obligations and the titles of columns use acronyms: {\small \textbf{PO}} - number of proof obligations; {\small \textbf{UPO}} - number of unproved proof obligations; {\small \textbf{WD PO}} - number of proof obligations  related to well definition lemmas; {\small \textbf{UWD PO}} - number of unproved proof obligations related to well definition lemmas.
+
+\begin{table} [h]
+\caption{Verification statistics }
+\label{tab:completestatistcs}
+\begin{center} 
+\begin{tabular}{|p{3.1cm}|c|c||c|c||c|c|c|}
+\hline 
+{\small \textbf{Component}}&{\small \textbf{PO}}&{\small \textbf{UPO}}&{\small \textbf{WD PO}}&{\small \textbf{UWD PO }}&{\small \textbf{Total PO}}&{\small \textbf{Total UPO}}& {\small \textbf{Total Percentage}}  \\\hline
+BIT&49&0&69&0&118&0&100\%\\\hline
+BV16&6&0&69&0&75&0&100\%\\\hline
+BYTE&18&0&136&0&154&0&100\%\\\hline
+POWER&3&0&4&0&7&0&100\%\\\hline
+POWER2&18&0 &0&0&18&0&100\%\\\hline
+SCHAR&4&0&26&0&30&0&100\%\\\hline
+SSHORT&4&0&58&0&62&0&100\%\\\hline
+TYPES&71&0&0&0&71&0&100\%\\\hline
+UCHAR&4&0&26&0&30&0&100\%\\\hline
+USHORT&4&0&58&0&62&0&100\%\\\hline
+ALU&51&22&117&47&168&69&59\%\\\hline
+MEMORY&13&0&0&0&13&0&100\%\\\hline
+Z80 -  {\scriptsize Initialization, Properties, Assertions} &66&9&150&11&216&20&91\%\\\hline
+Z80 - {\scriptsize Input and Output}&46&11&281&16&327&27&92\%\\\hline
+Z80 - {\scriptsize Logic Arithmetic 1}&88&33&343&84&431&117&73\%\\\hline
+Z80 - {\scriptsize Logic Arithmetic 2}&40&20&178&27&218&47&79\%\\\hline
+Z80 - {\scriptsize Bit Control}&111&34&763&157&874&191&78\%\\\hline
+Z80 - {\scriptsize External Actions}&31&0&60&38&91&38&59\%\\\hline
+Z80 - {\scriptsize General 1}&81&19&420&20&501&39&92\%\\\hline
+Z80 - {\scriptsize General 2}&26&10&139&14&165&24&86\%\\\hline
+
+\textbf{Total}&\textbf{734}&\textbf{158}&\textbf{2897}&\textbf{364}&\textbf{3631}&\textbf{569}&\textbf{84}\%\\\hline
+
+\end{tabular}
+\end{center} 
+\end{table}
+
+% Resumo de como aconteceu o processo de verificação, (melhor está após as estátisticas) ( )Atualizar dados
+However, few proof commands need to be used to prove most proof obligations. As there are many similar assembly instructions,
+some human-directed proofs, when replayed, could discharge other proof
+obligations. A good example is a set of 17 proof commands that quickly aided the
+verification of 99\% (2295) of WD proofs. We also set up a proving environment
+consisting of networked computers to take advantage of the distribution
+facilities now provided in the B development environment. Finally, all of the
+2926 proof obligations were proved using the tool support of the development
+environment.
+
+\subsection{Considerations about the Z80 model}
 
@@
 generated proof obligations guarantees: the correct use of data types, the developed
-security properties  and the well-defined of all the expressions. %\footnote{An expression is called ``well-defined'' (or unambiguous) if its definition assigns it a unique interpretation or value.}.
-Furthermore, the designer has a big flexibility to create new and specific security properties, this is very useful
+security properties  and the well-defined of all the expressions. 
+Furthermore, the designer has a big flexibility to create new and specific security properties, this is very useful %TODO: COLOCAR APÓS TODOS OS BENEFICIOS PROPORCIONADOS COM A SIMULAÇÃO e COMPATIBILIDADE COM O PROB.
+(OBTER DO PARAGRAFO ANTERIOR DA SEÇAO VERIFICATION PROCESS).
 to adjust the verification in accordance with the requirements. Moreover, this
@@
 assembly programmers. Z80 model is also useful to develop verified software 
-up to the assembly level. There are many benefits when a formal model is
-developed, but to verify the model is not an easy task, the first version of Z80 model have already verified completely,but the new Z80 model with support to animation in ProB has not been verified completely,
-because the changes in Z80 model are recent and the verification time is long.
+up to the assembly level.
 
@@
 since the first abstract B model up to B assembly model. This case study is an
-object of the pilot project developed to analyse a petroleum production test
-system in each oil field. Its modeling was developed in according to \cite{LAUT_SERGIO}
+object from the pilot project developed to analyse a petroleum production test
+system in each oil field. Its modeling was developed according to \cite{LAUT_SERGIO}
 and its B assembly model can be executed manually in ProB. The modeling of
@@
 
-\section{Related works}
+\section{Related work}
 \label{sec:relatedworks}
 
-There are in the literature several similar approaches to model hardware and instruction set using B.
-The paper \cite{springerlink:Yuan2011} reports a method to specify, design and construct sound and complete instruction set model by stepwise refinement and formal proof using the formal method Event-B.
-It also discuss about desirable properties of instruction set model.
-However, our work \cite{Valerio_SBMF09} and \cite{Subotic2010} use the B method, that seems more appropriated to software development, because it has implementable language defined,
-called B0, and tools to convert the models to a programming language.
-In \cite{Marc20113} is present an related experience on using B in the design of secure micro-controllers.
-It tries to show the feasibility of such a technique for high confidence trustful devices.
-The work \cite{Subotic2010} describes a similar approach with MIPS CPU architecture and it shows how the CPU is formally specified and implemented using the B method.
+There exists several approaches to model hardware and instruction set
+using B a.  The paper \cite{springerlink:Yuan2011} reports a method to
+specify, design and construct sound and complete instruction set model
+by stepwise refinement and formal proof using Event-B. It also
+discusses desirable properties of the instruction set model. However,
+our work \cite{Valerio_SBMF09} and \cite{Subotic2010} use the B
+method, that seems quite appropriated to software development, because
+it has an implementable language defined, called B0, a model-driven
+approach \cite{Dantas_SBMF08} to develop software from the functional
+specification level down to assembly, and tools to convert the models
+to a programming language.
+
+A related experience on using B in the design of secure
+micro-controllers is present in \cite{Marc20113}. It tries to show the
+feasibility of such a technique for high confidence trustful
+devices. The work \cite{Subotic2010} describes a similar approach with
+MIPS CPU architecture and it shows how the CPU is formally specified
+and implemented using the B method.
 
@@
 % Um pequeno resumo do trabalho
-This work has shown an approach to the formal modeling of the instruction set of
-microcontroller using the B method. During the construction of this model, some
-problems were found in the official reference for Z80
-microcontroller~\cite{Z80_manual}. The designer fixed the found problems in
-documentation and also developed a case study: a verified embedded software. This
-case study was interesting to analyse the technique of formal verification up to
-assembly level.
-
-% TODO: Adaptar para apenas fundamentar melhor o trabalho já desenvolvido. Citar o artigo SBMF e a dissertação. 
+This work has shown an approach to the formal modeling of the
+instruction set of microcontroller using the B method.  During the
+construction of this model, some problems were found in the official
+reference for Z80 microcontroller~\cite{Z80_manual}. We have consulted
+different sources \cite{Simulator_z80,UndocumentedZ80,Z80_manual} to
+avoid such problems in modeling and have developed a case study: a
+verified embedded software.  This case study was interesting to as a
+feasibility analysis of the approach to derive formally assembly code
+from a specification of functional requirements in the B method.
+
+%TODO: COLOCAR AQUI OS PRINCIPAIS BENEFÍCIOS PROPORCIONADOS PELO TRABALHO
+
+% fundamenta melhor o trabalho já desenvolvido. Citar o artigo SBMF e a dissertação. 
 % Os meus trabalhos já desenvolvidos/ resultados atuais.
-The next works quoted are directly related to the objective this paper.
-The first work \cite{DantasSemish2008} shows the developed metodology to verify
-software up to assembly level using B. A second work\cite{Dantas_SBMF08}
-presents more details about the verification approach and a small example
-software verified up to assembly level in three different platforms. 
-The last work \cite{Valerio_SBMF09} shows a general view of the first version of Z80
-model and techniques used in verification process.
-
+The following works are directly related to the objective this paper.
+The approach to develop verified software down to the assembly level
+using B was described in~\cite{DantasSemish2008}. A first case study
+for this approach was reported in~\cite{Dantas_SBMF08}, presenting
+more details as well as a small software that developed up to assembly
+level in three different platforms. A general view of a previous used
+in verification process were presented in~\cite{Valerio_SBMF09}. The
+model presented in the current paper added support for animation in
+ProB and the representation of external actions (interruptions).
 %Link com o parágrafo anterior 
-These works created a important step to software verification up to assembly
-language, they show the actuals difficulties and suggest improvements in the
-B tools.
-
-% TODO
+This set of papers present an important step to software verification
+up to assembly language, showing the current difficulties and
+suggesting improvements in tools supporting the B method.
+
 % [Trabalhos futuros ] LLVM:Existe também a pretensão de especificar as
-
-Future works comprise the development of software with the B method from
-functional specification to assembly level, using the Z80 model presented in this
-work. The mechanic compilation from B algorithmic constructs to assembly platform
-is also envisioned. Finally, another important activity is develop a formal model of a platform used in LLVM Compiler~\cite{DBLP:conf/cgo/LattnerA04}.
-
-\subsubsection{Acknowledgements:} This experiment could not have succeeded without the interesting work of Michael Leuschel who developed several improvements in the ProB and provided important aid to enable the animation of the Z80 modeling.
+In the future, we plan to improve the support for the development of
+software with the B method from functional specification to assembly
+level, using the Z80 model presented in this work. For instance, the
+mechanic compilation from B algorithmic constructs to assembly
+platform is also envisioned. Finally, another important activity is to
+develop a formal model of a platform used in LLVM
+Compiler~\cite{DBLP:conf/cgo/LattnerA04}, which would enable us to
+integrate different compilation techniques into our B-based framework.
+
+\subsubsection{Acknowledgements:} The animation of the model would not
+have been possible without the help of Michael Leuschel who kindly
+provided feedback and developed several improvements to ProB to meet our
+needs.
 

### 2012-05-15 papers/ArticleZ80/source/paper.tex
Adicionei as estatisticas de verificacao e fiz algumas correcoes

@@
 
-%[NEW PARAGRAPH-TO-REVIEW]
-%Similarly, several other general functions and lemmas were created for all other data types.
-%Some times, we need create arithmetic and logic functions that are more specific for a
-%determined computer architecture. These functions are developed in the module ALU 
-%(arithmetic  logic unit). The following shows some examples of specific functions used in
-%microcontroller model of Z80. For example, the function $\textit{instruction_next}$
-%receives a $\textit{BYTE}$ and returns a $\textit{BYTE}$ that contains a sequence of bit:
-%a signal bit,
-
-Nearly 50\% of all these proof obligations shown in the
-Table~\ref{tab:types} are verified automatically or with few proof
-commands\footnote{The proof commands are steps that help the prover to find the proof.} .
-However many others proof obligations are difficult to verify, as they involve functions
-that contain arithmetic expressions that are difficult to manipulate with the interactive
+%TODO:[NEW PARAGRAPH-TO-REVIEW]
+Similarly, several other general functions and lemmas were created for all other data types.
+Some times, we need  create arithmetic and logic functions that are more specific for a
+determined microprocessor, like Z80 that is presented in the section \ref{sec:z80}.
+
+%TODO:[NEW PARAGRAPH-TO-REVIEW]
+The definitions, functions and properties from hardware library and types library are difficult to verify,
+on the other hand  these libraries can be reused in several others  microprocessors without to verify again.
+Nearly 50\% of all these proof obligations shown in the Table~\ref{tab:types} are verified automatically 
+or with few user pass\footnote{The user pass are steps that aid the prover to find the proof.} .
+But many others proof obligations are difficult to verify, because they involve functions
+that contain arithmetic expressions that are complex to manipulate with the interactive
 prover of AtelierB.  However, all theses 531 proof obligations were verified completely.
@@
 
-This section shows the elements that make up the different types of instructions.
-The main elements are specified in the microcontroller module Z80 and parts of it
-are presented below.  Basically, the state of microcontroller is formed by data 
-from program counter, ports, registers, etc. The transitions between the states 
-are actioned by execution of instruction or an external action. Groups of registers
-are represented by the variables of the specification states that appear in clause
-\textit{VARIABLES}. The declaration of valid states of variables is
-in the \textit{INVARIANT} clause and the initial state is defined in the
-\textit{INITIALISATION} clause. The assembly instructions are defined by the
-clause \textit{OPERATIONS} and, in general, each instruction has three actions: update 
-the program counter, update the flag register and its main effect.
-%Several functions were created to Z80.
-General functions that can be used in other microcontroller models are defined in
-the modules of data types. Specific functions of the microcontroller are defined in the ALU (arithmetic logic unit) module.
-
-The main module includes an instance of the memory module and accesses the definitions from basic data
-types modules and the \textit{ALU} module.
+This section shows the elements that make up the different types of
+instructions. The main elements are specified in the microcontroller module Z80
+and parts of it are presented below.  Basically, the state of microcontroller is
+formed by data from program counter, ports, registers, etc. The transitions
+between the states are actioned by execution of instruction or an external
+action. Groups of registers are represented by the variables of the
+specification states that appear in clause \textit{VARIABLES}. The declaration
+of valid states of variables is in the \textit{INVARIANT} clause and the initial
+state is defined in the \textit{INITIALISATION} clause. The assembly
+instructions are defined by the clause \textit{OPERATIONS} and, in general, each
+instruction has three actions: update the program counter, update the flag
+register and its main effect.
+% Several functions were created to Z80.
+General functions that can be used in other microcontroller models are defined
+in the modules of data types. Specific functions of the microcontroller are
+defined in the ALU (arithmetic logic unit) module.
+
+The main module includes an instance of the memory module and accesses the
+definitions from basic data types modules and the \textit{ALU} module.
 
@@
 
-The Z80 CPU includes alternative set of accumulator, flag and general registers. The CPU contains a stack
-pointer ($\textit{sp}$), program counter ($\textit{pc}$), two index registers ($\textit{ix}$ and $\textit{iy}$), an
-interrupt register ($\textit{i\_}$), a refresh register ($\textit{r\_}$), two bits ($\textit{iff1}$,
-$\textit{iff2}$) used to control the interruptions, a pair of bits to define the interruption mode ($\textit{im}$)
-and the input and output ports ($\textit{i\_o\_ports}$). These definitions are represented by
+The Z80 CPU includes alternative set of accumulator, flag and general registers.
+The CPU contains a stack pointer ($\textit{sp}$), program counter
+($\textit{pc}$), two index registers ($\textit{ix}$ and $\textit{iy}$), an
+interrupt register ($\textit{i\_}$), a refresh register ($\textit{r\_}$), two
+bits ($\textit{iff1}$, $\textit{iff2}$) used to control the interruptions, a
+pair of bits to define the interruption mode ($\textit{im}$) and the input and
+output ports ($\textit{i\_o\_ports}$). These definitions are represented by
 \textit{INVARIANT}.
@@
 % Features of microcontroller
-  The internal registers contains 176 bits of read/write memory that are represented by
-identifiers used as parameters in the instructions . It includes two sets of six general purpose
-registers which may be used individually as 8-bit registers or as 16-bit register pairs.  The working
-registers are represented by variable $\textit{rgs8}$. The domain of $\textit{rgs8}$
-($\textit{id\_reg8}$) is a set formed by identifiers of registers of 8 bits. These registers are called
-main register set (\textit{a0,b0,c0,d0,e0,f0,h0,l0}) and alternate register set
-(\textit{a\_0,b\_0,c\_0,d\_0,e\_0,f\_0,h\_0,l\_0}). These registers can be accessed in pairs, forming
-16-bits, resulting in another set of identifiers of 16-bits registers, named $\textit{id\_reg16}$. These sets are represented below.
+The internal registers contains 176 bits of read/write memory that are
+represented by identifiers used as parameters in the instructions . It includes
+two sets of six general purpose registers which may be used individually as
+8-bit registers or as 16-bit register pairs.  The working registers are
+represented by variable $\textit{rgs8}$. The domain of $\textit{rgs8}$
+($\textit{id\_reg8}$) is a set formed by identifiers of registers of 8 bits.
+These registers are called main register set (\textit{a0,b0,c0,d0,e0,f0,h0,l0})
+and alternate register set (\textit{a\_0,b\_0,c\_0,d\_0,e\_0,f\_0,h\_0,l\_0}).
+These registers can be accessed in pairs, forming 16-bits, resulting in another
+set of identifiers of 16-bits registers, named $\textit{id\_reg16}$. These sets
+are represented below.
 
@@
 
-The external actions change the state of the microcontroller, for example, refreshing the I/O ports
-and interruption request. The external actions are also modeled by operations and they are named with the
-prefix ``$ext\_$'' and followed by the name of action. There are four external
-actions: $ext\_update\_io\_ports$, $ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The instruction
-$ext\_update\_io\_ports$ just changes the state of I/O port, see.
+The external actions change the state of the microcontroller, for example,
+refreshing the I/O ports and interruption request. The external actions are also
+modeled by operations and they are named with the prefix ``$ext\_$'' and
+followed by the name of action. There are four external actions:
+$ext\_update\_io\_ports$, $ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The
+instruction $ext\_update\_io\_ports$ just changes the state of I/O port, see.
 
@@
 
-The interruptions and the \textit{reset} action can change the state of
-program counter. Theses actions are modeled by B operations and its main
-effects are presented here\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$ holds the value of stack pointer minus 2,
-%=$\textit{dec\_BV16(dec\_BV16(sp))}$
+The interruptions and the \textit{reset} action can change the state of program
+counter. Theses actions are modeled by B operations and its main effects are
+presented here\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$
+holds the value of stack pointer minus 2,
+% =$\textit{dec\_BV16(dec\_BV16(sp))}$
  $\textit{sp\_minus\_one}$ is the value of stack pointer minus 1,
-%=$\textit{dec\_BV16(sp)}$
-$\textit{pc\_high}$ holds the most significant 8 bits and
-$\textit{pc\_low}$ holds the least significant 8 bits.}.
+% =$\textit{dec\_BV16(sp)}$
+$\textit{pc\_high}$ holds the most significant 8 bits and $\textit{pc\_low}$
+holds the least significant 8 bits.}.
 
@@
 
-  \textbf{INT} - Maskable Interrupt is usually reserved for important functions that can be enabled and
-  disabled by the programmer. When a maskable interrupt action happens, both $\textit{iff1}$ and $\textit{iff2}$ are
-  cleared, disabling the interrupts, $sp$ is pushed, the refresh register is updated and the other effects
-  depend on the interrupt mode register (\it im\rm).
+  \textbf{INT} - Maskable Interrupt is usually reserved for important functions
+  that can be enabled and disabled by the programmer. When a maskable interrupt
+  action happens, both $\textit{iff1}$ and $\textit{iff2}$ are cleared,
+  disabling the interrupts, $sp$ is pushed, the refresh register is updated and
+  the other effects depend on the interrupt mode register (\it im\rm).
 
@@
 
-  \item The mode 0 is compatible with 8080 and this mode is selected when \it im \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ).
-When a  non-maskable interrupt happens, the CPU fetches an instruction of one byte
-  from an external device, usually an RST instruction, and the CPU executes it.
-  The instruction code is received from an external device by data bus and it is represented by integer parameter called $\textit{byte\_bus}$.
-
-  \item The mode 1 is the simplest and this mode is selected when \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 1
-  \rm ). Simply, when a non-maskable interruption happens, the program counter
-  receives $38H$ (56 in decimal).
-
-  \item The mode 2 is the most flexible and this mode is selected when \it im \rm = \rm ( \rm 1  $\mapsto$
-  \rm 1 \rm ). When a non-maskable interruption happens, an indirect call can be
-  made to any address memory. The program counter receives a bit vector of size
-  16 composed with two parts: the most significant part of the $\textit{i\_}$
-  register and the least significant part of the $\textit{byte\_bus}$ with the
-  last bit cleared.
+  \item The mode 0 is compatible with 8080 and this mode is selected when \it im
+  \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a  non-maskable interrupt
+  happens, the CPU fetches an instruction of one byte from an external device,
+  usually an RST instruction, and the CPU executes it.
+  The instruction code is received from an external device by data bus and it is
+  represented by integer parameter called $\textit{byte\_bus}$.
+
+  \item The mode 1 is the simplest and this mode is selected when \it im \rm =
+  \rm ( \rm 0  $\mapsto$  \rm 1 \rm ). Simply, when a non-maskable interruption
+  happens, the program counter receives $38H$ (56 in decimal).
+
+  \item The mode 2 is the most flexible and this mode is selected when \it im
+  \rm = \rm ( \rm 1  $\mapsto$ \rm 1 \rm ). When a non-maskable interruption
+  happens, an indirect call can be made to any address memory. The program
+  counter receives a bit vector of size 16 composed with two parts: the most
+  significant part of the $\textit{i\_}$ register and the least significant part
+  of the $\textit{byte\_bus}$ with the last bit cleared.
 
@@
 
-The essential part of maskable interrupt is shown below, where $\textit{byte\_bus}$ is a parameter of the $\textit{INT}$ operation:
+The essential part of maskable interrupt is shown below, where
+$\textit{byte\_bus}$ is a parameter of the $\textit{INT}$ operation:
 
@@
 ProB requires more processing, specially when it works analysing  complex
-expressions or very large expressions. However, a user of ProB with this modeling
-can: execute manually a sequence of Z80's instructions that represent a program
-and view the visited states of ports, memory and registers. It is important to
-debug the effect action on the state space computed; create a tracing of the
-states; analyse constants, functions and expressions; and search possible
-violations.
+expressions or very large expressions. However, a user of ProB with this
+modeling can: execute manually a sequence of Z80's instructions that represent a
+program and view the visited states of ports, memory and registers. It is
+important to debug the effect action on the state space computed; create a
+tracing of the states; analyse constants, functions and expressions; and search
+possible violations.
 
 \subsection{Verification process}% ESTOU TRABALHANDO AQUI
-
-% parágrafo inicial
-This project has many proof obligations and its construction of then model was a difficult task, as several
-% iterations were needed to provide the good library definitions
-% as well as to optimize the definition of the functionality of
-% the microprocessor instructions by grouping common data
-% manipulation into auxiliary functions.
-
-% [O que as provas garantem? citar que dificeis, que é necessário buscar técnicas eficientes
-% para realizar as provas ]
-
-The proof obligations allow to verify the data types, important system
-properties and if the expressions are well-defined (WD)\footnote{An
-  expression is called ``well-defined'' (or unambiguous) if its
-  definition assigns it a unique interpretation or value.}. The
-properties provide additional guarantees, because they can set many
-safety rules. However, the model can be very difficult to prove.
-
-% Conclusão: toque final
-Several iterations were needed to provide the good library definitions as well as
-to fine-tune the model of the microcontroller instructions by factoring common
-functionalities into auxiliary definitions.
-
-%OK -> como as regras foram utilizadas ?  Respondido na seção das regras -
-
-%-> Quanto tempo foi gasto nessa verificação ? a minha dissertação diz que aproximadamente dois meses. Utilizando as regras este tempo caiu p/ 5 dias (linkar com este parágrafo)
-
-%-> Aproximadamente quanto tempo de processamento elas reduziram ?
-% Porque o modelo não foi verificado completamente ? Isto é viavel ?
- %There are many benefits when a formal model is
-%developed, but to verify the model is not an easy task, the first version of Z80 model have already verified %completely,but the new Z80 model with support to animation in ProB has not been verified completely,
-%because the changes in Z80 model are recent and the verification time is long.
-
-TODO:ENGLISH REVIEW
-
-%Decomposição do módulo Z80
-The Z80 module was decomposed in 8 modules of instructions, it was very helpful to decrease the time verification, because the AtelierB can start the simultaneously 8 instances of provers reducing the time verification by up to 8 times, when it uses a processor with 8 cores.
-
- %-> O que significa cada coluna e linha da tabela ?  	
-The follow table~\ref{tab:completestatistcs} contains the results of verification process, it shows only the number of no obvious proof obligations and the titles of columns use acronyms: {\small \textbf{PO}} - number of proof obligations; {\small \textbf{UPO}} - number of unproved proof obligations; {\small \textbf{WD PO}} - number of proof obligations  related to well definition lemmas; {\small \textbf{UWD PO}} - number of unproved proof obligations related to well definition lemmas.
+% OK -> como as regras foram utilizadas ?  Respondido na seção das regras
+% TODO:ENGLISH REVIEW parágrafo inicial: VISÃO GERAL
+This project has many proof obligations and its verification can be a hard task,
+hence this section explains about how happen the verification process of proof
+obligations and its importance. % [O que as provas garantem? citar que énecessário buscar técnicas eficientes para realizar as provas ]
+The proof obligations allow to verify the data types, important system properties and if
+the expressions are well-defined (WD)\footnote{An expression is called ``well-defined''
+(or unambiguous) if its definition assigns it a unique interpretation or value.}. The
+properties provide additional guarantees, because they can set many safety
+rules. However, the model can be very difficult to prove.
+
+ % -> O que significa cada coluna e linha da tabela ?
+The Table~\ref{tab:completestatistcs} contains some statistics of verification
+process, it shows only the number of no obvious proof
+obligations\footnote{Obvious proof obligations are simple formulas automatically eliminated by
+the tool.} and the titles of columns use acronyms: {\small \textbf{PO}} - number
+of proof obligations; {\small \textbf{UPO}} - number of unproved proof
+obligations; {\small \textbf{WD PO}} - number of proof obligations  related to
+well definition lemmas; {\small \textbf{UWD PO}} - number of unproved proof
+obligations related to well definition lemmas.
 
@@
 
-% Resumo de como aconteceu o processo de verificação, (melhor está após as estátisticas) ( )Atualizar dados
-However, few proof commands need to be used to prove most proof obligations. As there are many similar assembly instructions,
-some human-directed proofs, when replayed, could discharge other proof
-obligations. A good example is a set of 17 proof commands that quickly aided the
-verification of 99\% (2295) of WD proofs. We also set up a proving environment
-consisting of networked computers to take advantage of the distribution
-facilities now provided in the B development environment. Finally, all of the
-2926 proof obligations were proved using the tool support of the development
+%->Explicaçao de mais detalhes da tabela.
+
+% Porque o modelo não foi verificado completamente ?
+There are many benefits when a formal model is developed, but to verify the
+model is not an easy task, the first version of Z80 model have already verified
+completely, however this new Z80 model with support to animation in ProB has not
+been verified completely, because the changes in Z80 model are recent and the
+verification time is long.
+% ->Quanto tempo foi gasto nessa verificação ?  Aproximadamente quanto tempo de
+% processamento elas reduziram ?
+Nevertheless this project has 3631 proof obligations and the first version of
+Z80 model spent approximately two months of work to finish the verification
+process. The same  developed verification strategies were adjusted and reused in
+new Z80 model, so approximately  7 days of work was enough to verify 3062  proof
+obligations (84\%) and remaining only 569 proof obligations (16\%) not yet
+verified.
+
+Two development strategies were important to improve the verification of 3062
+proof obligations: the intensive use of parallelism and user pass. %Decomposição
+do módulo Z80 The Z80 module was decomposed in 8 modules of instructions, it was
+very helpful to decrease the time for verification, because AtelierB can start 
+simultaneously \textit{N} instances of provers, one per module, reducing the
+time verification by up to \textit{N}  times when a processor with \textit{N} 
+cores is used  or the process is submitted to \textit{N} networked computers. So
+we also set up a proving environment consisting of networked computers to take
+advantage of the distribution facilities provided in the B development
 environment.
+% Resumo de como aconteceu o processo de verificação,
+A second interesting strategy  was to reuse the user pass. Usually, an used
+user pass to aid the verification of one proof obligation can solve several
+similar proof obligations. Thus few generic user pass are needed to prove many
+proof obligations. As there are many similar assembly instructions, some
+human-directed proofs, when replayed, could discharge other proof obligations. A
+good example is a set of 17 user pass that quickly aided the verification of
+approximately 2533 of WD proofs.
 
@@
 and analyse several important properties.
+
+%\section{Case study}
+%\include{Petroleumproductiontestsystem}
 
@@
 integrate different compilation techniques into our B-based framework.
+
+% Conclusão: toque final
+%Several iterations were needed to provide the good library definitions as well as
+%to fine-tune the model of the microcontroller instructions by factoring common
+%functionalities into auxiliary definitions.
 

### 2012-05-15 trunk/papers/ArticleZ80/source/paper.tex
novos pequenos ajustes

@@
 
-The rigorous development of an embedded software in a microcontroller requires 
-a specification of its instruction set and the software requirements specification. 
-The actual industrial practice is to report the instruction set informally using natural language
-or with added semi-formal notation. This practice may leave space for errors and inconsistencies \cite{springerlink:Yuan2011}.
-This paper presents specifically a B formal model of the Z80 instruction set. 
-This model  can be  used on documentation, simulation, test and  a model-driven approach
-to design and develop software from the functional specification level down to assembly.
-%Using this model-driven approach and the instruction set model, a small case study to
-%analyse a petroleum production test system in oil field was developed.
-%  
+  This paper presents a formal model, developed in B, of the Z80
+  instruction set. The formal model includes different libraries that
+  specify entities commonly used in microcontrolers and
+  microprocessors. Not only this model can be used for documentation
+  purposes, but also to simulate and test the assembly code. It is also
+  a fundamental component to extend the reach of the B method from the
+  specification of the functional requirements of a system to an assembly-level
+  implementation thereof.
+% Using this model-driven approach and the instruction set model, a small case
+% study to analyse a petroleum production test system in oil field was developed.
 %  \paragraph[Keywords]{ \textbf{Keywords}: Embedded Software, Simulation, B
@@
 \end{abstract}
-
-%\keywords{ Validation, Verification, Microcontroller, Hardware}
-%
-%\setcounter{secnumdepth}{1}
-%\setcounter{chapter}{1}
-%\setcounter{section}{1}
 
@@
  \end{array}
- $
-
+$$
+
+%TODO:[NEW PARAGRAPH-TO-REVIEW]
 Similarly, several other general functions and lemmas were created for all other data types.
-%[NEW PARAGRAPH-TO-REVIEW]
-%Similarly, several other general functions and lemmas were created for all other data types.
-%Some times, we need create arithmetic and logic functions that are more specific for a
-%determined computer architecture. These functions are developed in the module ALU 
-%(arithmetic  logic unit). The following shows some examples of specific functions used in
-%microcontroller model of Z80. For example, the function $\textit{instruction_next}$
-%receives a $\textit{BYTE}$ and returns a $\textit{BYTE}$ that contains a sequence of bit:
-%a signal bit,
-
-Nearly 50\% of all these proof obligations shown in the table~\ref{tab:types} 
-are verified almost automatically. Although many others proof obligations are difficult to verify,
-because they are functions of type conversion and contain arithmetic expressions difficult 
-to manipulate with the interactive prover of AtelierB.
-
-\subsection{Verification in Hardware Library and Types Library} 
+Some times, we need  create arithmetic and logic functions that are more specific for a
+determined microprocessor, like Z80 that is presented in the section \ref{sec:z80}.
+
+%TODO:[NEW PARAGRAPH-TO-REVIEW]
+The definitions, functions and properties from hardware library and types library are difficult to verify,
+on the other hand  these libraries can be reused in several others  microprocessors without to verify again.
+Nearly 50\% of all these proof obligations shown in the Table~\ref{tab:types} are verified automatically 
+or with few user pass\footnote{The user pass are steps that aid the prover to find the proof.} .
+But many others proof obligations are difficult to verify, because they involve functions
+that contain arithmetic expressions that are complex to manipulate with the interactive
+prover of AtelierB.  However, all theses 531 proof obligations were verified completely.
+
+\subsection{Verification of hardware library and types library} 
 \label{sec:VerificationHardwareLibrary}
 
-\input{VerificationLibraryHardware}
+This section presents the verification of the aforementioned modules
+using AtelierB provers.  By default, AtelierB~\cite{atelierB} does not
+generate proof obligations to check the type of functions defined in
+the $PROPERTIES$ clause of a machine. For example, the definition $inc
+\in \{0,1\} \fun \{0,1\} \land inc = \lambda(x).( x \in \{0,1\} | x +
+1)$, does not create a proof obligation to check the consistency of
+the definition. However, this definition is inconsistent since
+$inc(1)=2$ and $ 2 \not\in \{0,1\}$.
+
+To avoid such inconsistencies, we state separately the required
+properties in the \textit{ASSERTIONS\/} clause: this causes
+corresponding proof obligations to be generated and verified. However,
+we have not been able to build the corresponding proofs using the 
+rule base included in the free distribution of AtelierB. So we
+formalized new proof rules to check these properties in AtelierB and
+used them in the interactive prover of AtelierB.
+
+The hardware library has many conversion functions with
+arithmetic,  for example, Bit\_vector to Natural, Natural to
+Bit\_vector, Integer to Bit\_vector and Bit\_vector to Integer. These
+functions are organized in modules and each module contains auxiliary
+functions for a data type. The modules are represented by gray
+rectangles in Figure \ref{DiagramTypesRules}.
+
+\begin{figure}[he]
+\centering
+\includegraphics[width=3.in]{images/Diagram_Types_and_Rules.pdf}
+\caption{Conversion functions}
+\label{DiagramTypesRules}
+\end{figure}
+
+The conversion functions need to satisfy two properties: bijectivity
+and equality to the reverse of the dual conversion function. Three
+such proof obligations were verified automatically with AtelierB: (\it
+uchar\_schar\rm \ $\in$ \textit{UCHAR} $\bij$ \textit{SCHAR}), (\it
+schar\_uchar\rm \ $\in$ \textit{SCHAR} $\bij$ \textit{UCHAR}) and (\it
+byte\_bv16\rm \ $=$ \it bv16\_byte\rm $^{-1}$).
+%\begin{enumerate}
+%
+%\item (\it uchar\_schar\rm \ $\in$ \textit{UCHAR} $\bij$ \textit{SCHAR})
+%\item (\it schar\_uchar\rm \ $\in$ \textit{SCHAR} $\bij$ \textit{UCHAR})
+%\item (\it byte\_bv16\rm \ $=$ \it bv16\_byte\rm $^{-1}$) 
+%
+%\end{enumerate}
+For the remaining proof obligations, we again had to create new proof
+rules. There are three kinds of proof rules, that we illustrate by giving
+one example each:
+
+%Four conditions are needed to be a bijective function:
+%
+%\begin{description}
+%\item $func$ is a function:\\
+%$\forall$ \rm (\it xx\rm ,\it yy\rm ,\it zz\rm )\rm .\rm ( \it xx  $\in$  \bf dom\rm (\it func\rm )  $\land$  \it yy $\in$  \bf ran\rm (\it func\rm )  $\land$ \\ \it zz $\in$  \bf ran\rm (\it func\rm )  $\land$ \rm ( \it xx $\mapsto$  \it yy\rm )  $\in$  \it func  $\land$  \rm ( \it xx $\mapsto$  \it zz\rm )  $\in$  \it func  $\implies$  \it yy\rm =\it zz \rm )
+%
+%\item $func$ is an injective function:\\
+%$\forall$ \rm (\it vv\rm ,\it xx\rm )\rm .\rm ( \it vv $\in$  \bf dom\rm (\it func\rm )  $\land$  \it xx  $\in$  \bf dom\rm (\it func\rm )  $\implies$  \it func\rm (\it vv\rm )\rm =\it func\rm (\it xx\rm ) \rm )
+%
+%\item $func$ is a total function:\\
+%$\forall$ \it xx\rm .\rm ( \it xx  $\in$  \bf dom\rm (\it func\rm )  $\implies$   $\exists$ \it yy\rm .\rm ( \it yy  $\in$  \bf ran\rm (\it func\rm )  $\land$  \it func\rm (\it xx\rm )\rm =\it yy \rm )\rm )
+%
+%\item $func$ is surjective function:\\
+% $\forall$ \it yy\rm .\rm ( \it yy $\in$  \bf ran\rm (\it func\rm ) $\implies$   $\exists$ \it xx\rm .\rm ( \it xx  $\in$  \bf dom\rm (\it func\rm ) $\land$  \it func\rm (\it xx\rm )\rm =\it yy \rm )\rm )
+%
+%
+%\end{description} 
+%
+%Here, two conversion functions between bit vectors and the natural are presented.
+These functions are finite and convert from \textit{BYTE} to \textit{UCHAR} and vice versa.\textit{UCHAR} is a range between 0 and 255 and \textit{BYTE} is a bit vector with eight elements enumerated starting from 1.
+We need to prove that $byte\_uchar$  and $uchar\_byte$  are bijective functions and  $byte\_uchar = uchar\_byte^{-1}$.
+These functions are defined as follows:
+%\begin{description}
+%
+%
+%\item\it byte\_uchar \rm = $\lambda$  \rm ( \it v0 \rm ) \rm .
+\rm ( \it v0  $\in$  \it BYTE  $\mid$ \hspace*{0.10in}\it 2$^{7}$ $\times$ \it v0\rm (\rm 8\rm )\rm +\it 2$^{6}$  $\times$ \it v0\rm (\rm 7\rm )\rm + \it 2$^{5}$ $\times$ \it v0\rm (\rm 6\rm )\rm +\it 2$^{4}$ $\times$ \it v0\rm (\rm 5\rm ) \rm + \it 2$^{3}$  $\times$  \it v0\rm (\rm 4 \rm ) \rm + \it 2$^{2}$  $\times$ \it v0\rm (\rm 3\rm ) \rm + \it 2$^{1}$  $\times$  \it v0\rm (\rm 2\rm ) \rm + \it 2$^{0}$  $\times$  \it v0\rm (\rm 1\rm )\rm)
+%
+%\item\it uchar\_byte \rm = $\lambda$  \rm ( \it v0 \rm ) \rm .
+\rm ( \it v0  $\in$  \it UCHAR $\mid$ \rm [\rm (\it v0   $\mod$ \it 2$^{1}$\rm ) $\div$ \it 2$^{0}$\rm , \rm (\it v0  $\mod$  \it 2$^{2}$\rm ) $\div$ \it 2$^{1}$\rm , \rm (\it v0  $\mod$  \it 2$^{3}$\rm ) $\div$ \it 2$^{2}$\rm , \rm (\it v0  $\mod$  \it 2$^{4}$\rm ) $\div$ \it 2$^{3}$\rm , \rm (\it v0  $\mod$  \it 2$^{5}$\rm ) $\div$ \it 2$^{4}$\rm , \rm(\it v0  $\mod$  \it 2$^{6}$\rm ) $\div$ \it 2$^{5}$\rm ,\rm (\it v0  $\mod$  \it 2$^{7}$\rm ) $\div$ \it 2$^{6}$\rm , \rm (\it v0  $\mod$  \it 2$^{8}$\rm ) $\div$ \it 2$^{7}$\rm \rm ] \rm)
+%
+%\end{description}
+%
+%The equivalent definitions using B sequence notation and set notation are:\\
+%\it uchar\_byte \rm = \rm \{ \rm 0 $\mapsto$ \rm \rm [\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm \rm ]\rm, \ldots  , \rm 2\rm 5\rm 5 $\mapsto$ \rm \rm [\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm \rm ]\rm \}  \\
+%\it byte\_uchar \rm =  \rm \{\rm \rm [\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm \rm ] $\mapsto$ \rm 0\rm, \ldots , \rm \rm [\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm \rm ] $\mapsto$ \rm 2\rm 5\rm 5 \rm \} \rm
+%
+%\paragraph{B Rules:}
+%These rules represent that  $byte\_uchar$  and $uchar\_byte$  are bijective functions and  $byte\_uchar = uchar\_byte^{-1}$. These rules are used on interactive prover of Atelierb.
+
+\begin{enumerate}
+
+\item For the proof obligation (\it byte\_uchar\rm \ $\in$ \textit{BYTE} $\bij$
+\textit{UCHAR}), the proof rule is:\\
+$\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}  $\mid$  \rm 1\rm 2\rm 8 $\times$ \it v0\rm (\rm 8\rm )\rm +\rm 6\rm 4 $\times$ \it v0\rm (\rm 7\rm )\rm +\rm 3\rm 2 $\times$ \it v0\rm (\rm 6\rm )\rm +\rm 1\rm 6 $\times$ \it v0\rm (\rm 5\rm )\rm +\rm 8 $\times$ \it v0\rm (\rm 4\rm )\rm +\rm 4 $\times$ \it v0\rm (\rm 3\rm )\rm +\rm 2 $\times$ \it v0\rm (\rm 2\rm )\rm +\rm 1 $\times$ \it v0\rm (\rm 1\rm )\rm ) \rm $\in$ \rm (\rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}\rm )  $\bij$  \rm 0 $\upto$ \rm 2\rm 5\rm 5 \\
+
+\item For the proof obligation (\it uchar\_byte\rm \ $\in$ \textit{UCHAR} $\bij$
+\textit{BYTE}), the proof rule is:\\ $\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 0$\upto$255  $\mid$  \rm \rm  [(\it v0  $\mod$  \rm 2)$\div$1,(\it v0  $\mod$  \rm 4)$\div$\rm 2\rm,(\it v0  $\mod$  \rm 8)$\div$\rm 4\rm,(\it v0  $\mod$  \rm 16)$\div$\rm 8\rm,(\it v0  $\mod$  \rm 32)$\div$\rm 1\rm 6\rm,(\it v0  $\mod$  \rm 6\rm 4)$\div$\rm 3\rm 2\rm,(\it v0  $\mod$  \rm 1\rm 2\rm 8)$\div$\rm 6\rm 4\rm,(\it v0  $\mod$  \rm 2\rm 5\rm 6)$\div$\rm 1\rm 2\rm 8\rm \rm ]\rm ) \rm   $\in$  \rm 0 $\upto$ \rm 2\rm 5\rm 5   $\bij$ \rm (\rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}\rm )\\
+
+\item For the proof obligation (\it byte\_uchar\rm  =  \it
+uchar\_byte\rm$^{-1}$), the proof rule is:\\ $\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}  $\mid$  \rm 1\rm 2\rm 8 $\times$ \it v0\rm (\rm 8\rm )\rm +\rm 6\rm 4 $\times$ \it v0\rm (\rm 7\rm )\rm +\rm 3\rm 2 $\times$ \it v0\rm (\rm 6\rm )\rm +\rm 1\rm 6 $\times$ \it v0\rm (\rm 5\rm )\rm +\rm 8 $\times$ \it v0\rm (\rm 4\rm )\rm +\rm 4 $\times$ \it v0\rm (\rm 3\rm )\rm +\rm 2 $\times$ \it v0\rm (\rm 2\rm )\rm +\rm 1 $\times$ \it v0\rm (\rm 1\rm )\rm )=($\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 0$\upto$255  $\mid$  \rm \rm  [(\it v0  $\mod$  \rm 2)$\div$1,(\it v0  $\mod$  \rm 4)$\div$\rm 2\rm,(\it v0  $\mod$ \rm 8) $\div$\rm 4\rm,(\it v0  $\mod$  \rm 16)$\div$\rm 8\rm,(\it v0  $\mod$  \rm 32)$\div$\rm 1\rm 6\rm,(\it v0  $\mod$ \rm 6\rm 4)$\div$\rm 3\rm 2\rm,(\it v0  $\mod$  \rm 1\rm 2\rm 8)$\div$\rm 6\rm 4\rm,(\it v0  $\mod$  \rm 2\rm 5\rm 6)$\div$\rm 1\rm 2\rm 8\rm \rm ]\rm )\rm)$^{-1}$
+\end{enumerate}
+
+Eight proof obligations require that conversion functions are finite. %: \it
+% byte\_uchar $\in$ $\finset$\rm (\it byte\_uchar\rm ); \it uchar\_byte  $\in$
+% $\finset$\rm (\it uchar\_byte\rm ); \it ushort\_bv16 $ \in \finset$\rm (\it
+% ushort\_bv16\rm ); \it bv16\_ushort $ \in \finset$\rm (\it bv16\_ushort\rm ); \it
+% byte\_schar $ \in \finset$\rm (\it byte\_schar\rm ); \it schar\_byte $ \in
+% \finset$\rm (\it schar\_byte\rm ); \it sshort\_bv16 $ \in \finset$\rm (\it
+% sshort\_bv16\rm ); and \it bv16\_sshort $ \in \finset$\rm (\it bv16\_sshort\rm).
+ These proof obligations were verified using the expression evaluator of ProB.
+Finally, we have included rules related to binary arithmetic as developed
+in~\cite{Leibnizens}. These rules are used in the interactive theorem prover with
+binary arithmetic to prove important properties~\cite{James2010,Dasgupta2006}.
 
@@
 
-This section shows the elements that make up the different types of instructions.
-The main elements are specified in the microcontroller module Z80 and parts of it
-are presented below.  Basically, the state of microcontroller is formed by data 
-from program counter, ports, registers, etc. The transitions between the states 
-are actioned by execution of instruction or an external action. Groups of registers
-are represented by the variables of the specification states that appear in clause
-\textit{VARIABLES}. The declaration of valid states of variables is
-in the \textit{INVARIANT} clause and the initial state is defined in the
-\textit{INITIALISATION} clause. The assembly instructions are defined by the
-clause \textit{OPERATIONS} and, in general, each instruction has three actions: update 
-the program counter, update the flag register and its main effect.
-%Several functions were created to Z80.
-General functions that can be used in others microcontroller models are defined in
-the modules of data types. Specific functions of the microcontroller are defined in the ALU (arithmetic logic unit) module.
-
-The main module includes an instance of the memory module and accesses the definitions from basic data
-types modules and the \textit{ALU} module.
+This section shows the elements that make up the different types of
+instructions. The main elements are specified in the microcontroller module Z80
+and parts of it are presented below.  Basically, the state of microcontroller is
+formed by data from program counter, ports, registers, etc. The transitions
+between the states are actioned by execution of instruction or an external
+action. Groups of registers are represented by the variables of the
+specification states that appear in clause \textit{VARIABLES}. The declaration
+of valid states of variables is in the \textit{INVARIANT} clause and the initial
+state is defined in the \textit{INITIALISATION} clause. The assembly
+instructions are defined by the clause \textit{OPERATIONS} and, in general, each
+instruction has three actions: update the program counter, update the flag
+register and its main effect.
+% Several functions were created to Z80.
+General functions that can be used in other microcontroller models are defined
+in the modules of data types. Specific functions of the microcontroller are
+defined in the ALU (arithmetic logic unit) module.
+
+The main module includes an instance of the memory module and accesses the
+definitions from basic data types modules and the \textit{ALU} module.
 
@@
 
-%
-
 % Features of microcontroller
-  The internal registers contain 176 bits of read/write memory that are represented by
-identifiers used as parameters in the instructions . It includes two sets of six general purpose
-registers which may be used individually as 8-bit registers or as 16-bit register pairs.  The working
-registers are represented by variable $\textit{rgs8}$. The domain of $\textit{rgs8}$
-($\textit{id\_regs8}$) is a set formed by identifiers of registers of 8 bits. These registers are called
-main register set (\textit{a0,b0,c0,d0,e0,f0,h0,l0}) and alternate register set
-(\textit{a\_0,b\_0,c\_0,d\_0,e\_0,f\_0,h\_0,l\_0}). These registers can be accessed in pairs, forming
-16-bits, resulting in another set of identifiers of 16-bits registers, named $\textit{id\_reg16}$. These sets are represented below.
+The internal registers contains 176 bits of read/write memory that are
+represented by identifiers used as parameters in the instructions . It includes
+two sets of six general purpose registers which may be used individually as
+8-bit registers or as 16-bit register pairs.  The working registers are
+represented by variable $\textit{rgs8}$. The domain of $\textit{rgs8}$
+($\textit{id\_reg8}$) is a set formed by identifiers of registers of 8 bits.
+These registers are called main register set (\textit{a0,b0,c0,d0,e0,f0,h0,l0})
+and alternate register set (\textit{a\_0,b\_0,c\_0,d\_0,e\_0,f\_0,h\_0,l\_0}).
+These registers can be accessed in pairs, forming 16-bits, resulting in another
+set of identifiers of 16-bits registers, named $\textit{id\_reg16}$. These sets
+are represented below.
 
@@
 
-The external actions change the state of the microcontroller, for example, refreshing the I/O ports
-and interruptions request. The external actions are also modeled by operations and they are named with the
-prefix ``$ext\_$'' and followed by the name of action. There are four external
-actions: $ext\_update\_io\_ports$, $ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The
-$ext\_update\_io\_ports$ just change the state of I/O port, see.
+The external actions change the state of the microcontroller, for example,
+refreshing the I/O ports and interruption request. The external actions are also
+modeled by operations and they are named with the prefix ``$ext\_$'' and
+followed by the name of action. There are four external actions:
+$ext\_update\_io\_ports$, $ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The
+instruction $ext\_update\_io\_ports$ just changes the state of I/O port, see.
 
@@
 
-The interruptions and the \textit{reset} action can change the state of
-program counter. Theses actions are modeled by B operations and its main
-effects are shown below\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$ holds the value of stack pointer minus 2,
-%=$\textit{dec\_BV16(dec\_BV16(sp))}$
+The interruptions and the \textit{reset} action can change the state of program
+counter. Theses actions are modeled by B operations and its main effects are
+presented here\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$
+holds the value of stack pointer minus 2,
+% =$\textit{dec\_BV16(dec\_BV16(sp))}$
  $\textit{sp\_minus\_one}$ is the value of stack pointer minus 1,
-%=$\textit{dec\_BV16(sp)}$
-$\textit{pc\_high}$ holds the most significant 8 bits and
-$\textit{pc\_low}$ holds the least significant 8 bits.}.
+% =$\textit{dec\_BV16(sp)}$
+$\textit{pc\_high}$ holds the most significant 8 bits and $\textit{pc\_low}$
+holds the least significant 8 bits.}.
 
@@
 
-  \item The mode 0 is compatible with 8080 and this mode is selected when \it im \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a 
-  non-maskable interrupt happens, the CPU fetches an instruction of one byte
-  from an external device, usually an RST instruction, and the CPU executes it.
-  The instruction code is received from an external device by data bus and it is represented by integer parameter called $\textit{byte\_bus}$ .
-
-  \item The mode 1 is the simplest and this mode is selected when \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 1
-  \rm ). Simply, when a non-maskable interruption happens, the program counter
-  receives $38H$ (56 in decimal).
-
-  \item The mode 2 is the most flexible and this mode is selected when \it im \rm = \rm ( \rm 1  $\mapsto$
-  \rm 1 \rm ). When a non-maskable interruption happens, an indirect call can be
-  made to any address memory. The program counter receives a bit vector of size
-  16 composed with two parts: the most significant part of the $\textit{i\_}$
-  register and the least significant part of the $\textit{byte\_bus}$ with the
-  last bit cleared.
+  \item The mode 0 is compatible with 8080 and this mode is selected when \it im
+  \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a  non-maskable interrupt
+  happens, the CPU fetches an instruction of one byte from an external device,
+  usually an RST instruction, and the CPU executes it.
+  The instruction code is received from an external device by data bus and it is
+  represented by integer parameter called $\textit{byte\_bus}$.
+
+  \item The mode 1 is the simplest and this mode is selected when \it im \rm =
+  \rm ( \rm 0  $\mapsto$  \rm 1 \rm ). Simply, when a non-maskable interruption
+  happens, the program counter receives $38H$ (56 in decimal).
+
+  \item The mode 2 is the most flexible and this mode is selected when \it im
+  \rm = \rm ( \rm 1  $\mapsto$ \rm 1 \rm ). When a non-maskable interruption
+  happens, an indirect call can be made to any address memory. The program
+  counter receives a bit vector of size 16 composed with two parts: the most
+  significant part of the $\textit{i\_}$ register and the least significant part
+  of the $\textit{byte\_bus}$ with the last bit cleared.
 
@@
 
-The modeling of instruction set, interruptions and input/output ports become the animation interesting.
-The animation in ProB version is not so fast and practical as a simulation, for example in \cite{Simulator_z80},
-because ProB works manipulating logic and math concepts that can have infinite sets and non determinism, then ProB requires more processing, specially when it works analysing  complex expressions or very large expressions.
-However, a user of ProB with this modeling can: execute manually a sequence of Z80's instructions that represent a program and view the visited states of ports, memory and registers, it is important to debug the effect action on the state space computed; create a tracing of the states; analyse constants, functions and expressions; and search possible violations.
-
-%\subsection{Considerations about the Z80 model}
+The modeling of instruction set, interruptions and input/output ports become the
+animation interesting. The animation in ProB is not so fast and practical as a
+simulation, for example in \cite{Simulator_z80}, because ProB works manipulating
+logic and math concepts that can have infinite sets and non determinism, then
+ProB requires more processing, specially when it works analysing  complex
+expressions or very large expressions. However, a user of ProB with this
+modeling can: execute manually a sequence of Z80's instructions that represent a
+program and view the visited states of ports, memory and registers. It is
+important to debug the effect action on the state space computed; create a
+tracing of the states; analyse constants, functions and expressions; and search
+possible violations.
+
+\subsection{Verification process}% ESTOU TRABALHANDO AQUI
+% OK -> como as regras foram utilizadas ?  Respondido na seção das regras
+% TODO:ENGLISH REVIEW parágrafo inicial: VISÃO GERAL
+This project has many proof obligations and its verification can be a hard task,
+hence this section explains about the verification process of proof
+obligations and its importance. % [O que as provas garantem? citar que énecessário buscar técnicas eficientes para realizar as provas ]
+The proof obligations allow to verify the data types, important system properties and if
+the expressions are well-defined (WD)\footnote{An expression is called ``well-defined''
+(or unambiguous) if its definition assigns it a unique interpretation or value.}. The
+properties provide additional guarantees because they can set many safety
+rules. However, the model can be very difficult to prove.
+
+ % -> O que significa cada coluna e linha da tabela ?
+The Table~\ref{tab:completestatistcs} contains some statistics of verification
+process, it shows only the number of no obvious proof
+obligations\footnote{Obvious proof obligations are simple formulas automatically eliminated by
+the tool.} and the titles of columns use acronyms: {\small \textbf{PO}} - number
+of proof obligations; {\small \textbf{UPO}} - number of unproved proof
+obligations; {\small \textbf{WD PO}} - number of proof obligations  related to
+well definition lemmas; {\small \textbf{UWD PO}} - number of unproved proof
+obligations related to well definition lemmas.
+
+\begin{table} [h]
+\caption{Verification statistics }
+\label{tab:completestatistcs}
+\begin{center} 
+\begin{tabular}{|p{3.1cm}|c|c||c|c||c|c|c|}
+\hline 
+{\small \textbf{Component}}&{\small \textbf{PO}}&{\small \textbf{UPO}}&{\small \textbf{WD PO}}&{\small \textbf{UWD PO }}&{\small \textbf{Total PO}}&{\small \textbf{Total UPO}}& {\small \textbf{Total Percentage}}  \\\hline
+BIT&49&0&69&0&118&0&100\%\\\hline
+BV16&6&0&69&0&75&0&100\%\\\hline
+BYTE&18&0&136&0&154&0&100\%\\\hline
+POWER&3&0&4&0&7&0&100\%\\\hline
+POWER2&18&0 &0&0&18&0&100\%\\\hline
+SCHAR&4&0&26&0&30&0&100\%\\\hline
+SSHORT&4&0&58&0&62&0&100\%\\\hline
+TYPES&71&0&0&0&71&0&100\%\\\hline
+UCHAR&4&0&26&0&30&0&100\%\\\hline
+USHORT&4&0&58&0&62&0&100\%\\\hline
+ALU&51&22&117&47&168&69&59\%\\\hline
+MEMORY&13&0&0&0&13&0&100\%\\\hline
+Z80 -  {\scriptsize Initialization, Properties, Assertions} &66&9&150&11&216&20&91\%\\\hline
+Z80 - {\scriptsize Input and Output}&46&11&281&16&327&27&92\%\\\hline
+Z80 - {\scriptsize Logic Arithmetic 1}&88&33&343&84&431&117&73\%\\\hline
+Z80 - {\scriptsize Logic Arithmetic 2}&40&20&178&27&218&47&79\%\\\hline
+Z80 - {\scriptsize Bit Control}&111&34&763&157&874&191&78\%\\\hline
+Z80 - {\scriptsize External Actions}&31&0&60&38&91&38&59\%\\\hline
+Z80 - {\scriptsize General 1}&81&19&420&20&501&39&92\%\\\hline
+Z80 - {\scriptsize General 2}&26&10&139&14&165&24&86\%\\\hline
+
+\textbf{Total}&\textbf{734}&\textbf{158}&\textbf{2897}&\textbf{364}&\textbf{3631}&\textbf{569}&\textbf{84}\%\\\hline
+
+\end{tabular}
+\end{center} 
+\end{table}
+
+%->Explicaçao de mais detalhes da tabela.
+
+% Porque o modelo não foi verificado completamente ?
+There are many benefits when a formal model is developed, but to verify the
+model is not an easy task, the first version of Z80 model have already verified
+completely. However, this new Z80 model with support to animation in ProB has not
+been verified completely because the changes in Z80 model are recent and the
+verification time is long.
+% ->Quanto tempo foi gasto nessa verificação ?  Aproximadamente quanto tempo de
+% processamento elas reduziram ?
+Nevertheless this project has 3631 proof obligations and the first version of
+Z80 model spent approximately two months of work to finish the verification
+process. The same  developed verification strategies were adjusted and reused in
+new Z80 model, so approximately  7 days of work was enough to verify 3062  proof
+obligations (84\%) and remaining only 569 proof obligations (16\%) not yet
+verified.
+
+Two development strategies were important to improve the verification of 3062
+proof obligations: the intensive use of parallelism and user pass. The Z80 module
+was decomposed in 8 modules of instructions, it was very helpful to decrease the
+time for verification. Because AtelierB can run simultaneously \textit{N} instances of 
+provers, one per module, reducing the time verification by up to \textit{N}  times when
+a processor with \textit{N} cores is used  or the process is submitted to \textit{N} 
+networked computers. So we also set up a proving environment consisting of networked
+computers to take advantage of the distribution facilities provided in the B development
+environment.
+% Resumo de como aconteceu o processo de verificação,
+A second interesting strategy  was to reuse the user pass. Usually, an used
+user pass to aid the verification of one proof obligation can solve several
+similar proof obligations. Thus, few generic user pass are needed to prove many
+proof obligations. As there are many similar assembly instructions, some
+human-directed proofs, when replayed, could discharge other proof obligations. A
+good example is a set of 17 user pass that quickly aided the verification of
+approximately 2533 of WD proofs.
+
+\subsection{Considerations about the Z80 model}
 
@@
 
-\section{Related works}
+%\section{Case study}
+%\include{Petroleumproductiontestsystem}
+
+\section{Related work}
 \label{sec:relatedworks}
 
-There are in the literature several similar approaches to model hardware and instruction set using B.
-The paper \cite{springerlink:Yuan2011} reports a method to specify, design and construct sound and complete instruction set model by stepwise refinement and formal proof using the formal method Event-B.
-It also discuss about desirable properties of instruction set model.
-However, our work \cite{Valerio_SBMF09} and \cite{Subotic2010} use the B method, that seems more appropriated to software development, because it has implementable language defined,
-called B0, and tools to convert the models to a programming language.
-In \cite{Marc20113} is present an related experience on using B in the design of secure micro-controllers.
-It tries to show the feasibility of such a technique for high confidence trustful devices.
-The work \cite{Subotic2010} describes a similar approach with MIPS CPU architecture and it shows how the CPU is formally specified and implemented using the B method.
+There exists several approaches to model hardware and instruction set
+using B a.  The paper \cite{springerlink:Yuan2011} reports a method to
+specify, design and construct sound and complete instruction set model
+by stepwise refinement and formal proof using Event-B. It also
+discusses desirable properties of the instruction set model. However,
+our work \cite{Valerio_SBMF09} and \cite{Subotic2010} use the B
+method, that seems quite appropriated to software development, because
+it has an implementable language defined, called B0, a model-driven
+approach \cite{Dantas_SBMF08} to develop software from the functional
+specification level down to assembly, and tools to convert the models
+to a programming language.
+
+A related experience on using B in the design of secure
+micro-controllers is present in \cite{Marc20113}. It tries to show the
+feasibility of such a technique for high confidence trustful
+devices. The work \cite{Subotic2010} describes a similar approach with
+MIPS CPU architecture and it shows how the CPU is formally specified
+and implemented using the B method.
 
@@
 % Um pequeno resumo do trabalho
-This work has shown an approach to the formal modeling of the instruction set of
-microcontroller using the B method. During the construction of this model, some
-problems were found in the official reference for Z80
-microcontroller~\cite{Z80_manual}. The designer fixed the found problems in
-documentation and also developed a case study: a verified embedded software. This
-case study was interesting to analyse the technique of formal verification up to
-assembly level.
-
-% TODO: Adaptar para apenas fundamentar melhor o trabalho já desenvolvido. Citar o artigo SBMF e a dissertação. 
+This work has shown an approach to the formal modeling of the
+instruction set of microcontroller using the B method.  During the
+construction of this model, some problems were found in the official
+reference for Z80 microcontroller~\cite{Z80_manual}. We have consulted
+different sources \cite{Simulator_z80,UndocumentedZ80,Z80_manual} to
+avoid such problems in modeling and have developed a case study: a
+verified embedded software.  This case study was interesting to as a
+feasibility analysis of the approach to derive formally assembly code
+from a specification of functional requirements in the B method.
+
+%TODO: COLOCAR AQUI OS PRINCIPAIS BENEFÍCIOS PROPORCIONADOS PELO TRABALHO
+
+% fundamenta melhor o trabalho já desenvolvido. Citar o artigo SBMF e a dissertação. 
 % Os meus trabalhos já desenvolvidos/ resultados atuais.
-The next works quoted are directly related to the objective this paper.
-The first work \cite{DantasSemish2008} shows the developed metodology to verify
-software up to assembly level using B. A second work\cite{Dantas_SBMF08}
-presents more details about the verification approach and a small example
-software verified up to assembly level in three different platforms. 
-The last work \cite{Valerio_SBMF09} shows a general view of the first version of Z80
-model and techniques used in verification process.
+The following works are directly related to the objective this paper.
+The approach to develop verified software down to the assembly level
+using B was described in~\cite{DantasSemish2008}. A first case study
+for this approach was reported in~\cite{Dantas_SBMF08}, presenting
+more details as well as a small software that developed up to assembly
+level in three different platforms. A general view of a previous used
+in verification process were presented in~\cite{Valerio_SBMF09}. The
+model presented in the current paper added support for animation in
+ProB and the representation of external actions (interruptions).
 %Link com o parágrafo anterior 
-These works created a important step to software verification up to assembly
-language, they show the actuals difficulties and suggest improvements in the
-B tools.
-
-% TODO
+This set of papers present an important step to software verification
+up to assembly language, showing the current difficulties and
+suggesting improvements in tools supporting the B method.
+
 % [Trabalhos futuros ] LLVM:Existe também a pretensão de especificar as
-
-Future works comprise the development of software with the B method from
-functional specification to assembly level, using the Z80 model presented in this
-work. The mechanic compilation from B algorithmic constructs to assembly platform
-is also envisioned. Finally, another important activity is develop a formal model of a platform used in LLVM Compiler~\cite{DBLP:conf/cgo/LattnerA04}.
-
-\subsubsection{Acknowledgements:} This experiment could not have succeeded without the interesting work of Michael Leuschel who developed several improvements in the ProB and provided important aid to enable the animation of the Z80 modeling.
+In the future, we plan to improve the support for the development of
+software with the B method from functional specification to assembly
+level, using the Z80 model presented in this work. For instance, the
+mechanic compilation from B algorithmic constructs to assembly
+platform is also envisioned. Finally, another important activity is to
+develop a formal model of a platform used in LLVM
+Compiler~\cite{DBLP:conf/cgo/LattnerA04}, which would enable us to
+integrate different compilation techniques into our B-based framework.
+
+% Conclusão: toque final
+%Several iterations were needed to provide the good library definitions as well as
+%to fine-tune the model of the microcontroller instructions by factoring common
+%functionalities into auxiliary definitions.
+
+\subsubsection{Acknowledgements:} The animation of the model would not
+have been possible without the help of Michael Leuschel who kindly
+provided feedback and developed several improvements to ProB to meet our
+needs.
 

### 2012-05-15 trunk/papers/ArticleZ80/source/paper.tex
novos pequenos ajustes2

@@
 user pass to aid the verification of one proof obligation can solve several
-similar proof obligations. Thus, few generic user pass are needed to prove many
+similar proof obligations. Thus, few generic user passes are needed to prove many
 proof obligations. As there are many similar assembly instructions, some
 human-directed proofs, when replayed, could discharge other proof obligations. A
-good example is a set of 17 user pass that quickly aided the verification of
+good example is a set of 17 user passes that quickly aided the verification of
 approximately 2533 of WD proofs.

### 2012-05-17 trunk/papers/ArticleZ80/source/paper.tex

--- 2012-05-15 trunk/papers/ArticleZ80/source/paper.tex
@@
 %%% PAGE DIMENSIONS
-\usepackage{geometry} % to change the page dimensions
+%\usepackage{geometry} % to change the page dimensions
 \usepackage{graphicx,array}
@@
 
-The modeling of instruction set, interruptions and input/output ports become the
-animation interesting. The animation in ProB is not so fast and practical as a
-simulation, for example in \cite{Simulator_z80}, because ProB works manipulating
-logic and math concepts that can have infinite sets and non determinism, then
-ProB requires more processing, specially when it works analysing  complex
-expressions or very large expressions. However, a user of ProB with this
-modeling can: execute manually a sequence of Z80's instructions that represent a
-program and view the visited states of ports, memory and registers. It is
-important to debug the effect action on the state space computed; create a
-tracing of the states; analyse constants, functions and expressions; and search
-possible violations.
-
-\subsection{Verification process}% ESTOU TRABALHANDO AQUI
+\subsection{Verification and validation}% ESTOU TRABALHANDO AQUI
 % OK -> como as regras foram utilizadas ?  Respondido na seção das regras
 % TODO:ENGLISH REVIEW parágrafo inicial: VISÃO GERAL
-This project has many proof obligations and its verification can be a hard task,
-hence this section explains about the verification process of proof
-obligations and its importance. % [O que as provas garantem? citar que énecessário buscar técnicas eficientes para realizar as provas ]
-The proof obligations allow to verify the data types, important system properties and if
-the expressions are well-defined (WD)\footnote{An expression is called ``well-defined''
-(or unambiguous) if its definition assigns it a unique interpretation or value.}. The
-properties provide additional guarantees because they can set many safety
-rules. However, the model can be very difficult to prove.
-
- % -> O que significa cada coluna e linha da tabela ?
-The Table~\ref{tab:completestatistcs} contains some statistics of verification
-process, it shows only the number of no obvious proof
-obligations\footnote{Obvious proof obligations are simple formulas automatically eliminated by
-the tool.} and the titles of columns use acronyms: {\small \textbf{PO}} - number
-of proof obligations; {\small \textbf{UPO}} - number of unproved proof
-obligations; {\small \textbf{WD PO}} - number of proof obligations  related to
-well definition lemmas; {\small \textbf{UWD PO}} - number of unproved proof
-obligations related to well definition lemmas.
+
+The model presented in this paper is suitable both for validation
+through animation with ProB, and for verification through the standard
+proof obligations generated by the application of the B method.
+
+The animation in ProB is not as fast and practical as a simulation
+with a dedicated tool (e.g. \cite{Simulator_z80}). ProB is a
+general-purpose animator of formal specification, and addresses
+computationally more difficult problems. Anyway, it is capable of
+animating, and also possibly model checking, a formal model of the
+instruction set of the Z80, which was our initial motivation in
+constructing that model. One may simulate step-by-step sequences of
+Z80 instructions and visualize the visited states of ports, memory and
+registers. It is a useful tool to debug by creating a trace of the
+visited states, offering to analyse constants, functions and
+expressions, and search possible violations.
+
+The construction of the model with B generates a large number of proof
+obligations and the verification is a time-consuming task. We describe
+here some details on this effort and some solutions we have devised to
+reduce the effort.  In B, the generated proof obligations allow to
+verify the data satisfy invariant properties, check important system
+properties and that the expressions are well-defined (WD)\footnote{An
+  expression is called ``well-defined'' (or unambiguous) if its
+  definition assigns it a unique interpretation or value.}.
+
+Table~\ref{tab:completestatistcs} displays statistics of the
+verification process. Note that the figures in it include only the
+number of \emph{non obvious} proof obligations\footnote{Obvious proof
+  obligations are simple formulas automatically eliminated by the
+  tool.}. The titles of columns use acronyms: {\small \textbf{PO}} -
+number of proof obligations; {\small \textbf{UPO}} - number of (yet)
+unproved proof obligations; {\small \textbf{WD PO}} - number of proof
+obligations related to well definition lemmas; {\small \textbf{UWD
+    PO}} - number of (yet) unproved proof obligations related to well
+definition lemmas.
 
@@
 % Porque o modelo não foi verificado completamente ?
-There are many benefits when a formal model is developed, but to verify the
-model is not an easy task, the first version of Z80 model have already verified
-completely. However, this new Z80 model with support to animation in ProB has not
-been verified completely because the changes in Z80 model are recent and the
-verification time is long.
+There is still a large number of open proof obligations, i.e. we have
+not yet found their proof, or they are not valid and our model is not
+correct. Each such proof obligation requires interaction with a proof
+assistant. We see that for a project of that size, a large effort
+needs to be dedicated to find those proofs. One interesting feature of
+the proof assistant is, once a proof has been found, it can
+automatically apply it to other proof obligations, as it is often the
+case that the same proof script discharges several proof obligations.
+
 % ->Quanto tempo foi gasto nessa verificação ?  Aproximadamente quanto tempo de
 % processamento elas reduziram ?
-Nevertheless this project has 3631 proof obligations and the first version of
-Z80 model spent approximately two months of work to finish the verification
-process. The same  developed verification strategies were adjusted and reused in
-new Z80 model, so approximately  7 days of work was enough to verify 3062  proof
-obligations (84\%) and remaining only 569 proof obligations (16\%) not yet
-verified.
-
-Two development strategies were important to improve the verification of 3062
-proof obligations: the intensive use of parallelism and user pass. The Z80 module
-was decomposed in 8 modules of instructions, it was very helpful to decrease the
-time for verification. Because AtelierB can run simultaneously \textit{N} instances of 
-provers, one per module, reducing the time verification by up to \textit{N}  times when
-a processor with \textit{N} cores is used  or the process is submitted to \textit{N} 
-networked computers. So we also set up a proving environment consisting of networked
-computers to take advantage of the distribution facilities provided in the B development
-environment.
-% Resumo de como aconteceu o processo de verificação,
-A second interesting strategy  was to reuse the user pass. Usually, an used
-user pass to aid the verification of one proof obligation can solve several
-similar proof obligations. Thus, few generic user pass are needed to prove many
-proof obligations. As there are many similar assembly instructions, some
-human-directed proofs, when replayed, could discharge other proof obligations. A
-good example is a set of 17 user pass that quickly aided the verification of
-approximately 2533 of WD proofs.
+We had a previous experience with a similar effot, where we were able
+to discharge all the proof obligations: that of the model of the Z80
+instruction set designed without restricting ourselves to constructs
+that may be animated with ProB, and without modeling external actions
+(as described in section~\ref{sec:externalactions}). In that first
+experience, the verification took us about two months. In that second
+project, we have been able to verify 3062 (e.g. roughly 80 \%) of 3631
+proof obligations in one week.
+
+In order to be able to verify such a large number of proof obligations
+in such a short term, we employed two strategies: parallelism and
+\emph{user pass}. AtelierB has an option to run any number of threads
+in parallel and also to send proof obligations to other, possibly
+remote, processes running AtelierB. With those options we had the
+possibility to have eight concurrent instances of the prover. We thus
+split the instruction set of the Z80 in eight sub-modules and checked
+them concurrently (each instance of AtelierB may only verify one module
+at a time).
+
+The second strategy, called \emph{user pass}, takes advantage of
+another feature of the AtelierB prover. We defined a number of proof
+scripts which are then applied to all remaining proof obligations. Few
+generic user passes are enough to prove many proof obligations. Also,
+as the models of many instructionsa are similar, proofs found in
+interactive mode may similarly be replayed and discharge other proof
+obligations. We have defined a set of 17 user passes that are able
+to discharge more than 2500 WD proof obliagations.
 
@@
 
-The Z80 formal model provides many benefits, because of the verification of
-generated proof obligations guarantees: the correct use of data types, the developed
-security properties  and the well-defined of all the expressions. 
-Furthermore, the designer has a big flexibility to create new and specific security properties, this is very useful %TODO: COLOCAR APÓS TODOS OS BENEFICIOS PROPORCIONADOS COM A SIMULAÇÃO e COMPATIBILIDADE COM O PROB.
-(OBTER DO PARAGRAFO ANTERIOR DA SEÇAO VERIFICATION PROCESS).
-to adjust the verification in accordance with the requirements. Moreover, this
-example of model could replace or improve the used documentation for users and
-assembly programmers. Z80 model is also useful to develop verified software 
-up to the assembly level.
-
-A case study using the first version of Z80 model was also created and verified
-since the first abstract B model up to B assembly model. This case study is an
-object from the pilot project developed to analyse a petroleum production test
-system in each oil field. Its modeling was developed according to \cite{LAUT_SERGIO}
-and its B assembly model can be executed manually in ProB. The modeling of
-instruction set and the B assembly model allow to simulate the execution of microcontroller
-and analyse several important properties.
-
-%\section{Case study}
-%\include{Petroleumproductiontestsystem}
+The Z80 formal model provides many benefits, because of the
+verification of generated proof obligations guarantees: the correct
+use of data types, the developed security properties and the
+well-defined of all the expressions.  Furthermore, the designer has a
+big flexibility to create new and specific security properties, this
+is very
+useful %TODO: COLOCAR APÓS TODOS OS BENEFICIOS PROPORCIONADOS COM A SIMULAÇÃO e COMPATIBILIDADE COM O PROB. (OBTER DO PARAGRAFO ANTERIOR DA SEÇAO VERIFICATION PROCESS).
+to adjust the verification in accordance with the
+requirements. Moreover, this example of model could complement the
+existing documentation for users and assembly programmers.
+
+But our goal is to use the model of the Z80 instruction set to extend
+the scope of the B method up to the assembly
+level~\cite{Dantas_SBMF08}. Indeed, we realized a case study where an
+abstract B model had been refined up to the algorithmic model using
+the classic development strategy. The algorithmic model was manually
+compiled to a Z80 assembly program. We then constructed a B module
+composed with the instructions of that program, and importing the
+formal model of the Z80 instruction set. We then showed that this
+assembly-level module was indeed a refinement of the original abstract
+model.
+
+This case study is a component of a pilot project to analyse a
+petroleum production test system. Its modeling was developed according
+to \cite{LAUT_SERGIO} and its B assembly model could also be animated
+with ProB.
 

### 2012-05-19 trunk/papers/ArticleZ80/source/paper.tex
Corrigido o formato. Estou resumindo o texto para as 16 páginas.

@@
 The new model supports animation in ProB~\cite{proB}.  Essentially, the
-new model\footnote{The interested reader in more details is invited to
-  visit our repository at: http://code.google.com/p/b2asm/.}  no
-longer contains definitions derived from infinite sets. Using B
+new model \footnote{The interested reader in more details is invited to visit our repository at: http://code.google.com/p/b2asm/.}
+no longer contains definitions derived from infinite sets. Using B
 constructs for modularity, auxiliary libraries of basic modules were
@@
 
-\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.03in} $\land$ \hspace*{0.0in}\it BYTE \rm =\rm (\it BYTE\_INDEX  $\fun$  \it BIT\rm )\hspace*{0.0in} $\land$
+\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.03in} $\land$
+
+\hspace*{0.0in}\it BYTE \rm =\rm (\it BYTE\_INDEX  $\fun$  \it BIT\rm )\hspace*{0.0in} $\land$
 
@@
 \item For the proof obligation (\it byte\_uchar\rm \ $\in$ \textit{BYTE} $\bij$
-\textit{UCHAR}), the proof rule is:\\
-$\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}  $\mid$  \rm 1\rm 2\rm 8 $\times$ \it v0\rm (\rm 8\rm )\rm +\rm 6\rm 4 $\times$ \it v0\rm (\rm 7\rm )\rm +\rm 3\rm 2 $\times$ \it v0\rm (\rm 6\rm )\rm +\rm 1\rm 6 $\times$ \it v0\rm (\rm 5\rm )\rm +\rm 8 $\times$ \it v0\rm (\rm 4\rm )\rm +\rm 4 $\times$ \it v0\rm (\rm 3\rm )\rm +\rm 2 $\times$ \it v0\rm (\rm 2\rm )\rm +\rm 1 $\times$ \it v0\rm (\rm 1\rm )\rm ) \rm $\in$ \rm (\rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}\rm )  $\bij$  \rm 0 $\upto$ \rm 2\rm 5\rm 5 \\
+\textit{UCHAR}), the proof rule is: $\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}  $\mid$  \rm 1\rm 2\rm 8 $\times$ \it v0\rm (\rm 8\rm )\rm +\rm 6\rm 4 $\times$ \it v0\rm (\rm 7\rm )\rm +\rm 3\rm 2 $\times$ \it v0\rm (\rm 6\rm )\rm +\rm 1\rm 6 $\times$ \it v0\rm (\rm 5\rm )\rm +\rm 8 $\times$ \it v0\rm (\rm 4\rm )\rm +\rm 4 $\times$ \it v0\rm (\rm 3\rm )\rm +\rm 2 $\times$ \it v0\rm (\rm 2\rm )\rm +\rm 1 $\times$ \it v0\rm (\rm 1\rm )\rm ) \rm $\in$ \rm (\rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}\rm )  $\bij$  \rm 0 $\upto$ \rm 2\rm 5\rm 5 \\
 
 \item For the proof obligation (\it uchar\_byte\rm \ $\in$ \textit{UCHAR} $\bij$
-\textit{BYTE}), the proof rule is:\\ $\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 0$\upto$255  $\mid$  \rm \rm  [(\it v0  $\mod$  \rm 2)$\div$1,(\it v0  $\mod$  \rm 4)$\div$\rm 2\rm,(\it v0  $\mod$  \rm 8)$\div$\rm 4\rm,(\it v0  $\mod$  \rm 16)$\div$\rm 8\rm,(\it v0  $\mod$  \rm 32)$\div$\rm 1\rm 6\rm,(\it v0  $\mod$  \rm 6\rm 4)$\div$\rm 3\rm 2\rm,(\it v0  $\mod$  \rm 1\rm 2\rm 8)$\div$\rm 6\rm 4\rm,(\it v0  $\mod$  \rm 2\rm 5\rm 6)$\div$\rm 1\rm 2\rm 8\rm \rm ]\rm ) \rm   $\in$  \rm 0 $\upto$ \rm 2\rm 5\rm 5   $\bij$ \rm (\rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}\rm )\\
+\textit{BYTE}), the proof rule is: $\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 0$\upto$255  $\mid$  \rm \rm  [(\it v0  $\mod$  \rm 2)$\div$1,(\it v0  $\mod$  \rm 4)$\div$\rm 2\rm,(\it v0  $\mod$  \rm 8)$\div$\rm 4\rm,(\it v0  $\mod$  \rm 16)$\div$\rm 8\rm,(\it v0  $\mod$  \rm 32)$\div$\rm 1\rm 6\rm,(\it v0  $\mod$  \rm 6\rm 4)$\div$\rm 3\rm 2\rm,(\it v0  $\mod$  \rm 1\rm 2\rm 8)$\div$\rm 6\rm 4\rm,(\it v0  $\mod$  \rm 2\rm 5\rm 6)$\div$\rm 1\rm 2\rm 8\rm \rm ]\rm ) \rm   $\in$  \rm 0 $\upto$ \rm 2\rm 5\rm 5   $\bij$ \rm (\rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}\rm )\\
 
@@
 in the modules of data types. Specific functions of the microcontroller are
-defined in the ALU (arithmetic logic unit) module.
-
-The main module includes an instance of the memory module and accesses the
+defined in the ALU (arithmetic logic unit) module.The main module includes an instance of the memory module and accesses the
 definitions from basic data types modules and the \textit{ALU} module.
@@
 
-\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ \it pc  $\in$  \it INSTRUCTION  $\land$  \it sp  $\in$  \it BV16 $\land$  \it ix  $\in$  \it BV16  $\land$
-
-\hspace*{0.10in}\it iy  $\in$  \it BV16  $\land$ \it i\_  $\in$  \it BYTE  $\land$  \it r\_ $\in$  \it BYTE  $\land$ \it iff1  $\in$  \it BIT  $\land$ \it iff2  $\in$  \it BIT  $\land$
-
-\hspace*{0.10in}\it im $\in$ (\it BIT $\times$ \it BIT\rm )  $\land$  \it i\_o\_ports  $\in$  \it BYTE  $\fun$  \it BYTE
+\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ \it pc  $\in$  \it INSTRUCTION  $\land$  \it sp  $\in$  \it BV16 $\land$
+
+\hspace*{0.10in}\it ix  $\in$  \it BV16  $\land$ \it iy  $\in$  \it BV16  $\land$ \it i\_  $\in$  \it BYTE  $\land$  \it r\_ $\in$  \it BYTE  $\land$ \it iff1  $\in$  \it BIT  $\land$
+
+\hspace*{0.10in}\it iff2  $\in$  \it BIT  $\land$ \it im $\in$ (\it BIT $\times$ \it BIT\rm )  $\land$  \it i\_o\_ports  $\in$  \it BYTE  $\fun$  \it BYTE
 \end{sloppypar}
@@
 \bf SETS\\
-\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0\rm, \it f0\rm, \it b0\rm, \it
+\hspace*{0.10in}\it id\_reg\_8\rm =\rm \{\it a0\rm, \it f0\rm, \it b0\rm, \it
 c0\rm, \it d0\rm, \it e0\rm, \it h0\rm, \it l0\rm, \it a\_0\rm,  \it f\_0\rm,
 \it b\_0\rm, \it c\_0\rm, \it d\_0\rm, \it e\_0\rm, \it h\_0\rm, \it l\_0 \rm
-\};\\ \hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{  \it AF \rm, \it BC\rm, \it
+\};\\ \hspace*{0.10in}\it id\_reg\_16\rm =\rm \{\it AF \rm, \it BC\rm, \it
 DE\rm, \it HL\rm, \it SP\rm\}
@@
 used as a \textbf{flag register}. This register uses only six bits to represent the execution result status of each instruction.
-According to the official manual the bits 3 and 5 are not used and the others bits have the follow meaning: %$\textit{bv\_get(rgs8(f0),0)}$ - the carry bit; $\textit{bv\_get(rgs8(f0),1)}$ - the add/subtract bit; $\textit{bv\_get(rgs8(f0),2)}$ - the parity or overflow bit; $\textit{bv\_get(rgs8(f0),4)}$ - the half carry bit; $\textit{bv\_get(rgs8(f0),6)}$ - the zero bit; and $\textit{bv\_get(rgs8(f0),7)}$ - the sign bit.
-\begin{description}
-  \item[$\textit{bv\_get(rgs8(f0),0)}$] - The Carry bit.
-  \item[$\textit{bv\_get(rgs8(f0),1)}$] - The Add/Subtract bit.
-  \item[$\textit{bv\_get(rgs8(f0),2)}$] - The Parity or Overflow bit.
-  \item[$\textit{bv\_get(rgs8(f0),4)}$] - The Half Carry bit.
-  \item[$\textit{bv\_get(rgs8(f0),6)}$] - The Zero bit.
-  \item[$\textit{bv\_get(rgs8(f0),7)}$] - The Sign bit.
-\end{description}
+According to the official manual the bits 3 and 5 are not used and the others bits have the follow meaning: $\textit{bv\_get(rgs8(f0),0)}$ - the carry bit; $\textit{bv\_get(rgs8(f0),1)}$ - the add/subtract bit; $\textit{bv\_get(rgs8(f0),2)}$ - the parity or overflow bit; $\textit{bv\_get(rgs8(f0),4)}$ - the half carry bit; $\textit{bv\_get(rgs8(f0),6)}$ - the zero bit; and $\textit{bv\_get(rgs8(f0),7)}$ - the sign bit.
+%\begin{description}
+%  \item[$\textit{bv\_get(rgs8(f0),0)}$] - The Carry bit.
+%  \item[$\textit{bv\_get(rgs8(f0),1)}$] - The Add/Subtract bit.
+%  \item[$\textit{bv\_get(rgs8(f0),2)}$] - The Parity or Overflow bit.
+%  \item[$\textit{bv\_get(rgs8(f0),4)}$] - The Half Carry bit.
+%  \item[$\textit{bv\_get(rgs8(f0),6)}$] - The Zero bit.
+%  \item[$\textit{bv\_get(rgs8(f0),7)}$] - The Sign bit.
+%\end{description}
 These bits can also be used to specify security properties for microcontroller programs.
@@
 \textbf{Assuring the absence of overflow:}
- \emph{To assure that an overflow does not happen, the developer can add this
- expression (\textit{bv\_get(rgs8(f0),0)} $\neq$ 1 $\land$
- \textit{bv\_get(rgs8(f0),2) $\neq$ 1}) in the invariant.
+ \emph{To assure that an overflow does not happen, the developer can add in the invariant this 
+ expression:\\(\textit{bv\_get(rgs8(f0),0)} $\neq$ 1 $\land$
+ \textit{bv\_get(rgs8(f0),2) $\neq$ 1}).
 % When a overflow happen,it can be dangerous. So the developer may restrict its use. However, this restriction can also
@@
 \subsection{Modeling the actions and instructions}
-
+%TODO: Fiz uma breve alteração
 Each instruction is represented by a B operation in Z80 module. The main
-module (\textit{Z80}) has three categories of operations: a category represents the 
-instructions of microcontrollers, a second category represents the input and output of data 
-(shown in ~\ref{sec:modelingIO}) and the last category represents the external
-actions (shown in~\ref{sec:externalactions}). A simple example of instruction is
-$\textit{LD\_(nn)\_A}$\footnoteremember{myfootnote}{The B tools does not allow
-to use parentheses in identifiers, and the characters \{``('',``)''\} are
+module (\textit{Z80}) has two categories of operations: a category represents the 
+instructions of microcontrollers and a category represents the external actions, e.g.
+interruptions actions. All 158 instructions and interruptions were modeled, but only
+two important examples are presented below.
+
+A simple example of instruction is $\textit{LD\_(nn)\_A}$\footnote{The B tools does not allow o use parentheses in identifiers, and the characters \{``('',``)''\} are
 replaced respectively by \{``9'',``0''\} in the actual specification.} shown below. The
@@
 integers values in decimal representation. This instruction use the
-$\textit{updateAddressMem}$ operation from \textit{Memory} module and it receives
-a address memory and its new memory value. It also increments the program
-counter ($\textit{pc}$) and update the refresh register ($\textit{r\_}$).
-The other instructions have a similar structure.
+$\textit{updateAddressMem}$ operation from \textit{Memory} module, it receives
+an address memory and its new memory value. $\textit{LD\_(nn)\_A}$ also increments the program
+counter ($\textit{pc}$) and updates the refresh register ($\textit{r\_}$). The other instructions have a similar structure.
 
@@
 
-\subsection{Modeling the input/output instructions}
-\label{sec:modelingIO}
-The Z80 has an extensive set of input and output (I/O) instructions and 256 ports for
-devices. This model can transfer data blocks between the I/O devices and any
-of the internal registers or memory address.
-
-The $\textit{IN\_r(C)}$\footnoterecall{myfootnote} instruction is represented
-by the following B operation.
-It receives a register identifier ``$\textit{rr}$'' and, it stores the value from ``rr'' in the $\textit{C}$ port address.
-Besides, it increments the program counter and updates the flag registers.
-
-\hspace*{0.0in}\bf IN\_r\_9C0 \rm ( \it rr \rm ) \rm =
-
-\hspace*{0.0in}\bf PRE \it rr  $\in$  \it id\_reg\_8  $\land$  \it rr  $\not =$  \it f0\hspace*{0.15in}\bf THEN
-
-\hspace*{0.20in}\bf ANY \hspace*{0.10in}\it negative \rm , \it zero \rm , \it half\_carry \rm , \it pv \rm , \it add\_sub \rm , \it carry
-
-\hspace*{0.20in}\bf WHERE
-
-\hspace*{0.40in}\it negative $\in$ \it BIT $\land$ \it zero $\in$ \it BIT $\land$ \it half\_carry $\in$ \it BIT 
-$\land$ \it pv $\in$ \it BIT $\land$
-
- \hspace*{0.40in}\it add\_sub $\in$ \it BIT $\land$ \it carry $\in$ \it BIT  $\land$
-
-\hspace*{0.40in}\it negative \rm = \it is\_negative \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$
-
-\hspace*{0.40in}\it zero \rm = \it is\_zero \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$ \it half\_carry \rm = \rm 0  $\land$
-
-\hspace*{0.40in}\it pv \rm = \it parity\_even \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) $\land$
-\it add\_sub \rm =\hspace*{0.10in}\rm 0  $\land$ 
-\it carry \rm = \it z\_c
-
-\hspace*{0.20in}\bf THEN
-
-\hspace*{0.40in}\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm ( \it rr  $\mapsto$  \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) \rm ,
-
-\hspace*{0.40in}\it update\_flag\_reg\rm (\it negative\rm,\it zero\rm,\it half\_carry\rm,\it pv\rm,\it add\_sub\rm,\it carry)\rm\}$\para$
-
-\hspace*{0.40in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
-
-\hspace*{0.20in}\bf END
-
-\hspace*{0.0in}\bf END \rm
-
-%The main I/O instructions work similarly, for example: $\textit{OUT(n),A}$  or $\textit{OUT (C ), r}$.
-
-% $
-% \begin{array}{l} 
-% \textit{IN\_r\_9C0} ( rr ) = \\
-% \quad   \PRE rr \in id\_reg\_8    \THEN\\
-% \quad\quad \ANY data\_in, negative , zero , half\_carry , pv , add\_sub , carry\\ 
-% \quad\quad \WHERE data\_in \in \textit{BYTE} \land negative \in \textit{BIT}\land\\
-% \quad\quad\quad carry \in \textit{BIT} \land half\_carry \in \textit{BIT} \land zero \in \textit{BIT} \land \\
-% \quad\quad\quad negative = is\_negative (data\_in) \land zero = is\_zero(data\_in ) \land\\
-% \quad\quad\quad half\_carry = 0 \land pv =parity\_even\_BYTE ( data\_in )    \land\\
-% \quad\quad\quad add\_sub =  0 \land carry = z\_c \\
-% \quad\quad  \THEN \\
-% \quad\quad\quad i\_o\_ports ( rgs8 ( c0 ) ) := data\_in ||\\
-% \quad\quad\quad rgs8 := rgs8 <+ \{ ( rr |-> data\_in ) ,\\
-% \quad\quad\quad get\_new\_flag\_register\_SZ\_H\_PvNC ( rgs8 , negative , zero, half\_carry , pv , add\_sub , carry ) \} ||\\
-% \quad\quad\quad pc := instruction\_next( pc )\\
-% \quad\quad \END\\
-% \quad \END\\
-% \end{array}
-% $
-
-\subsection{Modeling the external actions}
-\label{sec:externalactions}
-
+%\subsection{Modeling the input/output instructions}
+%\label{sec:modelingIO}
+%%The Z80 has an extensive set of input and output (I/O) instructions and 256 ports for
+%%devices. This model can transfer data blocks between the I/O devices and any
+%%of the internal registers or memory address. The $\textit{IN\_r(C)}$ instruction is
+%%represented by the following B operation. It receives a register identifier ``$\textit{rr}$''
+%%and, it stores the value from ``rr'' in the $\textit{C}$ port address. Besides,
+%%it increments the program counter and updates the flag registers.
+%%
+%%\hspace*{0.0in}\bf IN\_r\_9C0 \rm ( \it rr \rm ) \rm =
+%%
+%%\hspace*{0.0in}\bf PRE \it rr  $\in$  \it id\_reg\_8  $\land$  \it rr  $\not =$  \it f0\hspace*{0.15in}\bf THEN
+%%
+%%\hspace*{0.20in}\bf ANY \hspace*{0.10in}\it negative \rm , \it zero \rm , \it half\_carry \rm , \it pv \rm , \it add\_sub \rm , \it carry
+%%
+%%\hspace*{0.20in}\bf WHERE 
+%%
+%%\hspace*{0.40in}\it negative $\in$ \it BIT $\land$ \it zero $\in$ \it BIT $\land$ \it half\_carry $\in$ \it BIT 
+%%$\land$ \it pv $\in$ \it BIT $\land$
+%%
+%% \hspace*{0.40in}\it add\_sub $\in$ \it BIT $\land$ \it carry $\in$ \it BIT  $\land$
+%%
+%%\hspace*{0.40in}\it negative \rm = \it is\_negative \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$ 
+%%
+%%\hspace*{0.40in}\it zero \rm = \it is\_zero \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$ \it half\_carry \rm = \rm 0  $\land$ 
+%%
+%%\hspace*{0.40in}\it pv \rm = \it parity\_even \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) $\land$
+%%\it add\_sub \rm =\hspace*{0.10in}\rm 0 % $\land$ \it carry \rm = \it z\_c
+%%
+%%\hspace*{0.20in}\bf THEN
+%%
+%%\hspace*{0.40in}\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm ( \it rr  $\mapsto$  \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) \rm ,
+%%
+%%\hspace*{0.40in}\it update\_flag\_reg\rm (\it negative\rm,\it zero\rm,\it half\_carry\rm,\it pv\rm,\it add\_sub\rm,\it carry)\rm\}$\para$
+%%
+%%\hspace*{0.40in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
+%%
+%%\hspace*{0.20in}\bf END
+%%
+%%\hspace*{0.0in}\bf END \rm
+%%
+%%%The main input and output instructions work similarly, for example: $\textit{OUT(n),A}$  or $\textit{OUT (C ), r}$. 
+%%
+%% $
+%% \begin{array}{l} 
+%% \textit{IN\_r\_9C0} ( rr ) = \\
+%% \quad   \PRE rr \in id\_reg\_8    \THEN\\
+%% \quad\quad \ANY data\_in, negative , zero , half\_carry , pv , add\_sub , carry\\ 
+%% \quad\quad \WHERE data\_in \in \textit{BYTE} \land negative \in \textit{BIT}\land\\
+%% \quad\quad\quad carry \in \textit{BIT} \land half\_carry \in \textit{BIT} \land zero \in \textit{BIT} \land \\
+%% \quad\quad\quad negative = is\_negative (data\_in) \land zero = is\_zero(data\_in ) \land\\
+%% \quad\quad\quad half\_carry = 0 \land pv =parity\_even\_BYTE ( data\_in )    \land\\
+%% \quad\quad\quad add\_sub =  0 \land carry = z\_c \\
+%% \quad\quad  \THEN \\
+%% \quad\quad\quad i\_o\_ports ( rgs8 ( c0 ) ) := data\_in ||\\
+%% \quad\quad\quad rgs8 := rgs8 <+ \{ ( rr |-> data\_in ) ,\\
+%% \quad\quad\quad get\_new\_flag\_register\_SZ\_H\_PvNC ( rgs8 , negative , zero, half\_carry , pv , add\_sub , carry ) \} ||\\
+%% \quad\quad\quad pc := instruction\_next( pc )\\
+%% \quad\quad \END\\
+%% \quad \END\\
+%% \end{array}
+%% $
+%
+%
+%
+%
+%%\subsection{Modeling the external actions}
+%%\label{sec:externalactions}
+%
 The external actions change the state of the microcontroller, for example,
-refreshing the I/O ports and interruption request. The external actions are also
+refreshing the I/O ports and requesting interruptions. The external actions are also
 modeled by operations and they are named with the prefix ``$ext\_$'' and
@@
 devices. When a routine is finished, then the CPU comes back to the last routine
-that was interrupted.
-
-For the interrupts, the following elements are important:  the interrupt flip-flops
-($\textit{iff1}$ and $\textit{iff2}$), the types of interrupts (maskable and
-non-maskable), the interrupt mode (set with the $\textit{IM 0}$, $\textit{IM 1}$,
-$\textit{IM 2}$ instructions) and the $\textit{i\_}$ register.
-
-Flip-flops $\textit{iff1}$ and $\textit{iff2}$ control the maskable interrupts
-($\textit{INT}$). When $\textit{iff1}$ is set, the interrupt is enabled,
-otherwise it is disabled; $\textit{iff2}$ is used only as backup for $\textit{iff1}$. The
-instructions $\textit{EI}$ and $\textit{DI}$ respectively enable and disable the
-maskable interruptions, setting  $\textit{iff1}$ to 1 and 0.
-
-The interruptions and the \textit{reset} action can change the state of program
-counter. Theses actions are modeled by B operations and its main effects are
-presented here\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$
-holds the value of stack pointer minus 2,
-% =$\textit{dec\_BV16(dec\_BV16(sp))}$
- $\textit{sp\_minus\_one}$ is the value of stack pointer minus 1,
-% =$\textit{dec\_BV16(sp)}$
-$\textit{pc\_high}$ holds the most significant 8 bits and $\textit{pc\_low}$
-holds the least significant 8 bits.}.
-
- \textbf{NMI} - Non-maskable interrupts cannot be disabled
- by the programmer. Then, when a device makes a request, $sp$ is pushed,
- $pc$ receives $66H$ (102 in decimal), $\textit{iff1}$ is reset, $\textit{iff2}$ stores
- $\textit{iff1}$ and the refresh register is updated.
-
-\begin{sloppypar}
-\bf updateStack\rm (\rm \{ \rm (\it sp\_minus\_two  $\mapsto$  \it pc\_low\rm )\rm ,\rm (\it sp\_minus\_one  $\mapsto$ \it pc\_high \rm ) \rm \}\rm )$\para$
-
-\it sp \rm := \it sp\_minus\_two  $\para$ \it pc \rm := \rm 1\rm 0\rm 2 $\para$ \it iff1\rm :=\rm 0  $\para$  \it iff2\rm := \it iff1 $\para$
-
-\it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\\
-\end{sloppypar}
-
-  \textbf{INT} - Maskable Interrupt is usually reserved for important functions
-  that can be enabled and disabled by the programmer. When a maskable interrupt
-  action happens, both $\textit{iff1}$ and $\textit{iff2}$ are cleared,
-  disabling the interrupts, $sp$ is pushed, the refresh register is updated and
-  the other effects depend on the interrupt mode register (\it im\rm).
-
- \begin{itemize}
-
-  \item The mode 0 is compatible with 8080 and this mode is selected when \it im
-  \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a  non-maskable interrupt
-  happens, the CPU fetches an instruction of one byte from an external device,
-  usually an RST instruction, and the CPU executes it.
-  The instruction code is received from an external device by data bus and it is
-  represented by integer parameter called $\textit{byte\_bus}$.
-
-  \item The mode 1 is the simplest and this mode is selected when \it im \rm =
-  \rm ( \rm 0  $\mapsto$  \rm 1 \rm ). Simply, when a non-maskable interruption
-  happens, the program counter receives $38H$ (56 in decimal).
-
-  \item The mode 2 is the most flexible and this mode is selected when \it im
-  \rm = \rm ( \rm 1  $\mapsto$ \rm 1 \rm ). When a non-maskable interruption
-  happens, an indirect call can be made to any address memory. The program
-  counter receives a bit vector of size 16 composed with two parts: the most
-  significant part of the $\textit{i\_}$ register and the least significant part
-  of the $\textit{byte\_bus}$ with the last bit cleared.
-
- \end{itemize}
-
-The essential part of maskable interrupt is shown below, where
-$\textit{byte\_bus}$ is a parameter of the $\textit{INT}$ operation:
-
-\begin{sloppypar}
-
-\hspace*{0.00in}\bf IF \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 0 \rm ) \bf THEN
-
-\hspace*{0.10in}\bf IF\hspace*{0.10in}\it byte\_bus \rm $\in$  \it opcodes\_RST\_instruction
-
-\hspace*{0.10in}\bf THEN\hspace*{1.05in}
-
-\hspace*{0.40in}\it pc \rm := \it byte\_bus \rm - \rm 1\rm 9\rm
-9\hspace*{0.10in} $\para$
-
-\hspace*{0.40in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
-sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
-
-\hspace*{0.40in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
-\rm \} \rm )  $\para$
-
-\hspace*{0.40in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
-r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
-
-\hspace*{0.10in}\bf ELSIF \it byte\_bus \rm = \it opcode\_\ldots\_instruction
-
-\hspace*{0.40in}\bf \ldots
-
-\hspace*{0.10in}\bf END
-
-\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm =\hspace*{0.10in}\rm ( \rm
-0  $\mapsto$  \rm 1 \rm ) \bf THEN
-
-\hspace*{0.10in}\it pc \rm :=\hspace*{0.10in}\rm 5\rm 6  \hspace*{0.80in}
-$\para$
-
-\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm (\it
-sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
-
-\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
-\rm \} \rm )  $\para$
-
-\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
-r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\hspace*{0.85in}
-
-\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm = \rm ( \rm 1  $\mapsto$ 
-\rm 1 \rm ) \bf THEN  \hspace*{0.70in}
-
-\hspace*{0.10in}\it pc \rm := \it bv16\_ushort\rm (\it byte\_bv16\rm ( \it i\_
-\rm ,\it bv\_clear\rm (\it rotateleft\rm (\it uchar\_byte\rm (\it byte\_bus\rm )\rm )\rm ,\rm 0\rm )\rm )\rm )$\para$
-
-\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
-sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
-
-\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
-\rm \} \rm )  $\para$
-
-\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
-r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
-
-\hspace*{0.00in}\bf END\hspace*{0.10in}\\
-\end{sloppypar}
-
-\textbf{RESET}  - This just resets the registers related to the interruptions.
-
-\begin{sloppypar}
-\it iff1 \rm :=\rm 0 $\para$ \it iff2\rm :=\rm 0 $\para$  \it  im\rm := \rm (\rm 0 $\mapsto$ \rm 0\rm )  $\para$ 
-\it pc\rm :=\rm 0 $\para$ \it i\_ \rm := [0,0,0,0,0,0,0,0] $\para$
-
-\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm (\it a0  $\mapsto$ \rm [0,0,0,0,0,0,0,0] , \rm (\it
-f0 $\mapsto$ \rm [0,0,0,0,0,0,0,0] \rm \} $\para$
-
-\it r\_  \rm := [0,0,0,0,0,0,0,0] $\para$ \it sp \rm := [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0]\\
-\end{sloppypar}
+that was interrupted. Modeling details about interrupts can be consulted in
+public project's repository.
+
+%
+%For the interrupts, the following elements are important:  the interrupt flip-flops
+%($\textit{iff1}$ and $\textit{iff2}$), the types of interrupts (maskable and
+%non-maskable), the interrupt mode (set with the $\textit{IM 0}$, $\textit{IM 1}$,
+%$\textit{IM 2}$ instructions) and the $\textit{i\_}$ register.
+%
+%Flip-flops $\textit{iff1}$ and $\textit{iff2}$ control the maskable interrupts
+%($\textit{INT}$). When $\textit{iff1}$ is set, the interrupt is enabled,
+%otherwise it is disabled; $\textit{iff2}$ is used only as backup for $\textit{iff1}$. The
+%instructions $\textit{EI}$ and $\textit{DI}$ respectively enable and disable the
+%maskable interruptions, setting  $\textit{iff1}$ to 1 and 0.
+%
+%
+%The interruptions and the \textit{reset} action can change the state of program
+%counter. Theses actions are modeled by B operations and its main effects are
+%presented here\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$
+%holds the value of stack pointer minus 2,
+%% =$\textit{dec\_BV16(dec\_BV16(sp))}$
+% $\textit{sp\_minus\_one}$ is the value of stack pointer minus 1,
+%% =$\textit{dec\_BV16(sp)}$
+%$\textit{pc\_high}$ holds the most significant 8 bits and $\textit{pc\_low}$
+%holds the least significant 8 bits.}.
+%
+%
+% \textbf{NMI} - Non-maskable interrupts cannot be disabled
+% by the programmer. Then, when a device makes a request, $sp$ is pushed,
+% $pc$ receives $66H$ (102 in decimal), $\textit{iff1}$ is reset, $\textit{iff2}$ stores
+% $\textit{iff1}$ and the refresh register is updated.
+%  
+%\begin{sloppypar}
+%\bf updateStack\rm (\rm \{ \rm (\it sp\_minus\_two  $\mapsto$  \it pc\_low\rm )\rm ,\rm (\it sp\_minus\_one  $\mapsto$ \it pc\_high \rm ) \rm \}\rm )$\para$
+%
+%\it sp \rm := \it sp\_minus\_two  $\para$ \it pc \rm := \rm 1\rm 0\rm 2 $\para$ \it iff1\rm :=\rm 0  $\para$  \it iff2\rm := \it iff1 $\para$
+%
+%\it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\\
+%\end{sloppypar}
+%
+%  \textbf{INT} - Maskable Interrupt is usually reserved for important functions
+%  that can be enabled and disabled by the programmer. When a maskable interrupt
+%  action happens, both $\textit{iff1}$ and $\textit{iff2}$ are cleared,
+%  disabling the interrupts, $sp$ is pushed, the refresh register is updated and
+%  the other effects depend on the interrupt mode register (\it im\rm).
+% 
+%
+% \begin{itemize}
+%   
+%  \item The mode 0 is compatible with 8080 and this mode is selected when \it im
+%  \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a  non-maskable interrupt
+%  happens, the CPU fetches an instruction of one byte from an external device,
+%  usually an RST instruction, and the CPU executes it.
+%  The instruction code is received from an external device by data bus and it is
+%  represented by integer parameter called $\textit{byte\_bus}$.
+%
+% 
+%  \item The mode 1 is the simplest and this mode is selected when \it im \rm =
+%  \rm ( \rm 0  $\mapsto$  \rm 1 \rm ). Simply, when a non-maskable interruption
+%  happens, the program counter receives $38H$ (56 in decimal).
+%  
+%  \item The mode 2 is the most flexible and this mode is selected when \it im
+%  \rm = \rm ( \rm 1  $\mapsto$ \rm 1 \rm ). When a non-maskable interruption
+%  happens, an indirect call can be made to any address memory. The program
+%  counter receives a bit vector of size 16 composed with two parts: the most
+%  significant part of the $\textit{i\_}$ register and the least significant part
+%  of the $\textit{byte\_bus}$ with the last bit cleared.
+%  
+% \end{itemize}
+%
+%The essential part of maskable interrupt is shown below, where
+%$\textit{byte\_bus}$ is a parameter of the $\textit{INT}$ operation:
+% 
+%
+%
+%\begin{sloppypar}
+%
+%\hspace*{0.00in}\bf IF \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 0 \rm ) \bf THEN  
+%
+%\hspace*{0.10in}\bf IF\hspace*{0.10in}\it byte\_bus \rm $\in$  \it opcodes\_RST\_instruction
+%
+%\hspace*{0.10in}\bf THEN\hspace*{1.05in}
+%
+%\hspace*{0.40in}\it pc \rm := \it byte\_bus \rm - \rm 1\rm 9\rm
+%9\hspace*{0.10in} $\para$
+%
+%\hspace*{0.40in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
+%sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
+%
+%\hspace*{0.40in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
+%\rm \} \rm )  $\para$
+%
+%\hspace*{0.40in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
+%r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
+%
+%\hspace*{0.10in}\bf ELSIF \it byte\_bus \rm = \it opcode\_\ldots\_instruction
+%
+%\hspace*{0.40in}\bf \ldots 
+%
+%\hspace*{0.10in}\bf END
+%
+%\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm =\hspace*{0.10in}\rm ( \rm
+%0  $\mapsto$  \rm 1 \rm ) \bf THEN
+%
+%\hspace*{0.10in}\it pc \rm :=\hspace*{0.10in}\rm 5\rm 6  \hspace*{0.80in}
+%$\para$
+%
+%\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm (\it
+%sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
+%
+%\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
+%\rm \} \rm )  $\para$
+%
+%\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
+%r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\hspace*{0.85in}
+%
+%\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm = \rm ( \rm 1  $\mapsto$ 
+%\rm 1 \rm ) \bf THEN  \hspace*{0.70in}
+%
+%\hspace*{0.10in}\it pc \rm := \it bv16\_ushort\rm (\it byte\_bv16\rm ( \it i\_
+%\rm ,\it bv\_clear\rm (\it rotateleft\rm (\it uchar\_byte\rm (\it byte\_bus\rm )\rm )\rm ,\rm 0\rm )\rm )\rm )$\para$
+%
+%\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
+%sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,
+%
+%\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
+%\rm \} \rm )  $\para$
+%
+%\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
+%r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )
+%
+%\hspace*{0.00in}\bf END\hspace*{0.10in}\\
+%\end{sloppypar}
+%
+%
+%\textbf{RESET}  - This just resets the registers related to the interruptions.
+%
+%\begin{sloppypar}
+%\it iff1 \rm :=\rm 0 $\para$ \it iff2\rm :=\rm 0 $\para$  \it  im\rm := \rm (\rm 0 $\mapsto$ \rm 0\rm )  $\para$ 
+%\it pc\rm :=\rm 0 $\para$ \it i\_ \rm := [0,0,0,0,0,0,0,0] $\para$
+%
+%\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm (\it a0  $\mapsto$ \rm [0,0,0,0,0,0,0,0] , \rm (\it
+%f0 $\mapsto$ \rm [0,0,0,0,0,0,0,0] \rm \} $\para$
+%
+%\it r\_  \rm := [0,0,0,0,0,0,0,0] $\para$ \it sp \rm := [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0]\\
+%\end{sloppypar}
 
@@
 \hline 
-{\small \textbf{Component}}&{\small \textbf{PO}}&{\small \textbf{UPO}}&{\small \textbf{WD PO}}&{\small \textbf{UWD PO }}&{\small \textbf{Total PO}}&{\small \textbf{Total UPO}}& {\small \textbf{Total Percentage}}  \\\hline
+{\scriptsize \textbf{Component}}&{\scriptsize \textbf{PO}}&{\scriptsize \textbf{UPO}}&{\scriptsize \textbf{WD PO}}&{\scriptsize \textbf{UWD PO }}&{\scriptsize \textbf{Total PO}}&{\scriptsize \textbf{Total UPO}}& {\scriptsize \textbf{Total \%}}  \\\hline
 BIT&49&0&69&0&118&0&100\%\\\hline
@@
 instruction set designed without restricting ourselves to constructs
-that may be animated with ProB, and without modeling external actions
-(as described in section~\ref{sec:externalactions}). In that first
-experience, the verification took us about two months. In that second
+that may be animated with ProB, and without modeling external actions.
+In that first experience, the verification took us about two months. In that second
 project, we have been able to verify 3062 (e.g. roughly 80 \%) of 3631
@@
 
-\subsubsection{Acknowledgements:} The animation of the model would not
-have been possible without the help of Michael Leuschel who kindly
-provided feedback and developed several improvements to ProB to meet our
-needs.
+%\subsubsection{Acknowledgements:} The animation of the model would not
+%have been possible without the help of Michael Leuschel who kindly
+%provided feedback and developed several improvements to ProB to meet our
+%needs.
 

### 2012-05-20 trunk/papers/ArticleZ80/source/paper.tex
Mais proximo das 16 paginas.
Com 16 16 paginas, preciso de alguns ajustes na identacao das funcoes.

@@
 %\usepackage{supertabular} % used in B Symbols
-%\geometry{a4paper} % or letterpaper (US) or a5paper or....
 
@@
 The new model supports animation in ProB~\cite{proB}.  Essentially, the
-new model \footnote{The interested reader in more details is invited to visit our repository at: http://code.google.com/p/b2asm/.}
+new model\footnote{The interested reader in more details is invited to visit our repository at: http://code.google.com/p/b2asm/.}
 no longer contains definitions derived from infinite sets. Using B
@@
 \centering
-\includegraphics[width=.77\textwidth]{diagramaEstrutural_vertical_ProB.png}
+\includegraphics[width=.85\textwidth]{diagramaEstrutural_vertical_ProB2.png}
  \caption{Dependency diagram of the Z80 model.}
@@
 \textit{0..1}$. The negation is an unary function on bits and it is
-defined as:
+defined below. The module also provides lemmas on negation that may be useful for the
+users of the library to develop proofs.
 $$
@@
 \textit{bit\_not}  \in  \textit{BIT}  \fun  \textit{BIT}  \land 
-\textit{bit\_not} \rm = \rm \{ \rm 0  \mapsto  \rm 1 , \rm 1  \mapsto  \rm 0 \rm \}
-\end{array}
-$$
-%\forall ( \textit{bb}). (\textit{bb} \in \textit{BIT} \implies {bit\_not}(\textit{bb}) =1-\textit{bb})\\
-
-The module also provides lemmas on negation that may be useful for the
-users of the library to develop proofs:
-$$
-\begin{array}{l}
-%  \textit{bit\_not}(0) = 1;  \textit{bit\_not}(1) = 0; \\
+\textit{bit\_not} \rm = \rm \{ \rm 0  \mapsto  \rm 1 , \rm 1  \mapsto  \rm 0 \rm \} \land \\
 \forall (\textit{bb}).(\textit{bb} \in \textit{BIT} \implies \textit{bit\_not}(\textit{bit\_not}(\textit{bb})) = \textit{bb})
@@
 $$
-The module provides the following lemmas for commutativity and associativity:
-$$
-\begin{array}{l}
-%  \textit{bit\_and}(0,0) = 0;  \textit{bit\_and}(0,1) = 0; \\
-%  \textit{bit\_and}(1,0) = 0;  \textit{bit\_and}(1,1) = 1; \\
-\forall (\textit{b1},\textit{b2}).(\textit{b1} \in \textit{BIT} \land \textit{b2} \in \textit{BIT} \implies \\
-\quad (\textit{bit\_and}(\textit{b1}, \textit{b2}) = \textit{bit\_and}(\textit{b2},\textit{b1})))\land \\
-\forall (\textit{b1},\textit{b2},\textit{b3}).(\textit{b1} \in \textit{BIT} \land  \textit{b2} \in \textit{BIT} \land \textit{b3} \in \textit{BIT} \implies \\
-\quad (\textit{bit\_and}(\textit{b1}, \textit{bit\_and}(\textit{b2},\textit{b3})) = \textit{bit\_and}(\textit{bit\_and}(\textit{b1},\textit{b2}),\textit{b3})))\\
-% \forall (\textit{b1}).(\textit{b1} \in \textit{BIT} \implies (\textit{bit\_and}(\textit{b1}, 1) = \textit{b1})); \\
-% \forall (\textit{b1}).(\textit{b1} \in \textit{BIT} \implies (\textit{bit\_and}(\textit{b1}, 0) = 0));
-\end{array}
-$$
-The module provides definitions of $\textit{bit\_or}$ (disjunction)
+%The module provides the following lemmas for commutativity and associativity:
+%$$
+%\begin{array}{l}
+%%  \textit{bit\_and}(0,0) = 0;  \textit{bit\_and}(0,1) = 0; \\
+%%  \textit{bit\_and}(1,0) = 0;  \textit{bit\_and}(1,1) = 1; \\
+%\forall (\textit{b1},\textit{b2}).(\textit{b1} \in \textit{BIT} \land \textit{b2} \in \textit{BIT} \implies \\
+%\quad (\textit{bit\_and}(\textit{b1}, \textit{b2}) = \textit{bit\_and}(\textit{b2},\textit{b1})))\land \\
+%\forall (\textit{b1},\textit{b2},\textit{b3}).(\textit{b1} \in \textit{BIT} \land  \textit{b2} \in \textit{BIT} \land \textit{b3} \in \textit{BIT} \implies \\
+%\quad (\textit{bit\_and}(\textit{b1}, \textit{bit\_and}(\textit{b2},\textit{b3})) = \textit{bit\_and}(\textit{bit\_and}(\textit{b1},\textit{b2}),\textit{b3})))\\
+%% \forall (\textit{b1}).(\textit{b1} \in \textit{BIT} \implies (\textit{bit\_and}(\textit{b1}, 1) = \textit{b1})); \\
+%% \forall (\textit{b1}).(\textit{b1} \in \textit{BIT} \implies (\textit{bit\_and}(\textit{b1}, 0) = 0));
+%\end{array}
+%$$
+The module provides lemmas for commutativity and associativity, definitions of $\textit{bit\_or}$ (disjunction)
 and $\textit{bit\_xor}$ (exclusive disjunction), as well as lemmas on
@@
 
+%
 % $
@@
 % $
-
-\hspace*{0.00in} \it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$ \it
+\begin{flushleft}
+\hspace*{0.0in}\it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$ \it
 BIT\_VECTOR $\land$\\
-\hspace*{0.21in} \it bv\_catenate \rm =  $\lambda$  v\rm ,\it w \rm . \rm (\it v 
+\it bv\_catenate \rm =  $\lambda$  v\rm ,\it w \rm . \rm (\it v 
 $\in$  \it BIT\_VECTOR $\land$ \it w $\in$  \it BIT\_VECTOR  $\mid$  \it v\^ \it w\rm ) %$\cat$
+\\
+
+\end{flushleft}
 
@@
 is restricted to pairs of bit vectors of the same length:
+
+\begin{flushleft}
 $$
@@
 $$
+\end{flushleft}
+
 We provide several lemmas on bit vector operations. These lemmas
@@
 \begin{figure}[h] \centering
-\includegraphics[width=0.70\textwidth]{images/Architecture.png}
+\includegraphics[width=0.67\textwidth]{images/Architecture_compact.png}
 \caption{Simplified internal organization of Z80 CPU.}
@@
 	    \end{array}
-	     \hspace*{1cm} 
+	     \hspace*{0.2cm} 
 	    \begin{array}[t]{l}
@@
 devices. When a routine is finished, then the CPU comes back to the last routine
-that was interrupted. Modeling details about interrupts can be consulted in
+that was interrupted. Modeling details about interruptions can be consulted in
 public project's repository.
@@
 by stepwise refinement and formal proof using Event-B. It also
-discusses desirable properties of the instruction set model. However,
+discusses desirable properties of an instruction set model. However,
 our work \cite{Valerio_SBMF09} and \cite{Subotic2010} use the B
@@
 \end{document}
-
-%
-% \bibitem {clar:eke}
-% Clarke, F., Ekeland, I.:
-% Nonlinear oscillations and
-% boundary-value problems for Hamiltonian systems.
-% Arch. Rat. Mech. Anal. {\bf 78} (1982) 315--333
-% %
-% \bibitem {clar:eke:2}
-% Clarke, F., Ekeland, I.:
-% Solutions p\'{e}riodiques, du
-% p\'{e}riode donn\'{e}e, des \'{e}quations hamiltoniennes.
-% Note CRAS Paris {\bf 287} (1978) 1013--1015
-% %
-% \bibitem {mich:tar}
-% Michalek, R., Tarantello, G.:
-% Subharmonic solutions with prescribed minimal
-% period for nonautonomous Hamiltonian systems.
-% J. Diff. Eq. {\bf 72} (1988) 28--55
-% %
-% \bibitem {tar}
-% Tarantello, G.:
-% Subharmonic solutions for Hamiltonian
-% systems via a $\bbbz_{p}$ pseudoindex theory.
-% Annali di Matematica Pura (to appear)
-% %
-% \bibitem {rab}
-% Rabinowitz, P.:
-% On subharmonic solutions of a Hamiltonian system.
-% Comm. Pure Appl. Math. {\bf 33} (1980) 609--633
-% \end{thebibliography}
-%

### 2012-05-20 trunk/papers/ArticleZ80/source/paper.tex
Versao praticamente final.

@@
 users of the library to develop proofs.
+
 $$
@@
 that, given a bit vector and a valid position, returns the value of
-the bit at that position. Only the first definition is shown here:
-$$
-\begin{array}{l}
-\textit{bv\_set} \in \textit{BIT\_VECTOR} \times \nat \fun \textit{BIT\_VECTOR} \land \\ \textit{bv\_set} =
-\lambda v, n . (v \in \textit{BIT\_VECTOR} \land n \in \nat \land n <\textit{size}(v)
-\mid v \lover \{ n+1 \mapsto 1 \})
+the bit at that position. Only the first definition is shown here:\newline
+%
+%\vspace*{4mm}
+%\hspace*{-0.2in}\textit{bv\_set} $\in$ \textit{BIT\_VECTOR} $\times \nat \fun$ \textit{BIT\_VECTOR} $\land$\\
+%\it bv\_set \rm =  $\lambda$ \rm \it v\rm ,\it n\rm \rm .\rm (\it v $\in$  \it BIT\_VECTOR  $\land$  \it n  $\in$  \rm 0$\upto$\rm (\it size\rm (\it v\rm )\rm -\rm 1\rm )  $\mid$  \it v  $\lover$  \rm \{ \rm (\it n\rm +\rm 1\rm )  $\mapsto$  \rm 1 \rm \}\rm )
+%\vspace*{4mm}
+%
+%
+% \nat \land n <\textit{size}(v)
+$$
+\hspace*{0.1in}\begin{array}{l}
+\textit{bv\_set} \in \textit{BIT\_VECTOR} \times \nat \fun \textit{BIT\_VECTOR} \land \\
+ \textit{bv\_set} = \lambda v, n . (v \in \textit{BIT\_VECTOR} \land n \in   0..\textit{size}(v)-1
+  \mid v \lover \{n+1 \mapsto 1\})
 \end{array}
@@
 $w$, and returns the concatenation of $v$ and $w$, such
-that $v$ constitutes the most significant part of the result.
-
-%
-% $
-% \begin{array}{l}
+that $v$ constitutes the most significant part of the result. \newline
+
+%$$
+%\hspace*{-0.2in}\begin{array}{l}
 % \textit{bv\_catenate} \in \textit{BIT\_VECTOR} \times \textit{BIT\_VECTOR} \fun \textit{BIT\_VECTOR} \land \\
@@
 %   w)
-% \end{array}
-% $
-\begin{flushleft}
-\hspace*{0.0in}\it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$ \it
+%\end{array}
+%$$
+
+\hspace*{-0.0in}\it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$ \it
 BIT\_VECTOR $\land$\\
-\it bv\_catenate \rm =  $\lambda$  v\rm ,\it w \rm . \rm (\it v 
+\hspace*{0.2in}\it bv\_catenate \rm =  $\lambda$  v\rm ,\it w \rm . \rm (\it v 
 $\in$  \it BIT\_VECTOR $\land$ \it w $\in$  \it BIT\_VECTOR  $\mid$  \it v\^ \it w\rm ) %$\cat$
-\\
-
-\end{flushleft}
+\newline
 
@@
 the first two are presented here. The domain of the binary operators
-is restricted to pairs of bit vectors of the same length:
-
-\begin{flushleft}
-$$
-\begin{array}{l}
+is restricted to pairs of bit vectors of the same length:\newline
+$$
+\hspace*{-0.2in}\begin{array}{l}
 \textit{bv\_not} \in \textit{BIT\_VECTOR} \fun \textit{BIT\_VECTOR} \land \\
@@
 $$
-\end{flushleft}
 
@@
 interruptions actions. All 158 instructions and interruptions were modeled, but only
-two important examples are presented below.
+two important examples are presented here.
 
@@
 
-\hspace*{0.20in}\bf ext\_update\_io\_ports\rm (\it address\rm ,\it value\rm )\rm =
+\hspace*{0.00in}\bf ext\_update\_io\_ports\rm (\it address\rm ,\it value\rm )\rm =
 
@@
 
-\hspace*{0.40in}\it io\_ports \rm ( \it uchar\_byte \rm ( \it address \rm ) \rm ) \rm := \it schar\_byte \rm ( \it
+\hspace*{0.20in}\it io\_ports \rm ( \it uchar\_byte \rm ( \it address \rm ) \rm ) \rm := \it schar\_byte \rm ( \it
 value \rm )
 
-\hspace*{0.20in}\bf END\rm
+\hspace*{0.00in}\bf END\rm
 

### 2012-05-20 trunk/papers/ArticleZ80/source/paper.tex
Versao submetida.

@@
 \hyphenation{re-gis-ters}
+%\hyphenation{ge-ne-ra-ted}
 
@@
   specification of the functional requirements of a system to an assembly-level
-  implementation thereof.
+  implementation thereof.\newline %\keywords{Embedded Software, Simulation, B method and Verification}
 % Using this model-driven approach and the instruction set model, a small case
 % study to analyse a petroleum production test system in oil field was developed.
-%  \paragraph[Keywords]{ \textbf{Keywords}: Embedded Software, Simulation, B
-% method and Verification%}
+ { \textbf{Key words:} B method, Simulation, Verification and Z80}
+
 \end{abstract}
@@
 
-The new model supports animation in ProB~\cite{proB}.  Essentially, the
-new model\footnote{The interested reader in more details is invited to visit our repository at: http://code.google.com/p/b2asm/.}
-no longer contains definitions derived from infinite sets. Using B
-constructs for modularity, auxiliary libraries of basic modules were
-developed as part of the construction of the microcontroller model. Such
-a library has many definitions of concepts common to microcontrollers.
-Besides the new Z80 model, a petroleum production test system was
-developed using the Z80 assembly language to analyse the code simulation
-and the verification process.
+The new Z80 model supports animation in ProB~\cite{proB}.  Essentially, the
+new model\footnote{The interested reader in more details is invited to visit
+our repository at: http://code.google.com/p/b2asm/.} no longer contains
+definitions derived from infinite sets. Using B constructs for modularity,
+auxiliary libraries of basic modules were developed as part of the construction
+of the microcontroller model. Such a library has many definitions of 
+concepts common to microcontrollers. Besides the new Z80 model, a petroleum
+production test system was developed using the Z80 assembly language to
+analyse the code simulation and the verification process.
 
@@
 
-%TODO:[NEW PARAGRAPH-TO-REVIEW]
 Similarly, several other general functions and lemmas were created for all other data types.
-Some times, we need  create arithmetic and logic functions that are more specific for a
+Some times, we need to create arithmetic and logic functions that are more specific for a
 determined microprocessor, like Z80 that is presented in the section \ref{sec:z80}.
 
-%TODO:[NEW PARAGRAPH-TO-REVIEW]
-The definitions, functions and properties from hardware library and types library are difficult to verify,
-on the other hand  these libraries can be reused in several others  microprocessors without to verify again.
-Nearly 50\% of all these proof obligations shown in the Table~\ref{tab:types} are verified automatically 
-or with few user pass\footnote{The user pass are steps that aid the prover to find the proof.} .
-But many others proof obligations are difficult to verify, because they involve functions
-that contain arithmetic expressions that are complex to manipulate with the interactive
-prover of AtelierB.  However, all theses 531 proof obligations were verified completely.
+The definitions, functions and properties from hardware library and types library are difficult to verify, otherwise these libraries can be reused in several others microprocessors without to verify again.
+Nearly 50\% of all these proof obligations shown in the Table~\ref{tab:types} are verified automatically or with few user pass\footnote{The user pass allows to filter proof tactics that aid the prover to find the proof, this is a feature provided by Atelier B.}.
+But many others proof obligations are difficult to verify because they involve functions that contain complex arithmetic expressions to manipulate with the interactive prover of AtelierB.
+However, all these 531 proof obligations were verified completely.
 
@@
 \centering
-\includegraphics[width=3.in]{images/Diagram_Types_and_Rules.pdf}
+\includegraphics[width=2.8in]{images/Diagram_Types_and_Rules.pdf}
 \caption{Conversion functions}
@@
 
-There is a specific function to refresh the flag register, it is named $\textit{update\_reg\_flag}$.
+There is a specific function to refresh the flag register, it is named $\textit{update\_reg}$ $\textit{\_flag}$.
 It is typed and defined as follow:\it update\_flag\_reg \rm $\in$ \rm (\it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it
@@
 interruptions actions. All 158 instructions and interruptions were modeled, but only
-two important examples are presented here.
+two important operations are presented here.
 
 A simple example of instruction is $\textit{LD\_(nn)\_A}$\footnote{The B tools does not allow o use parentheses in identifiers, and the characters \{``('',``)''\} are
-replaced respectively by \{``9'',``0''\} in the actual specification.} shown below. The
+replaced respectively by \{``9'',``0''\} in the actual specification.} shown in this section. The
 pre-defined functions are necessary many times to model the instructions, these
@@
 more details as well as a small software that developed up to assembly
-level in three different platforms. A general view of a previous used
-in verification process were presented in~\cite{Valerio_SBMF09}. The
-model presented in the current paper added support for animation in
-ProB and the representation of external actions (interruptions).
+level in three different platforms. A short general view of a previous work
+were presented in~\cite{Valerio_SBMF09}. The model presented in the 
+current paper added support for animation in ProB and the representation
+of external actions (interruptions).
 %Link com o parágrafo anterior 
-This set of papers present an important step to software verification
+This set of papers present an alternative step to software verification
 up to assembly language, showing the current difficulties and

### 2012-05-21 trunk/papers/ArticleZ80/source/paper.tex

@@
 partial, Z80 model presented in \cite{Valerio_SBMF09} was fully
-verified, but it used constructs that made it impossible to animate.
+verified, but it used constructs that made it impossible to animate 
+with existing technology. %%DD
 
@@
 \item[Power] contains the standard definition of exponentiation, as
-  well as some basic properties, e.g. values of different powers of
-  two. Given this definition, these properties have been proved
-  interactively with the Atelier-B prover.
+  well as some basic lemmas, e.g. values of different powers of
+  two, allproved interactively with the Atelier-B prover.
+  This module is essential to establish the relationship between
+  bit vectors and integer arithmetics. %%DD
 
 \item[Hardware Library] defines bit vectors with size 8 and 16, basic
-  functions to manipulate bit vectors and important
-  assertions. Details are given in sections
-  \ref{subsec:HardwareLibrary1}, \ref{subsec:HardwareLibrary2} and
-  \ref{subsec:HardwareLibrary3}.
+  functions to manipulate bit vectors and important lemmas. Details
+  are given in sections \ref{subsec:HardwareLibrary1},
+  \ref{subsec:HardwareLibrary2} and \ref{subsec:HardwareLibrary3}.
 
@@
 We model bits as a set of integers: $\textit{BIT} =
-\textit{0..1}$. The negation is an unary function on bits and it is
-defined below. The module also provides lemmas on negation that may be useful for the
-users of the library to develop proofs.
+\textit{0..1}$. The negation is an unary function on bits; its definition
+follows. The module also provides lemmas on negation that may be
+useful for the users of the library to develop proofs.
 
@@
 
-Similarly, several other general functions and lemmas were created for all other data types.
-Some times, we need to create arithmetic and logic functions that are more specific for a
-determined microprocessor, like Z80 that is presented in the section \ref{sec:z80}.
-
-The definitions, functions and properties from hardware library and types library are difficult to verify, otherwise these libraries can be reused in several others microprocessors without to verify again.
-Nearly 50\% of all these proof obligations shown in the Table~\ref{tab:types} are verified automatically or with few user pass\footnote{The user pass allows to filter proof tactics that aid the prover to find the proof, this is a feature provided by Atelier B.}.
-But many others proof obligations are difficult to verify because they involve functions that contain complex arithmetic expressions to manipulate with the interactive prover of AtelierB.
-However, all these 531 proof obligations were verified completely.
+Similarly, several other general functions and lemmas were created for
+all other data types.  We sometimes need to create arithmetic and
+logic functions that are more specific for a determined
+architecture, as the Z80 presented in section \ref{sec:z80}.
+
+%%DD As primeiras frases do parágrafo abaixo não faziam sentido!
+The definitions, functions and properties from hardware library and
+types library may be reused to model different instruction set
+architectures. Nearly 50\% of all these proof obligations shown in
+Table~\ref{tab:types} are verified automatically or with few user
+pass\footnote{The user pass allows to filter proof tactics that aid
+  the prover to find the proof, this is a feature provided by Atelier
+  B.}. A large part of the proof obligations require long sessions
+with the interactive provers, mostly because they contain complex
+arithmetic expressions that the automatic provers in AtelierB have
+difficulty handling. However, all these 531 proof obligations were
+eventually verified.
 
@@
 
+%DD forneci mais detalhes no paragrafo seguinte
 To avoid such inconsistencies, we state separately the required
-properties in the \textit{ASSERTIONS\/} clause: this causes
-corresponding proof obligations to be generated and verified. However,
-we have not been able to build the corresponding proofs using the 
-rule base included in the free distribution of AtelierB. So we
-formalized new proof rules to check these properties in AtelierB and
-used them in the interactive prover of AtelierB.
+properties in the \textit{ASSERTIONS\/} clause: this changes the
+status of these properties from axioms to lemmas and causes
+corresponding proof obligations to be generated. However, we have not
+been able to build the corresponding proofs using the rule base
+included in the free distribution of AtelierB. In that situation, the
+used to common approach to define new proof rules to check these
+properties in AtelierB and to use them in the interactive prover of
+AtelierB.
 
@@
 
-The Z80 has 158 different instructions, including all the 78 from the Intel 8080
-microprocessor, and all of them were formally specified in B. These instructions are classified
-into these categories: load and exchange; block transfer and search; arithmetic
-and logical; rotate and shift; bit manipulation; jump, call and return;
-input/output; and basic cpu control. Each category of instruction has different
-elements of specification.
+The Z80 has 158 different instructions, including all the 78 from the
+Intel 8080 microprocessor, and all of them were formally specified in
+B. These instructions are classified into these categories: load and
+exchange; block transfer and search; arithmetic and logical; rotate
+and shift; bit manipulation; jump, call and return; input/output; and
+basic cpu control. Each category of instruction has different elements
+of specification.
 
 This section shows the elements that make up the different types of
-instructions. The main elements are specified in the microcontroller module Z80
-and parts of it are presented below.  Basically, the state of microcontroller is
-formed by data from program counter, ports, registers, etc. The transitions
-between the states are actioned by execution of instruction or an external
-action. Groups of registers are represented by the variables of the
-specification states that appear in clause \textit{VARIABLES}. The declaration
-of valid states of variables is in the \textit{INVARIANT} clause and the initial
-state is defined in the \textit{INITIALISATION} clause. The assembly
-instructions are defined by the clause \textit{OPERATIONS} and, in general, each
-instruction has three actions: update the program counter, update the flag
+instructions. The main elements are specified in the microcontroller
+module Z80 and parts of it are presented below.  Basically, the
+microcontroller state is formed by the program counter, ports,
+registers, etc. The stae transition occurs either as an instruction
+execution or an external action. Groups of registers are represented
+by the variables of the specification states that appear in clause
+\textit{VARIABLES}. The declaration of valid states of variables is in
+the \textit{INVARIANT} clause and the initial state is defined in the
+\textit{INITIALISATION} clause. The assembly instructions are defined
+by the clause \textit{OPERATIONS} and, in general, each instruction
+has three actions: update the program counter, update the flag
 register and its main effect.
 % Several functions were created to Z80.
-General functions that can be used in other microcontroller models are defined
-in the modules of data types. Specific functions of the microcontroller are
-defined in the ALU (arithmetic logic unit) module.The main module includes an instance of the memory module and accesses the
-definitions from basic data types modules and the \textit{ALU} module.
+The functions specific to the microcontroller are defined in the ALU
+(arithmetic logic unit) module. The main module includes an instance of
+the memory module and accesses the definitions from basic data types
+modules and the \textit{ALU} module.
 
@@
 
-The Z80 CPU includes alternative set of accumulator, flag and general registers.
-The CPU contains a stack pointer ($\textit{sp}$), program counter
-($\textit{pc}$), two index registers ($\textit{ix}$ and $\textit{iy}$), an
-interrupt register ($\textit{i\_}$), a refresh register ($\textit{r\_}$), two
-bits ($\textit{iff1}$, $\textit{iff2}$) used to control the interruptions, a
-pair of bits to define the interruption mode ($\textit{im}$) and the input and
-output ports ($\textit{i\_o\_ports}$). These definitions are represented by
+%%DD removi uma frase que não fazia sentido gramaticalmente.
+The Z80 CPU contains a stack pointer ($\textit{sp}$), program counter
+($\textit{pc}$), two index registers ($\textit{ix}$ and
+$\textit{iy}$), an interrupt register ($\textit{i\_}$), a refresh
+register ($\textit{r\_}$), two bits ($\textit{iff1}$, $\textit{iff2}$)
+used to control the interruptions, a pair of bits to define the
+interruption mode ($\textit{im}$) and the input and output ports
+($\textit{i\_o\_ports}$). These definitions are represented by
 \textit{INVARIANT}.
@@
 The internal registers contains 176 bits of read/write memory that are
-represented by identifiers used as parameters in the instructions . It includes
-two sets of six general purpose registers which may be used individually as
-8-bit registers or as 16-bit register pairs.  The working registers are
-represented by variable $\textit{rgs8}$. The domain of $\textit{rgs8}$
-($\textit{id\_reg8}$) is a set formed by identifiers of registers of 8 bits.
-These registers are called main register set (\textit{a0,b0,c0,d0,e0,f0,h0,l0})
-and alternate register set (\textit{a\_0,b\_0,c\_0,d\_0,e\_0,f\_0,h\_0,l\_0}).
-These registers can be accessed in pairs, forming 16-bits, resulting in another
-set of identifiers of 16-bits registers, named $\textit{id\_reg16}$. These sets
-are represented below.
+represented by identifiers used as parameters in the instructions. It
+includes two sets of six general purpose registers which may be used
+individually as 8-bit registers or as 16-bit register pairs.  The
+working registers are represented by variable $\textit{rgs8}$. The
+domain of $\textit{rgs8}$ ($\textit{id\_reg8}$) is a set formed by
+identifiers of registers of 8 bits.  These registers are called main
+register set (\textit{a0,b0,c0,d0,e0,f0,h0,l0}) and alternate register
+set (\textit{a\_0,b\_0,c\_0,d\_0,e\_0,f\_0,h\_0,l\_0}).  These
+registers can be accessed in pairs, forming 16-bits, resulting in
+another set of identifiers of 16-bits registers, named
+$\textit{id\_reg16}$. These sets are represented below.
 
@@
 
-The main working register of Z80 is the accumulator ($\textit{rgs8(a0)}$) used for arithmetic/logic, input/output and loading/storing operations.
-Another important element is the ``f'' register ($\textit{rgs8(f0)}$),  that is
-used as a \textbf{flag register}. This register uses only six bits to represent the execution result status of each instruction.
-According to the official manual the bits 3 and 5 are not used and the others bits have the follow meaning: $\textit{bv\_get(rgs8(f0),0)}$ - the carry bit; $\textit{bv\_get(rgs8(f0),1)}$ - the add/subtract bit; $\textit{bv\_get(rgs8(f0),2)}$ - the parity or overflow bit; $\textit{bv\_get(rgs8(f0),4)}$ - the half carry bit; $\textit{bv\_get(rgs8(f0),6)}$ - the zero bit; and $\textit{bv\_get(rgs8(f0),7)}$ - the sign bit.
+The main working register of Z80 is the accumulator
+($\textit{rgs8(a0)}$) used for arithmetic/logic, input/output and
+loading/storing operations.  Another important element is the ``f''
+register ($\textit{rgs8(f0)}$), that is used as a \textbf{flag
+  register}. This register uses only six bits to represent the
+execution result status of each instruction.  According to the
+official manual the bits 3 and 5 are not used and the others bits have
+the follow meaning: $\textit{bv\_get(rgs8(f0),0)}$ - the carry bit;
+$\textit{bv\_get(rgs8(f0),1)}$ - the add/subtract bit;
+$\textit{bv\_get(rgs8(f0),2)}$ - the parity or overflow bit;
+$\textit{bv\_get(rgs8(f0),4)}$ - the half carry bit;
+$\textit{bv\_get(rgs8(f0),6)}$ - the zero bit; and
+$\textit{bv\_get(rgs8(f0),7)}$ - the sign bit.
 %\begin{description}
@@
 %\end{description}
-These bits can also be used to specify security properties for microcontroller programs.
-
-\textbf{Assuring the absence of overflow:}
- \emph{To assure that an overflow does not happen, the developer can add in the invariant this 
+These bits can also be used to specify security properties for
+microcontroller programs.
+
+\textbf{Checking overflows:} \emph{To assure that an overflow does not
+  happen, the developer can add in the invariant this
   expression:\\(\textit{bv\_get(rgs8(f0),0)} $\neq$ 1 $\land$
- \textit{bv\_get(rgs8(f0),2) $\neq$ 1}).
-% When a overflow happen,it can be dangerous. So the developer may restrict its use. However, this restriction can also
-% become more difficult to verify the model.
-}
+  \textit{bv\_get(rgs8(f0),2) $\neq$ 1}).  }
 
@@
 
- There are some specific functions from Z80 to manipulate the data. In addressing mode, the function
-$\textit{bv\_ireg\_plus\_d}$ receives the value of a register ($\textit{ix}$ or $\textit{iy}$) 
-used as a base to point to a region in memory from which data is to be stored or retrieved.
+There are some specific functions from Z80 to manipulate the data. In
+addressing mode, the function $\textit{bv\_ireg\_plus\_d}$ receives
+the value of a register ($\textit{ix}$ or $\textit{iy}$) used as a
+base to point to a region in memory from which data is to be stored or
+retrieved.
 %and the displacement to return the sum, the result is the dislocated address memory, 
@@
 
-Another derived function is  $\textit{bv\_(ireg\_plus\_d)}$, which returns the value in the memory  address
-returned by the $\textit{bv\_ireg\_plus\_d}$ function and its definition is similar.
-
-There is a specific function to refresh the flag register, it is named $\textit{update\_reg}$ $\textit{\_flag}$.
-It is typed and defined as follow:\it update\_flag\_reg \rm $\in$ \rm (\it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it
+Another derived function is $\textit{bv\_(ireg\_plus\_d)}$, which
+returns the value in the memory address returned by the
+$\textit{bv\_ireg\_plus\_d}$ function and its definition is similar.
+
+There is a specific function to refresh the flag register, it is named
+$\textit{update\_reg}$ $\textit{\_flag}$. It is defined as follow:\it
+update\_flag\_reg \rm $\in$ \rm (\it BIT $\times$ \it BIT $\times$ \it
+BIT $\times$ \it
 BIT $\times$  \it BIT  $\times$  \it BIT\rm) $\fun$  \rm (\rm \{\it f0\rm \}  $\times$  \it BYTE\rm ).  This function is defined as:\\
@@
 
-The Z80 uses a unique memory for storing program instructions, data stack and
-general-purpose data. The memory has 16-bit addresses and each address holds a byte.
-Thus, the memory is very simple: \it mem  $\in$  \it BV16  $\fun$  \it BYTE\rm.
-
-In general, the instructions can access all memory addresses, but it is dangerous. The user may mistakenly
-access and change data in memory. For added  security, it is important that the program instructions has
-limited access by region. Thus the designer can specify address regions in a refinement model to restrict the access from
-instructions. The address regions can be specified using constants
-($\textit{PROGRAM\_R\_ADR,DATA\_R\_ADR,STACK\_R\_ADR}$), these define respectively a range of restricted
-address for: programs instructions, general purpose data and data stack.
+The Z80 uses a unique memory for storing program instructions, data
+stack and general-purpose data. The memory has 16-bit addresses and
+each address holds a byte.  Thus, the memory is very simple: \it mem
+$\in$ \it BV16 $\fun$ \it BYTE\rm.
+
+Instructions can access all the memory addresses and open the door to
+the programmer mistakenly access and change data in memory. For added
+security, it is desirable that the program instructions has limited
+access by region. Thus the designer can specify address regions in a
+refinement model to restrict the access from instructions. The address
+regions can be specified using constants
+($\textit{PROGRAM\_R\_ADR,DATA\_R\_ADR,STACK\_R\_ADR}$), these define
+respectively a range of restricted address for: programs instructions,
+general purpose data and data stack.
 %
@@
 
-\textbf{Preserving the memory consistency:} \emph{In general, the access to
-some address regions is dangerous. Then, each instruction has a specific pre-condition that verifies
-if the new address memory, that will be updated, is a member of its region. For example, the $\textit{PUSH}$
-program instruction permits write only in the region of stack ($\textit{STACK\_R\_ADR}$\rm).}
+\textbf{Controlling memory access:} \emph{Each instruction has a
+  specific pre-condition that verifies if the memory address that will
+  be updated belongs. For example, the $\textit{PUSH}$ instruction
+  permits write only in the stack area: $\textit{STACK\_R\_ADR}$.}
 
@@
 
-There are many functions in the module \textit{ALU}. In general, the definition of these 
-functions use basic definitions or previously defined functions. For example, the function
-$\textit{half8UCHAR}$ is used to get the half part of $\textit{UCHAR}$ value.
-It is important to know the half carry and it is used in the function $\textit{add8UCHAR}$. \\
+There are many functions in the module \textit{ALU}. In general, the
+definition of these functions use basic definitions or previously
+defined functions. For example, the function $\textit{half8UCHAR}$ is
+used to get the half part of $\textit{UCHAR}$ value. The value of the
+half carry is important as it is used in the function $\textit{add8UCHAR}$. \\
 
@@
 
-The function $\textit{add8UCHAR}$ receives a bit carry and two $\textit{UCHAR}$ values and returns respectively the 
-sum, the sign bit, the carry bit, the half carry bit and the zero bit. It is typed as follows: \it add8UCHAR \rm :
+The function $\textit{add8UCHAR}$ receives a bit carry and two
+$\textit{UCHAR}$ values and returns respectively the sum, the sign
+bit, the carry bit, the half carry bit and the zero bit. It is typed
+as follows: \it add8UCHAR \rm :
 \rm (\it BIT $\times$ \it UCHAR $\times$ \it UCHAR\rm ) $\fun$ \rm (\it UCHAR $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm ) and its definition is:\\
 
-\hspace*{0.0in}\it add8UCHAR\rm = $\lambda$ \rm(\it carry\rm,\it w1\rm, \it w2\rm)\rm.\rm
-
-\hspace*{0.0in}(\it carry $\in$  \it BIT  $\land$  \it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it UCHAR
-$\mid$
-
-\hspace*{0.40in}\rm(\rm(\rm(\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm ),
-
-\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it uchar\_schar\rm (\it w1\rm ) \rm + \it uchar\_schar \rm (\it
-w2\rm ) $<$ \rm 0\rm ),
-
-\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it w1 \rm + \it w2 $>$ \it UCHAR\_MAX\rm )\rm ,
-
-\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm) \rm + \it half8UCHAR\rm (\it w2\rm)  $\geq$  \it $2^{4}$\rm )\rm,
-
-\hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm
+\hspace*{0.0in}\it add8UCHAR\rm = $\lambda$ \rm(\it carry\rm,\it
+w1\rm, \it w2\rm)\rm.\rm
+
+\hspace*{0.0in}(\it carry $\in$ \it BIT $\land$ \it w1 $\in$ \it UCHAR
+$\land$ \it w2 $\in$ \it UCHAR $\mid$
+
+\hspace*{0.40in}\rm(\rm(\rm(\it carry \rm + \it w1 \rm + \it w2 \rm )
+$\mod$ \it $2^{8}$ \rm ),
+
+\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it uchar\_schar\rm
+(\it w1\rm ) \rm + \it uchar\_schar \rm (\it w2\rm ) $<$ \rm 0\rm ),
+
+\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it w1 \rm + \it w2
+$>$ \it UCHAR\_MAX\rm )\rm ,
+
+\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm
+(\it w1\rm) \rm + \it half8UCHAR\rm (\it w2\rm) $\geq$ \it $2^{4}$\rm
+)\rm,
+
+\hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1
+\rm + \it w2 \rm ) $\mod$ \it $2^{8}$ \rm
 )\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )\\
 
-A related function to subtract operation is $\textit{subtract8UCHAR}$. There are the same functions for the
-$\textit{SCHAR}$ type, they are respectively $\textit{add8SCHAR}$ and $\textit{subtract8SCHAR}$, all these
-functions are of 8 bits ($\textit{BYTE}$) and defined similarly. In the same way, the arithmetic functions for 16
-bits ($\textit{BV16}$) are defined.
-The module ALU has several others functions, for example:%but only some simplest functions also are explained below:
+A related function is subtraction: $\textit{subtract8UCHAR}$. There
+are also similar functions for the $\textit{SCHAR}$ type, they are
+respectively $\textit{add8SCHAR}$ and $\textit{subtract8SCHAR}$, all
+these functions are of 8 bits ($\textit{BYTE}$). In the same way, the
+arithmetic functions for 16 bits ($\textit{BV16}$) are defined.  The
+module ALU has several others functions, for
+example:%but only some simplest functions also are explained below:
 %\begin{itemize}
@@
 \it instruction\_next $\in$ USHORT $\fun$ USHORT \rm - It receives the
-  actual value from program counter register (\textit{pc}) and returns its increment;
+current value from program counter register (\textit{pc}) and returns
+its increment;
 % \item 
-and \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns the most significant bit,
- in other words, the signal bit.
+and \it is\_negative $\in$ \it BYTE $\fun$ \it BIT \rm - It returns
+the most significant bit, in other words, the signal bit.
   % \item \it is\_zero  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
@@
 %\end{itemize}
-As the logic functions that are defined in the \textit{BYTE} and \textit{BV16}
-module are included in the \textit{ALU} module, they can be seen and used
-directly in the \textit{ALU} and \textit{Z80} modules.
+As the logic functions that are defined in the \textit{BYTE} and
+\textit{BV16} module are included in the \textit{ALU} module, they can
+be seen and used directly in the \textit{ALU} and \textit{Z80}
+modules.
 
@@
 %TODO: Fiz uma breve alteração
-Each instruction is represented by a B operation in Z80 module. The main
-module (\textit{Z80}) has two categories of operations: a category represents the 
-instructions of microcontrollers and a category represents the external actions, e.g.
-interruptions actions. All 158 instructions and interruptions were modeled, but only
+Each instruction is represented by a B operation in Z80 module. The
+main module (\textit{Z80}) has two categories of operations: a
+category represents the instructions of microcontrollers and a
+category represents the external actions, e.g.  interruptions
+actions. All 158 instructions and interruptions were modeled, but only
 two important operations are presented here.
 
-A simple example of instruction is $\textit{LD\_(nn)\_A}$\footnote{The B tools does not allow o use parentheses in identifiers, and the characters \{``('',``)''\} are
-replaced respectively by \{``9'',``0''\} in the actual specification.} shown in this section. The
-pre-defined functions are necessary many times to model the instructions, these
-functions facilitate the construction of instruction set model. By default, all
-parameters in operations are either predefined elements in the model or
-integers values in decimal representation. This instruction use the
-$\textit{updateAddressMem}$ operation from \textit{Memory} module, it receives
-an address memory and its new memory value. $\textit{LD\_(nn)\_A}$ also increments the program
-counter ($\textit{pc}$) and updates the refresh register ($\textit{r\_}$). The other instructions have a similar structure.
+A simple example of instruction is $\textit{LD\_(nn)\_A}$\footnote{The
+  B tools does not allow o use parentheses in identifiers, and the
+  characters \{``('',``)''\} are replaced respectively by
+  \{``9'',``0''\} in the actual specification.} shown in this
+section. The pre-defined functions are necessary many times to model
+the instructions, these functions facilitate the construction of
+instruction set model. By default, all parameters in operations are
+either predefined elements in the model or integers values in decimal
+representation. This instruction use the $\textit{updateAddressMem}$
+operation from \textit{Memory} module, it receives an address memory
+and its new memory value. $\textit{LD\_(nn)\_A}$ also increments the
+program counter ($\textit{pc}$) and updates the refresh register
+($\textit{r\_}$). The other instructions have a similar structure.
 
@@
 %
-The external actions change the state of the microcontroller, for example,
-refreshing the I/O ports and requesting interruptions. The external actions are also
-modeled by operations and they are named with the prefix ``$ext\_$'' and
-followed by the name of action. There are four external actions:
-$ext\_update\_io\_ports$, $ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The
-instruction $ext\_update\_io\_ports$ just changes the state of I/O port, see.
+The external actions change the state of the microcontroller, for
+example, refreshing the I/O ports and requesting interruptions. The
+external actions are also modeled by operations and they are named
+with the prefix ``$ext\_$'' and followed by the name of action. There
+are four external actions: $ext\_update\_io\_ports$, $ext\_NMI$ and
+$ext\_INT$, $ext\_Reset$. The instruction $ext\_update\_io\_ports$
+just changes the state of I/O port, see.
 
@@
 
-The other external actions are related to interruptions. Interruptions allow
-devices to suspend a routine from CPU and start another service routine.
-This service routine can exchange data or signals between CPU and external
-devices. When a routine is finished, then the CPU comes back to the last routine
-that was interrupted. Modeling details about interruptions can be consulted in
+The other external actions are related to interrupts. Interrupts allow
+devices to suspend a routine from CPU and start a service routine.
+This routine can exchange data or signals between CPU and external
+devices. When it finishes, then the CPU comes back to the interrupted
+routine. Modeling details about interruptions can be consulted in
 public project's repository.
@@
 general-purpose animator of formal specification, and addresses
-computationally more difficult problems. Anyway, it is capable of
-animating, and also possibly model checking, a formal model of the
-instruction set of the Z80, which was our initial motivation in
+computationally more difficult problems than a simulator. It is anyway
+capable of animating, and also possibly model checking, a formal model
+of the instruction set of the Z80, which was our initial motivation in
 constructing that model. One may simulate step-by-step sequences of
@@
 here some details on this effort and some solutions we have devised to
-reduce the effort.  In B, the generated proof obligations allow to
-verify the data satisfy invariant properties, check important system
-properties and that the expressions are well-defined (WD)\footnote{An
-  expression is called ``well-defined'' (or unambiguous) if its
-  definition assigns it a unique interpretation or value.}.
+reduce the effort.  Recall that, in B, the generated proof obligations
+allow to verify the data satisfy invariant properties, check important
+system properties and that the expressions are well-defined
+(WD)\footnote{An expression is called ``well-defined'' (or
+  unambiguous) if its definition assigns it a unique interpretation or
+  value.}.
 
@@
 correct. Each such proof obligation requires interaction with a proof
-assistant. We see that for a project of that size, a large effort
-needs to be dedicated to find those proofs. One interesting feature of
-the proof assistant is, once a proof has been found, it can
-automatically apply it to other proof obligations, as it is often the
-case that the same proof script discharges several proof obligations.
+assistant. For a project of that size, a large effort needs to be
+dedicated to find those proofs. Note that, once a new proof has been
+found, the interactive prover can automatically apply it to other
+proof obligations. It is often the case that the same proof script
+discharges several proof obligations.
 
@@
 % processamento elas reduziram ?
-We had a previous experience with a similar effot, where we were able
+We had a previous experience with a similar effort, where we were able
 to discharge all the proof obligations: that of the model of the Z80
@@
 that may be animated with ProB, and without modeling external actions.
-In that first experience, the verification took us about two months. In that second
-project, we have been able to verify 3062 (e.g. roughly 80 \%) of 3631
-proof obligations in one week.
+In that first experience, the verification took us about two
+months. In that second project, we have been able to verify 3062
+(e.g. roughly 80 \%) of 3631 proof obligations in one week.
 
@@
 split the instruction set of the Z80 in eight sub-modules and checked
-them concurrently (each instance of AtelierB may only verify one module
-at a time).
+them concurrently (each instance of AtelierB may only verify one
+operation at a time).
 
@@
 use of data types, the developed security properties and the
-well-defined of all the expressions.  Furthermore, the designer has a
-big flexibility to create new and specific security properties, this
-is very
-useful %TODO: COLOCAR APÓS TODOS OS BENEFICIOS PROPORCIONADOS COM A SIMULAÇÃO e COMPATIBILIDADE COM O PROB. (OBTER DO PARAGRAFO ANTERIOR DA SEÇAO VERIFICATION PROCESS).
-to adjust the verification in accordance with the
-requirements. Moreover, this example of model could complement the
-existing documentation for users and assembly programmers.
-
-But our goal is to use the model of the Z80 instruction set to extend
-the scope of the B method up to the assembly
+well-defined of all the expressions.  Furthermore, the designer may
+create new and specific security properties to adjust the verification
+in accordance with the requirements of a specific
+application. Moreover, this model complements the existing
+documentation for users and assembly programmers.
+
+But our fundamental goal is to use the model of the Z80 instruction
+set to extend the scope of the B method up to the assembly
 level~\cite{Dantas_SBMF08}. Indeed, we realized a case study where an
@@
 There exists several approaches to model hardware and instruction set
-using B a.  The paper \cite{springerlink:Yuan2011} reports a method to
-specify, design and construct sound and complete instruction set model
-by stepwise refinement and formal proof using Event-B. It also
-discusses desirable properties of an instruction set model. However,
-our work \cite{Valerio_SBMF09} and \cite{Subotic2010} use the B
-method, that seems quite appropriated to software development, because
-it has an implementable language defined, called B0, a model-driven
-approach \cite{Dantas_SBMF08} to develop software from the functional
+using B, not to mention other formal modelling approaches.  The paper
+\cite{springerlink:Yuan2011} reports a method to specify, design and
+construct sound and complete instruction set model by stepwise
+refinement and formal proof using Event-B. It also discusses desirable
+properties of an instruction set model. However, our work
+\cite{Valerio_SBMF09} and \cite{Subotic2010} use the B method, that
+seems quite appropriated to software development, because it has an
+implementable language defined, called B0, a model-driven approach
+\cite{Dantas_SBMF08} to develop software from the functional
 specification level down to assembly, and tools to convert the models
@@
 
-% [Trabalhos futuros ] LLVM:Existe também a pretensão de especificar as
 In the future, we plan to improve the support for the development of
@@
 platform is also envisioned. Finally, another important activity is to
-develop a formal model of a platform used in LLVM
+develop a formal model of the internal representation used in LLVM
 Compiler~\cite{DBLP:conf/cgo/LattnerA04}, which would enable us to

### 2012-07-05 trunk/papers/ArticleZ80/source_full/paper.tex
Ajustando o repositorio para artigo short

--- 2012-05-21 trunk/papers/ArticleZ80/source/paper.tex
@@
   specification of the functional requirements of a system to an assembly-level
-  implementation thereof.\newline %\keywords{Embedded Software, Simulation, B method and Verification}
+  implementation thereof.
 % Using this model-driven approach and the instruction set model, a small case
 % study to analyse a petroleum production test system in oil field was developed.
- { \textbf{Key words:} B method, Simulation, Verification and Z80}
+% { \textbf{Key words:} B method, Simulation, Verification and Z80}
 
@@
   well as some basic lemmas, e.g. values of different powers of
-  two, allproved interactively with the Atelier-B prover.
+  two, all proved interactively with the Atelier-B prover.
   This module is essential to establish the relationship between
@@
 $$
-\begin{array}{l}
+\hspace*{-1.2in}\begin{array}{l}
 \textit{bit\_not}  \in  \textit{BIT}  \fun  \textit{BIT}  \land 
@@
 $$
-\begin{array}{l}
+\hspace*{-1.4in}\begin{array}{l}
 \textit{bit\_and} \in \textit{BIT} \times \textit{BIT} \fun \textit{BIT} \land \\
@@
 and $\textit{bit\_xor}$ (exclusive disjunction), as well as lemmas on
-those operators. These are standard and their expression in B is
+those operators. These are standard and their representations in B are
 similar to $\textit{bit\_and}$.
@@
 $\in$  \it BIT\_VECTOR $\land$ \it w $\in$  \it BIT\_VECTOR  $\mid$  \it v\^ \it w\rm ) %$\cat$
-\newline
+\vspace*{0.05in}
 
@@
 Bytes, i.e. bit vectors of length 8, are a common entity in hardware
-design. We provide the following definitions:
+design. We provide the following definitions:\\
 
@@
 \hspace*{0.0in}$\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BYTE  $\implies$  \bf size\rm (\it b1\rm )\rm =\it BYTE\_WIDTH  $\land$  \it b1  $\in$  \bf seq1\rm (\it BIT\rm )\rm ) 
-
+\\
 %\hspace*{0.0in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \} $\land$
@@
 $$
-\begin{array}{l}
+\hspace*{-0.3in}\begin{array}{l}
 \textit{bv\_to\_nat} \in \textit{BIT\_VECTOR} \fun \nat  \land \\
@@
 An important lemma associated to this definition is: 
-$$\forall n . (n \in \nat_1 \implies \textit{bv\_to\_nat}(\textit{nat\_to\_bv}(n)) = n).$$
+$$\hspace*{-1.59in}\forall n . (n \in \nat_1 \implies \textit{bv\_to\_nat}(\textit{nat\_to\_bv}(n)) = n).$$
 
@@
 \begin{array}{l}
-\textit{byte\_uchar} \in \textit{BYTE} \fun \nat \land \\
+\textit{byte\_uchar} \in \textit{BYTE} \fun \nat \land 
 \textit{byte\_uchar} = \lambda (v) . ( v \in BYTE | bv\_to\_nat(v) )
@@
 $$
- \begin{array}{l}
+\hspace*{-0.7in} \begin{array}{l}
   \forall (val) . (val \in \textit{UCHAR} |
@@
 \item For the proof obligation (\it byte\_uchar\rm \ $\in$ \textit{BYTE} $\bij$
-\textit{UCHAR}), the proof rule is: $\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}  $\mid$  \rm 1\rm 2\rm 8 $\times$ \it v0\rm (\rm 8\rm )\rm +\rm 6\rm 4 $\times$ \it v0\rm (\rm 7\rm )\rm +\rm 3\rm 2 $\times$ \it v0\rm (\rm 6\rm )\rm +\rm 1\rm 6 $\times$ \it v0\rm (\rm 5\rm )\rm +\rm 8 $\times$ \it v0\rm (\rm 4\rm )\rm +\rm 4 $\times$ \it v0\rm (\rm 3\rm )\rm +\rm 2 $\times$ \it v0\rm (\rm 2\rm )\rm +\rm 1 $\times$ \it v0\rm (\rm 1\rm )\rm ) \rm $\in$ \rm (\rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}\rm )  $\bij$  \rm 0 $\upto$ \rm 2\rm 5\rm 5 \\
+\textit{UCHAR}), the proof rule is: $\lambda$ \it v0\rm .\rm (\it v0 $\in$  \rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}  $\mid$  \rm 1\rm 2\rm 8 $\times$ \it v0\rm (\rm 8\rm )\rm +\rm 6\rm 4 $\times$ \it v0\rm (\rm 7\rm )\rm +\rm 3\rm 2 $\times$ \it v0\rm (\rm 6\rm )\rm +\rm 1\rm 6 $\times$ \it v0\rm (\rm 5\rm )\rm +\rm 8 $\times$ \it v0\rm (\rm 4\rm )\rm +\rm 4 $\times$ \it v0\rm (\rm 3\rm )\rm +\rm 2 $\times$ \it v0\rm (\rm 2\rm )\rm +\rm 1 $\times$ \it v0\rm (\rm 1\rm )\rm ) \rm $\in$ \rm (\rm 1 $\upto$ \rm 8  $\fun$  \rm \{\rm 0\rm ,\rm 1\rm \}\rm )  $\bij$  \rm (0 $\upto$ \rm 2\rm 5\rm 5) \\
 
@@
 microcontroller state is formed by the program counter, ports,
-registers, etc. The stae transition occurs either as an instruction
+registers, etc. The state transition occurs either as an instruction
 execution or an external action. Groups of registers are represented
@@
 ($\textit{i\_o\_ports}$). These definitions are represented by
-\textit{INVARIANT}.
+\textit{INVARIANT}.\\
 
 \begin{sloppypar}
-\bf INVARIANT
+%\bf INVARIANT
 
@@
 
-\hspace*{0.10in}\it iff2  $\in$  \it BIT  $\land$ \it im $\in$ (\it BIT $\times$ \it BIT\rm )  $\land$  \it i\_o\_ports  $\in$  \it BYTE  $\fun$  \it BYTE
+\hspace*{0.10in}\it iff2  $\in$  \it BIT  $\land$ \it im $\in$ (\it BIT $\times$ \it BIT\rm )  $\land$  \it i\_o\_ports  $\in$  \it BYTE  $\fun$  \it BYTE\\
 \end{sloppypar}
@@
 another set of identifiers of 16-bits registers, named
-$\textit{id\_reg16}$. These sets are represented below.
+$\textit{id\_reg16}$. These sets are represented by  \textit{SETS}.\\
 
@@
 
-\bf SETS\\
-\hspace*{0.10in}\it id\_reg\_8\rm =\rm \{\it a0\rm, \it f0\rm, \it b0\rm, \it
-c0\rm, \it d0\rm, \it e0\rm, \it h0\rm, \it l0\rm, \it a\_0\rm,  \it f\_0\rm,
-\it b\_0\rm, \it c\_0\rm, \it d\_0\rm, \it e\_0\rm, \it h\_0\rm, \it l\_0 \rm
-\};\\ \hspace*{0.10in}\it id\_reg\_16\rm =\rm \{\it AF \rm, \it BC\rm, \it
-DE\rm, \it HL\rm, \it SP\rm\}
+%\bf SETS\\
+\hspace*{0.0in}\it id\_reg\_8\rm =\rm \{\it a0\rm,\it f0\rm,\it b0\rm,\it c0\rm,\it d0\rm,\it e0\rm,\it h0\rm,\it l0\rm,\it a\_0\rm,  \it f\_0\rm,\it b\_0\rm,\it c\_0\rm,\it d\_0\rm,\it e\_0\rm,\it h\_0\rm,\it l\_0 \rm\};\\
+\hspace*{0.2in}\it id\_reg\_16\rm =\rm \{\it AF\rm,\it BC\rm,\it
+DE\rm,\it HL\rm,\it SP\rm\}\\
 
@@
 
+%There is a related function $\textit{subtract8UCHAR}$ and others
 A related function is subtraction: $\textit{subtract8UCHAR}$. There
-are also similar functions for the $\textit{SCHAR}$ type, they are
-respectively $\textit{add8SCHAR}$ and $\textit{subtract8SCHAR}$, all
+are also similar functions for the $\textit{SCHAR}$ type, for example,
+$\textit{add8SCHAR}$ and  $\textit{subtract8SCHAR}$. All
 these functions are of 8 bits ($\textit{BYTE}$). In the same way, the
@@
 module ALU has several others functions, for
-example:%but only some simplest functions also are explained below:
+example: %but only some simplest functions also are explained below:
 %\begin{itemize}
@@
 ($\textit{r\_}$). The other instructions have a similar structure.
-
-\hspace*{0.00in}\bf LD\_9nn0\_A \rm ( \it nn \rm ) \rm =
+\\ \\
+\hspace*{0.20in}\bf LD\_9nn0\_A \rm ( \it nn \rm ) \rm =
 
@@
 \hspace*{0.00in}\bf END\rm 
+\\
 
@@
 $ext\_INT$, $ext\_Reset$. The instruction $ext\_update\_io\_ports$
-just changes the state of I/O port, see.
+just changes the state of I/O port, see.\\
 
@@
 
-\hspace*{0.00in}\bf END\rm
+\hspace*{0.00in}\bf END\rm \\
 
@@
 in parallel and also to send proof obligations to other, possibly
-remote, processes running AtelierB. With those options we had the
+remote, processes AtelierB. With those options we had the
 possibility to have eight concurrent instances of the prover. We thus
@@
 them concurrently (each instance of AtelierB may only verify one
-operation at a time).
+module at a time).
 
@@
 
-%\subsubsection{Acknowledgements:} The animation of the model would not
-%have been possible without the help of Michael Leuschel who kindly
-%provided feedback and developed several improvements to ProB to meet our
-%needs.
+\textbf{Acknowledgements:} The animation of the model would not
+have been possible without the help of Michael Leuschel who kindly
+provided feedback and developed several improvements to ProB to meet our
+needs.
 

### 2012-07-08 papers/ArticleZ80/source/paper_full_fixed.tex
Apliquei os ajustes sugeridos na versão Full.

--- 2012-07-05 trunk/papers/ArticleZ80/source_full/paper.tex
@@
 
-%TODO: Verificar possíveis diferenças do FORMATO LNCS
+%TODO: Verificar poss√≠veis diferen√ßas do FORMATO LNCS
 
@@
 partial, Z80 model presented in \cite{Valerio_SBMF09} was fully
-verified, but it used constructs that made it impossible to animate 
+verified, but it used constructs that made it incapable to animate 
 with existing technology. %%DD
@@
 our repository at: http://code.google.com/p/b2asm/.} no longer contains
-definitions derived from infinite sets. Using B constructs for modularity,
+definitions derived from infinite sets and added support to complex
+interruptions.Using B constructs for modularity,
 auxiliary libraries of basic modules were developed as part of the construction
@@
 
-%%DD As primeiras frases do parágrafo abaixo não faziam sentido!
+%%DD As primeiras frases do par√°grafo abaixo n√£o faziam sentido!
 The definitions, functions and properties from hardware library and
@@
 Table~\ref{tab:types} are verified automatically or with few user
-pass\footnote{The user pass allows to filter proof tactics that aid
+pass\footnote{The user passes allows to filter proof tactics that aid
   the prover to find the proof, this is a feature provided by Atelier
@@
 \label{sec:VerificationHardwareLibrary}
-
+%candidato para simplificar/cortar: baixa depend√™ncia 
+% Coment√°rio do revisor:
+%Section 2.6, par 1: it on the edge of what can be understood without previous knowledge of B, which was not described here.
+If the problem raised % specific to Atelier B?
+If so, why not use another tool?
 This section presents the verification of the aforementioned modules
@@
 %\end{enumerate}
-For the remaining proof obligations, we again had to create new proof
+For the remaining proof obligations, we had to create new proof
 rules. There are three kinds of proof rules, that we illustrate by giving
@@
 
-Eight proof obligations require that conversion functions are finite. %: \it
-% byte\_uchar $\in$ $\finset$\rm (\it byte\_uchar\rm ); \it uchar\_byte  $\in$
-% $\finset$\rm (\it uchar\_byte\rm ); \it ushort\_bv16 $ \in \finset$\rm (\it
-% ushort\_bv16\rm ); \it bv16\_ushort $ \in \finset$\rm (\it bv16\_ushort\rm ); \it
-% byte\_schar $ \in \finset$\rm (\it byte\_schar\rm ); \it schar\_byte $ \in
-% \finset$\rm (\it schar\_byte\rm ); \it sshort\_bv16 $ \in \finset$\rm (\it
-% sshort\_bv16\rm ); and \it bv16\_sshort $ \in \finset$\rm (\it bv16\_sshort\rm).
- These proof obligations were verified using the expression evaluator of ProB.
-Finally, we have included rules related to binary arithmetic as developed
-in~\cite{Leibnizens}. These rules are used in the interactive theorem prover with
-binary arithmetic to prove important properties~\cite{James2010,Dasgupta2006}.
+%Eight proof obligations require that conversion functions are finite. 
+% \it byte\_uchar $\in$ $\finset$\rm (\it byte\_uchar\rm );
+% \it uchar\_byte  $\in$ $\finset$\rm (\it uchar\_byte\rm ); 
+%\it ushort\_bv16 $ \in \finset$\rm (\it  ushort\_bv16\rm ); 
+%\it bv16\_ushort $ \in \finset$\rm (\it bv16\_ushort\rm );
+% \it byte\_schar $ \in \finset$\rm (\it byte\_schar\rm ); 
+% \it schar\_byte $ \in \finset$\rm (\it schar\_byte\rm );
+% \it sshort\_bv16 $ \in \finset$\rm (\it  sshort\_bv16\rm ); 
+% and \it bv16\_sshort $ \in \finset$\rm (\it bv16\_sshort\rm).
+
+%%VM: Realizei ajustes p/ responder perguntas do revisor
+For the eight remaining proof obligations, we  had to create new proof
+rules. The expression evaluator of ProB verified all rules related to functions with 8-bit data,
+the rules to functions with 16-bit data were not verified in ProB, because these rules have
+extensive space of verification. However, these rules were built according 
+to binary system arithmetic developed by~\cite{Leibnizens} and in similar way as functions with 8-bit data.
+% These rules were verified using the expression evaluator of ProB.
+%Finally, we have included rules related to binary arithmetic as developed
+%in~\cite{Leibnizens}. 
+These rules are used in the interactive theorem prover with
+binary arithmetic to prove basic properties~\cite{James2010,Dasgupta2006}.
 
@@
 
-The \textit{Z80} is a CISC microcontroller developed by
-\textit{Zilog}~\cite{Z80_manual}. The Z80 is composed by different
-elements and a simplified internal organization is shown in the Figure
+The Z80 is a CISC microcontroller developed by \textit{Zilog}~\cite{Z80_manual}. 
+The Z80 is composed by different elements and a simplified internal organization is shown in the Figure
 \ref{fig:DiagramBlock}.  This figure has some important elements of
@@
 $\in$ \it BV16 $\fun$ \it BYTE\rm.
-
-Instructions can access all the memory addresses and open the door to
-the programmer mistakenly access and change data in memory. For added
-security, it is desirable that the program instructions has limited
-access by region. Thus the designer can specify address regions in a
-refinement model to restrict the access from instructions. The address
-regions can be specified using constants
-($\textit{PROGRAM\_R\_ADR,DATA\_R\_ADR,STACK\_R\_ADR}$), these define
+ %
+By default, the Z80 no limits the access memory. Instructions can access 
+all the memory addresses and open the door to the programmer mistakenly access and change data in memory. For added
+security, Z80 model supports to limit access by region. Thus the designer can
+specify address regions in an instance of Z80 model to restrict the access from instructions. The address
+regions can be specified using constants ($\textit{PROGRAM\_R\_ADR,DATA\_R\_ADR,STACK\_R\_ADR}$), these define
 respectively a range of restricted address for: programs instructions,
@@
 \subsection{Modeling the actions and instructions}
-%TODO: Fiz uma breve alteração
+%TODO: Fiz uma breve altera√ß√£o
 Each instruction is represented by a B operation in Z80 module. The
@@
 
-\subsection{Verification and validation}% ESTOU TRABALHANDO AQUI
-% OK -> como as regras foram utilizadas ?  Respondido na seção das regras
-% TODO:ENGLISH REVIEW parágrafo inicial: VISÃO GERAL
+\subsection{Verification and validation}
+% OK -> como as regras foram utilizadas ?
 
@@
 
-%->Explicaçao de mais detalhes da tabela.
+%->Explicação de mais detalhes da tabela.
 
@@
 generic user passes are enough to prove many proof obligations. Also,
-as the models of many instructionsa are similar, proofs found in
+as the models of many instructions are similar, proofs found in
 interactive mode may similarly be replayed and discharge other proof
@@
 
-There exists several approaches to model hardware and instruction set
+There exists several approaches to model hardware and instruction sets
 using B, not to mention other formal modelling approaches.  The paper
 \cite{springerlink:Yuan2011} reports a method to specify, design and
-construct sound and complete instruction set model by stepwise
+construct sound and complete instruction set models by stepwise
 refinement and formal proof using Event-B. It also discusses desirable
@@
 avoid such problems in modeling and have developed a case study: a
-verified embedded software.  This case study was interesting to as a
+verified embedded software.  This case study was interesting too as a
 feasibility analysis of the approach to derive formally assembly code
@@
 
-%TODO: COLOCAR AQUI OS PRINCIPAIS BENEFÍCIOS PROPORCIONADOS PELO TRABALHO
-
-% fundamenta melhor o trabalho já desenvolvido. Citar o artigo SBMF e a dissertação. 
-% Os meus trabalhos já desenvolvidos/ resultados atuais.
+%TODO: COLOCAR AQUI OS PRINCIPAIS BENEF√çCIOS PROPORCIONADOS PELO TRABALHO
+
+% fundamenta melhor o trabalho j√° desenvolvido. Citar o artigo SBMF e a disserta√ß√£o. 
+% Os meus trabalhos j√° desenvolvidos/ resultados atuais.
 The following works are directly related to the objective this paper.
@@
 of external actions (interruptions).
-%Link com o parágrafo anterior 
-This set of papers present an alternative step to software verification
+%Link com o par√°grafo anterior 
+This set of papers presents an alternative step to software verification
 up to assembly language, showing the current difficulties and
@@
 
-% Conclusão: toque final
+% Conclus√£o: toque final
 %Several iterations were needed to provide the good library definitions as well as
